### 2014-10-01 exploring_sem/ggm_pacakge.Rmd
file changes

---
title: "Intro to the `ggm` R package for graphical Markov models"
output: html_document
---

From the package description:
>This package provides functions for defining, manipulating and fitting graphical Markov models.

For application in structural equation modeling and path analysis I am initially interested in the functionality for defining graphs. Particularly, I have a complex time-series path analysis diagramming a 1-year lag-effect. 
![My almost graph]("CH2_DAG.svg")
I want to define the graph and independence claims it outlines.
```{r}
library(ggm) #Load library

```

### 2014-10-01 exploring_sem/lavaan_tutorial.Rmd
file changes

---
title: "Intro to `lavaan` pacakage for SEM - Tutorial"
output: html_document
---

Load `lavaan` package and create an SEM using appropriate syntax
```{r load lavaan, echo=TRUE}
library(lavaan)
myModel <- ' # regressions
             y1 + y2 ~ f1 + f2 + x1 + x2
                  f1 ~ f2 + f3
                  f2 ~ f3 + x1 + x2

             # latent variable definitions 
               f1 =~ y1 + y2 + y3 
               f2 =~ y4 + y5 + y6 
               f3 =~ y7 + y8 + y9 + y10

             # variances and covariances 
               y1 ~~ y1 
               y1 ~~ y2 
               f1 ~~ f2

             # intercepts 
               y1 ~ 1 
               f1 ~ 1
           '
```

Example 1: Fitting a confirmatory factor analysis (CFA) model

```{r CFA Model, echo=TRUE}
HS.model <- ' visual  =~ x1 + x2 + x3 
              textual =~ x4 + x5 + x6
              speed   =~ x7 + x8 + x9 '
fit <- cfa(HS.model, data = HolzingerSwineford1939)
summary(fit, fit.measures = TRUE) # Get summary of fitted model
```

Example 2: Fitting a structural equation model (SEM)

```{r SEM Model, echo=TRUE}
model <- '
  # measurement model (latent variable definitions, CFA)
    ind60 =~ x1 + x2 + x3
    dem60 =~ y1 + y2 + y3 + y4
    dem65 =~ y5 + y6 + y7 + y8

  # regressions
    dem60 ~ ind60
    dem65 ~ ind60 + dem60

  # residual correlations (covariances)
    y1 ~~ y5
    y2 ~~ y4 + y6
    y3 ~~ y7
    y4 ~~ y8
    y6 ~~ y8
'
fit <- sem(model, data = PoliticalDemocracy)
summary(fit, standardized = TRUE)
```

### 2014-10-01 exploring_sem/ggm_pacakge.Rmd
gym update

@@
 Particularly, I have a complex time-series path analysis diagramming a 1-year lag-effect.
-![My almost graph]("CH2_DAG.svg")
+![My not quite complete graph]("CH2_DAG.svg")
 I want to define the graph and independence claims it outlines.
-```{r}
+```{r create DAG matrix}
 library(ggm) #Load library
+# From the function description (some slight paraphrasing and defining): "The `DAG` function defines the adjacency matrix of a directed acyclic graph.
+An adjacency matrix is a square Boolean [e.g.
+0/1] matrix that is equal to the number of nodes of the graph, and with a one in a given position (i,j) if there is an arrow from i to j and zero otherwise.
+The row names of the adjacency matrix are the nodes of the DAG."
+# This means writing composing each the regressions between variables directly connected by arrows in the graph
+my_graph <- DAG(habitat ~ topo, # Static variables
+                temp04 ~ topo + habitat, temp05 ~ topo + habitat, temp06 ~ topo + habitat, temp07 ~ topo + habitat, temp08 ~ topo + habitat, temp09 ~ topo + habitat, temp10 ~ topo + habitat, temp11 ~ topo + habitat, temp12 ~ topo + habitat, temp14 ~ topo + habitat, # Regressions of temperature
+                ppt04 ~ topo, ppt05 ~ topo, ppt06 ~ topo, ppt07 ~ topo, ppt08 ~ topo, ppt09 ~ topo, ppt10 ~ topo, ppt11 ~ topo, ppt12 ~ topo, ppt14 ~ topo, # Regressions of precipitation
+                slc04 ~ temp04 + ppt04 + habitat, slc05 ~ slc04 + temp05 + ppt05 + habitat, slc06 ~ slc05 + temp06 + ppt06 + habitat, slc07 ~ slc06 + temp07 + ppt07 + habitat, slc07 ~ slc06 + temp07 + ppt07 + habitat, slc08 ~ slc07 + temp08 + ppt08 + habitat, slc09 ~ slc08 + temp09 + ppt09 + habitat, slc10 ~ slc09 + temp10 + ppt10 + habitat, slc11 ~ slc10 + temp11 + ppt11 + habitat, slc12 ~ slc11 + temp12 + ppt12 + habitat, slc14 ~ slc12 + temp14 + ppt14 + habitat, # Regressions of **symptomatic leaf count** (slc)
+                prev04 ~ slc04 + temp04 + ppt04 + habitat, prev05 ~ temp05 + ppt05 + ppt04 + slc05 + slc04 + habitat, prev06 ~ temp06 + ppt06 + ppt05 + slc06 + slc05 + habitat, prev07 ~ temp07 + ppt07 + ppt06 + slc07 + slc06 + habitat, prev08 ~ temp08 + ppt08 + ppt07 + slc08 + slc07 + habitat, prev09 ~ temp09 + ppt09 + ppt09 + slc09 + slc09 + habitat, prev10 ~ temp10 + ppt10 + ppt09 + slc10 + slc09 + habitat, prev11 ~ temp11 + ppt11 + ppt10 + slc11 + slc11 + habitat, prev12 ~ temp12 + ppt12 + ppt11 + slc12 + slc11 + habitat, prev14 ~ temp14 + ppt14 + ppt12 + slc14 + slc12 + habitat, # Regressions of **prevalence** (prev)
+                order=TRUE)
+```
 
+Now I need to find the basis set for the conditional independencies implied by the directed acyclic graph (Shipley 2000, 2009).
+This can be done using the `basiSet` function.
+The basis set is often (every time I have seen it) mathematically notated as *Bu*.
+```{r get basis set}
+basis_set <- basiSet(my_graph)
 ```
+
+The output of `basiSet` is a list, for my convoluted and complex graph of relationships across time it is pretty large.
+```{r print basis set}
+basis_set
+```

### 2014-10-10 exploring_sem/ggm_pacakge.Rmd
code edits

@@
 library(ggm) #Load library
-# From the function description (some slight paraphrasing and defining): "The `DAG` function defines the adjacency matrix of a directed acyclic graph.
+# From the function description (some slight paraphrasing and defining of terms): "The `DAG` function defines the adjacency matrix of a directed acyclic graph.
 An adjacency matrix is a square Boolean [e.g.
@@
 
-The output of `basiSet` is a list, for my convoluted and complex graph of relationships across time it is pretty large.
+The output of `basiSet` is a list, for my convoluted and complex graph of relationships across time; it is pretty large at 736 independence claims.
 ```{r print basis set}

### 2014-10-10 explore_data.R

@@ -12,9 +12,9 @@ str(stems)
 
 
 
-#### Need to do some data cleaning and prep ####
+### Need to do some data cleaning and prep ####
 
-######Create date variable as posix and integer/factor year variable ####
+#### Create date variable as posix and integer/factor year variable ####
 library(lubridate)
 stems$Date <- as.character(stems$Date)
 class(stems$Date)
@@ -22,12 +22,12 @@ stems$date <- gsub("/", "-",stems$Date)
 stems$date <- dmy_hms(stems$date)
 stems$year <- year(stems$date)
 
-######Load plyr, dplyr, tidyr packages ####
+#### Load plyr, dplyr, tidyr packages ####
 library(plyr)
 library(dplyr)
 library(tidyr)
 
-######Rename columns using `rename` funcion from `plyr` ####
+#### Rename columns using `rename` funcion from `plyr` ####
 stems <- rename(stems, c("PlotID" = "plot", "Date" = "old_date"))
 stems <- rename(stems, c("ClusterID" = "cluster", "TagNumber" = "tag", 
                           "SpeciesID" = "species", "StemStatus" = "status",
@@ -74,7 +74,7 @@ rainy_days<-with(plot_202_rain, colMeans(cbind(freq2004,freq2005,freq2006,freq20
 
 df_rain_slc<-as.data.frame(rbind(avg_rain_year,avg_slc_year))
 
-######These may be useful functions for reference #### 
+###### These may be useful functions for reference #### 
 matplot(cbind(avg_rain_year,avg_slc_year),type="b")
 matplot(cbind(rainy_days,avg_slc_year),type="b")
 

### 2014-10-14 exploring_sem/ggm_pacakge.Rmd
Figure and text edits

--- 2014-10-01 exploring_sem/ggm_pacakge.Rmd
@@
 Particularly, I have a complex time-series path analysis diagramming a 1-year lag-effect.
-![My almost graph]("CH2_DAG.svg")
+![My not quite complete graph]("CH2_DAG.svg")
 I want to define the graph and independence claims it outlines.
+```{r create DAG matrix}
+library(ggm) #Load library
+# From the function description (some slight paraphrasing and defining of terms): "The `DAG` function defines the adjacency matrix of a directed acyclic graph.
+An adjacency matrix is a square Boolean [e.g.
+0/1] matrix that is equal to the number of nodes of the graph, with a one in a given position (i,j) if there is an arrow from i to j and zero otherwise.
+The row names of the adjacency matrix are the nodes of the DAG."
+# This means composing each the regressions between variables directly connected by arrows in the graph
+my_graph <- DAG(habitat ~ topo, # Static variables, topography & habitat
+                # Regressions of temperature
+                temp04 ~ topo + habitat, temp05 ~ topo + habitat, temp06 ~ topo + habitat, temp07 ~ topo + habitat, temp08 ~ topo + habitat, temp09 ~ topo + habitat, temp10 ~ topo + habitat, temp11 ~ topo + habitat, temp12 ~ topo + habitat, temp14 ~ topo + habitat,
+                # Regressions of precipitation
+                ppt04 ~ topo, ppt05 ~ topo, ppt06 ~ topo, ppt07 ~ topo, ppt08 ~ topo, ppt09 ~ topo, ppt10 ~ topo, ppt11 ~ topo, ppt12 ~ topo, ppt14 ~ topo, 
+                # Regressions of **symptomatic leaf count** (slc)
+                slc04 ~ temp04 + ppt04 + habitat, slc05 ~ slc04 + temp05 + ppt05 + habitat, slc06 ~ slc05 + temp06 + ppt06 + habitat, slc07 ~ slc06 + temp07 + ppt07 + habitat, slc07 ~ slc06 + temp07 + ppt07 + habitat, slc08 ~ slc07 + temp08 + ppt08 + habitat, slc09 ~ slc08 + temp09 + ppt09 + habitat, slc10 ~ slc09 + temp10 + ppt10 + habitat, slc11 ~ slc10 + temp11 + ppt11 + habitat, slc12 ~ slc11 + temp12 + ppt12 + habitat, slc14 ~ slc12 + temp14 + ppt14 + habitat,
+                # Regressions of **prevalence** (prev)
+                prev04 ~ slc04 + temp04 + ppt04 + habitat, prev05 ~ temp05 + ppt05 + ppt04 + slc05 + slc04 + habitat, prev06 ~ temp06 + ppt06 + ppt05 + slc06 + slc05 + habitat, prev07 ~ temp07 + ppt07 + ppt06 + slc07 + slc06 + habitat, prev08 ~ temp08 + ppt08 + ppt07 + slc08 + slc07 + habitat, prev09 ~ temp09 + ppt09 + ppt09 + slc09 + slc09 + habitat, prev10 ~ temp10 + ppt10 + ppt09 + slc10 + slc09 + habitat, prev11 ~ temp11 + ppt11 + ppt10 + slc11 + slc11 + habitat, prev12 ~ temp12 + ppt12 + ppt11 + slc12 + slc11 + habitat, prev14 ~ temp14 + ppt14 + ppt12 + slc14 + slc12 + habitat,
+                order=TRUE)
+```
+
+Find out if the graph I have defined is in fact acyclic
+```{r isAcyclic}
+isAcyclic(my_graph)
+```
+It is, so I didn't screw that part up...
+
+```{r plot graph}
+library(tcltk)
+plotGraph(my_graph, dashed = TRUE, tcltk = TRUE, layout = layout.grid, directed = TRUE, noframe = FALSE, nodesize = 50, vld = 1, vc = "gray", vfc = "black", cex = 2)
+```
+That was super ugly, but a neat feature. The next example is much the same.
+```{r draw graph}
+drawGraph(my_graph, lwd = 2, ecol = "brown", left = TRUE)
+```
+
+```{r Rgraphviz & dynamicGraph packages, echo=FALSE}
+library(dynamicGraph)
+```
+
+Now I need to find the basis set for the conditional independencies implied by the directed acyclic graph (Shipley 2000, 2009).
+This can be done using the `basiSet` function.
+The basis set is often (every time I have seen it) mathematically notated as *Bu*.
+```{r get basis set}
+
+basis_set <- basiSet(my_graph)
+```
+
+The output of `basiSet` is a list, for my convoluted and complex graph of relationships across time; it is pretty large at 736 independence claims.
+```{r print basis set}
+basis_set
+```
+
+Find the list of all subsets of a set
 ```{r}
-library(ggm) #Load library
-
+powerset(my_graph)
 ```

### 2014-12-02 analysis/exploring_sem/ggm_pacakge.Rmd
revamped directory and made some changes to repeated measures and exploratory script files. plan to change these to Rmd

--- 2014-10-14 exploring_sem/ggm_pacakge.Rmd
@@
 From the package description:
->This package provides functions for defining, manipulating and fitting graphical Markov models.
+"This package provides functions for defining, manipulating and fitting graphical Markov models."
 
@@
 ```{r}
-powerset(my_graph)
+ps <- powerset(my_graph)
 ```

### 2014-12-02 analysis/exploring_sem/lavaan_tutorial.Rmd
revamped directory and made some changes to repeated measures and exploratory script files. plan to change these to Rmd

---
title: "Intro to `lavaan` pacakage for SEM - Tutorial"
output: html_document
---

#### Load `lavaan` package and create an SEM using appropriate syntax
```{r load lavaan, echo=TRUE}
library(lavaan)
myModel <- ' # regressions
             y1 + y2 ~ f1 + f2 + x1 + x2
                  f1 ~ f2 + f3
                  f2 ~ f3 + x1 + x2

             # latent variable definitions 
               f1 =~ y1 + y2 + y3 
               f2 =~ y4 + y5 + y6 
               f3 =~ y7 + y8 + y9 + y10

             # variances and covariances 
               y1 ~~ y1 
               y1 ~~ y2 
               f1 ~~ f2

             # intercepts 
               y1 ~ 1 
               f1 ~ 1
           '
```

#### Example 1: Fitting a confirmatory factor analysis (CFA) model
```{r CFA Model, echo=TRUE}
HS.model <- ' visual  =~ x1 + x2 + x3 
              textual =~ x4 + x5 + x6
              speed   =~ x7 + x8 + x9 '
fit <- cfa(HS.model, data = HolzingerSwineford1939)
summary(fit, fit.measures = TRUE) # Get summary of fitted model
```

#### Example 2: Fitting a structural equation model (SEM)
```{r SEM Model, echo=TRUE}
model <- '
  # measurement model (latent variable definitions, CFA)
    ind60 =~ x1 + x2 + x3
    dem60 =~ y1 + y2 + y3 + y4
    dem65 =~ y5 + y6 + y7 + y8

  # regressions
    dem60 ~ ind60
    dem65 ~ ind60 + dem60

  # residual correlations (covariances)
    y1 ~~ y5
    y2 ~~ y4 + y6
    y3 ~~ y7
    y4 ~~ y8
    y6 ~~ y8
'
fit <- sem(model, data = PoliticalDemocracy)
summary(fit, standardized = TRUE)
```

### 2014-12-02 analysis/explore_data.R

@@ -5,8 +5,7 @@
 
 ######Set working directory, read in all stem data, examine data structure ####
       #This is only the biologically related field data
-setwd("~/GitHub/superspreaders/analysis")
-stems <- read.csv("data/Stem_Summary_Query.txt")
+stems <- read.csv("analysis/data/stem_summary_query_2004-2014.txt")
 head(stems)
 str(stems)
 
@@ -22,12 +21,9 @@ stems$date <- gsub("/", "-",stems$Date)
 stems$date <- dmy_hms(stems$date)
 stems$year <- year(stems$date)
 
-#### Load plyr, dplyr, tidyr packages ####
-library(plyr)
-library(dplyr)
-library(tidyr)
 
 #### Rename columns using `rename` funcion from `plyr` ####
+library(plyr)
 stems <- rename(stems, c("PlotID" = "plot", "Date" = "old_date"))
 stems <- rename(stems, c("ClusterID" = "cluster", "TagNumber" = "tag", 
                           "SpeciesID" = "species", "StemStatus" = "status",
@@ -42,7 +38,9 @@ stems <- read.csv("data/all_stems.csv")
 
 
 
-####### Older code that can and will likely be abandoned ######
+### ### ### ### ### ### ### ### ### ### ### ### ### ### #
+### Older code that can and will likely be abandoned ####
+### ### ### ### ### ### ### ### ### ### ### ### ### ### # 
 library(foreign)
 plot_dat1<-read.dbf("analysis/data/UMCA_slc_dbh_04-12.dbf")
 str(plot_dat1)

### 2014-12-02 analysis/repeated_measures.R

@@ -1,6 +1,6 @@
-# Repeated measures analysis ####
-setwd("~/GitHub/superspreaders/analysis")
-stems <- read.csv("data/all_stems.csv")
+## Repeated measures analysis ####
+
+stems <- read.csv("analysis/data/all_stems.csv")
 str(stems)
 summary(stems)
 
@@ -12,10 +12,19 @@ library(dplyr)
 library(tidyr)
 library(ggplot2)
 
+names(stems)
+stems <- stems %>%
+      select(-X, -date, -year)
+class(stems$old_date)
+stems$old_date <- as.Date(stems$old_date, format = "%m/%d/%Y")
+stems$year <- as.factor(year(stems$old_date))
+class(stems$year)
+
 ### Create bay laurel only data frame ####
 umca <- as.tbl(stems) %>%
       select(plot, cluster, tag, species, status, slc, dbh, year) %>%
       filter(species == "UMCA")
+summary(umca)
 
 # Develop plot level summaries data frame and join to stems data frame
 umca <- left_join(umca, plot_umca <- umca %>%
@@ -26,7 +35,7 @@ umca <- left_join(umca, plot_umca <- umca %>%
                 live_dens = length(tag)), 
       by = c("plot", "year"))
 summary(umca)
-
+umca <- droplevels(umca)
 
 ###Linear Mixed Effects Method ####
 #### A couple of lines from Sarah's code to work from
@@ -52,7 +61,7 @@ summary(umca)
 # lmer1  4 3861.9 3882.1 -1927.0   3853.9                         
 # lmer2  6 3865.2 3895.4 -1926.6   3853.2 0.7407      2     0.6905
 
-#### Plot Level ####
+#### Plot Level Repeated Measures Analysis Using Multi-level Mixed-Effects Regression ####
 # Starting with plot level data b/c simpler w/fewer nested groups
 library(lme4)
 m0 <- lmer(tot_slc ~ 1 + (1 | year), data = plot_umca)
@@ -83,6 +92,7 @@ sigma(m0)
 sigma(m1)
 sigma(m2)
 #### Plot model profiles using `lattice` package
+library(lattice)
 xyplot(profile(m1), conf = c(50, 90, 95, 99)/100)
 densityplot(profile(m1))
 splom(profile(m1), conf = c(50, 90, 95, 99)/100)
[…]

### 2014-12-05 d-sep_test.Rmd
added directory for d-sep test exercise, d_sep_ex and the renamed d-sep_test.Rmd file

## Steps for Performing a d-Sep Test
###### (Shipley 2000, 2003, 2004, 2009)

1. Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph
<img src="figures/CH2_DAG.svg" width="600px" height="600px">
2. List each of the *k* pairs of variables in the graph that do not have an arrow between them:

	Row |Variable Pairs 1 | Variable Pairs 2 | Variable Pairs 3
--- | --- | --- | ---
1|*Topography, Leaf Count* | *Topography, Prevalence* | *Host Habitat, Precipitation*
2|*Temp2004, Temp2005*|*Temp2004, Precip2005*|*Precip2004, Precip2005*
3|*Temp2004, Temp2006*|*Temp2004, Precip2006*|*Precip2004, Precip2006*
4|*Temp2004, Temp2007*|*Temp2004, Precip2007*|*Precip2004, Precip2007*
5|*Temp2004, Temp2008*|*Temp2004, Precip2008*|*Precip2004, Precip2008*
6|*Temp2004, Temp2009*|*Temp2004, Precip2009*|*Precip2004, Precip2009*
7|*Temp2004, Temp2010*|*Temp2004, Precip2010*|*Precip2004, Precip2010*
8|*Temp2004, Temp2011*|*Temp2004, Precip2011*|*Precip2004, Precip2011*

This is just an example of what part of this table would look like. I can use the `ggm` package to find the entire list and define the basis set as described below.

3. For each of the *k* pairs of variables (*Xi,Xj*), list the set of other variables, {*Z*} that are direct causes of either *Xi* or *Xj*. The pair of variables, (*Xi,Xj*), along with its conditioning set, {*Z*}, define an independence claim, (*Xi,Xj*) | {*Z*}, and the full set of the *k* independence claims defines the basis set, *Bu*.

4. For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*. In other words, perform a regression model using an appropriate method.

5. Combine the *k* probabilities into the *C*-statistic:

$$
C = -2\Sigma_{i=1}^k ln(P_i)
$$

This methodology can be extended into the mixed-model context using standard statistical programs with these capabilities, because you are testing multiple regressions hypothesized to be independent of each other. If too many independence claims are violated, then the model will be found to not fit the data.

It is also possible to apply this method to multilevel Bayesian (MCMC) models (McCarthy 2007) by testing the independence of the partial slopes and determining the 95% credible intervals.

### Example from Shipley 2009 on how to perform the `d-sep` test in a generalized multilevel context 
This is to demonstrate how to assess a multilevel, repeated measures path model by applying mixed models in the tests of d-separation. I am using the supplementary material, including the data provided.

Load the two libraries and read in the data Shipley uses
```{r load libraries}
library(nlme)
library(lme4)# masked lmList from nlme package
data <- read.table("d_sep_ex/shipley2009.dat")
summary(data)
head(data)
```
Shipley describes these data in the text as a study beginning in 1970 where 20 sites are chosen with different latitudes.

### 2014-12-05 d-sep_test.Rmd
worked all the way through the d-sep test of a causal graph using the example from Shipley 2009. still want to figure out how to calculate cumulative Chi-square probability using the C-statistic and degress of freedom

@@
 
-### Example from Shipley 2009 on how to perform the `d-sep` test in a generalized multilevel context 
+### Example from Shipley 2009 on how to perform the d-sep test in a generalized multilevel context 
 This is to demonstrate how to assess a multilevel, repeated measures path model by applying mixed models in the tests of d-separation.
@@
 
+##### Path Model
+`Latitude` --> `Degree-days` --> `Date of bud burst` --> `Growth` --> `Live`
+
 Load the two libraries and read in the data Shipley uses
-```{r load libraries}
+```{r load libraries, eval=FALSE}
 library(nlme)
@@
 data <- read.table("d_sep_ex/shipley2009.dat")
+str(data)
 summary(data)
@@
 ```
-Shipley describes these data in the text as a study beginning in 1970 where 20 sites are chosen with different latitudes.
+Shipley (2009) describes these data in the text as a study beginning in 1970 where 20 sites are chosen with different latitudes.
+Five individual trees are chosen within each site, and each tree is followed every second year until 2006 or death.
+In each site, each year, for each individual four measurements are recorded:
+
+   1. `DD` - the number of growing degree days until bud break
+   2. `Date` - the Julian date (day of year) of bud break
+   3. `Growth` - individual tree growth (change in diameter)
+   4. `Live` - binary indicator of survival (1) or death (0)
+
+I am now going to load the `ggm` library for identifying the directed acyclic graph and defining the basis set (a.k.a. the condition independence claims) without writing it out by hand.
+
+"The `DAG` function defines the adjacency matrix of a directed acyclic graph.
+An adjacency matrix is a square Boolean [e.g.
+0/1] matrix that is equal to the number of nodes of the graph, with a one in a given position (i,j) if there is an arrow from i to j and zero otherwise.
+The row names of the adjacency matrix are the nodes of the DAG."
+
+This means composing each the regressions between variables directly connected by arrows in the graph, which is pretty easy in this simple example because each variable has no more than one direct cause.
+
+```{r get basis set, eval=TRUE}
+library(ggm)
+dag <- DAG(Live ~ Growth, Growth ~ Date, Date ~ DD, DD ~ Latitude)
+isAcyclic(dag)# Logical test to check if the translated graph is acyclic, though a cyclic graph should return a warning, e.g.:
+DAG(x~y, y~z, z~x)
+Bu <- basiSet(dag)# Get the basis set as a list of vectors.
+The first two items of each vector are the names of the two non-adjacent nodes, followed by the names of the nodes in the conditioning set, which may be empty (though not in this instance).
+```
+
+Following the directions of the Shipley (2009) supplement I will fit separate mixed models for each independence claim in the basis set. 
+```{r print basis set, echo=TRUE}
+print(Bu)
+```
+Six claims equals six models.
+
+The first independence claim is that `Date` is independent of `Latitude` after controlling for `DD`.
+So, I need to obtain the *precise* null probability for this claim.
+Choosing the correct modeling procedure requires identifying the appropriate distribution of the dependent variable, in this case `Date`.
+Using some helper fucntions and the `pairs()` function in the base `graphics` package I can look at the correlations and distributions simultaneously.
+
+```{r helper functions for scatterplot matrix, echo=FALSE}
+# These were taken from the help pages of the  `pairs` function with minor alterations
+panel.hist <- function(x, ...)
+{
+    usr <- par("usr"); on.exit(par(usr))
+    par(usr = c(usr[1:2], 0, 1.5) )
+    h <- hist(x, plot = FALSE)
+    breaks <- h$breaks; nB <- length(breaks)
+    y <- h$counts; y <- y/max(y)
+    rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
+}
+panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
+{
+    usr <- par("usr"); on.exit(par(usr))
+    par(usr = c(0, 1, 0, 1))
+    r <- abs(cor(x, y, use="na.or.complete"))
+    txt <- format(c(r, 0.123456789), digits = digits)[1]
+    txt <- paste0(prefix, txt)
+    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+    text(0.5, 0.5, txt, cex = cex.cor * r)
+}
+```
+```{r correlations & distributions}
+pairs(subset(data, select = c(lat, Date, DD, Growth, Live)),
+      diag.panel = panel.hist, lower.panel = panel.cor)
+```
+The `Date` variable appears close to a normal distribution, so either the `lmer()` in the `lme4` package or the `lme()` function in the `nlme` package are appropriate.
+Shipley says that the `lme()` prints out degrees of freedom but `lmer()` does not.
+I am going to check to see if that is still true.
+```{r independence claim 1 test}
+fit1 <- lme(Date ~ DD + lat, data = data, 
+            random = ~1|site/tree, 
+            na.action = na.omit)
+fit1lme4 <- lmer(Date ~ DD + lat + (1|site) + (1|tree), data = data,
+                 na.action = na.omit)
+summary(fit1nlme)
+summary(fit1lme4)
+```
+It is correct that `lmer()` does not provide degrees of freedom, and thus no probability estimate (p-value) for the coefficients.
+Aside from that it appears that the values estimate from both algorithms match up very well.
+Based on these results the null probability of observing a t-value at least as extreme for latitude (`lat`) is 0.9373, so pretty darn likely.
+That means that this independence claim is supported by the data.
+
+Now to do the same for the other normally distributed dependent variables
+```{r independence claim test 2 & 3}
+# Independence of Growth and lat conditioned on Date
+fit2 <- lme(Growth ~ Date + lat, data = data, 
+            random = ~1|site/tree, 
+            na.action = na.omit)
+# Independence of Growth and DD conditioned on Date and lat
+fit3 <- lme(Growth ~ Date + lat + DD, data = data, 
+            random = ~1|site/tree, 
+            na.action = na.omit)
+```
+
+The other three independence claims have `Live` as the dependent variable. This is binary so I need to use the binomial error distribution with the logit link, requiring the `lmer()` function.
+```{r binomial independence claim Live & lat}
+# Independence of Live and lat conditioned on Growth
+fit4 <- glmer(Live ~ scale(Growth) + scale(lat) + (1|site) + (1|tree),
+             data = data, na.action = na.omit, family = binomial)
+summary(fit4)
+```
+This currently throws a warning about a large eigenvalue ratio with suggestion to rescale the variables.
+
+Some of the current estimates are fairly different from those shown in the Shipley supplement, particularly for standard errors, z-values, and p-values.
+The actual coefficient estimates are much more similar.
+Nonetheless, the interpretation that the independence claim is supported holds true, with a null probability for observing a z-value associated with `lat` at least that extreme equaling 0.304.
+
+Now to test the remaining independence claims, `Live` & `DD` conditioned on `Growth` & `lat`, and `Live` & `Date` conditioned on `DD` & `Growth`.
+```{r binomial independence claims 5 & 6}
+# Independence of Live and DD conditioned on Growth and lat
+fit5 <- glmer(Live ~ scale(Growth) + scale(lat) + scale(DD) + (1|site) + (1|tree),
+             data = data, na.action = na.omit, family = binomial)
+# Indpendence of Live and Date conditioned on DD and Growth
+fit6 <- glmer(Live ~ scale(Growth) + scale(DD) + scale(Date) + (1|site) + (1|tree),
+             data = data, na.action = na.omit, family = binomial)
+summary(fit6)
+```
+Again got warnings about large eigenvalues, as well as failure of the model to converge, with suggestion to rescale the variables.
+
+Now I want to get the fixed-effects summaries for each model, and in particular the null probability estimates (p-values) for the pair of variables in each independence claim.
+```{r nlme model & probability summaries}
+# These four lines were to help me figure out how to access the right part of the summary object
+#sum_fit1 <- summary(fit1)
+#str(sum_fit1)
+#sum_fit1$coefficients$fixed# Same as fixed.effects(fit1)
+#sum_fit1$tTable# Gets the full t-test table including coefficient estimates and probabilities
+summary(fit1)$tTable# lat, Date independence
+```
+```{r}
+summary(fit2)$tTable# Growth, lat independence
+```
+```{r}
+summary(fit3)$tTable# Growth, DD indpendence
+```
+
+```{r lme4 model & probability summaries}
+#sum_fit4 <- summary(fit4)
+#str(sum_fit4)
+#sum_fit4$coefficients
+summary(fit4)$coefficients# Independence of Live and lat 
+```
+```{r}
+summary(fit5)$coefficients# Independence of Live and DD
+```
+```{r}
+summary(fit6)$coefficients# Independence of Live and Date
+```
+
+Create a vector of the probabilities for the independence claims (6 values).
+```{r vector of probabilities}
+pvalues <- c(summary(fit6)$coefficients[4,4], summary(fit5)$coefficients[4,4],
+             summary(fit4)$coefficients[3,4], summary(fit3)$tTable[3,5],
+             summary(fit2)$tTable[3,5], summary(fit1)$tTable[3,5])
+```
+
+Combine the p-values into the *C*-statistic
+```{r calculate C statistic}
+ln_pvalues <- log(pvalues)
+C_stat <- -2*sum(ln_pvalues)
+```
+I would like to figure out how to calculate the chi-square cumulative probability based on the degrees of freedom and *C*-statistic, a la this [website](http://stattrek.com/online-calculator/chi-square.aspx)

### 2014-12-08 d-sep_test.Rmd
language edit

@@
 ```
-I would like to figure out how to calculate the chi-square cumulative probability based on the degrees of freedom and *C*-statistic, a la this [website](http://stattrek.com/online-calculator/chi-square.aspx)
+I would like to learn how to calculate the chi-square cumulative probability based on the degrees of freedom and *C*-statistic, a la this [website](http://stattrek.com/online-calculator/chi-square.aspx)

### 2014-12-10 d-sep_test.Rmd
figured out stupd error on subset call during "knitting" was do to an "eval=FALSE" setting in the code chunk for reading in the data, so it didn't exist

@@
 Load the two libraries and read in the data Shipley uses
-```{r load libraries, eval=FALSE}
+```{r load libraries, eval=TRUE, message=FALSE}
 library(nlme)
 library(lme4)# masked lmList from nlme package
-data <- read.table("d_sep_ex/shipley2009.dat")
-str(data)
-summary(data)
-head(data)
+data1 <- read.table("d_sep_ex/shipley2009.dat")
+str(data1)
+summary(data1)
+head(data1)
 ```
@@
 ```
-Six claims equals six models.
+Six claims equals six models. The first two variables listed in each item are claimed to be independent based on the following set of zero or more variables.
 
@@
 Using some helper fucntions and the `pairs()` function in the base `graphics` package I can look at the correlations and distributions simultaneously.
+```{r helper functions for scatterplot matrix, echo=FALSE}
+# These were copied from the help pages of the `pairs` function with minor alterations
 
-```{r helper functions for scatterplot matrix, echo=FALSE}
-# These were taken from the help pages of the  `pairs` function with minor alterations
+# Make histograms to put in the diagonal panel
 panel.hist <- function(x, ...)
@@
 }
+
+# Print absolute correlations in either upper or lower panel; size proportianal to magnitude of the correlation
 panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
@@
 ```{r correlations & distributions}
-pairs(subset(data, select = c(lat, Date, DD, Growth, Live)),
-      diag.panel = panel.hist, lower.panel = panel.cor)
+#pairs(subset(data1, select = c(lat, Date, DD, Growth, Live)), diag.panel = panel.hist, lower.panel = panel.cor)
+sub_data1 <- data1[ ,c(3,5:7,9)] #subset(data1, , select = c("lat","Date","DD","Growth","Live"))
+#library(dplyr)
+#sub_data1 <- data1 %>% select(lat, Date, DD, Growth, Live)
+
+pairs(sub_data1, diag.panel = panel.hist, lower.panel = panel.cor)
 ```
+
 The `Date` variable appears close to a normal distribution, so either the `lmer()` in the `lme4` package or the `lme()` function in the `nlme` package are appropriate.
@@
 ```{r independence claim 1 test}
-fit1 <- lme(Date ~ DD + lat, data = data, 
+fit1 <- lme(Date ~ DD + lat, data = data1, 
             random = ~1|site/tree, 
             na.action = na.omit)
-fit1lme4 <- lmer(Date ~ DD + lat + (1|site) + (1|tree), data = data,
+fit1lme4 <- lmer(Date ~ DD + lat + (1|site) + (1|tree), data = data1,
                  na.action = na.omit)
-summary(fit1nlme)
+summary(fit1)
 summary(fit1lme4)
@@
 # Independence of Growth and lat conditioned on Date
-fit2 <- lme(Growth ~ Date + lat, data = data, 
+fit2 <- lme(Growth ~ Date + lat, data = data1, 
             random = ~1|site/tree, 
@@
 # Independence of Growth and DD conditioned on Date and lat
-fit3 <- lme(Growth ~ Date + lat + DD, data = data, 
+fit3 <- lme(Growth ~ Date + lat + DD, data = data1, 
             random = ~1|site/tree, 
@@
 fit4 <- glmer(Live ~ scale(Growth) + scale(lat) + (1|site) + (1|tree),
-             data = data, na.action = na.omit, family = binomial)
+             data = data1, na.action = na.omit, family = binomial)
 summary(fit4)
@@
 fit5 <- glmer(Live ~ scale(Growth) + scale(lat) + scale(DD) + (1|site) + (1|tree),
-             data = data, na.action = na.omit, family = binomial)
+             data = data1, na.action = na.omit, family = binomial)
 # Indpendence of Live and Date conditioned on DD and Growth
 fit6 <- glmer(Live ~ scale(Growth) + scale(DD) + scale(Date) + (1|site) + (1|tree),
-             data = data, na.action = na.omit, family = binomial)
+             data = data1, na.action = na.omit, family = binomial)
 summary(fit6)
@@
              summary(fit2)$tTable[3,5], summary(fit1)$tTable[3,5])
+print(pvalues)
 ```
+These values do not all match up with the estimates printed in Table 2 of Shipley (2009). Assuming I did the modeling correctly (a substantial assumption), what gives?
 
-Combine the p-values into the *C*-statistic
+Combine the p-values into the *C*-statistic and see what we get anyway.
 ```{r calculate C statistic}
@@
 C_stat <- -2*sum(ln_pvalues)
+print(C_stat)
 ```
-I would like to learn how to calculate the chi-square cumulative probability based on the degrees of freedom and *C*-statistic, a la this [website](http://stattrek.com/online-calculator/chi-square.aspx)
+Checking this value in a $\Chi^2$ look up table against 2*k*(12) degrees of freedom provides an estimated probability of a larger value being somewhere between 0.25 and 0.50.
+
+I would like to learn how to calculate the chi-square cumulative probability based on the degrees of freedom and *C*-statistic, a la the handy tool on this [website](http://stattrek.com/online-calculator/chi-square.aspx).
+
+It's probably simple...
+
+The tool from this website calculates the probability of observing a value smaller than this as 0.63, indicating the probability of observing a larger value is 0.37.

### 2014-12-18 ch2_path_appendix1.Rmd
new appendix applying the d-sep test process to my path model

---
title: "Appendix 1 - Fitting a path model"
author: "Whalen Dillon"
date: "December 17, 2014"
output: html_document
---
----
#### This appendix details steps 1-4 of calculating the *C*-statistic for fitting a path model in a generalized multilevel context (Shipley 2009).

I am going to start by using the path model from Fig. 7 as an example.

**Step 1:** Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).

![Figure 7 path model](figures/ch2_path_model.png)

**Step 2:** List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.

Variable 1 | Variable 2
--- | ---
Elevation | Topography index
Elevation | Oak infection
Topography | Precipitation
Topography | Oak infection
Diversity | Precipitation
Diversity | Elevation

**Step 3:** For each of the *k* pairs of variables (*Xi,Xj*), list the set of other variables {*Z*} that are direct causes of either *Xi* or *Xj*. The pair of variables, (*Xi,Xj*), along with its conditioning set, {*Z*}, define an independence claim, (*Xi,Xj*) | {*Z*}, and the full set of the *k* independence claims defines the basis set, *Bu*. The basis set can be determined by using functions in the `R` package `ggm`.

First I will load the library and define the DAG using regression models and the `DAG` function. From the accompanying documentation: "The `DAG` function defines the adjacency matrix of a directed acyclic graph. An adjacency matrix is a square Boolean [e.g. 0/1] matrix that is equal to the number of nodes of the graph, with a one in a given position (i,j) if there is an arrow from i to j and zero otherwise. The row names of the adjacency matrix are the nodes of the DAG."
```{r define the DAG, eval=TRUE, echo=TRUE}
library(ggm)
fig7 <- DAG(precip ~ elev, 
    diversity ~ topo,
    temp ~ precip + elev + topo + diversity,
    oak.inf ~ leaf.ct + diversity + temp + precip,
    leaf.ct ~ diversity + temp + precip + elev + topo)
```
I can apply the `basiSet` function to this DAG to generate the full set of independence claims.
```{r get basis set, echo=TRUE, eval=TRUE}
bu_fig7 <- basiSet(fig7)
bu_fig7
```
I cleaned this list up into a table:

**The Basis Set of Independence Claims**

Independence Set | Conditioning Set
--- | ---
topo, elev | *empty*
topo, precip | elev  
topo, oak.inf | diversity, precip, temp, leaf.ct
diversity, elev | topo
diversity, precip | topo, elev
elev, oak.inf | diversity, precip, temp, leaf.ct

**Step 4:** For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*. In other words, perform a regression model using an appropriate method. Here I am only going to formulate the models because I am still in the process of calculating variables, so I do not have values for calculating probabilities of the coefficients. The predictor variable whose independence is being checked in each model is bolded - this is the variable for which the exact probability must be recorded. I also hazard an educated guess at the distribution family of the response variable for each model.

Independence Set | Conditioning Set | Model | Distribution
--- | --- | --- | ---
topo, elev | *empty* | topo ~ **elev** | *normal*
precip, topo | elev | precip ~ elev + **topo** | *normal*
topo, oak.inf | diversity, precip, temp, leaf.ct | oak.inf ~ diversity + precip + temp + leaf.ct + **topo** | *binomial*
diversity, elev | topo | diversity ~ topo + **elev** | *normal*
diversity, precip | topo, elev | diversity ~ topo + elev + **precip** | *normal*
elev, oak.inf | diversity, precip, temp, leaf.ct | oak.inf ~ diversity + precip + temp + leaf.ct + **oak.inf** | binomial

**Step 5:** Combine the *k* probabilities into the *C*-statistic:

$$
C = -2\Sigma_{i=1}^k ln(P_i)
$$ 
Compare this *C*-statistic to a $\chi^2$ distribution with 2*k* degrees of freedom. If the model fit according to this test is sufficient (i.e. p > 0.05), then the coefficients from these models may be applied to the appropriate paths.

Noticeably, symptomatic leaf count is not part of the independence claims. If the complete model has sufficient fit then it would be modeled as either a Poisson or negative binomial distribution, and the coefficients applied to the appropriate paths.

### 2014-12-30 ch2_path_appendix2.Rmd
renamed file

---
title: "Appendix 2 - Fitting proposed path models"
author: "Whalen Dillon"
date: "December 17, 2014"
output: html_document
---
----
#### This appendix details steps 1-4 of calculating the *C*-statistic for fitting a path model in a generalized multilevel context (Shipley 2009).

I go through these steps for the path models in Fig. 7 and Fig. 8 (included here)  to produce the regression equations that are indicated by the path diagram.

#### First the model in Fig. 7
**Step 1:** Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).

![Figure 7 path model](figures/ch2_path_model.png)

**Step 2:** List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.

**Independence Set - Fig. 7**

Variable 1 | Variable 2
--- | ---
Elevation | Topography index
Elevation | Oak infection
Topography | Precipitation
Topography | Oak infection
Diversity | Precipitation
Diversity | Elevation

**Step 3:** For each of the *k* pairs of variables (*Xi,Xj*), list the set of other variables {*Z*} that are direct causes of either *Xi* or *Xj*. The pair of variables, (*Xi,Xj*), along with its conditioning set, {*Z*}, define an independence claim, (*Xi,Xj*) | {*Z*}, and the full set of the *k* independence claims defines the basis set, *Bu*. The basis set can be determined by using functions in the `R` package `ggm`.

First I will load the library and define the DAG using regression models and the `DAG` function. From the accompanying documentation: "The `DAG` function defines the adjacency matrix of a directed acyclic graph. An adjacency matrix is a square Boolean [e.g. 0/1] matrix that is equal to the number of nodes of the graph, with a one in a given position (i,j) if there is an arrow from i to j and zero otherwise. The row names of the adjacency matrix are the nodes of the DAG."
```{r define fig7 DAG, eval=TRUE, echo=TRUE}
library(ggm)
fig7 <- DAG(precip ~ elev, 
    diversity ~ topo,
    temp ~ precip + elev + topo + diversity,
    oak.inf ~ leaf.ct + diversity + temp + precip,
    leaf.ct ~ diversity + temp + precip + elev + topo)
```
I can apply the `basiSet` function to this DAG to generate the full set of independence claims.
```{r get fig 7 basis set, echo=TRUE, eval=TRUE}
bu_fig7 <- basiSet(fig7)
bu_fig7
```
I cleaned this list up into a table:

**The Basis Set of Independence Claims - Fig. 7**

Independence Set | Conditioning Set
--- | ---
topo, elev | *empty*
topo, precip | elev  
topo, oak.inf | diversity, precip, temp, leaf.ct
diversity, elev | topo
diversity, precip | topo, elev
elev, oak.inf | diversity, precip, temp, leaf.ct

**Step 4:** For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*. In other words, perform a regression model using an appropriate method. Here I am only going to formulate the models because I am still in the process of calculating variables, so I do not have values for calculating probabilities of the coefficients. The predictor variable whose independence is being checked in each model is bolded - this is the variable for which the exact probability must be recorded. I also hazard an educated guess at the distribution family of the response variable for each model.

**Model Set - Fig. 7**

Independence Set | Conditioning Set | Model | Distribution
--- | --- | --- | ---
topo, elev | *empty* | topo ~ **elev** + 1\|plot | *?*
precip, topo | elev | precip ~ elev + **topo** + 1\|plot + 1\|year | *?*
topo, oak.inf | diversity, precip, temp, leaf.ct | oak.inf ~ diversity + precip + temp + leaf.ct + **topo** + 1\|plot + 1\|year | *binomial*
diversity, elev | topo | diversity ~ topo + **elev** + 1\|plot | *?*
diversity, precip | topo, elev | diversity ~ topo + elev + **precip** + 1\|plot | *?*
elev, oak.inf | diversity, precip, temp, leaf.ct | oak.inf ~ diversity + precip + temp + leaf.ct + **elev** + 1\|plot + 1\|year | *binomial*

Noticeably, symptomatic leaf count is not part of the independence claims. If the complete model has sufficient fit then it would be modeled as either a Poisson or negative binomial distribution, and the coefficients applied to the appropriate paths.

This next step is currently not possible to do because I am still conducting data manipulations to calculate variables.

**Step 5:** Combine the *k* probabilities into the *C*-statistic:

$$
C = -2\Sigma_{i=1}^k ln(P_i)
$$ 
Compare this *C*-statistic to a $\chi^2$ distribution with 2*k* degrees of freedom. If the model fit according to this test is sufficient (i.e. p > 0.05), then the coefficients from these models may be applied to the appropriate paths.

#### The model in Fig. 8

**Step 1:** Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).

![Figure 8 path model](figures/ch2_path_model2.png)

This model is indicates a direct modeling of a lag effect from precipitation, temperature, and symptomatic leaf count during the preceding year. Temperature, precipitation, and symptomatic leaf count of the current and preceding year influence the oak infection of the current year. Temperature and precipitation of the current and preceding year also influence the symptomatic leaf count of the current year. This is not explicitly clear in the currently simplified diagram, but will become so with the statement of the independence claims.

**Step 2:** List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.

**Independence Sets - Fig. 8**

Variable 1 | Variable 2
--- | ---
Elevation | Topography index
Elevation | Oak infection
Topography | Precipitation$_t$
Topography | Precipitation$_t-1$
Topography | Oak infection
Diversity | Precipitation$_t$
Diversity | Precipitation$_t-1$
Diversity | Elevation

**Step 3:** For each of the *k* pairs of variables (*Xi,Xj*), list the set of other variables {*Z*} that are direct causes of either *Xi* or *Xj*. The pair of variables, (*Xi,Xj*), along with its conditioning set, {*Z*}, define an independence claim, (*Xi,Xj*) | {*Z*}, and the full set of the *k* independence claims defines the basis set, *Bu*. The basis set can be determined by using functions in the `R` package `ggm`.

```{r define fig 8 DAG, eval=TRUE, echo=TRUE}
library(ggm)
fig8 <- DAG(precip1 ~ elev, precip2 ~ elev, 
    diversity ~ topo,
    temp1 ~ precip1 + elev + topo + diversity,
    temp2 ~ precip2 + elev + topo + diversity,
    oak.inf ~ leaf.ct1 + leaf.ct2 + diversity + temp1 + temp2 + precip1 + precip2,
    leaf.ct1 ~ diversity + temp1 + precip1 + temp2 + precip2 + leaf.ct2 + elev + topo,
    leaf.ct2 ~ diversity + temp2 + precip2 + elev + topo)
```
Apply the `basiSet` function to this DAG to generate the full set of independence claims.
```{r get fig 8 basis set, echo=TRUE, eval=TRUE}
bu_fig8 <- basiSet(fig8)
```

**The Basis Set of Independence Claims - Fig. 8**

Independence Set | Conditioning Set
--- | ---
topo, elev | *empty*
topo, precip1 | elev
topo, precip2 | elev
topo, oak.inf | diversity, precip1, temp1, leaf.ct1, precip2, temp2, leaf.ct2
diversity, elev | topo
diversity, precip1 | topo, elev
diversity, precip2 | topo, elev
elev, oak.inf | diversity, precip1, temp1, leaf.ct1, precip2, temp2, leaf.ct2
precip1, precip2 | elev
precip2, temp1 | elev, topo, diversity, precip1
temp2, precip1 | topo, diversity, elev, precip2
temp2, temp1 | topo, diversity, elev, precip2, precip1
leaf.ct2, precip1 | topo, diversity, elev, precip2, temp2
leaf.ct2, temp1 | topo, diversity, elev, precip2, temp2, precip1

**Step 4:** For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*.

**Model set - Fig. 8**

Independence Set | Conditioning Set | Model | Distribution
--- | --- | --- | ---
topo, elev | *empty* | topo ~ **elev** + 1|plot | *?*
precip1, topo | elev | precip1 ~ elev + **topo** + 1\|plot + 1\|year | *?*
precip2, topo | elev | precip2 ~ elev + **topo** + 1\|plot + 1\|year | *?*
topo, oak.inf | diversity, precip1, temp1, precip2, temp2, leaf.ct | oak.inf ~ diversity + precip1 + temp1 + precip2 + temp2 + leaf.ct + **topo** + 1\|plot + 1\|year | *binomial*
diversity, elev | topo | diversity ~ topo + **elev** + 1\|plot | *?*
diversity, precip1 | topo, elev | diversity ~ topo + elev + **precip1** + 1\|plot | *?*
diversity, precip2 | topo, elev | diversity ~ topo + elev + **precip2** + 1\|plot | *?*
elev, oak.inf | diversity, precip1, temp1, precip2, temp2, leaf.ct1, leaf.ct2 | oak.inf ~ diversity + precip1 + temp1 + leaf.ct1 + precip2 + temp2 + leaf.ct2 + **elev** + 1\|plot + 1\|year | *binomial*
precip1, precip2 | elev | precip1 ~ elev + **precip2** + 1\|plot + 1\|year | *?*
precip2, temp1 | elev, topo, diversity, precip1 | temp1 ~ elev + topo + diversity + precip1 + **precip2** + 1\|plot + 1\|year | *?*
temp2, precip1 | topo, diversity, elev, precip2 | temp2 ~ topo + diversity + elev + precip2 + **precip1** + 1\|plot + 1\|year | *?*
temp2, temp1 | topo, diversity, elev, precip2, precip1 | temp1 ~ topo + diversity + elev + precip2 + precip1 + **temp2** + 1\|plot + 1\|year | *?*
leaf.ct2, precip1 | topo, diversity, elev, precip2, temp2 | leaf.ct2 ~ topo + diversity + elev + precip2 + temp2 + **precip1** + 1\|plot + 1\|year | *Poisson*
leaf.ct2, temp1 | topo, diversity, elev, precip2, temp2 + precip1 | leaf.ct2 ~ topo + diversity + elev + precip2 + temp2 + precip1 + **temp1** + 1\|plot + 1\|year | *Poisson*

### 2014-12-30 analysis/repeated_measures.R
minor formatting

@@ -91,6 +91,7 @@ confint(m2)
 sigma(m0)
 sigma(m1)
 sigma(m2)
+
 #### Plot model profiles using `lattice` package
 library(lattice)
 xyplot(profile(m1), conf = c(50, 90, 95, 99)/100)

### 2015-01-21 ch2_path_appendix2.Rmd
exploratory code revisions

@@
 ---
-title: "Appendix 2 - Fitting proposed path models"
+title: "Appendix 2 - Fitting proposed path models using the C-statistic"
 author: "Whalen Dillon"
 date: "December 17, 2014"
-output: html_document
+output: word_document
 ---

### 2015-01-21 analysis/explore_data.R

@@ -7,34 +7,64 @@
       #This is only the biologically related field data
 stems <- read.csv("analysis/data/stem_summary_query_2004-2014.txt")
 head(stems)
+summary(stems)
 str(stems)
 
 
 
 ### Need to do some data cleaning and prep ####
 
-#### Create date variable as posix and integer/factor year variable ####
-library(lubridate)
+## Create date variable as posix and integer/factor year variable ###
 stems$Date <- as.character(stems$Date)
 class(stems$Date)
-stems$date <- gsub("/", "-",stems$Date)
-stems$date <- dmy_hms(stems$date)
-stems$year <- year(stems$date)
-
-
-#### Rename columns using `rename` funcion from `plyr` ####
+library(lubridate)
+stems$year <- year(as.Date(stems$Date, format = "%m/%d/%Y"))
+summary(stems)
+## Why are there NA's in the `year` variable?
 library(plyr)
-stems <- rename(stems, c("PlotID" = "plot", "Date" = "old_date"))
-stems <- rename(stems, c("ClusterID" = "cluster", "TagNumber" = "tag", 
-                          "SpeciesID" = "species", "StemStatus" = "status",
-                          "AliveClass" = "alive_class", 
-                          "DeadClass" = "dead_class", "SympLeafCount" = "slc", 
-                          "CankerPresent" = "canker", "DBH" = "dbh",
-                          "LideFoliarSymp" = "lide_lf_symp", 
-                          "SOD_Dead" = "sod_dead", "Location" = "location"))
+library(dplyr)
+stems_sub <- filter(stems, is.na(year))
+stems_sub2 <- filter(stems, PlotID == "SECRET02", SpeciesID == "QUAG", year == 2012)
+# So the 19 NA's for in the table are from the SECRET02 plot and belong to 2012. I am not sure why these did not query correctly in the database because the data is entered with the proper date...
+# I replaced the NA's for the `year` variable with 2012
+stems$year[is.na(stems$year)] <- 2012
+summary(stems)
+rm(stems_sub)
+rm(stems_sub2)
+
+## Currently unused code for string substitution and date calculation
+# stems$date <- gsub("/", "-",stems$Date)
+# stems$date <- dmy_hms(stems$date)
+# stems$year <- year(stems$date)
+
+## Replace all `-9999` codes with NA ###
+stems[stems == -9999] <- NA
+summary(stems)
+
+
[…]

### 2015-01-23 report_dbh_problems.Rmd
dbh corrections
new dbh data (will be overwritten during corrections), and a report on
the dbh problems that i have addressed

## Report on DBH problems
###### Whalen Dillon

This is summarizing changes I made to the Sonoma Access Database DBH measurements based on data exploration in the `explore_data.R` document. There were 100 stems that had a negative change in DBH of greater than 5 cm, and 62 of these were alive in the 2014 remeasurement. This reports on those that I could logically make corrections to, and so avoid the incredible dramatically shrinking tree. Stems that are not explicitly listed

#### These stems all had a negative change in DBH > 15 cm (14)
 - For stems 4797 & 4795 I calculated an estimate for the original DBH based on a linear relationship between measurements from 2012 and 2014 because the original measurements recorded were most likelyl circumference. I initially did a simple coversion from circumference to diameter, but this resulted in values that were substantially hire than the measurements in either 2012 or 2014. New values for 2009 will be: stem 4797: 7.8, stem 4795: 10.0

 - Replaced 2004 value (31.5) for stem 945 by treating that as teh circumference, so new initial diameter is 10

 - Replaced 2004 value (69.25) for stem 423b by treating that as the circumference, so new inital diameter is 22

 - Replaced 2014 value (27.7) for stem 2367 with the measurement from 2012 because the change was ~50% of the initial measurement

 - Replaced 2004 value (40.9) for stem 1347 by treating that as the circumference, so new initial diameter is 13

 - Assumed 2004 measurement for stem 4318 missed a decimal point, so new value is 3.4

 - Same as above for stem 4319

 - Replaced 2004 value (38.8) by treating that as circumference, so new value is 12.4. No written record on 2004 or 2005 data sheets.

### 2015-01-23 analysis/explore_data.R

@@ -3,13 +3,13 @@
 # Superspreaders chapter of dissertation      
 --------
 
-######Set working directory, read in all stem data, examine data structure ####
-      #This is only the biologically related field data
+#### Read in all stem data, examine data structure ####
+#This is only the biologically related field data
+
 stems <- read.csv("analysis/data/stem_summary_query_2004-2014.txt")
 head(stems)
 summary(stems)
 str(stems)
-
 
 
 ### Need to do some data cleaning and prep ####
@@ -64,9 +64,131 @@ summary(stems$status) # There are a lot of different `X` statuses, which I am no
 class(stems$status)
 View(filter(stems, status == ""))
 ## A good number of these stems needed an `X` status, but there were at least 8 that needed correcting in the database. The primary cause for the blank status appeared to be a duplication of the visit date for a single stem. Once one of these records was deleted the data appeared. So, I am going to export the query from the data base again and reassess the data with these corrections.
+stems <- read.csv("analysis/data/stem_summary_qry2_04-14.csv")
+summary(stems)
 
+stems$Date <- as.character(stems$Date)
+class(stems$Date)
+library(lubridate)
+stems$year <- year(as.Date(stems$Date, format = "%m/%d/%Y"))
+summary(stems)
+stems$year[is.na(stems$year)] <- 2012
+stems[stems == -9999] <- NA
+summary(stems)
 
+stems <- rename(stems, plot = PlotID)
+stems <- rename(stems, cluster = ClusterID, tag = TagNumber, 
+                species = SpeciesID, status = StemStatus, 
+                alive_class = AliveClass, dead_class= DeadClass,
+                slc = SympLeafCount, canker = CankerPresent, dbh = DBH,
+                lide_lf_symp = LideFoliarSymp, sod_dead = SOD_Dead, 
+                location = Location, notes = StemSymptomNotes)
+str(stems)
+summary(stems)
+summary(stems$species) # 5 host species listed, so this appears legit
[…]

### 2015-01-23 analysis/explore_data.R
dbh exploration

@@ -186,9 +186,24 @@ stems <- left_join(stems, dbh, by = c("tag","plot"))
 str(stems)
 summary(stems)
 filter(stems, tag == 2285)
-stems %>% select(plot, cluster, tag, species, delta_dbh, year, status) %>% 
+stems %>% select(plot, cluster, tag, species, delta_dbh, year, year_dbh1, status) %>% 
       group_by(plot, cluster, tag) %>% 
-      filter(delta_dbh < -5, year == 2014, status == "Alive") 
+      filter(delta_dbh < -15, year == 2014, status == "Alive")
+# I changed values for 9 of these 14 stems
+
+stems %>% select(plot, cluster, tag, species, delta_dbh, year, year_dbh1, status) %>% group_by(plot, cluster, tag) %>% 
+      filter(between(delta_dbh, -15, -10), year == 2014, status == "Alive")
+# I changed values for 7 of 9 stems
+
+stems %>% select(plot, cluster, tag, species, delta_dbh, year, year_dbh1, status) %>% group_by(plot, cluster, tag) %>% 
+      filter(between(delta_dbh, -10, -7), year == 2014, status == "Alive")
+# I changed values for 11 of 20 stems
+
+stems %>% select(plot, cluster, tag, species, delta_dbh, year, year_dbh1, status) %>% group_by(plot, cluster, tag) %>% 
+      filter(between(delta_dbh, -6.9, -5), year == 2014, status == "Alive")
+# I changed values for 6 of 18 stems
+
+
 
 ### ### ### ### ### ### ### ### ### ### ### ### ### ### #
 ### Older code that can and will likely be abandoned ####

### 2015-01-23 report_dbh_problems.Rmd
more dbh exploration

@@
 There were 100 stems that had a negative change in DBH of greater than 5 cm, and 62 of these were alive in the 2014 remeasurement.
-This reports on those that I could logically make corrections to, and so avoid the incredible dramatically shrinking tree.
+This reports on the stems still alive in 2014 that I could logically make corrections to, and so avoid the incredible dramatically shrinking tree.
 Stems that are not explicitly listed
 
-#### These stems all had a negative change in DBH > 15 cm (14)
+#### Stems with a negative change in DBH > 15 cm (14)
  - For stems 4797 & 4795 I calculated an estimate for the original DBH based on a linear relationship between measurements from 2012 and 2014 because the original measurements recorded were most likelyl circumference.
@@
  - Replaced 2004 value (38.8) by treating that as circumference, so new value is 12.4. No written record on 2004 or 2005 data sheets.
+
+I changed values for 9 of these 14 stems
+
+#### Stems with a negative change in DBH between 10 cm & 15 cm (9)
+ - Assumed 2006 value (31.5) for stem 6023 was the circumference, so new value is 10 - no written record
+ - Assumed 2004 value (26.5) for stem 808 was the circumference, so new value is 8.4 - no written record
+ - Value was entered incorrectly for stem 2755, corrected based on written record to 11.2
+ - Value for stem 1883 was corrected based on note about needed tag switching. The new DBH values are much closer to making sense
+ - Assumed 2004 value (29.0) for stem 1880 was the circumference, so new value is 9.2 - this is contrary to the written record
+ - Assumed 2004 value (14.25) for stem 689 was the circumference, so new value is 4.5 - no written record
+ - Assumed 2004 value (33.5) for stem 576 was the circumference, so new value is 10.7
+
+I changed values for 7 of 9 stems
+
+#### Stems with a negative change in DBH between 7 cm & 10 cm (20 stems)
+ - Value was entered incorrectly for stem 3010, corrected based on written record to 20.6
+ - Value was entered incorrectly for stem 3087, corrected based on written record to 12.1
+ - Value was entered incorrectly for stem 3398, corrected based on written record to 10.9
+ - Value in 2004 (29.92, really? wtf?) for stem 157 was corrected based on linear estimate between 2012 and 2014 values, so new value is 19.3.
+ - Changed 2004 value (24.3) for stem 4323 to 14.3 - assuming written recording error
+ - Assumed 2006 value (9.0) for stem 4590 was circumference, so new value is 2.8
+ - Assumed 2004 value (23.2) for stem 1692 was entered or recorded incorrectly, new value is 13.2 - no written record found (missing data sheet)
+ - Same as above for stem 1686, but value changed from 24.7 to 14.7
+ - No written record for stem 604, so back calculated DBH for 2004, new value is 21.5
+ - Back calculated 2004 value (31.5) for stem 654 to new value of 21.9 - no written record
+ - Assumed 2014 value (7.8) was incorrectly recorded, changed to 17.8
+
+I changed values for 11 of 20 stems
+
+#### Stems with negative change in DBH between 5 cm and 6.9 cm (18 stems)
+ - Assumed 2004 value (12.0) was circumference, so new value is 3.8
+ - Corrected data entry error for stem 3022, DBH is 17.3
+ - Assumed 2004 value (9.25) for stem 4534 was circumference, new value is 2.9
+ - Assumed 2004 value (17.5) for stem 231 was circumference, new value is 5.6 - no written record
+ - Back calculated 2004 value for stem 529 based on 2012 & 2014 records, it was 19.75, now is 14.0 (19.75 could also have been circumference) - no written record
+ - Same situation as above for stem 4403, value was 19.2, now 8.3
+
+ I changed values for 6 of 18 stems

### 2015-01-23 analysis/explore_data.R

@@ -90,105 +90,111 @@ summary(stems$status) # Now there are no blanks, 36864 Alive records, 2482 dead 
 
 stems <- as.tbl(stems) # coerce to dplyr style data table
 class(stems)
-
-library(ggplot2)
-qplot(year, sum(dbh), data = stems)
-
-dbh <- filter(stems, dbh>0, year == 2014)
-length(unique(dbh$year))
-
 length(unique(stems$tag))
 # 4452 unique tag numbers in the data
 
+
+## Subset DBH data to create a data frame of remeasured stems ####
 dbh2014 <- stems %>% filter(dbh>0, year == 2014)
 length(unique(dbh2014$tag))
 # 4158, indicating there are 294 stems with no dbh measured in 2014
 
 dbh2003 <- stems %>% select(plot, tag, dbh, year) %>% 
       filter(dbh>0, year == 2003) # these 2 are the same as in 2005
-rm(dbh2003)
+# 11 observations
 
 dbh2004 <- stems %>% select(plot, tag, dbh, year) %>%
-      filter(dbh>0, year == 2004, plot == "ANN02")
+      filter(dbh>0, year == 2004)
 # 3109 observations
+anti_join(dbh2003, dbh2004, by = "tag")
+dbh_a <- union(dbh2004, dbh2003)
+
 
 dbh2005 <- stems %>% select(plot, tag, dbh, year) %>%
       filter(dbh>0, year == 2005)
 # 712 observations
-anti_join(dbh2004, dbh2005, by = "tag") # this indicates that the dbh data for 2004 and 2005 are unique
-dbh1 <- union(dbh2004, dbh2005)
+anti_join(dbh2005, dbh_a, by = "tag") # this indicates that the dbh data for 2004 and 2005 are unique
+dbh_a <- union(dbh_a, anti_join(dbh2005, dbh_a, by = "tag"))
 
 dbh2006 <- stems %>% select(plot, tag, dbh, year) %>%
       filter(dbh>0, year == 2006)
-summary(anti_join(dbh2006, dbh1, by = "tag")) # 2004, 2005, 2006 are independent sets
-dbh1 <- union(dbh1, dbh2006)
+anti_join(dbh2006, dbh_a, by = "tag") # 2004, 2005, 2006 are independent sets
+dbh_a <- union(dbh_a, anti_join(dbh2006, dbh_a, by = "tag"))
 
 dbh2007 <- stems %>% select(plot, tag, dbh, year) %>%
[…]

### 2015-03-17 report_dbh_problems.Rmd
Added review of `alive` stems that showed negative growth

@@
 This reports on the stems still alive in 2014 that I could logically make corrections to, and so avoid the incredible dramatically shrinking tree.
-Stems that are not explicitly listed
 
@@
 I initially did a simple coversion from circumference to diameter, but this resulted in values that were substantially hire than the measurements in either 2012 or 2014.
-New values for 2009 will be: stem 4797: 7.8, stem 4795: 10.0
+New values for 2009 will be: stem 4797 = 7.8, stem 4795 = 10.0
 
- - Replaced 2004 value (31.5) for stem 945 by treating that as teh circumference, so new initial diameter is 10
+ - Replaced 2004 value (31.5) for stem 945 by treating that as the circumference, so new initial diameter is 10
 
@@
 
- I changed values for 6 of 18 stems
+I changed values for 6 of 18 stems.
+
+In total, there are now 30 stems with 'Alive' status in 2014 that still have negative growth values > 5cm, so I was able to make corrections for just over half of the 62.
+
+So I think this covers all the stems I was considering for errors due to shrinkage. I still need to examine the outliers on the positive growth side.
+
+#### These stems had > 20cm growth
+plot    cluster   tag   species year year_dbh1 status delta_dbh dbh1  dbh2
+BOUV01       2    1321    QUAG 2014      2005  Alive      22.1 98.7 120.8 *ok*        
+***BOUV02       4    1325    QUAG 2014      2005  Alive      23.0 63.4  86.4***
+BOWES02      2    423b    UMCA 2014      2004  Alive      31.2 22.0  53.2 *ok*       
+COOK02       3    472     UMCA 2014      2005  Alive      30.0 16.1  46.1 *ok*       
+***LAZY05       2    2073    QUKE 2014      2004  Alive      34.0 63.8  97.8***      
+PFEND02      2    4516    UMCA 2014      2004  Alive      27.1 30.5  57.6 *ok*
+SKILE01      4    3288    UMCA 2014      2004  Alive      24.4 18.7  43.1 *ok*       
+SKILE01      4    3290    UMCA 2014      2004  Alive      21.7 41.2  62.9 *ok*
+
+The two rows that are ***bold italic*** are stems that I definitely want to double check the entries on the original data sheets.

### 2015-03-17 analysis/explore_data.R
Code added for checking large positive DBH changes after correcting as many of the negative DBH changes as I legitimately could. There are still some major outliers.

@@ -212,7 +212,7 @@ stems %>% select(plot, cluster, tag, species, delta_dbh, year, year_dbh1, status
 # I changed values for 6 of 18 stems
 
 
-#### Create a plot-level data frame ####
+#### Create a plot-level data frame of infection status ####
 library(plyr)
 library(dplyr)
 
@@ -232,6 +232,181 @@ filter(plots, infected_bay_ct == 0)
 
 write.csv(plots, "analysis/data/plots_umca_infection.csv")
 # I sent this file to Francesco for informing the spread model
+
+#### Checking large positive DBH changes ####
+## So at this point I think I have corrected all the negative dbh change values that I legitimately could in the database. I have gone back and exported the query again and use this file below, recycling code from the beginning of this document.
+stems <- read.csv("analysis/data/Stem_Summary_negDBH_corrected_201402.csv")
+summary(stems)
+
+stems$Date <- as.character(stems$Date)
+class(stems$Date)
+library(lubridate)
+stems$year <- year(as.Date(stems$Date, format = "%m/%d/%Y"))
+summary(stems)
+library(plyr)
+library(dplyr)
+stems_sub <- filter(stems, is.na(year))
+stems_sub2 <- filter(stems, PlotID == "SECRET02", SpeciesID == "QUAG", year == 2012)
+
+stems$year[is.na(stems$year)] <- 2012
+summary(stems)
+stems[stems == -9999] <- NA
+summary(stems)
+
+library(dplyr)
+stems <- rename(stems, plot = PlotID)
+stems <- rename(stems, cluster = ClusterID, tag = TagNumber, 
+                species = SpeciesID, status = StemStatus, 
+                alive_class = AliveClass, dead_class= DeadClass,
+                slc = SympLeafCount, canker = CankerPresent, dbh = DBH,
+                lide_lf_symp = LideFoliarSymp, sod_dead = SOD_Dead, 
+                location = Location)
+str(stems)
+stems <- rename(stems, notes = StemSymptomNotes)
+summary(stems)
+
+
+#### Subset DBH data to create a data frame of remeasured stems: Round 2 ####
+dbh2014 <- stems %>% filter(dbh>0, year == 2014)
+length(unique(dbh2014$tag))
[…]

### 2015-03-18 data_subset_wininger.R
R script file where I am providing a subset of the data for Kerry Winiger at SSU

@@ -1,0 +1,23 @@
+## Data subset for Kerry Wininger, SSU
+
+### "Would you be able to do me the favor of passing along any and all tree-level data you have from FOP, JROTH, and MROTH from last year?" *last year is referring to 2014
+
+load("~/GitHub/superspreaders/.RData")
+load("~/GitHub/superspreaders/stems_plots.Rdata")
+str(stems)
+str(plots)
+str(dbh2014)
+
+### Filter stem data to subset of requested plots and year
+library(dplyr)
+
+kw_data <- stems %>%
+      filter(plot %in% c("FOP04","FOP324","FOP333","FOP340","JROTH01","JROTH02","JROTH03","JROTH04","JROTH05","MROTH01","MROTH02","MROTH03"), year == 2014)
+str(kw_data)
+summary(kw_data)
+
+### "Plot temperature and precipitation points would be good too. Thanks for sending along some of this already." The temperature data is ready to go easily enough at the hourly level, but I still have some work to do on the precipitation data.
+
+
+
+

### 2015-03-18 data_subset_wininger.R
Data subsets from Sonoma SOD project for Kerry Wininger. The 2014 stem level data for 12 plots, and hourly temperature data for the same 12 plots from 2003-2014.

@@ -11,13 +11,17 @@ str(dbh2014)
 ### Filter stem data to subset of requested plots and year
 library(dplyr)
 
-kw_data <- stems %>%
+kw_stem_data <- stems %>%
       filter(plot %in% c("FOP04","FOP324","FOP333","FOP340","JROTH01","JROTH02","JROTH03","JROTH04","JROTH05","MROTH01","MROTH02","MROTH03"), year == 2014)
-str(kw_data)
-summary(kw_data)
+str(kw_stem_data)
+summary(kw_stem_data)
+write.csv(kw_stem_data, "deliverables/kw_stem_data.csv")
 
 ### "Plot temperature and precipitation points would be good too. Thanks for sending along some of this already." The temperature data is ready to go easily enough at the hourly level, but I still have some work to do on the precipitation data.
 
-
-
-
+### Read in the long format temperature data, it's a big file so might take a little longer than usual
+long_ppt <- readRDS("~/Documents/microclimate_tonini/Temperature/eof_pred_long.rds")
+str(long_ppt)
+kw_temp_data <- long_ppt %>% select(-posix_time) %>% filter(plotid %in% c("FOP04","FOP324","FOP333","FOP340","JROTH01","JROTH02","JROTH03","JROTH04","JROTH05","MROTH01","MROTH02","MROTH03"))
+summary(kw_temp_data)
+write.csv(kw_temp_data, "deliverables/kw_temp_data.csv")

### 2015-03-31 analysis/explore_data.R
Deleted some duplicated code that from correcting negative DBH changes that I had copied to build upon for looking at positive DBH changes

@@ -387,6 +387,7 @@ stems %>% select(plot, cluster, tag, species, year, year_dbh1, status, delta_dbh
       group_by(plot, cluster, tag) %>% 
       filter(delta_dbh > 20, year == 2014, status == "Alive")
 # 8 stems, 
+
 stems %>% select(plot, tag, cluster, species, year, year_dbh1, status, dbh, delta_dbh, dbh1, dbh2, notes) %>% 
       group_by(plot, cluster, tag) %>% 
       filter(plot == "LAZY05", species == "QUKE", status == "Alive")
@@ -399,13 +400,6 @@ stems %>% select(plot, cluster, tag, species, delta_dbh, dbh1, dbh2, year, year_
       filter(between(delta_dbh, 10, 20), year == 2014, status == "Alive")
 # 39 stems
 
-stems %>% select(plot, cluster, tag, species, delta_dbh, year, year_dbh1, status) %>% group_by(plot, cluster, tag) %>% 
-      filter(between(delta_dbh, -10, -7), year == 2014, status == "Alive")
-# I changed values for 11 of 20 stems
-
-stems %>% select(plot, cluster, tag, species, delta_dbh, year, year_dbh1, status) %>% group_by(plot, cluster, tag) %>% 
-      filter(between(delta_dbh, -6.9, -5), year == 2014, status == "Alive")
-# I changed values for 6 of 18 stems
 
 
 

### 2015-04-03 report_dbh_problems.Rmd
Code to subset tagged stem DBH data and send along for use in FVS simulation. Typo correction in the report Rmd file. Minor line corrections and edits to add/change titles of some of the plots in the explore_data script.

@@
 #### Stems with a negative change in DBH > 15 cm (14)
- - For stems 4797 & 4795 I calculated an estimate for the original DBH based on a linear relationship between measurements from 2012 and 2014 because the original measurements recorded were most likelyl circumference.
+ - For stems 4797 & 4795 I calculated an estimate for the original DBH based on a linear relationship between measurements from 2012 and 2014 because the original measurements recorded were most likely circumference.
 I initially did a simple coversion from circumference to diameter, but this resulted in values that were substantially hire than the measurements in either 2012 or 2014.

### 2015-04-03 analysis/explore_data.R

@@ -350,7 +350,9 @@ library(ggplot2)
 qplot(dbh1, dbh2, data = dbh)
 qplot(delta_dbh, dbh2, data = dbh, color = year_dbh1)
 
-#Join dbh measure, remeasure, and change data to `stems` data frame
+# Join dbh measure, remeasure, and change data to `stems` data frame
+dbh <- read.csv("analysis/data/dbh_remeasures.csv")
+dbh <- select(dbh, -X)
 stems <- left_join(stems, dbh, by = c("tag","plot"))
 str(stems)
 summary(stems)
@@ -365,22 +367,25 @@ qplot(delta_dbh, dbh1, data = stems %>% select(plot, cluster, tag, species, delt
 # 1. "Have you tried looking at relative growth instead of the absolute increment growth?  Either DBH2/DBH1, or the instantaneous rate ln(DBH2/DBH1) / (time2-time1).  20 cm growth means a lot more to a 5 cm tree than a 50 cm tree… either way, it's still big.  I'd also consider this species by species, because I bet some grow much faster than others."
 stems$dbh2_dbh1 <- stems$dbh2/stems$dbh1
 stems$inst_grwth_rate <- log(stems$dbh2/stems$dbh1) / (stems$year_dbh2 - stems$year_dbh1)
-qplot(stems$dbh2_dbh1)
+stems$time_diff <- (stems$year_dbh2 - stems$year_dbh1)
+summary(stems)
+
+qplot(stems$dbh2_dbh1, main = "Relative Growth")
 
 qplot(dbh1, dbh2_dbh1, data = stems %>% select(plot, cluster, tag, species, dbh2_dbh1, dbh1, year, status) %>% 
             group_by(plot, cluster, tag) %>%
             filter(year == 2014, status == "Alive" | status == "Dead"), 
-      color = status)
+      color = status, main = "Change in DBH vs. original DBH")
 
 qplot(dbh1, inst_grwth_rate, data = stems %>% select(plot, cluster, tag, species, inst_grwth_rate, dbh1, year, status) %>% 
             group_by(plot, cluster, tag) %>%
             filter(year == 2014, status == "Alive" | status == "Dead"), 
-      color = status)
+      color = status, main = "Instantaneous Growth Rate vs. Original DBH")
 
 qplot(dbh1, inst_grwth_rate, data = stems %>% select(plot, cluster, tag, species, inst_grwth_rate, dbh1, year_dbh1, year, status) %>% 
[…]

### 2015-04-03 fvs_dbh_subset.R

@@ -1,0 +1,48 @@
+# Tagged stem DBH data for Jim to implement in FVS
+### This uses the data where I corrected some of the negative growth values. There are still some with negative growth, but I had no way to provide reasonable replacement values. I haven't finished checking the positive growth outliers yet.
+
+#### Read in data with corrected DBH ####
+stems <- read.csv("analysis/data/Stem_Summary_negDBH_corrected_201402.csv")
+summary(stems)
+
+stems$Date <- as.character(stems$Date)
+class(stems$Date)
+library(lubridate)
+stems$year <- year(as.Date(stems$Date, format = "%m/%d/%Y"))
+summary(stems)
+library(plyr)
+library(dplyr)
+stems_sub <- filter(stems, is.na(year))
+stems_sub2 <- filter(stems, PlotID == "SECRET02", SpeciesID == "QUAG", year == 2012)
+
+stems$year[is.na(stems$year)] <- 2012
+summary(stems)
+stems[stems == -9999] <- NA
+summary(stems)
+
+library(dplyr)
+stems <- rename(stems, plot = PlotID)
+stems <- rename(stems, cluster = ClusterID, tag = TagNumber, 
+                species = SpeciesID, status = StemStatus, 
+                alive_class = AliveClass, dead_class= DeadClass,
+                slc = SympLeafCount, canker = CankerPresent, dbh = DBH,
+                lide_lf_symp = LideFoliarSymp, sod_dead = SOD_Dead, 
+                location = Location)
+str(stems)
+stems <- rename(stems, notes = StemSymptomNotes)
+summary(stems)
+
+#### Subset and structure data
+library(dplyr)
+library(tidyr)
+fvs_dbh <- stems %>% select(plot, year, tag, species, status, dbh) %>%
+      filter(dbh > 0)
+fvs_dbh <- as.tbl(fvs_dbh)
+class(fvs_dbh)
+fvs_dbh
+fvs_dbh  %>% spread(species, dbh)
+fvs_dbh[5391:5392,]
+fvs_dbh[c(5413:5414,5424:5425),]
+fvs_dbh2 <- fvs_dbh[-c(5414,5424:5425,5392),]
+fvs_dbh2 <- fvs_dbh2 %>% spread(species, dbh)
+write.csv(fvs_dbh2, "deliverables/fvs_tagged_dbh.csv")

### 2015-05-06 analysis/explore_data.R
The "stem growth" data file has all the stems and some dbh variables applicable to stems that were remeasured in 2014. The "dbh_2014_remeasures" has the the first measurement of and the remeasurement in 2014 (only includes stems remeasured in 2014). Edits to the "explore_data" file include fixing some minor errors in the data due to data entry, examining the positive dbh changes for outliers (there are a few, but I don't think they are entirely unreasonable), calculating some dbh variables (growth metrics) at the stem level and exploring to see if these show any relationship to leaf count on b

@@ -88,7 +88,9 @@ stems <- rename(stems, cluster = ClusterID, tag = TagNumber,
 str(stems)
 summary(stems)
 summary(stems$species) # 5 host species listed, so this appears legit
-summary(stems$status) # Now there are no blanks, 36864 Alive records, 2482 dead records
+summary(stems$status) # Now there are no blanks, 36864 Alive records, 2482 dead records, 1 "Nearly dead" so changing that to "Alive"
+stems$status[stems$status=="Nearly dead"] <- "Alive"
+stems <- droplevels(stems)# drop unused factor levels
 
 stems <- as.tbl(stems) # coerce to dplyr style data table
 class(stems)
@@ -182,7 +184,8 @@ library(ggplot2)
 qplot(dbh1, dbh2, data = dbh)
 qplot(delta_dbh, dbh2, data = dbh, color = year_dbh1)
 
-#Join dbh measure, remeasure and change data to `stems` data frame
+#Join dbh measure, remeasure and change data to `stems` data frame ####
+dbh <- read.csv("analysis/data/dbh_2014_remeasures.csv")
 stems <- left_join(stems, dbh, by = c("tag","plot"))
 str(stems)
 summary(stems)
@@ -217,13 +220,15 @@ stems <- read.csv("analysis/data/all_stems_corrected.csv")
 library(dplyr)
 
 plots_umca <- stems %>%
-      select(plot, Date, species, slc, year, status) %>%
-      group_by(plot, year) %>%
+      select(plot, date, species, slc, year, status) %>%
+      group_by(plot, year, species) %>%
       filter(species == "UMCA", status == "Alive") %>%
       summarise(uninfected_bay_ct = length(which(slc==0)), infected_bay_ct = length(which(slc > 0)), tot_bay = length(species)) 
 
-plots_umca$infected <- ifelse(plots_umca$infected_bay_ct==0, 0, 1)
+# plots_umca$infected <- ifelse(plots_umca$infected_bay_ct==0, 0, 1)
 summary(plots_umca)
+plots_umca <- droplevels(plots_umca)
+
 
 length(which(plots_umca$ct_bay_NA > 0))
 filter(plots_umca, infected == 0)
@@ -231,6 +236,28 @@ filter(plots_umca, infected_bay_ct == 0)
 
 write.csv(plots_umca, "analysis/data/plots_umca_infection.csv")
[…]

### 2015-05-06 search4superspreaders.Rmd
Making a presentation on data exploration to try and identify superspreader individuals

---
output:
  slidy_presentation:
    font_adjustment: 1
    footer: Center for Geospatial Analytics at NCSU
    highlight: haddock
    keep_md: yes
  beamer_presentation:
    highlight: haddock
---

## Searching for Superspreaders

**Whalen Dillon**

May 7, 2015

<img src=http://cnr.ncsu.edu/geospatial/wp-content/uploads/sites/6/2014/07/cgaDynamicB2REV.png width="800px" />

## Emerging Infectious Diseases

New in a population, or new but rapidly increasing

![](presentations_images/ebola_lyme_whitenose.png)
<img src=presentations_images/global_shipping.png width="500px" />

## Disease Triangle

![](presentations_images/disease_triangle.png)

## Sudden Oak Death Triangle

![](presentations_images/sod_triangle.png)

## Sudden Oak Death Hosts & Environment

46 wildland host species - 5 key host tree species

<img src=presentations_images/sod_host_distribution.png />

## Pathogen Spillover

**“when disease epidemics are driven by transmission from an alternate reservoir host population” – Power & Mitchell (2004)**
- characteristic of multihost disease systems

![](presentations_images/sod_spillover.png)

## Superspreaders

Individual(s) that disproportionately contribute to disease transmission
 - potential mechanism for pathogen spillover

![](presentations_images/bay_superspreader.png)

## Searching for superspreaders

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

## Species and Locations as Superspreaders

Disease hotspots as "superspreaders"

![](presentations_images/superspreader_hotspots.png)

### 2015-05-06 search4superspreaders.Rmd
Add images to introduction slides. Write code for calculating the upper quantile values

@@
 **“when disease epidemics are driven by transmission from an alternate reservoir host population” – Power & Mitchell (2004)**
+
 - characteristic of multihost disease systems
@@
 Individual(s) that disproportionately contribute to disease transmission
+
 - potential mechanism for pathogen spillover
@@
 
-```{r}
-summary(cars)
+The 90%, 95%, and 99% quantile threshold values for each year
+```{r Look at potential outliers, echo=FALSE, message=FALSE, results='asis'}
+load("~/GitHub/superspreaders/stem_level_data.RData")
+library(plyr); library(dplyr); library(xtable)
+tab <- xtable(bay_laurel %>% select(plot, tag, slc, status, year) %>% 
+      filter(status == "Alive") %>%
+      group_by(year) %>% 
+      summarise(#slc80 = quantile(slc, 0.80, na.rm = T),
+                #slc85 = quantile(slc, 0.85, na.rm = T), 
+                slc90 = quantile(slc, 0.90, na.rm = T), 
+                slc95 = quantile(slc, 0.95, na.rm = T),
+                slc99 = quantile(slc, 0.99, na.rm = T)))
+print(tab, type = "html")
+```
+
+```{r Subset data to potential outliers each year, echo=FALSE}
+slc90q <- bay_laurel %>% group_by(year) %>% 
+      filter(year == 2004 & slc >= 82 | year == 2005 & slc >= 63 | 
+                   year == 2006 & slc >= 112 | year == 2007 & slc >= 72 | 
+                   year == 2008 & slc >= 62 | year == 2009 & slc >= 65 | 
+                   year == 2010 & slc >= 104 | year == 2011 & slc >= 166 |
+                   year == 2012 & slc >= 126 | year == 2014 & slc >= 93)
+
+slc95q <- bay_laurel %>% group_by(year) %>% 
+      filter(year == 2004 & slc >= 114 | year == 2005 & slc >= 88 | 
+                   year == 2006 & slc >= 130 | year == 2007 & slc >= 86 | 
+                   year == 2008 & slc >= 76 | year == 2009 & slc >= 83 | 
+                   year == 2010 & slc >= 139 | year == 2011 & slc >= 190 |
+                   year == 2012 & slc >= 155 | year == 2014 & slc >= 128)
+
+slc99q <- bay_laurel %>% group_by(year) %>%
+      filter(year == 2004 & slc >= 178 | year == 2005 & slc >= 127 | 
+                   year == 2006 & slc >= 174 | year == 2007 & slc >= 113 | 
+                   year == 2008 & slc >= 110 | year == 2009 & slc >= 119 | 
+                   year == 2010 & slc >= 191 | year == 2011 & slc >= 214 |
+                   year == 2012 & slc >= 200 | year == 2014 & slc >= 181)
+
 ```
@@
 ```{r, echo=FALSE}
-plot(cars)
+
 ```

### 2015-05-08 search4superspreaders.Rmd
Created table of 90th, 95th, 99th quantile thresholds.  Plotted all 99th quantile stems, subsets of all that were in the 99th 2+ years, and all in the 99th in 3+ years. Ended presentation on hotspots as superspreaders.

--- 2015-05-06 search4superspreaders.Rmd
@@
   slidy_presentation:
+    fig_width: 8
+    fig_height: 5.5
     font_adjustment: 1
@@
     highlight: haddock
+    incremental: yes
     keep_md: yes
+  ioslides_presentation:
+    highlight: haddock
+    incremental: yes
   beamer_presentation:
@@
 
-<img src=http://cnr.ncsu.edu/geospatial/wp-content/uploads/sites/6/2014/07/cgaDynamicB2REV.png width="800px" />
+<img src=http://cnr.ncsu.edu/geospatial/wp-content/uploads/sites/6/2014/07/cgaDynamicB2REV.png width="800px" align="center" />
 
@@
 
-## Sudden Oak Death Triangle
+## Sudden Oak Death Disease Triangle
 
@@
 
-**“when disease epidemics are driven by transmission from an alternate reservoir host population” – Power & Mitchell (2004)**
-- characteristic of multihost disease systems
+**“when disease epidemics are driven by transmission from an alternate reservoir host population”**
+
+-- Power & Mitchell (2004) *Pathogen spillover in disease epidemics*
 
 ![](presentations_images/sod_spillover.png)
+
+A common characteristic of multihost disease systems
 
@@
 Individual(s) that disproportionately contribute to disease transmission
- - potential mechanism for pathogen spillover
+
+ -- potential mechanism for pathogen spillover
 
@@
 
+```{r load data & libraries, echo=FALSE, message=FALSE}
+load("~/GitHub/superspreaders/stem_level_data.RData")
+library(plyr); library(dplyr); library(xtable)
+```
+```{r subset to stems w/slc, echo=FALSE}
+slc <- bay_laurel %>% select(plot, tag, slc, date, status, year) %>% 
+      filter(status == "Alive" & slc != "NA")
+#summary(slc)
+slc <- droplevels(slc)
+```
+
+Number of bay stems in the study alive during at least one sampling year: **`r length(unique(slc$tag))`**
+
 ## Searching for superspreaders
 
-```{r}
-summary(cars)
+The 90%, 95%, and 99% quantile threshold values for each year
+```{r Look at potential outliers, echo=FALSE, message=FALSE, results='asis', out.extra=''}
+tab <- (bay_laurel %>% select(plot, tag, slc, date, status, year) %>% 
+      filter(status == "Alive") %>%
+      group_by(year) %>% 
+      summarise(#slc80 = quantile(slc, 0.80, na.rm = T),
+                #slc85 = quantile(slc, 0.85, na.rm = T), 
+                slc90 = quantile(slc, 0.90, na.rm = T), 
+                slc95 = quantile(slc, 0.95, na.rm = T),
+                slc99 = quantile(slc, 0.99, na.rm = T)))
+tab <- xtable(tab)
+print(tab, type = "html")
 ```
 
-## Slide with Plot
+## Searching for superspreaders
+```{r Plot slc99 by year, message=FALSE, echo=FALSE}
+library(ggplot2)
+g <- ggplot(tab, aes(year, slc99))
+g + geom_bar(stat = "identity", color = "blue", fill = "skyblue") + 
+      theme_bw() + theme(legend.position = "none") +
+      ggtitle("99th Percentiles of Symptomatic Leaf Counts by Year") +
+      ylab("Symptomatic Leaf Count")
+```
 
-```{r, echo=FALSE}
-plot(cars)
+```{r Subset data to potential outliers each year, echo=FALSE}
+slc90q <- bay_laurel %>% group_by(year) %>% 
+      filter(year == 2004 & slc >= 82 | year == 2005 & slc >= 63 | 
+                   year == 2006 & slc >= 112 | year == 2007 & slc >= 72 | 
+                   year == 2008 & slc >= 62 | year == 2009 & slc >= 65 | 
+                   year == 2010 & slc >= 104 | year == 2011 & slc >= 166 |
+                   year == 2012 & slc >= 126 | year == 2014 & slc >= 93)
+slc95q <- bay_laurel %>% group_by(year) %>% 
+      filter(year == 2004 & slc >= 114 | year == 2005 & slc >= 88 | 
+                   year == 2006 & slc >= 130 | year == 2007 & slc >= 86 | 
+                   year == 2008 & slc >= 76 | year == 2009 & slc >= 83 | 
+                   year == 2010 & slc >= 139 | year == 2011 & slc >= 190 |
+                   year == 2012 & slc >= 155 | year == 2014 & slc >= 128)
+slc99q <- bay_laurel %>% group_by(year) %>%
+      filter(year == 2004 & slc >= 178 | year == 2005 & slc >= 127 | 
+                   year == 2006 & slc >= 174 | year == 2007 & slc >= 113 | 
+                   year == 2008 & slc >= 110 | year == 2009 & slc >= 119 | 
+                   year == 2010 & slc >= 191 | year == 2011 & slc >= 214 |
+                   year == 2012 & slc >= 200 | year == 2014 & slc >= 181)
+```
+
+## Searching for Superspreaders - Stems in the 99th Percentile
+```{r subset 99th to 5 variables, echo=FALSE}
+slc99q <- slc99q %>% select(plot, tag, year, date, slc)
+```
+Number of observations in 99th percentiles: **`r length(slc99q$tag)`**
+
+```{r Plot slc99 data by year, echo=FALSE}
+qplot(date, slc, data = slc99q, geom = "jitter", color = tag) + 
+      geom_line(aes(color = tag)) +
+      theme_bw() + theme(legend.position = "none") +
+      ylab("Symptomatic Leaf Count") + 
+      ggtitle("Symptomatic Leaf Counts of Stems in 99th Percentile Each Year")
+slc99q <- ungroup(slc99q)
+```
+
+Number of unique stems in 99th percentiles: **`r length(unique(slc99q$tag))`**
+
+## Stems in the 99th Percentile 2+ Years
+```{r subset stems in 99th 2+ years,}
+d <- subset(slc99q, duplicated(slc99q$tag) | duplicated(slc99q$tag, fromLast = T))
+d <- arrange(d, tag, year)
+d
+```
+
+## Stems in the 99th Percentile 2+ Years
+```{r plot stems in 99th 2+ yrs, echo=FALSE}
+qplot(date, slc, data = d, color = tag, geom = "jitter") + 
+      geom_line(aes(color = tag)) +
+      #geom_text(aes(label = tag)) +
+      theme_bw() + theme(legend.position = "none") +
+      ylab("Symptomatic Leaf Count") +
+      ggtitle("Stems in the 99th Percentiles 2+ Years")
+```
+
+Number of unique stems in the 99th percentile 2+ years: **`r length(unique(d$tag))`**
+
+## Stems in the 99th Percentile 3+ Years
+```{r echo=FALSE}
+dt <- table(d$tag)
+d2 <- subset(d, tag %in% names(dt[dt > 2]))
+```
+
+Number of unique stems in the 99th percentile 3+ years: **`r length(unique(d2$tag))`**
+
+```{r echo=FALSE, eval=FALSE}
+#summary(d)
+#str(d)
+#d <- droplevels(d)
+#length(unique(d$plot))
+#length(unique(d$tag))
+
+dt <- table(d$tag)
+#dt > 2
+#dt[dt > 2]
+#names(dt[dt > 2])
+#d$tag %in% names(dt[dt > 2])
+d2 <- subset(d, tag %in% names(dt[dt > 2]))
+#d2
+length(unique(d2$tag))
+```
+```{r Plot stems 3+ yrs 99th, echo=FALSE}
+qplot(date, slc, data = d2, color = tag:plot, geom = "jitter") + 
+      geom_line(aes(color = tag:plot)) +
+      #geom_text(aes(label = plot)) +
+      theme_bw() +
+      ylab("Symptomatic Leaf Count") +
+      ggtitle("Stems in 99th Percentiles 3+ Years")
 ```
@@
 
-Disease hotspots as "superspreaders"
+More symptoms due to:
+
+ - individual characteristics
+ - environmental characteristics
+ - pathogen characteristics
 
 ![](presentations_images/superspreader_hotspots.png)
+
+Disease hotspots as "superspreaders"

### 2015-05-08 analysis/repeated_measures.R
Updated to use the most recently corrected data for bay laurel. Began adding assumptions of the linear mixed effects model.

@@ -1,45 +1,42 @@
+
 ## Repeated measures analysis ####
 
-stems <- read.csv("analysis/data/all_stems.csv")
+#### Read in data for all stems that has been processed in "explore_data.R"
+stems <- read.csv("analysis/data/stems_growth.csv")
 str(stems)
 summary(stems)
 
-### Change the assigned missing data value to NA across entire data frame
-stems[stems == -9999] <- NA
 
-library(plyr)
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-
+library(plyr); library(dplyr); library(tidyr); library(ggplot2)
 names(stems)
 stems <- stems %>%
-      select(-X, -date, -year)
-class(stems$old_date)
-stems$old_date <- as.Date(stems$old_date, format = "%m/%d/%Y")
-stems$year <- as.factor(year(stems$old_date))
-class(stems$year)
+      select(-X)# Remove the row names variable
+stems$date <- as.Date(stems$date)
+stems$year <- as.factor(stems$year)
 
 ### Create bay laurel only data frame ####
 umca <- as.tbl(stems) %>%
-      select(plot, cluster, tag, species, status, slc, dbh, year) %>%
-      filter(species == "UMCA")
+      select(plot, cluster, tag, species, status, slc, dbh, delta_dbh, dbh2_1_ratio, year, date) %>%
+      filter(species == "UMCA", status == "Alive" | status == "Dead")
 summary(umca)
+umca <- droplevels(umca)
 
 # Develop plot level summaries data frame and join to stems data frame
 umca <- left_join(umca, plot_umca <- umca %>%
       select(plot, cluster, tag, status, slc, year) %>% 
       group_by(plot, year) %>%
-      filter(slc >= 0, status == "Alive") %>%
+      filter(slc != "NA" & status == "Alive") %>%
       summarise(avg_slc = mean(slc), tot_slc = sum(slc), 
                 live_dens = length(tag)), 
       by = c("plot", "year"))
-summary(umca)
-umca <- droplevels(umca)
+summary(umca)# One NA in the summarized leaf count variables
+filter(umca, is.na(avg_slc))
+filter(umca, plot == "YAHNG02")# Looks like leaf count probably not done/recorded
 
-###Linear Mixed Effects Method ####
[…]

### 2015-05-08 SH_data/UMCA_stats.R
Deleted some unncessary carriage returns

@@ -206,8 +206,6 @@ x4=  x4 + geom_boxplot() +
 ggsave(plot=x4, filename="D:\\SOCO\\SOCO_space_time\\Figures\\annual_log10slc_boxplot_95CI.png", height=4, width=5)
 
 
-
-
 #A ribbon might look nicer:
 #ggplot(mtcars, aes(x=gear, y=mpg, group=as.factor(cyl), 
 #colour=as.factor(cyl))) + 
@@ -215,14 +213,3 @@ ggsave(plot=x4, filename="D:\\SOCO\\SOCO_space_time\\Figures\\annual_log10slc_bo
 #"ribbon", alpha = 0.6)+ 
 # stat_summary(fun.y=mean, geom="point") + 
 # stat_summary(fun.y=mean, geom="line") 
-
-
-
-
-
-
-
-
-
-
-

### 2015-05-09 analysis/repeated_measures.R
Changed project settings (Rproj file) to "No" for restoring workspace and saving workspace. Changed some aesthetic formatting of the code, trying to improve readability. Deleted the results included in Sarah's code. Noted the assumptions of the linear mixed effects model and reason for using random effects. Build models with 1 or 2 random intercepts, and an interaction between bay density (number of stems) and year. But, this should probably be with plot b/c thats where the variation would be.

@@ -1,7 +1,7 @@
 
 ## Repeated measures analysis ####
 
-#### Read in data for all stems that has been processed in "explore_data.R"
+## Read in data for all stems that has been processed in "explore_data.R"
 stems <- read.csv("analysis/data/stems_growth.csv")
 str(stems)
 summary(stems)
@@ -14,7 +14,7 @@ stems <- stems %>%
 stems$date <- as.Date(stems$date)
 stems$year <- as.factor(stems$year)
 
-### Create bay laurel only data frame ####
+#### Create bay laurel only data frame ####
 umca <- as.tbl(stems) %>%
       select(plot, cluster, tag, species, status, slc, dbh, delta_dbh, dbh2_1_ratio, year, date) %>%
       filter(species == "UMCA", status == "Alive" | status == "Dead")
@@ -35,46 +35,83 @@ filter(umca, plot == "YAHNG02")# Looks like leaf count probably not done/recorde
 
 #### Plot Level Repeated Measures Analysis Using Multi-level Mixed-Effects Regression ####
 
-## Linear Mixed Effects Method ####
-#### A couple of lines from Sarah's code to work from
+#### A couple of lines from Sarah's code to work from ####
+# lmer.null <- lmer(log(cum.slc + 1) ~ 1 + (1 | year))
+# lmer1 <- lmer(log(cum.slc + 1) ~ log(umca.ct + 1) + (1 | year))
 # anova(lmer.null, lmer1)
-# Models:
-# lmer.null: log(cum.slc + 1) ~ 1 + (1 | year)
-# lmer1: log(cum.slc + 1) ~ log(umca.ct + 1) + (1 | year)
-#          Df    AIC    BIC  logLik deviance  Chisq Chi Df Pr(>Chisq)    
-# lmer.null  3 4378.9 4394.0 -2186.4   4372.9                             
-# lmer1      4 3861.9 3882.1 -1927.0   3853.9 518.97      1  < 2.2e-16 ***
-#  ---
-#  Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
-
 # lmer2 <- lmer(log(cum.slc + 1) ~ log(umca.ct + 1) + (1 + log(umca.ct+1)|year), na.action=na.omit)
 # anova(lmer1, lmer2)
-# Data: 
-#  Models:
-#  lmer1: log(cum.slc + 1) ~ log(umca.ct + 1) + (1 | year)
[…]

### 2015-05-09 analysis/repeated_measures.R
Added models that use a densit:plot interaction and year random affect, i.e. switching the interaction random effect term.

@@ -65,32 +65,41 @@ library(lme4)
 
 ## Simplest models, with only one random effect (random intercept)
 m0 <- lmer(tot_slc ~ 1 + (1 | year), data = plot_umca)
+m0a <- lmer(tot_slc ~ 1 + (1 | plot), data = plot_umca)
 ## Add a second random effect where plot is nested within year
-m0a <- lmer(tot_slc ~ 1 + (1 | year) + (1 | plot), data = plot_umca)
+m0b <- lmer(tot_slc ~ 1 + (1 | year) + (1 | plot), data = plot_umca)
+ 
 ## One predictor + random effect (intercept)
 m1 <- lmer(tot_slc ~ live_dens + (1 | year), data = plot_umca)
+m1a <- lmer(tot_slc ~ live_dens + (1 | plot), data = plot_umca)
 ## One predictor + 2 random effects
-m1a <- lmer(tot_slc ~ live_dens + (1 | year) + (1 | plot), data = plot_umca)
+m1b <- lmer(tot_slc ~ live_dens + (1 | year) + (1 | plot), data = plot_umca)
 
 # Random slope, random intercept model
 ## The random effects term in the random slope, random intercept model accounts for the interaction between the predictor variable(s) and group in addition to the systematic variation between groups.
 ##    - similar to linear regresstion with nominal & continuous variables, but requires a lot of parameters to be estimated if there are many groups (1 for each)
 m2 <- lmer(tot_slc ~ live_dens + (1 + live_dens | year), data = plot_umca)
+m2a <- lmer(tot_slc ~ live_dens + (1 | year) + (1 + live_dens | plot), data = plot_umca)
 ## add the second random effect
-m2a <- lmer(tot_slc ~ live_dens + (1 + live_dens | year) + (1 | plot), data = plot_umca)
+m2b <- lmer(tot_slc ~ live_dens + (1 + live_dens | year) + (1 | plot), data = plot_umca)
+m2c <- lmer(tot_slc ~ live_dens + (1 | year) + (1 + live_dens | plot), data = plot_umca)
 
 ## Compare all models using Chi-square test of ANOVA
 ## Refits model(s) with maximum likelihood instead of REML
-anova(m0, m1, m2, m0a, m1a, m2a)
[…]

### 2015-05-21 search4superspreaders.Rmd
Added titles to some of the code chunks

@@
 ## Stems in the 99th Percentile 3+ Years
-```{r echo=FALSE}
+```{r Stems repeated 3+ years, echo=FALSE}
 dt <- table(d$tag)
@@
 
-```{r echo=FALSE, eval=FALSE}
+```{r Conditional subsetting data, echo=FALSE, eval=FALSE}
 #summary(d)

### 2015-05-22 analysis/explore_data.R
Exported the all stems summary from the database with the HostID field and edited code accordingly, i.e. replacing the 'ClusterID' with 'HostID' variable

@@ -260,20 +260,21 @@ tmp <- plot_qusp %>%
       dcast(year + plot ~ species, value.var = "infected_oak_ct")
 
 #### Checking large positive DBH changes ####
-## So at this point I think I have corrected all the negative dbh change values that I legitimately could in the database. I have gone back and exported the query again and use this file below, recycling code from the beginning of this document.
+## So at this point I think I have corrected all the negative dbh change values that I legitimately could in the database. I have gone back and exported the query again and use this file below, and then recycling code from the beginning of this document.
 stems <- read.csv("analysis/data/Stem_Summary_negDBH_corrected_201402.csv")
 summary(stems)
+str(stems)
 
 stems$Date <- as.character(stems$Date)
 class(stems$Date)
-library(lubridate)
-stems$year <- year(as.Date(stems$Date, format = "%m/%d/%Y"))
 stems$date <- as.Date(stems$Date, format = "%m/%d/%Y")
+library(lubridate)
+stems$year <- year(stems$date)
 summary(stems)
 library(plyr)
 library(dplyr)
 filter(stems, is.na(year))
-filter(stems, PlotID == "SECRET02", SpeciesID == "QUAG", year == 2012)
+filter(stems, PlotID == "SECRET02", SpeciesID == "QUAG")
 
 stems$year[is.na(stems$year)] <- 2012
 filter(stems, is.na(date))
@@ -283,7 +284,7 @@ stems[stems == -9999] <- NA
 summary(stems)
 
 stems <- rename(stems, plot = PlotID)
-stems <- rename(stems, cluster = ClusterID, tag = TagNumber, 
+stems <- rename(stems, cluster = HostID, tag = TagNumber, 
                 species = SpeciesID, status = StemStatus, 
                 alive_class = AliveClass, dead_class= DeadClass,
                 slc = SympLeafCount, canker = CankerPresent, dbh = DBH,
@@ -298,6 +299,9 @@ class(stems)
 stems <- as.tbl(stems)
 
 #### Subset DBH data to create a data frame of remeasured stems: Round 2 ####
+
+detach("package:lubridate", unload=TRUE)# b/c it has `union` function masking the one from `dplyr` that apparently behaves a little differently, creating a list instead of a data frame
+
 dbh2014 <- stems %>% filter(dbh>0, year == 2014)
 length(unique(dbh2014$tag))
[…]

### 2015-06-11 analysis/plotting.R
Changed commenting notation to be markdown compatible, i.e. inserting #' at the beginnings of lines

@@ -1,4 +1,4 @@
-## A lot of the plotting and code in this was written by Sarah Haas. I have made some additions and modifications while exploring ways to quantify what may constitute a superspreader individual.
+#' A lot of the plotting and code in this was written by Sarah Haas. I have made some additions and modifications while exploring ways to quantify what may constitute a superspreader individual.
 
 library(plyr)
 library(dplyr)
@@ -11,14 +11,12 @@ head(umca_stem)
 summary(umca_stem %>% group_by(year, tag))
 summary(umca_stem %>% group_by(year))
 
-# Some observations repeated in this file, so subset to unique
+#' Some observations repeated in this file, so subset to unique
 umca_stem <- unique(umca_stem[ , 1:8] )
 
-
-## Get specific quantile values for SLC for each year ####
-### If the same stems are consistently in the upper quantile then they may be superspreaders because they support more infections compared to the population. 
-
-# The `plyr` method with sample code from <http://stackoverflow.com/questions/5473537/how-to-calculate-95th-percentile-of-values-with-grouping-variable-in-r-or-excel>
+#' Get specific quantile values for SLC for each year ####
+#' If the same stems are consistently in the upper quantile then they may be superspreaders because they support more infections compared to the population. 
+#' The `plyr` method with sample code from <http://stackoverflow.com/questions/5473537/how-to-calculate-95th-percentile-of-values-with-grouping-variable-in-r-or-excel>
 library(plyr)
 #Random seed
 set.seed(42)
@@ -71,8 +69,8 @@ p2 <- p2 + geom_point(aes(color = plot.id), position = "jitter") +
 p2 + geom_text(aes(label = tag, color = ))
 
 
-## "Seasonality" and symptoms ####
-### Note that this has to be done within a sample season, so it's a bad assessment of true seasonality of symptoms.
+#' "Seasonality" and symptoms ####
+#' Note that this has to be done within a sample season, so it's a bad assessment of true seasonality of symptoms. Jennifer Davidson's work addressed this pretty well, so it would be a proper reference.
 library(dplyr)
 library(ggplot2)
[…]

### 2015-06-11 analysis/repeated_measures.R
Changed commenting format to #' to be compatible with markdown rendering

@@ -1,27 +1,29 @@
-
-## Repeated measures analysis ####
-
-## Read in data for all stems that has been processed in "explore_data.R"
+---
+#' Repeated measures analysis
+---
+#' Read in data for all stems that has been processed in "explore_data.R" ####
 stems <- read.csv("analysis/data/stems_growth.csv")
 str(stems)
 summary(stems)
 
-
+#' Load packages ####
 library(plyr); library(dplyr); library(tidyr); library(ggplot2)
+
+#' Examine stems data ####
 names(stems)
 stems <- stems %>%
       select(-X)# Remove the row names variable
 stems$date <- as.Date(stems$date)
 stems$year <- as.factor(stems$year)
 
-#### Create bay laurel only data frame ####
+#' Create bay laurel only data frame ####
 umca <- as.tbl(stems) %>%
       select(plot, cluster, tag, species, status, slc, dbh, delta_dbh, dbh2_1_ratio, year, date) %>%
       filter(species == "UMCA", status == "Alive" | status == "Dead")
 summary(umca)
 umca <- droplevels(umca)
 
-# Develop plot level summaries data frame and join to stems data frame
+#' Develop plot level summaries data frame and join to stems data frame ####
 umca <- left_join(umca, plot_umca <- umca %>%
       select(plot, cluster, tag, status, slc, year) %>% 
       group_by(plot, year) %>%
@@ -33,16 +35,16 @@ summary(umca)# One NA in the summarized leaf count variables
 filter(umca, is.na(avg_slc))
 filter(umca, plot == "YAHNG02")# Looks like leaf count probably not done/recorded
 
-#### Plot Level Repeated Measures Analysis Using Multi-level Mixed-Effects Regression ####
+#' Plot Level Repeated Measures Analysis Using Multi-level Mixed-Effects Regression ####
 
-#### A couple of lines from Sarah's code to work from ####
+#' A couple of lines from Sarah's code to work from
 # lmer.null <- lmer(log(cum.slc + 1) ~ 1 + (1 | year))
 # lmer1 <- lmer(log(cum.slc + 1) ~ log(umca.ct + 1) + (1 | year))
 # anova(lmer.null, lmer1)
 # lmer2 <- lmer(log(cum.slc + 1) ~ log(umca.ct + 1) + (1 + log(umca.ct+1)|year), na.action=na.omit)
 # anova(lmer1, lmer2)
 
-## Linear Mixed Effects Method ####
[…]

### 2015-06-11 oak_infection_model.Rmd
some off the cuff modeling of oak infection using bay laurel counts as predictors. trying to get an educated guess to inform the Tangible Landscape SOD Spread Model

---
title: "Predicting Oak Infection"
author: "Whalen Dillon"
date: "June 11, 2015"
output: html_document
---

I'm trying to define an empirical relationship between oak infection and bay laurel density, though I think this is something we can get from Sarah's survival analysis.

```{r load data}
load("stems_plots.RData")
summary(stems)
summary(plot_qusp)
summary(plots_umca)
```

```{r load packages, message=FALSE}
library(plyr); library(dplyr); library(lme4); library(ggplot2)
```

```{r manipulate data}
names(stems)
plot_aggregates <- stems %>%
      select(plot, species, status, slc, canker, sod_dead, dbh, year) %>%
      group_by(year, plot) %>% 
      summarise(slc_total = sum(slc, na.rm=T), slc_avg = mean(slc, na.rm=T), oak_inf = sum(canker, na.rm=T), oak_dead = sum(sod_dead, na.rm=T), dbh_total = sum(dbh, na.rm=T), dbh_avg = mean(dbh, na.rm=T))

summary(plot_aggregates)

summary(plot_qusp)
plot_aggregates <- left_join(plot_aggregates, select(plot_qusp, year, plot, uninfected_oak_ct, infected_oak_ct, tot_oak), by = c("year", "plot"))
summary(plots_umca)
plot_aggregates <- left_join(plot_aggregates, select(plots_umca, year, plot, uninfected_bay_ct, infected_bay_ct, tot_bay))
```

```{r model oak infection}
library(lme4)
oak2012 <- filter(plot_aggregates, year == 2012, infected_oak_ct > 0)
summary(oak2012)
plot(oak2012$tot_bay, oak2012$infected_oak_ct)

summary(filter(plot_aggregates, infected_oak_ct == 1))

ggplot(plot_aggregates, aes(tot_bay, infected_oak_ct)) +
      geom_point() +
      facet_grid(year ~ .)

oak.m1 <- glmer(cbind(infected_oak_ct, uninfected_oak_ct) ~ I(tot_bay/225) + (1 | plot), data = plot_aggregates, family = "binomial")
summary(oak.m1)

oak.m2 <- glmer(cbind(infected_oak_ct, uninfected_oak_ct) ~ log1p(tot_bay) + (1 | plot), data = plot_aggregates, family = "binomial")
summary(oak.m2)
```

### 2015-06-11 analysis/explore_data.R
changed line for looking at a temporary data frame

@@ -258,6 +258,7 @@ gather(plot_qusp, inf_status, count, ends_with("ct")) %>%
 library(reshape2)
 tmp <- plot_qusp %>%
       dcast(year + plot ~ species, value.var = "infected_oak_ct")
+head(tmp)
 
 #### Checking large positive DBH changes ####
 ## So at this point I think I have corrected all the negative dbh change values that I legitimately could in the database. I have gone back and exported the query again and use this file below, and then recycling code from the beginning of this document.

### 2015-06-24 iale2015_sem.Rmd
Starting analysis for IALE presentation. Trying to develop a path analysis to relate environmental variability, especially weather, to pathogen spillover

---
title: "IALE 2015 Analysis"
author: "Whalen Dillon"
date: "June 15, 2015"
output: html_document
---

```{r load data}
# Ecological data: leaf counts, infection
load("~/GitHub/superspreaders/stems_plots.RData")

# Physical environment: topography
library(rgdal)
plots_env <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
str(plots_env)
plot(plots_env, axes=T)
plots_env <- plots_env@data

# Temperature: rainy season hour summaries
plots_temps <- read.csv("analysis/data/rainy_season_temps.csv")
summary(plots_temps)
plots_temps <- select(plots_temps, -X)
```

Join the plot level environmental data (geographic/topographic & temperatures)
```{r join plot data, message=FALSE}
library(dplyr)
plots_env <- left_join(plots_env, plots_temps, by = c("PLOT_ID"="plotid"))
summary(plots_env)
str(plots_env)
plots_env <- plots_env %>% select(-cat) %>% rename(easting = X, northing = Y, plotid = PLOT_ID)
```

Investigating `NAs` in the data set
```{r investigate NAs}
summary(plots_env)
filter(plots_env, is.na(sample_year))

# BUSH01 plot included. I think we have limited ecological data for this plot, and no microclimate data. So, I am going to exclude it from all the data sets
plots_env <- filter(plots_env, plotid != "BUSH01")
stems <- rename(stems, plotid = plot) %>% filter(plotid != "BUSH01")

summary(plots_env) # A number of observations for above 25 threshold have NAs
# I believe that these are in fact truly zero observations. NAs occurred for the 2010, 2011, and 2014 seasons, and the only NAs in the data set are associated with the above 25 threshold. 
plots_env %>% filter(is.na(hrs_abv25))
# I am replacing the NAs with `0`, assuming true observations.
plots_env[is.na(plots_env)] <- 0
```

Make a pairs plot of the continuous variables, i.e. everything except `plotid`.
```{r pairs plot}
# Helper functions for `pairs`
panel.hist <- function(x, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(usr[1:2], 0, 1.5) )
      h <- hist(x, plot = FALSE)
      breaks <- h$breaks; nB <- length(breaks)
      y <- h$counts; y <- y/max(y)
      rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(0, 1, 0, 1))
      r <- abs(cor(x, y, use = "na.or.complete"))
      txt <- format(c(r, 0.123456789), digits = digits)[1]
      txt <- paste0(prefix, txt)
      if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
      text(0.5, 0.5, txt, cex = cex.cor * r)
}

pairs(select(plots_env, -plotid, -sample_year), lower.panel = panel.cor, diag.panel = panel.hist)

plots_env$avg_psi <- rowMeans(cbind(plots_env$psiSpr, plots_env$psiSum, plots_env$psiWin))
pairs(select(plots_env, -plotid, -sample_year, -starts_with("psi")), lower.panel = panel.cor, diag.panel = panel.hist)
```

### 2015-06-26 analysis/explore_data.R
deleted an old and outdated code chunk

@@ -556,40 +556,6 @@ summarise(bay_laurel %>% filter(year == 2005), mdoy = max(doy))
 filter(bay_laurel, year == 2005, doy == 146)
 
 
-### ### ### ### ### ### ### ### ### ### ### ### ### ### #
-### Older code that can and will likely be abandoned ####
-### ### ### ### ### ### ### ### ### ### ### ### ### ### # 
-library(foreign)
-plot_dat1<-read.dbf("analysis/data/UMCA_slc_dbh_04-12.dbf")
-str(plot_dat1)
-summary(plot_dat1)
-head(plot_dat1)
-
-plot_dat1$avg_slc_04_12<-with(plot_dat1, rowMeans(cbind(avgslc04,avgslc05,avgslc06,avgslc07,avgslc08,avgslc09,avgslc10,avgslc11,avgslc12)))
-
-years<-c("2004","2005","2006","2007","2008","2009","2010","2011","2012")
-with(plot_dat1, boxplot(avgslc04,avgslc05,avgslc06,avgslc07,avgslc08,
-                        avgslc09,avgslc10,avgslc11,avgslc12,names=years,xlab="Year",ylab="Mean SLC"))
-
-
-library(reshape)
-plot_dat1<-rename(plot_dat1,c(PLOT_ID="plotid"))
-library(dplyr)
-plot_dat1<-left_join(plot_dat1,plot_202_rain,by="plotid")
-
-avg_slc_year<-with(plot_dat1, colMeans(cbind(avgslc04,avgslc05,avgslc06,avgslc07,avgslc08,
-                                             avgslc09,avgslc10,avgslc11,avgslc12)))
-
-
-plot(years,avg_slc_year,xlab="Year",ylab="Mean SLC")
-
-avg_rain_year<-with(plot_202_rain, colMeans(cbind(amou2004,amou2005,amou2006,amou2007,amou2008,
-                                                  amou2009,amou2010,amou2011,amou2012)))
-rainy_days<-with(plot_202_rain, colMeans(cbind(freq2004,freq2005,freq2006,freq2007,freq2008,
-                                               freq2009,freq2010,freq2011,freq2012)))
-
-df_rain_slc<-as.data.frame(rbind(avg_rain_year,avg_slc_year))
-
 ###### These may be useful functions for reference #### 
 matplot(cbind(avg_rain_year,avg_slc_year),type="b")
 matplot(cbind(rainy_days,avg_slc_year),type="b")

### 2015-06-27 explore_veg_data.Rmd
add new file where i explore the data from vegetation transects

---
title: "Explore Vegetation Data"
author: "Whalen Dillon"
date: "June 26, 2015"
output: html_document
---

In this document I am exploring vegetation data from the Sonoma County plot network. These data were collected usin vegetation transects across each of the plots circa plot establishment (2003/2004) and in 2011. The DBH of alternate and non-host species were also recorded in 2006, 2012, and 2014.

Read in general vegetation data related to each plot
```{r read in data}
canopy_cover <- read.csv("analysis/data/canopy_cover_qry.csv")
veg_transects <- rbind(read.csv("analysis/data/veg_transects_subset_qry_pre2011.csv"), read.csv("analysis/data/veg_transects_2011.csv"))
alt_sp_dbh <- read.csv("analysis/data/alt_species_dbh.csv")
```

What am I looking at here?
```{r examine canopy cover data}
library(dplyr)
str(canopy_cover)
head(canopy_cover)
# I am interested in the date and total estimated cover fore each plot
canopy_cover <- canopy_cover %>% select(PlotID, Date, PctTotalCover)
canopy_cover$Date <- as.Date(as.character(canopy_cover$Date), format = "%m/%d/%Y")
canopy_cover <- rename(canopy_cover, plotid = PlotID, date = Date, pct_cvr = PctTotalCover)
summary(canopy_cover)
filter(canopy_cover, is.na(date))
filter(canopy_cover, plotid == "APGAR01") # appears to be a duplicate entry...
canopy_cover <- filter(canopy_cover, is.na(date) == "FALSE")

library(lubridate)
canopy_cover$year <- year(canopy_cover$date)
```

The subset veg transect data `veg2` is approximately half of the original data points, selecting every other sample point along the transect. About twice as many points were sampled during plot establishment compared to the 2011 sampling
```{r examine veg transect data}
str(veg_transects)
veg_transects$Date <- as.Date(as.character(veg_transects$Date), format = "%m/%d/%Y")
veg_transects$year <- year(veg_transects$Date)
summary(veg_transects)
head(veg_transects, 20)
unique(veg_transects$year)

dim(filter(veg_transects, year != 2011))

unique(filter(veg_transects, year == 2005 | year == 2004 | year == 2003)$PlotID) # 205 plot IDs
unique(filter(veg_transects, year == 2011)$PlotID) # 199 plot IDs

```

### 2015-06-30 explore_veg_data.Rmd
split veg transect data into overstory and understory data; fixed duplicate species due to different naming conventions and typos. summarized data to number of times each species was observed in each plot and calculated a ratio of number of points it was observed at to the total number of points sampled along the transects

@@
 
-What am I looking at here?
+## Canopy Cover - Densiometers
+
 ```{r examine canopy cover data}
-library(dplyr)
+library(plyr); library(dplyr)
 str(canopy_cover)
@@
 dim(filter(veg_transects, year != 2011))
-
 unique(filter(veg_transects, year == 2005 | year == 2004 | year == 2003)$PlotID) # 205 plot IDs
 unique(filter(veg_transects, year == 2011)$PlotID) # 199 plot IDs
+veg_transects$PlotID <- tolower(veg_transects$PlotID)
 
+detach("package:lubridate", unload = T)
 ```
+
+## Overstory Vegetation
+
+```{r overstory veg}
+veg_os <- select(veg_transects, -UnderstorySp)
+summary(veg_os)
+
+# veg_us <- veg_us %>% (UnderstorySp != "")
+# summary(veg_us)
+
+# veg_os <- veg_os %>% filter(OverstorySp != "")
+veg_os[veg_os == -9999.00] <- NA
+filter(veg_os, is.na(OverstorySp))
+#veg_os <- na.omit(veg_os)
+summary(veg_os) # Percent coverage max is greater than 100
+filter(veg_os, PercentCoverage > 100)
+veg_os$PercentCoverage[veg_os$PercentCoverage > 100] <- 80
+summary(veg_os)
+
+veg_os <- as.tbl(veg_os)
+(veg_os)
+veg_os$OverstorySp <- tolower(veg_os$OverstorySp)
+unique(veg_os$OverstorySp)
+veg_os$OverstorySp[veg_os$OverstorySp == "arcto sp."] <- "arcto"
+veg_os$OverstorySp[veg_os$OverstorySp == "quke x quag"] <- "quag x quke"
+veg_os$OverstorySp[veg_os$OverstorySp == "none"] <- "open"
+veg_os$OverstorySp[veg_os$OverstorySp == "visp"] <- "vinca sp."
+veg_os$OverstorySp[veg_os$OverstorySp == ""] <- NA
+filter(veg_os, OverstorySp == "uq1")
+unique((filter(veg_os, PlotID == "sugar04", year == 2004))$OverstorySp)
+unique((filter(veg_os, PlotID == "sugar04", year == 2011))$OverstorySp)
+# This hints at the "uq1" from 2004 being the hybrid recorded in 2011. Note that the "open" value will be dropped later, but right now is being used as a place holder.
+veg_os$OverstorySp[veg_os$OverstorySp == "uq1"] <- "quke x quwi"
+# I'm also making the assumption of the "quwi" recorded in 2004 also actually being a hybrid because we didn't record a "quwi" in 2011
+veg_os$OverstorySp[veg_os$PlotID == "sugar04" & veg_os$year == 2004 & veg_os$OverstorySp == "quwi"] <- "quke x quwi"
+unique(veg_os$OverstorySp)
+# What's the deal with "unknown sp." & "unknown sp. 1"?
+filter(veg_os, OverstorySp == "unknown sp.")
+filter(veg_os, OverstorySp == "unknown sp. 1")
+# I think I can assume that the "unknowns" are the same within a plot, but probably different across the plot network.
+So, I can use them for calculating richness or diversity within a plot, but should probably exclude them from any other metric.
+
+# I want to get this arranged so that I have the number of observations for each species in each plot
+
+veg_os_summ <- left_join(veg_os %>% filter(OverstorySp != "OPEN") %>% 
+      group_by(PlotID, year, OverstorySp) %>% 
+      summarise(ptct = length(OverstorySp)),
+      veg_os %>% group_by(PlotID, year) %>% summarise(totpts = length(unique(PtNum)) * length(unique(Transect))),
+      by = c("PlotID", "year"))
+summary(veg_os_summ)
+```
+
+Why is the minimum for `totpts` 45 and the maximum 80? All *should* be 60, though I would allow for a little wiggle room. If I can't figure anything out then I'm going to call it good enough
+```{r overstory outliers}
+filter(veg_os_summ, totpts == 45)
+filter(veg_os_summ, PlotID == "ann25")
+(unique(filter(veg_os, PlotID == "ann25", year == 2004)$Transect))
+(unique(filter(veg_os, PlotID == "ann25", year == 2004)$PtNum))
+(unique(filter(veg_transects, PlotID == "ann25", year == 2004)$PtNum))
+
+filter(veg_os_summ, PlotID == "mutte02")
+unique(filter(veg_os, PlotID == "mutte02", year == 2004)$Transect)
+unique(filter(veg_os, PlotID == "mutte02", year == 2011)$Transect)
+(unique(filter(veg_os, PlotID == "mutte02", year == 2004)$PtNum))
+(unique(filter(veg_transects, PlotID == "mutte02", year == 2004)$PtNum))
+(unique(filter(veg_os, PlotID == "mutte02", year == 2011)$PtNum))
+(unique(filter(veg_transects, PlotID == "mutte02", year == 2011)$PtNum))
+
+filter(veg_os_summ, PlotID == "sugar32")
+unique(filter(veg_os, PlotID == "sugar32", year == 2004)$Transect)
+unique(filter(veg_os, PlotID == "sugar32", year == 2004)$PtNum)
+# I don't see much I can do about these.
+
+filter(veg_os_summ, totpts == 80)
+filter(veg_os_summ, PlotID == "jlsp05")
+# The ratios from this sampling are nearly the same (within < 1%), so that should be fine.
+```
+
+Here I am adding a column that calculates the proportional coverage for each species based on the number of points they were observed in throughout the plot and the total number of points sampled in the plot.
+```{r calculate overstory proportianal coverage}
+mutate(veg_os_summ, pcov = ptct/totpts)
+```
+
+## Understory Vegetation
+
+Now on to working through the understory veg data.
+I think this will actually be the more important because I will have DBH measurements for most of the species that would be observed in the overstory.
+Though there will as always be a few exceptions, e.g.
+TODI.
+```{r understory veg}
+veg_us <- select(veg_transects, -OverstorySp, -PercentCoverage)
+summary(veg_us)
+
+veg_us <- as.tbl(veg_us)
+veg_us
+veg_us$UnderstorySp <- tolower(veg_us$UnderstorySp)
+summary(veg_us)
+unique(veg_us$UnderstorySp)
+# summary(filter(veg_us, UnderstorySp == "None"))
+# summary(filter(veg_us, UnderstorySp == "OPEN"))
+veg_us[veg_us == -9999] <- NA
+veg_us$UnderstorySp[veg_us$UnderstorySp == "none"] <- "open"
+veg_us$UnderstorySp[veg_us$UnderstorySp == "arcto sp."] <- "arcto"
+filter(veg_us, UnderstorySp == "vasp")
+
+unique(filter(veg_us, PlotID == "ann33", year == 2005)$UnderstorySp)
+unique(filter(veg_us, PlotID == "ann33", year == 2011)$UnderstorySp)
+veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "losp"] <- "lohi"
+veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "syal"] <- "symo"
+veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "roca"] <- "rogy"
+
+(unique(veg_us$UnderstorySp))
+
+filter(veg_us, UnderstorySp == "bacspp") # Assuming this is a Baccharis species
+veg_us$UnderstorySp[veg_us$UnderstorySp == "bacspp"] <- "bapi"
+
+filter(veg_us, UnderstorySp == " to") # Maybe it's supposed to say "arcto" or "todi"?
+veg_us$UnderstorySp[veg_us$UnderstorySp == " to"] <- "todi"
+filter(veg_us, UnderstorySp == "4") # this is probably a typo; it should be "umca"
+veg_us$UnderstorySp[veg_us$UnderstorySp == "4"] <- "umca"
+
+veg_us$UnderstorySp[veg_us$UnderstorySp == "quke x quag"] <- "quag x quke"
+
+filter(veg_us, UnderstorySp == "losp")
+summary(filter(veg_us, UnderstorySp == "losp"))
+unique(filter(veg_us, UnderstorySp == "losp")$PlotID)
+unique(filter(veg_us, PlotID == "ann14", year == 2011)$UnderstorySp)
+unique(filter(veg_us, PlotID == "ann14", year == 2003)$UnderstorySp)
+# I wonder if all the "losp" we did in 2011 are reallly just "lohi"
+unique(filter(veg_us, year < 2011)$UnderstorySp)
+unique(filter(veg_us, year >= 2011)$UnderstorySp)
+filter(veg_us, year < 2011, UnderstorySp == "losp")
+filter(veg_us, year < 2011, UnderstorySp == "lonicera sp.")
+# Changing all to lohi
+veg_us$UnderstorySp[veg_us$UnderstorySp == "losp"] <- "lohi"
+
+filter(veg_us, UnderstorySp == "quag x")
+unique(filter(veg_us, PlotID == "mcneil01", year == 2005)$UnderstorySp)
+unique(filter(veg_us, PlotID == "mcneil01", year == 2011)$UnderstorySp)
+# McNeil01 was decommissioned prior to 2011, so not sure what the "quag x" was.
+But if this is in fact a quag-hybrid, then it there should probably be an adult tagged stem.
+Going to check the database...this plot only had adult UMCA tagged.
+So, I'm going to assume this was just a (perhaps funny looking) quag seedling/sapling.
+veg_us$UnderstorySp[veg_us$UnderstorySp == "quag x"] <- "quag"
+
+unique((veg_us$UnderstorySp))
+filter(veg_us, UnderstorySp == "todye") # assume that this should be "todi"
+veg_us$UnderstorySp[veg_us$UnderstorySp == "todye"] <- "todi"
+
+filter(veg_us, UnderstorySp == "arsp?") # assume "arcto" b/c ARME would be pretty obvious for ID
+veg_us$UnderstorySp[veg_us$UnderstorySp == "arsp?"] <- "arcto"
+
+filter(veg_us, UnderstorySp == "unk1")
+filter(veg_us, UnderstorySp == "unknown sp. 1")
+veg_us$UnderstorySp[veg_us$UnderstorySp == "unk1"] <- "unknown sp. 1"
+veg_us$UnderstorySp[veg_us$UnderstorySp == "unk2"] <- "unknown sp. 2"
+
+filter(veg_us, UnderstorySp == "uq1")
+filter(veg_us, UnderstorySp == "us#1")
+veg_us$UnderstorySp[veg_us$UnderstorySp == "uq1" | veg_us$UnderstorySp == "us#1" | veg_us$UnderstorySp == "unknown sp."] <- "unknown sp. 1"
+
+veg_us$UnderstorySp[veg_us$UnderstorySp == "vinca sp"] <- "vinca sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "lusp" | veg_us$UnderstorySp == "lupinus sp"] <- "lupine sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "vasp"] <- "vaccinium sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "gasp"] <- "garrya sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "qusp"] <- "quercus sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "risp" | veg_us$UnderstorySp == "ribes sp"] <- "ribes sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "prunus sp"] <- "prunus sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "rosa" | veg_us$UnderstorySp == "rosp"] <- "rose sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "rubu sp."] <- "rusp"
+veg_us$UnderstorySp[veg_us$UnderstorySp == "rhsp"] <- "rhamnus sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "garry sp."] <- "garrya sp."
+
+filter(veg_us, UnderstorySp == "asca") # I suspect this was supposed to be "aeca", even thou "asca" is written on the data sheet
+filter(veg_us, PlotID == "bouv03", PtNum == 6)
+unique(filter(veg_us, PlotID == "bouv03", year == 2005)$UnderstorySp)
+unique(filter(veg_us, PlotID == "bouv03", year == 2011)$UnderstorySp)
+veg_us$UnderstorySp[veg_us$UnderstorySp == "asca"] <- "aeca"
+```
+
+Okay, I think I have solved all (most?) of the duplicate species issues. Now to summarise the understory vegetation data similar to the overstory data.
+```{r summarise veg understory data}
+veg_us_summ <- left_join(veg_us %>% filter(UnderstorySp != "", UnderstorySp != "open") %>% 
+      group_by(PlotID, year, UnderstorySp) %>% 
+      summarise(ptct = length(UnderstorySp)),
+      veg_us %>% group_by(PlotID, year) %>% 
+            summarise(totpts = length(unique(PtNum)) * length(unique(Transect))),
+      by = c("PlotID", "year"))
+# filter(veg_us, PlotID == "ANN05", year == 2011)
+
+veg_us_summ <- mutate(veg_us_summ, pcov = ptct/totpts)
+veg_us_summ
+summary(veg_us_summ)
+unique(veg_us_summ$UnderstorySp)
+
+veg_us_2011 <- filter(veg_us_summ, year == 2011)
+veg_us_2005 <- filter(veg_us_summ, year < 2011)
+```

### 2015-06-30 explore_veg_data.Rmd
exported understory vegetation for 2005 and 2011 - complete sample years

@@
 veg_us_2005 <- filter(veg_us_summ, year < 2011)
+
+write.csv(veg_us_2005, "analysis/data/understory_veg_2005.csv")
+write.csv(veg_us_2011, "analysis/data/understory_veg_2011.csv")
 ```

### 2015-06-30 explore_veg_data.Rmd
Added chunk to summarise the untagged species dbh by plot, species, year, and alive/dead status. Calculated counts, and average and total dbh

@@
 ```
+
+Now I need to examine the data for alternate and non-host species DBH. These are the stems that are untagged and meet a 5cm DBH threshold.
+
+```{r untagged species}
+summary(alt_sp_dbh)
+library(lubridate)
+alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
+filter(alt_sp_dbh, is.na(date)) # this should be 5/22/2012
+alt_sp_dbh$Date[alt_sp_dbh$Date == ""] <- "5/22/2012 0:00:00"
+alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
+summary(alt_sp_dbh)
+alt_sp_dbh$year <- year(alt_sp_dbh$date)
+alt_sp_dbh <- alt_sp_dbh %>% select(-PlotVisit_PlotVisitID, -Date) %>% 
+      rename(dbh = Measure)
+unique(as.character(alt_sp_dbh$Species))
+
+alt_sp_dbh$Species[alt_sp_dbh$Species == "PSME"] <- "psme"
+alt_sp_dbh$Species <- tolower(alt_sp_dbh$Species)
+summary(alt_sp_dbh)
+unique(alt_sp_dbh$Species)
+unique(alt_sp_dbh$Dead)
+alt_sp_dbh$Dead <- ifelse(alt_sp_dbh$Dead == 0, "no", "yes")
+
+alt_sp_live <- filter(alt_sp_dbh, Dead == "no")
+alt_sp_dead <- filter(alt_sp_dbh, Dead == "yes")
+
+alt_sp_summ <- left_join(
+      # First data frame
+      alt_sp_live %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
+      summarise(live_count = length(Species), live_avg_dbh = mean(dbh), live_tot_dbh = sum(dbh)),
+      # Second data frame
+      alt_sp_dead %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
+      summarise(dead_count = length(Species), dead_avg_dbh = mean(dbh), dead_tot_dbh = sum(dbh))
+)
+
+write.csv(alt_sp_summ, "analysis/data/untagged_dbh_summary.csv")
+```

### 2015-07-03 iale2015_sem.Rmd
Calculated plot level leaf count variables, joined plot-level environmental; calculated understory richness from vegetation transects, and overstory richness from tagged and untagged stem DBH data and saved this into a "richness" data frame. Calculated multiple DBH variables for tagged & untagged stems, combined those data frames together, and calculated relative density, relative dominance, and importance value for each species.

@@
 load("~/GitHub/superspreaders/stems_plots.RData")
+# Vegetation data
+veg_us_2005 <- read.csv("analysis/data/understory_veg_2005.csv")
+veg_us_2011 <- read.csv("analysis/data/understory_veg_2011.csv")
+# Untagged dbh data
+untagged_dbh <- read.csv("analysis/data/untagged_dbh_summary.csv")
 
@@
 library(rgdal)
-plots_env <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
-str(plots_env)
-plot(plots_env, axes=T)
-plots_env <- plots_env@data
+plots_shp <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
+summary(plots_shp)
+plot(plots_shp, axes=T)
+plots_env <- plots_shp@data
 
 # Temperature: rainy season hour summaries
+library(plyr); library(dplyr)
 plots_temps <- read.csv("analysis/data/rainy_season_temps.csv")
@@
 ```{r join plot data, message=FALSE}
-library(dplyr)
 plots_env <- left_join(plots_env, plots_temps, by = c("PLOT_ID"="plotid"))
@@
 
-Investigating `NAs` in the data set
+Investigating `NAs` in the `plots_env` data set
 ```{r investigate NAs}
@@
 
-Make a pairs plot of the continuous variables, i.e. everything except `plotid`.
+Examine the bay laurel data at individual and plot levels
+```{r examine umca data}
+summary(plots_umca)
+# This only has number of infected, uninfected, and total, which must apply to living stems only. I want to analyze leaf counts.
+summary(bay_laurel)
+# This has stem-level data on bay laurel.
+```
+
+I can use this to calculate plot level leaf count variables (sum, average) for each plot and year
+```{r calculate plot level leaf counts}
+plots_umca <- left_join(plots_umca, bay_laurel %>% filter(status == "Alive") %>% 
+      group_by(plot, year) %>% 
+      summarise(
+            avg_lfct = mean(slc, na.rm = T),
+            tot_lfct = sum(slc, na.rm = T)),
+      by = c("plot", "year"))
+summary(plots_umca)
+```
+
+```{r examine lfct NAs}
+filter(plots_umca, is.na(avg_lfct))
+# I'm not sure why this wasn't excluded from the calculations with the na.rm = T option
+filter(bay_laurel, plot == "YAHNG02", status == "Alive")
+# I'm removing it manually from the `plots_umca` data frame
+plots_umca <- plots_umca %>% filter(avg_lfct != "NaN")
+summary(plots_umca)
+```
+
+So now I have a plot-level of UMCA infection levels for each year. I want to join this with some of the environmental data.
+
+```{r join env and umca data}
+summary(plots_env)
+plots_env$avg_psi <- rowMeans(cbind(plots_env$psiSpr, plots_env$psiSum, plots_env$psiWin))
+summary(plots_umca)
+# plot == plotid, year == sample_year
+plots_umca <- left_join(select(plots_umca, -species), plots_env, by = c("plot" = "plotid", "year" = "sample_year"))
+summary(plots_umca)
+filter(plots_umca, is.na(easting))
+# NAs are due to no microclimate data for year 2003, and abandoned plots BUSH01 & SUGAR02
+```
+
+Now I also want to calculate some plot-level community variables from the vegetation data. First, understory richness from the vegetation transects.
+```{r calculate understory richness}
+summary(veg_us_2005)
+veg_us_2005 <- as.tbl(select(veg_us_2005, -X))
+summary(veg_us_2011)
+veg_us_2011 <- as.tbl(select(veg_us_2011, -X))
+veg_us_2011
+veg_us_2005
+
+veg <- rbind(veg_us_2005, veg_us_2011)
+veg
+unique(veg$year)
+veg$year[veg$year < 2011] <- 2005 # Creates the two most complete veg sampling
+veg <- veg %>% rename(us_species = UnderstorySp, plotid = PlotID)
+richness <- veg %>% group_by(plotid, year) %>% summarise(us_rich = length(unique(us_species)))
+richness
+```
+
+Next, overstory richness from the DBH data for tagged & untagged stems.
+
+```{r calculate overstory richness}
+summary(untagged_dbh)
+untagged_dbh <- as.tbl(untagged_dbh %>% select(-X) %>% rename(plotid = PlotID, species = Species))
+untagged_dbh
+unique(untagged_dbh$year)
+untagged_dbh$year[untagged_dbh$year < 2014] <- 2005 # Create two most complete groups
+untagged_dbh$plotid <- tolower(untagged_dbh$plotid)
+
+untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species)))
+
+stems <- as.tbl(stems)
+summary(stems)
+unique(stems$status)
+stems <- stems %>% select(-X) %>% filter(status == "Alive" | status == "Dead")
+unique(stems$species)
+stems$species <- tolower(stems$species)
+stems$species[stems$species == "quke x quag"] <- "quag x quke"
+stems$plotid <- tolower(stems$plotid)
+stems %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species)))
+
+os_richness <- left_join(
+      stems %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species))),
+      untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species))), by = c("plotid", "year")
+) %>% mutate(os_rich = tagged_rich + untagged_rich)
+os_richness
+richness <- left_join(os_richness, richness, by = c("plotid","year"))
+richness
+summary(richness)
+```
+
+I also want to calculate some variables related to DBH, and in particular the "importance value" for each species. Importance value is calculated in the literature as:
+
+`IV = 0.5 * (relative density / relative dominance)` where 
+`Relative Density = # stems of species / total # of stems` and
+`Relative Dominance = basal area of species / total basal area`
+
+```{r calculate importance value}
+unique(untagged_dbh$species)
+untagged_dbh
+unique(stems$species)
+
+tagged_dbh <- left_join(stems %>% 
+      filter(status == "Alive", location == "In") %>%
+      select(plotid, year, species, dbh) %>% 
+      group_by(plotid, species, year) %>% 
+      summarise(live_count = length(species), live_avg_dbh = mean(dbh), live_tot_dbh = sum(dbh)),
+      stems %>% 
+      filter(status == "Dead", location == "In") %>%
+      select(plotid, year, species, dbh) %>% 
+      group_by(plotid, species, year) %>% 
+      summarise(dead_count = length(species), dead_avg_dbh = mean(dbh), dead_tot_dbh = sum(dbh)),
+      by = c("plotid", "species", "year")
+)
+tagged_dbh <- as.tbl(tagged_dbh)
+summary(tagged_dbh)
+
+tagged_dbh
+untagged_dbh
+
+stems_dbh <- rbind(tagged_dbh, untagged_dbh)
+
+stems_dbh <- stems_dbh %>% group_by(plotid, year) %>% mutate(all_live_count = sum(live_count), all_live_dbh = sum(live_tot_dbh, na.rm = T))
+summary(stems_dbh)
+filter(stems_dbh, all_live_dbh > 800)
+
+stems_dbh <- stems_dbh %>% mutate(rel_dens = live_count / all_live_count, rel_dom = live_tot_dbh / all_live_dbh, imp_value = 0.5*(rel_dens + rel_dom))
+summary(stems_dbh)
+```
+
+Make a pairs plot of the continuous variables for UMCA plots.
 ```{r pairs plot}
@@
 
-pairs(select(plots_env, -plotid, -sample_year), lower.panel = panel.cor, diag.panel = panel.hist)
+plots_umca <- ungroup(plots_umca) %>% select(-starts_with("psi"), -plot, -year)
 
-plots_env$avg_psi <- rowMeans(cbind(plots_env$psiSpr, plots_env$psiSum, plots_env$psiWin))
-pairs(select(plots_env, -plotid, -sample_year, -starts_with("psi")), lower.panel = panel.cor, diag.panel = panel.hist)
+pairs(select(plots_umca, -plot, -year, -starts_with("psi")), lower.panel = panel.cor, diag.panel = panel.hist)
 ```

### 2015-07-03 iale2015_sem.Rmd
Changed the calculation of understory richness by limiting the number of different species. Moved where I did the join of richness and UMCA data. Created an oak infection data frame joining UMCA infection, richness, and environmental variables. Created a pairs plot of this data frame.

@@
 veg <- rbind(veg_us_2005, veg_us_2011)
-veg
+
 unique(veg$year)
 veg$year[veg$year < 2011] <- 2005 # Creates the two most complete veg sampling
+unique(veg$us_species)
 veg <- veg %>% rename(us_species = UnderstorySp, plotid = PlotID)
-richness <- veg %>% group_by(plotid, year) %>% summarise(us_rich = length(unique(us_species)))
+
+# I want to match the species to those used by Sarah in her second chapter
+local_woody <- read.csv("analysis/data/presence_local_woody_n197.csv")
+summary(local_woody)
+unique(local_woody$spp.local)
+woody_spp <- as.character(local_woody$spp.local)
+woody_spp[woody_spp == "prunus"] <- "prunus sp."
+sort(woody_spp)
+sort(unique(veg$us_species))
+# It looks like I will want to add a few more to her list
+
+filter(veg, us_species == "ribes sp.")
+filter(veg, us_species == "ribes")
+veg$us_species[veg$us_species == "ribes"] <- "ribes sp."
+
+woody_spp <- c(woody_spp, "ceanothus sp.", "cebe", "cecu", "frla", "garrya sp.", "gemo", "juca", "mafu", "mane", "mimulus sp", "pisa", "quag x quke", "qube", "quercus sp.", "quwi", "rhamnus sp.", "rhcr", "ribes sp.", "rica", "risa", "roca", "rosa sp.", "rubus sp.", "rudi", "salix sp.", "syal", "vaccinium sp.")
+
+veg_sub <- filter(veg, veg$us_species %in% woody_spp)
+
+richness <- veg_sub %>%
+      group_by(plotid, year) %>% 
+      summarise(us_rich = length(unique(us_species)))
 richness
@@
 unique(stems$status)
+
 stems <- stems %>% select(-X) %>% filter(status == "Alive" | status == "Dead")
@@
 os_richness
+
 richness <- left_join(os_richness, richness, by = c("plotid","year"))
@@
 summary(richness)
+```
+
+Now I want to join the richness data set to the `plots_umca` data frame that currently has the environmental and leaf count/infection variables.
+
+```{r join richness data}
+plots_umca
+richness
+plots_umca <- rename(plots_umca, plotid = plot)
+plots_umca$plotid <- tolower(plots_umca$plotid)
+
+plots_umca <- left_join(plots_umca, richness, by = c("plotid", "year"))
+```
+
+Make a pairs plot of the continuous variables for UMCA plots.
+```{r pairs plot}
+# Helper functions for `pairs`
+panel.hist <- function(x, ...)
+{
+      usr <- par("usr"); on.exit(par(usr))
+      par(usr = c(usr[1:2], 0, 1.5) )
+      h <- hist(x, plot = FALSE)
+      breaks <- h$breaks; nB <- length(breaks)
+      y <- h$counts; y <- y/max(y)
+      rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
+}
+panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
+{
+      usr <- par("usr"); on.exit(par(usr))
+      par(usr = c(0, 1, 0, 1))
+      r <- abs(cor(x, y, use = "na.or.complete"))
+      txt <- format(c(r, 0.123456789), digits = digits)[1]
+      txt <- paste0(prefix, txt)
+      if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+      text(0.5, 0.5, txt, cex = cex.cor * r)
+}
+
+plots_umca <- ungroup(plots_umca)
+
+pairs(select(plots_umca, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist)
+```
+
+Now that I have summarized some vegetation (not species specific), environmental, and bay laurel disease data to the plot-level I want to join this to the plot-level oak infection data.
+
+```{r join umca and qusp data}
+plots_umca
+plot_qusp # right now this is broken down by oak species
+summary(plot_qusp)
+oak_sod <- plot_qusp %>% group_by(plot, year) %>%
+      summarise(uninf_oak_ct = sum(uninfected_oak_ct), inf_oak_ct = sum(infected_oak_ct), tot_oak_ct = sum(c(uninfected_oak_ct, infected_oak_ct)))
+oak_sod
+oak_sod <- rename(oak_sod, plotid = plot)
+oak_sod$plotid <- tolower(oak_sod$plotid)
+
+oak_sod <- left_join(oak_sod, plots_umca[1:7], 
+                     by = c("plotid", "year"))
+summary(oak_sod) # the NA's for the bay laurel count variables are legitimately zeroes b/c there are no bay in some of these plots
+oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
+oak_sod$infected_bay_ct[is.na(oak_sod$infected_bay_ct)] <- 0
+oak_sod$tot_bay[is.na(oak_sod$tot_bay)] <- 0
+
+head(plots_env)
+plots_env$plotid <- tolower(plots_env$plotid)
+oak_sod <- left_join(oak_sod, plots_env, by = c("plotid", "year" = "sample_year"))
+oak_sod
+
+richness
+oak_sod <- left_join(oak_sod, richness, by = c("plotid", "year"))
+summary(oak_sod)
+summary(filter(oak_sod, is.na(avg_tmax)))
+filter(oak_sod, is.na(avg_tmax), year == 2005) # BUSH01 plot; I'm dropping that from the data set
+oak_sod <- filter(oak_sod, plotid != "bush01")
+summary(filter(oak_sod, is.na(avg_tmax)))
+filter(oak_sod, is.na(avg_tmax), year == 2004) # PONTI01 and SWEET01 were other early abandoned plots
+oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01")
+summary(filter(oak_sod, is.na(avg_tmax))) # Now the NAs in the environmental variables are just for year 2003 when we had severely incomplete temperature data
+summary(oak_sod)
+
+filter(oak_sod, plotid == "ann05") # There is no 2007 record for ann05 in the UMCA form, but there is a plot visit. I will need to come back and investigate this further for correcting the database.
+oak_sod$infected_bay_ct[oak_sod$plotid == "ann05" & oak_sod$year == 2007] <- 15
+oak_sod$tot_bay[oak_sod$plotid == "ann05" & oak_sod$year == 2007] <- 15
+# I filled in the counts based on the values in the immediately preceding and following years, which I think is pretty solid reasoning.
+However, there is nothing I can do about the leaf counts right now.
+summary(oak_sod)
+
+filter(oak_sod, year == 2003) # 87 records
+filter(oak_sod, year == 2004) # 90 records
+# 2005 appears to be the first full year of observing all the plots
+filter(oak_sod, year == 2005) # 171 records
+filter(oak_sod, year == 2006) # 172 records
+filter(oak_sod, year == 2007) # 171 records
+filter(oak_sod, year == 2008) # 171 records
+filter(oak_sod, year == 2009) # 170 records
+filter(oak_sod, year == 2010) # 170 records
+filter(oak_sod, year == 2011) # 170 records; final year of Sarah's 2nd chapter
+filter(oak_sod, year == 2012) # 169 records
+filter(oak_sod, year == 2014) # 168 records
+# In Sarah's analysis we end up using only 163 plots
+```
+
+Now to do a pairs plot of the `oak_sod` data frame
+
+```{r oak_sod pairs plot}
+class(oak_sod)
+oak_sod <- ungroup(oak_sod)
+
+pairs(select(oak_sod, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist)
 ```
@@
 summary(stems_dbh)
+write.csv(stems_dbh, "analysis/data/species_dbh_impval_plot.csv")
 ```
-
-Make a pairs plot of the continuous variables for UMCA plots.
-```{r pairs plot}
-# Helper functions for `pairs`
-panel.hist <- function(x, ...)
-{
-      usr <- par("usr"); on.exit(par(usr))
-      par(usr = c(usr[1:2], 0, 1.5) )
-      h <- hist(x, plot = FALSE)
-      breaks <- h$breaks; nB <- length(breaks)
-      y <- h$counts; y <- y/max(y)
-      rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
-}
-panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
-{
-      usr <- par("usr"); on.exit(par(usr))
-      par(usr = c(0, 1, 0, 1))
-      r <- abs(cor(x, y, use = "na.or.complete"))
-      txt <- format(c(r, 0.123456789), digits = digits)[1]
-      txt <- paste0(prefix, txt)
-      if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
-      text(0.5, 0.5, txt, cex = cex.cor * r)
-}
-
-plots_umca <- ungroup(plots_umca) %>% select(-starts_with("psi"), -plot, -year)
-
-pairs(select(plots_umca, -plot, -year, -starts_with("psi")), lower.panel = panel.cor, diag.panel = panel.hist)
-```

### 2015-07-03 iale2015_sem.Rmd
added titles to the calls to the pairs plotting functions for the oak and bay laurel data frames

@@
 Make a pairs plot of the continuous variables for UMCA plots.
-```{r pairs plot}
+```{r umca sod pairs plot}
 # Helper functions for `pairs`
@@
 
-pairs(select(plots_umca, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist)
+pairs(select(plots_umca, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist, main = "UMCA SOD Data")
 ```
@@
 oak_sod <- ungroup(oak_sod)
-
-pairs(select(oak_sod, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist)
+pairs(select(oak_sod, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist, main = "Oak_SOD Data")
 ```

### 2015-07-03 iale2015_sem.Rmd
added canopy cover estimates for the plots from the spherical densiometer calculations

@@
 untagged_dbh <- read.csv("analysis/data/untagged_dbh_summary.csv")
+canopy_cover <- read.csv("analysis/data/plot_canopy_cover_dens.csv")
 
@@
 
-Now I want to join the richness data set to the `plots_umca` data frame that currently has the environmental and leaf count/infection variables.
+Now I want to join the richness data set and the canopy cover estimates to the `plots_umca` data frame that currently has the environmental and leaf count/infection variables.
 
@@
 plots_umca <- left_join(plots_umca, richness, by = c("plotid", "year"))
+head(canopy_cover)
+canopy_cover <- select(canopy_cover, plotid, year, pct_cvr)
+plots_umca <- left_join(plots_umca, canopy_cover, by = c("plotid", "year"))
 ```
@@
 
-Now that I have summarized some vegetation (not species specific), environmental, and bay laurel disease data to the plot-level I want to join this to the plot-level oak infection data.
+Now that I have summarized some vegetation (not species specific), environmental, and bay laurel disease data to the plot-level I want to join this and the canopy cover data to the plot-level oak infection data.
 
@@
 
-filter(oak_sod, year == 2003) # 87 records
-filter(oak_sod, year == 2004) # 90 records
+oak_sod <- left_join(oak_sod, canopy_cover, by = c("plotid", "year"))
+
+filter(oak_sod, year == 2003) # 86 records
+filter(oak_sod, year == 2004) # 88 records
 # 2005 appears to be the first full year of observing all the plots
-filter(oak_sod, year == 2005) # 171 records
+filter(oak_sod, year == 2005) # 170 records
 filter(oak_sod, year == 2006) # 172 records

### 2015-07-11 iale2015_sem.Rmd
Path analysis of 2005 & 2014 data for IALE 2015 World Congress presentation. I still need to do further data prep in order to achieve a repeated measures analysis. Considering limiting to data that I have measures for each sampling year

@@
 stems$plotid <- tolower(stems$plotid)
-stems %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species)))
+
+tagged_dbh <- stems %>% filter(status == "Alive") %>% group_by(plotid, year, species) %>% summarise(live_count = length(species))
+unique(tagged_dbh$species)
+os_richness <- rbind(tagged_dbh, select(untagged_dbh, plotid, species, year, live_count))
 
 os_richness <- left_join(
-      stems %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species))),
+      stems %>% filter(year == 2005 | year == 2014) %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species))),
       untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species))), by = c("plotid", "year")
-) %>% mutate(os_rich = tagged_rich + untagged_rich)
+) %>% mutate(os_rich = rowSums(cbind(tagged_rich,untagged_rich), na.rm = T))
 os_richness
@@
 
-```{r join richness data}
+```{r join richness and umca data}
 plots_umca
@@
 
+## Path Analysis Oak Infection 2005
+
+Here I am using the `oak_sod` data set which will be the oak infection (binomial) as the terminal response variable.
+
+```{r oak_sod path model 2005}
+library(ggm)
+names(oak_sod)
+oak_path1 <- DAG(
+      tmi ~ elevation,
+      os_rich ~ tmi,
+      hrs_14_20 ~ elevation + d2c + avg_psi + tmi,
+      tot_lfct ~ hrs_14_20 + os_rich + d2c,
+      inf_oak_ct ~ tot_lfct + os_rich + hrs_14_20 + d2c
+      )
+isAcyclic(oak_path1)
+bu_oak_path1 <- basiSet(oak_path1)
+bu_oak_path1
+
+oak_2005 <- filter(oak_sod, year == 2005)
+summary(oak_2005)
+oak_2005 <- select(oak_2005, plotid, uninf_oak_ct, inf_oak_ct, tot_lfct, hrs_14_20, elevation, tmi, avg_psi, os_rich, d2c)
+oak_2005$tot_lfct[is.na(oak_2005$tot_lfct)] <- 0
+filter(oak_2005, is.na(os_rich))
+oak_2005 <- na.omit(oak_2005)
+oak_2005$plotid <- as.factor(oak_2005$plotid)
+
+# This model makes Gaussian assumptions, which should be incorrect for these data
+shipley.test(oak_path1, cov(select(oak_2005, elevation, tmi, os_rich, hrs_14_20, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 170)
+# The p-value of 0.1114 indicates that there is no severe violation of the conditional independence assumptions with the application of a Gaussian model
+```
+
+~~Violation of conditional independence supports addressing the complexity of the data using different model distributions to deal with the data.~~
+
+Still I am going to fully asses the conditional independence claims
+
+```{r assess conditional independence claims 2005}
+bu_oak_path1 # 15 independence claims means 15 models
+library(nlme)
+library(lme4)
+
+f1 <- lme(avg_psi ~ d2c, data = oak_2005, random = ~1|plotid)
+summary(f1) # p = 0.5091
+f2 <- lme(avg_psi ~ elevation, data = oak_2005, random = ~1|plotid)
+summary(f2) # p = 0.2013
+f3 <- lme(avg_psi ~ tmi + elevation, data = oak_2005, random = ~1|plotid)
+summary(f3) # tmi; p = 0.1982
+f4 <- lme(os_rich ~ avg_psi + tmi, data = oak_2005, random = ~1|plotid)
+summary(f4) # avg_psi; p = 0.1066
+f5 <- lm(log1p(tot_lfct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
+# f5 <- lme(log1p(tot_lfct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, random = ~1|plotid)
+# f5 <- glmer(tot_lfct ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2005, family = "poisson")
+summary(f5) # avg_psi; p = 0.5993
+# f6 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "binomial")
+# f6 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "quasibinomial")
+f6 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
+summary(f6) # avg_psi; p = 0.08970
+f7 <- lm(elevation ~ d2c, data = oak_2005)
+# f7 <- lme(elevation ~ d2c, data = oak_2005, random = ~1|plotid)
+summary(f7) # p = 0.03163
+f8 <- lm(log(tmi) ~ d2c + elevation, data = oak_2005)
+# f8 <- lme(log(tmi) ~ d2c + elevation, data = oak_2005, random = ~1|plotid)
+summary(f8) # d2c; p = 0.4375
+f9 <- lm(os_rich ~ d2c + tmi, data = oak_2005)
+# f9 <- lme(os_rich ~ d2c + tmi, data = oak_2005, random = ~1|plotid)
+summary(f9) # d2c; p = 0.997
+f10 <- lm(os_rich ~ elevation + tmi, data = oak_2005)
+# f10 <- lme(os_rich ~ elevation + tmi, data = oak_2005, random = ~1|plotid)
+summary(f10) # elevation; p = 0.260
+
+# f11 <- lm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
+# f11 <- lm(log1p(tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
+f11 <- glm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, family = "poisson")
+chat <- deviance(f11)/df.residual(f11)
+QAIC(update(f11, family = x.quasipoisson), chat = chat)
+f11 <- glm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, family = "quasipoisson")
+# f11 <- lme(log1p(tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), random = ~1|plotid, data = oak_2005)
+# f11 <- glmer(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2005, family = "poisson")
+summary(f11) # elevation; p = 0.353425
+
+# f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "binomial")
+# chat <- deviance(f12)/df.residual(f12)
+# QAIC(update(f12, family = x.quasibinomial), chat = chat)
+# f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "quasibinomial")
+f12 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
+summary(f12) # elevation; p = 0.45826
+
+f13 <- lm(log1p(tot_lfct) ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
+# f13 <- lme(log1p(tot_lfct) ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, random = ~1|plotid)
+# f13 <- glmer(tot_lfct ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2005, family = "poisson")
+summary(f13) # tmi; p = 0.113050
+
+# f14 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "binomial")
+# chat <- deviance(f14)/df.residual(f14)
+# QAIC(update(f14, family = x.quasibinomial), chat = chat)
+# f14 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "quasibinomial")
+
+f14 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
+summary(f14) # tmi; p = 0.67609
+
+f15 <- lm(os_rich ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2005)
+# f15 <- lme(os_rich ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2005, random = ~1|plotid)
+summary(f15) # hrs_14_20; p = 0.355
+
+# Calculate C-statistic
+pvalues <- c(summary(f1)$tTable[2,5], summary(f2)$tTable[2,5], summary(f3)$tTable[2,5], summary(f4)$tTable[2,5], summary(f5)$coefficients[2,4], summary(f6)$coefficients[2,4], summary(f7)$tTable[2,5], summary(f8)$tTable[2,5], summary(f9)$tTable[2,5], summary(f10)$tTable[2,5], summary(f11)$coefficients[2,4], summary(f12)$coefficients[2,4], summary(f13)$coefficients[2,4], summary(f14)$coefficients[2,4], summary(f15)$tTable[2,5])
+
+cstat <- -2 * sum(log(pvalues))
+cstat
+```
+Given the C-statistic of 26.9 and that 2k degrees of freedom would be 30 I think that this meets the criteria for supporting the conditional independence claims.
+So, this supports my path model, now I need to run the models to produce estimates for the path coefficients.
+This is the model set that I need to get the estimated path coefficients:
+
+      1. os_rich ~ tmi,
+      2. hrs_14_20 ~ elevation + d2c + avg_psi + tmi,
+      3. tot_lfct ~ hrs_14_20 + os_rich + d2c,
+      4. infected oak ~ tot_lfct + os_rich + hrs_14_20 + d2c
+
+```{r estimate path coefficients 2005}
+# f0 <- fitted(mlme1, level = 0)# Values from population model
+# f1 <- fitted(mlme1, level = 1)# Within beach fitted values
+
+# fit1 <- lm(tmi ~ scale(elevation), data = oak_2005)
+fit1 <- lm(log(tmi) ~ scale(elevation), data = oak_2005) # best model
+# fit1 <- lme(log(tmi) ~ scale(elevation), data = oak_2005, random = ~1|plotid, method = "ML")
+summary(fit1)
+par(mfrow = c(2,2))
+plot(fit1)
+
+# fit2 <- lm(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2005)
+# fit2 <- lme(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2005, random = ~1|plotid) # best model
+fit2 <- glmer(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi) + (1|plotid), data = oak_2005, family = "poisson")
+# fit2 <- glmer.nb(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi) + (1|plotid), data = oak_2005)
+summary(fit2)
+plot(fit2)
+
+# fit3 <- lm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c), data = oak_2005)
+fit3 <- lm(log1p(tot_lfct) ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c), data = oak_2005)
+fit3 <- glm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c), data = oak_2005, family = "poisson")
+(chat <- deviance(fit3) / df.residual(fit3))
+x.quasipoisson<-function(...){
+  res<-quasipoisson(...)
+  res$aic<-poisson(...)$aic
+  res
+}
+library(MuMIn)
+QAIC(update(fit3, family = x.quasipoisson), chat = chat) # 178.3
+fit3 <- glm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c), data = oak_2005, family = "quasipoisson") # best model
+# fit3 <- glmer(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + (1|plotid), data = oak_2005, family = "poisson")
+summary(fit3)
+
+# fit4 <- lm((inf_oak_ct/(inf_oak_ct+uninf_oak_ct)) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c), data = oak_2005)
+# par(mfrow = c(2,2))
+# plot(fit4)
+
+# fit4 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c), data = oak_2005, family = "binomial")
+# (chat <- deviance(fit4) / df.residual(fit4))
+# x.quasibinomial<-function(...){
+#   res<-quasibinomial(...)
+#   res$aic<-binomial(...)$aic
+#   res
+# }
+# library(MuMIn)
+# QAIC(update(fit4, family = x.quasibinomial), chat = chat)
+# fit4 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c), data = oak_2005, family = "quasibinomial")
+# summary(fit4)
+
+fit4 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(d2c) + (1|plotid), data = oak_2005, family = "binomial") # best model
+summary(fit4)
+
+fit5 <- lm(os_rich ~ tmi, data = oak_2005); AIC(fit5)
+# fit5 <- lme(os_rich ~ tmi, data = oak_2005, random = ~1|plotid)
+summary(fit5)
+
+# Trying to plot the fitted model by each group...
+# f0 <- fitted(fit4, level = 0)
+# f1 <- fitted(fit4, level = 1)
+# 
+# f0[I]
+# 
+# I <- order(oak_2005$tot_lfct) ; lfct <- sort(oak_2005$tot_lfct)
+# plot(lfct, f0[I], lwd = 4, type = "l",
+#      ylab = "oak infection", xlab = "total leaf count")
+# for(i in 1:length(f1)){
+#       x1 <- oak_2005$hrs_14_20[oak_2005$plotid == i]
+#       y1 <- f1[oak_2005$hrs_14_20 == i]
+#       K <- order(x1)
+#       lines(sort(x1), y1[K], col = "red")
+# }
+# text(oak_2005$hrs_14_20, oak_2005$tot_lfct, oak_2005$plotid, cex = 0.9)
+```
+
+## Path Analysis Oak Infection - 2014
+Now I am going to fit the same path model to data from 2011.
+
+```{r oak_sod path model 2014}
+oak_path1 <- DAG(
+      tmi ~ elevation,
+      os_rich ~ tmi,
+      hrs_14_20 ~ elevation + d2c + avg_psi + tmi,
+      tot_lfct ~ hrs_14_20 + os_rich + d2c,
+      inf_oak_ct ~ tot_lfct + os_rich + hrs_14_20 + d2c
+      )
+
+oak_2014 <- filter(oak_sod, year == 2014)
+oak_2014 %>% select(plotid, tagged_rich, untagged_rich, os_rich) %>%
+      filter(is.na(os_rich))
+summary(oak_2014)
+oak_2014 <- select(oak_2014, plotid, uninf_oak_ct, inf_oak_ct, tot_lfct, hrs_14_20, hrs_blw10, hrs_abv25, elevation, tmi, avg_psi, os_rich, d2c)
+oak_2014$tot_lfct[is.na(oak_2014$tot_lfct)] <- 0
+oak_2014 <- na.omit(oak_2014)
+oak_2014$plotid <- as.factor(oak_2014$plotid)
+
+shipley.test(oak_path1, cov(select(oak_2014, elevation, tmi, os_rich, hrs_14_20, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 168)
+# Strongly significant p-value
+
+# shipley.test(DAG(
+#       tmi ~ elevation,
+#       os_rich ~ tmi,
+#       hrs_blw10 ~ elevation + d2c + avg_psi + tmi,
+#       tot_lfct ~ hrs_blw10 + os_rich + d2c,
+#       inf_oak_ct ~ tot_lfct + os_rich + hrs_blw10 + d2c), 
+#       cov(select(oak_2014, elevation, tmi, os_rich, hrs_blw10, d2c, avg_psi, tot_lfct, inf_oak_ct)), 
+#       n = 168)
+```
+
+```{r assess conditional independence 2014}
+f1 <- lme(avg_psi ~ d2c, data = oak_2014, random = ~1|plotid)
+summary(f1) # p = 0.591
+f2 <- lme(avg_psi ~ elevation, data = oak_2014, random = ~1|plotid)
+summary(f2) # p = 0.2171
+f3 <- lme(avg_psi ~ tmi + elevation, data = oak_2014, random = ~1|plotid)
+summary(f3) # tmi; p = 0.1889
+f4 <- lm(os_rich ~ avg_psi + tmi, data = oak_2014)
+# f4 <- lme(os_rich ~ avg_psi + tmi, data = oak_2014, random = ~1|plotid)
+summary(f4) # avg_psi; p = 0.156
+f5 <- lm(log1p(tot_lfct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014)
+# f5 <- lme(log1p(tot_lfct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, random = ~1|plotid)
+# f5 <- glmer(tot_lfct ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2014, family = "poisson")
+summary(f5) # avg_psi; p = 0.265
+
+f6 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "binomial")
+chat <- deviance(f6)/df.residual(f6)
+QAIC(update(f6, family = x.quasibinomial), chat = chat)
+f6 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "quasibinomial")
+# f6 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2014, family = "binomial")
+summary(f6) # avg_psi; p = 0.000722
+
+f7 <- lm(log(elevation) ~ d2c, data = oak_2014)
+# f7 <- lme(log(elevation) ~ d2c, data = oak_2014, random = ~1|plotid)
+summary(f7) # p = 0.0101
+
+f8 <- lm(log(tmi) ~ d2c + elevation, data = oak_2014)
+# f8 <- lme(log(tmi) ~ d2c + elevation, data = oak_2014, random = ~1|plotid)
+summary(f8) # d2c; p = 0.65795
+
+f9 <- lm(log1p(os_rich) ~ d2c + tmi, data = oak_2014)
+# f9 <- lme(log1p(os_rich) ~ d2c + tmi, data = oak_2014, random = ~1|plotid)
+summary(f9) # d2c; p = 0.73
+
+f10 <- lm(log1p(os_rich) ~ elevation + tmi, data = oak_2014)
+# f10 <- lme(log1p(os_rich) ~ elevation + tmi, data = oak_2014, random = ~1|plotid)
+summary(f10) # elevation; p = 0.126
+
+# f11 <- lm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014)
+f11 <- lm(log1p(tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014)
+f11 <- glm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, family = "poisson")
+chat <- deviance(f11)/df.residual(f11)
+QAIC(update(f11, family = x.quasipoisson), chat = chat)
+f11 <- glm((tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, family = "quasipoisson")
+# f11 <- lme(log1p(tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), random = ~1|plotid, data = oak_2014)
+# f11 <- glmer(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2014, family = "poisson")
+summary(f11) # elevation; p = 0.0170
+
+f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "binomial")
+chat <- deviance(f12)/df.residual(f12)
+QAIC(update(f12, family = x.quasibinomial), chat = chat)
+f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "quasibinomial")
+# f12 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2014, family = "binomial")
+summary(f12) # elevation; p = 0.00627
+
+f13 <- lm(log1p(tot_lfct) ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2014)
+f13 <- glm(tot_lfct ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, family = "poisson")
+chat <- deviance(f13)/df.residual(f13)
+QAIC(update(f13, family = x.quasipoisson), chat = chat)
+f13 <- glm(tot_lfct ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, family = "quasipoisson")
+
+# f13 <- lme(log1p(tot_lfct) ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, random = ~1|plotid)
+# f13 <- glmer(tot_lfct ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2014, family = "poisson")
+summary(f13) # tmi; p = 0.2562
+
+f14 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "binomial")
+chat <- deviance(f14)/df.residual(f14)
+QAIC(update(f14, family = x.quasibinomial), chat = chat)
+f14 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "quasibinomial")
+# f14 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2014, family = "binomial")
+summary(f14) # tmi; p = 0.7633
+
+f15 <- lm(log1p(os_rich) ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2014)
+# f15 <- lme(log1p(os_rich) ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2014, random = ~1|plotid)
+summary(f15) # hrs_14_20; p = 0.1039
+
+# Calculate C-statistic
+pvalues <- c(summary(f1)$tTable[2,5], summary(f2)$tTable[2,5], summary(f3)$tTable[2,5], summary(f4)$coefficients[2,4], summary(f5)$coefficients[2,4], summary(f6)$coefficients[2,4], summary(f7)$coefficients[2,4], summary(f8)$coefficients[2,4], summary(f9)$coefficients[2,4], summary(f10)$coefficients[2,4], summary(f11)$coefficients[2,4], summary(f12)$coefficients[2,4], summary(f13)$coefficients[2,4], summary(f14)$coefficients[2,4], summary(f15)$coefficients[2,4])
+
+cstat <- -2 * sum(log(pvalues))
+cstat # 69.18
+chisq.test(c(cstat,30))
+qchisq(pvalues, df = 30)
+chisq.test(cbind(pvalues, qchisq(pvalues, df = 30)))
+```
+
+This C-statistic is much larger than the degrees of freedom, so I'm not confident in the conditional independencies of the structural model being met for the 2014 data.
+In fact, based on the C-test value from the earlier call to `shipleys.test` on these data this value is significantly different.
+
+The models testing for conditional independence indicate direct effects of PSI and elevation on oak infection, as well as a direct effect of elevation on symptomatic leaf count.
+This is indicated by the p-values showing strong statistical significance in the relationship between these variables.
+Below I have generated a new path model that accounts for the unassumed direct effects, which means they become assumed direct effects.
+
+```{r oak path 2}
+oak_path2 <- DAG(
+      tmi ~ elevation,
+      os_rich ~ tmi,
+      hrs_14_20 ~ elevation + d2c + avg_psi + tmi,
+      tot_lfct ~ hrs_14_20 + os_rich + d2c + elevation,
+      inf_oak_ct ~ tot_lfct + avg_psi + os_rich + hrs_14_20 + d2c + elevation
+      )
+isAcyclic(oak_path2)
+bu_oak_path2 <- basiSet(oak_path2)
+bu_oak_path2
+
+shipley.test(oak_path2, cov(select(oak_2014, elevation, tmi, os_rich, hrs_14_20, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 168)
+
+plotGraph(oak_path2)
+```
+
+This means commenting out models `f6` (oak ~ PSI), `f12` (oak ~ elevation), and `f11` (leaf count ~ elevation) in the `pvalues` summary.
+
+```{r recalculate c-statistic}
+pvalues <- c(
+      summary(f1)$tTable[2,5], summary(f2)$tTable[2,5], summary(f3)$tTable[2,5],
+      summary(f4)$coefficients[2,4], summary(f5)$coefficients[2,4], 
+             # summary(f6)$coefficients[2,4], 
+      summary(f7)$coefficients[2,4], summary(f8)$coefficients[2,4],
+      summary(f9)$coefficients[2,4], summary(f10)$coefficients[2,4],
+      # summary(f11)$coefficients[2,4], 
+      # summary(f12)$coefficients[2,4], 
+      summary(f13)$coefficients[2,4], summary(f14)$coefficients[2,4], summary(f15)$coefficients[2,4])
+
+cstat <- -2 * sum(log(pvalues))
+cstat # 36.56767
+chisq.test(c(cstat,30))
+```
+I think this would be evidence for wanting/needing to use the repeated measures approach across all the data to capture interannual variation in the relationships.
+
+I am continuing with fitting the models to estimate the path coefficients for 2014 data based on this redrawn path model.
+
+```{r estimate path coefficients 2014}
+# f0 <- fitted(mlme1, level = 0)# Values from population model
+# f1 <- fitted(mlme1, level = 1)# Within beach fitted values
+
+# fit1_2014 <- lm(tmi ~ scale(elevation), data = oak_2014)
+fit1_2014 <- lm(log(tmi) ~ scale(elevation), data = oak_2014) # best model
+# fit1_2014 <- lme(log(tmi) ~ scale(elevation), data = oak_2014, random = ~1|plotid)
+summary(fit1_2014)
+par(mfrow = c(2,2))
+plot(fit1_2014)
+
+# fit2_2014 <- lm(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2014)
+# fit2_2014 <- lme(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2014, random = ~1|plotid) # best model
+fit2_2014 <- glm(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2014, family = "poisson")
+chat <- deviance(fit2_2014)/df.residual(fit2_2014)
+QAIC(update(fit2_2014, family = x.quasipoisson), chat = chat)
+fit2_2014 <- glm(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2014, family = "quasipoisson")
+# fit2_2014 <- glmer(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi) + (1|plotid), data = oak_2014, family = "poisson")
+summary(fit2_2014)
+plot(fit2_2014)
+
+# fit3_2014 <- lm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation), data = oak_2014)
+# fit3_2014 <- lm(log1p(tot_lfct) ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation), data = oak_2014)
+fit3_2014 <- glm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation), data = oak_2014, family = "poisson")
+(chat <- deviance(fit3_2014) / df.residual(fit3_2014))
+x.quasipoisson<-function(...){
+  res<-quasipoisson(...)
+  res$aic<-poisson(...)$aic
+  res
+}
+library(MuMIn)
+QAIC(update(fit3_2014, family = x.quasipoisson), chat = chat) # 178.3
+fit3_2014 <- glm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation), data = oak_2014, family = "quasipoisson") # best model
+# fit3_2014 <- glmer(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation) + (1|plotid), data = oak_2014, family = "poisson")
+summary(fit3_2014)
+
+# fit4_2014 <- lm((inf_oak_ct/(inf_oak_ct+uninf_oak_ct)) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c) + scale(elevation) + scale(avg_psi), data = oak_2014)
+par(mfrow = c(2,2))
+plot(fit4_2014)
+
+fit4_2014 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c) + scale(elevation) + scale(avg_psi), data = oak_2014, family = "binomial")
+(chat <- deviance(fit4_2014) / df.residual(fit4_2014))
+x.quasibinomial<-function(...){
+  res<-quasibinomial(...)
+  res$aic<-binomial(...)$aic
+  res
+}
+library(MuMIn)
+QAIC(update(fit4_2014, family = x.quasibinomial), chat = chat)
+fit4_2014 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c) + scale(elevation) + scale(avg_psi), data = oak_2014, family = "quasibinomial")
+
+# fit4_2014 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c) + (1|plotid), data = oak_2014, family = "binomial") # best model
+summary(fit4_2014)
+
+# fit5_2014 <- lm(os_rich ~ tmi, data = oak_2014); AIC(fit5_2014)
+fit5_2014 <- lm(log(os_rich) ~ tmi, data = oak_2014); AIC(fit5_2014)
+# fit5_2014 <- lme(os_rich ~ tmi, data = oak_2014, random = ~1|plotid)
+# fit5_2014 <- lme(log(os_rich) ~ tmi, data = oak_2014, random = ~1|plotid)
+summary(fit5_2014)
+```
+
 I also want to calculate some variables related to DBH, and in particular the "importance value" for each species. Importance value is calculated in the literature as:

### 2015-08-06 sem_development.Rmd
Change file name via Save As...
I saved the 'IALE2015_sem.Rmd' file under a new name, while still retaining the original in order to maintain some clarity about what precisely I did for that presentation.

--- 2015-07-11 iale2015_sem.Rmd
@@
 
-```{r load data}
+#### Load data
+```{r load ecological data, echo=FALSE}
 # Ecological data: leaf counts, infection
@@
 canopy_cover <- read.csv("analysis/data/plot_canopy_cover_dens.csv")
+```
+```{r load physical data, echo=FALSE}
+# Physical environment: topography
+library(rgdal); library(plyr); library(dplyr)
+plots_shp <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
+plots_env <- plots_shp@data
+summary(plots_env)
+plots_env <- select(plots_env, -cat)
 
-# Physical environment: topography
-library(rgdal)
-plots_shp <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
-summary(plots_shp)
-plot(plots_shp, axes=T)
-plots_env <- plots_shp@data
+# Elevation and vegetation height are in feet, slope is degrees, and canopy density is the proportion of the cell covered by canopy (0-1)
+plots_topo.shp <- readOGR("analysis/data/plot_gis/", "plots202_vegtopo")
+plots_vegtopo <- plots_topo.shp@data
+summary(plots_vegtopo)
+plots_vegtopo <- select(plots_vegtopo, -X, -Y)
 
-# Temperature: rainy season hour summaries
-library(plyr); library(dplyr)
+plots_env <- inner_join(plots_env, plots_vegtopo, by = "PLOT_ID")
+summary(plots_env)
+plots_env$elev_m <- plots_env$elev15m * .3048
+plots_env <- select(plots_env, -elevation, -tmi) #remove elevation, tmi from old derivation (I think a legacy 10m USGS DEM)
+plots_env$veght_m <- plots_env$veght15m * .3048
+```
+```{r load weather data}
+# Temperature: rainy season hour summaries (long format for each year)
 plots_temps <- read.csv("analysis/data/rainy_season_temps.csv")
@@
 plots_temps <- select(plots_temps, -X)
+
+# Rainfall: interpolated rainfall for each season
+plots_rain <- read.csv("analysis/data/plot202rain.csv")
+summary(plots_rain)
+plots_rain <- select(plots_rain, -cat)
+# pairs(select(plots_rain, -PLOT_ID), 
+#       lower.panel = panel.cor, 
+#       diag.panel = panel.hist)
 ```
 
-Join the plot level environmental data (geographic/topographic & temperatures)
+#### Join the plot level environmental data 
+(geographic, topographic, rainfall, temperatures)
 ```{r join plot data, message=FALSE}
+plots_env <- left_join(plots_env, plots_rain, by = "PLOT_ID")
 plots_env <- left_join(plots_env, plots_temps, by = c("PLOT_ID"="plotid"))
@@
 str(plots_env)
-plots_env <- plots_env %>% select(-cat) %>% rename(easting = X, northing = Y, plotid = PLOT_ID)
+plots_env <- plots_env %>% rename(plotid = PLOT_ID, easting = X, northing = Y)
 ```
 
-Investigating `NAs` in the `plots_env` data set
+#### Investigating `NAs` in the `plots_env` data set
 ```{r investigate NAs}
@@
 
-Examine the bay laurel data at individual and plot levels
+#### Examine the bay laurel data at individual and plot levels
 ```{r examine umca data}

### 2015-08-16 analysis/explore_data.R
Update notes on DBH data quality check

@@ -159,7 +159,7 @@ dbh2014 <- stems %>% select(plot, tag, dbh, year) %>%
       filter(dbh>0, year == 2014)
 anti_join(dbh2014, dbh_a, by = "tag") # this inidcates that there are 113 stems that were measured for the first time in 2014
 
-#### Explore DBH data: change in DBH ####
+#### Explore DBH data: negative change in DBH ####
 library(ggplot2)
 # Scatterplot of first dbh measurement (y-axis) against year
 qplot(year, dbh, data = dbh_a, geom="jitter")
@@ -410,8 +410,11 @@ qplot(delta_dbh, dbh1, data = stems %>% select(plot, cluster, tag, species, delt
 # A visual check shows that there are a handful of stems with growth > 20cm
 
 #### Suggestions from Margaret about exploring DBH changes ####
-# I emailed Richard and Margaret about thresholds for where I should really look more closely at the measurements. Richard noted that there isn't much out there on growth rates for these species. Margaret suggested a few things, including other ways of looking at the data:
+# I emailed Richard and Margaret about thresholds for where I should really look more closely at the measurements. Richard noted that there isn't much out there on growth rates for these species. I did find something done for tree species of the Northeastern U.S. <http://www.fs.fed.us/ne/newtown_square/publications/research_papers/pdfs/scanned/OCR/ne_rp649.pdf>
+
+# Margaret suggested a few things, including other ways of looking at the data:
 # 1. "Have you tried looking at relative growth instead of the absolute increment growth?  Either DBH2/DBH1, or the instantaneous rate ln(DBH2/DBH1) / (time2-time1).  20 cm growth means a lot more to a 5 cm tree than a 50 cm tree… either way, it's still big.  I'd also consider this species by species, because I bet some grow much faster than others."
+
 stems$dbh2_1_ratio <- stems$dbh2/stems$dbh1# absolute increment growth
 stems$inst_grwth_rate <- log(stems$dbh2/stems$dbh1) / (stems$year_dbh2 - stems$year_dbh1)
 #stems$time_diff <- (stems$year_dbh2 - stems$year_dbh1)

### 2015-08-16 sem_development.Rmd
Rearrange code and add new titles/headings.
Explore relationship of contributing area and slope in TWI

--- 2015-07-11 iale2015_sem.Rmd
@@
 ---
-title: "IALE 2015 Analysis"
+title: "Path Analysis - Chapter 2"
 author: "Whalen Dillon"
@@
 
-```{r load data}
+#### Load data
+```{r pairs helper functions, echo=FALSE}
+panel.hist <- function(x, ...)
+{
+      usr <- par("usr"); on.exit(par(usr))
+      par(usr = c(usr[1:2], 0, 1.5) )
+      h <- hist(x, plot = FALSE)
+      breaks <- h$breaks; nB <- length(breaks)
+      y <- h$counts; y <- y/max(y)
+      rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
+}
+panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
+{
+      usr <- par("usr"); on.exit(par(usr))
+      par(usr = c(0, 1, 0, 1))
+      r <- abs(cor(x, y, use = "na.or.complete"))
+      txt <- format(c(r, 0.123456789), digits = digits)[1]
+      txt <- paste0(prefix, txt)
+      if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
+      text(0.5, 0.5, txt, cex = cex.cor * r)
+}
+```
+```{r load ecological data, echo=FALSE}
 # Ecological data: leaf counts, infection
@@
 canopy_cover <- read.csv("analysis/data/plot_canopy_cover_dens.csv")
+```
+```{r load physical data, echo=FALSE}
+# Physical environment: topography
+library(rgdal); library(plyr); library(dplyr)
+plots_shp <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
+plots_env <- plots_shp@data
+summary(plots_env)
+plots_env <- select(plots_env, -cat)
 
-# Physical environment: topography
-library(rgdal)
-plots_shp <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
-summary(plots_shp)
-plot(plots_shp, axes=T)
-plots_env <- plots_shp@data
-
-# Temperature: rainy season hour summaries
-library(plyr); library(dplyr)
+# Elevation and vegetation height are in feet, slope is degrees, and canopy density is the proportion of the cell covered by canopy (0-1)
+plots_topo.shp <- readOGR("analysis/data/plot_gis/", "plots202_vegtopo")
+plots_vegtopo <- plots_topo.shp@data
+summary(plots_vegtopo)
+plots_vegtopo <- select(plots_vegtopo, -X, -Y)
+```
+```{r load weather data, echo=FALSE}
+# Temperature: rainy season hour summaries (long format for each year)
 plots_temps <- read.csv("analysis/data/rainy_season_temps.csv")
@@
 plots_temps <- select(plots_temps, -X)
+
+# Rainfall: interpolated rainfall for each season
+plots_rain <- read.csv("analysis/data/plot202rain.csv")
+summary(plots_rain)
+plots_rain <- select(plots_rain, -cat)
+# pairs(select(plots_rain, -PLOT_ID), 
+#       lower.panel = panel.cor, 
+#       diag.panel = panel.hist)
+plots_rain <- rename(plots_rain, plotid = PLOT_ID)
 ```
 
-Join the plot level environmental data (geographic/topographic & temperatures)
-```{r join plot data, message=FALSE}
-plots_env <- left_join(plots_env, plots_temps, by = c("PLOT_ID"="plotid"))
+#### Join remotely sensed and derived physical data
+LiDAR products, variables derived from DEMs
+```{r join remotely sensed data}
+plots_env <- inner_join(plots_env, plots_vegtopo, by = "PLOT_ID")
+summary(plots_env)
+plots_env$elev_m <- plots_env$elev15m * .3048
+plots_env <- select(plots_env, -elevation, -tmi) #remove elevation, tmi from old derivation (I think a legacy 10m USGS DEM)
+plots_env$veght_m <- plots_env$veght15m * .3048
+rm(plots_vegtopo)
+```
+
+#### Convert rain data from wide to long
+```{r rain wide to long}
+library(tidyr)
+rain2d <- plots_rain %>% select(plotid, ends_with("2d"))
+names(rain2d) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain2d <- rain2d %>% gather(year, rain_tot_2d, -plotid)
+
+rain3d <- plots_rain %>% select(plotid, ends_with("3d"))
+names(rain3d) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain3d <- rain3d %>% gather(year, rain_tot_3d, -plotid)
+
+rain_v <- plots_rain %>% select(plotid, ends_with("v"))
+names(rain_v) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain_v <- rain_v %>% gather(year, rain_tot_v, -plotid)
+
+rain_data <-  left_join(rain2d, 
+                        left_join(rain3d, rain_v, by = c("plotid","year")),
+                        by = c("plotid","year"))
+summary(rain_data)
+rm(rain_v); rm(rain2d); rm(rain3d)
+write.csv(rain_data, "analysis/data/rain_data_long.csv", row.names = F)
+```
+
+#### Explore data for completeness and matching
+```{r explore plot_env data, message=FALSE}
 summary(plots_env)
 str(plots_env)
-plots_env <- plots_env %>% select(-cat) %>% rename(easting = X, northing = Y, plotid = PLOT_ID)
+plots_env <- plots_env %>% rename(plotid = PLOT_ID, easting = X, northing = Y)
+plots_env <- select(plots_env, -cat)
+write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
+plots_env <- read.csv("analysis/data/env_gis_variables.csv")
 ```
 
-Investigating `NAs` in the `plots_env` data set
-```{r investigate NAs}
-summary(plots_env)
-filter(plots_env, is.na(sample_year))
+Examine temperature data, which is in long format and covers sample years 2004 through 2014. This means that I will need to exclude disease-related sampling from 2003.
+```{r examine temperature & rain data}
+summary(plots_temps)
+summary(rain_data)
+unique(plots_temps$plotid)# 201 levels
+unique(rain_data$plotid)# 202 levels
+tempsid <- as.character(unique(plots_temps$plotid))
+rainid <- as.character(unique(rain_data$plotid))
+rainid %in% tempsid
+rainid[62]# BUSH01 still in the rain data
+rain_data <- filter(rain_data, plotid != "BUSH01")
+# Temperature and rainfall data now have same number of observations
+rm(tempsid); rm(rainid)
+```
+
+#### Investigating `NAs` in the `plots_temps` data set
+A number of observations for above 25 threshold have NAs.
+I believe that these are in fact truly zero observations.
+NAs occurred for the 2010, 2011, and 2014 seasons, and the only NAs in the data set are associated with the above 25 threshold.
+I do believe that it is possible that the temperature was never above 25 C during the wet season November - May in some years.
+```{r investigate NAs, eval=FALSE, echo=FALSE}
+# summary(plots_env)
+# filter(plots_env, plotid == "BUSH01")
 
 # BUSH01 plot included. I think we have limited ecological data for this plot, and no microclimate data. So, I am going to exclude it from all the data sets
-plots_env <- filter(plots_env, plotid != "BUSH01")
+# plots_env <- filter(plots_env, plotid != "BUSH01")
 stems <- rename(stems, plotid = plot) %>% filter(plotid != "BUSH01")
 
-summary(plots_env) # A number of observations for above 25 threshold have NAs
-# I believe that these are in fact truly zero observations. NAs occurred for the 2010, 2011, and 2014 seasons, and the only NAs in the data set are associated with the above 25 threshold. 
-plots_env %>% filter(is.na(hrs_abv25))
+plots_temps %>% filter(is.na(hrs_abv25))
 # I am replacing the NAs with `0`, assuming true observations.
-plots_env[is.na(plots_env)] <- 0
+plots_temps[is.na(plots_temps)] <- 0
 ```
 
-Examine the bay laurel data at individual and plot levels
+#### Save data with BUSH01 excluded
+```{r save data}
+write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
+write.csv(rain_data, "analysis/data/rain_data_long.csv", row.names = F)
+write.csv(plots_temps, "analysis/data/temp_data_long.csv", row.names = F)
+write.csv(stems, "analysis/data/stems_noBUSH01.csv", row.names = F)
+rm(plots_temps); rm(plots_rain)
+```
+
+#### Reload data that has BUSH01 excluded
+```{r reload data}
+plots_env <- read.csv("analysis/data/env_gis_variables.csv")
+rain_data <- read.csv("analysis/data/rain_data_long.csv")
+temps_data <- read.csv("analysis/data/temp_data_long.csv")
+stems <- read.csv("analysis/data/stems_noBUSH01.csv")
+```
+
+#### Plot contributing area vs. slope
+This is a suggestion from Ryan to plot the contributing area vs.
+the slope for each of the plots in order to tease out how either influences infection and hosts.
+Decomposing TWI to determine which factor has a greater influence on disease dynamics.
+```{r decomposing TWI influence}
+library(ggplot2)
+summary(plots_vegtopo)
+par(mfrow = c(3,1))
+qplot(slope15m, flowacc15m, data = plots_env) +
+      geom_text(aes(label = plotid))
+qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m < 20), 
+      main = "Flow Accumulation < 5") +
+      geom_text(aes(label = plotid))
+qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 20 & flowacc15m < 1000), main = "Flow Accumulation 5 - 1000") +
+      geom_text(aes(label = plotid))
+qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 1000), 
+      main = "Flow Accumulation > 1000") +
+      geom_text(aes(label = plotid))
+
+qplot(slope15m, facc_ws15m, data = plots_env)
+qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m < 15), 
+      main = "facc < 15") +
+      geom_text(aes(label = plotid))
+qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 15 & facc_ws15m < 50), 
+      main = "facc 15 - 50") +
+      geom_text(aes(label = plotid))
+qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 50 & facc_ws15m < 200), 
+      main = "facc 50 - 200") +
+      geom_text(aes(label = plotid))
+qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 200 & facc_ws15m < 1000), 
+      main = "facc 200 - 1000") +
+      geom_text(aes(label = plotid))
+qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 1000), 
+      main = "facc >1000") +
+      geom_text(aes(label = plotid))
+
+# pairs(select(plots_env, -plotid, -starts_with("rain")), 
+#       lower.panel = panel.cor, diag.panel = panel.hist)
+# pairs(select(plots_env, -plotid, -starts_with("rain")) %>% 
+#             filter(flowacc15m < 5), 
+#       lower.panel = panel.cor, diag.panel = panel.hist,
+#       main = "Flow accumualtion < 5")
+# pairs(select(plots_env, -plotid, -starts_with("rain")) %>% 
+#             filter(flowacc15m > 100 & flowacc15m < 1000), 
+#       lower.panel = panel.cor, diag.panel = panel.hist,
+#       main = "Flow accumulation 100-1000")
+# pairs(select(plots_env, -plotid, -starts_with("rain")) %>% 
+#             filter(flowacc15m > 1000), 
+#       lower.panel = panel.cor, diag.panel = panel.hist,
+#       main = "Flow accumulation > 1000")
+
+plot(flowacc15m ~ slope15m, data = filter(plots_env, flowacc15m < 30))
+pairs(select(plots_env, -plotid) %>% filter(flowacc15m < 30), 
+      lower.panel = panel.cor, diag.panel = panel.hist)
+
+library(manipulate)
+manipulate(
+      qplot(slope15m, flowacc15m, data = plots_env,
+           axes = axes,
+           cex = cex,
+           pch = if(pch) 19 else 1
+           ),
+      axes = checkbox(TRUE, "Show axes"),
+      cex = slider(0, 5, initial = 1, step = 0.1, label = "Point size"),
+      pch = button("Fill points")
+)
+
+dataplot <- function(dat){
+      name <- sys.call()[[2]]
+      vars <- as.list(names(dat))
+      e <- new.env()
+      e$data <- name
+      manipulate(
+            {
+                  form = as.formula(paste(y,x, sep = "~"))
+                  plot(form, data = dat, main = as.character(name), las = 1)
+                  e$form <- form
+            },
+            x = do.call(picker, c(vars, initial = vars[1])),
+            y = do.call(picker, c(vars, initial = vars[2]))
+      )
+      invisible(e)
+}
+f <- dataplot(plots_env)
+```
+
+#### Examine the bay laurel data at individual and plot levels
 ```{r examine umca data}
@@
 # This only has number of infected, uninfected, and total, which must apply to living stems only. I want to analyze leaf counts.
-summary(bay_laurel)
-# This has stem-level data on bay laurel.
+summary(bay_laurel) # This has stem-level data on bay laurel.
 ```
@@
 So now I have a plot-level of UMCA infection levels for each year. I want to join this with some of the environmental data.
-
 ```{r join env and umca data}
@@
 # plot == plotid, year == sample_year
-plots_umca <- left_join(select(plots_umca, -species), plots_env, by = c("plot" = "plotid", "year" = "sample_year"))
+plots_umca <- left_join(select(plots_umca, -species), plots_env, by = c("plot" = "plotid"))
 summary(plots_umca)
 filter(plots_umca, is.na(easting))
-# NAs are due to no microclimate data for year 2003, and abandoned plots BUSH01 & SUGAR02
+# NAs due to no topo data for abandoned SUGAR02
+plots_umca <- na.omit(plots_umca)
+
+plots_temps <- as.tbl(plots_temps)
+plots_temps <- plots_temps %>% rename(year = sample_year)
+plots_temps$plotid <- tolower(plots_temps$plotid)
+
+plots_umca <- left_join(plots_umca, plots_temps, by = c("plotid", "year"))
 ```
@@
 Now I also want to calculate some plot-level community variables from the vegetation data. First, understory richness from the vegetation transects.
+
 ```{r calculate understory richness}
@@
 veg$year[veg$year < 2011] <- 2005 # Creates the two most complete veg sampling
+veg <- veg %>% rename(us_species = UnderstorySp, plotid = PlotID)
 unique(veg$us_species)
-veg <- veg %>% rename(us_species = UnderstorySp, plotid = PlotID)
 
@@
 woody_spp[woody_spp == "prunus"] <- "prunus sp."
-sort(woody_spp)
+sort(woody_spp) # only 29 species
 sort(unique(veg$us_species))
@@
 
-tagged_dbh <- stems %>% filter(status == "Alive") %>% group_by(plotid, year, species) %>% summarise(live_count = length(species))
-unique(tagged_dbh$species)
-os_richness <- rbind(tagged_dbh, select(untagged_dbh, plotid, species, year, live_count))
+live_tagged_dbh <- stems %>% filter(status == "Alive") %>% group_by(plotid, year, species) %>% summarise(live_count = length(species))
+unique(live_tagged_dbh$species)
+os_richness <- rbind(live_tagged_dbh, select(untagged_dbh, plotid, species, year, live_count))
 
 os_richness <- left_join(
-      stems %>% filter(year == 2005 | year == 2014) %>% group_by(plotid, year) %>% summarise(tagged_rich = length(unique(species))),
-      untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species))), by = c("plotid", "year")
-) %>% mutate(os_rich = rowSums(cbind(tagged_rich,untagged_rich), na.rm = T))
+      stems %>% filter(year == 2005 | year == 2014) %>% 
+            group_by(plotid, year) %>% 
+            summarise(tagged_rich = length(unique(species))),
+      untagged_dbh %>% 
+            group_by(plotid, year) %>% 
+            summarise(untagged_rich = length(unique(species))), by = c("plotid", "year")) %>% 
+      mutate(os_rich = rowSums(cbind(tagged_rich,untagged_rich), na.rm = T))
 os_richness
@@
 
-pairs(select(plots_umca, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist, main = "UMCA SOD Data")
+# pairs(select(plots_umca, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist, main = "UMCA SOD Data")
 ```
@@
 oak_sod <- plot_qusp %>% group_by(plot, year) %>%
-      summarise(uninf_oak_ct = sum(uninfected_oak_ct), inf_oak_ct = sum(infected_oak_ct), tot_oak_ct = sum(c(uninfected_oak_ct, infected_oak_ct)))
+      summarise(uninf_oak_ct = sum(uninfected_oak_ct), 
+                inf_oak_ct = sum(infected_oak_ct), 
+                tot_oak_ct = sum(c(uninfected_oak_ct, infected_oak_ct)))
 oak_sod
@@
 
-oak_sod <- left_join(oak_sod, plots_umca[1:7], 
+oak_sod <- left_join(oak_sod, select(plots_umca, plotid, year, uninfected_bay_ct, infected_bay_ct, tot_bay, avg_lfct, tot_lfct), 
                      by = c("plotid", "year"))
-summary(oak_sod) # the NA's for the bay laurel count variables are legitimately zeroes b/c there are no bay in some of these plots
+summary(oak_sod) # the NA's for the bay laurel variables are legitimately zeroes b/c there are no bay in some of these plots
 oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
@@
 
-head(plots_env)
-plots_env$plotid <- tolower(plots_env$plotid)
-oak_sod <- left_join(oak_sod, plots_env, by = c("plotid", "year" = "sample_year"))
+oak_sod <- left_join(oak_sod, plots_env, by = "plotid")
+oak_sod <- left_join(oak_sod, plots_temps, 
+                     by = c("plotid", "year" = "year"))
 oak_sod
@@
 oak_sod <- filter(oak_sod, plotid != "bush01")
-summary(filter(oak_sod, is.na(avg_tmax)))
 filter(oak_sod, is.na(avg_tmax), year == 2004) # PONTI01 and SWEET01 were other early abandoned plots
 oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01")
-summary(filter(oak_sod, is.na(avg_tmax))) # Now the NAs in the environmental variables are just for year 2003 when we had severely incomplete temperature data
+summary(filter(oak_sod, is.na(avg_tmax))) # Now the NAs in the temperature variables are just for year 2003 when we had severely incomplete temperature data
 summary(oak_sod)
+filter(oak_sod, is.na(easting))
+oak_sod <- filter(oak_sod, plotid != "duer01")
 
@@
 oak_sod$tot_bay[oak_sod$plotid == "ann05" & oak_sod$year == 2007] <- 15
-# I filled in the counts based on the values in the immediately preceding and following years, which I think is pretty solid reasoning.
-However, there is nothing I can do about the leaf counts right now.
+# I filled in the counts based on the values in the immediately preceding and following years, which I think is pretty solid reasoning. However, there is nothing I can do about the leaf counts.
 summary(oak_sod)
@@
 
-filter(oak_sod, year == 2003) # 86 records
+filter(oak_sod, year == 2003) # 85 records
 filter(oak_sod, year == 2004) # 88 records
@@
 filter(oak_sod, year == 2010) # 170 records
-filter(oak_sod, year == 2011) # 170 records; final year of Sarah's 2nd chapter
+filter(oak_sod, year == 2011) # 170 records; final year for Sarah's 2nd chapter
 filter(oak_sod, year == 2012) # 169 records
@@
 oak_path1 <- DAG(
-      tmi ~ elevation,
-      os_rich ~ tmi,
-      hrs_14_20 ~ elevation + d2c + avg_psi + tmi,
+      twi15m ~ elev15m,
+      os_rich ~ twi15m,
+      hrs_14_20 ~ elev15m + d2c + avg_psi + twi15m,
       tot_lfct ~ hrs_14_20 + os_rich + d2c,
@@
 
+oak_path2 <- DAG(
+      twi15m ~ elev15m,
+      os_rich ~ twi15m,
+      hrs_blw10 ~ elev15m + d2c + avg_psi + twi15m,
+      tot_lfct ~ hrs_blw10 + os_rich + d2c,
+      inf_oak_ct ~ tot_lfct + os_rich + hrs_blw10 + d2c
+      )
+isAcyclic(oak_path2)
+
 oak_2005 <- filter(oak_sod, year == 2005)
 summary(oak_2005)
-oak_2005 <- select(oak_2005, plotid, uninf_oak_ct, inf_oak_ct, tot_lfct, hrs_14_20, elevation, tmi, avg_psi, os_rich, d2c)
+oak_2005 <- select(oak_2005, plotid, uninf_oak_ct, inf_oak_ct, tot_lfct, hrs_14_20, hrs_blw10, hrs_abv25, avg_tmax, avg_tmin, rain05_2d, rain05_v, elev15m, twi15m, avg_psi, os_rich, d2c)
 oak_2005$tot_lfct[is.na(oak_2005$tot_lfct)] <- 0
 filter(oak_2005, is.na(os_rich))
-oak_2005 <- na.omit(oak_2005)
+# oak_2005 <- na.omit(oak_2005)
 oak_2005$plotid <- as.factor(oak_2005$plotid)
@@
 # This model makes Gaussian assumptions, which should be incorrect for these data
-shipley.test(oak_path1, cov(select(oak_2005, elevation, tmi, os_rich, hrs_14_20, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 170)
-# The p-value of 0.1114 indicates that there is no severe violation of the conditional independence assumptions with the application of a Gaussian model
+shipley.test(oak_path1, cov(select(oak_2005, elev15m, twi15m, os_rich, hrs_14_20, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 170)
+# The p-value of 0.08 indicates that there is not a severe violation of the conditional independence assumptions with the application of a Gaussian model
+
+shipley.test(oak_path2, cov(select(oak_2005, elev15m, twi15m, os_rich, hrs_blw10, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 170)
+# The p-value of .04 indicates that the conditional independence assumptions under the application of a Gaussian model are questionable.
 ```
 
-~~Violation of conditional independence supports addressing the complexity of the data using different model distributions to deal with the data.~~
-
-Still I am going to fully asses the conditional independence claims
+Violation or questioning of conditional independence supports addressing the complexity of the data using different model distributions to deal with the data.
 
@@
 bu_oak_path1 # 15 independence claims means 15 models
-library(nlme)
-library(lme4)
+library(nlme); library(lme4); library(MuMIn)
 
@@
 summary(f1) # p = 0.5091
-f2 <- lme(avg_psi ~ elevation, data = oak_2005, random = ~1|plotid)
+f2 <- lme(avg_psi ~ elev15m, data = oak_2005, random = ~1|plotid)
 summary(f2) # p = 0.2013
-f3 <- lme(avg_psi ~ tmi + elevation, data = oak_2005, random = ~1|plotid)
-summary(f3) # tmi; p = 0.1982
-f4 <- lme(os_rich ~ avg_psi + tmi, data = oak_2005, random = ~1|plotid)
-summary(f4) # avg_psi; p = 0.1066
+f3 <- lme(avg_psi ~ twi15m + elev15m, data = oak_2005, random = ~1|plotid)
+summary(f3) # twi15m; p = 0.0959
+f4 <- lme(os_rich ~ avg_psi + twi15m, data = oak_2005, random = ~1|plotid)
+summary(f4) # avg_psi; p = 0.1712
 f5 <- lm(log1p(tot_lfct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
@@
 summary(f6) # avg_psi; p = 0.08970
-f7 <- lm(elevation ~ d2c, data = oak_2005)
-# f7 <- lme(elevation ~ d2c, data = oak_2005, random = ~1|plotid)
-summary(f7) # p = 0.03163
-f8 <- lm(log(tmi) ~ d2c + elevation, data = oak_2005)
-# f8 <- lme(log(tmi) ~ d2c + elevation, data = oak_2005, random = ~1|plotid)
-summary(f8) # d2c; p = 0.4375
-f9 <- lm(os_rich ~ d2c + tmi, data = oak_2005)
-# f9 <- lme(os_rich ~ d2c + tmi, data = oak_2005, random = ~1|plotid)
-summary(f9) # d2c; p = 0.997
-f10 <- lm(os_rich ~ elevation + tmi, data = oak_2005)
-# f10 <- lme(os_rich ~ elevation + tmi, data = oak_2005, random = ~1|plotid)
-summary(f10) # elevation; p = 0.260
+f7 <- lm(elev15m ~ d2c, data = oak_2005)
+# f7 <- lme(elev15m ~ d2c, data = oak_2005, random = ~1|plotid)
+summary(f7) # p = 0.03184
+f8 <- lm(log(twi15m) ~ d2c + elev15m, data = oak_2005)
+# f8 <- lme(log(twi15m) ~ d2c + elev15m, data = oak_2005, random = ~1|plotid)
+summary(f8) # d2c; p = 0.3693
+f9 <- lm(os_rich ~ d2c + twi15m, data = oak_2005)
+# f9 <- lme(os_rich ~ d2c + twi15m, data = oak_2005, random = ~1|plotid)
+summary(f9) # d2c; p = 0.9370
+f10 <- lm(os_rich ~ elev15m + twi15m, data = oak_2005)
+# f10 <- lme(os_rich ~ elev15m + twi15m, data = oak_2005, random = ~1|plotid)
+summary(f10) # elev15m; p = 0.19174
 
-# f11 <- lm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
-# f11 <- lm(log1p(tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
-f11 <- glm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, family = "poisson")
+# f11 <- lm(tot_lfct ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
+# f11 <- lm(log1p(tot_lfct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
+f11 <- glm(tot_lfct ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, family = "poisson")
 chat <- deviance(f11)/df.residual(f11)
+library(MuMIn)
+x.quasipoisson<-function(...){
+  res<-quasipoisson(...)
+  res$aic<-poisson(...)$aic
+  res
+}
 QAIC(update(f11, family = x.quasipoisson), chat = chat)
-f11 <- glm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, family = "quasipoisson")
-# f11 <- lme(log1p(tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), random = ~1|plotid, data = oak_2005)
-# f11 <- glmer(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2005, family = "poisson")
-summary(f11) # elevation; p = 0.353425
+f11 <- glm(tot_lfct ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, family = "quasipoisson")
+# f11 <- lme(log1p(tot_lfct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), random = ~1|plotid, data = oak_2005)
+# f11 <- glmer(tot_lfct ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2005, family = "poisson")
+summary(f11) # elev15m; p = 0.360630
 
-# f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "binomial")
+# f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "binomial")
 # chat <- deviance(f12)/df.residual(f12)
 # QAIC(update(f12, family = x.quasibinomial), chat = chat)
-# f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "quasibinomial")
-f12 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
-summary(f12) # elevation; p = 0.45826
+# f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "quasibinomial")
+f12 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
+summary(f12) # elevation; p = 0.44861
 
-f13 <- lm(log1p(tot_lfct) ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
+f13 <- lm(log1p(tot_lfct) ~ scale(twi15m) + scale(d2c) + scale(elev15m) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
 # f13 <- lme(log1p(tot_lfct) ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, random = ~1|plotid)
 # f13 <- glmer(tot_lfct ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2005, family = "poisson")
-summary(f13) # tmi; p = 0.113050
+summary(f13) # twi15m; p = 0.801
 
@@
 
-f14 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
-summary(f14) # tmi; p = 0.67609
+f14 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(twi15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
+summary(f14) # twi15m; p = 0.61177
 
-f15 <- lm(os_rich ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2005)
+f15 <- lm(os_rich ~ hrs_14_20 + avg_psi + d2c + elev15m + twi15m, data = oak_2005)
 # f15 <- lme(os_rich ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2005, random = ~1|plotid)
-summary(f15) # hrs_14_20; p = 0.355
+summary(f15) # hrs_14_20; p = 0.08468
 
@@
 ```
+
 Given the C-statistic of 26.9 and that 2k degrees of freedom would be 30 I think that this meets the criteria for supporting the conditional independence claims.
@@
 
-      1. os_rich ~ tmi,
-      2. hrs_14_20 ~ elevation + d2c + avg_psi + tmi,
+      1. os_rich ~ twi15m,
+      2. hrs_14_20 ~ elev15m + d2c + avg_psi + twi15m,
       3. tot_lfct ~ hrs_14_20 + os_rich + d2c,

### 2015-12-21 explore_veg_data.Rmd
added a write out of canopy cover estimated for each plot via the spherical densiometer
Minor changes to enhance data summaries.

--- 2015-06-30 explore_veg_data.Rmd
@@
 canopy_cover$year <- year(canopy_cover$date)
+canopy_cover <- select(canopy_cover, -date)
+canopy_cover$plotid <- tolower(canopy_cover$plotid)
+write.csv(canopy_cover, "analysis/data/plot_canopy_cover_dens.csv")
 ```
@@
 Here I am adding a column that calculates the proportional coverage for each species based on the number of points they were observed in throughout the plot and the total number of points sampled in the plot.
-```{r calculate overstory proportianal coverage}
-mutate(veg_os_summ, pcov = ptct/totpts)
+```{r calculate overstory proportianal coverage at each point}
+veg_os_summ <- mutate(veg_os_summ, pcov = ptct/totpts)
+veg_os_summ
 ```
@@
 unique(filter(veg_us, PlotID == "ann33", year == 2011)$UnderstorySp)
+filter(veg_us, UnderstorySp == "")
 veg_us$UnderstorySp[veg_us$PlotID == "ann33" & veg_us$year == 2011 & veg_us$UnderstorySp == "losp"] <- "lohi"
@@
 ```
+
+Now I need to examine the data for alternate and non-host species DBH. These are the stems that are untagged and meet a 5cm DBH threshold.
+
+```{r untagged species}
+summary(alt_sp_dbh)
+library(lubridate)
+alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
+filter(alt_sp_dbh, is.na(date)) # this should be 5/22/2012
+alt_sp_dbh$Date[alt_sp_dbh$Date == ""] <- "5/22/2012 0:00:00"
+alt_sp_dbh$date <- as.Date(alt_sp_dbh$Date, format = "%m/%d/%Y")
+summary(alt_sp_dbh)
+alt_sp_dbh$year <- year(alt_sp_dbh$date)
+alt_sp_dbh <- alt_sp_dbh %>% select(-PlotVisit_PlotVisitID, -Date) %>% 
+      rename(dbh = Measure)
+unique(as.character(alt_sp_dbh$Species))
+
+alt_sp_dbh$Species[alt_sp_dbh$Species == "PSME"] <- "psme"
+alt_sp_dbh$Species <- tolower(alt_sp_dbh$Species)
+summary(alt_sp_dbh)
+unique(alt_sp_dbh$Species)
+unique(alt_sp_dbh$Dead)
+# alt_sp_dbh$Dead <- ifelse(alt_sp_dbh$Dead == 0, "no", "yes")
+alt_sp_dbh <- as.tbl(alt_sp_dbh)
+
+alt_sp_live <- filter(alt_sp_dbh, Dead == 0)
+alt_sp_dead <- filter(alt_sp_dbh, Dead == 1)
+
+alt_sp_summ <- left_join(
+      # First data frame
+      alt_sp_live %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
+      summarise(live_count = length(Species), live_avg_dbh = mean(dbh), live_tot_dbh = sum(dbh)),
+      # Second data frame
+      alt_sp_dead %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
+      summarise(dead_count = length(Species), dead_avg_dbh = mean(dbh), dead_tot_dbh = sum(dbh))
+)
+summary(alt_sp_summ)
+
+write.csv(alt_sp_summ, "analysis/data/untagged_dbh_summary.csv")
+```

### 2015-12-21 oak_infection_model.Rmd
Adjust data and models.

@@
 summary(stems)
-summary(plot_qusp)
-summary(plots_umca)
+summary(qusp_plots)
+qusp_plots$plot <- tolower(qusp_plots$plot)
+qusp_plots <- rename(qusp_plots, plotid = plot)
+summary(umca_plots)
 ```
@@
 names(stems)
+
+inf_oak_stems <- stems %>% 
+      select(plotid, year, species, tag, status, canker, dbh2) %>% 
+      filter(species != "umca", species != "lide", species != "unknown sp.", canker != 0)
+
+summary(inf_oak_stems)
+summary(filter(inf_oak_stems, year == 2011))
+filter(inf_oak_stems, year == 2014)
+qplot(dbh2, data = inf_oak_stems)
+
 plot_aggregates <- stems %>%
-      select(plot, species, status, slc, canker, sod_dead, dbh, year) %>%
-      group_by(year, plot) %>% 
+      select(plotid, slc, canker, sod_dead, dbh, year) %>%
+      group_by(year, plotid) %>% 
       summarise(slc_total = sum(slc, na.rm=T), slc_avg = mean(slc, na.rm=T), oak_inf = sum(canker, na.rm=T), oak_dead = sum(sod_dead, na.rm=T), dbh_total = sum(dbh, na.rm=T), dbh_avg = mean(dbh, na.rm=T))
@@
 
-summary(plot_qusp)
-plot_aggregates <- left_join(plot_aggregates, select(plot_qusp, year, plot, uninfected_oak_ct, infected_oak_ct, tot_oak), by = c("year", "plot"))
-summary(plots_umca)
-plot_aggregates <- left_join(plot_aggregates, select(plots_umca, year, plot, uninfected_bay_ct, infected_bay_ct, tot_bay))
+summary(qusp_plots)
+plot_aggregates <- left_join(plot_aggregates, select(qusp_plots, year, plotid, uninfected_oak_ct, infected_oak_ct, tot_oak), by = c("year", "plotid"))
+summary(umca_plots)
+plot_aggregates <- left_join(plot_aggregates, select(umca_plots, year, plotid, uninfected_bay_ct, infected_bay_ct, tot_bay))
+
+plot_aggregates
+filter(plot_aggregates, year == 2011) %>% group_by(plotid) %>% summarise(inf_oak_ct = sum(infected_oak_ct, na.rm = T))
+
+plot_aggregates %>% group_by(year) %>% summarise(oak_inf = sum(oak_inf), oak_dead = sum(oak_dead))
+sum(plot_aggregates$oak_inf, na.rm = T)
 ```
@@
       facet_grid(year ~ .)
+ggplot(plot_aggregates)
 
-oak.m1 <- glmer(cbind(infected_oak_ct, uninfected_oak_ct) ~ I(tot_bay/225) + (1 | plot), data = plot_aggregates, family = "binomial")
+oak.m1 <- glmer(cbind(infected_oak_ct, uninfected_oak_ct) ~ I(tot_bay/225) + (1 | plotid), data = plot_aggregates, family = "binomial")
 summary(oak.m1)
 
-oak.m2 <- glmer(cbind(infected_oak_ct, uninfected_oak_ct) ~ log1p(tot_bay) + (1 | plot), data = plot_aggregates, family = "binomial")
+oak.m2 <- glmer(cbind(infected_oak_ct, uninfected_oak_ct) ~ log1p(tot_bay) + (1 | plotid), data = plot_aggregates, family = "binomial")
 summary(oak.m2)

### 2015-12-21 sem_development.Rmd
Update rainfall data sources, rename some objects.
Saved data file 'data_201508.RData' holding the R data objects.

@@
 summary(plots_vegtopo)
-plots_vegtopo <- select(plots_vegtopo, -X, -Y)
+plots_vegtopo <- select(plots_vegtopo, -cat) %>% rename(lon_utm = X, lat_utm = Y)
 ```
@@
 
-# Rainfall: interpolated rainfall for each season
-plots_rain <- read.csv("analysis/data/plot202rain.csv")
-summary(plots_rain)
-plots_rain <- select(plots_rain, -cat)
+# Rainfall: interpolated/estimated rainfall for each season
+load("~/Documents/soco_ppt/plot_ppt_estimates.RData")
+summary(ppt_days)
+summary(ppt_total)
+# plots_rain <- read.csv("analysis/data/plot202rain.csv")
+# summary(plots_rain)
+# plots_rain <- select(plots_rain, -cat)
 # pairs(select(plots_rain, -PLOT_ID), 
@@
 library(ggplot2)
-summary(plots_vegtopo)
-par(mfrow = c(3,1))
+# summary(plots_vegtopo)
+# par(mfrow = c(3,1))
 qplot(slope15m, flowacc15m, data = plots_env) +
@@
 qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m < 20), 
-      main = "Flow Accumulation < 5") +
+      main = "Flow Accumulation < 20") +
       geom_text(aes(label = plotid))
-qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 20 & flowacc15m < 1000), main = "Flow Accumulation 5 - 1000") +
+qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 20 & flowacc15m < 1000), main = "Flow Accumulation 20 - 1000") +
       geom_text(aes(label = plotid))
@@
 # This only has number of infected, uninfected, and total, which must apply to living stems only. I want to analyze leaf counts.
+unique(plots_umca$plot)
+umca_plots <- filter(plots_umca, plot != "BUSH01")
+rm(plots_umca)
 summary(bay_laurel) # This has stem-level data on bay laurel.
+unique(bay_laurel$plot)
+bay_laurel <- filter(bay_laurel, plot != "BUSH01")
 ```
@@
 ```{r calculate plot level leaf counts}
-plots_umca <- left_join(plots_umca, bay_laurel %>% filter(status == "Alive") %>% 
-      group_by(plot, year) %>% 
-      summarise(
+umca_plots <- left_join(umca_plots, bay_laurel %>% filter(status == "Alive") %>% group_by(plot, year) %>%  summarise(
             avg_lfct = mean(slc, na.rm = T),
@@
       by = c("plot", "year"))
-summary(plots_umca)
+summary(umca_plots)
 ```
@@
 ```{r examine lfct NAs}
-filter(plots_umca, is.na(avg_lfct))
+filter(umca_plots, is.na(avg_lfct))
 # I'm not sure why this wasn't excluded from the calculations with the na.rm = T option
@@
 # I'm removing it manually from the `plots_umca` data frame
-plots_umca <- plots_umca %>% filter(avg_lfct != "NaN")
-summary(plots_umca)
+umca_plots <- umca_plots %>% filter(avg_lfct != "NaN")
+summary(umca_plots)
 ```
@@
 
+#### Community vegetation data and variables
 Now I also want to calculate some plot-level community variables from the vegetation data. First, understory richness from the vegetation transects.
@@
 veg <- rbind(veg_us_2005, veg_us_2011)
-
 unique(veg$year)
@@
 woody_spp <- c(woody_spp, "ceanothus sp.", "cebe", "cecu", "frla", "garrya sp.", "gemo", "juca", "mafu", "mane", "mimulus sp", "pisa", "quag x quke", "qube", "quercus sp.", "quwi", "rhamnus sp.", "rhcr", "ribes sp.", "rica", "risa", "roca", "rosa sp.", "rubus sp.", "rudi", "salix sp.", "syal", "vaccinium sp.")
-
 veg_sub <- filter(veg, veg$us_species %in% woody_spp)
+veg_sub <- droplevels(veg_sub)
 
@@
 untagged_dbh$plotid <- tolower(untagged_dbh$plotid)
-
+unique(untagged_dbh$plotid)
+untagged_dbh <- filter(untagged_dbh, plotid != "bush01")
 untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species)))
@@
 unique(stems$status)
-
 stems <- stems %>% select(-X) %>% filter(status == "Alive" | status == "Dead")
 unique(stems$species)
+unique(stems$plotid)
 stems$species <- tolower(stems$species)
@@
 stems$plotid <- tolower(stems$plotid)
+stems <- droplevels(stems)
 
-live_tagged_dbh <- stems %>% filter(status == "Alive") %>% group_by(plotid, year, species) %>% summarise(live_count = length(species))
-unique(live_tagged_dbh$species)
-os_richness <- rbind(live_tagged_dbh, select(untagged_dbh, plotid, species, year, live_count))
+live_tagged <- stems %>% filter(status == "Alive") %>% group_by(plotid, year, species) %>% summarise(live_count = length(species))
+unique(live_tagged$species)
+os_richness <- rbind(live_tagged, select(untagged_dbh, plotid, species, year, live_count))
 
@@
 
-Now I want to join the richness data set and the canopy cover estimates to the `plots_umca` data frame that currently has the environmental and leaf count/infection variables.
+Now I want to join the richness data set and the canopy cover estimates to the `umca_plots` data frame that currently has the environmental and leaf count/infection variables.
 
 ```{r join richness and umca data}
-plots_umca
+umca_plots
 richness
-plots_umca <- rename(plots_umca, plotid = plot)
-plots_umca$plotid <- tolower(plots_umca$plotid)
+umca_plots <- rename(umca_plots, plotid = plot)
+umca_plots$plotid <- tolower(umca_plots$plotid)
 
-plots_umca <- left_join(plots_umca, richness, by = c("plotid", "year"))
+umca_plots <- left_join(umca_plots, richness, by = c("plotid", "year"))
 head(canopy_cover)
-canopy_cover <- select(canopy_cover, plotid, year, pct_cvr)
-plots_umca <- left_join(plots_umca, canopy_cover, by = c("plotid", "year"))
+canopy_cover <- select(canopy_cover, -X)
+umca_plots <- left_join(umca_plots, canopy_cover, by = c("plotid", "year"))
 ```
@@
 
-plots_umca <- ungroup(plots_umca)
+umca_plots <- ungroup(umca_plots)
 
@@
 ```{r join umca and qusp data}
-plots_umca
+umca_plots
 plot_qusp # right now this is broken down by oak species
-summary(plot_qusp)
-oak_sod <- plot_qusp %>% group_by(plot, year) %>%
+qusp_plots <- plot_qusp
+rm(plot_qusp)
+summary(qusp_plots)
+qusp_plots
+oak_sod <- qusp_plots %>% group_by(plot, year) %>%
       summarise(uninf_oak_ct = sum(uninfected_oak_ct), 
@@
 
-oak_sod <- left_join(oak_sod, select(plots_umca, plotid, year, uninfected_bay_ct, infected_bay_ct, tot_bay, avg_lfct, tot_lfct), 
+oak_sod <- left_join(oak_sod, select(umca_plots, plotid, year, uninfected_bay_ct, infected_bay_ct, tot_bay, avg_lfct, tot_lfct), 
                      by = c("plotid", "year"))
 summary(oak_sod) # the NA's for the bay laurel variables are legitimately zeroes b/c there are no bay in some of these plots
+filter(oak_sod, is.na(tot_bay))
 oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
@@
 
+plots_env
+plots_env$plotid <- tolower(plots_env$plotid)
 oak_sod <- left_join(oak_sod, plots_env, by = "plotid")
-oak_sod <- left_join(oak_sod, plots_temps, 
-                     by = c("plotid", "year" = "year"))
+head(temps_data, 15)
+temps_data$plotid <- tolower(temps_data$plotid)
+temps_data <- rename(temps_data, year = sample_year)
+oak_sod <- left_join(oak_sod, temps_data, by = c("plotid", "year"))
 oak_sod
+names(oak_sod)
 
@@
 summary(oak_sod)
+filter(oak_sod, is.na(elev15m))
+oak_sod <- filter(oak_sod, plotid != "duer01", plotid != "ponti01", plotid != "sweet01")
+
 summary(filter(oak_sod, is.na(avg_tmax)))
@@
 oak_sod <- left_join(oak_sod, canopy_cover, by = c("plotid", "year"))
+summary(rain_data)
+rain_data$plotid <- tolower(rain_data$plotid)
+oak_sod <- left_join(oak_sod, rain_data, by = c("plotid", "year"))
+summary(filter(oak_sod, is.na(avg_tmax))) # all obs from 2003
 
@@
 
-## Path Analysis Oak Infection 2005
+# Path Analysis Presented at 2015 IALE World Congress
 
+#### Path Analysis Plot-Level Oak Infection 2005
 Here I am using the `oak_sod` data set which will be the oak infection (binomial) as the terminal response variable.
@@
 
-## Path Analysis Oak Infection - 2014
+#### Path Analysis Oak Infection - 2014
 Now I am going to fit the same path model to data from 2011.
@@
 
-```{r oak path 2}
+```{r oak path 2014 v2}
 oak_path2 <- DAG(
@@
 
+#### Importance Value
 I also want to calculate some variables related to DBH, and in particular the "importance value" for each species. Importance value is calculated in the literature as:
@@
 `IV = 0.5 * (relative density / relative dominance)` where 
-`Relative Density = # stems of species / total # of stems` and
-`Relative Dominance = basal area of species / total basal area`
+ - `Relative Density = # stems of species / total # of stems` and
+ - `Relative Dominance = basal area of species / total basal area`
 
@@
 
-stems_dbh <- rbind(tagged_dbh, untagged_dbh)
+species_dbh <- rbind(tagged_dbh, untagged_dbh)
+species_dbh
+unique(species_dbh$species)
+species_dbh %>% group_by(plotid, species, year)
 
-stems_dbh <- stems_dbh %>% group_by(plotid, year) %>% mutate(all_live_count = sum(live_count), all_live_dbh = sum(live_tot_dbh, na.rm = T))
+stems_dbh <- species_dbh %>% group_by(plotid, year) %>% mutate(all_live_count = sum(live_count), all_live_dbh = sum(live_tot_dbh, na.rm = T), all_dead_count =  sum(dead_count, na.rm = T), all_dead_dbh = sum(dead_tot_dbh, na.rm = T))
+stems_dbh
 summary(stems_dbh)
@@
 ```
+
+```{r pare down stems dataframe}
+stems %>% select(year, tag, species, status, notes) %>% filter(species == "umca") 
+summary(stems) # Limited to "Alive/Dead"" status, "Missing" statuses excluded
+unique(stems$species)
+unique(stems$plotid)
+summary(bay_laurel) # Still includes stems with "Missing" status
+unique(bay_laurel$plot)
+
+names(stems)
+stemssub <- select(stems, plotid, year, cluster, tag, species, status, slc, canker, dbh, lide_lf_symp, sod_dead, location, dbh1, dbh2, delta_dbh, dbh2_1_ratio, inst_grwth_rate, doy, date)
+summary(stemssub)
+str(stemssub)
+
+summary(oak_sod) # plots with oak stems
+summary(umca_plots) # plots with bay laurel stems
+summary(rain_data)
+summary(temps_data)
+summary(plots_env)
+summary(veg_sub)
+summary(richness)
+summary(stems_dbh)
+
+save.image("~/GitHub/superspreaders/data_201508.RData")
+```
+
+```{r}
+load("~/GitHub/superspreaders/data_201508.RData")
+stemssub$cluster <- as.factor(stemssub$cluster)
+stemssub$tag <- as.factor(stemssub$tag)
+stemssub$canker <- as.factor(stemssub$canker)
+stemssub$sod_dead <- as.factor(stemssub$sod_dead)
+stemssub$date <- as.Date(stemssub$date)
+str(stemssub)
+```

### 2015-12-21 pathogen_biomass.R
Add script exploring pathogen ~ biomass relationship.

@@ -1,0 +1,29 @@
+# Pathogen ~ Biomass relationship for parameterizing LANDIS-II extension
+# From Francesco's email: "if we come up with a decent relationship to use biomass related to pathogen production we can improve the current model"
+load("~/GitHub/superspreaders/data_201508.RData")
+
+summary(bay_laurel)
+summary(stems_dbh)
+summary(stems)
+
+library(ggplot2); library(dplyr)
+
+qplot(dbh1, data = bay_laurel %>% filter(status == "Alive", dbh1 >= 2)) +
+      geom_vline(xintercept = c(10.2, 13.34))
+
+qplot(dbh2, data = bay_laurel %>% filter(status == "Alive", dbh2 >= 2)) +
+      geom_vline(xintercept = c(11.1, 14.64), colour = "red", linetype = "longdash")
+qplot(log(dbh2), data = bay_laurel %>% filter(status == "Alive", dbh2 >= 2))
+
+qplot(slc, data = bay_laurel %>% filter(status == "Alive", dbh2 >= 2, year > 2003))
+qplot(log1p(slc), data = bay_laurel %>% filter(status == "Alive", dbh2 >= 2, year > 2003))
+
+p1 <- ggplot(filter(bay_laurel, status == "Alive", dbh2 > 2, year > 2003), 
+             aes(x = log(dbh2), y = log1p(slc))) +
+      geom_point() +
+      facet_grid(year ~ .)
+
+quantile(filter(bay_laurel, dbh2 >=2)$slc, na.rm = T)
+quantile(filter(bay_laurel, dbh2 >= 2)$dbh2, 0.9, na.rm = T)
+
+log1p(filter(bay_laurel, dbh2 >=2)$slc, na.rm = T)

### 2015-12-21 vegan_diversity.Rmd
Add script exploring diversity analysis using 'vegan' package.

---
title: "vegan Diversity"
author: "Whalen Dillon"
date: "August 24, 2015"
output: html_document
---

```{r load vegan package, echo=FALSE, results='hide'}
library(vegan)
data(BCI)
summary(BCI)
str(BCI)
head(BCI)
names(BCI)
```

The data example is a wide format, with each species a column and the number (abundunce) of that species observed in each row. Each row corresponds to the observation unit, in this case 1-hectare plots.

The `diversity` function finds the Shannon-Wiener, Simpson, and inverse Simpson diversity indices:

  - Shannon-Wiener: $H = -\sum_{i = 1}^{S} {(p_{i}) (log_b p_i)}$
  - Simpson: $D_1 = 1 - \sum_{i = 1}^{S} p_i^2$ 
  - inverse Simpson: $D_2 = 1 / \sum_{i = 1}^{S} p_i^2$

     - $p_i$ = proportion of species *i*
     - *S* = number of species
     - *b* = base of logarithm

The default uses the natural logarithm.

```{r calculate diversity}
H <- diversity(BCI, index = "shannon", MARGIN = 1, base = exp(1))
D1 <- diversity(BCI, index = "simpson")
D2 <- diversity(BCI, index = "invsimpson")
H
D1
D2
```

#### Species Abundance Models

Fisher's log-series estimates the expected number of species with *n* individuals, plotting species by frequencies. 
```{r Fishers log series}
k <- sample(nrow(BCI), 1)
fish <- fisherfit(BCI[k,])
plot(fish)
fish
```

Preston's log-normal model bins species into frequency classes of increasing sizes instead of plotting species by frequencies. There are two alternative fucntions in the `vegan` pacakge for fitting the lognormal model:

  1. `prestonfit` uses binning approach with arbitrary choices of limits and ties
     - for ties between adjacent octaves half were in 1st octave and half transferred to next octave
     - can either split ties or keep all limit cases in lower octave
  2. `prestondistr` maximizes truncated log-normal likelihood without binning data
     - recommended in `vegan` vignette paper

```{r Prestons lognormal}
prest <- prestondistr(BCI[k,])
plot(prest)
prest
```

This uses maximum-likelihood method and, I think, log base 2 (?)

#### Whittaker Plots - Ranked Abundance Distribution
The `vegan` function `radfit` uses maximum likelihood to fit popular models:

  - brokenstick: null model with no estimated parameters in `vegan`
  - preemption: one estimated parameter ($\alpha$)
  - log-normal: two estimated parameters ($\mu, sigma$)
  - Zipf: two estimated parameters (p1, $\gamma$)
  - Zipf-Mandlebrot: three estimated parameters (*c*, $\beta, gamma$)

The models are customarily defined for proportions, but `radfit` works directly with abundance data.

```{r Whittaker plots}
rad <- radfit(BCI[k,], family = poisson)
plot(rad)
radlattice(rad)
rad
```

This function compares models using AIC or BIC (based on log-likelihood, penalized by number of estimated parameters).

### 2015-12-21 sem_development.Rmd
Update rainfall data sourcing.

@@
 ```
-```{r load physical data, echo=FALSE}
+```{r load physical data, echo=FALSE, message=F}
 # Physical environment: topography
@@
 #       diag.panel = panel.hist)
-plots_rain <- rename(plots_rain, plotid = PLOT_ID)
+ppt_days@data <- rename(ppt_days@data, plotid = PLOT_ID)
+ppt_total@data <- rename(ppt_total@data, plotid = PLOT_ID)
 ```
@@
 library(tidyr)
-rain2d <- plots_rain %>% select(plotid, ends_with("2d"))
-names(rain2d) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain2d <- rain2d %>% gather(year, rain_tot_2d, -plotid)
-
-rain3d <- plots_rain %>% select(plotid, ends_with("3d"))
-names(rain3d) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain3d <- rain3d %>% gather(year, rain_tot_3d, -plotid)
-
-rain_v <- plots_rain %>% select(plotid, ends_with("v"))
-names(rain_v) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain_v <- rain_v %>% gather(year, rain_tot_v, -plotid)
-
-rain_data <-  left_join(rain2d, 
-                        left_join(rain3d, rain_v, by = c("plotid","year")),
+rain2d_total <- ppt_total@data %>% select(plotid, ends_with("2d"))
+names(rain2d_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain2d_total <- rain2d_total %>% gather(year, rain_tot_2d, -plotid)
+
+rain3d_total <- ppt_total@data %>% select(plotid, ends_with("3d"))
+names(rain3d_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain3d_total <- rain3d_total %>% gather(year, rain_tot_3d, -plotid)
+
+rain_v_total <- ppt_total@data %>% select(plotid, ends_with("v"))
+names(rain_v_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain_v_total <- rain_v_total %>% gather(year, rain_tot_v, -plotid)
+
+rain_r_total <- ppt_total@data %>% select(plotid, ends_with("r"))
+names(rain_r_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain_r_total <- rain_r_total %>% gather(year, rain_tot_r, -plotid)
+
+rain_psm_total <- ppt_total@data %>% select(plotid, contains("sum"))
+names(rain_psm_total) <- c("plotid", paste(c(2004:2010), sep = ","))
+rain_psm_total <- rain_psm_total %>% gather(year, rain_tot_psm, -plotid)
+
+rain2d_days <- ppt_days@data %>% select(plotid, ends_with("2d"))
+names(rain2d_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain2d_days <- rain2d_days %>% gather(year, rain_days_2d, -plotid)
+
+rain3d_days <- ppt_days@data %>% select(plotid, ends_with("3d"))
+names(rain3d_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain3d_days <- rain3d_days %>% gather(year, rain_days_3d, -plotid)
+
+rain_v_days <- ppt_days@data %>% select(plotid, ends_with("v"))
+names(rain_v_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain_v_days <- rain_v_days %>% gather(year, rain_days_v, -plotid)
+
+rain_r_days <- ppt_days@data %>% select(plotid, ends_with("r"))
+names(rain_r_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+rain_r_days <- rain_r_days %>% gather(year, rain_days_r, -plotid)
+
+rain_total <- left_join(rain2d_total, 
+                        left_join(rain3d_total, rain_v_total, 
+                                  by = c("plotid","year")),
                         by = c("plotid","year"))
-summary(rain_data)
-rm(rain_v); rm(rain2d); rm(rain3d)
-write.csv(rain_data, "analysis/data/rain_data_long.csv", row.names = F)
+rain_total <- left_join(rain_total, rain_r_total, by = c("plotid", "year"))
+rain_total <- left_join(rain_total, rain_psm_total, by = c("plotid", "year"))
+
+rain_days <- left_join(rain2d_days,
+                       left_join(rain3d_days, rain_v_days, 
+                                 by = c("plotid", "year")),
+                       by = c("plotid", "year"))
+rain_days <- left_join(rain_days, rain_r_days, by = c("plotid", "year"))
+
+rm(list = ls(pattern = "^rain_v"))
+rm(list = ls(pattern = "^rain2d"))
+rm(list = ls(pattern = "^rain3d"))
+rm(list = ls(pattern = "^rain_r"))
+rm(rain_psm_total)
+
+write.csv(rain_total, "analysis/data/rain_total_long.csv", row.names = F)
+write.csv(rain_days, "analysis/data/rain_days_long.csv", row.names = F)
 ```

### 2015-12-21 sem_development.Rmd
Edit rainfall data frame filtering for days and totals.

@@
 summary(plots_env)
-plots_env <- select(plots_env, -cat)
+plots_env <- select(plots_env, -cat, -X, -Y)
 
@@
 plots_env$elev_m <- plots_env$elev15m * .3048
-plots_env <- select(plots_env, -elevation, -tmi) #remove elevation, tmi from old derivation (I think a legacy 10m USGS DEM)
+# plots_env <- select(plots_env, -elevation, -tmi) #remove elevation, tmi from old derivation (I think a legacy 10m USGS DEM)
 plots_env$veght_m <- plots_env$veght15m * .3048
@@
 str(plots_env)
-plots_env <- plots_env %>% rename(plotid = PLOT_ID, easting = X, northing = Y)
-plots_env <- select(plots_env, -cat)
+plots_env <- plots_env %>% rename(plotid = PLOT_ID)
+
 write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
@@
 summary(plots_temps)
-summary(rain_data)
+summary(rain_days)
+summary(rain_total)
 unique(plots_temps$plotid)# 201 levels
-unique(rain_data$plotid)# 202 levels
+unique(rain_days$plotid)# 202 levels
+unique(rain_total$plotid)# 202 levels
 tempsid <- as.character(unique(plots_temps$plotid))
-rainid <- as.character(unique(rain_data$plotid))
+rainid <- as.character(unique(rain_days$plotid))
 rainid %in% tempsid
 rainid[62]# BUSH01 still in the rain data
-rain_data <- filter(rain_data, plotid != "BUSH01")
+rain_days <- filter(rain_days, plotid != "BUSH01")
+rain_total <- filter(rain_total, plotid != "BUSH01")
 # Temperature and rainfall data now have same number of observations

### 2015-12-23 sem_development.Rmd
Update temperature data sources.

@@
 # Temperature: rainy season hour summaries (long format for each year)
-plots_temps <- read.csv("analysis/data/rainy_season_temps.csv")
-summary(plots_temps)
-plots_temps <- select(plots_temps, -X)
+rs_temps <- read.csv("analysis/data/rs_select_temps.csv")
+summary(rs_temps)
+rs_temps <- select(rs_temps, -X)
+
+summer_temps <- read.csv("analysis/data/summer_select_temps.csv")
+summary(summer_temps)
+summer_temps <- select(summer_temps, -X)
 

### 2015-12-23 analysis/explore_data.R
Aesthetic changes to code (spacing).

@@ -564,17 +564,12 @@ matplot(cbind(avg_rain_year,avg_slc_year),type="b")
 matplot(cbind(rainy_days,avg_slc_year),type="b")
 
 library(plotrix)
-twoord.stackplot(lx=years,rx=years,ldata=avg_slc_year,rdata=avg_rain_year,lcol="red",rcol="blue",
-                 ltype="b",rtype="b",rylab="Mean SLC",lylab="Mean Rainfall",xlab="Year")
+twoord.stackplot(lx=years, rx=years, ldata=avg_slc_year, rdata=avg_rain_year, lcol="red", rcol="blue", ltype="b", rtype="b", rylab="Mean SLC", lylab="Mean Rainfall", xlab="Year")
 
 library(ggplot2)
-qplot(years,avg_slc_year,geom=c("point","line"))
+qplot(years, avg_slc_year, geom = c("point", "line"))
 
 ggplot() +
-      geom_line(data=bay_dat1,aes(x=years,y=colMeans(cbind(amou2004,amou2005,amou2006,amou2007,amou2008,
-                                                           amou2009,amou2010,amou2011,amou2012)),
-                                  colour="blue")) +
-      geom_line(data=bay_dat1,aes(x=years,y=colMeans(cbind(cbind(avgslc04,avgslc05,avgslc06,avgslc07,avgslc08,
-                                                                 avgslc09,avgslc10,avgslc11,avgslc12))),
-                                  colour="red"))
+      geom_line(data = bay_dat1, aes(x = years, y = colMeans(cbind(amou2004, amou2005, amou2006, amou2007, amou2008, amou2009, amou2010, amou2011, amou2012)), colour="blue")) +
+      geom_line(data=bay_dat1,aes(x=years,y=colMeans(cbind(cbind(avgslc04, avgslc05, avgslc06, avgslc07, avgslc08, avgslc09, avgslc10, avgslc11, avgslc12))), colour="red"))
 

### 2016-01-07 explore_veg_data.Rmd
Recalculate/correct alt species abundances.

@@
 
-alt_sp_live <- filter(alt_sp_dbh, Dead == 0)
-alt_sp_dead <- filter(alt_sp_dbh, Dead == 1)
+summary(filter(alt_sp_dbh, year == 2003))
+unique(droplevels(filter(alt_sp_dbh, year == 2003)$PlotID))
+pid2003 <- as.character(unique(filter(alt_sp_dbh, year == 2003)$PlotID))
+alt_sp_2003 <- filter(alt_sp_dbh, year == 2003)
 
-alt_sp_summ <- left_join(
+summary(filter(alt_sp_dbh, year == 2004))
+unique(droplevels(filter(alt_sp_dbh, year == 2004)$PlotID))
+pid2004 <- as.character(unique(filter(alt_sp_dbh, year == 2004)$PlotID))
+alt_sp_2004 <- filter(alt_sp_dbh, year == 2004)
+
+pid2003 %in% pid2004
+pid2004[47]; pid2004[69]
+filter(alt_sp_dbh, year == 2003, PlotID == "MITSU02")
+filter(alt_sp_dbh, year == 2004, PlotID == "MITSU02")
+
+summary(filter(alt_sp_dbh, year == 2005))
+unique(droplevels(filter(alt_sp_dbh, year == 2005)$PlotID))
+pid2005 <- as.character(unique(filter(alt_sp_dbh, year == 2005)$PlotID))
+alt_sp_2005 <- filter(alt_sp_dbh, year == 2005)
+
+# I think I should work this backwards so that the most recent sampling is added to
+pid2005 %in% pid2004# COOK02 is the only duplicate in the 2004 sampling
+filter(alt_sp_2004, PlotID=="COOK02")
+filter(alt_sp_2005, PlotID=="COOK02")
+# But there are a lot more records in the 2004 data, which is closer to the 2012 & 2014 sampling. I could combine the 2004 & 2005 data, including the potentially duplicated COOK02 data.
+summary(filter(alt_sp_2012, PlotID=="COOK02"))
+filter(alt_sp_2014, PlotID=="COOK02")
+
+alt_sp_45 <- rbind(alt_sp_2005, alt_sp_2004)
+summary(alt_sp_45)
+pid45 <- as.character(unique(alt_sp_45$PlotID))# 144 unique plots
+pid45 %in% pid2003
+pid45[71]; pid45[93]; pid45[133]
+filter(alt_sp_45, PlotID == "MITSU02")
+filter(alt_sp_2003, PlotID == "MITSU02")
+# Exclude MITSU02 from the 2003 data
+alt_sp_2003 <- filter(alt_sp_2003, PlotID != "MITSU02")
+
+filter(alt_sp_45, PlotID == "LINS01")
+filter(alt_sp_2003, PlotID == "LINS01")
+# Keep LINS01 from 2003 data, add to other data
+filter(alt_sp_45, PlotID == "GANAY02")
+filter(alt_sp_2003, PlotID == "GANAY02")
+# Exclude GANAY02 from the 2004, keep the 2003 data
+alt_sp_45 <- filter(alt_sp_45, PlotID != "GANAY02")
+unique(alt_sp_2003$PlotID) %in% unique(alt_sp_45$PlotID)
+alt_sp_45 <- rbind(alt_sp_45, alt_sp_2003)
+alt_sp_2005 <- alt_sp_45
+rm(alt_sp_45)
+
+summary(filter(alt_sp_dbh, year == 2012))
+pid2012 <- unique(droplevels(filter(alt_sp_dbh, year == 2012)$PlotID))# 171 plots
+alt_sp_2012 <- filter(alt_sp_dbh, year == 2012)
+
+summary(filter(alt_sp_dbh, year == 2014))
+pid2014 <- unique(droplevels(filter(alt_sp_dbh, year == 2014)$PlotID))# 164 plots
+alt_sp_2014 <- filter(alt_sp_dbh, year == 2014)
+
+# Determine which plots from 2012 (the one with more records) aren't in the 2014 data.
+pid2012 %in% pid2014
+pid2012sub <- droplevels(pid2012[c(4,12:13,20:21,52:53,111,114)])
+alt_sp_2012 <- filter(alt_sp_2012, PlotID == "ARBIT04" | PlotID == "LINS01" | PlotID == "LUPIN01" | PlotID == "MITSU02" | PlotID == "MITSU04" | PlotID == "SKILE01" | PlotID == "SPAUL01" | PlotID == "GANAY02" | PlotID == "HALOW02")
+
+alt_sp_2014 <- rbind(alt_sp_2014, alt_sp_2012)
+summary(alt_sp_2014)
+summary(alt_sp_2005)
+alt_sp_2005$year <- 2005
+
+alt_sp_2014 <- select(alt_sp_2014, -date)
+alt_sp_2005 <- select(alt_sp_2005, -date)
+
+# Create data frames of live and dead stems for '2005' and '2014' samplings.
+In reality the 2005 grouping has data from visits during 2003 and 2004, and the 2014 grouping includes data from some visits in 2012.
+alt_sp_2014_live <- filter(alt_sp_2014, Dead == 0)
+alt_sp_2014_live <- select(alt_sp_2014_live, -Dead)
+alt_sp_2005_live <- filter(alt_sp_2005, Dead == 0)
+alt_sp_2005_live <- select(alt_sp_2005_live, -Dead)
+alt_sp_2014_dead <- filter(alt_sp_2014, Dead == 1)
+alt_sp_2014_dead <- select(alt_sp_2014_dead, -Dead)
+alt_sp_2005_dead <- filter(alt_sp_2005, Dead == 1)
+alt_sp_2005_dead <- select(alt_sp_2005_dead, -Dead)
+
+# Confirm year assignments are correct/consistent.
+unique(alt_sp_2014_live$year)
+unique(alt_sp_2005_live$year)
+unique(alt_sp_2005_dead$year)
+unique(alt_sp_2014_dead$year)
+
+alt_sp_summ_2005 <- left_join(
       # First data frame
-      alt_sp_live %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
-      summarise(live_count = length(Species), live_avg_dbh = mean(dbh), live_tot_dbh = sum(dbh)),
+      alt_sp_2005_live %>%
+            group_by(PlotID, Species, year) %>% 
+            summarise(
+                  live_count = length(Species),
+                  live_avg_dbh = mean(dbh),
+                  live_tot_dbh = sum(dbh)),
       # Second data frame
-      alt_sp_dead %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
-      summarise(dead_count = length(Species), dead_avg_dbh = mean(dbh), dead_tot_dbh = sum(dbh))
+      alt_sp_2005_dead %>%
+            group_by(PlotID, Species, year) %>% 
+            summarise(
+                  dead_count = length(Species), 
+                  dead_avg_dbh = mean(dbh), 
+                  dead_tot_dbh = sum(dbh))
 )
-summary(alt_sp_summ)
+summary(alt_sp_summ_2005)
 
-write.csv(alt_sp_summ, "analysis/data/untagged_dbh_summary.csv")
+alt_sp_summ_2014 <- left_join(
+      # First data frame
+      alt_sp_2014_live %>%
+            group_by(PlotID, Species, year) %>% 
+            summarise(
+                  live_count = length(Species),
+                  live_avg_dbh = mean(dbh),
+                  live_tot_dbh = sum(dbh)),
+      # Second data frame
+      alt_sp_2014_dead %>%
+            group_by(PlotID, Species, year) %>% 
+            summarise(
+                  dead_count = length(Species), 
+                  dead_avg_dbh = mean(dbh), 
+                  dead_tot_dbh = sum(dbh))
+)
+summary(alt_sp_summ_2014)
+
+# Write out summary data frames of the untagged DBH species
+write.csv(alt_sp_summ_2005, "analysis/data/untagged_dbh2005_summary.csv", row.names = F)
+write.csv(alt_sp_summ_2014, "analysis/data/untagged_dbh2014_summary.csv", row.names = F)
 ```
+
+```{r make wide format species abundance}
+unique(alt_sp_2014_live$Species)
+alt_sp_2014_live$Species[alt_sp_2014_live$Species=="arcto sp."] <- "arcto"
+alt_sp_2014_live$Species[alt_sp_2014_live$Species=="prunus sp."] <- "prunus"
+alt_sp_2014_live$Species[alt_sp_2014_live$Species=="unknown sp. 1"] <- "unk1"
+alt_sp_2014_live$Species[alt_sp_2014_live$Species=="quercus sp."] <- "qusp"
+
+unique(alt_sp_2005_live$Species)
+alt_sp_2005_live$Species[alt_sp_2005_live$Species=="quercus sp."] <- "qusp"
+alt_sp_2005_live$Species[alt_sp_2005_live$Species=="arcto sp."] <- "arcto"
+alt_sp_2005_live$Species[alt_sp_2005_live$Species=="prunus sp."] <- "prunus"
+
+# Aggregate to species counts by plot
+alt_sp_05_live_abund <- alt_sp_2005_live %>% select(-dbh, -year) %>% 
+      group_by(PlotID, Species) %>% summarise(count = length(Species))
+# Spread from long to wide format
+alt_sp_05_live_abund <- alt_sp_05_live_abund %>% spread(Species, count)
+
+alt_sp_14_live_abund <- alt_sp_2014_live %>% select(-dbh, -year) %>% 
+      group_by(PlotID, Species) %>% summarise(count = length(Species))
+alt_sp_14_live_abund <- alt_sp_14_live_abund %>% spread(Species, count)
+
+# Create species abundance data frames of dead species
+alt_sp_05_dead_abund <- alt_sp_2005_dead %>% select(-dbh, -year) %>% 
+      group_by(PlotID, Species) %>% summarise(count = length(Species))
+alt_sp_05_dead_abund <- alt_sp_05_dead_abund %>% spread(Species, count)
+
+alt_sp_14_dead_abund <- alt_sp_2014_dead %>% select(-dbh, -year) %>% 
+      group_by(PlotID, Species) %>% summarise(count = length(Species))
+alt_sp_14_dead_abund <- alt_sp_14_dead_abund %>% spread(Species, count)
+
+# Write out wide format abundance data frames
+write.csv(alt_sp_05_live_abund, "analysis/data/altsp05abund_live.csv")
+write.csv(alt_sp_14_live_abund, "analysis/data/altsp14abund_live.csv")
+write.csv(alt_sp_05_dead_abund, "analysis/data/altsp05abund_dead.csv")
+write.csv(alt_sp_14_dead_abund, "analysis/data/altsp14abund_dead.csv")
+```

### 2016-01-07 analysis/species_diversity_calc.R
Add new script for calculating diversity metrics.

@@ -1,0 +1,180 @@
+#' # Species Diversity Metrics
+#' Now I also want to calculate some plot-level community variables from the vegetation data: diversity and evenness metrics from the vegetation transects
+#' 
+#'  1. over and understory richness; count of number of species
+#'  2. Shannon-Weiner Evenness & Diversity Indices
+#'  3. Simpson's Diversity Index
+#' 
+#' The Shannon and Simpson's indices can be calculated using functions in the `vegan` package. I have done some exploring of this package in the 'vegan_diversity.Rmd' script. 
+#' 
+#+ load plot sampling data
+## Ecological data: leaf counts, infection
+load("~/GitHub/superspreaders/stems_plots.RData")
+
+## Vegetation transect data
+veg_us_2005 <- read.csv("analysis/data/understory_veg_2005.csv")
+veg_us_2011 <- read.csv("analysis/data/understory_veg_2011.csv")
+canopy_cover <- read.csv("analysis/data/plot_canopy_cover_dens.csv")
+
+## Untagged dbh data
+untagged_dbh_2005 <- read.csv("analysis/data/untagged_dbh2005_summary.csv")
+untagged_dbh_2014 <- read.csv("analysis/data/untagged_dbh2014_summary.csv")
+alt05live <- read.csv("analysis/data/altsp05abund_live.csv")
+alt05dead <- read.csv("analysis/data/altsp05abund_dead.csv")
+alt14live <- read.csv("analysis/data/altsp14abund_live.csv")
+alt14dead <- read.csv("analysis/data/altsp14abund_dead.csv")
+#'
+#'
+#' Calculate understory richness
+#+ understory species coverage data from transects
+
+summary(veg_us_2005)# data from vegetation transects 2003, 2004, 2005
+veg_us_2005 <- as.tbl(select(veg_us_2005, -X))
+summary(veg_us_2011)# data from vegetation transects 2011
+veg_us_2011 <- as.tbl(select(veg_us_2011, -X))
+veg_us_2011# 1300 records
+veg_us_2005# 1249 records
+
+# Each record is a species in a plot, so it was observed in at least one sample point.
+
+library(tidyr)
+# 2005 Vegetation Data
+veg_us_2005_wide1 <- select(veg_us_2005, PlotID, UnderstorySp, pcov)
+veg_us_2005_wide1 <- veg_us_2005_wide1 %>% spread(UnderstorySp, pcov)
+summary(veg_us_2005_wide1)
[…]

### 2016-01-07 analysis/species_diversity_calc.R
Edit vegetation transect data.
Change some of the species data, subsetting the understory vegetation transect data.

@@ -8,7 +8,8 @@
 #' The Shannon and Simpson's indices can be calculated using functions in the `vegan` package. I have done some exploring of this package in the 'vegan_diversity.Rmd' script. 
 #' 
 #+ load plot sampling data
-## Ecological data: leaf counts, infection
+
+## Ecological data: leaf counts, infection, at stem and plot levels
 load("~/GitHub/superspreaders/stems_plots.RData")
 
 ## Vegetation transect data
@@ -35,21 +36,26 @@ veg_us_2011 <- as.tbl(select(veg_us_2011, -X))
 veg_us_2011# 1300 records
 veg_us_2005# 1249 records
 
-# Each record is a species in a plot, so it was observed in at least one sample point.
+# Each record is a species in a plot, so it was observed in at least one transect sample point.
 
+#' Transpose vegetation transect data from long to wide format. The wide format preps the data for the `diversity` function in the `vegan` package.
+#' 
+#+ transpose data long to wide
 library(tidyr)
-# 2005 Vegetation Data
+# 2005 vegetation data proportional species coverage
 veg_us_2005_wide1 <- select(veg_us_2005, PlotID, UnderstorySp, pcov)
 veg_us_2005_wide1 <- veg_us_2005_wide1 %>% spread(UnderstorySp, pcov)
 summary(veg_us_2005_wide1)
+# 2005 vegetation data point count species coverage
 veg_us_2005_wide2 <- select(veg_us_2005, PlotID, UnderstorySp, ptct)
 veg_us_2005_wide2 <- veg_us_2005_wide2 %>% spread(UnderstorySp, ptct)
 summary(veg_us_2005_wide2)
 
-# 2011 Vegetation Data
+# 2011 vegetation data proportianal species coverage
 veg_us_2011_wide1 <- select(veg_us_2011, PlotID, UnderstorySp, pcov)
 veg_us_2011_wide1 <- veg_us_2011_wide1 %>% spread(UnderstorySp, pcov)
 summary(veg_us_2011_wide1)
+# 2011 vegetation data point count species coverage
 veg_us_2011_wide2 <- select(veg_us_2011, PlotID, UnderstorySp, ptct)
 veg_us_2011_wide2 <- veg_us_2011_wide2 %>% spread(UnderstorySp, ptct)
 summary(veg_us_2011_wide2)
[…]

### 2016-01-09 analysis/species_diversity_calc.R
Calculate diversity indices.
I calculated Shannon's and Simpson's diversity indices, and Pielou's evenness. I put the diversity metrics and the straight species richness counts into one data frame and exported it to a CSV file 'os_us_diversity'.

@@ -7,7 +7,7 @@
 #' 
 #' The Shannon and Simpson's indices can be calculated using functions in the `vegan` package. I have done some exploring of this package in the 'vegan_diversity.Rmd' script. 
 #' 
-#+ load plot sampling data
+#+ load plot sampling data ####
 
 ## Ecological data: leaf counts, infection, at stem and plot levels
 load("~/GitHub/superspreaders/stems_plots.RData")
@@ -27,7 +27,7 @@ alt14dead <- read.csv("analysis/data/altsp14abund_dead.csv")
 #'
 #'
 #' Calculate understory richness
-#+ understory species coverage data from transects
+#+ understory species coverage data from transects ####
 
 summary(veg_us_2005)# data from vegetation transects 2003, 2004, 2005
 veg_us_2005 <- as.tbl(select(veg_us_2005, -X))
@@ -38,28 +38,7 @@ veg_us_2005# 1249 records
 
 # Each record is a species in a plot, so it was observed in at least one transect sample point.
 
-#' Transpose vegetation transect data from long to wide format. The wide format preps the data for the `diversity` function in the `vegan` package.
-#' 
-#+ transpose data long to wide
 library(tidyr)
-# 2005 vegetation data proportional species coverage
-veg_us_2005_wide1 <- select(veg_us_2005, PlotID, UnderstorySp, pcov)
-veg_us_2005_wide1 <- veg_us_2005_wide1 %>% spread(UnderstorySp, pcov)
-summary(veg_us_2005_wide1)
-# 2005 vegetation data point count species coverage
-veg_us_2005_wide2 <- select(veg_us_2005, PlotID, UnderstorySp, ptct)
-veg_us_2005_wide2 <- veg_us_2005_wide2 %>% spread(UnderstorySp, ptct)
-summary(veg_us_2005_wide2)
-
-# 2011 vegetation data proportianal species coverage
-veg_us_2011_wide1 <- select(veg_us_2011, PlotID, UnderstorySp, pcov)
-veg_us_2011_wide1 <- veg_us_2011_wide1 %>% spread(UnderstorySp, pcov)
-summary(veg_us_2011_wide1)
-# 2011 vegetation data point count species coverage
-veg_us_2011_wide2 <- select(veg_us_2011, PlotID, UnderstorySp, ptct)
-veg_us_2011_wide2 <- veg_us_2011_wide2 %>% spread(UnderstorySp, ptct)
[…]

### 2016-01-09 explore_veg_data.Rmd
Changes to investigation lines of code.

@@
 # But there are a lot more records in the 2004 data, which is closer to the 2012 & 2014 sampling. I could combine the 2004 & 2005 data, including the potentially duplicated COOK02 data.
-summary(filter(alt_sp_2012, PlotID=="COOK02"))
-filter(alt_sp_2014, PlotID=="COOK02")
+# summary(filter(alt_sp_2012, PlotID=="COOK02"))
+# filter(alt_sp_2014, PlotID=="COOK02")
 
@@
 summary(alt_sp_2014)
+alt_sp_2014$year <- 2014
 summary(alt_sp_2005)
@@
 alt_sp_2014_dead <- select(alt_sp_2014_dead, -Dead)
-alt_sp_2005_dead <- filter(alt_sp_2005, Dead == 1)
+alt_sp_2005_dead <- filter(alt_sp_2005, Dead == 1)# only 12 obs, all at GANAY02
 alt_sp_2005_dead <- select(alt_sp_2005_dead, -Dead)

### 2016-01-09 sem_development.Rmd
Remove richness/diversity code to other script.
I also revised names of the rainfall and temperature data objects.

@@
 load("~/GitHub/superspreaders/stems_plots.RData")
-# Vegetation data
+
+# Vegetation transect data
 veg_us_2005 <- read.csv("analysis/data/understory_veg_2005.csv")
 veg_us_2011 <- read.csv("analysis/data/understory_veg_2011.csv")
+
 # Untagged dbh data
-untagged_dbh <- read.csv("analysis/data/untagged_dbh_summary.csv")
+untagged_dbh_2005 <- read.csv("analysis/data/untagged_dbh2005_summary.csv")
+untagged_dbh_2014 <- read.csv("analysis/data/untagged_dbh2014_summary.csv")
 canopy_cover <- read.csv("analysis/data/plot_canopy_cover_dens.csv")
@@
 rm(plots_vegtopo)
+
 ```
@@
 ```{r examine temperature & rain data}
-summary(plots_temps)
-summary(rain_days)
-summary(rain_total)
-unique(plots_temps$plotid)# 201 levels
-unique(rain_days$plotid)# 202 levels
-unique(rain_total$plotid)# 202 levels
-tempsid <- as.character(unique(plots_temps$plotid))
-rainid <- as.character(unique(rain_days$plotid))
+summary(rs_temps)
+summary(summer_temps)
+
+summary(ppt_days)
+summary(ppt_total)
+rain_df <- merge(ppt_days@data, ppt_total@data)
+summary(rain_df)
+
+unique(rs_temps$plotid)# 201 levels
+unique(summer_temps$plotid)# 201 levels
+unique(rain_df$plotid)# 202 levels
+
+tempsid <- as.character(unique(rs_temps$plotid))
+rainid <- as.character(unique(rain_df$plotid))
 rainid %in% tempsid
 rainid[62]# BUSH01 still in the rain data
-rain_days <- filter(rain_days, plotid != "BUSH01")
-rain_total <- filter(rain_total, plotid != "BUSH01")
+rain_df <- filter(rain_df, plotid != "BUSH01")
+
 # Temperature and rainfall data now have same number of observations
@@
 # plots_env <- filter(plots_env, plotid != "BUSH01")
+stems <- as.tbl(stems)
+stems
+stems <- select(stems, -X) 
 stems <- rename(stems, plotid = plot) %>% filter(plotid != "BUSH01")
 
-plots_temps %>% filter(is.na(hrs_abv25))
-# I am replacing the NAs with `0`, assuming true observations.
-plots_temps[is.na(plots_temps)] <- 0
+rs_temps %>% filter(is.na(hrs_abv25))
+# I am replacing the NAs with 0, assuming true observations, i.e. there were zero hours where temperatures were above 25C
+rs_temps[is.na(rs_temps)] <- 0
 ```
@@
 write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
-write.csv(rain_data, "analysis/data/rain_data_long.csv", row.names = F)
-write.csv(plots_temps, "analysis/data/temp_data_long.csv", row.names = F)
+write.csv(rain_df, "analysis/data/rain_data_long.csv", row.names = F)
+write.csv(rs_temps, "analysis/data/rs_temps_data.csv", row.names = F)
+write.csv(summer_temps, "analysis/data/summer_temps_data.csv", row.names = F)
 write.csv(stems, "analysis/data/stems_noBUSH01.csv", row.names = F)
@@
 
-#### Reload data that has BUSH01 excluded
+### Reload data that has BUSH01 excluded
 ```{r reload data}
 plots_env <- read.csv("analysis/data/env_gis_variables.csv")
-rain_data <- read.csv("analysis/data/rain_data_long.csv")
-temps_data <- read.csv("analysis/data/temp_data_long.csv")
+rain_df <- read.csv("analysis/data/rain_data_long.csv")
+rs_temps <- read.csv("analysis/data/rs_temps_data.csv")
+summer_temps <- read.csv("analysis/data/summer_temps_data.csv")
 stems <- read.csv("analysis/data/stems_noBUSH01.csv")
@@
 
-#### Plot contributing area vs. slope
+### Plot contributing area vs. slope
 This is a suggestion from Ryan to plot the contributing area vs.
@@
 # par(mfrow = c(3,1))
-qplot(slope15m, flowacc15m, data = plots_env) +
+names(plots_env)
+qplot(flowacc15m, facc_ws15m, data = plots_env) #+
       geom_text(aes(label = plotid))
+
+qplot(slope15m, flowacc15m, data = plots_env) #+
+      geom_text(aes(label = plotid))
+
 qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m < 20), 
-      main = "Flow Accumulation < 20") +
+      main = "Flow Accumulation < 20") #+
       geom_text(aes(label = plotid))
-qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 20 & flowacc15m < 1000), main = "Flow Accumulation 20 - 1000") +
+
+qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 20 & flowacc15m < 1000), main = "Flow Accumulation 20 - 1000") #+
       geom_text(aes(label = plotid))
+
 qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 1000), 
-      main = "Flow Accumulation > 1000") +
+      main = "Flow Accumulation > 1000") #+
       geom_text(aes(label = plotid))
 
+# I think the 'facc_ws15m' is the estimate from the 'r.watershed' function in GRASS GIS
 qplot(slope15m, facc_ws15m, data = plots_env)
@@
 # This only has number of infected, uninfected, and total, which must apply to living stems only. I want to analyze leaf counts.
+
 unique(plots_umca$plot)
@@
 
-I can use this to calculate plot level leaf count variables (sum, average) for each plot and year
+I can use this to calculate plot level leaf count variables (sum, average) for each plot and year.
 ```{r calculate plot level leaf counts}
-umca_plots <- left_join(umca_plots, bay_laurel %>% filter(status == "Alive") %>% group_by(plot, year) %>%  summarise(
-            avg_lfct = mean(slc, na.rm = T),
+umca_plots <- left_join(
+      umca_plots, bay_laurel %>% filter(status == "Alive") %>% 
+            group_by(plot, year) %>% 
+            summarise(avg_lfct = mean(slc, na.rm = T),
                       tot_lfct = sum(slc, na.rm = T)),
@@
 
+Explore the NAs in the UMCA leaf counts.
 ```{r examine lfct NAs}
@@
 filter(bay_laurel, plot == "YAHNG02", status == "Alive")
-# I'm removing it manually from the `plots_umca` data frame
+# I'm removing it manually from the `umca_plots` data frame
 umca_plots <- umca_plots %>% filter(avg_lfct != "NaN")
@@
 
-So now I have a plot-level of UMCA infection levels for each year. I want to join this with some of the environmental data.
+So now I have a plot-level of UMCA infection levels for each year in terms of symptomatic leaf counts and number of un/infected stems.
+I want to join this with some of the plot-level environmental data.
 ```{r join env and umca data}
 summary(plots_env)
-plots_env$avg_psi <- rowMeans(cbind(plots_env$psiSpr, plots_env$psiSum, plots_env$psiWin))
-summary(plots_umca)
+
+# Calculate a mean potential solar radiation
+plots_env$avg_psi <- rowMeans(
+      cbind(plots_env$psiSpr, plots_env$psiSum, plots_env$psiWin))
+summary(umca_plots)
 # plot == plotid, year == sample_year
-plots_umca <- left_join(select(plots_umca, -species), plots_env, by = c("plot" = "plotid"))
-summary(plots_umca)
-filter(plots_umca, is.na(easting))
-# NAs due to no topo data for abandoned SUGAR02
-plots_umca <- na.omit(plots_umca)
-
-plots_temps <- as.tbl(plots_temps)
-plots_temps <- plots_temps %>% rename(year = sample_year)
-plots_temps$plotid <- tolower(plots_temps$plotid)
-
-plots_umca <- left_join(plots_umca, plots_temps, by = c("plotid", "year"))
-```
-
-#### Community vegetation data and variables
-Now I also want to calculate some plot-level community variables from the vegetation data. First, understory richness from the vegetation transects.
-
-```{r calculate understory richness}
-summary(veg_us_2005)
-veg_us_2005 <- as.tbl(select(veg_us_2005, -X))
-summary(veg_us_2011)
-veg_us_2011 <- as.tbl(select(veg_us_2011, -X))
-veg_us_2011
-veg_us_2005
-
-veg <- rbind(veg_us_2005, veg_us_2011)
-unique(veg$year)
-veg$year[veg$year < 2011] <- 2005 # Creates the two most complete veg sampling
-veg <- veg %>% rename(us_species = UnderstorySp, plotid = PlotID)
-unique(veg$us_species)
-
-# I want to match the species to those used by Sarah in her second chapter
-local_woody <- read.csv("analysis/data/presence_local_woody_n197.csv")
-summary(local_woody)
-unique(local_woody$spp.local)
-woody_spp <- as.character(local_woody$spp.local)
-woody_spp[woody_spp == "prunus"] <- "prunus sp."
-sort(woody_spp) # only 29 species
-sort(unique(veg$us_species))
-# It looks like I will want to add a few more to her list
-
-filter(veg, us_species == "ribes sp.")
-filter(veg, us_species == "ribes")
-veg$us_species[veg$us_species == "ribes"] <- "ribes sp."
-
-woody_spp <- c(woody_spp, "ceanothus sp.", "cebe", "cecu", "frla", "garrya sp.", "gemo", "juca", "mafu", "mane", "mimulus sp", "pisa", "quag x quke", "qube", "quercus sp.", "quwi", "rhamnus sp.", "rhcr", "ribes sp.", "rica", "risa", "roca", "rosa sp.", "rubus sp.", "rudi", "salix sp.", "syal", "vaccinium sp.")
-veg_sub <- filter(veg, veg$us_species %in% woody_spp)
-veg_sub <- droplevels(veg_sub)
-
-richness <- veg_sub %>%
-      group_by(plotid, year) %>% 
-      summarise(us_rich = length(unique(us_species)))
-richness
-```
-
-Next, overstory richness from the DBH data for tagged & untagged stems.
-
-```{r calculate overstory richness}
-summary(untagged_dbh)
-untagged_dbh <- as.tbl(untagged_dbh %>% select(-X) %>% rename(plotid = PlotID, species = Species))
-untagged_dbh
-unique(untagged_dbh$year)
-untagged_dbh$year[untagged_dbh$year < 2014] <- 2005 # Create two most complete groups
-untagged_dbh$plotid <- tolower(untagged_dbh$plotid)
-unique(untagged_dbh$plotid)
-untagged_dbh <- filter(untagged_dbh, plotid != "bush01")
-untagged_dbh %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species)))
-
-stems <- as.tbl(stems)
-summary(stems)
-unique(stems$status)
-stems <- stems %>% select(-X) %>% filter(status == "Alive" | status == "Dead")
-unique(stems$species)
-unique(stems$plotid)
-stems$species <- tolower(stems$species)
-stems$species[stems$species == "quke x quag"] <- "quag x quke"
-stems$plotid <- tolower(stems$plotid)
-stems <- droplevels(stems)
-
-live_tagged <- stems %>% filter(status == "Alive") %>% group_by(plotid, year, species) %>% summarise(live_count = length(species))
-unique(live_tagged$species)
-os_richness <- rbind(live_tagged, select(untagged_dbh, plotid, species, year, live_count))
-
-os_richness <- left_join(
-      stems %>% filter(year == 2005 | year == 2014) %>% 
-            group_by(plotid, year) %>% 
-            summarise(tagged_rich = length(unique(species))),
-      untagged_dbh %>% 
-            group_by(plotid, year) %>% 
-            summarise(untagged_rich = length(unique(species))), by = c("plotid", "year")) %>% 
-      mutate(os_rich = rowSums(cbind(tagged_rich,untagged_rich), na.rm = T))
-os_richness
-
-richness <- left_join(os_richness, richness, by = c("plotid","year"))
-richness
-summary(richness)
+
+# Join the environmental data to the bay laurel infection data
+umca_plots <- left_join(select(umca_plots, -species), plots_env, by = c("plot" = "plotid"))
+summary(umca_plots)
+filter(umca_plots, is.na(lon_utm))
+# NAs due to no topo data for abandoned SUGAR02; remove SUGAR02
+umca_plots <- na.omit(umca_plots)
+umca_plots <- rename(umca_plots, plotid = plot, sample_year = year)
+umca_plots$plotid <- tolower(umca_plots$plotid)
+
+# Join rainy season and summer temperatures to bay laurel infection data
+rs_temps <- as.tbl(rs_temps)
+rs_temps
+rs_temps$plotid <- tolower(rs_temps$plotid)
+
+summer_temps <- as.tbl(summer_temps)
+summer_temps
+summer_temps$plotid <- tolower(summer_temps$plotid)
+
+umca_plots <- left_join(umca_plots, rs_temps, by = c("plotid", "sample_year"))
+umca_plots <- left_join(umca_plots, summer_temps, by = c("plotid", "sample_year"))
+umca_plots
 ```

### 2016-01-11 sem_development.Rmd
Munge data to create plot-level data frames.
Changed names on temperature and rainfall data to maintain as unique variables. Joined environmental and ecological data at the plot level - completed for plots with oak trees.

@@
 
-Examine temperature data, which is in long format and covers sample years 2004 through 2014. This means that I will need to exclude disease-related sampling from 2003.
+Examine temperature data, which is in long format and covers sample years 2004 through 2014.
+This means that I will need to exclude disease-related sampling from 2003 of the ecological data (NAs should fill-in if missing data when joining).
 ```{r examine temperature & rain data}
@@
 
-### Reload data that has BUSH01 excluded
+### Starting Point with Revised Data: Reload data that has BUSH01 excluded
+Still need to load the .RData source at the beginning of this document, line 32.
 ```{r reload data}
@@
 
-unique(plots_umca$plot)
-umca_plots <- filter(plots_umca, plot != "BUSH01")
+unique(plots_umca$plot)# 185 plots with bay laurel
+umca_plots <- filter(plots_umca, plot != "BUSH01")# now 184 plots
 rm(plots_umca)
@@
 
-# Join rainy season and summer temperatures to bay laurel infection data
+# Join rainy season and summer temperatures to bay laurel infection data.
+## I need to rename temperature variables to differentiate rainy vs. dry season
 rs_temps <- as.tbl(rs_temps)
-rs_temps
 rs_temps$plotid <- tolower(rs_temps$plotid)
+rs_temps <- rename(rs_temps, 
+                   hrs_14_20_rs = hrs_14_20, hrs_abv25rs = hrs_abv25,
+                   hrs_abv20_rs = hrs_abv20, hrs_blw10rs = hrs_blw10,
+                   avg_tmax_rs = avg_tmax, avg_tmin_rs = avg_tmin)
 
@@
 summer_temps$plotid <- tolower(summer_temps$plotid)
-
+summer_temps <- rename(summer_temps, 
+                       hrs_abv25ds = hrs_abv25, hrs_abv20ds = hrs_abv20,
+                       hrs_blw18ds = hrs_blw18, avg_tmax_ds = avg_tmax,
+                       avg_tmin_ds = avg_tmin)
+
+# Now I can join the data without duplicating variable names
 umca_plots <- left_join(umca_plots, rs_temps, by = c("plotid", "sample_year"))
@@
 
+### Load and Join Richness, Diversity, Evenness Data
 Now I want to join the richness data set and the canopy cover estimates to the `umca_plots` data frame that currently has the environmental and leaf count/infection variables.
@@
 ```{r join richness and umca data}
+os.us.diversity <- as.tbl(read.csv("analysis/data/os_us_diversity.csv"))
+os.us.diversity
 umca_plots
-richness
-umca_plots <- rename(umca_plots, plotid = plot)
-umca_plots$plotid <- tolower(umca_plots$plotid)
-
-umca_plots <- left_join(umca_plots, richness, by = c("plotid", "year"))
+
+umca_plots <- left_join(umca_plots, os.us.diversity, by = c("plotid"))
 head(canopy_cover)
 canopy_cover <- select(canopy_cover, -X)
-umca_plots <- left_join(umca_plots, canopy_cover, by = c("plotid", "year"))
+canopy_cover <- rename(canopy_cover, sample_year = year)
+unique(filter(canopy_cover, year == 2006)$plotid)
+
+umca_plots <- left_join(umca_plots, canopy_cover, by = c("plotid", "sample_year"))
 ```
@@
 summary(qusp_plots)
-qusp_plots
+
 oak_sod <- qusp_plots %>% group_by(plot, year) %>%
@@
 oak_sod
-oak_sod <- rename(oak_sod, plotid = plot)
+oak_sod <- rename(oak_sod, plotid = plot, sample_year = year)
 oak_sod$plotid <- tolower(oak_sod$plotid)
 
-oak_sod <- left_join(oak_sod, select(umca_plots, plotid, year, uninfected_bay_ct, infected_bay_ct, tot_bay, avg_lfct, tot_lfct), 
-                     by = c("plotid", "year"))
+unique(umca_plots$plotid) %in% unique(oak_sod$plotid)
+
+# Add bay laurel infection data to oak plots
+names(umca_plots)
+oak_sod <- left_join(oak_sod, 
+                     select(umca_plots, plotid, sample_year, uninfected_bay_ct, infected_bay_ct, tot_bay, avg_lfct, tot_lfct), 
+                     by = c("plotid", "sample_year"))
 summary(oak_sod) # the NA's for the bay laurel variables are legitimately zeroes b/c there are no bay in some of these plots
-filter(oak_sod, is.na(tot_bay))
+filter(oak_sod, is.na(tot_bay), sample_year != 2003)
+# Change NAs of bay laurel in oak only plots to zeroes
 oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
@@
 
+# Add environmental data to oak plot infection data.
 plots_env
 plots_env$plotid <- tolower(plots_env$plotid)
-oak_sod <- left_join(oak_sod, plots_env, by = "plotid")
-head(temps_data, 15)
-temps_data$plotid <- tolower(temps_data$plotid)
-temps_data <- rename(temps_data, year = sample_year)
-oak_sod <- left_join(oak_sod, temps_data, by = c("plotid", "year"))
+oak_sod <- left_join(oak_sod, plots_env, by = "plotid")# Topo data
+
+rs_temps
+summer_temps
+oak_sod <- left_join(oak_sod, rs_temps, by = c("plotid", "sample_year"))
+oak_sod <- left_join(oak_sod, summer_temps, by = c("plotid", "sample_year"))
 oak_sod
@@
 
-richness
-oak_sod <- left_join(oak_sod, richness, by = c("plotid", "year"))
+os.us.diversity
+oak_sod <- left_join(oak_sod, os.us.diversity, by = c("plotid"))
 summary(oak_sod)
 filter(oak_sod, is.na(elev15m))
+# NAs are plots that were abandoned early in the study
 oak_sod <- filter(oak_sod, plotid != "duer01", plotid != "ponti01", plotid != "sweet01")
 
-summary(filter(oak_sod, is.na(avg_tmax)))
-filter(oak_sod, is.na(avg_tmax), year == 2005) # BUSH01 plot; I'm dropping that from the data set
+summary(filter(oak_sod, is.na(avg_tmax_rs)))
+# Many of the NAs are due to the 2003 sample year
+oak_sod <- filter(oak_sod, sample_year != 2003)
+filter(oak_sod, is.na(avg_tmax_rs))
+# BUSH01 plot; I'm dropping that from the data set
 oak_sod <- filter(oak_sod, plotid != "bush01")
-filter(oak_sod, is.na(avg_tmax), year == 2004) # PONTI01 and SWEET01 were other early abandoned plots
-oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01")
-summary(filter(oak_sod, is.na(avg_tmax))) # Now the NAs in the temperature variables are just for year 2003 when we had severely incomplete temperature data
 summary(oak_sod)
-filter(oak_sod, is.na(easting))
-oak_sod <- filter(oak_sod, plotid != "duer01")
-
-filter(oak_sod, plotid == "ann05") # There is no 2007 record for ann05 in the UMCA form, but there is a plot visit. I will need to come back and investigate this further for correcting the database.
-oak_sod$infected_bay_ct[oak_sod$plotid == "ann05" & oak_sod$year == 2007] <- 15
-oak_sod$tot_bay[oak_sod$plotid == "ann05" & oak_sod$year == 2007] <- 15
-# I filled in the counts based on the values in the immediately preceding and following years, which I think is pretty solid reasoning. However, there is nothing I can do about the leaf counts.
+
+oak_sod <- left_join(oak_sod, canopy_cover, by = c("plotid", "sample_year"))
+
+summary(rain_total)
+rain_total$plotid <- tolower(rain_total$plotid)
+rain_total <- rename(rain_total, sample_year = year)
+# rain_total$sample_year <- as.numeric(rain_total$sample_year)
+summary(rain_days)
+rain_days$plotid <- tolower(rain_days$plotid)
+rain_days <- rename(rain_days, sample_year = year)
+#rain_days$sample_year <- as.numeric(as.character(rain_days$sample_year))
+
+oak_sod <- left_join(oak_sod, rain_total, by = c("plotid", "sample_year"))
+oak_sod <- left_join(oak_sod, rain_days, by = c("plotid", "sample_year"))
 summary(oak_sod)
 
-oak_sod <- left_join(oak_sod, canopy_cover, by = c("plotid", "year"))
-summary(rain_data)
-rain_data$plotid <- tolower(rain_data$plotid)
-oak_sod <- left_join(oak_sod, rain_data, by = c("plotid", "year"))
-summary(filter(oak_sod, is.na(avg_tmax))) # all obs from 2003
-
-filter(oak_sod, year == 2003) # 85 records
-filter(oak_sod, year == 2004) # 88 records
+summary(filter(oak_sod, is.na(rain_tot_psm))) # all obs after 2010
+
+filter(oak_sod, sample_year == 2004) # 88 records
 # 2005 appears to be the first full year of observing all the plots
-filter(oak_sod, year == 2005) # 170 records
-filter(oak_sod, year == 2006) # 172 records
-filter(oak_sod, year == 2007) # 171 records
-filter(oak_sod, year == 2008) # 171 records
-filter(oak_sod, year == 2009) # 170 records
-filter(oak_sod, year == 2010) # 170 records
-filter(oak_sod, year == 2011) # 170 records; final year for Sarah's 2nd chapter
-filter(oak_sod, year == 2012) # 169 records
-filter(oak_sod, year == 2014) # 168 records
+filter(oak_sod, sample_year == 2005) # 170 records
+filter(oak_sod, sample_year == 2006) # 172 records
+filter(oak_sod, sample_year == 2007) # 171 records
+filter(oak_sod, sample_year == 2008) # 171 records
+filter(oak_sod, sample_year == 2009) # 170 records
+filter(oak_sod, sample_year == 2010) # 170 records
+filter(oak_sod, sample_year == 2011) # 170 records; final year for Sarah's 2nd chapter
+filter(oak_sod, sample_year == 2012) # 169 records
+filter(oak_sod, sample_year == 2014) # 168 records
 # In Sarah's analysis we end up using only 163 plots

### 2016-01-12 sem_development.Rmd
Add rainfall data to UMCA plots data frame.

@@
 umca_plots <- left_join(umca_plots, summer_temps, by = c("plotid", "sample_year"))
-umca_plots
+names(umca_plots)
+
+# Add rainfall data to UMCA plots data frame
+names(rain_total)
+names(rain_days)
+umca_plots <- left_join(umca_plots, rain_total, by = c("plotid", "sample_year"))
+umca_plots <- left_join(umca_plots, rain_days, by = c("plotid", "sample_year"))
+summary(umca_plots)
+
+filter(umca_plots, is.na(rain_tot_2d))
+# NAs in rainfall data are from 2003
+filter(umca_plots, sample_year == 2003)
+# Only 7 plots w/UMCA sampled in 2003, so removing these records
+umca_plots <- filter(umca_plots, sample_year != 2003)
 ```

### 2016-01-12 sem__data_development.Rmd
Rename file to 'sem_data_development'.
Since the majority of this file is actually data munging I removed the code of my initial path model develoment efforts to another script. This compartmentalization will hopefully make it easier for  any necessary redoing of data prep.

--- 2016-01-12 sem_development.Rmd
@@
 ---
-title: "Path Analysis - Chapter 2"
+title: "Path Analysis Data Development - Chapter 2"
 author: "Whalen Dillon"
@@
 
-Now to do a pairs plot of the `oak_sod` data frame
+Now to do a pairs plot of the variables from the `oak_sod` data frame.
 
@@
 oak_sod <- ungroup(oak_sod)
-pairs(select(oak_sod, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist, main = "Oak_SOD Data")
+pairs(select(oak_sod, elev_m, starts_with("hrs"), starts_with("avg_t"), contains("rain_tot")), lower.panel = panel.cor, diag.panel = panel.hist, main = "Oak-SOD Rainfall & Temperature Data")
 ```
 
-# Path Analysis Presented at 2015 IALE World Congress
-
-#### Path Analysis Plot-Level Oak Infection 2005
-Here I am using the `oak_sod` data set which will be the oak infection (binomial) as the terminal response variable.
-
-```{r oak_sod path model 2005}
-library(ggm)
-names(oak_sod)
-oak_path1 <- DAG(
-      twi15m ~ elev15m,
-      os_rich ~ twi15m,
-      hrs_14_20 ~ elev15m + d2c + avg_psi + twi15m,
-      tot_lfct ~ hrs_14_20 + os_rich + d2c,
-      inf_oak_ct ~ tot_lfct + os_rich + hrs_14_20 + d2c
-      )
-isAcyclic(oak_path1)
-bu_oak_path1 <- basiSet(oak_path1)
-bu_oak_path1
-
-oak_path2 <- DAG(
-      twi15m ~ elev15m,
-      os_rich ~ twi15m,
-      hrs_blw10 ~ elev15m + d2c + avg_psi + twi15m,
-      tot_lfct ~ hrs_blw10 + os_rich + d2c,
-      inf_oak_ct ~ tot_lfct + os_rich + hrs_blw10 + d2c
-      )
-isAcyclic(oak_path2)
-
-oak_2005 <- filter(oak_sod, year == 2005)
-summary(oak_2005)
-oak_2005 <- select(oak_2005, plotid, uninf_oak_ct, inf_oak_ct, tot_lfct, hrs_14_20, hrs_blw10, hrs_abv25, avg_tmax, avg_tmin, rain05_2d, rain05_v, elev15m, twi15m, avg_psi, os_rich, d2c)
-oak_2005$tot_lfct[is.na(oak_2005$tot_lfct)] <- 0
-filter(oak_2005, is.na(os_rich))
-# oak_2005 <- na.omit(oak_2005)
-oak_2005$plotid <- as.factor(oak_2005$plotid)
-
-# This model makes Gaussian assumptions, which should be incorrect for these data
-shipley.test(oak_path1, cov(select(oak_2005, elev15m, twi15m, os_rich, hrs_14_20, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 170)
-# The p-value of 0.08 indicates that there is not a severe violation of the conditional independence assumptions with the application of a Gaussian model
-
-shipley.test(oak_path2, cov(select(oak_2005, elev15m, twi15m, os_rich, hrs_blw10, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 170)
-# The p-value of .04 indicates that the conditional independence assumptions under the application of a Gaussian model are questionable.
+### Save patha analysis data environment
+```{r save data environment}
+save.image("~/GitHub/superspreaders/pathmodel_data_20160112.RData")
 ```
-
-Violation or questioning of conditional independence supports addressing the complexity of the data using different model distributions to deal with the data.
-
-```{r assess conditional independence claims 2005}
-bu_oak_path1 # 15 independence claims means 15 models
-library(nlme); library(lme4); library(MuMIn)
-
-f1 <- lme(avg_psi ~ d2c, data = oak_2005, random = ~1|plotid)
-summary(f1) # p = 0.5091
-f2 <- lme(avg_psi ~ elev15m, data = oak_2005, random = ~1|plotid)
-summary(f2) # p = 0.2013
-f3 <- lme(avg_psi ~ twi15m + elev15m, data = oak_2005, random = ~1|plotid)
-summary(f3) # twi15m; p = 0.0959
-f4 <- lme(os_rich ~ avg_psi + twi15m, data = oak_2005, random = ~1|plotid)
-summary(f4) # avg_psi; p = 0.1712
-f5 <- lm(log1p(tot_lfct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
-# f5 <- lme(log1p(tot_lfct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, random = ~1|plotid)
-# f5 <- glmer(tot_lfct ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2005, family = "poisson")
-summary(f5) # avg_psi; p = 0.5993
-# f6 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "binomial")
-# f6 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "quasibinomial")
-f6 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
-summary(f6) # avg_psi; p = 0.08970
-f7 <- lm(elev15m ~ d2c, data = oak_2005)
-# f7 <- lme(elev15m ~ d2c, data = oak_2005, random = ~1|plotid)
-summary(f7) # p = 0.03184
-f8 <- lm(log(twi15m) ~ d2c + elev15m, data = oak_2005)
-# f8 <- lme(log(twi15m) ~ d2c + elev15m, data = oak_2005, random = ~1|plotid)
-summary(f8) # d2c; p = 0.3693
-f9 <- lm(os_rich ~ d2c + twi15m, data = oak_2005)
-# f9 <- lme(os_rich ~ d2c + twi15m, data = oak_2005, random = ~1|plotid)
-summary(f9) # d2c; p = 0.9370
-f10 <- lm(os_rich ~ elev15m + twi15m, data = oak_2005)
-# f10 <- lme(os_rich ~ elev15m + twi15m, data = oak_2005, random = ~1|plotid)
-summary(f10) # elev15m; p = 0.19174
-
-# f11 <- lm(tot_lfct ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
-# f11 <- lm(log1p(tot_lfct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
-f11 <- glm(tot_lfct ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, family = "poisson")
-chat <- deviance(f11)/df.residual(f11)
-library(MuMIn)
-x.quasipoisson<-function(...){
-  res<-quasipoisson(...)
-  res$aic<-poisson(...)$aic
-  res
-}
-QAIC(update(f11, family = x.quasipoisson), chat = chat)
-f11 <- glm(tot_lfct ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, family = "quasipoisson")
-# f11 <- lme(log1p(tot_lfct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), random = ~1|plotid, data = oak_2005)
-# f11 <- glmer(tot_lfct ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2005, family = "poisson")
-summary(f11) # elev15m; p = 0.360630
-
-# f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "binomial")
-# chat <- deviance(f12)/df.residual(f12)
-# QAIC(update(f12, family = x.quasibinomial), chat = chat)
-# f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "quasibinomial")
-f12 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elev15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
-summary(f12) # elevation; p = 0.44861
-
-f13 <- lm(log1p(tot_lfct) ~ scale(twi15m) + scale(d2c) + scale(elev15m) + scale(hrs_14_20) + scale(os_rich), data = oak_2005)
-# f13 <- lme(log1p(tot_lfct) ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2005, random = ~1|plotid)
-# f13 <- glmer(tot_lfct ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2005, family = "poisson")
-summary(f13) # twi15m; p = 0.801
-
-# f14 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "binomial")
-# chat <- deviance(f14)/df.residual(f14)
-# QAIC(update(f14, family = x.quasibinomial), chat = chat)
-# f14 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2005, family = "quasibinomial")
-
-f14 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(twi15m) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2005, family = "binomial")
-summary(f14) # twi15m; p = 0.61177
-
-f15 <- lm(os_rich ~ hrs_14_20 + avg_psi + d2c + elev15m + twi15m, data = oak_2005)
-# f15 <- lme(os_rich ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2005, random = ~1|plotid)
-summary(f15) # hrs_14_20; p = 0.08468
-
-# Calculate C-statistic
-pvalues <- c(summary(f1)$tTable[2,5], summary(f2)$tTable[2,5], summary(f3)$tTable[2,5], summary(f4)$tTable[2,5], summary(f5)$coefficients[2,4], summary(f6)$coefficients[2,4], summary(f7)$tTable[2,5], summary(f8)$tTable[2,5], summary(f9)$tTable[2,5], summary(f10)$tTable[2,5], summary(f11)$coefficients[2,4], summary(f12)$coefficients[2,4], summary(f13)$coefficients[2,4], summary(f14)$coefficients[2,4], summary(f15)$tTable[2,5])
-
-cstat <- -2 * sum(log(pvalues))
-cstat
-```
-
-Given the C-statistic of 26.9 and that 2k degrees of freedom would be 30 I think that this meets the criteria for supporting the conditional independence claims.
-So, this supports my path model, now I need to run the models to produce estimates for the path coefficients.
-This is the model set that I need to get the estimated path coefficients:
-
-      1. os_rich ~ twi15m,
-      2. hrs_14_20 ~ elev15m + d2c + avg_psi + twi15m,
-      3. tot_lfct ~ hrs_14_20 + os_rich + d2c,
-      4. infected oak ~ tot_lfct + os_rich + hrs_14_20 + d2c
-
-```{r estimate path coefficients 2005}
-# f0 <- fitted(mlme1, level = 0)# Values from population model
-# f1 <- fitted(mlme1, level = 1)# Within beach fitted values
-
-# fit1 <- lm(tmi ~ scale(elevation), data = oak_2005)
-fit1 <- lm(log(tmi) ~ scale(elevation), data = oak_2005) # best model
-# fit1 <- lme(log(tmi) ~ scale(elevation), data = oak_2005, random = ~1|plotid, method = "ML")
-summary(fit1)
-par(mfrow = c(2,2))
-plot(fit1)
-
-# fit2 <- lm(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2005)
-# fit2 <- lme(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2005, random = ~1|plotid) # best model
-fit2 <- glmer(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi) + (1|plotid), data = oak_2005, family = "poisson")
-# fit2 <- glmer.nb(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi) + (1|plotid), data = oak_2005)
-summary(fit2)
-plot(fit2)
-
-# fit3 <- lm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c), data = oak_2005)
-fit3 <- lm(log1p(tot_lfct) ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c), data = oak_2005)
-fit3 <- glm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c), data = oak_2005, family = "poisson")
-(chat <- deviance(fit3) / df.residual(fit3))
-x.quasipoisson<-function(...){
-  res<-quasipoisson(...)
-  res$aic<-poisson(...)$aic
-  res
-}
-library(MuMIn)
-QAIC(update(fit3, family = x.quasipoisson), chat = chat) # 178.3
-fit3 <- glm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c), data = oak_2005, family = "quasipoisson") # best model
-# fit3 <- glmer(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + (1|plotid), data = oak_2005, family = "poisson")
-summary(fit3)
-
-# fit4 <- lm((inf_oak_ct/(inf_oak_ct+uninf_oak_ct)) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c), data = oak_2005)
-# par(mfrow = c(2,2))
-# plot(fit4)
-
-# fit4 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c), data = oak_2005, family = "binomial")
-# (chat <- deviance(fit4) / df.residual(fit4))
-# x.quasibinomial<-function(...){
-#   res<-quasibinomial(...)
-#   res$aic<-binomial(...)$aic
-#   res
-# }
-# library(MuMIn)
-# QAIC(update(fit4, family = x.quasibinomial), chat = chat)
-# fit4 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c), data = oak_2005, family = "quasibinomial")
-# summary(fit4)
-
-fit4 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(d2c) + (1|plotid), data = oak_2005, family = "binomial") # best model
-summary(fit4)
-
-fit5 <- lm(os_rich ~ tmi, data = oak_2005); AIC(fit5)
-# fit5 <- lme(os_rich ~ tmi, data = oak_2005, random = ~1|plotid)
-summary(fit5)
-
-# Trying to plot the fitted model by each group...
-# f0 <- fitted(fit4, level = 0)
-# f1 <- fitted(fit4, level = 1)
-# 
-# f0[I]
-# 
-# I <- order(oak_2005$tot_lfct) ; lfct <- sort(oak_2005$tot_lfct)
-# plot(lfct, f0[I], lwd = 4, type = "l",
-#      ylab = "oak infection", xlab = "total leaf count")
-# for(i in 1:length(f1)){
-#       x1 <- oak_2005$hrs_14_20[oak_2005$plotid == i]
-#       y1 <- f1[oak_2005$hrs_14_20 == i]
-#       K <- order(x1)
-#       lines(sort(x1), y1[K], col = "red")
-# }
-# text(oak_2005$hrs_14_20, oak_2005$tot_lfct, oak_2005$plotid, cex = 0.9)
-```
-
-#### Path Analysis Oak Infection - 2014
-Now I am going to fit the same path model to data from 2011.
-
-```{r oak_sod path model 2014}
-oak_path1 <- DAG(
-      tmi ~ elevation,
-      os_rich ~ tmi,
-      hrs_14_20 ~ elevation + d2c + avg_psi + tmi,
-      tot_lfct ~ hrs_14_20 + os_rich + d2c,
-      inf_oak_ct ~ tot_lfct + os_rich + hrs_14_20 + d2c
-      )
-
-oak_2014 <- filter(oak_sod, year == 2014)
-oak_2014 %>% select(plotid, tagged_rich, untagged_rich, os_rich) %>%
-      filter(is.na(os_rich))
-summary(oak_2014)
-oak_2014 <- select(oak_2014, plotid, uninf_oak_ct, inf_oak_ct, tot_lfct, hrs_14_20, hrs_blw10, hrs_abv25, elevation, tmi, avg_psi, os_rich, d2c)
-oak_2014$tot_lfct[is.na(oak_2014$tot_lfct)] <- 0
-oak_2014 <- na.omit(oak_2014)
-oak_2014$plotid <- as.factor(oak_2014$plotid)
-
-shipley.test(oak_path1, cov(select(oak_2014, elevation, tmi, os_rich, hrs_14_20, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 168)
-# Strongly significant p-value
-
-# shipley.test(DAG(
-#       tmi ~ elevation,
-#       os_rich ~ tmi,
-#       hrs_blw10 ~ elevation + d2c + avg_psi + tmi,
-#       tot_lfct ~ hrs_blw10 + os_rich + d2c,
-#       inf_oak_ct ~ tot_lfct + os_rich + hrs_blw10 + d2c), 
-#       cov(select(oak_2014, elevation, tmi, os_rich, hrs_blw10, d2c, avg_psi, tot_lfct, inf_oak_ct)), 
-#       n = 168)
-```
-
-```{r assess conditional independence 2014}
-f1 <- lme(avg_psi ~ d2c, data = oak_2014, random = ~1|plotid)
-summary(f1) # p = 0.591
-f2 <- lme(avg_psi ~ elevation, data = oak_2014, random = ~1|plotid)
-summary(f2) # p = 0.2171
-f3 <- lme(avg_psi ~ tmi + elevation, data = oak_2014, random = ~1|plotid)
-summary(f3) # tmi; p = 0.1889
-f4 <- lm(os_rich ~ avg_psi + tmi, data = oak_2014)
-# f4 <- lme(os_rich ~ avg_psi + tmi, data = oak_2014, random = ~1|plotid)
-summary(f4) # avg_psi; p = 0.156
-f5 <- lm(log1p(tot_lfct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014)
-# f5 <- lme(log1p(tot_lfct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, random = ~1|plotid)
-# f5 <- glmer(tot_lfct ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2014, family = "poisson")
-summary(f5) # avg_psi; p = 0.265
-
-f6 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "binomial")
-chat <- deviance(f6)/df.residual(f6)
-QAIC(update(f6, family = x.quasibinomial), chat = chat)
-f6 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "quasibinomial")
-# f6 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(avg_psi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2014, family = "binomial")
-summary(f6) # avg_psi; p = 0.000722
-
-f7 <- lm(log(elevation) ~ d2c, data = oak_2014)
-# f7 <- lme(log(elevation) ~ d2c, data = oak_2014, random = ~1|plotid)
-summary(f7) # p = 0.0101
-
-f8 <- lm(log(tmi) ~ d2c + elevation, data = oak_2014)
-# f8 <- lme(log(tmi) ~ d2c + elevation, data = oak_2014, random = ~1|plotid)
-summary(f8) # d2c; p = 0.65795
-
-f9 <- lm(log1p(os_rich) ~ d2c + tmi, data = oak_2014)
-# f9 <- lme(log1p(os_rich) ~ d2c + tmi, data = oak_2014, random = ~1|plotid)
-summary(f9) # d2c; p = 0.73
-
-f10 <- lm(log1p(os_rich) ~ elevation + tmi, data = oak_2014)
-# f10 <- lme(log1p(os_rich) ~ elevation + tmi, data = oak_2014, random = ~1|plotid)
-summary(f10) # elevation; p = 0.126
-
-# f11 <- lm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014)
-f11 <- lm(log1p(tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014)
-f11 <- glm(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, family = "poisson")
-chat <- deviance(f11)/df.residual(f11)
-QAIC(update(f11, family = x.quasipoisson), chat = chat)
-f11 <- glm((tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, family = "quasipoisson")
-# f11 <- lme(log1p(tot_lfct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich), random = ~1|plotid, data = oak_2014)
-# f11 <- glmer(tot_lfct ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2014, family = "poisson")
-summary(f11) # elevation; p = 0.0170
-
-f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "binomial")
-chat <- deviance(f12)/df.residual(f12)
-QAIC(update(f12, family = x.quasibinomial), chat = chat)
-f12 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "quasibinomial")
-# f12 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(elevation) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2014, family = "binomial")
-summary(f12) # elevation; p = 0.00627
-
-f13 <- lm(log1p(tot_lfct) ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2014)
-f13 <- glm(tot_lfct ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, family = "poisson")
-chat <- deviance(f13)/df.residual(f13)
-QAIC(update(f13, family = x.quasipoisson), chat = chat)
-f13 <- glm(tot_lfct ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, family = "quasipoisson")
-
-# f13 <- lme(log1p(tot_lfct) ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich), data = oak_2014, random = ~1|plotid)
-# f13 <- glmer(tot_lfct ~ scale(tmi) + scale(d2c) + scale(elevation) + scale(hrs_14_20) + scale(os_rich) + (1|plotid), data = oak_2014, family = "poisson")
-summary(f13) # tmi; p = 0.2562
-
-f14 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "binomial")
-chat <- deviance(f14)/df.residual(f14)
-QAIC(update(f14, family = x.quasibinomial), chat = chat)
-f14 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct), data = oak_2014, family = "quasibinomial")
-# f14 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tmi) + scale(d2c) + scale(hrs_14_20) + scale(os_rich) + scale(tot_lfct) + (1|plotid), data = oak_2014, family = "binomial")
-summary(f14) # tmi; p = 0.7633
-
-f15 <- lm(log1p(os_rich) ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2014)
-# f15 <- lme(log1p(os_rich) ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2014, random = ~1|plotid)
-summary(f15) # hrs_14_20; p = 0.1039
-
-# Calculate C-statistic
-pvalues <- c(summary(f1)$tTable[2,5], summary(f2)$tTable[2,5], summary(f3)$tTable[2,5], summary(f4)$coefficients[2,4], summary(f5)$coefficients[2,4], summary(f6)$coefficients[2,4], summary(f7)$coefficients[2,4], summary(f8)$coefficients[2,4], summary(f9)$coefficients[2,4], summary(f10)$coefficients[2,4], summary(f11)$coefficients[2,4], summary(f12)$coefficients[2,4], summary(f13)$coefficients[2,4], summary(f14)$coefficients[2,4], summary(f15)$coefficients[2,4])
-
-cstat <- -2 * sum(log(pvalues))
-cstat # 69.18
-chisq.test(c(cstat,30))
-qchisq(pvalues, df = 30)
-chisq.test(cbind(pvalues, qchisq(pvalues, df = 30)))
-```
-
-This C-statistic is much larger than the degrees of freedom, so I'm not confident in the conditional independencies of the structural model being met for the 2014 data.
-In fact, based on the C-test value from the earlier call to `shipleys.test` on these data this value is significantly different.
-
-The models testing for conditional independence indicate direct effects of PSI and elevation on oak infection, as well as a direct effect of elevation on symptomatic leaf count.
-This is indicated by the p-values showing strong statistical significance in the relationship between these variables.
-Below I have generated a new path model that accounts for the unassumed direct effects, which means they become assumed direct effects.
-
-```{r oak path 2014 v2}
-oak_path2 <- DAG(
-      tmi ~ elevation,
-      os_rich ~ tmi,
-      hrs_14_20 ~ elevation + d2c + avg_psi + tmi,
-      tot_lfct ~ hrs_14_20 + os_rich + d2c + elevation,
-      inf_oak_ct ~ tot_lfct + avg_psi + os_rich + hrs_14_20 + d2c + elevation
-      )
-isAcyclic(oak_path2)
-bu_oak_path2 <- basiSet(oak_path2)
-bu_oak_path2
-
-shipley.test(oak_path2, cov(select(oak_2014, elevation, tmi, os_rich, hrs_14_20, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 168)
-
-plotGraph(oak_path2)
-```
-
-This means commenting out models `f6` (oak ~ PSI), `f12` (oak ~ elevation), and `f11` (leaf count ~ elevation) in the `pvalues` summary.
-
-```{r recalculate c-statistic}
-pvalues <- c(
-      summary(f1)$tTable[2,5], summary(f2)$tTable[2,5], summary(f3)$tTable[2,5],
-      summary(f4)$coefficients[2,4], summary(f5)$coefficients[2,4], 
-             # summary(f6)$coefficients[2,4], 
-      summary(f7)$coefficients[2,4], summary(f8)$coefficients[2,4],
-      summary(f9)$coefficients[2,4], summary(f10)$coefficients[2,4],
-      # summary(f11)$coefficients[2,4], 
-      # summary(f12)$coefficients[2,4], 
-      summary(f13)$coefficients[2,4], summary(f14)$coefficients[2,4], summary(f15)$coefficients[2,4])
-
-cstat <- -2 * sum(log(pvalues))
-cstat # 36.56767
-chisq.test(c(cstat,30))
-```
-I think this would be evidence for wanting/needing to use the repeated measures approach across all the data to capture interannual variation in the relationships.
-
-I am continuing with fitting the models to estimate the path coefficients for 2014 data based on this redrawn path model.
-
-```{r estimate path coefficients 2014}
-# f0 <- fitted(mlme1, level = 0)# Values from population model
-# f1 <- fitted(mlme1, level = 1)# Within beach fitted values
-
-# fit1_2014 <- lm(tmi ~ scale(elevation), data = oak_2014)
-fit1_2014 <- lm(log(tmi) ~ scale(elevation), data = oak_2014) # best model
-# fit1_2014 <- lme(log(tmi) ~ scale(elevation), data = oak_2014, random = ~1|plotid)
-summary(fit1_2014)
-par(mfrow = c(2,2))
-plot(fit1_2014)
-
-# fit2_2014 <- lm(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2014)
-# fit2_2014 <- lme(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2014, random = ~1|plotid) # best model
-fit2_2014 <- glm(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2014, family = "poisson")
-chat <- deviance(fit2_2014)/df.residual(fit2_2014)
-QAIC(update(fit2_2014, family = x.quasipoisson), chat = chat)
-fit2_2014 <- glm(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi), data = oak_2014, family = "quasipoisson")
-# fit2_2014 <- glmer(hrs_14_20 ~ scale(elevation) + scale(d2c) + scale(avg_psi) + scale(tmi) + (1|plotid), data = oak_2014, family = "poisson")
-summary(fit2_2014)
-plot(fit2_2014)
-
-# fit3_2014 <- lm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation), data = oak_2014)
-# fit3_2014 <- lm(log1p(tot_lfct) ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation), data = oak_2014)
-fit3_2014 <- glm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation), data = oak_2014, family = "poisson")
-(chat <- deviance(fit3_2014) / df.residual(fit3_2014))
-x.quasipoisson<-function(...){
-  res<-quasipoisson(...)
-  res$aic<-poisson(...)$aic
-  res
-}
-library(MuMIn)
-QAIC(update(fit3_2014, family = x.quasipoisson), chat = chat) # 178.3
-fit3_2014 <- glm(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation), data = oak_2014, family = "quasipoisson") # best model
-# fit3_2014 <- glmer(tot_lfct ~ scale(hrs_14_20) + scale(os_rich) + scale(d2c) + scale(elevation) + (1|plotid), data = oak_2014, family = "poisson")
-summary(fit3_2014)
-
-# fit4_2014 <- lm((inf_oak_ct/(inf_oak_ct+uninf_oak_ct)) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c) + scale(elevation) + scale(avg_psi), data = oak_2014)
-par(mfrow = c(2,2))
-plot(fit4_2014)
-
-fit4_2014 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c) + scale(elevation) + scale(avg_psi), data = oak_2014, family = "binomial")
-(chat <- deviance(fit4_2014) / df.residual(fit4_2014))
-x.quasibinomial<-function(...){
-  res<-quasibinomial(...)
-  res$aic<-binomial(...)$aic
-  res
-}
-library(MuMIn)
-QAIC(update(fit4_2014, family = x.quasibinomial), chat = chat)
-fit4_2014 <- glm(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c) + scale(elevation) + scale(avg_psi), data = oak_2014, family = "quasibinomial")
-
-# fit4_2014 <- glmer(cbind(inf_oak_ct, uninf_oak_ct) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c) + (1|plotid), data = oak_2014, family = "binomial") # best model
-summary(fit4_2014)
-
-# fit5_2014 <- lm(os_rich ~ tmi, data = oak_2014); AIC(fit5_2014)
-fit5_2014 <- lm(log(os_rich) ~ tmi, data = oak_2014); AIC(fit5_2014)
-# fit5_2014 <- lme(os_rich ~ tmi, data = oak_2014, random = ~1|plotid)
-# fit5_2014 <- lme(log(os_rich) ~ tmi, data = oak_2014, random = ~1|plotid)
-summary(fit5_2014)
-```
-
-#### Importance Value
-I also want to calculate some variables related to DBH, and in particular the "importance value" for each species. Importance value is calculated in the literature as:
-
-`IV = 0.5 * (relative density / relative dominance)` where 
- - `Relative Density = # stems of species / total # of stems` and
- - `Relative Dominance = basal area of species / total basal area`
-
-```{r calculate importance value}
-unique(untagged_dbh$species)
-untagged_dbh
-unique(stems$species)
-
-tagged_dbh <- left_join(stems %>% 
-      filter(status == "Alive", location == "In") %>%
-      select(plotid, year, species, dbh) %>% 
-      group_by(plotid, species, year) %>% 
-      summarise(live_count = length(species), live_avg_dbh = mean(dbh), live_tot_dbh = sum(dbh)),
-      stems %>% 
-      filter(status == "Dead", location == "In") %>%
-      select(plotid, year, species, dbh) %>% 
-      group_by(plotid, species, year) %>% 
-      summarise(dead_count = length(species), dead_avg_dbh = mean(dbh), dead_tot_dbh = sum(dbh)),
-      by = c("plotid", "species", "year")
-)
-tagged_dbh <- as.tbl(tagged_dbh)
-summary(tagged_dbh)
-
-tagged_dbh
-untagged_dbh
-
-species_dbh <- rbind(tagged_dbh, untagged_dbh)
-species_dbh
-unique(species_dbh$species)
-species_dbh %>% group_by(plotid, species, year)
-
-stems_dbh <- species_dbh %>% group_by(plotid, year) %>% mutate(all_live_count = sum(live_count), all_live_dbh = sum(live_tot_dbh, na.rm = T), all_dead_count =  sum(dead_count, na.rm = T), all_dead_dbh = sum(dead_tot_dbh, na.rm = T))
-stems_dbh
-summary(stems_dbh)
-filter(stems_dbh, all_live_dbh > 800)
-
-stems_dbh <- stems_dbh %>% mutate(rel_dens = live_count / all_live_count, rel_dom = live_tot_dbh / all_live_dbh, imp_value = 0.5*(rel_dens + rel_dom))
-summary(stems_dbh)
-write.csv(stems_dbh, "analysis/data/species_dbh_impval_plot.csv")
-```
-
-```{r pare down stems dataframe}
-stems %>% select(year, tag, species, status, notes) %>% filter(species == "umca") 
-summary(stems) # Limited to "Alive/Dead"" status, "Missing" statuses excluded
-unique(stems$species)
-unique(stems$plotid)
-summary(bay_laurel) # Still includes stems with "Missing" status
-unique(bay_laurel$plot)
-
-names(stems)
-stemssub <- select(stems, plotid, year, cluster, tag, species, status, slc, canker, dbh, lide_lf_symp, sod_dead, location, dbh1, dbh2, delta_dbh, dbh2_1_ratio, inst_grwth_rate, doy, date)
-summary(stemssub)
-str(stemssub)
-
-summary(oak_sod) # plots with oak stems
-summary(umca_plots) # plots with bay laurel stems
-summary(rain_data)
-summary(temps_data)
-summary(plots_env)
-summary(veg_sub)
-summary(richness)
-summary(stems_dbh)
-
-save.image("~/GitHub/superspreaders/data_201508.RData")
-```
-
-```{r}
-load("~/GitHub/superspreaders/data_201508.RData")
-stemssub$cluster <- as.factor(stemssub$cluster)
-stemssub$tag <- as.factor(stemssub$tag)
-stemssub$canker <- as.factor(stemssub$canker)
-stemssub$sod_dead <- as.factor(stemssub$sod_dead)
-stemssub$date <- as.Date(stemssub$date)
-str(stemssub)
-```

### 2016-01-12 pathmodel_development.R
Add script containing code for path model development.
This code was orignially in the file 'sem_development' (now 'sem_data_development') that ended up being mostly data munging. To better separate the data munging from modeling I removed the code from that file to this one.

@@ -1,0 +1,516 @@
+#' # Path Analysis Presented at 2015 IALE World Congress
+#' Final details for the IALE presentation are in the `iale2015_sem.Rmd` file. 
+
+#' ## Path Analysis of Plot-Level Oak Infection 2005
+#' Here I am using the `oak_sod` data set which will be the oak infection (binomial) as the terminal response variable.
+
+#+ oak_sod path model 2005
+# load data
+load("pathmodel_data_20160112.RData")
+library(ggm)
+names(oak_sod)
+oak_path1 <- DAG(
+      twi15m ~ elev15m,
+      os_rich ~ twi15m,
+      hrs_14_20 ~ elev15m + d2c + avg_psi + twi15m,
+      tot_lfct ~ hrs_14_20 + os_rich + d2c,
+      inf_oak_ct ~ tot_lfct + os_rich + hrs_14_20 + d2c
+)
+isAcyclic(oak_path1)
+bu_oak_path1 <- basiSet(oak_path1)
+bu_oak_path1
+
+oak_path2 <- DAG(
+      twi15m ~ elev15m,
+      os_rich ~ twi15m,
+      hrs_blw10 ~ elev15m + d2c + avg_psi + twi15m,
+      tot_lfct ~ hrs_blw10 + os_rich + d2c,
+      inf_oak_ct ~ tot_lfct + os_rich + hrs_blw10 + d2c
+)
+isAcyclic(oak_path2)
+
+oak_2005 <- filter(oak_sod, year == 2005)
+summary(oak_2005)
+oak_2005 <- select(oak_2005, plotid, uninf_oak_ct, inf_oak_ct, tot_lfct, hrs_14_20, hrs_blw10, hrs_abv25, avg_tmax, avg_tmin, rain05_2d, rain05_v, elev15m, twi15m, avg_psi, os_rich, d2c)
+oak_2005$tot_lfct[is.na(oak_2005$tot_lfct)] <- 0
+filter(oak_2005, is.na(os_rich))
+# oak_2005 <- na.omit(oak_2005)
+oak_2005$plotid <- as.factor(oak_2005$plotid)
+
+# This model makes Gaussian assumptions, which should be incorrect for these data
+shipley.test(oak_path1, cov(select(oak_2005, elev15m, twi15m, os_rich, hrs_14_20, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 170)
[…]

### 2016-01-12 pathmodel_development.R
Edit commenting and text to be knitr::spin compatible.

@@ -7,6 +7,7 @@
 #+ oak_sod path model 2005
 # load data
 load("pathmodel_data_20160112.RData")
+
 library(ggm)
 names(oak_sod)
 oak_path1 <- DAG(
@@ -43,12 +44,11 @@ shipley.test(oak_path1, cov(select(oak_2005, elev15m, twi15m, os_rich, hrs_14_20
 
 shipley.test(oak_path2, cov(select(oak_2005, elev15m, twi15m, os_rich, hrs_blw10, d2c, avg_psi, tot_lfct, inf_oak_ct)), n = 170)
 # The p-value of .04 indicates that the conditional independence assumptions under the application of a Gaussian model are questionable.
-```
 
 
-Violation or questioning of conditional independence supports addressing the complexity of the data using different model distributions to deal with the data.
+#' Violation or questioning of conditional independence supports addressing the complexity of the data using different model distributions to deal with the data.
 
-```{r assess conditional independence claims 2005}
+#+ assess conditional independence claims 2005
 bu_oak_path1 # 15 independence claims means 15 models
 library(nlme); library(lme4); library(MuMIn)
 
@@ -120,22 +120,22 @@ summary(f14) # twi15m; p = 0.61177
 f15 <- lm(os_rich ~ hrs_14_20 + avg_psi + d2c + elev15m + twi15m, data = oak_2005)
 # f15 <- lme(os_rich ~ hrs_14_20 + avg_psi + d2c + elevation + tmi, data = oak_2005, random = ~1|plotid)
 summary(f15) # hrs_14_20; p = 0.08468
-
-# Calculate C-statistic
+#'
+#'
+#+ Calculate C-statistic for 2005 data model
 pvalues <- c(summary(f1)$tTable[2,5], summary(f2)$tTable[2,5], summary(f3)$tTable[2,5], summary(f4)$tTable[2,5], summary(f5)$coefficients[2,4], summary(f6)$coefficients[2,4], summary(f7)$tTable[2,5], summary(f8)$tTable[2,5], summary(f9)$tTable[2,5], summary(f10)$tTable[2,5], summary(f11)$coefficients[2,4], summary(f12)$coefficients[2,4], summary(f13)$coefficients[2,4], summary(f14)$coefficients[2,4], summary(f15)$tTable[2,5])
 
 cstat <- -2 * sum(log(pvalues))
 cstat
-```
-
[…]

### 2016-01-13 analysis/species_diversity_calc.R
Correct NAs generated in data, examine correlations.
To correct some of the NAs generated in calculating diversity metrics I needed to remove some abandoned plots from the data. In a couple instances zeroes needed to be assigned to NA values for richness, because some counts were actually zero.

I added pairs plots to examine correlations between diversity metrics within and between years of overstory and understory measurements, and overstory vs. understory metrics.

@@ -28,7 +28,7 @@ alt14dead <- read.csv("analysis/data/altsp14abund_dead.csv")
 #'
 #' Calculate understory richness
 #+ understory species coverage data from transects ####
-
+library(plyr); library(dplyr)
 summary(veg_us_2005)# data from vegetation transects 2003, 2004, 2005
 veg_us_2005 <- as.tbl(select(veg_us_2005, -X))
 summary(veg_us_2011)# data from vegetation transects 2011
@@ -88,12 +88,27 @@ veg_sub <- filter(veg_us, veg_us$us_species %in% woody_spp)# 56 species
 veg_sub <- droplevels(veg_sub)# drop the unused levels from veg species
 
 # Calculate basic richness as number of different species in each plot each year
-us_richness <- left_join(
+us_richness <- full_join(
       veg_sub %>% filter(year == 2005) %>% group_by(plotid) %>% 
             summarise(us_rich05 = length(unique(us_species))),
       veg_sub %>% filter(year == 2011) %>% group_by(plotid) %>% 
-            summarise(us_rich14 = length(unique(us_species))))
-us_richness
+            summarise(us_rich11 = length(unique(us_species))),
+      by = "plotid")
+summary(us_richness)
+filter(us_richness, is.na(us_rich11))
+# remove plots abandoned early in study
+us_richness <- filter(us_richness, 
+                      plotid != "bush01", plotid != "halow01",
+                      plotid != "sweet01")
+# mcneil01 was lost in 2009 (road built through plot), so NA for later years
+# sumtv02 was abandoned after 2008, so it should be an NA for later years
+# ann06 and mroth03 are actually zeroes for understory species in 2011
+us_richness$us_rich11[us_richness$plotid=="ann06"] <- 0
+us_richness$us_rich11[us_richness$plotid=="mroth03"] <- 0
+# This leaves to legitimate NAs or understory species in 2011
+
+filter(us_richness, is.na(us_rich05))# this should be zero for 2005
+us_richness$us_rich05[us_richness$plotid=="jlsp05"] <- 0
 
 
 # 2005 vegetation data proportional species coverage
@@ -109,6 +124,7 @@ veg_us_2005_ptct <- veg_sub %>% filter(year == 2005) %>%
[…]

### 2016-01-13 sem__data_development.Rmd
Update data cleaning, improve aesthetics.
I re-examined data and reproduced plot level data sets for oak and UMCA plots. I added some pairs plots of various subsets of data from UMCA and oak plots to examine correlations between variables. I resaved the path model .RData file and the stems data subset to exclude the BUSH01 plot.

@@
 date: "June 15, 2015"
-output: html_document
+output: 
+  html_document:
+    toc: true
 ---
 
-#### Load data
+## Load data
 ```{r pairs helper functions, echo=FALSE}
@@
 
-#### Join remotely sensed and derived physical data
+## Join remotely sensed and derived physical data
 LiDAR products, variables derived from DEMs
@@
 rm(plots_vegtopo)
-
 ```
 
-#### Convert rain data from wide to long
+## Convert rain data from wide to long
 ```{r rain wide to long}
@@
 rain_total <- left_join(rain_total, rain_psm_total, by = c("plotid", "year"))
+summary(rain_total)# NAs are due to current PRISM data only going through 2010
+unique(rain_total$plotid)# 202 plots
+unique(rain_total$year)# 10 sampling seasons
 
@@
 rain_days <- left_join(rain_days, rain_r_days, by = c("plotid", "year"))
+summary(rain_days)
+unique(rain_days$plotid)# 202 plots
+unique(rain_days$year)# 10 sampling seasons
 
@@
 
-#### Explore data for completeness and matching
+## Explore data for completeness and matching
 ```{r explore plot_env data, message=FALSE}
@@
 ```{r examine temperature & rain data}
+# Temperature Data
 summary(rs_temps)
@@
 
+# Rainfall Data
 summary(ppt_days)
@@
 
-unique(rs_temps$plotid)# 201 levels
-unique(summer_temps$plotid)# 201 levels
-unique(rain_df$plotid)# 202 levels
+unique(rs_temps$plotid)# 201 plots
+unique(summer_temps$plotid)# 201 plots
+unique(rain_df$plotid)# 202 plots
 
@@
 
-#### Investigating `NAs` in the `plots_temps` data set
+## Investigating `NAs` in the `plots_temps` data set
 A number of observations for above 25 threshold have NAs.
@@
 
-#### Save data with BUSH01 excluded
+### Save data with BUSH01 excluded
 ```{r save data}
@@
 
-### Starting Point with Revised Data: Reload data that has BUSH01 excluded
+## Starting Point with Revised Data: Reload data that has BUSH01 excluded
 Still need to load the .RData source at the beginning of this document, line 32.
@@
 
-### Plot contributing area vs. slope
+## Plot Contributing Area vs. Slope
 This is a suggestion from Ryan to plot the contributing area vs.
@@
 library(ggplot2)
-# summary(plots_vegtopo)
 # par(mfrow = c(3,1))
 names(plots_env)
+summary(plots_env)
+## 'facc_ws15m' is the estimate from the 'r.watershed' function in GRASS GIS
+
+# Plot estimates from two different functions agains each other.
 qplot(flowacc15m, facc_ws15m, data = plots_env) #+
       geom_text(aes(label = plotid))
+cor(plots_env$flowacc15m, plots_env$facc_ws15m)
+## There are 9 plots that really stand out from the others.
+## The r.watershed calculation provides similar, but generally smaller values.
+qplot(flowacc15m, facc_ws15m, data = filter(plots_env, flowacc15m > 300),
+      main = "'High' Flow Accumulation Value Plots") +
+      geom_text(aes(label = plotid))
+qplot(flowacc15m, facc_ws15m, data = filter(plots_env, flowacc15m < 300),
+      main = "'Low' Flow Accumulation Value Plots") +
+      geom_text(aes(label = plotid))
+## Much of the variability between the two estimates stems from a few plots.
 
+# Plot flow accumulation vs. slope
 qplot(slope15m, flowacc15m, data = plots_env) #+
       geom_text(aes(label = plotid))
+## Can't see much pattern (if there is any) because of a few big values
+      cor(plots_env$slope15m, plots_env$flowacc15m)
+      ## negative, but weak
 
-qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m < 20), 
-      main = "Flow Accumulation < 20") #+
+qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m < 300), 
+      main = "Flow Accumulation <300") #+
+      geom_text(aes(label = plotid))
+## Generally, a shallower slope means higher flow accumulation, but it's not a strong pattern
+
+qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 300), 
+      main = "Flow Accumulation >300 ") #+
+      geom_text(aes(label = plotid))
+## The set of high flow accumulation plots shows an opposite general trend in the relationship between flow accumulation and slope.
+It's a small sample size, which leads me to consider the general trend from the large sample size being closer to the real relationship.
+
+qplot(slope15m, facc_ws15m, data = plots_env)
+cor(plots_env$slope15m, plots_env$facc_ws15m)
+## weakly negative correlation between slope and flow accumulation
+
+qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m < 300), 
+      main = "Flow Accumulation (r.watershed) <300") #+
+      geom_text(aes(label = plotid))
+qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 300), 
+      main = "Flow Accumulation (r.watershed) >300") #+
       geom_text(aes(label = plotid))
 
-qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 20 & flowacc15m < 1000), main = "Flow Accumulation 20 - 1000") #+
-      geom_text(aes(label = plotid))
-
-qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 1000), 
-      main = "Flow Accumulation > 1000") #+
-      geom_text(aes(label = plotid))
-
-# I think the 'facc_ws15m' is the estimate from the 'r.watershed' function in GRASS GIS
-qplot(slope15m, facc_ws15m, data = plots_env)
-qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m < 15), 
-      main = "facc < 15") +
-      geom_text(aes(label = plotid))
-qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 15 & facc_ws15m < 50), 
-      main = "facc 15 - 50") +
-      geom_text(aes(label = plotid))
 qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 50 & facc_ws15m < 200), 
@@
 
-plot(flowacc15m ~ slope15m, data = filter(plots_env, flowacc15m < 30))
-pairs(select(plots_env, -plotid) %>% filter(flowacc15m < 30), 
+plot(flowacc15m ~ slope15m, data = filter(plots_env, flowacc15m > 300))
+pairs(select(plots_env, flowacc15m, slope15m) %>% filter(flowacc15m > 300), 
       lower.panel = panel.cor, diag.panel = panel.hist)
@@
            axes = axes,
-           cex = cex,
-           pch = if(pch) 19 else 1
+           cex = cex
+           #pch = if(pch) 19 else 1
            ),
@@
 
-#### Examine the bay laurel data at individual and plot levels
+## Examine the bay laurel data at individual and plot levels
 ```{r examine umca data}
@@
 
-I can use this to calculate plot level leaf count variables (sum, average) for each plot and year.
+I can use the stem-level data to calculate plot level leaf count variables (sum, average) for each plot and year.
 ```{r calculate plot level leaf counts}
+live_umca <- bay_laurel %>% filter(status == "Alive")
+filter(live_umca, is.na(slc))
+# Spot checking some of these shows that NAs are reasonable
 umca_plots <- left_join(
@@
 filter(umca_plots, is.na(avg_lfct))
+filter(bay_laurel, plot == "YAHNG02", status == "Alive")
 # I'm not sure why this wasn't excluded from the calculations with the na.rm = T option
-filter(bay_laurel, plot == "YAHNG02", status == "Alive")
 # I'm removing it manually from the `umca_plots` data frame
@@
 
-So now I have a plot-level of UMCA infection levels for each year in terms of symptomatic leaf counts and number of un/infected stems.
+So now I have a plot-level data frame of UMCA infection for each year in terms of symptomatic leaf counts and number of un/infected stems.
 I want to join this with some of the plot-level environmental data.
@@
 rs_temps$plotid <- tolower(rs_temps$plotid)
+summary(rs_temps)
 rs_temps <- rename(rs_temps, 
@@
 summer_temps <- as.tbl(summer_temps)
-summer_temps
 summer_temps$plotid <- tolower(summer_temps$plotid)
+summary(summer_temps)
 summer_temps <- rename(summer_temps, 
@@
 umca_plots <- left_join(umca_plots, summer_temps, by = c("plotid", "sample_year"))
-names(umca_plots)
+summary(umca_plots)# 7 NAs introduced into the temperature variables
+filter(umca_plots, is.na(avg_tmax_rs))# NAs due to records from 2003
+## I don't think these will end up in my analysis, but no reason to remove them at this time.
 
 # Add rainfall data to UMCA plots data frame
-names(rain_total)
-names(rain_days)
+summary(rain_total)
+unique(rain_total$plotid)
+rain_total$plotid <- tolower(rain_total$plotid)
+summary(rain_days)
+unique(rain_days$plotid)
+rain_days$plotid <- tolower(rain_days$plotid)
+
+rain_total <- rename(rain_total, sample_year = year)
+rain_total$sample_year <- as.character(rain_total$sample_year)
+rain_days <- rename(rain_days, sample_year = year)
+rain_days$sample_year <- as.character(rain_days$sample_year)
+
+summary(umca_plots)
+umca_plots$sample_year <- as.character(umca_plots$sample_year)
 umca_plots <- left_join(umca_plots, rain_total, by = c("plotid", "sample_year"))
@@
 filter(umca_plots, is.na(rain_tot_2d))
-# NAs in rainfall data are from 2003
-filter(umca_plots, sample_year == 2003)
-# Only 7 plots w/UMCA sampled in 2003, so removing these records
-umca_plots <- filter(umca_plots, sample_year != 2003)
+# Same case of NAs in rainfall due to data from 2003
 ```
 
-### Load and Join Richness, Diversity, Evenness Data
+## Load and Join Richness, Diversity, Evenness Data
 Now I want to join the richness data set and the canopy cover estimates to the `umca_plots` data frame that currently has the environmental and leaf count/infection variables.
@@
 os.us.diversity <- as.tbl(read.csv("analysis/data/os_us_diversity.csv"))
-os.us.diversity
+summary(os.us.diversity)
+# variables prefixed with 'us' indicate understory metrics
 umca_plots
@@
 umca_plots <- left_join(umca_plots, os.us.diversity, by = c("plotid"))
+summary(umca_plots)
+filter(umca_plots, is.na(H.2014))$plotid
+## NAs due to plots abandoned/lost between 2005 and 2014; 4 unique plots
+filter(umca_plots, is.na(us.H.2005))
+## NAs due to plot having '0' for understory richness
+filter(umca_plots, is.na(us.H.2011))
+## NAs due to 2 plots being abaondoned prior to 2011 & one plot having '0' for understory richness
+
 head(canopy_cover)
@@
 canopy_cover <- rename(canopy_cover, sample_year = year)
-unique(filter(canopy_cover, year == 2006)$plotid)
+summary(canopy_cover)
+unique(filter(canopy_cover, sample_year == 2006)$plotid)
+canopy_cover$sample_year <- as.character(canopy_cover$sample_year)
 
 umca_plots <- left_join(umca_plots, canopy_cover, by = c("plotid", "sample_year"))
+summary(umca_plots)# NAs because canopy cover was not measured every year
 ```
@@
 
-# pairs(select(plots_umca, -plotid, -year, -starts_with("psi"), -tagged_rich, -untagged_rich), lower.panel = panel.cor, diag.panel = panel.hist, main = "UMCA SOD Data")
+pairs(select(umca_plots, candens15m, pct_cvr, veght15m), 
+      lower.panel = panel.cor, diag.panel = panel.hist, 
+      main = "Canopy Data - UMCA Plots")
+cor(umca_plots$candens15m, umca_plots$pct_cvr, use = "complete.obs")
+cor(umca_plots$veght15m, umca_plots$pct_cvr, use = "complete.obs")
+
+pairs(na.omit(select(umca_plots, avg_lfct, tot_lfct, contains("hrs"), contains("rain_tot"), H.2005, H.2014)),
+      lower.panel = panel.cor, diag.panel = panel.hist, 
+      main = "Leaf Count, Temperature, Rainfall, Diversity - UMCA Plots")
+## Rainfall and temperature all have fairly weak correlations (<0.36)
+## The diversity metrics are the only correlates with leaf count >0.20
 ```
@@
 umca_plots
+umca_plots$sample_year <- as.integer(umca_plots$sample_year)
 plot_qusp # right now this is broken down by oak species
@@
 summary(oak_sod)
-filter(oak_sod, is.na(elev15m))
+filter(oak_sod, is.na(elev_m))
 # NAs are plots that were abandoned early in the study
@@
 
+filter(oak_sod, is.na(H.2014))$plotid
+## plots that were decommissioned prior to 2014
+filter(oak_sod, is.na(us.H.2011))$plotid
+## plots that have legit 0 for richness or were decommissioned prior to 2011
+
+canopy_cover$sample_year <- as.integer(canopy_cover$sample_year)
 oak_sod <- left_join(oak_sod, canopy_cover, by = c("plotid", "sample_year"))
@@
 summary(rain_total)
-rain_total$plotid <- tolower(rain_total$plotid)
-rain_total <- rename(rain_total, sample_year = year)
-# rain_total$sample_year <- as.numeric(rain_total$sample_year)
+rain_total$sample_year <- as.integer(rain_total$sample_year)
+
 summary(rain_days)
-rain_days$plotid <- tolower(rain_days$plotid)
-rain_days <- rename(rain_days, sample_year = year)
-#rain_days$sample_year <- as.numeric(as.character(rain_days$sample_year))
+rain_days$sample_year <- as.integer(rain_days$sample_year)
 
@@
 summary(filter(oak_sod, is.na(rain_tot_psm))) # all obs after 2010
+summary(filter(oak_sod, is.na(us.J.2011)))
+filter(oak_sod, is.na(us.J.2011))$plotid# ann06, mroth03, sumtv02
+# ann06 & mroth03 are legitimate zeroes for understory richness in 2011
+# sumtv02 was abandoned after 2008, so it should be an NA for later years
 
@@
 oak_sod <- ungroup(oak_sod)
+summary(oak_sod)
 pairs(select(oak_sod, elev_m, starts_with("hrs"), starts_with("avg_t"), contains("rain_tot")), lower.panel = panel.cor, diag.panel = panel.hist, main = "Oak-SOD Rainfall & Temperature Data")
+
+pairs(na.omit(select(oak_sod, c(3:4,7:8,9:10), H.2005, H.2014)),
+      lower.panel = panel.cor, diag.panel = panel.hist, 
+      main = "Oak-SOD Infection & Diversity Data")
 ```
 
-### Save patha analysis data environment
+### Save path analysis data environment
 ```{r save data environment}

### 2016-01-26 ch2_path_appendix2.Rmd
Edit to incorporat table of contents.

--- 2014-12-30 ch2_path_appendix2.Rmd
@@
 ---
-title: "Appendix 2 - Fitting proposed path models"
+title: "Appendix 2 - Fitting proposed path models using the C-statistic"
 author: "Whalen Dillon"
 date: "December 17, 2014"
-output: html_document
+output: 
+  word_document:
+    toc: true
 ---
 ----
-#### This appendix details steps 1-4 of calculating the *C*-statistic for fitting a path model in a generalized multilevel context (Shipley 2009).
 
-I go through these steps for the path models in Fig. 7 and Fig. 8 (included here)  to produce the regression equations that are indicated by the path diagram.
+This appendix details steps 1-4 of calculating the *C*-statistic for fitting a path model in a generalized multilevel context (Shipley 2009).
+I go through these steps for the path models in Fig.
+7 and Fig.
+8 (included here)  to produce the regression equations that are indicated by the path diagram.
 
-#### First the model in Fig. 7
-**Step 1:** Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).
+# The model in Fig. 7
+## Step 1
+Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).
 
@@
 
-**Step 2:** List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.
+## Step 2:
+List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.
 
-**Independence Set - Fig. 7**
+### Independence Set - Fig. 7
 
@@
 
-**Step 3:** For each of the *k* pairs of variables (*Xi,Xj*), list the set of other variables {*Z*} that are direct causes of either *Xi* or *Xj*.
+## Step 3
+For each of the *k* pairs of variables (*Xi,Xj*), list the set of other variables {*Z*} that are direct causes of either *Xi* or *Xj*.
 The pair of variables, (*Xi,Xj*), along with its conditioning set, {*Z*}, define an independence claim, (*Xi,Xj*) | {*Z*}, and the full set of the *k* independence claims defines the basis set, *Bu*.
@@
 
-**The Basis Set of Independence Claims - Fig. 7**
+### The Basis Set of Independence Claims - Fig. 7
 
@@
 
-**Step 4:** For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*.
+## Step 4
+For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*.
 In other words, perform a regression model using an appropriate method.
@@
 
-**Model Set - Fig. 7**
+### Model Set - Fig. 7
 
@@
 
-**Step 5:** Combine the *k* probabilities into the *C*-statistic:
+## Step 5 
+Combine the *k* probabilities into the *C*-statistic:
 
@@
 
-#### The model in Fig. 8
+# The model in Fig. 8
 
-**Step 1:** Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).
+## Step 1
+Express the hypothesized causal relationships between the variables in the form of a directed acyclic graph (DAG).
 
@@
 
-**Step 2:** List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.
+## Step 2
+List each of the *k* pairs of variables in the graph that do not have an arrow directly between them. The topology of the graph indicates that these variables should be independent.
 
-**Independence Sets - Fig. 8**
+### Independence Sets - Fig. 8
 
@@
 
-**The Basis Set of Independence Claims - Fig. 8**
+### The Basis Set of Independence Claims - Fig. 8
 
@@
 
-**Step 4:** For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*.
+## Step 4
+For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*.
 
-**Model set - Fig. 8**
+### Model set - Fig. 8
 

### 2016-01-26 d-sep_test.Rmd
knitr output maybe?
substantial changes
new data, RStudio files, updated code in multiple files
Edit to be compatible with table of contents.

@@
-## Steps for Performing a d-Sep Test
-###### (Shipley 2000, 2003, 2004, 2009)
+---
+output: html_document
+---
+
+## Steps for Performing a d-Sep Test (Shipley 2000, 2003, 2004, 2009)
 
@@
 <img src="figures/CH2_DAG.svg" width="600px" height="600px">
+
 2. List each of the *k* pairs of variables in the graph that do not have an arrow between them:
@@
 
-### Example from Shipley 2009 on how to perform the d-sep test in a generalized multilevel context 
+## Example from Shipley 2009 on how to perform the d-sep test in a generalized multilevel context 
 This is to demonstrate how to assess a multilevel, repeated measures path model by applying mixed models in the tests of d-separation.
@@
 
-##### Path Model
+### Path Model
 `Latitude` --> `Degree-days` --> `Date of bud burst` --> `Growth` --> `Live`
@@
 
+### Establish Basis Set
 ```{r get basis set, eval=TRUE}
@@
 
+### Assess Independence Claims
 The first independence claim is that `Date` is independent of `Latitude` after controlling for `DD`.
@@
 Now I want to get the fixed-effects summaries for each model, and in particular the null probability estimates (p-values) for the pair of variables in each independence claim.
+
+### Independence Claim Model Summaries
 ```{r nlme model & probability summaries}
@@
 I would like to learn how to calculate the chi-square cumulative probability based on the degrees of freedom and *C*-statistic, a la the handy tool on this [website](http://stattrek.com/online-calculator/chi-square.aspx).
-
 It's probably simple...

### 2016-01-26 pathmodel_development.R
Revise to new variable names, apply piecewiseSEM package.

@@ -7,15 +7,17 @@
 #+ oak_sod path model 2005
 # load data
 load("pathmodel_data_20160112.RData")
+library(plyr); library(dplyr)
+pairs(select(oak_sod, elev_m, starts_with("hrs"), starts_with("avg_t"), contains("rain_tot")), lower.panel = panel.cor, diag.panel = panel.hist, main = "Oak-SOD Rainfall & Temperature Data")
 
 library(ggm)
 names(oak_sod)
 oak_path1 <- DAG(
-      twi15m ~ elev15m,
-      os_rich ~ twi15m,
-      hrs_14_20 ~ elev15m + d2c + avg_psi + twi15m,
-      tot_lfct ~ hrs_14_20 + os_rich + d2c,
-      inf_oak_ct ~ tot_lfct + os_rich + hrs_14_20 + d2c
+      twi15m ~ elev_m,
+      os_rich05 ~ twi15m,
+      hrs_14_20_rs ~ elev_m + d2c + avg_psi + twi15m,
+      tot_lfct ~ hrs_14_20_rs + os_rich05 + d2c,
+      inf_oak_ct ~ tot_lfct + os_rich05 + hrs_14_20_rs + d2c
 )
 isAcyclic(oak_path1)
 bu_oak_path1 <- basiSet(oak_path1)
@@ -30,13 +32,36 @@ oak_path2 <- DAG(
 )
 isAcyclic(oak_path2)
 
-oak_2005 <- filter(oak_sod, year == 2005)
+oak_2005 <- filter(oak_sod, sample_year == 2005)
 summary(oak_2005)
-oak_2005 <- select(oak_2005, plotid, uninf_oak_ct, inf_oak_ct, tot_lfct, hrs_14_20, hrs_blw10, hrs_abv25, avg_tmax, avg_tmin, rain05_2d, rain05_v, elev15m, twi15m, avg_psi, os_rich, d2c)
+# oak_2005 <- select(oak_2005, plotid, uninf_oak_ct, inf_oak_ct, tot_lfct, hrs_14_20, hrs_blw10, hrs_abv25, avg_tmax, avg_tmin, rain05_2d, rain05_v, elev15m, twi15m, avg_psi, os_rich, d2c)
+filter(oak_2005, is.na(oak_2005$tot_lfct))
+
 oak_2005$tot_lfct[is.na(oak_2005$tot_lfct)] <- 0
-filter(oak_2005, is.na(os_rich))
[…]

### 2016-01-26 climate-path-model.R
Add script for path modeling using temperature and precipitation.

@@ -1,0 +1,7 @@
+#' # Path Modeling of Climatic Influence on Spillover
+#' I'm using the `piecewiseSEM package, which operationalized the methods from Shipley 2009 & 2013 in particular for fitting and assessing path models that enable applying hierarchical modeling methods.
+#' 
+#+ load data and packages
+load("pathmodel_data_20160112.RData")
+library(plyr); library(dplyr); library(piecewiseSEM)
+library(lmerTest); library(nlme)

### 2016-01-26 iale2015_sem.Rmd
Edit to knit reproducibly.

@@
 
-```{r load data}
+```{r load data, echo=F, message=F}
 # Ecological data: leaf counts, infection
@@
 library(rgdal)
-plots_shp <- readOGR("analysis/data/plot_gis/", "soco_plots_metric")
+plots_shp <- readOGR("analysis/data/plot_gis/", "soco_plots_metric", verbose = F)
 summary(plots_shp)
@@
 library(plyr); library(dplyr)
-plots_temps <- read.csv("analysis/data/rainy_season_temps.csv")
+plots_temps <- read.csv("analysis/data/rs_temps_data.csv")
 summary(plots_temps)
-plots_temps <- select(plots_temps, -X)
+# plots_temps <- select(plots_temps, -X)
 ```
@@
 Join the plot level environmental data (geographic/topographic & temperatures)
-```{r join plot data, message=FALSE}
+```{r join plot data, message=FALSE, echo=F}
 plots_env <- left_join(plots_env, plots_temps, by = c("PLOT_ID"="plotid"))
@@
 Investigating `NAs` in the `plots_env` data set
-```{r investigate NAs}
+```{r investigate NAs, echo=F}
 summary(plots_env)
@@
 Examine the bay laurel data at individual and plot levels
-```{r examine umca data}
+```{r examine umca data, echo=F}
 summary(plots_umca)
@@
 I can use this to calculate plot level leaf count variables (sum, average) for each plot and year
-```{r calculate plot level leaf counts}
+```{r calculate plot level leaf counts, echo=F}
 plots_umca <- left_join(plots_umca, bay_laurel %>% filter(status == "Alive") %>% 
@@
 
-```{r examine lfct NAs}
+```{r examine lfct NAs, echo=F, results='hide'}
 filter(plots_umca, is.na(avg_lfct))
@@
 
-```{r join env and umca data}
+```{r join env and umca data, echo=F, results='hide'}
 summary(plots_env)
@@
 Now I also want to calculate some plot-level community variables from the vegetation data. First, understory richness from the vegetation transects.
-```{r calculate understory richness}
+```{r calculate understory richness, echo=F, results='hide'}
 summary(veg_us_2005)
@@
 
-```{r calculate overstory richness}
+```{r calculate overstory richness, echo=F, results='hide'}
 summary(untagged_dbh)
@@
 
-```{r join richness and umca data}
+```{r join richness and umca data, echo=F, results='hide'}
 plots_umca
@@
 Make a pairs plot of the continuous variables for UMCA plots.
-```{r umca sod pairs plot}
+```{r umca sod pairs plot, echo=F, results='hide'}
 # Helper functions for `pairs`
@@
 
-```{r join umca and qusp data}
+```{r join umca and qusp data, echo=F, results='hide'}
 plots_umca
@@
 
-```{r oak_sod pairs plot}
+```{r oak_sod pairs plot, echo=F}
 class(oak_sod)
@@
 
-```{r assess conditional independence claims 2005}
+```{r assess conditional independence claims 2005, message=F}
 bu_oak_path1 # 15 independence claims means 15 models
-library(nlme)
-library(lme4)
+library(nlme); library(lme4); library(MuMIn)
+
+x.quasibinomial <- function(...) {
+    res <- quasibinomial(...)
+    res$aic <- binomial(...)$aic
+    res
+}
+
+x.quasipoisson <- function(...) {
+    res <- quasipoisson(...)
+    res$aic <- poisson(...)$aic
+    res
+}
 
@@
 
-```{r estimate path coefficients 2005}
+```{r estimate path coefficients 2005, message=F}
 # f0 <- fitted(mlme1, level = 0)# Values from population model
@@
 
-```{r assess conditional independence 2014}
+```{r assess conditional independence 2014, message=F}
 f1 <- lme(avg_psi ~ d2c, data = oak_2014, random = ~1|plotid)
@@
 
-```{r recalculate c-statistic}
+```{r recalculate c-statistic, echo=F}
 pvalues <- c(
@@
 
-```{r estimate path coefficients 2014}
+```{r estimate path coefficients 2014, message=F}
 # f0 <- fitted(mlme1, level = 0)# Values from population model
@@
 # fit4_2014 <- lm((inf_oak_ct/(inf_oak_ct+uninf_oak_ct)) ~ scale(tot_lfct) + scale(os_rich) + scale(hrs_14_20) + scale(d2c) + scale(elevation) + scale(avg_psi), data = oak_2014)
-par(mfrow = c(2,2))
-plot(fit4_2014)
 
@@
 summary(fit4_2014)
+par(mfrow = c(2,2))
+plot(fit4_2014)
 
@@
 
-```{r calculate importance value}
+```{r calculate importance value, eval=F, echo=F}
 unique(untagged_dbh$species)

### 2016-01-28 climate-path-model.R
Code for assessing plot-level path model for oak plots.
Fitting of initial model was unacceptable at p=.05 threshold. I revised, compar 'glmer' with Poisson family to 'lme' and a log-transformed response for modeling total bay count. The AIC of the SEM using 'lme' was lower, so I used that method. Missing paths were revealed so I added those and refit the model (lines149-156).

@@ -1,7 +1,171 @@
 #' # Path Modeling of Climatic Influence on Spillover
 #' I'm using the `piecewiseSEM package, which operationalized the methods from Shipley 2009 & 2013 in particular for fitting and assessing path models that enable applying hierarchical modeling methods.
 #' 
-#+ load data and packages
+#+ load data and packages ####
 load("pathmodel_data_20160112.RData")
 library(plyr); library(dplyr); library(piecewiseSEM)
-library(lmerTest); library(nlme)
+library(nlme); library(lme4); library(lmerTest)
+
+#+ panel functions for pairs plots
+panel.cor <- function(x, y, digits = 2, cex.cor, ...)
+{
+      usr <- par("usr"); on.exit(par(usr))
+      par(usr = c(0, 1, 0, 1))
+      # correlation coefficient
+      r <- cor(x, y)
+      txt <- format(c(r, 0.123456789), digits = digits)[1]
+      txt <- paste("r= ", txt, sep = "")
+      text(0.5, 0.6, txt)
+      
+      # p-value calculation
+      p <- cor.test(x, y)$p.value
+      txt2 <- format(c(p, 0.123456789), digits = digits)[1]
+      txt2 <- paste("p= ", txt2, sep = "")
+      if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
+      text(0.5, 0.4, txt2)
+}
+
+panel.hist <- function(x, ...)
+{
+      usr <- par("usr"); on.exit(par(usr))
+      par(usr = c(usr[1:2], 0, 1.5) )
+      h <- hist(x, plot = FALSE)
+      breaks <- h$breaks; nB <- length(breaks)
+      y <- h$counts; y <- y/max(y)
+      rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
+}
+
+#+ examine oak plots data ####
+summary(oak_sod)
+oak_sod$plotid <- as.factor(oak_sod$plotid)
+str(oak_sod)
+
+#+ pairs plots of oak plots data ####
+pairs(na.omit(select(oak_sod, starts_with("rain"))), 
+      lower.panel = panel.cor, diag.panel = panel.hist,
+      main = "Oak Plots Rainfall Variables")
+# The regression variables are the least correlated within totals & days.
[…]

### 2016-02-09 sem__data_development.Rmd
Update and rename path model data.

@@
 ```
+
+### Update path analysis data
+Updating the path analysis data with the temperature during wet days variables.
+```{r update data}
+load("pathmodel_data_20160112.RData")
+wet_hrs <- read.csv("analysis/data/wet-days-temperature-variables.csv")
+save.image("~/GitHub/superspreaders/pathmodel_data.RData")
+```

### 2016-02-09 analysis/explore_data.R
Update variable name and add start of t-test analysis.

@@ -403,9 +403,9 @@ str(stems)
 summary(stems)
 
 # Exploratory plot of dbh data
-qplot(delta_dbh, dbh1, data = stems %>% select(plot, cluster, tag, species, delta_dbh, dbh1, year, status) %>% 
-            group_by(plot, cluster, tag) %>%
-            filter(year == 2014, status == "Alive" | status == "Dead"), 
+qplot(delta_dbh, dbh1, data = stems %>% select(plotid, cluster, tag, species, delta_dbh, dbh1, year, status) %>% 
+            group_by(plotid, cluster, tag) %>%
+            filter(year == 2014, status == "Alive" | status == "Dead", species == "UMCA"), 
       color = status)
 # A visual check shows that there are a handful of stems with growth > 20cm
 
@@ -425,27 +425,47 @@ write.csv(stems, "analysis/data/stems_growth.csv")
 summary(stems$status)
 summary(stems$tag)
 
-qplot(dbh2_1_ratio, data = stems %>% select(plot, tag, species, dbh2_1_ratio, dbh1, year, status, doy) %>% 
-            group_by(plot, tag) %>% 
+qplot(dbh2_1_ratio, data = stems %>% select(plotid, tag, species, dbh2_1_ratio, dbh1, year, status, doy) %>% 
+            group_by(plotid, tag) %>% 
             filter(status == "Alive" | status == "Dead", dbh1 >= 2.0),
       color = status)
 
 qplot(stems$dbh2_1_ratio, main = "Relative Growth")
 
-qplot(dbh1, dbh2_1_ratio, data = stems %>% select(plot, cluster, tag, species, dbh2_1_ratio, dbh1, year, status, dbh) %>% 
-            group_by(plot, cluster, tag) %>%
+qplot(dbh1, dbh2_1_ratio, data = stems %>% select(plotid, cluster, tag, species, dbh2_1_ratio, dbh1, year, status, dbh) %>% 
+            group_by(plotid, cluster, tag) %>%
             filter(year == 2014, status == "Alive" | status == "Dead", dbh1 >= 2.0), 
       color = status, main = "Change in DBH vs. Original DBH")
 
-qplot(dbh1, inst_grwth_rate, data = stems %>% select(plot, cluster, tag, species, inst_grwth_rate, dbh1, year, status) %>% 
+qplot(dbh1, inst_grwth_rate, data = stems %>% select(plotid, cluster, tag, species, inst_grwth_rate, dbh1, year, status) %>% 
[…]

### 2016-02-09 climate-path-model.R
Update data source for script.

@@ -2,7 +2,7 @@
 #' I'm using the `piecewiseSEM package, which operationalized the methods from Shipley 2009 & 2013 in particular for fitting and assessing path models that enable applying hierarchical modeling methods.
 #' 
 #+ load data and packages ####
-load("pathmodel_data_20160112.RData")
+load("pathmodel_data.RData")
 library(plyr); library(dplyr); library(piecewiseSEM)
 library(nlme); library(lme4); library(lmerTest)
 

### 2016-02-12 analysis/explore_data.R
Update DBH checking.
I added checks for duplicate plots, discovering that SUGAR02 from 2004 == SUGAR30 from 2012. Started on some code to compare DBH measurements between establishment and remeasurement in 2014 (line 393).

@@ -220,7 +220,7 @@ stems <- read.csv("analysis/data/all_stems_corrected.csv")
 library(dplyr)
 
 plots_umca <- stems %>%
-      select(plot, date, species, slc, year, status) %>%
+      select(plot, Date, species, slc, year, status) %>%
       group_by(plot, year, species) %>%
       filter(species == "UMCA", status == "Alive") %>%
       summarise(uninfected_bay_ct = length(which(slc==0)), infected_bay_ct = length(which(slc > 0)), tot_bay = length(species)) 
@@ -231,8 +231,8 @@ plots_umca <- droplevels(plots_umca)
 
 
 length(which(plots_umca$ct_bay_NA > 0))
-filter(plots_umca, infected == 0)
-filter(plots_umca, infected_bay_ct == 0)
+# filter(plots_umca, infected == 0)
+summary(filter(plots_umca, infected_bay_ct == 0))# 49 plots, max year = 2010
 
 write.csv(plots_umca, "analysis/data/plots_umca_infection.csv")
 # I sent this file to Francesco for informing the spread model
@@ -260,7 +260,7 @@ tmp <- plot_qusp %>%
       dcast(year + plot ~ species, value.var = "infected_oak_ct")
 head(tmp)
 
-#### Checking large positive DBH changes ####
+# Checking large positive DBH changes ####
 ## So at this point I think I have corrected all the negative dbh change values that I legitimately could in the database. I have gone back and exported the query again and use this file below, and then recycling code from the beginning of this document.
 stems <- read.csv("analysis/data/Stem_Summary_negDBH_corrected_201402.csv")
 summary(stems)
@@ -284,7 +284,7 @@ summary(stems)
 stems[stems == -9999] <- NA
 summary(stems)
 
-stems <- rename(stems, plot = PlotID)
+stems <- rename(stems, plotid = PlotID)
 stems <- rename(stems, cluster = HostID, tag = TagNumber, 
                 species = SpeciesID, status = StemStatus, 
                 alive_class = AliveClass, dead_class= DeadClass,
@@ -303,65 +303,113 @@ stems <- as.tbl(stems)
 
[…]

### 2016-02-12 climate-path-model.R
Update 'oak_sod' data frame with 'wet-hours' data.

@@ -40,20 +40,31 @@ summary(oak_sod)
 oak_sod$plotid <- as.factor(oak_sod$plotid)
 str(oak_sod)
 
+## join wet-hours data
+summary(wet_hrs)# since these are counts the NAs are in fact zeroes
+wet_hrs[is.na(wet_hrs)] <- 0# introduce a lot of zeroes for some columns
+oak_sod <- left_join(
+      oak_sod, select(wet_hrs, -X), by = c("plotid","sample_year"))
+oak_sod$plotid <- as.factor(oak_sod$plotid)
+
 #+ pairs plots of oak plots data ####
-pairs(na.omit(select(oak_sod, starts_with("rain"))), 
+pairs(na.omit(select(oak_sod, starts_with("rain"), ends_with("wet"))), 
       lower.panel = panel.cor, diag.panel = panel.hist,
       main = "Oak Plots Rainfall Variables")
 # The regression variables are the least correlated within totals & days.
 # The 2D/3D interpolations are strongly correlated with each other, and strongly correlated with with the Voronoi polygon extractions.
 # The days & total precipitation variables are fairly strongly correlated.
+# Total wet hours & hours below 10 & 14 are strongly correlated with rainy days.
+# Wet hours below 14 & total hours are correlated with total rainfall r = 0.66 - 0.76
+# Wet hours below 10 is correlated with total rainfall r = 0.61 - 0.68
 
-pairs(select(oak_sod, ends_with("rs"), ends_with("ds")),
+pairs(select(oak_sod, ends_with("rs"), ends_with("ds"), ends_with("wet")),
       lower.panel = panel.cor, diag.panel = panel.hist,
       main = "Oak Plots Temperature Variables")
 # Overall, correlations among variables are relatively weak with a few strong pairs. For specific pairs of variables it may be okay to use one from the dry season and one from the rainy season.
+# 
 
-pairs(select(oak_sod, ends_with("rs"), starts_with("rain")),
+pairs(select(oak_sod, ends_with("rs"), starts_with("rain"), ends_with("wet")),
       lower.panel = panel.cor, diag.panel = panel.hist,
       main = "Oak Plots RS Temperature and Rainfall Variables")
[…]

### 2016-02-12 analysis/repeated_measures.R
Update variable names.

@@ -18,19 +18,19 @@ stems$year <- as.factor(stems$year)
 
 #' Create bay laurel only data frame ####
 umca <- as.tbl(stems) %>%
-      select(plot, cluster, tag, species, status, slc, dbh, delta_dbh, dbh2_1_ratio, year, date) %>%
+      select(plotid, cluster, tag, species, status, slc, dbh, delta_dbh, dbh2_1_ratio, year, date) %>%
       filter(species == "UMCA", status == "Alive" | status == "Dead")
 summary(umca)
 umca <- droplevels(umca)
 
 #' Develop plot level summaries data frame and join to stems data frame ####
 umca <- left_join(umca, plot_umca <- umca %>%
-      select(plot, cluster, tag, status, slc, year) %>% 
-      group_by(plot, year) %>%
+      select(plotid, cluster, tag, status, slc, year) %>% 
+      group_by(plotid, year) %>%
       filter(slc != "NA" & status == "Alive") %>%
       summarise(avg_slc = mean(slc), tot_slc = sum(slc), 
                 live_dens = length(tag)), 
-      by = c("plot", "year"))
+      by = c("plotid", "year"))
 summary(umca)# One NA in the summarized leaf count variables
 filter(umca, is.na(avg_slc))
 filter(umca, plot == "YAHNG02")# Looks like leaf count probably not done/recorded
@@ -51,7 +51,7 @@ filter(umca, plot == "YAHNG02")# Looks like leaf count probably not done/recorde
 ##     2. E ~ N(0, var), errors "white noise"
 ##     3. These terms are independent in their distributions
 
-## The use of random effects prevent systematic variation between groups from ending up in the residual errors "white noise"
+## The use of random effects prevent systematic variation between groups from ending up in the residual errors ("white noise").
 
 #' Starting with plot level data b/c simpler w/fewer nested groups
 library(lme4)
@@ -67,24 +67,24 @@ library(lme4)
 
 #' Simplest models, with only one random effect (random intercept)
 m0 <- lmer(tot_slc ~ 1 + (1 | year), data = plot_umca)
-m0a <- lmer(tot_slc ~ 1 + (1 | plot), data = plot_umca)
[…]

### 2016-02-15 explore_veg_data.Rmd
Update untagged species DBH data.
Made more generalized data set of alternate species DBH and individual stem and plot-levels. I assigned sampling years of 2005 and 2014, corresponding to plot establishment and the most recent remeasurement.

@@
 
+# Untagged Overstory
 Now I need to examine the data for alternate and non-host species DBH. These are the stems that are untagged and meet a 5cm DBH threshold.
@@
 alt_sp_dbh$year <- year(alt_sp_dbh$date)
+head(alt_sp_dbh)
+
 alt_sp_dbh <- alt_sp_dbh %>% select(-PlotVisit_PlotVisitID, -Date) %>% 
@@
 unique(alt_sp_dbh$Dead)
+unique(alt_sp_dbh$year)
 # alt_sp_dbh$Dead <- ifelse(alt_sp_dbh$Dead == 0, "no", "yes")
@@
 summary(filter(alt_sp_dbh, year == 2003))
-unique(droplevels(filter(alt_sp_dbh, year == 2003)$PlotID))
+unique(droplevels(filter(alt_sp_dbh, year == 2003)$PlotID))# 32 plots
 pid2003 <- as.character(unique(filter(alt_sp_dbh, year == 2003)$PlotID))
@@
 summary(filter(alt_sp_dbh, year == 2004))
-unique(droplevels(filter(alt_sp_dbh, year == 2004)$PlotID))
+unique(droplevels(filter(alt_sp_dbh, year == 2004)$PlotID))# 120 plots
 pid2004 <- as.character(unique(filter(alt_sp_dbh, year == 2004)$PlotID))
@@
 summary(filter(alt_sp_dbh, year == 2005))
-unique(droplevels(filter(alt_sp_dbh, year == 2005)$PlotID))
+unique(droplevels(filter(alt_sp_dbh, year == 2005)$PlotID))# 25 plots
 pid2005 <- as.character(unique(filter(alt_sp_dbh, year == 2005)$PlotID))
@@
 alt_sp_45 <- filter(alt_sp_45, PlotID != "GANAY02")
-unique(alt_sp_2003$PlotID) %in% unique(alt_sp_45$PlotID)
+unique(alt_sp_2003$PlotID) %in% unique(alt_sp_45$PlotID)# LINS01 still there
 alt_sp_45 <- rbind(alt_sp_45, alt_sp_2003)
@@
 
+alt_species <- rbind(alt_sp_2005, alt_sp_2014)
+write.csv(alt_species, "analysis/data/alt-species-dbh-corrected.csv", row.names = F)
+```
+
+```{r group alt dbh to plot and year}
+alt_species
+alt_species$Dead[alt_species$Dead==0] <- "Alive"
+alt_species$Dead[alt_species$Dead==1] <- "Dead"
+alt_species <- rename(alt_species, status = Dead, plotid = PlotID, species = Species)
+alt_dbh_plot <- alt_species %>% 
+      group_by(plotid, species, status, year) %>% 
+      summarise(tot_dbh = sum(dbh, na.rm = T), avg_dbh = mean(dbh, na.rm = T),
+                tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
+                avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T),
+                abundance = length(dbh))
+write.csv(alt_dbh_plot, "analysis/data/alt-dbh-plot.csv", row.names = F)
+```
+
+```{r }
 # Create data frames of live and dead stems for '2005' and '2014' samplings.
@@
 ```{r make wide format species abundance}
+alt_species$PlotID <- tolower(alt_species$PlotID)
+unique(alt_species$Species)
+alt_species$Species[alt_species$Species=="arcto sp."] <- "arcto"
+alt_species$Species[alt_species$Species=="prunus sp."] <- "prunus"
+alt_species$Species[alt_species$Species=="unknown sp. 1"] <- "unk1"
+alt_species$Species[alt_species$Species=="quercus sp."] <- "qusp"
+alt_species
+
 unique(alt_sp_2014_live$Species)

### 2016-02-15 analysis/explore_data.R
Update code for assessing DBH data of tagged stems.
Edited to include other potential individual stem characteristics (because not all have a leaf count or canker). Assessed resulting data for duplicated tag numbers and made corrections; I believe if the data were exported from the current database this would not be necessary (i.e. I think the corrections have been made in the database). Developed a dataset for comparing plot establishment measurements (2005) to the most recent measurements (2014), calculating basal area and abundance by species and year for each plot.

@@ -299,66 +299,67 @@ stems <- droplevels(stems)
 class(stems)
 stems <- as.tbl(stems)
 
-#### Subset DBH data to create a data frame of remeasured stems: Round 2 ####
+# Subset DBH data to create a data frame of remeasured stems: Round 2 ####
 
 detach("package:lubridate", unload=TRUE)# b/c it has `union` function masking the one from `dplyr` that apparently behaves a little differently, creating a list instead of a data frame
 
 
-dbh2003 <- stems %>% select(plotid, tag, species, dbh, year, date) %>% 
+dbh2003 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp) %>% 
       filter(dbh>0, year == 2003) # these 2 are the same as in 2005
 # 11 observations
 
-dbh2004 <- stems %>% select(plotid, tag, species, dbh, year, date) %>%
+dbh2004 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp) %>%
       filter(dbh>0, year == 2004)
 # 3109 observations
-anti_join(dbh2003, dbh2004, by = c("plotid", "tag"))# 11 new DBH observations in 2004
+anti_join(dbh2003, dbh2004, by = c("plotid", "tag"))# 11 DBH observations different between 2003 and 2004
 dbh_a <- union(dbh2004, dbh2003)
 
-dbh2005 <- stems %>% select(plotid, tag, species, dbh, year, date) %>%
+dbh2005 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp) %>%
       filter(dbh>0, year == 2005)
 # 704 observations
 anti_join(dbh2005, dbh_a, by = c("plotid","tag")) # this indicates that the dbh data for 2003/04 and 2005 are nearly unique, 702 new DBH values at 109 plots in 2005
+# The 'union' with the 'anti_join' selects only the new records from 2005
 dbh_a <- union(dbh_a, anti_join(dbh2005, dbh_a, by = c("plotid","tag")))
 unique(dbh_a$plotid)# 204 plots on record at this point
[…]

### 2016-02-15 explore_veg_data.Rmd
Update wide formatting data code.

@@
 
-```{r group alt dbh to plot and year}
+```{r group alt species dbh to plot and year}
 alt_species
@@
 ```{r make wide format species abundance}
-alt_species$PlotID <- tolower(alt_species$PlotID)
-unique(alt_species$Species)
-alt_species$Species[alt_species$Species=="arcto sp."] <- "arcto"
-alt_species$Species[alt_species$Species=="prunus sp."] <- "prunus"
-alt_species$Species[alt_species$Species=="unknown sp. 1"] <- "unk1"
-alt_species$Species[alt_species$Species=="quercus sp."] <- "qusp"
-alt_species
+library(tidyr)
+alt_dbh_plot <- read.csv("analysis/data/alt-dbh-plot.csv")
+# alt_species$PlotID <- tolower(alt_species$PlotID)
+# unique(alt_species$species)
+# alt_species$Species[alt_species$Species=="arcto sp."] <- "arcto"
+# alt_species$Species[alt_species$Species=="prunus sp."] <- "prunus"
+# alt_species$Species[alt_species$Species=="unknown sp. 1"] <- "unk1"
+# alt_species$Species[alt_species$Species=="quercus sp."] <- "qusp"
+# head(alt_species)
 
-unique(alt_sp_2014_live$Species)
-alt_sp_2014_live$Species[alt_sp_2014_live$Species=="arcto sp."] <- "arcto"
-alt_sp_2014_live$Species[alt_sp_2014_live$Species=="prunus sp."] <- "prunus"
-alt_sp_2014_live$Species[alt_sp_2014_live$Species=="unknown sp. 1"] <- "unk1"
-alt_sp_2014_live$Species[alt_sp_2014_live$Species=="quercus sp."] <- "qusp"
+head(alt_dbh_plot)
+unique(alt_dbh_plot$species)
+alt_dbh_plot.wide <- as.tbl(alt_dbh_plot %>% 
+      select(plotid, species, status, year, abundance) %>% 
+      spread(species, abundance))
 
-unique(alt_sp_2005_live$Species)
-alt_sp_2005_live$Species[alt_sp_2005_live$Species=="quercus sp."] <- "qusp"
-alt_sp_2005_live$Species[alt_sp_2005_live$Species=="arcto sp."] <- "arcto"
-alt_sp_2005_live$Species[alt_sp_2005_live$Species=="prunus sp."] <- "prunus"
+# unique(alt_sp_2014_live$Species)
+# alt_sp_2014_live$Species[alt_sp_2014_live$Species=="arcto sp."] <- "arcto"
+# alt_sp_2014_live$Species[alt_sp_2014_live$Species=="prunus sp."] <- "prunus"
+# alt_sp_2014_live$Species[alt_sp_2014_live$Species=="unknown sp. 1"] <- "unk1"
+# alt_sp_2014_live$Species[alt_sp_2014_live$Species=="quercus sp."] <- "qusp"
+# 
+# unique(alt_sp_2005_live$Species)
+# alt_sp_2005_live$Species[alt_sp_2005_live$Species=="quercus sp."] <- "qusp"
+# alt_sp_2005_live$Species[alt_sp_2005_live$Species=="arcto sp."] <- "arcto"
+# alt_sp_2005_live$Species[alt_sp_2005_live$Species=="prunus sp."] <- "prunus"
 

### 2016-02-15 veg-change-assessment.Rmd
Add veg change assessment document.

---
title: "Vegetation Change Assessment - SOD Plots in Sonoma County"
author: "Whalen Dillon"
date: "February 15, 2016"
output: html_document

---

```{r}
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Alive", species == "UMCA"),
        ylab = "m^2",
        main = "Live UMCA Basal Area")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Dead", species == "UMCA"),
        ylab = "m^2",
        main = "Dead UMCA Basal Area")

boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
        ylab = "m^2",
        main = "Live QUAG Basal Area")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Dead", species == "QUAG"),
        ylab = "m^2",
        main = "Dead QUAG Basal Area")

ggplot(data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
       aes(year, tot_ba_m2, group = year)) +
      geom_boxplot()

boxplot(abundance ~ year, 
        data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
        ylab = "Number of stems",
        main = "Live QUAG Abundance")
boxplot(tot_ba_m2 ~ year, 
        data = basal_abund %>% filter(status == "Dead", species == "QUAG"),
        ylab = "Number of stems",
        main = "Dead QUAG Abundance")

ggplot(data = basal_abund,
       aes(species, tot_ba_m2, label = plotid)) +
      geom_boxplot() +
      facet_grid(status ~ year) +
      geom_text() +
      ylab("Basal Area (m^2)")

filter(dbhs, plotid == "SUGAR05", species == "QUAG")
# new giant dead stem in 2014, was probably dead entire study period

ggplot(data = basal_abund,
       aes(species, abundance)) +
      geom_boxplot() +
      facet_grid(status ~ year) +
      #geom_text() +
      ylab("Abundance")

ggplot(basal_abund %>% filter(status == "Alive"), aes(tot_ba_m2)) + 
      geom_histogram() +
      facet_grid(species ~ year)
ggplot(basal_abund %>% filter(status == "Alive"), aes(avg_ba_m2)) + 
      geom_histogram() +
      facet_grid(species ~ year)
ggplot(basal_abund %>% filter(status == "Alive"), aes(abundance)) + 
      geom_histogram() +
      facet_grid(species ~ year)
```

### 2016-02-15 analysis/species_diversity_calc.R
Update diversity metric calculations.
I added a calculation of Simpson's Evenness to the understory metrics since the observations are not in fact observations of numbers of individuals.

@@ -86,6 +86,7 @@ sort(unique(veg_us$us_species))# 79 species
 woody_spp <- c(woody_spp, "ceanothus", "cebe", "cecu", "frla", "garrya", "gemo", "juca", "mafu", "mane", "mimulus", "pisa", "quagXquke", "qube", "qusp", "quwi", "rhamnus", "rhcr", "ribes", "rica", "risa", "roca", "rosa", "rubus", "rudi", "salix", "syal", "vacc")
 veg_sub <- filter(veg_us, veg_us$us_species %in% woody_spp)# 56 species
 veg_sub <- droplevels(veg_sub)# drop the unused levels from veg species
+write.csv(veg_sub, "analysis/data/us-veg-sub.csv")
 
 # Calculate basic richness as number of different species in each plot each year
 us_richness <- full_join(
@@ -105,7 +106,7 @@ us_richness <- filter(us_richness,
 # ann06 and mroth03 are actually zeroes for understory species in 2011
 us_richness$us_rich11[us_richness$plotid=="ann06"] <- 0
 us_richness$us_rich11[us_richness$plotid=="mroth03"] <- 0
-# This leaves to legitimate NAs or understory species in 2011
+# This leaves two legitimate NAs for understory species in 2011
 
 filter(us_richness, is.na(us_rich05))# this should be zero for 2005
 us_richness$us_rich05[us_richness$plotid=="jlsp05"] <- 0
@@ -194,7 +195,7 @@ untagged_dbh_2005 %>% group_by(plotid, year) %>% summarise(untagged_rich = lengt
 untagged_dbh_2014 %>% group_by(plotid, year) %>% summarise(untagged_rich = length(unique(species)))
 
 
-# Calculate the richness of the tagged stems
+# Calculate the richness of the tagged stems ####
 stems <- as.tbl(stems)
 summary(stems)
 unique(stems$status)
@@ -263,6 +264,10 @@ os_richness <- filter(os_richness, plotid != "ann28", plotid != "halow01",
 # Join overstory richness and understory richness (counts) into single data frame
 richness <- full_join(os_richness, us_richness, by = c("plotid"))
 summary(richness)
+filter(richness, is.na(us_rich05))
+filter(richness, is.na(us_rich11))
+filter(richness, is.na(os_rich14))
+richness <- filter(richness, plotid != "bush01")
 
 
 #' ## Calculate Diversity Indices
[…]

### 2016-02-15 analysis/species_diversity_calc.R
Update data with new understory evenness metric.

@@ -86,7 +86,7 @@ sort(unique(veg_us$us_species))# 79 species
 woody_spp <- c(woody_spp, "ceanothus", "cebe", "cecu", "frla", "garrya", "gemo", "juca", "mafu", "mane", "mimulus", "pisa", "quagXquke", "qube", "qusp", "quwi", "rhamnus", "rhcr", "ribes", "rica", "risa", "roca", "rosa", "rubus", "rudi", "salix", "syal", "vacc")
 veg_sub <- filter(veg_us, veg_us$us_species %in% woody_spp)# 56 species
 veg_sub <- droplevels(veg_sub)# drop the unused levels from veg species
-write.csv(veg_sub, "analysis/data/us-veg-sub.csv")
+write.csv(veg_sub, "analysis/data/us-veg-sub.csv", row.names = F)
 
 # Calculate basic richness as number of different species in each plot each year
 us_richness <- full_join(

### 2016-02-16 analysis/species_diversity_calc.R
Update species data "arma" to "arcto".
This is an edit to generalize all the arctostaphylus to genus. Some had been identified as A. manzanita during the establishment phase, but people did not have the confidence for species-level identification at subsequent visits.

@@ -75,7 +75,7 @@ sort(unique(veg_us$us_species))
 # I want to match the species to those used by Sarah in her second chapter
 local_woody <- read.csv("analysis/data/presence_local_woody_n197.csv")
 summary(local_woody)
-unique(local_woody$spp.local)
+sort(unique(local_woody$spp.local))
 woody_spp <- as.character(local_woody$spp.local)
 
 sort(woody_spp) # only 29 species
@@ -86,6 +86,10 @@ sort(unique(veg_us$us_species))# 79 species
 woody_spp <- c(woody_spp, "ceanothus", "cebe", "cecu", "frla", "garrya", "gemo", "juca", "mafu", "mane", "mimulus", "pisa", "quagXquke", "qube", "qusp", "quwi", "rhamnus", "rhcr", "ribes", "rica", "risa", "roca", "rosa", "rubus", "rudi", "salix", "syal", "vacc")
 veg_sub <- filter(veg_us, veg_us$us_species %in% woody_spp)# 56 species
 veg_sub <- droplevels(veg_sub)# drop the unused levels from veg species
+sort(unique(veg_sub$us_species))
+# change all arctostaphylus identifications to the genus
+veg_sub[veg_sub=="arma"] <- "arcto"
+
 write.csv(veg_sub, "analysis/data/us-veg-sub.csv", row.names = F)
 
 # Calculate basic richness as number of different species in each plot each year
@@ -157,12 +161,12 @@ untagged_dbh_2014 <- as.tbl(untagged_dbh_2014 %>% rename(plotid = PlotID, specie
 untagged_dbh_2005$plotid <- tolower(untagged_dbh_2005$plotid)
 untagged_dbh_2014$plotid <- tolower(untagged_dbh_2014$plotid)
 
-unique(untagged_dbh_2005$species)
-unique(untagged_dbh_2014$species)
+sort(unique(untagged_dbh_2005$species))
+sort(unique(untagged_dbh_2014$species))
 untagged_dbh_2005$species <- as.character(untagged_dbh_2005$species)
 untagged_dbh_2014$species <- as.character(untagged_dbh_2014$species)
 untagged_dbh_2005[untagged_dbh_2005=="quercus sp."] <- "qusp"
-untagged_dbh_2005[untagged_dbh_2005=="arcto sp."] <- "arcto"
+untagged_dbh_2005[untagged_dbh_2005=="arcto sp." | untagged_dbh_2005=="arma"] <- "arcto"
[…]

### 2016-02-16 explore_veg_data.Rmd
Update overstory species list.
Changed species ID 'arma' to genus ID 'arcto' for overstory species data.

@@
 alt_species <- rbind(alt_sp_2005, alt_sp_2014)
+sort(unique(alt_species$species))
+alt_species[alt_species=="arma"] <- "arcto"
 write.csv(alt_species, "analysis/data/alt-species-dbh-corrected.csv", row.names = F)

### 2016-02-16 veg-change-assessment.Rmd
Update script.
Added read-in of data and initial examination of overstory/understory species data. Created some plots of species abundance: number of plots a species was recorded in, and the total number of a species recorded across the entire plot network. These are split by year, plot establishment (2005) and the most recent sampling season (2014).

@@
 date: "February 15, 2016"
-output: html_document
-
+output: 
+  html_document:
+    toc: true
 ---
 
-```{r}
+```{r read in data, echo=F, results='hide', message=FALSE}
+library(plyr, verbose = F); library(dplyr, verbose = F);
+library(ggplot2, verbose = F); library(tidyr, verbose = F)
+
+# Tagged Stems
+stems <- as.tbl(read.csv("analysis/data/stems_noBUSH01.csv"))
+stems
+stems$plotid <- tolower(stems$plotid)
+stems$species <- tolower(stems$species)
+
+# Overstory species data of tagged and untagged stems
+tag_dbh_plot <- as.tbl(read.csv("analysis/data/tag-dbh-plot.csv"))
+tag_dbh_plot
+tag_dbh_plot$plotid <- tolower(tag_dbh_plot$plotid)
+tag_dbh_plot$species <- tolower(tag_dbh_plot$species)
+tag_dbh_stem <- as.tbl(read.csv("analysis/data/tag-dbh_0514-corrected.csv"))
+tag_dbh_stem
+tag_dbh_stem$plotid <- tolower(tag_dbh_stem$plotid)
+tag_dbh_stem$species <- tolower(tag_dbh_stem$species)
+alt_dbh_plot <- as.tbl(read.csv("analysis/data/alt-dbh-plot.csv"))
+alt_dbh_plot
+alt_dbh_stem <- as.tbl(read.csv("analysis/data/alt-species-dbh-corrected.csv"))
+alt_dbh_stem
+
+unique(alt_dbh_plot$plotid) %in% unique(tag_dbh_plot$plotid)
+unique(alt_dbh_plot$plotid)[c(25,47,61,74,112,174,180)]
+# ann28: abandoned early b/c no hosts, remove
+# badge01: no hosts, kept in study
+# bush01, halow01, sweet01: abandoned early, remove
+# mcneil01, votru01: not sampled in 2014, remove
+alt_dbh_plot <- filter(alt_dbh_plot, plotid != "ann28", plotid != "bush01", plotid != "halow01", plotid != "sweet01", plotid != "mcneil01", plotid != "votru01")
+os_dbh_plot <- rbind(
+      select(tag_dbh_plot, plotid, species, status, year, everything()), 
+      select(alt_dbh_plot, -tot_dbh, -avg_dbh))
+sort(unique(os_dbh_plot$species))
+
+# Understory species data from microplots of veg transects
+veg_sub <- as.tbl(read.csv("analysis/data/us-veg-sub.csv"))
+veg_sub
+sort(unique(veg_sub$us_species))
+
+# Understory and overstory diversity/evenness calculations
+richness <- as.tbl(read.csv("analysis/data/os_us_diversity.csv"))
+richness
+
+# Importance Values
+imp_val_plot <- as.tbl(read.csv("analysis/data/species_dbh_impval_plot.csv"))
+imp_val_plot
+imp_val_plot <- select(imp_val_plot, -X)
+summary(imp_val_plot)
+```
+
+```{r number of species, echo=F}
+length(unique(os_dbh_plot$species))
+length(unique(veg_sub$us_species))
+ggplot(veg_sub %>% group_by(plotid, year, us_species), 
+       aes(us_species)) +
+      geom_bar()
+
+# Get total number of plots each understory species occurred in at least once
+ggplot(veg_sub %>% group_by(year, us_species) %>% 
+             summarise(abundance = length(plotid)), 
+       aes(us_species, abundance)) +
+      geom_bar(stat = "identity") +
+      facet_grid(year ~.)
+
+# Plot total number of plots each overstory species occurred in at least once
+ggplot(os_dbh_plot %>% filter(status == "Alive") %>% 
+             group_by(year, species) %>% 
+             summarise(abundance = length(plotid)), 
+       aes(species, abundance)) +
+      geom_bar(stat = "identity") +
+      facet_grid(year ~.)
+```
+
+```{r number of overstory species}
+# Plot total number of living stems for each overstory species by year
+ggplot(os_dbh_plot %>% filter(status == "Alive") %>% 
+             group_by(year, species), 
+       aes(species, abundance)) +
+      geom_bar(stat = "identity") +
+      facet_grid(year ~.)
+
+os_abund <- os_dbh_plot %>% 
+      filter(status == "Alive") %>%
+      group_by(year, species) %>% 
+      summarise(abundance = sum(abundance)) %>% 
+      spread(species, abundance)
+
+## same as above, but filtering out the super-abundant UMCA
+ggplot(os_dbh_plot %>% filter(status == "Alive", species != "umca") %>% 
+             group_by(year, species), 
+       aes(species, abundance)) +
+      geom_bar(stat = "identity") +
+      facet_grid(year ~.)
+```
+
+```{r plotting, echo=F}
+ggplot(data = os_dbh_plot,
+       aes(species, tot_ba_m2, label = plotid)) +
+      geom_boxplot() +
+      facet_grid(status ~ year) +
+      #geom_text() +
+      ylab("Basal Area (m^2)")
+
+filter(dbhs, plotid == "SUGAR05", species == "QUAG")
+# new giant dead stem in 2014, was probably dead entire study period
+
+ggplot(data = basal_abund,
+       aes(species, abundance)) +
+      geom_boxplot() +
+      facet_grid(status ~ year) +
+      #geom_text() +
+      ylab("Abundance")
+
+ggplot(basal_abund %>% filter(status == "Alive"), aes(tot_ba_m2)) + 
+      geom_histogram() +
+      facet_grid(species ~ year)
+ggplot(basal_abund %>% filter(status == "Alive"), aes(avg_ba_m2)) + 
+      geom_histogram() +
+      facet_grid(species ~ year)
+ggplot(basal_abund %>% filter(status == "Alive"), aes(abundance)) + 
+      geom_histogram() +
+      facet_grid(species ~ year)
+
 boxplot(tot_ba_m2 ~ year, 
@@
         main = "Dead QUAG Abundance")
-
-ggplot(data = basal_abund,
-       aes(species, tot_ba_m2, label = plotid)) +
-      geom_boxplot() +
-      facet_grid(status ~ year) +
-      geom_text() +
-      ylab("Basal Area (m^2)")
-
-filter(dbhs, plotid == "SUGAR05", species == "QUAG")
-# new giant dead stem in 2014, was probably dead entire study period
-
-ggplot(data = basal_abund,
-       aes(species, abundance)) +
-      geom_boxplot() +
-      facet_grid(status ~ year) +
-      #geom_text() +
-      ylab("Abundance")
-
-ggplot(basal_abund %>% filter(status == "Alive"), aes(tot_ba_m2)) + 
-      geom_histogram() +
-      facet_grid(species ~ year)
-ggplot(basal_abund %>% filter(status == "Alive"), aes(avg_ba_m2)) + 
-      geom_histogram() +
-      facet_grid(species ~ year)
-ggplot(basal_abund %>% filter(status == "Alive"), aes(abundance)) + 
-      geom_histogram() +
-      facet_grid(species ~ year)
 ```

### 2016-02-16 analysis/explore_data.R
Add 'location' variable to DBH data.

@@ -304,56 +304,56 @@ stems <- as.tbl(stems)
 detach("package:lubridate", unload=TRUE)# b/c it has `union` function masking the one from `dplyr` that apparently behaves a little differently, creating a list instead of a data frame
 
 
-dbh2003 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp) %>% 
+dbh2003 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp, location) %>% 
       filter(dbh>0, year == 2003) # these 2 are the same as in 2005
 # 11 observations
 
-dbh2004 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp) %>%
+dbh2004 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp, location) %>%
       filter(dbh>0, year == 2004)
 # 3109 observations
 anti_join(dbh2003, dbh2004, by = c("plotid", "tag"))# 11 DBH observations different between 2003 and 2004
 dbh_a <- union(dbh2004, dbh2003)
 
-dbh2005 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp) %>%
+dbh2005 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp, location) %>%
       filter(dbh>0, year == 2005)
-# 704 observations
+# 702 observations
 anti_join(dbh2005, dbh_a, by = c("plotid","tag")) # this indicates that the dbh data for 2003/04 and 2005 are nearly unique, 702 new DBH values at 109 plots in 2005
 # The 'union' with the 'anti_join' selects only the new records from 2005
 dbh_a <- union(dbh_a, anti_join(dbh2005, dbh_a, by = c("plotid","tag")))
 unique(dbh_a$plotid)# 204 plots on record at this point
 dbh_t1 <- dbh_a# data frame for DBH at time 1
 
-dbh2006 <- stems %>% select(plotid, tag, species, status, dbh, year, date, slc, canker, sod_dead, lide_lf_symp) %>%
[…]

### 2016-02-16 analysis/species_diversity_calc.R
Recalculate overstory diversity.
I filtered the tagged stems data to reflect only the stems rooted in the plot for the diversity calculations.

@@ -44,7 +44,7 @@ veg_us <- rbind(veg_us_2005, veg_us_2011)
 unique(veg_us$year)
 veg_us$year[veg_us$year < 2011] <- 2005 # Creates the two most complete veg sampling
 veg_us <- veg_us %>% rename(us_species = UnderstorySp, plotid = PlotID)
-unique(veg_us$us_species)
+sort(unique(veg_us$us_species))
 
 # Change species ID to be continuous (no spaces) for plants identified to genus
 veg_us$us_species <- as.character(veg_us$us_species)
@@ -70,6 +70,7 @@ veg_us[veg_us=="cotoneaster sp"] <- "cotoneaster"
 veg_us[veg_us=="quke x quwi"] <- "qukeXquwi"
 veg_us[veg_us=="quke x quch"] <- "qukeXquch"
 veg_us[veg_us=="rosa sp."] <- "rosa"
+veg_us[veg_us=="arma"] <- "arcto"
 sort(unique(veg_us$us_species))
 
 # I want to match the species to those used by Sarah in her second chapter
@@ -209,7 +210,7 @@ stems
 stems <- select(stems, -X) %>% rename(plotid = plot)
 stems$plotid <- tolower(stems$plotid)
 unique(stems$plotid)
-stem <- filter(stems, plotid != "bush01")
+stems <- filter(stems, plotid != "bush01")
 unique(stems$species)
 stems$species <- tolower(stems$species)
 stems$species[stems$species == "quke x quag"] <- "quagXquke"
@@ -219,7 +220,7 @@ stems <- droplevels(stems)
 summary(stems)
 
 
-live_tagged <- stems %>% filter(status == "Alive") %>% 
+live_tagged <- stems %>% filter(status == "Alive", location == "In") %>% 
       group_by(plotid, year, species) %>% 
       summarise(live_count = length(species))
 unique(live_tagged$species)

### 2016-02-17 analysis/species_diversity_calc.R
Recalculate diversity metrics.
I recalculated diversity metrics, particularly for the overstory limiting the DBH of included stems to be >= 2.0 cm.

@@ -71,23 +71,26 @@ veg_us[veg_us=="quke x quwi"] <- "qukeXquwi"
 veg_us[veg_us=="quke x quch"] <- "qukeXquch"
 veg_us[veg_us=="rosa sp."] <- "rosa"
 veg_us[veg_us=="arma"] <- "arcto"
+veg_us <- filter(veg_us, us_species != "dead wood")
 sort(unique(veg_us$us_species))
+write.csv(veg_us, "analysis/data/us-veg.csv", row.names = F)
 
 # I want to match the species to those used by Sarah in her second chapter
 local_woody <- read.csv("analysis/data/presence_local_woody_n197.csv")
 summary(local_woody)
 sort(unique(local_woody$spp.local))
 woody_spp <- as.character(local_woody$spp.local)
+woody_spp <- woody_spp[woody_spp != "arma" & woody_spp != "other"]
 
-sort(woody_spp) # only 29 species
+sort(woody_spp) # 27 species
 sort(unique(veg_us$us_species))# 79 species
 # It looks like I will want to add a few more to her list
 
 # Add other selected woody species to the list for inclusion in species richness/diversity/evenness calculations
 woody_spp <- c(woody_spp, "ceanothus", "cebe", "cecu", "frla", "garrya", "gemo", "juca", "mafu", "mane", "mimulus", "pisa", "quagXquke", "qube", "qusp", "quwi", "rhamnus", "rhcr", "ribes", "rica", "risa", "roca", "rosa", "rubus", "rudi", "salix", "syal", "vacc")
-veg_sub <- filter(veg_us, veg_us$us_species %in% woody_spp)# 56 species
+veg_sub <- filter(veg_us, veg_us$us_species %in% woody_spp)
 veg_sub <- droplevels(veg_sub)# drop the unused levels from veg species
-sort(unique(veg_sub$us_species))
+sort(unique(veg_sub$us_species))# 54
 # change all arctostaphylus identifications to the genus
 veg_sub[veg_sub=="arma"] <- "arcto"
 
@@ -95,15 +98,30 @@ write.csv(veg_sub, "analysis/data/us-veg-sub.csv", row.names = F)
 
 # Calculate basic richness as number of different species in each plot each year
 us_richness <- full_join(
+      veg_us %>% filter(year == 2005) %>% group_by(plotid) %>% 
[…]

### 2016-02-17 analysis/explore_data.R
Update data to DBH >= 2.0 cm.

@@ -450,8 +450,8 @@ boxplot(dbh ~ year,
 
 
 basal_abund <- dbhs %>% 
-      #filter(status == "Alive") %>% 
-      group_by(plotid, species, year, status) %>% 
+      filter(dbh >= 2.0) %>% 
+      group_by(plotid, species, year, status, location) %>% 
       summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
                 tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
                 abundance = length(dbh))

### 2016-02-17 explore_veg_data.Rmd
Revise species names.

@@
 veg_us$UnderstorySp[veg_us$UnderstorySp == "none"] <- "open"
-veg_us$UnderstorySp[veg_us$UnderstorySp == "arcto sp."] <- "arcto"
+veg_us$UnderstorySp[veg_us$UnderstorySp == "arcto sp." | veg_us$UnderstorySp == "arma"] <- "arcto"
 filter(veg_us, UnderstorySp == "vasp")
@@
 veg_us$UnderstorySp[veg_us$UnderstorySp == "garry sp."] <- "garrya sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "rose sp."] <- "rosa sp."
+veg_us$UnderstorySp[veg_us$UnderstorySp == "ribes sp."] <- "ribes"
 
@@
 summary(veg_us_summ)
-unique(veg_us_summ$UnderstorySp)
+sort(unique(veg_us_summ$UnderstorySp))
 

### 2016-02-17 veg-change-assessment.Rmd
Add narrative text, update data sources.

@@
 alt_dbh_plot <- filter(alt_dbh_plot, plotid != "ann28", plotid != "bush01", plotid != "halow01", plotid != "sweet01", plotid != "mcneil01", plotid != "votru01")
+
+# Overstory species w/measured DBH rooted in the plot
+os_dbh_in_plot <- rbind(
+      tag_dbh_plot %>% filter(location == "In") %>% 
+            select(plotid, species, status, year, everything(), -location), 
+      select(alt_dbh_plot, -tot_dbh, -avg_dbh))
+sort(unique(os_dbh_in_plot$species))
+
+# Overstory species w/measured DBH in or overhanging the plot
 os_dbh_plot <- rbind(
-      select(tag_dbh_plot, plotid, species, status, year, everything()), 
+      tag_dbh_plot %>% group_by(plotid, species, status, year) %>% 
+            summarise(tot_ba_m2 = sum(tot_ba_m2), avg_ba_m2 = sum(tot_ba_m2)/sum(abundance),
+                      abundance = sum(abundance)), 
       select(alt_dbh_plot, -tot_dbh, -avg_dbh))
-sort(unique(os_dbh_plot$species))
 
 # Understory species data from microplots of veg transects
+veg_us <- as.tbl(read.csv("analysis/data/us-veg.csv"))
+veg_us
+sort(unique(veg_us$us_species))
 veg_sub <- as.tbl(read.csv("analysis/data/us-veg-sub.csv"))
@@
 # Importance Values
-imp_val_plot <- as.tbl(read.csv("analysis/data/species_dbh_impval_plot.csv"))
-imp_val_plot
-imp_val_plot <- select(imp_val_plot, -X)
-summary(imp_val_plot)
+# imp_val_plot <- as.tbl(read.csv("analysis/data/species_dbh_impval_plot.csv"))
+# imp_val_plot# importance value by plot-species-year
+# imp_val_plot <- select(imp_val_plot, -X)
+# summary(imp_val_plot)
+# arrange(select(imp_val_plot, plotid, species, year, live.ct = live_count, all.live.ct = all_live_count, live.tot.dbh = live_tot_dbh, all.live.dbh = all_live_dbh, imp.val = imp_value) %>% filter(year == 2014), plotid)
+# unique(imp_val_plot$plotid)
+
+impt_vals <- rbind(tag_dbh_stem %>% filter(location == "In") %>% 
+                         select(plotid, species, dbh, status, year), 
+                   alt_dbh_stem) %>% 
+      group_by(plotid, species, status, year) %>% 
+      summarise(sp_abund = length(species), sp_dbh = sum(dbh)) %>% 
+      group_by(plotid, status, year) %>% 
+      mutate(total_abund = sum(sp_abund), total_dbh = sum(sp_dbh), 
+                imp_val =  0.5*( (sp_abund / total_abund) + (sp_dbh / total_dbh) ) )
+summary(impt_vals)
+unique(impt_vals$plotid)
 ```
+
+## Visualize Species Occurrence and Abundance
+We collected data on overstory and understory woody plant species to varying degrees throughout the study period.
+Overstory host species were assessed every year, so recruitment of new stems or death of existing stems was recorded, however, DBH measurements were initially only recorded at the first instance of a stem entering the study.
+We measured the DBH of untagged stems rooted in the plot and all tagged stems again in 2012.
+At this time we switched from an annual to bi-annual sampling cycle and decided that each time we visited the plots we would measure the DBH of all the overstory stems.
+Thus, the most recent DBH measurements are from the 2014 sampling season.
+
+The understory species data was collected at two time points, 2005 and 2011.
+In 2005, Xnumber transects with Ynumber microplots each...
+Prior to doing the understory vegetation assessment in 2011 we conducted an analysis to assess the effort necessary to sufficiently collect comparable data.
+We repeatedly randomly sampled 50% of the entire dataset from 2005 and compared univariate statistics between this and the full dataset.
+Correlations were very strong (r > 0.9) and rare species were seldom missing from the smaller dataset.
+
+During the 2011 season we used five transects across each plot with 0.5m radius microplots spaced at 1.3m intervals resulting in ~60 microplots per plot (12 per transect).
+Baselines were run between posts placed at plot plot establishment to clearly demarcate the plot edges.
+In some cases the distance between baselines was slightly greater or less than 15m, resulting in more or total number of microplots.
+At each microplot we recorded the occurrence of woody plant species that were rooted in or overhanging the microplot, and did not meet DBH requirements.
+
+Initially 80 "species" in the understory vegetation data set, however, this list included two unknowns and "dead wood." We grouped some species to the genus level due to low confidence in species level identification, e.g.
+*Arctostaphylus* and removed the "dead wood" observation, reducing this to 77 species.
 

### 2016-02-20 veg-change-assessment.Rmd
Add new plotting and rearrange data imports and editing.

--- 2016-02-15 veg-change-assessment.Rmd
@@
 date: "February 15, 2016"
-output: html_document
-
+output: 
+  html_document:
+    toc: true
 ---
 
-```{r}
+## Load Packages and Data
+```{r load packages, echo=F, results='hide', message=FALSE}
+library(plyr, verbose = F); library(dplyr, verbose = F);
+library(ggplot2, verbose = F); library(tidyr, verbose = F)
+```
+
+```{r load overstory tree data, echo=F, results='hide', message=FALSE}
+# Tagged Stems
+stems <- as.tbl(read.csv("analysis/data/stems_noBUSH01.csv"))
+stems
+stems$plotid <- tolower(stems$plotid)
+stems$species <- tolower(stems$species)
+
+# Overstory species data of tagged and untagged stems
+tag_dbh_plot <- as.tbl(read.csv("analysis/data/tag-dbh-plot.csv"))
+tag_dbh_plot
+tag_dbh_plot$plotid <- tolower(tag_dbh_plot$plotid)
+tag_dbh_plot$species <- tolower(tag_dbh_plot$species)
+tag_dbh_stem <- as.tbl(read.csv("analysis/data/tag-dbh_0514-corrected.csv"))
+tag_dbh_stem
+tag_dbh_stem$plotid <- tolower(tag_dbh_stem$plotid)
+tag_dbh_stem$species <- tolower(tag_dbh_stem$species)
+alt_dbh_plot <- as.tbl(read.csv("analysis/data/alt-dbh-plot.csv"))
+alt_dbh_plot
+alt_dbh_stem <- as.tbl(read.csv("analysis/data/alt-species-dbh-corrected.csv"))
+alt_dbh_stem
+
+dbh_stem <- rbind(select(tag_dbh_stem, plotid, species, dbh, status, year),
+                  alt_dbh_stem)
+dbh_stem$plotid <- as.factor(dbh_stem$plotid)
+dbh_stem$status <- as.factor(dbh_stem$plotid)
+dbh_stem$year <- as.factor(dbh_stem$year)
+
+unique(alt_dbh_plot$plotid) %in% unique(tag_dbh_plot$plotid)
+unique(alt_dbh_plot$plotid)[c(25,47,61,74,112,174,180)]
+# ann28: abandoned early b/c no hosts, remove
+# badge01: no hosts, kept in study
+# bush01, halow01, sweet01: abandoned early, remove
+# mcneil01, votru01: not sampled in 2014, remove
+alt_dbh_plot <- filter(alt_dbh_plot, plotid != "ann28", plotid != "bush01", plotid != "halow01", plotid != "sweet01", plotid != "mcneil01", plotid != "votru01")
+```
+
+```{r recalculate plot data, echo=F}
+# Recalculate tag-dbh-plot to exclude new large dead quag stem in 2014 at sugar05
+filter(tag_dbh_stem, plotid == "sugar05", year == 2014)
+tag_dbh_plot <- tag_dbh_stem %>% 
+      filter(dbh >= 2.0, tag != 7794) %>% 
+      group_by(plotid, species, year, status, location) %>% 
+      summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
+                tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
+                abundance = length(dbh))
+tag_dbh_plot <- ungroup(tag_dbh_plot)
+
+# Overstory species w/measured DBH rooted in the plot
+os_dbh_in_plot <- rbind(
+      tag_dbh_plot %>% filter(location == "In") %>% 
+            select(plotid, species, status, year, everything(), -location), 
+      select(alt_dbh_plot, -tot_dbh, -avg_dbh))
+sort(unique(os_dbh_in_plot$species))
+
+# Overstory species w/measured DBH in or overhanging the plot
+os_dbh_plot <- rbind(
+      tag_dbh_plot %>% group_by(plotid, species, status, year) %>% 
+            summarise(tot_ba_m2 = sum(tot_ba_m2), avg_ba_m2 = sum(tot_ba_m2)/sum(abundance),
+                      abundance = sum(abundance)), 
+      select(alt_dbh_plot, -tot_dbh, -avg_dbh))
+```
+
+```{r recalculate importance values, echo=F}
+# Importance Values
+# imp_val_plot <- as.tbl(read.csv("analysis/data/species_dbh_impval_plot.csv"))
+# imp_val_plot# importance value by plot-species-year
+# imp_val_plot <- select(imp_val_plot, -X)
+# summary(imp_val_plot)
+# arrange(select(imp_val_plot, plotid, species, year, live.ct = live_count, all.live.ct = all_live_count, live.tot.dbh = live_tot_dbh, all.live.dbh = all_live_dbh, imp.val = imp_value) %>% filter(year == 2014), plotid)
+# unique(imp_val_plot$plotid)
+
+impt_vals <- rbind(tag_dbh_stem %>% filter(location == "In", dbh >= 2.0) %>% 
+                         select(plotid, species, dbh, status, year), 
+                   alt_dbh_stem) %>% 
+      group_by(plotid, species, status, year) %>% 
+      summarise(sp_abund = length(species), 
+                sp_basal_area = sum(sum(pi*dbh^2/40000, na.rm = T))) %>% 
+      group_by(plotid, status, year) %>% 
+      mutate(total_abund = sum(sp_abund), 
+             total_basal_area = sum(sp_basal_area), 
+             imp_val =  
+                   0.5 * ( (sp_abund / total_abund) + 
+                                 (sp_basal_area / total_basal_area) ) )
+summary(impt_vals)
+impt_vals$plotid <- as.factor(impt_vals$plotid)
+impt_vals$species <- as.factor(impt_vals$species)
+impt_vals$year <- as.factor(impt_vals$year)
+unique(impt_vals$plotid)
+```
+
+```{r load understory veg data, echo=F, results='hide', message=FALSE}
+# Understory species data from microplots of veg transects
+veg_us <- as.tbl(read.csv("analysis/data/us-veg.csv"))
+veg_us
+sort(unique(veg_us$us_species))
+veg_sub <- as.tbl(read.csv("analysis/data/us-veg-sub.csv"))
+veg_sub
+sort(unique(veg_sub$us_species))
+```
+
+```{r load diversity metrics os and us}
+# Understory and overstory diversity/evenness metrics
+## These are based on the full set of understory species (n=77) and "overstory" species with DBH >= 2.0cm (tagged hosts) or DBH >= 5.0cm (untagged stems).
+richness <- as.tbl(read.csv("analysis/data/os_us_diversity.csv"))
+richness
+```
+
+## Visualize Species Occurrence and Abundance
+We collected data on overstory and understory woody plant species to varying degrees throughout the study period.
+Overstory host species were assessed every year, so recruitment of new stems or death of existing stems was recorded, however, DBH measurements were initially only recorded at the first instance of a tagged host stem entering the study (DBH ≥ 2.0cm).
+We measured the DBH of untagged stems rooted in the plot and all tagged stems again in 2012.
+At this time we switched from an annual to bi-annual sampling cycle and decided that each time we visited the plots we would measure the DBH of all the overstory stems.
+Thus, the most recent DBH measurements are from the 2014 sampling season.
+
+The understory species data was collected at two time points, 2005 and 2011.
+In 2005, Xnumber transects with Ynumber microplots each...
+Prior to doing the understory vegetation assessment in 2011 we conducted an analysis to assess the effort necessary to sufficiently collect comparable data.
+We repeatedly randomly sampled 50% of the entire dataset from 2005 and compared univariate statistics between this and the full dataset.
+Correlations were very strong (r > 0.9) and rare species were seldom missing from the smaller dataset.
+
+During the 2011 season we used five transects across each plot with 0.5m radius microplots spaced at 1.3m intervals resulting in ~60 microplots per plot (12 per transect).
+Baselines were run between posts placed at plot plot establishment to clearly demarcate the plot edges.
+In some cases the distance between baselines was slightly greater or less than 15m, resulting in more or total number of microplots.
+At each microplot we recorded the occurrence of woody plant species that were rooted in or overhanging the microplot, and did not meet DBH requirements.
+
+Initially there were 80 "species" in the understory vegetation data set as exported from the database, however, this list included two "unknowns" and "dead wood." We grouped some species to the genus level due to low confidence in species level identification, e.g.
+*Arctostaphylus* and removed the "dead wood" observation, reducing this to `r length(sort(unique(veg_us$us_species)))`.
+
+### Understory Species
+```{r number of understory species, echo=F}
+length(sort(unique(veg_us$us_species)))
+length(sort(unique(veg_sub$us_species)))
+
+# Total number of plots each understory species occurred in at least once
+ggplot(veg_us %>% group_by(year, us_species) %>% 
+             summarise(occurrence = length(plotid)), 
+       aes(us_species, occurrence, label = occurrence)) +
+      geom_bar(stat = "identity") +
+      geom_text(nudge_y = 3) +
+      facet_grid(year ~ .) +
+      ylab("# of plots") +
+      ggtitle("Understory Species Occurrence (all species)")
+ggplot(veg_us %>% group_by(year, us_species) %>% 
+             summarise(occurrence = length(plotid)) %>% 
+             filter(occurrence > 1), 
+       aes(us_species, occurrence, label = occurrence)) +
+      geom_bar(stat = "identity") +
+      geom_text(nudge_y = 3) +
+      facet_grid(year ~ .) +
+      ylab("# of plots") +
+      ggtitle("Understory Species Occurrence (2+ plots)")
+length(
+      unique(
+            (veg_us %>% group_by(year, us_species) %>% 
+             summarise(abundance = length(plotid)) %>% 
+             filter(abundance > 1))$us_species
+             )
+       )
+```
+
+### Overstory Species
+```{r number of overstory species rooted in the plot}
+length(sort(unique(os_dbh_in_plot$species)))
+summary(os_dbh_in_plot)
+filter(os_dbh_in_plot, plotid == "sugar05")# dropped large, long-dead QUAG first recorded in 2014
+
+# Total number of plots each overstory species occurred in at least once based on stems rooted in the plot
+ggplot(os_dbh_in_plot %>% 
+             filter(species != "unk1", species != "unknown sp.") %>% 
+             group_by(year, species, status) %>% 
+             summarise(occurrence = length(plotid)), 
+       aes(species, occurrence, label = occurrence)) +
+      geom_bar(stat = "identity") +
+      geom_text(nudge_y = 3) +
+      facet_grid(status ~ year) +
+      ylab("# of plots") +
+      ggtitle("Overstory Species Occurrence")
+
+# Total number of living stems for each overstory species by year across all plots
+ggplot(os_dbh_in_plot %>%
+             filter(species != "unk1", species != "unknown sp.") %>% 
+             group_by(year, species, status) %>% 
+             summarise(abundance = sum(abundance)), 
+       aes(species, abundance, label = abundance)) +
+      geom_bar(stat = "identity") +
+      facet_grid(status ~ year) +
+      ylab("abundance") +
+      ggtitle("Overstory Species Abundance") +
+      geom_text(aes(vjust = 0), nudge_y = 10)
+
+## same as above, but filtering out the super-abundant UMCA
+ggplot(os_dbh_in_plot %>% 
+             filter(species != "umca", species != "unk1", 
+                    species != "unknown sp.") %>% 
+             group_by(year, species, status) %>% 
+             summarise(abundance = sum(abundance)), 
+       aes(species, abundance, label = abundance)) +
+      geom_bar(stat = "identity") +
+      facet_grid(status ~ year) +
+      ylab("abundance") +
+      ggtitle("Overstory Species Abundance (minus UMCA)") +
+      geom_text(aes(vjust = 0), nudge_y = 6)
+
+# Total numbers only for tagged host species.
+I think for 'Dead' status it is probably only fair to look at tagged host species because I don't think the dead stems of untagged species were frequently recorded earlier in the study.
+## abundance
+ggplot(os_dbh_in_plot %>% 
+             filter(species == "lide" | species == "quag" | species == "quch" |
+                          species == "quke" | species == "quag x quwi" | 
+                          species == "quke x quag" | species == "umca") %>%
+             group_by(species, status, year) %>% 
+             summarise(abundance = sum(abundance)),
+       aes(species, abundance, label = abundance)) +
+      geom_bar(stat = "identity") +
+      facet_grid(status ~ year) +
+      ylab("abundance") +
+      ggtitle("Tagged Overstory Species Abundance") +
+      geom_text(aes(vjust = 0), nudge_y = 6)
+
+qusp_dbh_plot <- tag_dbh_stem %>% 
+      filter(species == "quke" | species == "quag" | species == "quch") %>%
+            group_by(plotid, species, status, year, sod_dead) %>% 
+            summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
+                tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
+                abundance = length(dbh))
+ggplot(qusp_dbh_plot, aes(species, abundance, label = abundance)) +
+      geom_bar(stat = "identity") +
+      facet_grid(sod_dead ~ year) +
+      ylab("abundance") +
+      ggtitle("Quercus Species SOD Killed and Living Abundance") #+
+      geom_text(nudge_y = 6)
+```
+
+#### Total basal area of tagged species rooted in the plot
+```{r basal area tagged species}
+ggplot(os_dbh_in_plot %>% 
+             filter(species == "lide" | species == "quag" | species == "quch" |
+                          species == "quke" | species == "quag x quwi" | 
+                          species == "quke x quag" | species == "umca") %>%
+             group_by(species, status, year) %>% 
+             summarise(basal_area = round(sum(tot_ba_m2), digits = 4)),
+       aes(species, basal_area, label = basal_area)) +
+      geom_bar(stat = "identity") +
+      facet_grid(status ~ year) +
+      ylab("basal area (m^2)") +
+      ggtitle("Tagged Overstory Species Basal Area") +
+      geom_text(aes(vjust = 0), nudge_y = 3)
+
+# boxplot of total basal area for tagged species
+ggplot(os_dbh_in_plot %>% 
+             filter(species == "lide" | species == "quag" | species == "quch" |
+                          species == "quke" | species == "quag x quwi" | 
+                          species == "quke x quag" | species == "umca"),
+       aes(species, tot_ba_m2, label = plotid)) +
+      geom_boxplot() +
+      facet_grid(status ~ year) +
+      ylab("basal area (m^2)") +
+      ggtitle("Tagged Overstory Species Basal Area") +
+      geom_text(nudge_y = .2)
+
+# Sudden oak death mortality for QUAG, QUKE, QUCH
+tag_dbh_stem$sod_dead[tag_dbh_stem$sod_dead==0] <- "alive"
+tag_dbh_stem$sod_dead[tag_dbh_stem$sod_dead==1] <- "killed"
+summary(tag_dbh_stem %>% 
+      filter(species == "quke" | species == "quag" | species == "quch"))
+qusp_dbh_plot <- tag_dbh_stem %>% 
+      filter(species == "quke" | species == "quag" | species == "quch") %>%
+            group_by(plotid, species, status, year, sod_dead) %>% 
+            summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
+                tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
+                abundance = length(dbh))
+
+ggplot(qusp_dbh_plot, aes(species, abundance, label = plotid)) +
+      geom_boxplot() +
+      facet_grid(sod_dead ~ year) +
+      ylab("# of stems") +
+      ggtitle("SOD Killed and Living Abundances (QUAG, QUKE, QUCH") #+
+      geom_text(nudge_y = .2)
+
+ggplot(qusp_dbh_plot, aes(species, tot_ba_m2, label = plotid)) +
+      geom_boxplot() +
+      facet_grid(sod_dead ~ year) +
+      ylab("basal area (m^2)") +
+      ggtitle("SOD Killed and Living Basal Area (QUAG, QUKE, QUCH") +
+      geom_text(nudge_y = .2)
+```
+
+#### Species Importance Values
+```{r importance values}
+impt_vals <- ungroup(impt_vals)
+unique(impt_vals$species)
+
+ggplot(impt_vals %>% 
+             filter(species == "lide" | species == "quag" | species == "quch" |
+                          species == "quke" | species == "quag x quwi" | 
+                          species == "quke x quag" | species == "umca"),
+       aes(species, imp_val, label = plotid)) +
+      geom_boxplot() +
+      facet_grid(status ~ year) +
+      ylab("importance value") +
+      ggtitle("Tagged Overstory Species Importance Value") #+
+      geom_text(nudge_y = .2)
+```
+
+#### Species Richness
+```{r diversity plots}
+shannons_os <- richness %>% gather(key = year, value = diversity, -plotid) %>% 
+      filter(year == "H.2005" | year == "H.2014")
+shannons_os$year <- as.factor(ifelse(shannons_os$year == "H.2005", 2005, 2014))
+summary(shannons_os)
+pielous_os <- richness %>% gather(key = year, value = evenness, -plotid) %>% 
+      filter(year == "J.2005" | year == "J.2014")
+pielous_os$year <- as.factor(ifelse(pielous_os$year == "J.2005", 2005, 2014))
+
+ggplot(shannons_os, aes(year, diversity, label = plotid)) +
+      geom_boxplot() +
+      # facet_grid(status ~ year) +
+      ylab("Shannon's Diversity Index") +
+      ggtitle("Overstory Species Diversity") #+
+      geom_text(nudge_y = .2)
+
+ggplot(pielous_os, aes(year, evenness, label = plotid)) +
+      geom_boxplot() +
+      # facet_grid(status ~ year) +
+      ylab("Pielou's Evenness Index") +
+      ggtitle("Overstory Species Evenness") #+
+      geom_text(nudge_y = .2)
+```
+No difference in diversity across the entire plot network.
+No difference in average evenness, but evenness does appear to decrease for a subset of plots?
+Or, these plots are simply identified as outliers in 2014, but not in 2005.
+I am going to look more closely at those plots with an evenness index < 0.25 in the next chunk.
+
+```{r evenness less than 0.25}
+filter(pielous_os, evenness < 0.25, year == 2005)
+filter(pielous_os, evenness < 0.25, year == 2014)
+
+# plots that only occur in the 2005 data
+anti_join(filter(pielous_os, evenness < 0.25, year == 2005),
+          filter(pielous_os, evenness < 0.25, year == 2014),
+          by = "plotid")
+filter(pielous_os, plotid == "votru01", year == 2014)# plot decommisioned in 2011
+filter(pielous_os, plotid == "cook02", year == 2014)# suspicious change in evenness, 0 to 0.34
+filter(pielous_os, plotid == "backm03", year == 2014)# evenness increased from 0.19 to 0.33
+filter(pielous_os, plotid == "athea02", year == 2014)# suspicious change in evenness, 0 to 0.62
+
+# plots that only occur in the 2014 data
+anti_join(filter(pielous_os, evenness < 0.25, year == 2014),
+          filter(pielous_os, evenness < 0.25, year == 2005),
+          by = "plotid")
+filter(pielous_os, plotid == "pfend02", year == 2005)# decrease from 0.43 to 0.23
+filter(pielous_os, plotid == "jroth03", year == 2005)# decrease from 0.57 to 0
+filter(pielous_os, plotid == "jlsp05", year == 2005)# decrease from 0.43 to 0
+filter(pielous_os, plotid == "hood02", year == 2005)# decrease from 0.30 to 0.25
+filter(pielous_os, plotid == "ann22", year == 2005)# decrease from 0.46 to 0
+```
+
+The large changes toward or away from zero values for evenness are suspicious, and warrant investigating the original data.
+It seems unlikely that we would go from a single overstory species to multiple overstory species, and vice-versa, from multiple overstory species to a single overstory species.
+
+```{r stems in low evenness plots}
+summary(dbh_stem)
+unique(filter(dbh_stem, plotid == "ann01")$species)
+## plot dominated by umca, psme also present
+unique(filter(dbh_stem, plotid == "ann01", year == 2005)$species)
+## only umca in 2005
+filter(dbh_stem, plotid == "ann01", year == 2014, species == "psme")
+## single very large psme recorded in 2014, so it should have been recorded at some point earlier; first obs 2012, so made linear estimate for 2005 based on these two points and added to database.
+```
+
+#### Other Plots
+```{r plotting, echo=F}
+ggplot(data = os_dbh_in_plot,
+       aes(species, tot_ba_m2, label = plotid)) +
+      geom_boxplot() +
+      facet_grid(status ~ year) +
+      #geom_text() +
+      ylab("Basal Area (m^2)")
+
+filter(dbhs, plotid == "SUGAR05", species == "QUAG")
+# new giant dead stem in 2014, was probably dead entire study period
+
+ggplot(data = os_dbh_in_plot,
+       aes(species, abundance)) +
+      geom_boxplot() +
+      facet_grid(status ~ year) +
+      #geom_text() +
+      ylab("Abundance")
+
+ggplot(os_dbh_in_plot %>% filter(status == "Alive"), aes(tot_ba_m2)) + 
+      geom_histogram() +
+      facet_grid(species ~ year)
+ggplot(os_dbh_in_plot %>% filter(status == "Alive"), aes(avg_ba_m2)) + 
+      geom_histogram() +
+      facet_grid(species ~ year)
+ggplot(os_dbh_in_plot %>% filter(status == "Alive"), aes(abundance)) + 
+      geom_histogram() +
+      facet_grid(species ~ year)
+
 boxplot(tot_ba_m2 ~ year, 
-        data = basal_abund %>% filter(status == "Alive", species == "UMCA"),
+        data = os_dbh_in_plot %>% filter(status == "Alive", species == "UMCA"),
         ylab = "m^2",
@@
 boxplot(tot_ba_m2 ~ year, 
-        data = basal_abund %>% filter(status == "Dead", species == "UMCA"),
+        data = os_dbh_in_plot %>% filter(status == "Dead", species == "UMCA"),
         ylab = "m^2",
@@
 boxplot(tot_ba_m2 ~ year, 
-        data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
+        data = os_dbh_in_plot %>% filter(status == "Alive", species == "QUAG"),
         ylab = "m^2",
@@
 boxplot(tot_ba_m2 ~ year, 
-        data = basal_abund %>% filter(status == "Dead", species == "QUAG"),
+        data = os_dbh_in_plot %>% filter(status == "Dead", species == "QUAG"),
         ylab = "m^2",

### 2016-02-20 climate-path-model.R
Add notes and edits from meeting with Beth.
We went through the model fits, assessing the residuals that looked crazy. She guided me in interpreting the patterns in the residuals.

I also began putting together a data-set to do stem-level analyses.

@@ -3,8 +3,8 @@
 #' 
 #+ load data and packages ####
 load("pathmodel_data.RData")
-library(plyr); library(dplyr); library(piecewiseSEM)
-library(nlme); library(lme4); library(lmerTest)
+library(plyr); library(dplyr); library(nlme); library(lme4)
+library(piecewiseSEM)
 
 #+ panel functions for pairs plots
 panel.cor <- function(x, y, digits = 2, cex.cor, ...)
@@ -62,6 +62,9 @@ pairs(select(oak_sod, ends_with("rs"), ends_with("ds"), ends_with("wet")),
       lower.panel = panel.cor, diag.panel = panel.hist,
       main = "Oak Plots Temperature Variables")
 # Overall, correlations among variables are relatively weak with a few strong pairs. For specific pairs of variables it may be okay to use one from the dry season and one from the rainy season.
+pairs(select(oak_sod, ends_with("ds"), ends_with("wet")),
+      lower.panel = panel.cor, diag.panel = panel.hist,
+      main = "Oak Plots Temperature Variables")
 # 
 
 pairs(select(oak_sod, ends_with("rs"), starts_with("rain"), ends_with("wet")),
@@ -88,8 +91,11 @@ pairs(na.omit(select(oak_sod, inf_oak_ct, tot_bay, tot_lfct, avg_lfct, H.2005, H
 # oak_sod$oak.inf <- cbind(oak_sod$inf_oak_ct, oak_sod$uninf_oak_ct)
 
 ## subset data frame and transform/rescale variables
-oakplots.sub <- select(oak_sod, plotid:tot_lfct, candens15m, twi15m, elev_m, veght_m, hrs_14_20_rs:rain_tot_v, rain_tot_psm:rain_days_v)
+oakplots.sub <- select(oak_sod, plotid:tot_lfct, candens15m, twi15m, elev_m, veght_m, hrs_14_20_rs:us_rich11, rain_tot_2d, rain_tot_v, rain_days_v, ends_with("wet"))
 summary(oakplots.sub)
+pairs(select(oakplots.sub, -plotid), lower.panel = panel.cor, diag.panel = panel.hist)
+
+# binomial variable for infected/uninfected oaks in each plot
 oakplots.sub$oak.inf <- cbind(oakplots.sub$inf_oak_ct, oakplots.sub$uninf_oak_ct)
 summary(filter(oak_sod, is.na(tot_lfct)))# lfct NAs are 0 bay plots
[…]

### 2016-02-20 analysis/repeated_measures.R
Edit code to match data variables.

@@ -18,19 +18,19 @@ stems$year <- as.factor(stems$year)
 
 #' Create bay laurel only data frame ####
 umca <- as.tbl(stems) %>%
-      select(plotid, cluster, tag, species, status, slc, dbh, delta_dbh, dbh2_1_ratio, year, date) %>%
+      select(plot, cluster, tag, species, status, slc, dbh, delta_dbh, dbh2_1_ratio, year, date) %>%
       filter(species == "UMCA", status == "Alive" | status == "Dead")
 summary(umca)
 umca <- droplevels(umca)
 
 #' Develop plot level summaries data frame and join to stems data frame ####
 umca <- left_join(umca, plot_umca <- umca %>%
-      select(plotid, cluster, tag, status, slc, year) %>% 
-      group_by(plotid, year) %>%
+      select(plot, cluster, tag, status, slc, year) %>% 
+      group_by(plot, year) %>%
       filter(slc != "NA" & status == "Alive") %>%
       summarise(avg_slc = mean(slc), tot_slc = sum(slc), 
                 live_dens = length(tag)), 
-      by = c("plotid", "year"))
+      by = c("plot", "year"))
 summary(umca)# One NA in the summarized leaf count variables
 filter(umca, is.na(avg_slc))
 filter(umca, plot == "YAHNG02")# Looks like leaf count probably not done/recorded
@@ -67,24 +67,24 @@ library(lme4)
 
 #' Simplest models, with only one random effect (random intercept)
 m0 <- lmer(tot_slc ~ 1 + (1 | year), data = plot_umca)
-m0a <- lmer(tot_slc ~ 1 + (1 | plotid), data = plot_umca)
+m0a <- lmer(tot_slc ~ 1 + (1 | plot), data = plot_umca)
 #' Add a second random effect where plot is nested within year
-m0b <- lmer(tot_slc ~ 1 + (1 | year) + (1 | plotid), data = plot_umca)
+m0b <- lmer(tot_slc ~ 1 + (1 | year) + (1 | plot), data = plot_umca)
  
 #' One predictor + random effect (intercept)
 m1 <- lmer(tot_slc ~ live_dens + (1 | year), data = plot_umca)
-m1a <- lmer(tot_slc ~ live_dens + (1 | plotid), data = plot_umca)
+m1a <- lmer(tot_slc ~ live_dens + (1 | plot), data = plot_umca)
[…]

### 2016-02-22 analysis/explore_data.R
Update code with revised data from database.
I exported data from the database that should incorporate many of the corrections I made during data exploration process. Some code was no longer needed and so I removed it or commented it out, so far.

@@ -3,161 +3,160 @@
 # Superspreaders chapter of dissertation      
 --------
 
-#### Read in all stem data, examine data structure ####
-#This is only the biologically related field data
+## Read in all tagged stem data, examine data structure ####
+# This is only the biologically related field data
 
-stems <- read.csv("analysis/data/stem_summary_query_2004-2014.txt")
-head(stems)
-summary(stems)
-str(stems)
+tagged_stems <- read.csv("analysis/data/TaggedStemQuery.csv")
+head(tagged_stems)
+summary(tagged_stems)
+str(tagged_stems)
+
+untagged_stems <- read.csv("analysis/data/Veg_TblVegDBH_Summary.csv")
+head(untagged_stems)
+summary(untagged_stems)
+str(untagged_stems)
 
 
-### Need to do some data cleaning and prep ####
+## Create 'Date' formatted date variable and year variable ####
+tagged_stems$date <- as.Date(as.character(tagged_stems$Date), format = "%m/%d/%Y")
+class(tagged_stems$date)
+untagged_stems$date <- as.Date(as.character(untagged_stems$Date), format = "%m/%d/%Y")
 
-## Create date variable as posix and integer/factor year variable ###
-stems$Date <- as.character(stems$Date)
-class(stems$Date)
 library(lubridate)
-stems$year <- year(as.Date(stems$Date, format = "%m/%d/%Y"))
-summary(stems)
-## Why are there NA's in the `year` variable?
-library(plyr)
-library(dplyr)
-stems_sub <- filter(stems, is.na(year))
-stems_sub2 <- filter(stems, PlotID == "SECRET02", SpeciesID == "QUAG", year == 2012)
-# So the 19 NA's for in the table are from the SECRET02 plot and belong to 2012. I am not sure why these did not query correctly in the database because the data is entered with the proper date...
-# I replaced the NA's for the `year` variable with 2012
-stems$year[is.na(stems$year)] <- 2012
-summary(stems)
-rm(stems_sub)
-rm(stems_sub2)
+tagged_stems$year <- year(tagged_stems$date)
+summary(tagged_stems)
+untagged_stems$year <- year(untagged_stems$date)
+summary(untagged_stems)
+detach("package:lubridate", unload=TRUE)
 
-## Currently unused code for string substitution and date calculation
[…]

### 2016-02-22 climate-path-model.R
Add narrative notes.
I added some notes from my meeting with Beth.

@@ -232,10 +232,18 @@ sem.coefs(oakplots.modlist, oakplots.sub, standardize = 'scale')
 # Path model using dry season temperature variable ####
 ## model set with sample year as random slope
 
-# I went through these with Beth on 2/18/2016 and we looked at the difficult to interpret residuals together. The patterning is likely due to the large number of 0 values for symptomatic leaf count, number of bay laurel, and number of infected oak stems. Specifically, the residuals are off from the fitted values because of the many zeroes in the data, seen especially for the values near zero.
+#' I went through these with Beth on 2/18/2016 and we looked at the difficult to interpret residuals together. The patterning is likely due to the large number of 0 values for symptomatic leaf count, number of bay laurel, and number of infected oak stems. Specifically, the residuals are off from the fitted values because of the many zeroes in the data, seen especially for the values near zero. The models seem to be properly parameterized with the available data, so Beth did not express too much concern with the residuals in the end. Sometimes funny looking residuals are just the way the data are, in this case, a lot of zero values.
+#' 
+#' Beth made a couple of suggestions
+#' 
+#'  1. Standardize the variables and rerun the models. Double check that the `scale` function does what it's supposed to do: rescale each variable to a mean of zero and variance of one. This will put the estimates into standard deviations from zero, but should not change the p-values.
+#'  2. Look at the different components of each multivariate regression model as bivariate regressions. This can help understand the contributions of each variable to the model and their relationship with the response.
+#'  3. Look up the calculation of Pearson's, and specifically how they are interpreted for a binomial model.
 
 # Pearson Residuals in Binomial Models ####
-# Pearson's residuals are reported for the binomial model, and Beth wanted me to check how these are calculated to make sure the interpretation was correct. 
+#' Pearson's residuals are reported for the binomial model, and Beth wanted me to check how these are calculated to make sure the interpretation was correct. A google search lead me here, [](http://data.princeton.edu/wws509/notes/c3s8.html), which says Pearson residuals are, "*the difference between observed and fitted values divided by an estimate of the standard deviation of the observed value*", and, "They take into account the fact that different observations have different variances, but they make no allowance for additional variation arising from estimation of the parameters, in the way studentized residuals in classical linear models do." So, they are not fully standardized residuals, but this is the common standardization for binomial models. 
+#' 
+#' For base GLM there are five types of residuals available, c("deviance", "pearson", "working","response", "partial")
 
[…]

### 2016-02-23 analysis/explore_data.R
Revert "Update data to DBH >= 2.0 cm."
This reverts commit a4e65d4d3e971deccd534d15ef2f19e91af576ab.

@@ -449,8 +449,8 @@ boxplot(dbh ~ year,
 
 
 basal_abund <- dbhs %>% 
-      filter(dbh >= 2.0) %>% 
-      group_by(plotid, species, year, status, location) %>% 
+      #filter(status == "Alive") %>% 
+      group_by(plotid, species, year, status) %>% 
       summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
                 tot_ba_m2 = sum(pi*dbh^2/40000, na.rm = T),
                 abundance = length(dbh))

### 2016-02-23 analysis/explore_data.R
Revert "Update code with revised data from database."
This reverts commit 37232c5ac76e4018f022b00cdbe6e871e0f138ab.

@@ -3,160 +3,161 @@
 # Superspreaders chapter of dissertation      
 --------
 
-## Read in all tagged stem data, examine data structure ####
-# This is only the biologically related field data
+#### Read in all stem data, examine data structure ####
+#This is only the biologically related field data
 
-tagged_stems <- read.csv("analysis/data/TaggedStemQuery.csv")
-head(tagged_stems)
-summary(tagged_stems)
-str(tagged_stems)
-
-untagged_stems <- read.csv("analysis/data/Veg_TblVegDBH_Summary.csv")
-head(untagged_stems)
-summary(untagged_stems)
-str(untagged_stems)
+stems <- read.csv("analysis/data/stem_summary_query_2004-2014.txt")
+head(stems)
+summary(stems)
+str(stems)
 
 
-## Create 'Date' formatted date variable and year variable ####
-tagged_stems$date <- as.Date(as.character(tagged_stems$Date), format = "%m/%d/%Y")
-class(tagged_stems$date)
-untagged_stems$date <- as.Date(as.character(untagged_stems$Date), format = "%m/%d/%Y")
+### Need to do some data cleaning and prep ####
 
+## Create date variable as posix and integer/factor year variable ###
+stems$Date <- as.character(stems$Date)
+class(stems$Date)
 library(lubridate)
-tagged_stems$year <- year(tagged_stems$date)
-summary(tagged_stems)
-untagged_stems$year <- year(untagged_stems$date)
-summary(untagged_stems)
-detach("package:lubridate", unload=TRUE)
+stems$year <- year(as.Date(stems$Date, format = "%m/%d/%Y"))
+summary(stems)
+## Why are there NA's in the `year` variable?
+library(plyr)
+library(dplyr)
+stems_sub <- filter(stems, is.na(year))
+stems_sub2 <- filter(stems, PlotID == "SECRET02", SpeciesID == "QUAG", year == 2012)
+# So the 19 NA's for in the table are from the SECRET02 plot and belong to 2012. I am not sure why these did not query correctly in the database because the data is entered with the proper date...
+# I replaced the NA's for the `year` variable with 2012
+stems$year[is.na(stems$year)] <- 2012
+summary(stems)
+rm(stems_sub)
+rm(stems_sub2)
 
-## Replace all `-9999` codes with NA ####
[…]

### 2016-02-23 analysis/explore_data.R
Make aesthetic changes to code.

@@ -56,7 +56,7 @@ str(stems)
 write.csv(stems,"analysis/data/all_stems.csv") # Write the revised stem data to CSV
 
 #### Further exploration of revised stem data ####
-stems <- read.csv("analysis/data/all_stems.csv") # Read in if necessary
+# stems <- read.csv("analysis/data/all_stems.csv") # Read in if necessary
 str(stems)
 summary(stems)
 summary(stems$species) # 5 host species listed, so this appears legit
@@ -215,7 +215,7 @@ stems %>% select(plot, cluster, tag, species, delta_dbh, year, year_dbh1, status
 # I changed values for 6 of 18 stems
 
 
-#### Create a plot-level data frame of infection status ####
+## Create a plot-level data frame of infection status ####
 stems <- read.csv("analysis/data/all_stems_corrected.csv")
 library(dplyr)
 
@@ -260,8 +260,8 @@ tmp <- plot_qusp %>%
       dcast(year + plot ~ species, value.var = "infected_oak_ct")
 head(tmp)
 
-# Checking large positive DBH changes ####
-## So at this point I think I have corrected all the negative dbh change values that I legitimately could in the database. I have gone back and exported the query again and use this file below, and then recycling code from the beginning of this document.
+## Checking large positive DBH changes ####
+# So at this point I think I have corrected all the negative dbh change values that I legitimately could in the database. I have gone back and exported the query again and use this file below, and then recycling code from the beginning of this document.
 stems <- read.csv("analysis/data/Stem_Summary_negDBH_corrected_201402.csv")
 summary(stems)
 str(stems)

### 2016-02-23 corrected-dbh.R
Add script for assessing and analyzing DBH data.

@@ -1,0 +1,221 @@
+#' # DBH Data
+#' 
+#' 
+
+## Read in all tagged stem data, examine data structure ####
+# This is only the biologically related field data
+
+tagged_stems <- read.csv("analysis/data/TaggedStemQuery.csv")
+head(tagged_stems)
+summary(tagged_stems)
+str(tagged_stems)
+
+untagged_stems <- read.csv("analysis/data/Veg_TblVegDBH_Summary.csv")
+head(untagged_stems)
+summary(untagged_stems)
+str(untagged_stems)
+
+
+## Create 'Date' formatted date variable and year variable ####
+tagged_stems$date <- as.Date(as.character(tagged_stems$Date), format = "%m/%d/%Y")
+class(tagged_stems$date)
+untagged_stems$date <- as.Date(as.character(untagged_stems$Date), format = "%m/%d/%Y")
+
+library(lubridate)
+tagged_stems$year <- year(tagged_stems$date)
+summary(tagged_stems)
+untagged_stems$year <- year(untagged_stems$date)
+summary(untagged_stems)
+detach("package:lubridate", unload=TRUE)
+
+## Replace all `-9999` codes with NA ####
+tagged_stems[tagged_stems == -9999] <- NA
+summary(tagged_stems)
+
+
+## Rename columns using `rename` funcion from `dplyr` ####
+library(plyr); library(dplyr)
+tagged_stems <- select(tagged_stems, -Date)
+names(tagged_stems) <- tolower(names(tagged_stems))
+
+tagged_stems <- rename(tagged_stems, plot_status = status,
+                       cluster = clusterid, species = speciesid,
+                       tag = tagnumber, initial_dbh = dbhcm,  
+                       stem_status = stemstatus, sod_killed = sod_dead, 
+                       umca_slc = sympleafcount, canker = cankerpresent,
+                       lide_slc = lideslc, lide_leaf_symp = lidefoliarsymp, 
+                       notes = stemsymptomnotes)
+str(tagged_stems)
+
+unique(tagged_stems$stem_status)
+tagged_stems$plotid <- tolower(tagged_stems$plotid)
+tagged_stems$species <- tolower(tagged_stems$species)
+tagged_stems$sod_killed <- ifelse(tagged_stems$sod_killed == 0, "No", "Yes")
[…]

### 2016-02-23 corrected-dbh.R
Fix missing plot data.
Data for the 2014 plot visit to MROTH03 was missing from the data base query; I don't know why. I added it to the stem dataframes. This commit also includes some additional code cut from the "explore_data.R" script.

@@ -67,7 +67,7 @@ summary(untagged_stems)
 ## Write out raw revised data sets ####
 write.csv(tagged_stems, "analysis/data/tagged_stems.csv", row.names = F)
 write.csv(untagged_stems, "analysis/data/untagged_stems.csv", row.names = F)
-write.csv(rbind(
+write.csv(all_stems <- rbind(
       select(tagged_stems, plotid, species, dbh, stem_status, date, year),
       untagged_stems),
       "analysis/data/all_stems.csv", row.names = F) 
@@ -188,11 +188,15 @@ tagged_stems <- as.tbl(left_join(tagged_stems,
                                  select(dbh_a, plotid, tag, dbh_entry),
                                  by = c("plotid", "tag")))
 summary(tagged_stems)
-summary(filter(tagged_stems, is.na(dbh_entry)))# NA are due to stems not meeting the 2.0 cm DBH requirement.
+summary(too_small_stems <- filter(tagged_stems, is.na(dbh_entry)))# NA are due to stems not meeting the 2.0 cm DBH requirement, except at least one curiosity for the initial_dbh
+filter(tagged_stems, is.na(dbh_entry), initial_dbh>2)
+# this is a plot that was abandoned very early in the study, may have been visited only one time
+# Take a closer look at these too small stems
+# fix(too_small_stems)
+# many are from abandoned plots, or are missing stems, and mostly did not meet DBH of 2.0cm upon first measurement
 
-# dropping stems that don't meet DBH threshold requirement
-tagged_stems <- filter(tagged_stems, dbh >= 2.0)
-write.csv(tagged_stems, "analysis/data/tagged-stems-corrected.csv")
+# dropping 194 stem records that don't meet DBH threshold requirement on their entry DBH measurement
+tagged_stems_corrected <- filter(tagged_stems, is.na(dbh_entry)==F)
 
 
 ## Explore DBH data: negative change in DBH ####
@@ -204,18 +208,87 @@ qplot(year, dbh_entry, data = dbh_a, geom="jitter", color = stem_status)
 # Dotplot of most recent dbh measurement
 qplot(stem_status, dbh, data = dbh2014, geom = "jitter", color = stem_status)
 
-# Data frame of tagged stems that were measured or remeasured in 2014
+
+
[…]

### 2016-02-23 analysis/explore_data.R
Cut code (moved to new script).
I cut out code on comparing species basal area and abundances and moved this to the "corrected-dbh.R" script file. I'm hoping this adds some clarity to each script, where this one focused on the initial data exploration resulting in numerous corrections to the database. The new script starts out with these corrected data.

@@ -391,73 +391,6 @@ no2014 <- anti_join(dbh_a, dbh2014, by = "tag")# this indicates that there were 
 
 summary(no2014); unique(no2014$plotid)# 73 plots; for proper comparison, whether using 2012 or 2014 data, I need to limit the set to plots that were in the establishment and final sampling.
 
-# Compare Basal & Species Abundances ####
-head(dbh_t1)
-dbh_t1 <- select(dbh_t1, -date) %>% arrange(plotid)
-dbh_t1$year <- 2005# change all to last sampling season included
-dbh_t1 <- droplevels(dbh_t1)
-head(dbh2014)
-dbh2014 <- select(dbh2014, -date) %>% arrange(plotid)
-dbh2014 <- droplevels(dbh2014)
-
-# Match plots between establishment and 2014 remeasurement
-length(unique(dbh_t1$plotid))# 204 plots
-length(unique(dbh2014$plotid))# 195 plots
-tmp1 <- droplevels(unique(dbh_t1$plotid))
-tmp2 <- droplevels(unique(dbh2014$plotid))
-setdiff(tmp1,tmp2)# identify plots not in 2014 data, note that one is SUGAR02
-filter(dbh_t1, plotid == "SUGAR02" | plotid == "SUGAR30")
-# change SUGAR02 to SUGAR30 for early data
-dbh_t1$plotid[dbh_t1$plotid == "SUGAR02"] <- "SUGAR30"
-
-tmp1 <- droplevels(unique(dbh_t1$plotid))
-tmp2 <- droplevels(unique(dbh2014$plotid))
-setdiff(tmp1,tmp2)
-dbh2014 <- rbind(dbh2014, select(dbh2012, -date) %>% 
-                       filter(plotid == "MROTH03"))
-# enter 2014 DBH values from database
-dbh2014$dbh[dbh2014$tag==1668] <- 68
-dbh2014$dbh[dbh2014$tag==1669] <- 49.2
-dbh2014$dbh[dbh2014$tag==1670] <- 56.1
-dbh2014$dbh[dbh2014$tag==1671] <- 51.2
-dbh2014$dbh[dbh2014$tag==1672] <- 66.9
-dbh2014$year <- 2014
-
-tmp1 <- droplevels(unique(dbh_t1$plotid))
-tmp2 <- droplevels(unique(dbh2014$plotid))
-setdiff(tmp1,tmp2)# reduced to 7 missing, 3 lost "late" in study
[…]

### 2016-02-23 corrected-dbh.R
Correct missing stems data.
I filtered out plots that were only assessed once and then abandoned, added missing data to the 'all_stems' data frame and CSV, and created dataframes of tagged and untagged stems each filtered so that there is a measurement from 2005 and 2014.

@@ -65,12 +65,12 @@ untagged_stems$species <- tolower(untagged_stems$species)
 summary(untagged_stems)
 
 ## Write out raw revised data sets ####
-write.csv(tagged_stems, "analysis/data/tagged_stems.csv", row.names = F)
-write.csv(untagged_stems, "analysis/data/untagged_stems.csv", row.names = F)
+write.csv(tagged_stems, "analysis/data/tagged_stems_revised.csv", row.names = F)
+write.csv(untagged_stems, "analysis/data/untagged_stems_revised.csv", row.names = F)
 write.csv(all_stems <- rbind(
       select(tagged_stems, plotid, species, dbh, stem_status, date, year),
       untagged_stems),
-      "analysis/data/all_stems.csv", row.names = F) 
+      "analysis/data/all_stems_revised.csv", row.names = F) 
 
 ## Further exploration of revised stem data ####
 # tagged_stems <- read.csv("analysis/data/tagged_stems.csv") # Read in if necessary
@@ -97,7 +97,6 @@ length(unique(filter(tagged_stems, species == "quke x quag")$tag))# 2 stems
 
 
 ## Subset DBH data to create a data frame of remeasured tagged_stems ####
-
 dbh2003 <- tagged_stems %>% 
       select(plotid, plot_status, species:dbh, stem_status, year) %>% 
       filter(dbh >= 2.0, year == 2003)
@@ -199,7 +198,20 @@ filter(tagged_stems, is.na(dbh_entry), initial_dbh>2)
 tagged_stems_corrected <- filter(tagged_stems, is.na(dbh_entry)==F)
 
 
-## Explore DBH data: negative change in DBH ####
+## Examine untagged stems data ####
+summary(untagged_stems)
+length(unique(untagged_stems$plotid))# 185 plots with untagged stems recorded
+length(unique(tagged_stems_corrected$plotid))# 203 plots
+unique(untagged_stems$plotid) %in% unique(tagged_stems_corrected$plotid)
+# there are three plots in the untagged stems data that aren't in the corrected tagged stems data
+unique(untagged_stems$plotid)[c(28,163,170)]
+# badge01 does not have any tagged hosts; ann28 and halow01 were abandoned in 2003/2004
+# remove the abandoned plots, retain badge01
[…]

### 2016-02-23 corrected-dbh.R
Correct missing stems data.
I filtered out plots that were only assessed once and then abandoned, added missing data to the 'all_stems' data frame and CSV, and created dataframes of tagged and untagged stems each filtered so that there is a measurement from 2005 and 2014.

@@ -65,12 +65,12 @@ untagged_stems$species <- tolower(untagged_stems$species)
 summary(untagged_stems)
 
 ## Write out raw revised data sets ####
-write.csv(tagged_stems, "analysis/data/tagged_stems.csv", row.names = F)
-write.csv(untagged_stems, "analysis/data/untagged_stems.csv", row.names = F)
+write.csv(tagged_stems, "analysis/data/tagged_stems_revised.csv", row.names = F)
+write.csv(untagged_stems, "analysis/data/untagged_stems_revised.csv", row.names = F)
 write.csv(all_stems <- rbind(
       select(tagged_stems, plotid, species, dbh, stem_status, date, year),
       untagged_stems),
-      "analysis/data/all_stems.csv", row.names = F) 
+      "analysis/data/all_stems_revised.csv", row.names = F) 
 
 ## Further exploration of revised stem data ####
 # tagged_stems <- read.csv("analysis/data/tagged_stems.csv") # Read in if necessary
@@ -97,7 +97,6 @@ length(unique(filter(tagged_stems, species == "quke x quag")$tag))# 2 stems
 
 
 ## Subset DBH data to create a data frame of remeasured tagged_stems ####
-
 dbh2003 <- tagged_stems %>% 
       select(plotid, plot_status, species:dbh, stem_status, year) %>% 
       filter(dbh >= 2.0, year == 2003)
@@ -121,7 +120,7 @@ anti_join(dbh2005, dbh_a, by = "tag")# this indicates that 2 stems were measured
 dbh_a <- union(dbh_a, anti_join(dbh2005, dbh_a, by = "tag"))
 # I think that this is the complete set of stems that makes up the "plot-establishment" data set. 
 dbh_establishment <- dbh_a
-write.csv(dbh_establishment, "analysis/data/tagged-stems-2003_2005.csv")
+write.csv(dbh_establishment, "analysis/data/tagged-stems-2003_2005.csv", row.names = F)
 
 
 dbh2006 <- tagged_stems %>% 
@@ -199,7 +198,20 @@ filter(tagged_stems, is.na(dbh_entry), initial_dbh>2)
 tagged_stems_corrected <- filter(tagged_stems, is.na(dbh_entry)==F)
 
 
-## Explore DBH data: negative change in DBH ####
+## Examine untagged stems data ####
+summary(untagged_stems)
+length(unique(untagged_stems$plotid))# 185 plots with untagged stems recorded
[…]

### 2016-02-23 corrected-dbh.R
Comment plotting code that is throwing an error.

@@ -319,15 +319,15 @@ qplot(year, abs_incr_grwth, data = tagged_dbh_changes, geom = "boxplot")
 
 
 ## Compare Basal & Species Abundances ####
-par(mfrow=c(2,1))
-boxplot(dbh ~ year, 
-        data = tagged_dbh_changes %>% 
-              filter(species == "UMCA", stem_status == "Alive"), 
-        main = "Live UMCA DBHs")
-boxplot(dbh ~ year, 
-        data = tagged_dbh_changes %>% 
-              filter(species == "UMCA", stem_status == "Dead"), 
-        main = "Dead UMCA DBHs")
+# par(mfrow=c(2,1))
+# boxplot(dbh ~ year, 
+#         data = tagged_dbh_changes %>% 
+#               filter(species == "umca", stem_status == "Alive"), 
+#         main = "Live UMCA DBHs")
+# boxplot(dbh ~ year, 
+#         data = tagged_dbh_changes %>% 
+#               filter(species == "umca", stem_status == "Dead"), 
+#         main = "Dead UMCA DBHs")
 
 qplot(factor(year), dbh, data = tagged_dbh_changes, geom = "jitter", color = stem_status, facets = .~species)
 

### 2016-02-24 sem__data_development.Rmd
Update using corrected stems data.
I used the corrected stems data developed in the 'corrected-dbh.R' script, which should be more accurate (i.e. have more corrections). I also removed the decomposition of topographic wetness index to its own script.

--- 2016-01-13 sem__data_development.Rmd
@@
 load("~/GitHub/superspreaders/stems_plots.RData")
+stems <- read.csv("analysis/data/tagged-stems-corrected.csv")
 
@@
 write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
-plots_env <- read.csv("analysis/data/env_gis_variables.csv")
+# plots_env <- read.csv("analysis/data/env_gis_variables.csv")
 ```
@@
 stems
-stems <- select(stems, -X) 
-stems <- rename(stems, plotid = plot) %>% filter(plotid != "BUSH01")
+# stems <- select(stems, -X) 
+unique(filter(stems, plot_status == "Abandoned")$plotid)
+stems <- stems %>% filter(plotid != "bush01", plotid != "ponti01", 
+                                plotid != "sweet01")
+stems$plotid <- droplevels(stems$plotid)
 
@@
 write.csv(summer_temps, "analysis/data/summer_temps_data.csv", row.names = F)
-write.csv(stems, "analysis/data/stems_noBUSH01.csv", row.names = F)
-rm(plots_temps); rm(plots_rain)
+# write.csv(stems, "analysis/data/stems_noBUSH01.csv", row.names = F)
+# rm(plots_temps); rm(plots_rain)
 ```
@@
 summer_temps <- read.csv("analysis/data/summer_temps_data.csv")
-stems <- read.csv("analysis/data/stems_noBUSH01.csv")
-```
-
-## Plot Contributing Area vs. Slope
-This is a suggestion from Ryan to plot the contributing area vs.
-the slope for each of the plots in order to tease out how either influences infection and hosts.
-Decomposing TWI to determine which factor has a greater influence on disease dynamics.
-```{r decomposing TWI influence}
-library(ggplot2)
-# par(mfrow = c(3,1))
-names(plots_env)
-summary(plots_env)
-## 'facc_ws15m' is the estimate from the 'r.watershed' function in GRASS GIS
-
-# Plot estimates from two different functions agains each other.
-qplot(flowacc15m, facc_ws15m, data = plots_env) #+
-      geom_text(aes(label = plotid))
-cor(plots_env$flowacc15m, plots_env$facc_ws15m)
-## There are 9 plots that really stand out from the others.
-## The r.watershed calculation provides similar, but generally smaller values.
-qplot(flowacc15m, facc_ws15m, data = filter(plots_env, flowacc15m > 300),
-      main = "'High' Flow Accumulation Value Plots") +
-      geom_text(aes(label = plotid))
-qplot(flowacc15m, facc_ws15m, data = filter(plots_env, flowacc15m < 300),
-      main = "'Low' Flow Accumulation Value Plots") +
-      geom_text(aes(label = plotid))
-## Much of the variability between the two estimates stems from a few plots.
-
-# Plot flow accumulation vs. slope
-qplot(slope15m, flowacc15m, data = plots_env) #+
-      geom_text(aes(label = plotid))
-## Can't see much pattern (if there is any) because of a few big values
-      cor(plots_env$slope15m, plots_env$flowacc15m)
-      ## negative, but weak
-
-qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m < 300), 
-      main = "Flow Accumulation <300") #+
-      geom_text(aes(label = plotid))
-## Generally, a shallower slope means higher flow accumulation, but it's not a strong pattern
-
-qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 300), 
-      main = "Flow Accumulation >300 ") #+
-      geom_text(aes(label = plotid))
-## The set of high flow accumulation plots shows an opposite general trend in the relationship between flow accumulation and slope.
-It's a small sample size, which leads me to consider the general trend from the large sample size being closer to the real relationship.
-
-qplot(slope15m, facc_ws15m, data = plots_env)
-cor(plots_env$slope15m, plots_env$facc_ws15m)
-## weakly negative correlation between slope and flow accumulation
-
-qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m < 300), 
-      main = "Flow Accumulation (r.watershed) <300") #+
-      geom_text(aes(label = plotid))
-qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 300), 
-      main = "Flow Accumulation (r.watershed) >300") #+
-      geom_text(aes(label = plotid))
-
-qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 50 & facc_ws15m < 200), 
-      main = "facc 50 - 200") +
-      geom_text(aes(label = plotid))
-qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 200 & facc_ws15m < 1000), 
-      main = "facc 200 - 1000") +
-      geom_text(aes(label = plotid))
-qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 1000), 
-      main = "facc >1000") +
-      geom_text(aes(label = plotid))
-
-# pairs(select(plots_env, -plotid, -starts_with("rain")), 
-#       lower.panel = panel.cor, diag.panel = panel.hist)
-# pairs(select(plots_env, -plotid, -starts_with("rain")) %>% 
-#             filter(flowacc15m < 5), 
-#       lower.panel = panel.cor, diag.panel = panel.hist,
-#       main = "Flow accumualtion < 5")
-# pairs(select(plots_env, -plotid, -starts_with("rain")) %>% 
-#             filter(flowacc15m > 100 & flowacc15m < 1000), 
-#       lower.panel = panel.cor, diag.panel = panel.hist,
-#       main = "Flow accumulation 100-1000")
-# pairs(select(plots_env, -plotid, -starts_with("rain")) %>% 
-#             filter(flowacc15m > 1000), 
-#       lower.panel = panel.cor, diag.panel = panel.hist,
-#       main = "Flow accumulation > 1000")
-
-plot(flowacc15m ~ slope15m, data = filter(plots_env, flowacc15m > 300))
-pairs(select(plots_env, flowacc15m, slope15m) %>% filter(flowacc15m > 300), 
-      lower.panel = panel.cor, diag.panel = panel.hist)
-
-library(manipulate)
-manipulate(
-      qplot(slope15m, flowacc15m, data = plots_env,
-           axes = axes,
-           cex = cex
-           #pch = if(pch) 19 else 1
-           ),
-      axes = checkbox(TRUE, "Show axes"),
-      cex = slider(0, 5, initial = 1, step = 0.1, label = "Point size"),
-      pch = button("Fill points")
-)
-
-dataplot <- function(dat){
-      name <- sys.call()[[2]]
-      vars <- as.list(names(dat))
-      e <- new.env()
-      e$data <- name
-      manipulate(
-            {
-                  form = as.formula(paste(y,x, sep = "~"))
-                  plot(form, data = dat, main = as.character(name), las = 1)
-                  e$form <- form
-            },
-            x = do.call(picker, c(vars, initial = vars[1])),
-            y = do.call(picker, c(vars, initial = vars[2]))
-      )
-      invisible(e)
-}
-f <- dataplot(plots_env)
+# stems <- read.csv("analysis/data/stems_noBUSH01.csv")
 ```
@@
 ```{r examine umca data}
-summary(plots_umca)
-# This only has number of infected, uninfected, and total, which must apply to living stems only. I want to analyze leaf counts.
+bay_laurel <- filter(stems, species == "umca") %>% 
+      select(plotid, cluster:stem_status, umca_slc, notes:dbh_entry)
+bay_laurel$species <- droplevels(bay_laurel$species)
+summary(bay_laurel) # This has stem-level data on bay laurel.
+length(unique(bay_laurel$plotid))# 183 plots with bay laurel
+# bay_laurel <- filter(bay_laurel, plot != "BUSH01")
 
-unique(plots_umca$plot)# 185 plots with bay laurel
-umca_plots <- filter(plots_umca, plot != "BUSH01")# now 184 plots
+# Data frame of UMCA plots only
+umca_plots <- bay_laurel %>%
+      # select(plotid, date, species, umca_slc, year, stem_status) %>%
+      group_by(plotid, year, species) %>%
+      filter(species == "umca", stem_status == "Alive") %>%
+      summarise(uninfected_bay_ct = length(which(umca_slc == 0)), 
+                infected_bay_ct = length(which(umca_slc > 0)), 
+                tot_bay = length(species)) 
+summary(umca_plots)
+
+# length(unique(umca_plots$plotid))# 183 plots with bay laurel
+# umca_plots <- filter(plots_umca, plot != "BUSH01")
 rm(plots_umca)
-summary(bay_laurel) # This has stem-level data on bay laurel.
-unique(bay_laurel$plot)
-bay_laurel <- filter(bay_laurel, plot != "BUSH01")
 ```
@@
 ```{r calculate plot level leaf counts}
-live_umca <- bay_laurel %>% filter(status == "Alive")
-filter(live_umca, is.na(slc))
+live_umca <- bay_laurel %>% filter(stem_status == "Alive")
+filter(live_umca, is.na(umca_slc))
 # Spot checking some of these shows that NAs are reasonable
 umca_plots <- left_join(
-      umca_plots, bay_laurel %>% filter(status == "Alive") %>% 
-            group_by(plot, year) %>% 
-            summarise(avg_lfct = mean(slc, na.rm = T),
-                      tot_lfct = sum(slc, na.rm = T)),
-      by = c("plot", "year"))
+      umca_plots, 
+      live_umca %>% 
+            group_by(plotid, year) %>% 
+            summarise(avg_lfct = mean(umca_slc, na.rm = T),
+                      tot_lfct = sum(umca_slc, na.rm = T)),
+      by = c("plotid", "year"))
 summary(umca_plots)
@@
 filter(umca_plots, is.na(avg_lfct))
-filter(bay_laurel, plot == "YAHNG02", status == "Alive")
+filter(bay_laurel, plotid == "yahng02", stem_status == "Alive")
 # I'm not sure why this wasn't excluded from the calculations with the na.rm = T option
@@
 summary(plots_env)
+plots_env$plotid <- tolower(plots_env$plotid)
 
@@
 # Join the environmental data to the bay laurel infection data
-umca_plots <- left_join(select(umca_plots, -species), plots_env, by = c("plot" = "plotid"))
+umca_plots <- left_join(select(umca_plots, -species), plots_env, by = "plotid")
 summary(umca_plots)
-filter(umca_plots, is.na(lon_utm))
+# filter(umca_plots, is.na(lon_utm))
 # NAs due to no topo data for abandoned SUGAR02; remove SUGAR02
-umca_plots <- na.omit(umca_plots)
-umca_plots <- rename(umca_plots, plotid = plot, sample_year = year)
-umca_plots$plotid <- tolower(umca_plots$plotid)
+# umca_plots <- na.omit(umca_plots)
+umca_plots <- rename(umca_plots, sample_year = year)
+# umca_plots$plotid <- tolower(umca_plots$plotid)
 
@@
 plot_qusp # right now this is broken down by oak species
-qusp_plots <- plot_qusp
-rm(plot_qusp)
-summary(qusp_plots)
+qusp4_plots <- stems %>%
+      select(plotid, date, species, canker, year, stem_status) %>%
+      filter(species == "quag" | species == "quke" | species == "quch" |
+                   species == "quwi", stem_status == "Alive") %>% 
+      group_by(plotid, year, species) %>% 
+      summarise(uninfected_oak_ct = length(which(canker==0)), 
+                infected_oak_ct = length(which(canker > 0)), 
+                tot_oak = length(species))
 
-oak_sod <- qusp_plots %>% group_by(plot, year) %>%
+# rm(plot_qusp)
+summary(qusp4_plots)
+
+oak_sod <- qusp4_plots %>% group_by(plotid, year) %>%
       summarise(uninf_oak_ct = sum(uninfected_oak_ct), 
@@
 oak_sod
-oak_sod <- rename(oak_sod, plotid = plot, sample_year = year)
-oak_sod$plotid <- tolower(oak_sod$plotid)
+oak_sod <- rename(oak_sod, sample_year = year)
+# oak_sod$plotid <- tolower(oak_sod$plotid)
 
-unique(umca_plots$plotid) %in% unique(oak_sod$plotid)
+# unique(umca_plots$plotid) %in% unique(oak_sod$plotid)
 
@@
 summary(oak_sod)
-filter(oak_sod, is.na(elev_m))
-# NAs are plots that were abandoned early in the study
-oak_sod <- filter(oak_sod, plotid != "duer01", plotid != "ponti01", plotid != "sweet01")
+filter(oak_sod, is.na(avg_tmax_rs))# NAs are plots sampled in 2003
+filter(oak_sod, is.na(H.2014))# NAs are plots lost before 2014
+summary(filter(oak_sod, is.na(us.J.2011)))# zero or missing data for understory richness
+# oak_sod <- filter(oak_sod, plotid != "duer01", plotid != "ponti01", plotid != "sweet01")
 
-summary(filter(oak_sod, is.na(avg_tmax_rs)))
-# Many of the NAs are due to the 2003 sample year
 oak_sod <- filter(oak_sod, sample_year != 2003)
@@
 # BUSH01 plot; I'm dropping that from the data set
-oak_sod <- filter(oak_sod, plotid != "bush01")
+# oak_sod <- filter(oak_sod, plotid != "bush01")
 summary(oak_sod)
@@
 # 2005 appears to be the first full year of observing all the plots
-filter(oak_sod, sample_year == 2005) # 170 records
-filter(oak_sod, sample_year == 2006) # 172 records
-filter(oak_sod, sample_year == 2007) # 171 records
-filter(oak_sod, sample_year == 2008) # 171 records
-filter(oak_sod, sample_year == 2009) # 170 records
-filter(oak_sod, sample_year == 2010) # 170 records
-filter(oak_sod, sample_year == 2011) # 170 records; final year for Sarah's 2nd chapter
-filter(oak_sod, sample_year == 2012) # 169 records
-filter(oak_sod, sample_year == 2014) # 168 records
-# In Sarah's analysis we end up using only 163 plots
+filter(oak_sod, sample_year == 2005) # 165 records
+filter(oak_sod, sample_year == 2006) # 167 records
+filter(oak_sod, sample_year == 2007) # 164 records
+filter(oak_sod, sample_year == 2008) # 163 records
+filter(oak_sod, sample_year == 2009) # 162 records
+filter(oak_sod, sample_year == 2010) # 161 records
+filter(oak_sod, sample_year == 2011) # 159 records; final year for Sarah's 2nd chapter
+filter(oak_sod, sample_year == 2012) # 156 records
+filter(oak_sod, sample_year == 2014) # 156 records
+# In Sarah's analysis we end up using 163 plots
 ```
@@
 summary(oak_sod)
+# drop rainy days interpolations, they aren't reliable; keep voronoi neighbor method
+oak_sod <- select(oak_sod, -rain_days_2d, -rain_days_3d, -rain_days_r)
 pairs(select(oak_sod, elev_m, starts_with("hrs"), starts_with("avg_t"), contains("rain_tot")), lower.panel = panel.cor, diag.panel = panel.hist, main = "Oak-SOD Rainfall & Temperature Data")
@@
 ```{r save data environment}
-save.image("~/GitHub/superspreaders/pathmodel_data_20160112.RData")
+# save.image("~/GitHub/superspreaders/pathmodel_data_20160112.RData")
 ```
+
+### Update path analysis data
+Updating the path analysis data with the temperature during wet days variables.
+```{r update data}
+# load("pathmodel_data_20160112.RData")
+wet_hrs <- read.csv("analysis/data/wet-days-temperature-variables.csv")
+save.image("~/GitHub/superspreaders/pathmodel_data.RData")
+```

### 2016-02-24 decompose-twi.Rmd
Add script for decomposition of topographic wetness index.

---
title: "Decomposing TWI Influence"
author: "Whalen Dillon"
date: "February 23, 2016"
output: html_document
---

## Plot Contributing Area vs. Slope
This is a suggestion from Ryan to plot the contributing area vs. the slope for each of the plots in order to tease out how either influences infection and hosts. Decomposing TWI to determine which factor has a greater influence on disease dynamics. 
```{r decomposing TWI influence}
library(ggplot2)
# par(mfrow = c(3,1))
names(plots_env)
summary(plots_env)
## 'facc_ws15m' is the estimate from the 'r.watershed' function in GRASS GIS

# Plot estimates from two different functions agains each other.
qplot(flowacc15m, facc_ws15m, data = plots_env) #+
      geom_text(aes(label = plotid))
cor(plots_env$flowacc15m, plots_env$facc_ws15m)
## There are 9 plots that really stand out from the others.
## The r.watershed calculation provides similar, but generally smaller values.
qplot(flowacc15m, facc_ws15m, data = filter(plots_env, flowacc15m > 300),
      main = "'High' Flow Accumulation Value Plots") +
      geom_text(aes(label = plotid))
qplot(flowacc15m, facc_ws15m, data = filter(plots_env, flowacc15m < 300),
      main = "'Low' Flow Accumulation Value Plots") +
      geom_text(aes(label = plotid))
## Much of the variability between the two estimates stems from a few plots.

# Plot flow accumulation vs. slope
qplot(slope15m, flowacc15m, data = plots_env) #+
      geom_text(aes(label = plotid))
## Can't see much pattern (if there is any) because of a few big values
      cor(plots_env$slope15m, plots_env$flowacc15m)
      ## negative, but weak

qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m < 300), 
      main = "Flow Accumulation <300") #+
      geom_text(aes(label = plotid))
## Generally, a shallower slope means higher flow accumulation, but it's not a strong pattern

qplot(slope15m, flowacc15m, data = filter(plots_env, flowacc15m > 300), 
      main = "Flow Accumulation >300 ") #+
      geom_text(aes(label = plotid))
## The set of high flow accumulation plots shows an opposite general trend in the relationship between flow accumulation and slope. It's a small sample size, which leads me to consider the general trend from the large sample size being closer to the real relationship.

qplot(slope15m, facc_ws15m, data = plots_env)
cor(plots_env$slope15m, plots_env$facc_ws15m)
## weakly negative correlation between slope and flow accumulation

qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m < 300), 
      main = "Flow Accumulation (r.watershed) <300") #+
      geom_text(aes(label = plotid))
qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 300), 
      main = "Flow Accumulation (r.watershed) >300") #+
      geom_text(aes(label = plotid))

qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 50 & facc_ws15m < 200), 
      main = "facc 50 - 200") +
      geom_text(aes(label = plotid))
qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 200 & facc_ws15m < 1000), 
      main = "facc 200 - 1000") +
      geom_text(aes(label = plotid))
qplot(slope15m, facc_ws15m, data = filter(plots_env, facc_ws15m > 1000), 
      main = "facc >1000") +
      geom_text(aes(label = plotid))

# pairs(select(plots_env, -plotid, -starts_with("rain")), 
#       lower.panel = panel.cor, diag.panel = panel.hist)
# pairs(select(plots_env, -plotid, -starts_with("rain")) %>% 
#             filter(flowacc15m < 5), 
#       lower.panel = panel.cor, diag.panel = panel.hist,
#       main = "Flow accumualtion < 5")
# pairs(select(plots_env, -plotid, -starts_with("rain")) %>% 
#             filter(flowacc15m > 100 & flowacc15m < 1000), 
#       lower.panel = panel.cor, diag.panel = panel.hist,
#       main = "Flow accumulation 100-1000")
# pairs(select(plots_env, -plotid, -starts_with("rain")) %>% 
#             filter(flowacc15m > 1000), 
#       lower.panel = panel.cor, diag.panel = panel.hist,
#       main = "Flow accumulation > 1000")

plot(flowacc15m ~ slope15m, data = filter(plots_env, flowacc15m > 300))
pairs(select(plots_env, flowacc15m, slope15m) %>% filter(flowacc15m > 300), 
      lower.panel = panel.cor, diag.panel = panel.hist)

library(manipulate)
manipulate(
      qplot(slope15m, flowacc15m, data = plots_env,
           axes = axes,
           cex = cex
           #pch = if(pch) 19 else 1
           ),
      axes = checkbox(TRUE, "Show axes"),
      cex = slider(0, 5, initial = 1, step = 0.1, label = "Point size"),
      pch = button("Fill points")
)

dataplot <- function(dat){
      name <- sys.call()[[2]]
      vars <- as.list(names(dat))
      e <- new.env()
      e$data <- name
      manipulate(
            {
                  form = as.formula(paste(y,x, sep = "~"))
                  plot(form, data = dat, main = as.character(name), las = 1)
                  e$form <- form
            },
            x = do.call(picker, c(vars, initial = vars[1])),
            y = do.call(picker, c(vars, initial = vars[2]))
      )
      invisible(e)
}
f <- dataplot(plots_env)
```

### 2016-02-24 analysis/explore_data.R
Update code aesthetics.

@@ -608,7 +608,7 @@ summarise(bay_laurel %>% filter(year == 2005), mdoy = max(doy))
 filter(bay_laurel, year == 2005, doy == 146)
 
 
-###### These may be useful functions for reference #### 
+## These may be useful functions for reference #### 
 matplot(cbind(avg_rain_year,avg_slc_year),type="b")
 matplot(cbind(rainy_days,avg_slc_year),type="b")
 

### 2016-02-25 esa-2016-abstract.Rmd
Add ESA 2016 abstract.

---
title: "Effects of diversity, topography, and interannual climate variability on pathogen spillover"
author: "Whalen Dillon"
date: "February 25, 2016"
output: html_document
---

## Background/Question/Methods
Variation in disease transmission is caused by heterogeneity of individuals, species, and landscapes across space and time. Becoming knowledgeable of the dynamic relationships among these factors is key for understanding and successfully managing any disease.

can be especially challenging for multihost disease systems due to challenges of spatial and temporal scale.  are frequently characterized by pathogen spillover, where a single amplifying host species drives disease transmission.

This is the case of the forest disease sudden oak death, where pathogen amplification on the highly competent alternate host species California bay laurel leads to spillover and infection of nearby oak trees.

When these hosts aggregate under environmental conditions favorable to the pathogen disease transmission may be amplified.

## Results/Conclusions

Results from our path model of disease observations aggregated to the plot level revealed that diversity moderated pathogen spillover through a relatively strong direct negative effect on oak infection. Potential inoculum load had a direct positive effect on oak infection moderated by temperature and topography. Temperature and rainfall had relatively weaker influences compared to diversity and inoculum load in the context of pathogen spillover. Although diversity also had a positive effect on inoculum load, its negative effect on oak infection is consistent with the dilution effect found in other studies. Topography also had substantial influence on diversity and inoculum load, where higher values of the wetness index tended to have lower values of diversity, but higher inoculum load. This is consistent with areas where moisture is likely to accumulate and persist providing a more favorable environment for *P. ramorum* sporulation.

### 2016-02-25 esa-2016-abstract.Rmd
Update background/methods.

@@
 Variation in disease transmission is caused by heterogeneity of individuals, species, and landscapes across space and time.
-Becoming knowledgeable of the dynamic relationships among these factors is key for understanding and successfully managing any disease.
-
-can be especially challenging for multihost disease systems due to challenges of spatial and temporal scale.
-are frequently characterized by pathogen spillover, where a single amplifying host species drives disease transmission.
-
-This is the case of the forest disease sudden oak death, where pathogen amplification on the highly competent alternate host species California bay laurel leads to spillover and infection of nearby oak trees.
-
-When these hosts aggregate under environmental conditions favorable to the pathogen disease transmission may be amplified.
+Multi-host disease systems are often characterized by pathogen spillover due to differences between host-pathogen interactions among the various hosts.
+Sudden oak death is a multi-host forest disease characterized by pathogen spillover, where a single host species (California bay laurel) amplifies the pathogen, which spills over to infect nearby oak trees.
+We applied a path analysis to analyze the direct and indirect effects of diversity, topography, and interannual climate variability on pathogen spillover for sudden oak death.
+The path analysis is a set of hypothesis built on findings from experiments and observations of the disease.
+We applied the path analysis framework to 10 years of data from a network of disease-monitoring plots distributed across a 275-km2 area in southeastern Sonoma County, California.
+During each year, we measured temperature, rainfall, pathogen load, and oak infection, while plant community diversity was measured once at each plot in this low-turnover landscape.
 
 ## Results/Conclusions
-
 Results from our path model of disease observations aggregated to the plot level revealed that diversity moderated pathogen spillover through a relatively strong direct negative effect on oak infection.

### 2016-02-25 sem-data-development.Rmd
Add code for calculating lagged rainfall variables.

--- 2016-02-24 sem__data_development.Rmd
@@
 
-# Untagged dbh data
-untagged_dbh_2005 <- read.csv("analysis/data/untagged_dbh2005_summary.csv")
-untagged_dbh_2014 <- read.csv("analysis/data/untagged_dbh2014_summary.csv")
+# Canopy cover from densiometer readings
 canopy_cover <- read.csv("analysis/data/plot_canopy_cover_dens.csv")
@@
 
-pairs(na.omit(select(oak_sod, c(3:4,7:8,9:10), H.2005, H.2014)),
+filter(oak_sod, is.na(elevation))
+oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01")
+filter(oak_sod, is.na(hrs_14_20_rs))
+oak_sod <- filter(oak_sod, plotid != "bush01")
+filter(oak_sod, is.na(H.2005))$plotid
+oak_sod <- filter(oak_sod, plotid != "mcneil03", plotid != "sumtv02",
+                  plotid != "votru01")
+filter(oak_sod, is.na(hrs_abv25rs))# I believe these locations actually did not have any temperatures above 25 C during the rainy season
+oak_sod$hrs_abv25rs[is.na(oak_sod$hrs_abv25rs)] <- 0
+
+length(unique(oak_sod$plotid))# 164 plots remaining
+
+pairs(na.omit(select(oak_sod, c(3,4,7:10), H.2005, H.2014)),
       lower.panel = panel.cor, diag.panel = panel.hist, 
       main = "Oak-SOD Infection & Diversity Data")
+```
+
+I'm going to add some lag variables, e.g. average rainfall for time t + t-1 or t-1 + t-2.
+```{r create lag variables}
+library(tidyr)
+summary(rain_total)
+unique(rain_total$sample_year)
+rain_lag <- as.tbl(select(rain_total, plotid, sample_year, rain_tot_v))
+rain_lag1 <- as.tbl(select(rain_total, plotid, sample_year))
+rain_lag1 <- rain_lag1 %>% 
+      mutate(sample_year_minus1 = sample_year - 1, 
+             sample_year_minus2 = sample_year - 2,
+             sample_year_minus3 = sample_year - 3) 
+rain_lag1 <- left_join(rain_lag1, rain_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(rain_tot_t_minus1 = rain_tot_v)
+rain_lag1 <- left_join(rain_lag1, rain_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(rain_tot_t_minus2 = rain_tot_v)
+rain_lag1 <- left_join(rain_lag1, rain_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(rain_tot_t_minus3 = rain_tot_v)
+rain_lag1 <- left_join(rain_lag1, rain_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+rain_lag1 <- select(rain_lag1, plotid, sample_year, starts_with("rain"))
+
+rain_lag1 <- rain_lag1 %>% 
+      mutate(avg_rain_t_minus1 = 
+                   rowMeans(cbind(rain_tot_v, rain_tot_t_minus1)),
+             avg_rain_t_minus2 =
+                   rowMeans(cbind(rain_tot_v, rain_tot_t_minus1, rain_tot_t_minus2)),
+             avg_rain_t1t2 = rowMeans(cbind(rain_tot_t_minus1, rain_tot_t_minus2)))
+summary(rain_lag1)
 ```
@@
 # load("pathmodel_data_20160112.RData")
+# Add untagged dbh data
+untagged_stems <- read.csv("analysis/data/untagged_stems_corrected.csv")
+untagged_dbh_2005 <- read.csv("analysis/data/untagged_dbh2005_summary.csv")
+untagged_dbh_2014 <- read.csv("analysis/data/untagged_dbh2014_summary.csv")
+
 wet_hrs <- read.csv("analysis/data/wet-days-temperature-variables.csv")

### 2016-02-25 veg-change-assessment.Rmd
Rename data frames.

@@
 # Tagged Stems
-stems <- as.tbl(read.csv("analysis/data/stems_noBUSH01.csv"))
-stems
-stems$plotid <- tolower(stems$plotid)
-stems$species <- tolower(stems$species)
+tagged_stems <- as.tbl(read.csv("analysis/data/tagged-stems-corrected.csv"))
+tagged_stems
+tagged_stems$plotid <- tolower(tagged_stems$plotid)
+tagged_stems$species <- tolower(tagged_stems$species)
 
@@
 tag_dbh_plot$species <- tolower(tag_dbh_plot$species)
+
 tag_dbh_stem <- as.tbl(read.csv("analysis/data/tag-dbh_0514-corrected.csv"))
@@
 tag_dbh_stem$species <- tolower(tag_dbh_stem$species)
+
 alt_dbh_plot <- as.tbl(read.csv("analysis/data/alt-dbh-plot.csv"))
 alt_dbh_plot
+
 alt_dbh_stem <- as.tbl(read.csv("analysis/data/alt-species-dbh-corrected.csv"))
@@
 
-dbh_stem <- rbind(select(tag_dbh_stem, plotid, species, dbh, status, year),
-                  alt_dbh_stem)
-dbh_stem$plotid <- as.factor(dbh_stem$plotid)
-dbh_stem$status <- as.factor(dbh_stem$plotid)
-dbh_stem$year <- as.factor(dbh_stem$year)
+all_stems <- as.tbl(read.csv("analysis/data/all_stems_corrected.csv"))
+summary(all_stems)
+      # rbind(select(tag_dbh_stem, plotid, species, dbh, status, year),
+      #             alt_dbh_stem)
+all_stems$plotid <- as.factor(all_stems$plotid)
+all_stems$status <- as.factor(all_stems$plotid)
+all_stems$year <- as.factor(all_stems$year)
 

### 2016-02-25 climate-path-model.R
Update models with standardized variables.

@@ -37,7 +37,7 @@ panel.hist <- function(x, ...)
 
 #+ examine oak plots data ####
 summary(oak_sod)
-oak_sod$plotid <- as.factor(oak_sod$plotid)
+# oak_sod$plotid <- as.factor(oak_sod$plotid)
 str(oak_sod)
 
 ## join wet-hours data
@@ -46,6 +46,10 @@ wet_hrs[is.na(wet_hrs)] <- 0# introduce a lot of zeroes for some columns
 oak_sod <- left_join(
       oak_sod, select(wet_hrs, -X), by = c("plotid","sample_year"))
 oak_sod$plotid <- as.factor(oak_sod$plotid)
+
+## drop 3d, PRISM, and regression rainfall variables based on discussion with Ryan; use a simple 2d spline interpolation or the Voronoi polygon values
+oak_sod <- select(oak_sod, -rain_tot_3d, -rain_tot_r, -rain_tot_psm)
+oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01", plotid != "bush01")
 
 #+ pairs plots of oak plots data ####
 pairs(na.omit(select(oak_sod, starts_with("rain"), ends_with("wet"))), 
@@ -57,11 +61,13 @@ pairs(na.omit(select(oak_sod, starts_with("rain"), ends_with("wet"))),
 # Total wet hours & hours below 10 & 14 are strongly correlated with rainy days.
 # Wet hours below 14 & total hours are correlated with total rainfall r = 0.66 - 0.76
 # Wet hours below 10 is correlated with total rainfall r = 0.61 - 0.68
+# To me, this indicates the possibility of using the wet-hours count combination variable of hours below 14 C in place of rainfall. This is only weakly correlated with the average temperature variables calculated for wet days. The number of hours below 14 C on wet days definitely captures the number of rainy days, as it should (r = 0.94).
 
-pairs(select(oak_sod, ends_with("rs"), ends_with("ds"), ends_with("wet")),
+pairs(na.omit(select(oak_sod, ends_with("rs"), ends_with("ds"), hrsblw14_wet, avgtmin_wet, avgtmax_wet)),
[…]

### 2016-02-25 esa-2016-abstract.Rmd
Update backgroun/methods.

@@
 Variation in disease transmission is caused by heterogeneity of individuals, species, and landscapes across space and time.
-Multi-host disease systems are often characterized by pathogen spillover due to differences between host-pathogen interactions among the various hosts.
+For example, multi-host disease systems are often characterized by pathogen spillover due to differences between host-pathogen interactions among the various hosts.
+Understanding the complexities of these disease dynamics at meaningful scales is a continuing challenge for disease ecologists.
 Sudden oak death is a multi-host forest disease characterized by pathogen spillover, where a single host species (California bay laurel) amplifies the pathogen, which spills over to infect nearby oak trees.
-We applied a path analysis to analyze the direct and indirect effects of diversity, topography, and interannual climate variability on pathogen spillover for sudden oak death.
-The path analysis is a set of hypothesis built on findings from experiments and observations of the disease.
-We applied the path analysis framework to 10 years of data from a network of disease-monitoring plots distributed across a 275-km2 area in southeastern Sonoma County, California.
+We used path analysis to analyze the direct and indirect effects of diversity, topography, and interannual climate variability on pathogen spillover (oak infection).
+The hypothesized relationships laid out by our path model for sudden oak death were based on published findings from experiments and observations of the disease during the past 15 years.
+We applied our path analysis framework to 10 years of data from a network of disease-monitoring plots distributed across a 275-km2 area in southeastern Sonoma County, California.
 During each year, we measured temperature, rainfall, pathogen load, and oak infection, while plant community diversity was measured once at each plot in this low-turnover landscape.

### 2016-02-25 climate-path-model.R
Update models with standardized variables.

@@ -111,25 +111,26 @@ pairs(select(oakplots.sub, -plotid), lower.panel = panel.cor, diag.panel = panel
 oakplots.sub$oak.inf <- cbind(oakplots.sub$inf_oak_ct, oakplots.sub$uninf_oak_ct)
 summary(filter(oak_sod, is.na(tot_lfct)))# lfct NAs are 0 bay plots
 oakplots.sub$tot_lfct[is.na(oakplots.sub$tot_lfct)] <- 0
-oakplots.sub$avg_lfct[is.na(oakplots.sub$avg_lfct)] <- 0
+# oakplots.sub$avg_lfct[is.na(oakplots.sub$avg_lfct)] <- 0
 
 oakplots.sub$rain_tot_v.cm <- oakplots.sub$rain_tot_v/10
-oakplots.sub$rain_tot_2d.cm <- oakplots.sub$rain_tot_2d/10
+# oakplots.sub$rain_tot_2d.cm <- oakplots.sub$rain_tot_2d/10
 # oakplots.sub$rain_tot_3d.cm <- oakplots.sub$rain_tot_3d/10
 
 oakplots.sub$dys_14_20_rs <- oakplots.sub$hrs_14_20_rs/24
-oakplots.sub$dys_blw10_rs <- oakplots.sub$hrs_blw10rs/24
-oakplots.sub$dys_abv25_rs <- oakplots.sub$hrs_abv25rs/24
+# oakplots.sub$dys_blw10_rs <- oakplots.sub$hrs_blw10rs/24
+# oakplots.sub$dys_abv25_rs <- oakplots.sub$hrs_abv25rs/24
 # oakplots.sub$tot_lfct.10 <- oakplots.sub$tot_lfct/10
 oakplots.sub$tot_lfct.100 <- oakplots.sub$tot_lfct/100
 oakplots.sub$tot_lfct.log <- log1p(oakplots.sub$tot_lfct)
 oakplots.sub$tot_bay.log <- log1p(oakplots.sub$tot_bay)
+oakplots.sub$twi15m.log <- log(oakplots.sub$twi15m)
 # oakplots.sub$avg_lfct.log <- log1p(oakplots.sub$avg_lfct)
 # oakplots.sub$sample_year <- as.factor(oakplots.sub$sample_year)
 
 ## rough path-model 1, how does this work ####
 oakplots.modlist <- list(
-      lme(tot_bay.log ~ twi15m + H.2005, random = 1|sample_year/plotid, 
[…]

### 2016-02-25 esa-2016-abstract.Rmd
Update results/conclusions.

@@
 ## Results/Conclusions
-Results from our path model of disease observations aggregated to the plot level revealed that diversity moderated pathogen spillover through a relatively strong direct negative effect on oak infection.
-Potential inoculum load had a direct positive effect on oak infection moderated by temperature and topography.
-Temperature and rainfall had relatively weaker influences compared to diversity and inoculum load in the context of pathogen spillover.
-Although diversity also had a positive effect on inoculum load, its negative effect on oak infection is consistent with the dilution effect found in other studies.
-Topography also had substantial influence on diversity and inoculum load, where higher values of the wetness index tended to have lower values of diversity, but higher inoculum load.
-This is consistent with areas where moisture is likely to accumulate and persist providing a more favorable environment for *P.
-ramorum* sporulation.
+We assessed our initial path model using measurements summarized for each plot.
+Our analysis indicated that diversity reduced pathogen spillover through a relatively strong direct negative effect on oak infection, however, this was tempered by more diverse plots also having a higher pathogen load.
+Pathogen load had the strongest direct positive effect on oak infection, which was countered by direct negative influences from temperature, and interestingly, bay laurel density.
+Our temperature variable was calcuated as the averaged daily maximum temperature during June-September preceding the spring sampling season, a time when bay laurel will drop damaged leaves.
+Our hypothesis was that higher temperatures leads to more leaf abscision and fewer infections persisting until the wet season.
+We tested the number of rainy days during November-May for each sampling season as our initial rainfall variable, because persistent moisture is more favorable for pathogen reproduction.
+However, this had a relatively weak effect on pathogen load.
+So far, we have confirmed a dilution effect via diversity reducing likelihood of oak infection, pathogen load has the single greatest effect on pathogen spillover, and discovered a surprising relationship between bay laurel and oak infection warranting further exploration.

### 2016-03-14 sem-data-development.Rmd
Add lagged temperature and rainfall variables.

--- 2016-02-24 sem__data_development.Rmd
@@
 
-# Untagged dbh data
-untagged_dbh_2005 <- read.csv("analysis/data/untagged_dbh2005_summary.csv")
-untagged_dbh_2014 <- read.csv("analysis/data/untagged_dbh2014_summary.csv")
-canopy_cover <- read.csv("analysis/data/plot_canopy_cover_dens.csv")
+# Canopy cover from densiometer readings
+## I've decided to not use these since we have high-resolution LiDAR data of canopy density
+# canopy_cover <- read.csv("analysis/data/plot_canopy_cover_dens.csv")
 ```
@@
 summary(plots_vegtopo)
-plots_vegtopo <- select(plots_vegtopo, -cat) %>% rename(lon_utm = X, lat_utm = Y)
+plots_vegtopo <- select(plots_vegtopo, -cat, -flowacc15m, -facc_ws15m, -slope15m) %>% rename(lon_utm = X, lat_utm = Y)
 ```
@@
 summary(rs_temps)
-rs_temps <- select(rs_temps, -X)
+# rs_temps <- select(rs_temps, -X)
 
@@
 summary(summer_temps)
-summer_temps <- select(summer_temps, -X)
+# summer_temps <- select(summer_temps, -X)
+
+wet_hrs <- read.csv("analysis/data/wet-days-temperature-variables.csv")
 
@@
 load("~/Documents/soco_ppt/plot_ppt_estimates.RData")
-summary(ppt_days)
-summary(ppt_total)
+rm(ppt_days)
+rm(ppt_total)
+summary(rainfall)# use newly calculated values
+rainfall$plotid <- tolower(rainfall$plotid)
 # plots_rain <- read.csv("analysis/data/plot202rain.csv")
@@
 #       diag.panel = panel.hist)
-ppt_days@data <- rename(ppt_days@data, plotid = PLOT_ID)
-ppt_total@data <- rename(ppt_total@data, plotid = PLOT_ID)
 ```
@@
 plots_env$veght_m <- plots_env$veght15m * .3048
+plots_env <- select(plots_env, -elevation, -elev15m, -veght15m, -tmi) %>% 
+      rename(plotid = PLOT_ID)
+plots_env$plotid <- tolower(plots_env$plotid)
 rm(plots_vegtopo)
@@
 library(tidyr)
-rain2d_total <- ppt_total@data %>% select(plotid, ends_with("2d"))
-names(rain2d_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain2d_total <- rain2d_total %>% gather(year, rain_tot_2d, -plotid)
+rain2d_rs_total <- rainfall@data %>% 
+      select(plotid, contains("rs"),  -ends_with("v"))
+names(rain2d_rs_total) <- c("plotid",paste(c(2004:2014), sep = ","))
+rain2d_rs_total <- rain2d_rs_total %>% gather(year, rain2d_total, -plotid)
 
-rain3d_total <- ppt_total@data %>% select(plotid, ends_with("3d"))
-names(rain3d_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain3d_total <- rain3d_total %>% gather(year, rain_tot_3d, -plotid)
+rain2d_ss_total <- rainfall@data %>% 
+      select(plotid, contains("ss"),  -ends_with("v"))
+names(rain2d_ss_total) <- c("plotid",paste(c(2004:2014), sep = ","))
+rain2d_ss_total <- rain2d_ss_total %>% gather(year, rain2d_total, -plotid)
 
-rain_v_total <- ppt_total@data %>% select(plotid, ends_with("v"))
-names(rain_v_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain_v_total <- rain_v_total %>% gather(year, rain_tot_v, -plotid)
+# rain3d_total <- ppt_total@data %>% select(plotid, ends_with("3d"))
+# names(rain3d_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+# rain3d_total <- rain3d_total %>% gather(year, rain_tot_3d, -plotid)
 
-rain_r_total <- ppt_total@data %>% select(plotid, ends_with("r"))
-names(rain_r_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain_r_total <- rain_r_total %>% gather(year, rain_tot_r, -plotid)
+rain_rs_v_total <- rainfall@data %>% 
+      select(plotid, ends_with("v")) %>% 
+      select(plotid, contains("rs")) %>%
+      select(plotid, starts_with("rain"))
+names(rain_rs_v_total) <- c("plotid",paste(c(2004:2014), sep = ","))
+rain_rs_v_total <- rain_rs_v_total %>% gather(year, rain_v_total, -plotid)
 
-rain_psm_total <- ppt_total@data %>% select(plotid, contains("sum"))
-names(rain_psm_total) <- c("plotid", paste(c(2004:2010), sep = ","))
-rain_psm_total <- rain_psm_total %>% gather(year, rain_tot_psm, -plotid)
+rain_ss_v_total <- rainfall@data %>% 
+      select(plotid, ends_with("v")) %>% 
+      select(plotid, contains("ss")) %>%
+      select(plotid, starts_with("rain"))
+names(rain_ss_v_total) <- c("plotid",paste(c(2004:2014), sep = ","))
+rain_ss_v_total <- rain_ss_v_total %>% gather(year, rain_v_total, -plotid)
 
-rain2d_days <- ppt_days@data %>% select(plotid, ends_with("2d"))
-names(rain2d_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain2d_days <- rain2d_days %>% gather(year, rain_days_2d, -plotid)
+# rain_r_total <- ppt_total@data %>% select(plotid, ends_with("r"))
+# names(rain_r_total) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+# rain_r_total <- rain_r_total %>% gather(year, rain_tot_r, -plotid)
 
-rain3d_days <- ppt_days@data %>% select(plotid, ends_with("3d"))
-names(rain3d_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain3d_days <- rain3d_days %>% gather(year, rain_days_3d, -plotid)
+# rain_psm_total <- ppt_total@data %>% select(plotid, contains("sum"))
+# names(rain_psm_total) <- c("plotid", paste(c(2004:2010), sep = ","))
+# rain_psm_total <- rain_psm_total %>% gather(year, rain_tot_psm, -plotid)
 
-rain_v_days <- ppt_days@data %>% select(plotid, ends_with("v"))
-names(rain_v_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain_v_days <- rain_v_days %>% gather(year, rain_days_v, -plotid)
+# rain2d_days <- ppt_days@data %>% select(plotid, ends_with("2d"))
+# names(rain2d_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+# rain2d_days <- rain2d_days %>% gather(year, rain_days_2d, -plotid)
 
-rain_r_days <- ppt_days@data %>% select(plotid, ends_with("r"))
-names(rain_r_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
-rain_r_days <- rain_r_days %>% gather(year, rain_days_r, -plotid)
+# rain3d_days <- ppt_days@data %>% select(plotid, ends_with("3d"))
+# names(rain3d_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+# rain3d_days <- rain3d_days %>% gather(year, rain_days_3d, -plotid)
 
-rain_total <- left_join(rain2d_total, 
-                        left_join(rain3d_total, rain_v_total, 
-                                  by = c("plotid","year")),
+rain_rs_v_days <- rainfall@data %>% 
+      select(plotid, contains("dys")) %>% 
+      select(plotid, ends_with("v")) %>% 
+      select(plotid, contains("rs"))
+names(rain_rs_v_days) <- c("plotid",paste(c(2004:2014), sep = ","))
+rain_rs_v_days <- rain_rs_v_days %>% gather(year, rain_days_v, -plotid)
+
+rain_ss_v_days <- rainfall@data %>% 
+      select(plotid, contains("dys")) %>% 
+      select(plotid, ends_with("v")) %>% 
+      select(plotid, contains("ss"))
+names(rain_ss_v_days) <- c("plotid",paste(c(2004:2014), sep = ","))
+rain_ss_v_days <- rain_ss_v_days %>% gather(year, rain_days_v, -plotid)
+
+# rain_r_days <- ppt_days@data %>% select(plotid, ends_with("r"))
+# names(rain_r_days) <- c("plotid",paste(c(2004:2012,2014), sep = ","))
+# rain_r_days <- rain_r_days %>% gather(year, rain_days_r, -plotid)
+
+rain_rs_total <- left_join(rain2d_rs_total,
+                           rain_rs_v_total,
                            by = c("plotid","year"))
-rain_total <- left_join(rain_total, rain_r_total, by = c("plotid", "year"))
-rain_total <- left_join(rain_total, rain_psm_total, by = c("plotid", "year"))
-summary(rain_total)# NAs are due to current PRISM data only going through 2010
-unique(rain_total$plotid)# 202 plots
-unique(rain_total$year)# 10 sampling seasons
 
-rain_days <- left_join(rain2d_days,
-                       left_join(rain3d_days, rain_v_days, 
-                                 by = c("plotid", "year")),
-                       by = c("plotid", "year"))
-rain_days <- left_join(rain_days, rain_r_days, by = c("plotid", "year"))
-summary(rain_days)
-unique(rain_days$plotid)# 202 plots
-unique(rain_days$year)# 10 sampling seasons
+rain_ss_total <- left_join(rain2d_ss_total,
+                           rain_ss_v_total,
+                           by = c("plotid","year"))
 
-rm(list = ls(pattern = "^rain_v"))
-rm(list = ls(pattern = "^rain2d"))
-rm(list = ls(pattern = "^rain3d"))
-rm(list = ls(pattern = "^rain_r"))
-rm(rain_psm_total)
+# rain_total <- left_join(rain_total, rain_r_total, by = c("plotid", "year"))
+# rain_total <- left_join(rain_total, rain_psm_total, by = c("plotid", "year"))
+# summary(rain_total)# NAs are due to current PRISM data only going through 2010
+# unique(rain_total$plotid)# 202 plots
+# unique(rain_total$year)# 10 sampling seasons
 
-write.csv(rain_total, "analysis/data/rain_total_long.csv", row.names = F)
-write.csv(rain_days, "analysis/data/rain_days_long.csv", row.names = F)
+# rain_days <- left_join(rain2d_days,
+#                        left_join(rain3d_days, rain_v_days, 
+#                                  by = c("plotid", "year")),
+#                        by = c("plotid", "year"))
+# rain_days <- left_join(rain_days, rain_r_days, by = c("plotid", "year"))
+# summary(rain_days)
+# unique(rain_days$plotid)# 202 plots
+# unique(rain_days$year)# 10 sampling seasons
+
+# rm(list = ls(pattern = "^rain_v"))
+# rm(list = ls(pattern = "^rain2d"))
+# rm(list = ls(pattern = "^rain3d"))
+# rm(list = ls(pattern = "^rain_r"))
+# rm(rain_psm_total)
+
+write.csv(rain_rs_total, "analysis/data/rainy_season_rainfall_total_long.csv", row.names = F)
+write.csv(rain_ss_total, "analysis/data/sampling_season_rainfall_total_long.csv", row.names = F)
+write.csv(rain_rs_v_days, "analysis/data/rainy_season_wet_days_long.csv", row.names = F)
+write.csv(rain_ss_v_days, "analysis/data/sampling_season_wet_days_long.csv", row.names = F)
 ```
@@
 str(plots_env)
-plots_env <- plots_env %>% rename(plotid = PLOT_ID)
+# plots_env <- plots_env %>% rename(plotid = PLOT_ID)
 
-write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
+# write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
 # plots_env <- read.csv("analysis/data/env_gis_variables.csv")
@@
 summary(rs_temps)
+rs_temps$plotid <- tolower(rs_temps$plotid)
 summary(summer_temps)
+summer_temps$plotid <- tolower(summer_temps$plotid)
 
 # Rainfall Data
-summary(ppt_days)
-summary(ppt_total)
-rain_df <- merge(ppt_days@data, ppt_total@data)
-summary(rain_df)
+rainy_season_ppt <- merge(rain_rs_total, rain_rs_v_days)
+summary(rainy_season_ppt)
 
-unique(rs_temps$plotid)# 201 plots
-unique(summer_temps$plotid)# 201 plots
-unique(rain_df$plotid)# 202 plots
+length(unique(rs_temps$plotid))# 201 plots
+length(unique(summer_temps$plotid))# 201 plots
+length(unique(rainy_season_ppt$plotid))# 202 plots
 
-tempsid <- as.character(unique(rs_temps$plotid))
-rainid <- as.character(unique(rain_df$plotid))
+tempsid <- tolower(as.character(unique(rs_temps$plotid)))
+rainid <- as.character(unique(rainy_season_ppt$plotid))
 rainid %in% tempsid
 rainid[62]# BUSH01 still in the rain data
-rain_df <- filter(rain_df, plotid != "BUSH01")
+rainy_season_ppt <- filter(rainy_season_ppt, plotid != "bush01")
 
@@
 
-## Investigating `NAs` in the `plots_temps` data set
+## Investigating `NAs` in the temperature data set
 A number of observations for above 25 threshold have NAs.
@@
 ```{r investigate NAs, eval=FALSE, echo=FALSE}
-# summary(plots_env)
-# filter(plots_env, plotid == "BUSH01")
-
-# BUSH01 plot included. I think we have limited ecological data for this plot, and no microclimate data. So, I am going to exclude it from all the data sets
-# plots_env <- filter(plots_env, plotid != "BUSH01")
-stems <- as.tbl(stems)
-stems
-# stems <- select(stems, -X) 
-unique(filter(stems, plot_status == "Abandoned")$plotid)
-stems <- stems %>% filter(plotid != "bush01", plotid != "ponti01", 
-                                plotid != "sweet01")
-stems$plotid <- droplevels(stems$plotid)
-
 rs_temps %>% filter(is.na(hrs_abv25))
@@
 ```{r save data}
-write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
-write.csv(rain_df, "analysis/data/rain_data_long.csv", row.names = F)
-write.csv(rs_temps, "analysis/data/rs_temps_data.csv", row.names = F)
-write.csv(summer_temps, "analysis/data/summer_temps_data.csv", row.names = F)
+stems <- as.tbl(stems)
+stems
+# stems <- select(stems, -X) 
+# unique(filter(stems, plot_status == "Abandoned")$plotid)
+stems <- stems %>% filter(plotid != "bush01", plotid != "ponti01", 
+                                plotid != "sweet01")
+stems$plotid <- droplevels(stems$plotid)
+# write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
+# write.csv(rainy_season_ppt, "analysis/data/rain_data_long.csv", row.names = F)
+# write.csv(rs_temps, "analysis/data/rs_temps_data.csv", row.names = F)
+# write.csv(summer_temps, "analysis/data/summer_temps_data.csv", row.names = F)
 # write.csv(stems, "analysis/data/stems_noBUSH01.csv", row.names = F)
@@
 ```{r reload data}
-plots_env <- read.csv("analysis/data/env_gis_variables.csv")
-rain_df <- read.csv("analysis/data/rain_data_long.csv")
-rs_temps <- read.csv("analysis/data/rs_temps_data.csv")
-summer_temps <- read.csv("analysis/data/summer_temps_data.csv")
+# plots_env <- read.csv("analysis/data/env_gis_variables.csv")
+# rain_df <- read.csv("analysis/data/rain_data_long.csv")
+# rs_temps <- read.csv("analysis/data/rs_temps_data.csv")
+# summer_temps <- read.csv("analysis/data/summer_temps_data.csv")
 # stems <- read.csv("analysis/data/stems_noBUSH01.csv")
@@
 filter(live_umca, is.na(umca_slc))
-# Spot checking some of these shows that NAs are reasonable
+# Spot checking some of these shows that NAs are due to missing data, i.e. they should not be changed to zeroes
 umca_plots <- left_join(
@@
 umca_plots <- umca_plots %>% filter(avg_lfct != "NaN")
+# filter(umca_plots, plotid == "yahng02")
 summary(umca_plots)
@@
 # Add rainfall data to UMCA plots data frame
-summary(rain_total)
-unique(rain_total$plotid)
-rain_total$plotid <- tolower(rain_total$plotid)
-summary(rain_days)
-unique(rain_days$plotid)
-rain_days$plotid <- tolower(rain_days$plotid)
+summary(rain_rs_total)
+summary(rain_ss_total)
+# unique(rain_rs_total$plotid)
+# unique(rain_ss_total$plotid)
 
-rain_total <- rename(rain_total, sample_year = year)
-rain_total$sample_year <- as.character(rain_total$sample_year)
-rain_days <- rename(rain_days, sample_year = year)
-rain_days$sample_year <- as.character(rain_days$sample_year)
+summary(rain_rs_v_days)
+summary(rain_ss_v_days)
+# unique(rain_rs_v_days$plotid)
+# unique(rain_ss_v_days$plotid)
+
+rain_rs_total <- rename(rain_rs_total, sample_year = year,
+                        rs2d_total = rain2d_total,
+                        rs_v_total = rain_v_total)
+rain_ss_total <- rename(rain_ss_total, sample_year = year,
+                        ss2d_total = rain2d_total,
+                        ss_v_total = rain_v_total)
+rain_total <- left_join(rain_rs_total, rain_ss_total, by = c("plotid", "sample_year"))
+# rain_total$sample_year <- as.character(rain_total$sample_year)
+rain_rs_v_days <- rename(rain_rs_v_days, sample_year = year, rs_days = rain_days_v)
+rain_ss_v_days <- rename(rain_ss_v_days, sample_year = year, ss_days = rain_days_v)
+rain_days <- left_join(rain_rs_v_days, rain_ss_v_days, by = c("plotid", "sample_year"))
+# rain_days$sample_year <- as.character(rain_days$sample_year)
 
@@
 
-filter(umca_plots, is.na(rain_tot_2d))
-# Same case of NAs in rainfall due to data from 2003
+filter(umca_plots, is.na(rs2d_total))
+# Same case of NAs in rainfall due to data sampled in 2003
 ```
@@
 summary(umca_plots)
-filter(umca_plots, is.na(H.2014))$plotid
+unique(filter(umca_plots, is.na(H.2014))$plotid)
 ## NAs due to plots abandoned/lost between 2005 and 2014; 4 unique plots
@@
 
-head(canopy_cover)
-canopy_cover <- select(canopy_cover, -X)
-canopy_cover <- rename(canopy_cover, sample_year = year)
-summary(canopy_cover)
-unique(filter(canopy_cover, sample_year == 2006)$plotid)
-canopy_cover$sample_year <- as.character(canopy_cover$sample_year)
-
-umca_plots <- left_join(umca_plots, canopy_cover, by = c("plotid", "sample_year"))
-summary(umca_plots)# NAs because canopy cover was not measured every year
+# head(canopy_cover)
+# canopy_cover <- select(canopy_cover, -X)
+# canopy_cover <- rename(canopy_cover, sample_year = year)
+# summary(canopy_cover)
+# unique(filter(canopy_cover, sample_year == 2006)$plotid)
+# canopy_cover$sample_year <- as.character(canopy_cover$sample_year)
+# 
+# umca_plots <- left_join(umca_plots, canopy_cover, by = c("plotid", "sample_year"))
+# summary(umca_plots)# NAs because canopy cover was not measured every year
 ```
@@
 
-pairs(select(umca_plots, candens15m, pct_cvr, veght15m), 
+pairs(select(umca_plots, candens15m, veght_m), 
       lower.panel = panel.cor, diag.panel = panel.hist, 
       main = "Canopy Data - UMCA Plots")
-cor(umca_plots$candens15m, umca_plots$pct_cvr, use = "complete.obs")
-cor(umca_plots$veght15m, umca_plots$pct_cvr, use = "complete.obs")
+# cor(umca_plots$candens15m, umca_plots$pct_cvr, use = "complete.obs")
+# cor(umca_plots$veght15m, umca_plots$pct_cvr, use = "complete.obs")
 
-pairs(na.omit(select(umca_plots, avg_lfct, tot_lfct, contains("hrs"), contains("rain_tot"), H.2005, H.2014)),
+pairs(na.omit(select(umca_plots, avg_lfct, tot_lfct, contains("hrs"), contains("_total"), H.2005, H.2014)),
       lower.panel = panel.cor, diag.panel = panel.hist, 
       main = "Leaf Count, Temperature, Rainfall, Diversity - UMCA Plots")
-## Rainfall and temperature all have fairly weak correlations (<0.36)
-## The diversity metrics are the only correlates with leaf count >0.20
+plot(umca_plots$hrs_abv25ds, log1p(umca_plots$tot_lfct))
+cor(umca_plots$hrs_abv25ds, log1p(umca_plots$tot_lfct), use = "complete")
 ```
@@
 oak_sod <- left_join(oak_sod, 
-                     select(umca_plots, plotid, sample_year, uninfected_bay_ct, infected_bay_ct, tot_bay, avg_lfct, tot_lfct), 
+                     select(umca_plots, plotid, sample_year, uninfected_bay_ct, infected_bay_ct, tot_bay, tot_lfct), 
                      by = c("plotid", "sample_year"))
@@
 summary(oak_sod)
-filter(oak_sod, is.na(avg_tmax_rs))# NAs are plots sampled in 2003
-filter(oak_sod, is.na(H.2014))# NAs are plots lost before 2014
+summary(filter(oak_sod, is.na(avg_tmax_rs)))# NAs are plots sampled in 2003
+unique(filter(oak_sod, is.na(H.2014))$plotid)# NAs are plots lost before 2014
 summary(filter(oak_sod, is.na(us.J.2011)))# zero or missing data for understory richness
@@
 
-filter(oak_sod, is.na(H.2014))$plotid
-## plots that were decommissioned prior to 2014
+unique(filter(oak_sod, is.na(H.2014))$plotid)
+## plots that were decommissioned prior to 2014 sampling
 filter(oak_sod, is.na(us.H.2011))$plotid
-## plots that have legit 0 for richness or were decommissioned prior to 2011
+## plots that have legit 0 for understory richness or were decommissioned prior to 2011
 
-canopy_cover$sample_year <- as.integer(canopy_cover$sample_year)
-oak_sod <- left_join(oak_sod, canopy_cover, by = c("plotid", "sample_year"))
+# canopy_cover$sample_year <- as.integer(canopy_cover$sample_year)
+# oak_sod <- left_join(oak_sod, canopy_cover, by = c("plotid", "sample_year"))
 
@@
 
-summary(filter(oak_sod, is.na(rain_tot_psm))) # all obs after 2010
-summary(filter(oak_sod, is.na(us.J.2011)))
-filter(oak_sod, is.na(us.J.2011))$plotid# ann06, mroth03, sumtv02
+# summary(filter(oak_sod, is.na(rain_tot_psm))) # all obs after 2010
+# summary(filter(oak_sod, is.na(us.J.2011)))
+unique(filter(oak_sod, is.na(us.J.2011))$plotid)# ann06, mroth03, sumtv02
 # ann06 & mroth03 are legitimate zeroes for understory richness in 2011
+# i must have excluded mcneil03 from the diversity calculations
 # sumtv02 was abandoned after 2008, so it should be an NA for later years
@@
 summary(oak_sod)
+unique(filter(oak_sod, is.na(tot_lfct))$plotid)
+filter(oak_sod, plotid == "belta07")[c(1,2,5,8,9)]# no bay in plot 2004-14
+oak_sod$tot_lfct[oak_sod$plotid == "belta07" & is.na(oak_sod$tot_lfct)] <- 0
+filter(oak_sod, plotid == "ann23")[c(1,2,5,8,9)]# no bay in plot 2005-06
+oak_sod$tot_lfct[oak_sod$plotid == "ann23" & is.na(oak_sod$tot_lfct)] <- 0
+filter(oak_sod, plotid == "arbit05")[c(1,2,5,8,9)]# no bay 2004-2010
+oak_sod$tot_lfct[oak_sod$plotid == "arbit05" & is.na(oak_sod$tot_lfct)] <- 0
+filter(oak_sod, plotid == "backm02")[c(1,2,5,8,9)]# no bay in plot 2004-14
+oak_sod$tot_lfct[oak_sod$plotid == "backm02" & is.na(oak_sod$tot_lfct)] <- 0
+filter(oak_sod, plotid == "belta07")[c(1,2,5,8,9)]# no bay in plot 2004-14
+oak_sod$tot_lfct[oak_sod$plotid == "belta07" & is.na(oak_sod$tot_lfct)] <- 0
+filter(oak_sod, plotid == "hallo01")[c(1,2,5,8,9)]# no bay 2004-2012
+filter(oak_sod, plotid == "jlsp06")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "kunde03")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "mccor02")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "mroth03")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "peter01")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "petsn02")[c(1,2,5,8,9)]# no bay 2005-14
+filter(oak_sod, plotid == "pfend01")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "phill01")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "phill02")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "schiek02")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "sdc04")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "sugar24")[c(1,2,5,8,9)]# no bay 2005-14
+filter(oak_sod, plotid == "sumtv01")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "sumtv02")[c(1,2,5,8,9)]# no bay after 2006
+filter(oak_sod, plotid == "svrp01")[c(1,2,5,8,9)]# no bay 2005-14
+filter(oak_sod, plotid == "tell03")[c(1,2,5,8,9)]# no bay 2004-14
+filter(oak_sod, plotid == "terri03")[c(1,2,5,8,9)]# no bay 2005-2010
+filter(oak_sod, plotid == "yahng02")[c(1,2,5,8,9)]# missed bay in 2010 visit
+
+oak_sod$tot_lfct[is.na(oak_sod$tot_lfct)] <- 0
+oak_sod$tot_lfct[oak_sod$plotid == "yahng02" & oak_sod$sample_year == 2010] <- NA
+oak_sod$tot_bay[oak_sod$plotid == "yahng02" & oak_sod$sample_year == 2010] <- NA
+
 # drop rainy days interpolations, they aren't reliable; keep voronoi neighbor method
-oak_sod <- select(oak_sod, -rain_days_2d, -rain_days_3d, -rain_days_r)
-pairs(select(oak_sod, elev_m, starts_with("hrs"), starts_with("avg_t"), contains("rain_tot")), lower.panel = panel.cor, diag.panel = panel.hist, main = "Oak-SOD Rainfall & Temperature Data")
+# oak_sod <- select(oak_sod, -rain_days_2d, -rain_days_3d, -rain_days_r)
+pairs(select(oak_sod, elev_m, starts_with("hrs"), starts_with("avg_t"), contains("_total")), lower.panel = panel.cor, diag.panel = panel.hist, main = "Oak-SOD Rainfall & Temperature Data")
 
-pairs(na.omit(select(oak_sod, c(3:4,7:8,9:10), H.2005, H.2014)),
+# filter(oak_sod, is.na(elev_m))
+# oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01")
+# filter(oak_sod, is.na(hrs_14_20_rs))
+# oak_sod <- filter(oak_sod, plotid != "bush01")
+unique(filter(oak_sod, is.na(H.2005))$plotid)
+## removing b/c not in the most recent data set (2014)
+# oak_sod <- filter(oak_sod, plotid != "mcneil03", plotid != "sumtv02",
+                  # plotid != "votru01")
+# filter(oak_sod, is.na(hrs_abv25rs))# I believe these locations actually did not have any temperatures above 25 C during the rainy season
+# oak_sod$hrs_abv25rs[is.na(oak_sod$hrs_abv25rs)] <- 0
+
+length(unique(oak_sod$plotid))# 164 plots remaining if I remove mcneil03, sumtv02, votru01
+
+pairs(na.omit(select(oak_sod, uninf_oak_ct:tot_oak_ct, infected_bay_ct:tot_lfct, H.2005, H.2014)),
       lower.panel = panel.cor, diag.panel = panel.hist, 
       main = "Oak-SOD Infection & Diversity Data")
+```
+
+### Calculate "Lagged" Rainfall & Temperature Variables
+I'm going to add some lag variables, e.g. average rainfall for time t + t-1 or t-1 + t-2.
+```{r create rainfall lag variables}
+library(tidyr)
+summary(rain_total)
+unique(rain_total$sample_year)
+rs2d_lag <- as.tbl(select(rain_total, plotid, sample_year, rs2d_total))
+rs2d_lag1 <- as.tbl(select(rain_total, plotid, sample_year))
+rs2d_lag1 <- rs2d_lag1 %>% 
+      mutate(sample_year_minus1 = sample_year - 1, 
+             sample_year_minus2 = sample_year - 2,
+             sample_year_minus3 = sample_year - 3) 
+rs2d_lag1 <- left_join(rs2d_lag1, rs2d_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(rs2d_total_t_minus1 = rs2d_total)
+rs2d_lag1 <- left_join(rs2d_lag1, rs2d_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(rs2d_total_t_minus2 = rs2d_total)
+rs2d_lag1 <- left_join(rs2d_lag1, rs2d_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(rs2d_total_t_minus3 = rs2d_total)
+rs2d_lag1 <- left_join(rs2d_lag1, rs2d_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+rs2d_lag1 <- select(rs2d_lag1, plotid, sample_year, starts_with("rs2d"))
+
+rs2d_lag1 <- rs2d_lag1 %>% 
+      mutate(avg_rs2d_t_minus1 = 
+                   rowMeans(cbind(rs2d_total, rs2d_total_t_minus1)),
+             avg_rs2d_t_minus2 =
+                   rowMeans(cbind(rs2d_total, rs2d_total_t_minus1, rs2d_total_t_minus2)),
+             avg_rs2d_t1t2 = rowMeans(cbind(rs2d_total_t_minus1, rs2d_total_t_minus2)))
+summary(rs2d_lag1)
+rm(rs2d_lag)
+```
+
+```{r create wet-hours temperature lag variables}
+summary(wet_hrs); unique(wet_hrs$sample_year)
+
+hrsblw14_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrsblw14_wet))
+avgtmax_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, avgtmax_wet))
+hrs1422_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrs1422_wet))
+wet_hrs_lag <- select(wet_hrs_lag, plotid, sample_year)
+wet_hrs_lag <- wet_hrs_lag %>% 
+      mutate(sample_year_minus1 = sample_year - 1,
+             sample_year_minus2 = sample_year - 2,
+             sample_year_minus3 = sample_year - 3)
+
+## Number of wet-hours below 14 C lagged ####
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw14_wet_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrsblw14_wet_tminus1 = hrsblw14_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw14_wet_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrsblw14_wet_tminus2 = hrsblw14_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw14_wet_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrsblw14_wet_tminus3 = hrsblw14_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw14_wet_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## Average daily maximum temperature on wet days lagged ####
+wet_hrs_lag <- left_join(wet_hrs_lag, avgtmax_wet_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(avgtmax_wet_tminus1 = avgtmax_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, avgtmax_wet_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(avgtmax_wet_tminus2 = avgtmax_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, avgtmax_wet_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(avgtmax_wet_tminus3 = avgtmax_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, avgtmax_wet_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## Number of hours 14-22 C on wet days ####
+wet_hrs_lag <- left_join(wet_hrs_lag, hrs1422_wet_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrs1422_wet_tminus1 = hrs1422_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrs1422_wet_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrs1422_wet_tminus2 = hrs1422_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrs1422_wet_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrs1422_wet_tminus3 = hrs1422_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrs1422_wet_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## ####
+wet_hrs_lag <- select(wet_hrs_lag, plotid, sample_year, starts_with("hrs"), starts_with("avg"))
+wet_hrs_lag <- wet_hrs_lag %>% 
+      mutate(
+            avg_hrsblw14_wet_tminus1 = rowMeans(cbind(hrsblw14_wet, hrsblw14_wet_tminus1)),
+             avg_hrsblw14_wet_tminus2 = rowMeans(cbind(hrsblw14_wet, hrsblw14_wet_tminus1, hrsblw14_wet_tminus2)),
+             avg_hrsblw14_wet_tminus3 = rowMeans(cbind(hrsblw14_wet, hrsblw14_wet_tminus1, hrsblw14_wet_tminus2, hrsblw14_wet_tminus3)),
+
+            avgtmax_wet_t_t1 = rowMeans(cbind(avgtmax_wet, avgtmax_wet_tminus1)),
+             avgtmax_wet_t_t1t2 = rowMeans(cbind(avgtmax_wet, avgtmax_wet_tminus1, avgtmax_wet_tminus2)),
+             avgtmax_wet_t1t2 = rowMeans(cbind(avgtmax_wet_tminus1, avgtmax_wet_tminus2)),
+
+            avg_hrs1422_wet_t_t1 = rowMeans(cbind(hrs1422_wet, hrs1422_wet_tminus1)),
+             avg_hrs1422_wet_t_t1t2 = rowMeans(cbind(hrs1422_wet, hrs1422_wet_tminus1, hrs1422_wet_tminus2)),
+             avg_hrs1422_wet_t1t2 = rowMeans(cbind(hrs1422_wet_tminus1, hrs1422_wet_tminus2))
+            )
+```
+
+```{r create rainy season temperature lagged variables}
+summary(rs_temps); unique(rs_temps$sample_year)
+
+hrs1420_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, hrs_14_20_rs))
+avgtmax_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, avg_tmax_rs))
+hrsabv20_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, hrs_abv20_rs))
+hrsblw10_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, hrs_blw10rs))
+
+rs_temps_lag <- select(rs_temps, plotid, sample_year)
+rs_temps_lag <- rs_temps_lag %>% 
+      mutate(sample_year_minus1 = sample_year - 1,
+             sample_year_minus2 = sample_year - 2,
+             sample_year_minus3 = sample_year - 3)
+
+## Number of rainy season hours 14-20 C ####
+rs_temps_lag <- left_join(rs_temps_lag, hrs1420_rs_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrs1420_rs_tminus1 = hrs_14_20_rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrs1420_rs_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrs1420_rs_tminus2 = hrs_14_20_rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrs1420_rs_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrs1420_rs_tminus3 = hrs_14_20_rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrs1420_rs_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## Average daily maximum temperature during rainy season lagged ####
+rs_temps_lag <- left_join(rs_temps_lag, avgtmax_rs_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(avgtmax_rs_tminus1 = avg_tmax_rs)
+rs_temps_lag <- left_join(rs_temps_lag, avgtmax_rs_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(avgtmax_rs_tminus2 = avg_tmax_rs)
+rs_temps_lag <- left_join(rs_temps_lag, avgtmax_rs_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(avgtmax_rs_tminus3 = avg_tmax_rs)
+rs_temps_lag <- left_join(rs_temps_lag, avgtmax_rs_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## Number of hours above 20 C during rainy season lagged ####
+rs_temps_lag <- left_join(rs_temps_lag, hrsabv20_rs_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrsabv20_rs_tminus1 = hrs_abv20_rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsabv20_rs_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrsabv20_rs_tminus2 = hrs_abv20_rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsabv20_rs_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrsabv20_rs_tminus3 = hrs_abv20_rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsabv20_rs_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## Number of hours below 10 C during rainy season lagged ####
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw10_rs_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrsblw10_rs_tminus1 = hrs_blw10rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw10_rs_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrsblw10_rs_tminus2 = hrs_blw10rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw10_rs_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrsblw10_rs_tminus3 = hrs_blw10rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw10_rs_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## ####
+rs_temps_lag <- select(rs_temps_lag, plotid, sample_year, starts_with("hrs"), starts_with("avg"))
+rs_temps_lag <- rs_temps_lag %>% 
+      mutate(
+            avgtmax_rs_t_t1 = rowMeans(cbind(avg_tmax_rs, avgtmax_rs_tminus1)),
+             avgtmax_rs_t_t1t2 = rowMeans(cbind(avg_tmax_rs, avgtmax_rs_tminus1, avgtmax_rs_tminus2)),
+             avgtmax_rs_t_t1t2t3 = rowMeans(cbind(avg_tmax_rs, avgtmax_rs_tminus1, avgtmax_rs_tminus2, avgtmax_rs_tminus3)),
+             avgtmax_rs_t1t2 = rowMeans(cbind(avgtmax_rs_tminus1, avgtmax_rs_tminus2)),
+
+            avg_hrs1420_rs_tminus1 = rowMeans(cbind(hrs_14_20_rs, hrs1420_rs_tminus1)),
+             avg_hrs1420_rs_tminus2 = rowMeans(cbind(hrs_14_20_rs, hrs1420_rs_tminus1, hrs1420_rs_tminus2)),
+             avg_hrs1420_rs_tminus3 = rowMeans(cbind(hrs_14_20_rs, hrs1420_rs_tminus1, hrs1420_rs_tminus2, hrs1420_rs_tminus3)),
+             avg_hrs1420_rs_t1t2 = rowMeans(cbind(hrs1420_rs_tminus1, hrs1420_rs_tminus2)),
+
+            avg_hrsblw10_rs_tminus1 = rowMeans(cbind(hrs_blw10rs, hrsblw10_rs_tminus1)),
+             avg_hrsblw10_rs_tminus2 = rowMeans(cbind(hrs_blw10rs, hrsblw10_rs_tminus1, hrsblw10_rs_tminus2)),
+             avg_hrsblw10_rs_tminus3 = rowMeans(cbind(hrs_blw10rs, hrsblw10_rs_tminus1, hrsblw10_rs_tminus2, hrsblw10_rs_tminus3)),
+             avg_hrsblw10_rs_t1t2 = rowMeans(cbind(hrsblw10_rs_tminus1, hrsblw10_rs_tminus2)),
+
+             avg_hrsabv20_rs_tminus1 = rowMeans(cbind(hrs_abv20_rs, hrsabv20_rs_tminus1)),
+             avg_hrsabv20_rs_tminus2 = rowMeans(cbind(hrs_abv20_rs, hrsabv20_rs_tminus1, hrsabv20_rs_tminus2)),
+             avg_hrsabv20_rs_tminus3 = rowMeans(cbind(hrs_abv20_rs, hrsabv20_rs_tminus1, hrsabv20_rs_tminus2, hrsabv20_rs_tminus3)),
+             avg_hrsabv20_rs_t1t2 = rowMeans(cbind(hrsabv20_rs_tminus1, hrsabv20_rs_tminus2))
+             )
+```
+
+```{r create summer season temperature lagged variables}
+summary(summer_temps); unique(summer_temps$sample_year)
+
+hrs_abv25ds_lag <- as.tbl(select(summer_temps, plotid, sample_year, hrs_abv25ds))
+avg_tmax_ds_lag <- as.tbl(select(summer_temps, plotid, sample_year, avg_tmax_ds))
+hrs_blw18ds_lag <- as.tbl(select(summer_temps, plotid, sample_year, hrs_blw18ds))
+
+summer_temps_lag <- select(rs_temps, plotid, sample_year)
+summer_temps_lag <- summer_temps_lag %>% 
+      mutate(sample_year_minus1 = sample_year - 1,
+             sample_year_minus2 = sample_year - 2,
+             sample_year_minus3 = sample_year - 3)
+
+## Number of dry-season hours above 25 C lagged ####
+summer_temps_lag <- left_join(summer_temps_lag, hrs_abv25ds_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrs_abv25ds_tminus1 = hrs_abv25ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_abv25ds_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrs_abv25ds_tminus2 = hrs_abv25ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_abv25ds_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrs_abv25ds_tminus3 = hrs_abv25ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_abv25ds_lag,
+                              by = c("plotid", "sample_year"="sample_year"))
+
+## Average daily maximum temperature during dry-season lagged ####
+summer_temps_lag <- left_join(summer_temps_lag, avg_tmax_ds_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(avg_tmax_ds_tminus1 = avg_tmax_ds)
+summer_temps_lag <- left_join(summer_temps_lag, avg_tmax_ds_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(avg_tmax_ds_tminus2 = avg_tmax_ds)
+summer_temps_lag <- left_join(summer_temps_lag, avg_tmax_ds_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(avg_tmax_ds_tminus3 = avg_tmax_ds)
+summer_temps_lag <- left_join(summer_temps_lag, avg_tmax_ds_lag,
+                              by = c("plotid","sample_year"="sample_year"))
+
+## Number of dry-season hours below 18 C lagged ####
+summer_temps_lag <- left_join(summer_temps_lag, hrs_blw18ds_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrs_blw18ds_tminus1 = hrs_blw18ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_blw18ds_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrs_blw18ds_tminus2 = hrs_blw18ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_blw18ds_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrs_blw18ds_tminus3 = hrs_blw18ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_blw18ds_lag,
+                              by = c("plotid", "sample_year"="sample_year"))
+
+## ####
+summer_temps_lag <- select(summer_temps_lag, plotid, sample_year, starts_with("hrs"), starts_with("avg"))
+summer_temps_lag <- summer_temps_lag %>% 
+      mutate(
+            avg_hrsblw18_ds_tminus1 = rowMeans(cbind(hrs_blw18ds, hrs_blw18ds_tminus1)),
+             avg_hrsblw18_ds_tminus2 = rowMeans(cbind(hrs_blw18ds, hrs_blw18ds_tminus1, hrs_blw18ds_tminus2)),
+             avg_hrsblw18_ds_tminus3 = rowMeans(cbind(hrs_blw18ds, hrs_blw18ds_tminus1, hrs_blw18ds_tminus2, hrs_blw18ds_tminus3)),
+             avg_hrsblw18_ds_t1t2 = rowMeans(cbind(hrs_blw18ds_tminus1, hrs_blw18ds_tminus2)),
+
+            avgtmax_ds_t_t1 = rowMeans(cbind(avg_tmax_ds, avg_tmax_ds_tminus1)),
+             avgtmax_ds_t_t1t2 = rowMeans(cbind(avg_tmax_ds, avg_tmax_ds_tminus1, avg_tmax_ds_tminus2)),
+             avgtmax_ds_t_t1t2t3 = rowMeans(cbind(avg_tmax_ds, avg_tmax_ds_tminus1, avg_tmax_ds_tminus2, avg_tmax_ds_tminus3)),
+             avgtmax_ds_t1t2 = rowMeans(cbind(avg_tmax_ds_tminus1, avg_tmax_ds_tminus2)),
+
+            avg_hrsabv25_ds_tminus1 = rowMeans(cbind(hrs_abv25ds, hrs_abv25ds_tminus1)),
+             avg_hrsabv25_ds_tminus2 = rowMeans(cbind(hrs_abv25ds, hrs_abv25ds_tminus1, hrs_abv25ds_tminus2)),
+             avg_hrsabv25_ds_tminus3 = rowMeans(cbind(hrs_abv25ds, hrs_abv25ds_tminus1, hrs_abv25ds_tminus2, hrs_abv25ds_tminus3)),
+             avg_hrsabv25_ds_t1t2 = rowMeans(cbind(hrs_abv25ds_tminus1, hrs_abv25ds_tminus2))
+            )
 ```
@@
 
-### Update path analysis data
+### Update path analysis .Rdata object
 Updating the path analysis data with the temperature during wet days variables.
 ```{r update data}
-# load("pathmodel_data_20160112.RData")
-wet_hrs <- read.csv("analysis/data/wet-days-temperature-variables.csv")
+# load("~/GitHub/superspreaders/pathmodel_data.RData")
+
+# Add untagged dbh data
+untagged_stems <- read.csv("analysis/data/untagged_stems_corrected.csv")
+untagged_dbh_2005 <- read.csv("analysis/data/untagged_dbh2005_summary.csv")
+untagged_dbh_2014 <- read.csv("analysis/data/untagged_dbh2014_summary.csv")
+
+# remove unnecessary data
+rm(list = ls(pattern = "^rain_rs"))
+rm(list = ls(pattern = "^rain_ss"))
+rm(list = ls(pattern = "^rain2d"))
+rm(list = ls(pattern = "*ds_lag"))
+rm(list = ls(pattern = "*0_rs_lag")); rm(avgtmax_rs_lag)
+rm(list = ls(pattern = "*wet_lag"))
+
 save.image("~/GitHub/superspreaders/pathmodel_data.RData")

### 2016-03-14 climate-path-model.R
Add ideas on basal area vs. density.
I also updated varables with lagged rainfall, and added variable scaling, e.g. hours to days and mm to cm. This commit also includes single season models of disease prevalence (plot-level infection observed each year) through 2011.

@@ -40,19 +40,48 @@ summary(oak_sod)
 # oak_sod$plotid <- as.factor(oak_sod$plotid)
 str(oak_sod)
 
-## join wet-hours data
+#+ join wet-hours and lagged rainfall data ####
 summary(wet_hrs)# since these are counts the NAs are in fact zeroes
 wet_hrs[is.na(wet_hrs)] <- 0# introduce a lot of zeroes for some columns
 oak_sod <- left_join(
       oak_sod, select(wet_hrs, -X), by = c("plotid","sample_year"))
 oak_sod$plotid <- as.factor(oak_sod$plotid)
 
+summary(rs2d_lag1)
+oak_sod <- left_join(
+      oak_sod, select(rs2d_lag1, -rs2d_total), by = c("plotid","sample_year"))
+oak_sod$plotid <- as.factor(oak_sod$plotid)
+
 ## drop 3d, PRISM, and regression rainfall variables based on discussion with Ryan; use a simple 2d spline interpolation or the Voronoi polygon values
-oak_sod <- select(oak_sod, -rain_tot_3d, -rain_tot_r, -rain_tot_psm)
-oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01", plotid != "bush01")
+# oak_sod <- select(oak_sod, -rain_tot_3d, -rain_tot_r, -rain_tot_psm)
+# oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01", plotid != "bush01")
+
[…]

### 2016-03-14 corrected-dbh.R
Update to match column names.
I exported a fresh TaggedStemQuery data set from the database and needed to update the column naming accordingly.

@@ -31,6 +31,14 @@ detach("package:lubridate", unload=TRUE)
 ## Replace all `-9999` codes with NA ####
 tagged_stems[tagged_stems == -9999] <- NA
 summary(tagged_stems)
+tagged_stems[tagged_stems == -999] <- NA
+# filter(tagged_stems, is.na(year))
+# filter(tagged_stems, TagNumber == 2147)
+# filter(tagged_stems, TagNumber == 5843)
+# filter(tagged_stems, TagNumber == 5808)
+# filter(tagged_stems, TagNumber == 1374)
+# filter(tagged_stems, TagNumber == 4263)
+
 
 
 ## Rename columns using `rename` funcion from `dplyr` ####
@@ -38,8 +46,7 @@ library(plyr); library(dplyr)
 tagged_stems <- select(tagged_stems, -Date)
 names(tagged_stems) <- tolower(names(tagged_stems))
 
-tagged_stems <- rename(tagged_stems, plot_status = status,
-                       cluster = clusterid, species = speciesid,
+tagged_stems <- rename(tagged_stems, cluster = clusterid, species = speciesid,
                        tag = tagnumber, initial_dbh = dbhcm,  
                        stem_status = stemstatus, sod_killed = sod_dead, 
                        umca_slc = sympleafcount, canker = cankerpresent,
@@ -51,8 +58,10 @@ unique(tagged_stems$stem_status)
 tagged_stems$plotid <- tolower(tagged_stems$plotid)
 tagged_stems$species <- tolower(tagged_stems$species)
 tagged_stems$sod_killed <- ifelse(tagged_stems$sod_killed == 0, "No", "Yes")
+# tagged_stems$stem_status[tagged_stems$stem_status == "Nearly dead"] <- "Alive"
 tagged_stems <- droplevels(filter(tagged_stems, stem_status == "Alive" | stem_status == "Dead"))
 summary(tagged_stems)
+filter(tagged_stems, is.na(initial_dbh))# 4 stems did not have DBH measured prior to 2012
 
 
 untagged_stems <- select(untagged_stems, -Date, -TreeID, -StemID)
@@ -98,13 +107,13 @@ length(unique(filter(tagged_stems, species == "quke x quag")$tag))# 2 stems
 
[…]

### 2016-03-14 analysis/species_diversity_calc.R
Update based on revised tagged stems data.

@@ -10,7 +10,8 @@
 #+ load plot sampling data ####
 
 ## Ecological data: leaf counts, infection, at stem and plot levels
-load("~/GitHub/superspreaders/stems_plots.RData")
+# load("~/GitHub/superspreaders/stems_plots.RData")
+tagged_stems <- as.tbl(read.csv("analysis/data/tagged-stems-corrected.csv"))
 
 ## Vegetation transect data
 veg_us_2005 <- read.csv("analysis/data/understory_veg_2005.csv")
@@ -18,8 +19,13 @@ veg_us_2011 <- read.csv("analysis/data/understory_veg_2011.csv")
 canopy_cover <- read.csv("analysis/data/plot_canopy_cover_dens.csv")
 
 ## Untagged dbh data
-untagged_dbh_2005 <- read.csv("analysis/data/untagged_dbh2005_summary.csv")
-untagged_dbh_2014 <- read.csv("analysis/data/untagged_dbh2014_summary.csv")
+untagged_stems <- as.tbl(read.csv("analysis/data/untagged-stems_0514-corrected.csv"))
+untagged_dbh_2005 <- filter(untagged_stems, year == 2005)
+untagged_dbh_2014 <- filter(untagged_stems, year == 2014)
+                            
+# untagged_dbh_2005 <- read.csv("analysis/data/untagged_dbh2005_summary.csv")
+# untagged_dbh_2014 <- read.csv("analysis/data/untagged_dbh2014_summary.csv")
+
 alt05live <- read.csv("analysis/data/altsp05abund_live.csv")
 alt05dead <- read.csv("analysis/data/altsp05abund_dead.csv")
 alt14live <- read.csv("analysis/data/altsp14abund_live.csv")
@@ -109,7 +115,9 @@ us_richness <- filter(us_richness, plotid != "bush01", plotid != "halow01",
                       plotid != "sweet01")
 us_richness$us_rich05[us_richness$plotid=="jlsp05"] <- 0
 us_richness$us_rich05[us_richness$plotid=="ann06"] <- 0
+us_richness$us_rich11[us_richness$plotid=="ann06"] <- 0
 us_richness$us_rich05[us_richness$plotid=="mroth03"] <- 0
+us_richness$us_rich11[us_richness$plotid=="mroth03"] <- 0
 
 
 us_sub_richness <- full_join(
@@ -127,9 +135,9 @@ us_sub_richness <- filter(us_sub_richness,
 # mcneil01 was lost in 2009 (road built through plot), so NA for later years
[…]

### 2016-03-14 lag-effects.Rmd
Explore correlation of symptomatic leaf count between years.

---
title: "Correlation of Data Between Years - Lag Effects?"
author: "Whalen Dillon"
date: "February 18, 2016"
output: html_document
---

Examine correlation of the same measurements between years to explore where a lag-effect might be at work.

```{r}
load("pathmodel_data.RData")
library(plyr); library(dplyr); library(tidyr); library(ggplot2)
panel.cor <- function(x, y, digits = 2, cex.cor, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(0, 1, 0, 1))
      # correlation coefficient
      r <- cor(x, y)
      txt <- format(c(r, 0.123456789), digits = digits)[1]
      txt <- paste("r= ", txt, sep = "")
      text(0.5, 0.6, txt)

      # p-value calculation
      p <- cor.test(x, y)$p.value
      txt2 <- format(c(p, 0.123456789), digits = digits)[1]
      txt2 <- paste("p= ", txt2, sep = "")
      if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
      text(0.5, 0.4, txt2)
}
panel.hist <- function(x, ...)
{
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(usr[1:2], 0, 1.5) )
      h <- hist(x, plot = FALSE)
      breaks <- h$breaks; nB <- length(breaks)
      y <- h$counts; y <- y/max(y)
      rect(breaks[-nB], 0, breaks[-1], y, col = "cyan", ...)
}
```

```{r correlation between years for umca plot variables}
summary(umca_plots)
umca_plots$tot_lfct_log <- log1p(umca_plots$tot_lfct)
boxplot(tot_lfct_log ~ sample_year, data = umca_plots)

slc_plot_year <- umca_plots %>% 
      select(plotid, sample_year, tot_lfct_log) %>%
      filter(sample_year != 2003) %>% 
      spread(., sample_year, tot_lfct_log)
summary(slc_plot_year)
names(slc_plot_year) <- c("plotid", "slc04","slc05","slc06","slc07","slc08","slc09","slc10","slc11","slc12","slc14")

pairs(na.omit(select(slc_plot_year, -plotid)),
      lower.panel = panel.cor, diag.panel = panel.hist, 
      main = "Log Total Symptomatic Leaf Count Across Years")

plot(lm(slc05 ~ slc04, data = slc_plot_year))

```

### 2016-03-22 sem-data-development.Rmd
Add infection rule.
I'm not yet confident in this calculation.

@@
 plot_qusp # right now this is broken down by oak species
+
+summary(stems)
+# calculate the number of times a canker was observed on each stem
+cankers_observed <- stems %>%
+      select(plotid, year, species, tag, canker, stem_status) %>% 
+      filter(species == "quag" | species == "quke" | species == "quch" |
+                   species == "quwi", stem_status == "Alive") %>% 
+      group_by(plotid, species, tag) %>% 
+      summarise(times_canker_observed = sum(canker, na.rm = T))
+summary(cankers_observed)
+cankers_observed <- droplevels(cankers_observed)
+
+## If canker was observed at least twice for a stem then "infection confirmed"
+## This is Sarah's "twice cankered always infected rule" - more conservative
+cankers_observed$infection_confirmed_2plus <- ifelse(
+      cankers_observed$times_canker_observed >= 2, 1, 0
+)
+## If canker was observed just once then "infection confirmed"
+## This is a "once cankered always infected rule" - less conservative
+cankers_observed$infection_confirmed_1plus <- ifelse(
+      cankers_observed$times_canker_observed >= 1, 1, 0
+)
+## Both of these approaches are an effort to better represent oak infection, instead of just based on whether or not a canker was observed for a given year.
+
+stems <- left_join(stems, cankers_observed, by = c("plotid", "species", "tag"))
+
+stems <- stems %>% mutate(
+      inf_2plus = ifelse(
+            year == 2003 & canker == 1 & infection_confirmed_2plus == 1, 1, 0)
+) %>% mutate(inf_2plus = ifelse(
+      year == 2004 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
+      mutate(inf_2plus = ifelse(
+            year == 2005 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
+      mutate(inf_2plus = ifelse(
+            year == 2006 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
+      mutate(inf_2plus = ifelse(
+            year == 2007 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
+      mutate(inf_2plus = ifelse(
+            year == 2008 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
+      mutate(inf_2plus = ifelse(
+            year == 2009 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>%
+      mutate(inf_2plus = ifelse(
+            year == 2010 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
+      mutate(inf_2plus = ifelse(
+            year == 2011 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
+      mutate(inf_2plus = ifelse(
+            year == 2012 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0)) %>% 
+      mutate(inf_2plus = ifelse(
+            year == 2014 & canker == 1 & infection_confirmed_2plus == 1 & inf_2plus != 1, 1, 0))
+
+qusp_stems <- stems %>% 
+      select(plotid, year, tag, species, canker, stem_status, infection_confirmed_1plus, infection_confirmed_2plus) %>% 
+      filter(species == "quag" | species == "quke" | species == "quch" |
+                   species == "quwi", stem_status == "Alive")
+
 qusp4_plots <- stems %>%
-      select(plotid, date, species, canker, year, stem_status) %>%
+      select(plotid, species, canker, year, stem_status, infection_confirmed_1plus, infection_confirmed_2plus) %>%
       filter(species == "quag" | species == "quke" | species == "quch" |
@@
       group_by(plotid, year, species) %>% 
-      summarise(uninfected_oak_ct = length(which(canker==0)), 
+      summarise(uninfected_oak_ct = length(which(canker == 0)), 
                 infected_oak_ct = length(which(canker > 0)), 
-                tot_oak = length(species))
+                tot_oak = length(species),
+                inf_oak_1plus_ct = length(which(infection_confirmed_1plus > 0)),
+                inf_oak_2plus_ct = length(which(infection_confirmed_2plus > 0)))
+summary(qusp4_plots)
+# write.csv(qusp4_plots, "analysis/data/qusp4_infections.csv", row.names = F)
+qusp4_plots <- read.csv("analysis/data/qusp4_infections.csv")
 
@@
                 inf_oak_ct = sum(infected_oak_ct), 
-                tot_oak_ct = sum(c(uninfected_oak_ct, infected_oak_ct)))
+                tot_oak_ct = sum(c(uninfected_oak_ct, infected_oak_ct)),
+                inf_oak_1plus_ct = sum(inf_oak_1plus_ct),
+                inf_oak_2plus_ct = sum(inf_oak_2plus_ct))
 oak_sod
 oak_sod <- rename(oak_sod, sample_year = year)
+summary(oak_sod)
 # oak_sod$plotid <- tolower(oak_sod$plotid)
@@
 hrs1422_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrs1422_wet))
-wet_hrs_lag <- select(wet_hrs_lag, plotid, sample_year)
+wet_hrs_lag <- select(wet_hrs, plotid, sample_year)
 wet_hrs_lag <- wet_hrs_lag %>% 
@@
 rm(list = ls(pattern = "*0_rs_lag")); rm(avgtmax_rs_lag)
-rm(list = ls(pattern = "*wet_lag"))
+rm(list = ls(pattern = "*wet_lag")); #rm(wet_hrs_lag1)
 

### 2016-03-22 veg-change-assessment.Rmd
Update variable names.

@@
 tag_dbh_plot$species <- tolower(tag_dbh_plot$species)
+tag_dbh_plot <- rename(tag_dbh_plot, stem_status = status)
 
@@
 tag_dbh_stem$species <- tolower(tag_dbh_stem$species)
+tag_dbh_stem <- rename(tag_dbh_stem, stem_status = status)
+tag_dbh_stem <- left_join(tag_dbh_stem, select(tagged_stems, plotid, cluster, species, tag, location, stem_status, year, sod_killed))
 
@@
 alt_dbh_plot
+alt_dbh_plot <- rename(alt_dbh_plot, stem_status = status)
 
@@
 alt_dbh_stem
+alt_dbh_stem <- rename(alt_dbh_stem, stem_status = status)
 
@@
 all_stems$plotid <- as.factor(all_stems$plotid)
-all_stems$status <- as.factor(all_stems$plotid)
+all_stems$stem_status <- as.factor(all_stems$stem_status)
 all_stems$year <- as.factor(all_stems$year)
@@
       filter(dbh >= 2.0, tag != 7794) %>% 
-      group_by(plotid, species, year, status, location) %>% 
+      group_by(plotid, species, year, stem_status, location) %>% 
       summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
@@
       tag_dbh_plot %>% filter(location == "In") %>% 
-            select(plotid, species, status, year, everything(), -location), 
+            select(plotid, species, stem_status, year, everything(), -location), 
       select(alt_dbh_plot, -tot_dbh, -avg_dbh))
@@
 os_dbh_plot <- rbind(
-      tag_dbh_plot %>% group_by(plotid, species, status, year) %>% 
+      tag_dbh_plot %>% group_by(plotid, species, stem_status, year) %>% 
             summarise(tot_ba_m2 = sum(tot_ba_m2), avg_ba_m2 = sum(tot_ba_m2)/sum(abundance),
@@
 impt_vals <- rbind(tag_dbh_stem %>% filter(location == "In", dbh >= 2.0) %>% 
-                         select(plotid, species, dbh, status, year), 
+                         select(plotid, species, dbh, stem_status, year), 
                    alt_dbh_stem) %>% 
-      group_by(plotid, species, status, year) %>% 
+      group_by(plotid, species, stem_status, year) %>% 
       summarise(sp_abund = length(species), 
                 sp_basal_area = sum(sum(pi*dbh^2/40000, na.rm = T))) %>% 
-      group_by(plotid, status, year) %>% 
+      group_by(plotid, stem_status, year) %>% 
       mutate(total_abund = sum(sp_abund), 
@@
              filter(species != "unk1", species != "unknown sp.") %>% 
-             group_by(year, species, status) %>% 
+             group_by(year, species, stem_status) %>% 
              summarise(occurrence = length(plotid)), 
@@
       geom_text(nudge_y = 3) +
-      facet_grid(status ~ year) +
+      facet_grid(stem_status ~ year) +
       ylab("# of plots") +
@@
              filter(species != "unk1", species != "unknown sp.") %>% 
-             group_by(year, species, status) %>% 
+             group_by(year, species, stem_status) %>% 
              summarise(abundance = sum(abundance)), 
@@
       geom_bar(stat = "identity") +
-      facet_grid(status ~ year) +
+      facet_grid(stem_status ~ year) +
       ylab("abundance") +
@@
                     species != "unknown sp.") %>% 
-             group_by(year, species, status) %>% 
+             group_by(year, species, stem_status) %>% 
              summarise(abundance = sum(abundance)), 
@@
       geom_text(aes(vjust = 0), nudge_y = 6)
+
+## select overstory species abundance, abundance ≥ 100 stems
+ggplot(os_dbh_in_plot %>%
+             filter(species == "umca" | species == "quag" | species == "quga" | species == "psme" | species == "arme" | species == "quke" | species == "quch") %>% 
+             group_by(year, species, stem_status) %>% 
+             summarise(abundance = sum(abundance)), 
+       aes(species, abundance, label = abundance)) +
+      geom_bar(stat = "identity") +
+      facet_grid(stem_status ~ year, scales = "free") +
+      ylab("abundance") +
+      ggtitle("Overstory Species Abundance") +
+      geom_text(aes(vjust = 0), nudge_y = 8) +
+      theme_bw()
+
+ggplot(os_dbh_in_plot %>%
+             filter(species == "umca" | species == "quag" | species == "quga" | species == "psme" | species == "arme" | species == "quke" | species == "quch" | species == "sese") %>% 
+             group_by(year, species, stem_status) %>% 
+             summarise(tot_ba_m2 = round(sum(tot_ba_m2), 2)), 
+       aes(species, tot_ba_m2, label = tot_ba_m2)) +
+      geom_bar(stat = "identity") +
+      facet_grid(stem_status ~ year, scales = "free") +
+      ylab("basal area") +
+      ggtitle("Overstory Species Basal Area") +
+      geom_text(aes(vjust = 0), nudge_y = 1.5) +
+      theme_bw()
 
@@
              filter(species == "lide" | species == "quag" | species == "quch" |
-                          species == "quke" | species == "quag x quwi" | 
-                          species == "quke x quag" | species == "umca") %>%
-             group_by(species, status, year) %>% 
+                          species == "quke" | species == "umca") %>%
+             group_by(species, stem_status, year) %>% 
              summarise(abundance = sum(abundance)),
@@
       filter(species == "quke" | species == "quag" | species == "quch") %>%
-            group_by(plotid, species, status, year, sod_dead) %>% 
+            group_by(plotid, species, stem_status, year, sod_killed) %>% 
             summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
@@
                 abundance = length(dbh))
+qusp_dbh_plot <- ungroup(qusp_dbh_plot)
+summary(qusp_dbh_plot)
+filter(qusp_dbh_plot, is.na(sod_killed))
+qusp_dbh_plot$sod_killed[is.na(qusp_dbh_plot$sod_killed)] <- "No"
+
 ggplot(qusp_dbh_plot, aes(species, abundance, label = abundance)) +
       geom_bar(stat = "identity") +
-      facet_grid(sod_dead ~ year) +
+      facet_grid(sod_killed ~ year, scales = "free") +
       ylab("abundance") +
@@
                           species == "quke x quag" | species == "umca") %>%
-             group_by(species, status, year) %>% 
+             group_by(species, stem_status, year) %>% 
              summarise(basal_area = round(sum(tot_ba_m2), digits = 4)),
@@
       geom_bar(stat = "identity") +
-      facet_grid(status ~ year) +
+      facet_grid(stem_status ~ year) +
       ylab("basal area (m^2)") +
@@
       geom_boxplot() +
-      facet_grid(status ~ year) +
+      facet_grid(stem_status ~ year) +
       ylab("basal area (m^2)") +
@@
       filter(species == "quke" | species == "quag" | species == "quch") %>%
-            group_by(plotid, species, status, year, sod_dead) %>% 
+            group_by(plotid, species, stem_status, year, sod_dead) %>% 
             summarise(avg_ba_m2 = mean(pi*dbh^2/40000, na.rm = T), 
@@
       geom_boxplot() +
-      facet_grid(status ~ year) +
+      facet_grid(stem_status ~ year) +
       ylab("importance value") +
@@
       geom_boxplot() +
-      # facet_grid(status ~ year) +
+      # facet_grid(stem_status ~ year) +
       ylab("Shannon's Diversity Index") +
@@
       geom_boxplot() +
-      # facet_grid(status ~ year) +
+      # facet_grid(stem_status ~ year) +
       ylab("Pielou's Evenness Index") +
@@
       geom_boxplot() +
-      facet_grid(status ~ year) +
+      facet_grid(stem_status ~ year) +
       #geom_text() +
@@
 
-ggplot(os_dbh_in_plot %>% filter(status == "Alive"), aes(tot_ba_m2)) + 
+ggplot(os_dbh_in_plot %>% filter(stem_status == "Alive"), aes(tot_ba_m2)) + 
       geom_histogram() +
       facet_grid(species ~ year)
-ggplot(os_dbh_in_plot %>% filter(status == "Alive"), aes(avg_ba_m2)) + 
+ggplot(os_dbh_in_plot %>% filter(stem_status == "Alive"), aes(avg_ba_m2)) + 
       geom_histogram() +
       facet_grid(species ~ year)
-ggplot(os_dbh_in_plot %>% filter(status == "Alive"), aes(abundance)) + 
+ggplot(os_dbh_in_plot %>% filter(stem_status == "Alive"), aes(abundance)) + 
       geom_histogram() +
@@
 boxplot(tot_ba_m2 ~ year, 
-        data = os_dbh_in_plot %>% filter(status == "Alive", species == "UMCA"),
+        data = os_dbh_in_plot %>% filter(stem_status == "Alive", species == "UMCA"),
         ylab = "m^2",
@@
 boxplot(tot_ba_m2 ~ year, 
-        data = os_dbh_in_plot %>% filter(status == "Dead", species == "UMCA"),
+        data = os_dbh_in_plot %>% filter(stem_status == "Dead", species == "UMCA"),
         ylab = "m^2",
@@
 boxplot(tot_ba_m2 ~ year, 
-        data = os_dbh_in_plot %>% filter(status == "Alive", species == "QUAG"),
+        data = os_dbh_in_plot %>% filter(stem_status == "Alive", species == "QUAG"),
         ylab = "m^2",
@@
 boxplot(tot_ba_m2 ~ year, 
-        data = os_dbh_in_plot %>% filter(status == "Dead", species == "QUAG"),
+        data = os_dbh_in_plot %>% filter(stem_status == "Dead", species == "QUAG"),
         ylab = "m^2",
@@
 
-ggplot(data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
+ggplot(data = basal_abund %>% filter(stem_status == "Alive", species == "QUAG"),
        aes(year, tot_ba_m2, group = year)) +
@@
 boxplot(abundance ~ year, 
-        data = basal_abund %>% filter(status == "Alive", species == "QUAG"),
+        data = basal_abund %>% filter(stem_status == "Alive", species == "QUAG"),
         ylab = "Number of stems",
@@
 boxplot(tot_ba_m2 ~ year, 
-        data = basal_abund %>% filter(status == "Dead", species == "QUAG"),
+        data = basal_abund %>% filter(stem_status == "Dead", species == "QUAG"),
         ylab = "Number of stems",

### 2016-03-22 climate-path-model.R
Update path models.
Added another infection variable and developed separate models using wet-hours and dry-season temperature variables. This led to reordering some of the code.

@@ -4,7 +4,7 @@
 #+ load data and packages ####
 load("pathmodel_data.RData")
 library(plyr); library(dplyr); library(nlme); library(lme4)
-library(piecewiseSEM)
+library(lavaan); library(piecewiseSEM)
 
 #+ panel functions for pairs plots
 panel.cor <- function(x, y, digits = 2, cex.cor, ...)
@@ -44,13 +44,28 @@ str(oak_sod)
 summary(wet_hrs)# since these are counts the NAs are in fact zeroes
 wet_hrs[is.na(wet_hrs)] <- 0# introduce a lot of zeroes for some columns
 oak_sod <- left_join(
-      oak_sod, select(wet_hrs, -X), by = c("plotid","sample_year"))
+      oak_sod, select(wet_hrs, -X, -hrsabv22_wet, -avgtemp_wet), by = c("plotid","sample_year"))
 oak_sod$plotid <- as.factor(oak_sod$plotid)
 
 summary(rs2d_lag1)
 oak_sod <- left_join(
       oak_sod, select(rs2d_lag1, -rs2d_total), by = c("plotid","sample_year"))
 oak_sod$plotid <- as.factor(oak_sod$plotid)
+
+summary(rs_temps_lag)
+oak_sod <- left_join(
+      oak_sod, 
+      select(rs_temps_lag, plotid, sample_year, contains("tminus"), contains("t1")), by = c("plotid","sample_year"))
+
+summary(summer_temps_lag)
+oak_sod <- left_join(
+      oak_sod,
+      select(summer_temps_lag, plotid, sample_year, contains("tminus"), contains("t1")), by = c("plotid","sample_year"))
+
+summary(wet_hrs_lag)
+oak_sod <- left_join(
+      oak_sod,
+      select(wet_hrs_lag, plotid, sample_year, contains("tminus"), contains("t1")), by = c("plotid","sample_year"))
 
 ## drop 3d, PRISM, and regression rainfall variables based on discussion with Ryan; use a simple 2d spline interpolation or the Voronoi polygon values
 # oak_sod <- select(oak_sod, -rain_tot_3d, -rain_tot_r, -rain_tot_psm)
@@ -77,9 +92,13 @@ pairs(na.omit(select(oak_sod, inf_oak_ct, tot_bay, tot_lfct, avg_lfct, umca_basa
[…]

### 2016-03-22 veg-change-assessment.Rmd
Add plot of subset of species only in oak plots.

@@
 ggplot(os_dbh_in_plot %>%
-             filter(species == "umca" | species == "quag" | species == "quga" | species == "psme" | species == "arme" | species == "quke" | species == "quch" | species == "sese") %>% 
+             filter(species == "quag" | species == "quke" | species == "quga" | species == "umca" | species == "psme" | species == "arme" |  species == "quch" | species == "sese") %>% 
              group_by(year, species, stem_status) %>% 
@@
       ggtitle("Overstory Species Basal Area") +
-      geom_text(aes(vjust = 0), nudge_y = 1.5) +
+      geom_text(aes(vjust = 0), nudge_y = 1) +
+      theme_bw()
+
+qu_plot <- na.omit(left_join(select(oak_sod, plotid, sample_year),
+                     os_dbh_in_plot, 
+                     by = c("plotid","sample_year"="year")))
+summary(qu_plot)
+
+ggplot(qu_plot %>%
+             filter(species == "quag" | species == "quke" | species == "quga" | species == "umca" | species == "psme" | species == "arme" |  species == "quch") %>% 
+             group_by(sample_year, species, stem_status) %>% 
+             summarise(tot_ba_m2 = round(sum(tot_ba_m2), 2)), 
+       aes(species, tot_ba_m2, label = tot_ba_m2)) +
+      geom_bar(stat = "identity") +
+      facet_grid(stem_status ~ sample_year, scales = "free") +
+      ylab("basal area") +
+      ggtitle("Overstory Species Basal Area in Quercus Plots") +
+      geom_text(aes(vjust = 0), nudge_y = 1) +
       theme_bw()

### 2016-04-14 path-temp-compare.Rmd
Add Rmd to build multiple temperature path models.
This Rmd will be the report on my analysis of multiple path models examining the relationships between disease and microclimate across multiple temperature variables.

---
title: "Path Model Temperature Comparisons"
author: "Whalen Dillon"
date: "April 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Based on previous studies, I expect that the summer season temperature and wet season conditions will influence leaf symptoms on California bay laurel. Specifically, hotter and drier summers will result in lower symptomatic leaf counts, because damaged leaves are more likely to abscise in these conditions. Increased abscision of infected leaves results in lower inoculum load to seed the seasonal epidemic during the following wet season. A wet season where rainfall extends into the late spring, when temperatures are warmer, will result in higher symptomatic leaf counts. These climate effects on California bay laurel leaf symptoms should manifest rapidly.

In contrast, symptoms on oak trees (bleeding or oozing cankers) are unlikely to be immediately detectable. Although infection of oak trees most likely happens when there is high inoculum load on California bay laurel seldom will the infection be visibly detectable during the same year. I also assume that conditions need to be favorable for the infection to establish in the oak tree to a degree sufficient to cause visible symptoms. This is supported by Haas et al. (2016), who showed that oak trees experiencing two consecutive years of warm and wet conditions during November-May had an increased probability of exhibiting symptoms of infection. Since I am assessing infection based on visual symptoms I believe that the influence of conditions during the sampling season, particularly with respect to rainfall, are an effect on our ability to visually detect infection as opposed to a true effect on infection of oak trees.

Based on these assessments of the differential relationships between our visual assessments of infection, and how climate and weather effect true infection, I developed temperature and moisture metrics to capture the effect on true infection. These metrics include dry-season temperature variables (the number of hours above or below a threshold temperature) that is intended to capture the influence on the amount of inoculum that over-summers on California bay laurel; a combined temperature/moisture variable that is the number of hours above, below, or within a temperature range during days that had measurable rainfall. Based on Haas et al. (2016), I expected the wet-hours variable averaged over the current and previous wet-seasons to have an effect on oak infection, but a limited statistical influence on California bay laurel leaf symptoms.

While temperature thresholds for growth, reproduction, and survival of *P. ramorum* have been experimentally assessed in laboratory settings, the conditions used have seldom corresponded to natural settings. For example, *P. ramorum* has a growth and reproduction optimum around 18-20 C in laboratory experiments when humdity is 60-70% or greater. In the mediterranean climate of California rarely does a temperature above 18 C correspond to humidity in that range. Still, these experiments and additional observations provided a starting point for developing a set of temperature thresholds to examine using a degree-day approach. By examining the relationships across multiple temperature I am attempting to identify temperatures that may in fact be threshold temperatures where the inferential relationship between disease symptoms and climate changes.

### 2016-06-27 path-temp-compare.Rmd
Add narrative context and mutiple models.
I built a number of regression models to explore the relationship between different temperature thresholds and symptomatic leaf counts observed in the field.

@@
 ---
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
 
@@
 In contrast, symptoms on oak trees (bleeding or oozing cankers) are unlikely to be immediately detectable.
-Although infection of oak trees most likely happens when there is high inoculum load on California bay laurel seldom will the infection be visibly detectable during the same year.
+Although infection of oak trees most likely happens when there is high inoculum load on California bay laurel, seldom will the infection be visibly detectable during the same year.
 I also assume that conditions need to be favorable for the infection to establish in the oak tree to a degree sufficient to cause visible symptoms.
@@
 (2016), who showed that oak trees experiencing two consecutive years of warm and wet conditions during November-May had an increased probability of exhibiting symptoms of infection.
-Since I am assessing infection based on visual symptoms I believe that the influence of conditions during the sampling season, particularly with respect to rainfall, are an effect on our ability to visually detect infection as opposed to a true effect on infection of oak trees.
+Since I am assessing infection based on visual symptoms I believe that the influence of conditions during the sampling season, particularly with respect to rainfall, are an effect on our ability to visually detect infection as opposed to a biological effect on infection of oak trees.
 
-Based on these assessments of the differential relationships between our visual assessments of infection, and how climate and weather effect true infection, I developed temperature and moisture metrics to capture the effect on true infection.
+Based on these assessments of the differential relationships between our visual assessments of infection, and how climate and weather affect true infection, I developed temperature and moisture metrics to capture the effect on true infection.
 These metrics include dry-season temperature variables (the number of hours above or below a threshold temperature) that is intended to capture the influence on the amount of inoculum that over-summers on California bay laurel; a combined temperature/moisture variable that is the number of hours above, below, or within a temperature range during days that had measurable rainfall.
@@
 For example, *P.
-ramorum* has a growth and reproduction optimum around 18-20 C in laboratory experiments when humdity is 60-70% or greater.
-In the mediterranean climate of California rarely does a temperature above 18 C correspond to humidity in that range.
+ramorum* has a growth and reproduction optimum around 18-20C in laboratory experiments when humdity is 60-70% or greater.
+In the mediterranean climate of California rarely does a temperature above 18C correspond to humidity in that range.
 Still, these experiments and additional observations provided a starting point for developing a set of temperature thresholds to examine using a degree-day approach.
-By examining the relationships across multiple temperature I am attempting to identify temperatures that may in fact be threshold temperatures where the inferential relationship between disease symptoms and climate changes.
+By examining the relationships across multiple temperature cutoffs or ranges I am attempting to identify temperatures that may in fact be threshold temperatures where the inferential relationship between disease symptoms and climate variation changes.
+
+```{r load data, echo=F}
+load("climate-path-model-data.RData")
+library(piecewiseSEM); library(dplyr); library(lme4); library(lmerTest)
+oakplots.sub$umca_density_m2 <- oakplots.sub$tot_bay / (15*15)
+oakplots.sub$umca_density_m2.sqrt <- sqrt(oakplots.sub$umca_density_m2)
+```
+
+```{r rescale variables}
+## rescale variables
+# oakplots.sub2$inf_bay.scl <- scale((oakplots.sub2$infected_bay_ct))
+oakplots.sub$tot_bay.scl <- scale(oakplots.sub$tot_bay)
+oakplots.sub$tot_bay.log.scl <- scale(oakplots.sub$tot_bay.log)
+oakplots.sub$umca_basal_area.scl <- scale(oakplots.sub$umca_basal_area)
+oakplots.sub$umca_ba.log.scl <- scale(oakplots.sub$umca_ba.log)
+oakplots.sub$tot_lfct.scl <- scale(oakplots.sub$tot_lfct)
+oakplots.sub$tot_lfct.log.scl <- scale(oakplots.sub$tot_lfct.log)
+# oakplots.sub$avg_lfct.scl <- scale(oakplots.sub$avg_lfct)
+oakplots.sub$hrs_abv25ds.scl <- scale(oakplots.sub$hrs_abv25ds)
+oakplots.sub$avg_tmax_ds.scl <- scale(oakplots.sub$avg_tmax_ds)
+oakplots.sub$avg_hrsblw14_wet_tminus1.scl <- scale(oakplots.sub$avg_hrsblw14_wet_tminus1)
+oakplots.sub$hrs1422_wet_tminus1.log.scl <- scale(oakplots.sub$hrs1422_wet_tminus1.log)
+# oakplots.sub$avg_hrs1422_wet_t_t1.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1)
+oakplots.sub$avg_hrs1422_wet_t_t1.log.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1.log)
+
+## static variables
+oakplots.sub$candens15m.scl <- scale(oakplots.sub$candens15m)
+oakplots.sub$shannons2005.scl <- scale(oakplots.sub$H.2005)
+oakplots.sub$shannons2014.scl <- scale(oakplots.sub$H.2014)
+oakplots.sub$twi15m.scl <- scale(oakplots.sub$twi15m)
+oakplots.sub$twi15m.log.scl <- scale(oakplots.sub$twi15m.log)
+```
+
+This code chunk describes the general structure for the path model for application in the `piecewiseSEM` package methods.
+```{r general path model structure, eval=F}
+model_list <- list(
+      m1 <- lm(Diversity ~ Topographic_Index, data = data),
+      m2 <- lm(Bay_Laurel_Density ~ Topographic_Index, data = data),
+      m3 <- lm(Symptomatic_Leaf_Count ~ Bay_Laurel_Density + Dry_Season_Temperature + Topographic_Index, data = data),
+      # m6 <- lm(candens15m ~ tot_bay.log, data = oakplots.2007),
+      m6 <- lm(Wet_Hours_Temperature ~ Bay_Laurel_Density, data = data),
+      m4 <- lm(Dry_Season_Temperature ~ Bay_Laurel_Density + Topographic_Index, data = data),
+      m5 <- glm(Oak_Disease_Prevalence ~ Symptomatic_Leaf_Count + Wet_Hours_Temperature + Diversity, family = binomial(link = "logit"), data = data)
+)
+
+# model_list <- list(
+#       m1 <- lm(H.2005 ~ twi15m.log, data = oakplots.2007),
+#       m2 <- lm(tot_bay.log ~ twi15m.log, data = oakplots.2007),
+#       m3 <- lm(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + twi15m.log, data = oakplots.2007),
+#       # m6 <- lm(candens15m ~ tot_bay.log, data = oakplots.2007),
+#       m6 <- lm(avg_hrs1422_wet_t_t1 ~ tot_bay.log, data = oakplots.2007),
+#       m4 <- lm(hrs_abv25ds ~ candens15m + twi15m.log, data = oakplots.2007),
+#       m5 <- glm(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1 + H.2005 + candens15m, family = binomial(link = "logit"), data = oakplots.2007)
+# )
+```
+
+Using data on oak infection at the plot-level:
+```{r model pieces}
+names(oakplots.sub)
+oak_and_bay_plots <- filter(oakplots.sub, tot_bay > 0)
+
+model_list1 <- list(
+      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
+
+      umca_density <- lm(tot_bay.log ~ twi15m.log, data = oakplots.sub),
+
+      slc <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      avgtmax_ds <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+)
+model_list1.scl <- list(
+      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),
+
+      umca_density <- lm(tot_bay.log.scl ~ twi15m.log.scl, data = oakplots.sub),
+
+      slc <- lmer(tot_lfct.log.scl ~ tot_bay.log.scl + hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ tot_bay.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      avgtmax_ds <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+)
+
+# Model set reduced to plots that have oaks & at least 1 bay laurel
+oak_and_bay_list1 <- list(
+      shannons <- lm(H.2005 ~ twi15m.log, data = oak_and_bay_plots),
+      #umca_density <- lm(tot_bay.log ~ twi15m.log, data = oak_and_bay_plots),
+      slc <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
+      hrs_abv25 <- lmer(hrs_abv25ds ~ tot_bay.log + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oak_and_bay_plots)
+)
+
+# Comparing fits of the piecewise models using all oak plots vs. plots with oaks and at least one bay laurel
+sem.fit(model_list1, oakplots.sub, 
+        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+sem.fit(model_list1.scl, oakplots.sub,
+          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+
+sem.fit(oak_and_bay_list1, oak_and_bay_plots, 
+        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
+                        "tot_bay.log ~~ H.2005"))
+
+sem.coefs(model_list1, oakplots.sub,
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+sem.coefs(model_list1.scl, oakplots.sub,
+          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+
+sem.coefs(oak_and_bay_list1, oak_and_bay_plots, standardize = "none",
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
+                          "tot_bay.log ~~ H.2005"))
+
+# Model set using average daily maximum temperature during the dry season.
+model_avgtmax_ds <- list(
+      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
+      # umca_density <- lm(tot_bay.log ~ twi15m.log, data = oakplots.sub),
+
+      slc <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      avgtmax_ds <- lmer(avg_tmax_ds ~ tot_bay.log + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + avg_tmax_ds + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+)
+
+model_avgtmax_ds.scl <- list(
+      shannons <- lm(shannons2005.scl ~ twi15m.log.scl + (1|sample_year), data = oakplots.sub),
+      # umca_density <- lm(tot_bay.log ~ twi15m.log, data = oakplots.sub),
+
+      slc <- lmer(tot_lfct.log.scl ~ tot_bay.log.scl + avg_tmax_ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ tot_bay.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      avgtmax_ds <- lmer(avg_tmax_ds.scl ~ tot_bay.log.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + avg_tmax_ds.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+)
+
+oak_and_bay_avgtmax_ds <- list(
+      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oak_and_bay_plots)),
+      # umca_density <- lm(tot_bay.log ~ twi15m.log, data = oak_and_bay_plots),
+
+      slc <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
+
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
+
+      avgtmax_ds <- lmer(avg_tmax_ds ~ tot_bay.log + twi15m.log + (1|plotid) + (1|sample_year), data = oak_and_bay_plots),
+
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + avg_tmax_ds + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oak_and_bay_plots)
+)
+
+sem.fit(model_avgtmax_ds, oakplots.sub,
+        corr.errors = c(
+              "avg_tmax_ds ~~ avg_hrs1422_wet_t_t1.log",
+              "H.2005 ~~ avg_tmax_ds"
+                        )
+        )
+sem.fit(oak_and_bay_avgtmax_ds, oak_and_bay_plots,
+        corr.errors = c("avg_tmax_ds ~~ avg_hrs1422_wet_t_t1.log"
+                        ,"H.2005 ~~ tot_bay.log"
+                        # ,"H.2005 ~~ avg_tmax_ds"
+                        )
+        )
+
+sem.coefs(model_avgtmax_ds, oakplots.sub, standardize = "none",
+          corr.errors = c(
+                "avg_tmax_ds ~~ avg_hrs1422_wet_t_t1.log",
+                "H.2005 ~~ avg_tmax_ds"))
+sem.coefs(model_avgtmax_ds.scl, oakplots.sub, standardize = "none",
+          corr.errors = c(
+                "avg_tmax_ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl",
+                "shannons2005.scl ~~ avg_tmax_ds.scl"))
+
+sem.coefs(oak_and_bay_avgtmax_ds, oak_and_bay_plots, standardize = "none",
+          corr.errors = c("avg_tmax_ds ~~ avg_hrs1422_wet_t_t1.log",
+                        "H.2005 ~~ tot_bay.log"))
+```
+
+Exploratory modeling of symptomatic leaf count with different temperature variables
+```{r slc with dry-season temperature variables}
+# Models using dry-season temperatures (June-September)
+slc_abv25ds <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_abv20ds <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv20ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_blw18ds <- lmer(tot_lfct.log ~ tot_bay.log + hrs_blw18ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avgtmax_ds <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avgtmin_ds <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmin_ds + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+AIC(slc_avgtmax_ds)
+AIC(slc_avgtmin_ds)
+AIC(slc_abv25ds)
+AIC(slc_abv20ds)
+AIC(slc_blw18ds)
+## The best model based on AIC is avgtmax_ds.
+Interestingly, abv20ds & blw18ds have precisely equal support, but show opposite inference.
+More hours above 20C during the dry season reduces symptomatic leaf count the following spring, while more hours below 18C increases symptomatic leaf count.
+library(ggplot2)
+par(mfrow = c(2,1))
+boxplot(tot_lfct.log ~ sample_year, data = oakplots.sub,
+        xlab = "sample year", ylab = "ln(total symptomatic leaf count)")
+boxplot(hrs_abv25ds ~ sample_year, data = oakplots.sub,
+        xlab = "sample_year", ylab = "number of dry-season hours >25C")
+boxplot(avg_lfct.log ~ sample_year, data = oakplots.sub,
+        xlab = "sample year", ylab = "ln(total symptomatic leaf count)")
+
+ggplot(oak_and_bay_plots, aes(hrs_abv25ds, tot_lfct.log)) +
+      geom_point() +
+      ggtitle("Oak & Bay laurel Plots") +
+      xlab("Dry-season Hours >25C") + ylab("log(Total Symptomatic Leaf Count")
+ggplot(oakplots.sub, aes(hrs_abv25ds, tot_lfct.log)) +
+      geom_point() +
+      ggtitle("Oak Plots") +
+      xlab("Dry-season Hours >25C") + ylab("log(Total Symptomatic Leaf Count")
+
+ggplot(oakplots.sub, aes(hrs_abv20ds, tot_lfct.log)) +
+      geom_point() +
+      ggtitle("Oak Plots") +
+      xlab("Dry-season Hours >20C") + ylab("log(Total Symptomatic Leaf Count")
+ggplot(oakplots.sub, aes(hrs_blw18ds, tot_lfct.log)) +
+      geom_point() +
+      ggtitle("Oak Plots") +
+      xlab("Dry-season Hours <18C") + ylab("log(Total Symptomatic Leaf Count")
+
+ggplot(oakplots.sub, aes(avg_tmax_ds, tot_lfct.log)) +
+      geom_point() +
+      ggtitle("Oak Plots") +
+      xlab("Dry-season Average Daily Maximum Temperature (C)") + 
+      ylab("log(Total Symptomatic Leaf Count")
+
+ggplot(oakplots.sub, aes(avg_tmin_ds, tot_lfct.log)) +
+      geom_point() +
+      ggtitle("Oak Plots") +
+      xlab("Dry-season Average Daily Minimum Temperature (C)") + 
+      ylab("log(Total Symptomatic Leaf Count")
+```
+
+## Models of total symptomatic leaf-count using rainy season temperatures
+These models use different rainy-season (November-May) temperature variables to model total symptomatic leaf count on California bay laurel
+```{r slc models using rainy season temperature}
+# Models using rainy-season temperatures (November-May)
+slc_avgtmax_rs <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avgtmin_rs <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmin_rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrs1420_rs <- lmer(tot_lfct.log ~ tot_bay.log + hrs_14_20_rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrs_abv25rs <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrs_abv20_rs <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv20_rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrs_blw10rs <- lmer(tot_lfct.log ~ tot_bay.log + hrs_blw10rs + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+AIC(slc_avgtmax_rs)
+AIC(slc_avgtmin_rs)
+AIC(slc_hrs1420_rs)
+AIC(slc_hrs_abv25rs)
+AIC(slc_hrs_abv20_rs)
+AIC(slc_hrs_blw10rs)
+```
+
+## Models of symptomatic leaf count using temperature on wet days
+These models use different temperature variables calculated for days that had measurable rainfall during the rainy season, based on plots being assigned to the nearest rain gauge.
+```{r slc models using wet-days temperatures}
+# Models using wet-days temperatures
+slc_hrsblw14_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrsblw14_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrsabv22_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrsabv22_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrs1422_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrs1422_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrs1020_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrs1020_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrsabv20_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrsabv20_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrsblw10_wet <- lmer(tot_lfct.log ~ tot_bay.log + hrsblw10_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avgtmin_wet <- lmer(tot_lfct.log ~ tot_bay.log + avgtmin_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avgtmax_wet <- lmer(tot_lfct.log ~ tot_bay.log + avgtmax_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avgtemp_wet <- lmer(tot_lfct.log ~ tot_bay.log + avgtemp_wet + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+AIC(slc_hrsblw14_wet)
+AIC(slc_hrs1422_wet)
+AIC(slc_hrsabv22_wet)
+AIC(slc_hrsblw10_wet)
+AIC(slc_hrs1020_wet)
+AIC(slc_hrsabv20_wet)
+AIC(slc_avgtmax_wet)
+AIC(slc_avgtmin_wet)
+AIC(slc_avgtemp_wet)
+```
+
+## Models using lagged dry-season temperature variables
+
+```{r slc models using lagged dry-season temperature}
+# Models using lagged dry-season temperature variables
+slc_hrs_abv25ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_hrs_blw18ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + hrs_blw18ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avg_tmax_ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + avg_tmax_ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avg_hrsblw18_ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + avg_hrsblw18_ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avgtmax_ds_t_t1 <- lmer(tot_lfct.log ~ tot_bay.log + avgtmax_ds_t_t1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+slc_avg_hrsabv25_ds_tminus1 <- lmer(tot_lfct.log ~ tot_bay.log + avg_hrsabv25_ds_tminus1 + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub)
+
+AIC(slc_hrs_abv25ds_tminus1)
+AIC(slc_hrs_blw18ds_tminus1)
+AIC(slc_avg_tmax_ds_tminus1)
+AIC(slc_avg_hrsblw18_ds_tminus1)
+AIC(slc_avg_hrsabv25_ds_tminus1)
+AIC(slc_avgtmax_ds_t_t1)
+```

### 2016-06-27 sem-data-development.Rmd
Add Julian day and week of year to dataframe for path model.

@@
 stems$plotid <- droplevels(stems$plotid)
+stems$date <- as.Date(stems$date)
+library(lubridate)
+stems$julian_day <- yday(stems$date)
+stems$month <- month(stems$date)
+stems$week <- week(stems$date)
+summary(stems)
+detach("package:lubridate", unload=TRUE)
 # write.csv(plots_env, "analysis/data/env_gis_variables.csv", row.names = F)
@@
 
+```{r plot visit dates}
+plot_visit_dates <- stems %>% 
+      select(plotid, year, julian_day:week) %>% 
+      group_by(plotid, year) %>% 
+      summarise(julian_day = mean(julian_day), month = mean(month), week = mean(week))
+plot_visit_dates <- ungroup(plot_visit_dates)
+```
+
 ## Examine the bay laurel data at individual and plot levels
@@
 bay_laurel <- filter(stems, species == "umca") %>% 
-      select(plotid, cluster:stem_status, umca_slc, notes:dbh_entry)
+      select(plotid, cluster:stem_status, umca_slc, notes:week)
 bay_laurel$species <- droplevels(bay_laurel$species)
@@
                       tot_lfct = sum(umca_slc, na.rm = T)),
-      by = c("plotid", "year"))
+      by = c("plotid","year"))
+umca_plots <- left_join(
+      umca_plots,
+      plot_visit_dates,
+      by = c("plotid","year")
+)
 summary(umca_plots)
@@
 qusp_stems <- stems %>% 
-      select(plotid, year, tag, species, canker, stem_status, infection_confirmed_1plus, infection_confirmed_2plus) %>% 
+      select(plotid, year, tag, species, canker, stem_status, infection_confirmed_1plus, infection_confirmed_2plus, dbh, dbh_entry, julian_day:week) %>% 
       filter(species == "quag" | species == "quke" | species == "quch" |
@@
 oak_sod <- left_join(oak_sod, 
-                     select(umca_plots, plotid, sample_year, uninfected_bay_ct, infected_bay_ct, tot_bay, tot_lfct), 
+                     select(umca_plots, plotid, sample_year, uninfected_bay_ct:tot_lfct ), 
                      by = c("plotid", "sample_year"))
 summary(oak_sod) # the NA's for the bay laurel variables are legitimately zeroes b/c there are no bay in some of these plots
+oak_sod <- left_join(
+      oak_sod,
+      plot_visit_dates,
+      by = c("plotid","sample_year"="year"))
+
 filter(oak_sod, is.na(tot_bay), sample_year != 2003)
@@
 # sumtv02 was abandoned after 2008, so it should be an NA for later years
+filter(oak_sod, is.na(hrs_abv25rs))
+oak_sod$hrs_abv25rs[is.na(oak_sod$hrs_abv25rs)] <- 0
 

### 2016-06-27 climate-path-model.R
Add additional seasonal and climate variables to data subset, and variogram analysis of 2005 data.

@@ -39,18 +39,22 @@ panel.hist <- function(x, ...)
 summary(oak_sod)
 # oak_sod$plotid <- as.factor(oak_sod$plotid)
 str(oak_sod)
+anyDuplicated(oak_sod)
 
 #+ join wet-hours and lagged rainfall data ####
 summary(wet_hrs)# since these are counts the NAs are in fact zeroes
 wet_hrs[is.na(wet_hrs)] <- 0# introduce a lot of zeroes for some columns
+anyDuplicated(wet_hrs)
 oak_sod <- left_join(
-      oak_sod, select(wet_hrs, -X, -hrsabv22_wet, -avgtemp_wet), by = c("plotid","sample_year"))
-oak_sod$plotid <- as.factor(oak_sod$plotid)
+      oak_sod, select(wet_hrs, -X), by = c("plotid","sample_year"))
+anyDuplicated(oak_sod)
+# oak_sod$plotid <- as.factor(oak_sod$plotid)
 
 summary(rs2d_lag1)
 oak_sod <- left_join(
       oak_sod, select(rs2d_lag1, -rs2d_total), by = c("plotid","sample_year"))
-oak_sod$plotid <- as.factor(oak_sod$plotid)
+
+# oak_sod$plotid <- as.factor(oak_sod$plotid)
 
 summary(rs_temps_lag)
 oak_sod <- left_join(
@@ -67,6 +71,8 @@ oak_sod <- left_join(
       oak_sod,
       select(wet_hrs_lag, plotid, sample_year, contains("tminus"), contains("t1")), by = c("plotid","sample_year"))
 
+# oak_sod$plotid <- as.factor(oak_sod$plotid)
+
 ## drop 3d, PRISM, and regression rainfall variables based on discussion with Ryan; use a simple 2d spline interpolation or the Voronoi polygon values
 # oak_sod <- select(oak_sod, -rain_tot_3d, -rain_tot_r, -rain_tot_psm)
 # oak_sod <- filter(oak_sod, plotid != "ponti01", plotid != "sweet01", plotid != "bush01")
@@ -77,43 +83,41 @@ oak_sod <- left_join(
 #' 
 #+ calculate & add umca basal area ####
 umca_ba <- stems %>% filter(year > 2003, species == "umca", stem_status == "Alive", location == "In") %>% select(plotid, year, dbh_entry) %>% group_by(plotid, year) %>% summarise(umca_basal_area = sum(pi*dbh_entry))
[…]

### 2016-07-13 path-temp-compare.Rmd
Add chunk that combines data for all plots.

@@
 oakplots.sub$umca_density_m2.sqrt <- sqrt(oakplots.sub$umca_density_m2)
+```
+
+```{r combine plots data, eval=F, echo=F}
+soco_plots <- stems %>%
+      select(plotid, species, year) %>%
+      group_by(plotid, year) %>%
+      # filter(species == "UMCA", status == "Alive") %>%
+      summarise(stem_total = length(species))
+
+soco_plots <- rename(soco_plots, sample_year = year)
+
+soco_plots <- left_join(soco_plots, select(umca_plots, plotid:tot_lfct), by = c("plotid", "sample_year"))
+soco_plots <- left_join(soco_plots, select(oakplots.sub, plotid:inf_oak_ct), by = c("plotid","sample_year"))
+soco_plots <- left_join(soco_plots, plots_env, by = "plotid")
+summary(soco_plots)
+soco_plots <- select(soco_plots, -stem_total)
+
+write.csv(soco_plots, "analysis/data/soco_plots.csv")
 ```

### 2016-07-15 sem-data-development.Rmd
Recalculate umca rooted within plots to use for calculating umca density.

@@
 ```{r examine umca data}
+# All tagged bay laurel
 bay_laurel <- filter(stems, species == "umca") %>% 
@@
 
-# Data frame of UMCA plots only
+# Only bay laurel rooted within the plot boundaries
+bay_laurel_in <- filter(stems, species == "umca", location == "In") %>% 
+      select(plotid, cluster:stem_status, umca_slc, notes:week)
+bay_laurel_in$species <- droplevels(bay_laurel_in$species)
+summary(bay_laurel_in) # This has stem-level data on bay laurel.
+
+# Data frames of UMCA-only plots, summarizing UMCA abundance and symptomatic leaf count
 umca_plots <- bay_laurel %>%
       # select(plotid, date, species, umca_slc, year, stem_status) %>%
-      group_by(plotid, year, species) %>%
-      filter(species == "umca", stem_status == "Alive") %>%
+      group_by(plotid, year) %>%
+      filter(stem_status == "Alive") %>%
       summarise(uninfected_bay_ct = length(which(umca_slc == 0)), 
                 infected_bay_ct = length(which(umca_slc > 0)), 
-                tot_bay = length(species)) 
+                tot_bay = infected_bay_ct + uninfected_bay_ct,
+                avg_lfct = mean(umca_slc, na.rm = T),
+                tot_lfct = sum(umca_slc, na.rm = T)) 
 summary(umca_plots)
 
+umca_in_plots <- bay_laurel_in %>%
+      # select(plotid, date, species, umca_slc, year, stem_status) %>%
+      group_by(plotid, year) %>%
+      filter(stem_status == "Alive") %>%
+      summarise(uninfected_bay_ct = length(which(umca_slc == 0)), 
+                infected_bay_ct = length(which(umca_slc > 0)), 
+                tot_bay = uninfected_bay_ct + infected_bay_ct,
+                avg_lfct = mean(umca_slc, na.rm = T),
+                tot_lfct = sum(umca_slc, na.rm = T))
+summary(umca_in_plots)
 # length(unique(umca_plots$plotid))# 183 plots with bay laurel
@@
 
-I can use the stem-level data to calculate plot level leaf count variables (sum, average) for each plot and year.
 ```{r calculate plot level leaf counts}
-live_umca <- bay_laurel %>% filter(stem_status == "Alive")
-filter(live_umca, is.na(umca_slc))
-# Spot checking some of these shows that NAs are due to missing data, i.e. they should not be changed to zeroes
-umca_plots <- left_join(
-      umca_plots, 
-      live_umca %>% 
-            group_by(plotid, year) %>% 
-            summarise(avg_lfct = mean(umca_slc, na.rm = T),
-                      tot_lfct = sum(umca_slc, na.rm = T)),
-      by = c("plotid","year"))
 umca_plots <- left_join(
@@
 summary(umca_plots)
+
+umca_in_plots <- left_join(
+      umca_in_plots,
+      plot_visit_dates,
+      by = c("plotid","year")
+)
+summary(umca_in_plots)
 ```
@@
 # Join the environmental data to the bay laurel infection data
-umca_plots <- left_join(select(umca_plots, -species), plots_env, by = "plotid")
+umca_plots <- left_join(umca_plots, plots_env, by = "plotid")
 summary(umca_plots)
@@
 # umca_plots$plotid <- tolower(umca_plots$plotid)
-
+umca_in_plots <- left_join(umca_in_plots, plots_env, by = "plotid")
+umca_in_plots <- rename(umca_in_plots, sample_year = year)
 # Join rainy season and summer temperatures to bay laurel infection data.
@@
 ## I don't think these will end up in my analysis, but no reason to remove them at this time.
+umca_in_plots <- left_join(umca_in_plots, rs_temps, by = c("plotid", "sample_year"))
+umca_in_plots <- left_join(umca_in_plots, summer_temps, by = c("plotid", "sample_year"))
 
@@
 
+umca_in_plots$sample_year <- as.character(umca_in_plots$sample_year)
+umca_in_plots <- left_join(umca_in_plots, rain_total, by = c("plotid", "sample_year"))
+umca_in_plots <- left_join(umca_in_plots, rain_days, by = c("plotid", "sample_year"))
+
 filter(umca_plots, is.na(rs2d_total))
@@
 umca_plots <- left_join(umca_plots, os.us.diversity, by = c("plotid"))
+umca_in_plots <- left_join(umca_in_plots, os.us.diversity, by = c("plotid"))
 summary(umca_plots)
@@
 oak_sod <- left_join(oak_sod, 
-                     select(umca_plots, plotid, sample_year, uninfected_bay_ct:tot_lfct ), 
-                     by = c("plotid", "sample_year"))
+                     select(umca_plots, plotid, sample_year, 
+                            avg_lfct, tot_lfct ),
+                     by = c("plotid","sample_year"))
+names(umca_in_plots)
+umca_in_plots$sample_year <- as.character(umca_in_plots$sample_year)
+oak_sod$sample_year <- as.character(oak_sod$sample_year)
+oak_sod <- left_join(oak_sod,
+                     select(umca_in_plots, plotid, sample_year,
+                            tot_bay, infected_bay_ct),
+                     by = c("plotid","sample_year"))
 summary(oak_sod) # the NA's for the bay laurel variables are legitimately zeroes b/c there are no bay in some of these plots
+plot_visit_dates$year <- as.character(plot_visit_dates$year)
 oak_sod <- left_join(
@@
 # Change NAs of bay laurel in oak only plots to zeroes
-oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
+# oak_sod$uninfected_bay_ct[is.na(oak_sod$uninfected_bay_ct)] <- 0
 oak_sod$infected_bay_ct[is.na(oak_sod$infected_bay_ct)] <- 0
@@
 summer_temps
+oak_sod$sample_year <- as.integer(oak_sod$sample_year)
 oak_sod <- left_join(oak_sod, rs_temps, by = c("plotid", "sample_year"))

### 2016-07-15 analysis/explore_data.R
Revise calculation of plot-level umca data.

@@ -223,7 +223,7 @@ plots_umca <- stems %>%
       select(plot, Date, species, slc, year, status) %>%
       group_by(plot, year, species) %>%
       filter(species == "UMCA", status == "Alive") %>%
-      summarise(uninfected_bay_ct = length(which(slc==0)), infected_bay_ct = length(which(slc > 0)), tot_bay = length(species)) 
+      summarise(uninfected_bay_ct = length(which(slc==0)), infected_bay_ct = length(which(slc > 0)), tot_bay = length(species))
 
 # plots_umca$infected <- ifelse(plots_umca$infected_bay_ct==0, 0, 1)
 summary(plots_umca)

### 2016-07-15 analysis/species_diversity_calc.R
Comment out code that removed decommissioned plots.

@@ -275,7 +275,7 @@ untagged_dbh_2014 %>% group_by(plotid, year) %>% summarise(untagged_rich = lengt
 # Calculate the richness of the tagged stems ####
 stems <- as.tbl(read.csv("analysis/data/tag-dbh_0514-corrected.csv"))
 summary(stems)
-unique(live_tagged$plotid)
+# unique(live_tagged$plotid)
 # stems$plotid <- tolower(stems$plotid)
 # stems$species <- tolower(stems$species)
 
@@ -305,16 +305,23 @@ summary(tagged_abund_2014)
 # Join tagged and untagged stems for 2005
 tagged_abund_2005; unique(tagged_abund_2005$species)
 untagged_dbh_2005; unique(untagged_dbh_2005$species)
+untagged_dbh_2005 <- ungroup(untagged_dbh_2005) %>% 
+      select(plotid, species, live_count)
 
 ## 2005 data
 abund_2005 <- rbind(tagged_abund_2005, select(untagged_dbh_2005, plotid, species, live_count))
-sort(unique(abund_2005$species))# 25 species
+sort(unique(abund_2005$species))# 26 species, remove 'unknown sp.'
+abund_2005 <- droplevels(filter(abund_2005, species != "unknown sp."))
 
+# reformat so that abundance of each species is shown in each plot
 abund_2005_wide <- ungroup(abund_2005) %>% spread(species, live_count)
 
 ## 2014 data
-abund_2014 <- rbind(tagged_abund_2014, select(untagged_dbh_2014, plotid, species, live_count))
-sort(unique(abund_2014$species))# 24 species
+abund_2014 <- rbind(tagged_abund_2014, ungroup(select(untagged_dbh_2014, plotid, species, live_count)))
+sort(unique(abund_2014$species))# 25 species, need to remove 'unknown sp.'
+abund_2014 <- droplevels(filter(abund_2014, species != "unknown sp."))
+
+# reformat so that abundance of each species is shown in each plot
 abund_2014_wide <- abund_2014 %>% spread(species, live_count)
 
 # Summarize into overstory species richness - number of each species in each plot for each survey
@@ -332,8 +339,9 @@ unique(os_rich2014$plotid)
 os_richness <- left_join(os_rich2005, os_rich2014, by = "plotid")
 summary(os_richness)
 filter(os_richness, is.na(os_rich14))
+
 # remove abandoned plots
-os_richness <- filter(os_richness, plotid != "bush01", plotid != "ann28", plotid != "halow01", plotid != "sweet01")
[…]

### 2016-07-15 path-temp-compare.Rmd
Add calculation of absolute umca density variables.

@@
 oakplots.sub$umca_density_m2 <- oakplots.sub$tot_bay / (15*15)
+oakplots.sub$umca_density_ha <- (oakplots.sub$tot_bay / (15*15))*10000
 oakplots.sub$umca_density_m2.sqrt <- sqrt(oakplots.sub$umca_density_m2)
+oakplots.sub$umca_density_ha.sqrt <- sqrt(oakplots.sub$umca_density_ha)
 ```
@@
 oakplots.sub$tot_bay.log.scl <- scale(oakplots.sub$tot_bay.log)
+oakplots.sub$umca_density_m2.scl <- scale(oakplots.sub$umca_density_m2)
+oakplots.sub$umca_density_m2.sqrt.scl <- scale(oakplots.sub$umca_density_m2.sqrt)
 oakplots.sub$umca_basal_area.scl <- scale(oakplots.sub$umca_basal_area)

### 2016-07-15 climate-path-model.R
Comment out pairs plots.
This will make for more efficient code sourcing as needed.

@@ -81,7 +81,7 @@ oak_sod <- left_join(
 #' 
 #' A single large tree supporting a lot of infection can have a substantial epidemiological role in the plot, and likely beyond the plot. I think total leaf count underrepresents the role of these stems, whereas the average leaf count may underrepresent the role of many small stems in producing a high total inoculum load. I guess the leads to the question of whether I want to use the average inoculum load on bay laurel in the plot, capturing the role of a few large trees with potentially high inoculum load, or the total inoculum load on bay laurel, which ignores the role of a few large trees while capturing some of the effect of stem density. 
 #' 
-#+ calculate & add umca basal area ####
+#+ calculate & add "in-plot" living umca basal area ####
 umca_ba <- stems %>% filter(year > 2003, species == "umca", stem_status == "Alive", location == "In") %>% select(plotid, year, dbh_entry) %>% group_by(plotid, year) %>% summarise(umca_basal_area = sum(pi*dbh_entry))
 oak_sod <- left_join(oak_sod, ungroup(umca_ba), by = c("plotid","sample_year" = "year"))
 
@@ -120,17 +120,17 @@ oak_sod$avg_lfct[oak_sod$plotid == "yahng02" & oak_sod$sample_year == 2010] <- N
 #       lower.panel = panel.cor, diag.panel = panel.hist)
 
 #+ pairs plots of oak plots data ####
-pairs(na.omit(select(oak_sod, inf_oak_ct, tot_bay, tot_lfct, avg_lfct, hrsblw14_wet, hrsblw10_wet, hrs1020_wet, hrsabv20_wet, avgtmax_wet, avgtmin_wet, rs2d_total, ss2d_total, rs_days, ss_days)), 
-      lower.panel = panel.cor, diag.panel = panel.hist,
-      main = "Oak Plots Current Year Rainfall Variables")
-
-pairs(na.omit(select(oak_sod, inf_oak_ct, tot_bay, tot_lfct, avg_lfct, starts_with("avg_rs2d"), contains("rs2d_t"))), 
-      lower.panel = panel.cor, diag.panel = panel.hist,
-      main = "Oak Plots Lagged Rainfall Variables")
-
[…]

### 2016-07-15 path-temp-compare.Rmd
Revise model formulae and add regression residual plots.

--- 2016-07-13 path-temp-compare.Rmd
@@
 library(piecewiseSEM); library(dplyr); library(lme4); library(lmerTest)
-oakplots.sub$umca_density_m2 <- oakplots.sub$tot_bay / (15*15)
-oakplots.sub$umca_density_m2.sqrt <- sqrt(oakplots.sub$umca_density_m2)
+
+# Calculate oak density variable as stems per hectare 
+oakplots.sub$umca_density_ha <- (oakplots.sub$tot_bay / (15*15))*10000
+oakplots.sub$umca_density_ha.log <- log(oakplots.sub$umca_density_ha+1)
 ```
@@
 
-```{r rescale variables}
+```{r rescale variables, echo=F}
 ## rescale variables
@@
 oakplots.sub$tot_bay.log.scl <- scale(oakplots.sub$tot_bay.log)
+oakplots.sub$umca_density_ha.scl <- scale(oakplots.sub$umca_density_ha)
+oakplots.sub$umca_density_ha.log.scl <- scale(oakplots.sub$umca_density_ha.log)
 oakplots.sub$umca_basal_area.scl <- scale(oakplots.sub$umca_basal_area)
@@
 oakplots.sub$hrs_abv25ds.scl <- scale(oakplots.sub$hrs_abv25ds)
+oakplots.sub$dys_abv25ds.scl <- scale(oakplots.sub$dys_abv25ds)
 oakplots.sub$avg_tmax_ds.scl <- scale(oakplots.sub$avg_tmax_ds)
@@
 Using data on oak infection at the plot-level:
-```{r model pieces}
-names(oakplots.sub)
-oak_and_bay_plots <- filter(oakplots.sub, tot_bay > 0)
+```{r hours >25 path model all oak plots unscaled, echo=F}
+model_list1 <- list(
+      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.sub, sample_year == 2005)),
 
-model_list1 <- list(
-      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
-
-      umca_density <- lm(tot_bay.log ~ twi15m.log, data = oakplots.sub),
+      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
@@
 
-      avgtmax_ds <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
@@
 )
+```
+
+Path model with the continuous variables standardized using the default settings of the `scale` function, resulting in each having a mean of zero and standard deviation of one.
+Each variable is first  centered by subtracting its mean and then scaled by dividing by its standard deviation.
+```{r hours >25 path model all oak plots scaled}
 model_list1.scl <- list(
@@
 
-      umca_density <- lm(tot_bay.log.scl ~ twi15m.log.scl, data = oakplots.sub),
+      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
@@
 
-      avgtmax_ds <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
@@
 )
+```
 
-# Model set reduced to plots that have oaks & at least 1 bay laurel
+Examining the residual plots of each piece of the path model
+```{r shannons residuals plot}
+par(mfrow=c(2,2))
+plot(shannons, main = "OLS regression of diversity ~ wetness index")
+```
+
+```{r bay laurel density residuals plot}
+plot(umca_density, main = "LMM of bay laurel density")
+```
+
+```{r symptomatic leaf count residuals plot}
+plot(slc, main = "LMM of symptomatic leaf count")
+```
+
+```{r wet-hours residuals plot}
+plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year")
+```
+
+```{r dry-season hours residuals plot}
+plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season")
+```
+
+```{r oak infection residuals plot}
+plot(oak_infection, main = "Binomial GLMM of oak infection")
+```
+
+Model set reduced to plots that have oaks & at least 1 bay laurel
+```{r path model with oak plots having >0 bay laurel}
+oak_and_bay_plots <- filter(oakplots.sub, tot_bay > 0)
 oak_and_bay_list1 <- list(
@@
 )
+```
 
+```{r fit and compare all oak plots vs >0 bay laurel oak plots}
 # Comparing fits of the piecewise models using all oak plots vs. plots with oaks and at least one bay laurel
@@
                         "tot_bay.log ~~ H.2005"))
+```
 
+Get the path coeffcients for the fitted path models
+```{r get path coefficients}
 sem.coefs(model_list1, oakplots.sub,
@@
                           "tot_bay.log ~~ H.2005"))
+```
 
+```{r path models using avg dry season tmax}
 # Model set using average daily maximum temperature during the dry season.

### 2016-07-15 path-temp-compare.Rmd
Add residuals vs. predictor plots for umca density and diversity models.

@@
 
-Using data on oak infection at the plot-level:
-```{r hours >25 path model all oak plots unscaled, echo=F}
-model_list1 <- list(
-      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.sub, sample_year == 2005)),
-
-      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
-
-      slc <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
-
-      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
-
-      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
-
-      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
-)
-```
-
 Path model with the continuous variables standardized using the default settings of the `scale` function, resulting in each having a mean of zero and standard deviation of one.
@@
 
+```{r hours >25 path model all oak plots unscaled, echo=F}
+model_list1 <- list(
+      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
+
+      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      slc <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+)
+```
+
 Examining the residual plots of each piece of the path model
@@
 par(mfrow=c(2,2))
-plot(shannons, main = "OLS regression of diversity ~ wetness index")
+plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
+par(mfrow=c(1,1))
+plot(shannons$model$twi15m.log, resid(shannons),
+     xlab = "wetness index",
+     ylab = "model residuals",
+     main = "Diversity model residuals vs. wetness index predictor")
 ```
@@
 ```{r bay laurel density residuals plot}
-plot(umca_density, main = "LMM of bay laurel density")
+plot(umca_density, main = "LMM of bay laurel density residuals")
+plot(umca_density@frame$twi15m.log, resid(umca_density),
+     xlab = "wetness index",
+     ylab = "model residuals",
+     main = "Bay laurel density model residuals vs. wetness index predictor")
 ```
@@
 ```{r symptomatic leaf count residuals plot}
-plot(slc, main = "LMM of symptomatic leaf count")
+plot(slc, main = "LMM of symptomatic leaf count residuals")
 ```
@@
 ```{r wet-hours residuals plot}
-plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year")
+plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
 ```
@@
 ```{r dry-season hours residuals plot}
-plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season")
+plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
 ```
@@
 ```{r oak infection residuals plot}
-plot(oak_infection, main = "Binomial GLMM of oak infection")
+plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
 ```

### 2016-07-18 path-temp-compare.Rmd
Add residual plots for remaining models.
I also had to revert to the 1.1.1 version of the 'piecewiseSEM' package because of an error about matrices mis-matching when I used the most recent version. The error arises in the 'sem.missing.paths' function I think, but that's where my investigating ended.

@@
 load("climate-path-model-data.RData")
-library(piecewiseSEM); library(dplyr); library(lme4); library(lmerTest)
+library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
+library(dplyr)
+library(lme4)
+library(lmerTest)
 
@@
 # oakplots.sub2$inf_bay.scl <- scale((oakplots.sub2$infected_bay_ct))
-oakplots.sub$tot_bay.scl <- scale(oakplots.sub$tot_bay)
-oakplots.sub$tot_bay.log.scl <- scale(oakplots.sub$tot_bay.log)
-oakplots.sub$umca_density_ha.scl <- scale(oakplots.sub$umca_density_ha)
-oakplots.sub$umca_density_ha.log.scl <- scale(oakplots.sub$umca_density_ha.log)
-oakplots.sub$umca_basal_area.scl <- scale(oakplots.sub$umca_basal_area)
-oakplots.sub$umca_ba.log.scl <- scale(oakplots.sub$umca_ba.log)
-oakplots.sub$tot_lfct.scl <- scale(oakplots.sub$tot_lfct)
-oakplots.sub$tot_lfct.log.scl <- scale(oakplots.sub$tot_lfct.log)
+oakplots.sub$tot_bay.scl <- as.numeric(scale(oakplots.sub$tot_bay))
+oakplots.sub$tot_bay.log.scl <- as.numeric(scale(oakplots.sub$tot_bay.log))
+oakplots.sub$umca_density_ha.scl <- as.numeric(scale(oakplots.sub$umca_density_ha))
+oakplots.sub$umca_density_ha.log.scl <- as.numeric(scale(oakplots.sub$umca_density_ha.log))
+oakplots.sub$umca_basal_area.scl <- as.numeric(scale(oakplots.sub$umca_basal_area))
+oakplots.sub$umca_ba.log.scl <- as.numeric(scale(oakplots.sub$umca_ba.log))
+oakplots.sub$tot_lfct.scl <- as.numeric(scale(oakplots.sub$tot_lfct))
+oakplots.sub$tot_lfct.log.scl <- as.numeric(scale(oakplots.sub$tot_lfct.log))
 # oakplots.sub$avg_lfct.scl <- scale(oakplots.sub$avg_lfct)
-oakplots.sub$hrs_abv25ds.scl <- scale(oakplots.sub$hrs_abv25ds)
-oakplots.sub$dys_abv25ds.scl <- scale(oakplots.sub$dys_abv25ds)
-oakplots.sub$avg_tmax_ds.scl <- scale(oakplots.sub$avg_tmax_ds)
-oakplots.sub$avg_hrsblw14_wet_tminus1.scl <- scale(oakplots.sub$avg_hrsblw14_wet_tminus1)
-oakplots.sub$hrs1422_wet_tminus1.log.scl <- scale(oakplots.sub$hrs1422_wet_tminus1.log)
+oakplots.sub$hrs_abv25ds.scl <- as.numeric(scale(oakplots.sub$hrs_abv25ds))
+oakplots.sub$dys_abv25ds.scl <- as.numeric(scale(oakplots.sub$dys_abv25ds))
+oakplots.sub$avg_tmax_ds.scl <- as.numeric(scale(oakplots.sub$avg_tmax_ds))
+oakplots.sub$avg_hrsblw14_wet_tminus1.scl <- as.numeric(scale(oakplots.sub$avg_hrsblw14_wet_tminus1))
+oakplots.sub$hrs1422_wet_tminus1.log.scl <- as.numeric(scale(oakplots.sub$hrs1422_wet_tminus1.log))
 # oakplots.sub$avg_hrs1422_wet_t_t1.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1)
-oakplots.sub$avg_hrs1422_wet_t_t1.log.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1.log)
+oakplots.sub$avg_hrs1422_wet_t_t1.log.scl <- as.numeric(scale(oakplots.sub$avg_hrs1422_wet_t_t1.log))
 
 ## static variables
-oakplots.sub$candens15m.scl <- scale(oakplots.sub$candens15m)
-oakplots.sub$shannons2005.scl <- scale(oakplots.sub$H.2005)
-oakplots.sub$shannons2014.scl <- scale(oakplots.sub$H.2014)
-oakplots.sub$twi15m.scl <- scale(oakplots.sub$twi15m)
-oakplots.sub$twi15m.log.scl <- scale(oakplots.sub$twi15m.log)
+oakplots.sub$candens15m.scl <- as.numeric(scale(oakplots.sub$candens15m))
+oakplots.sub$shannons2005.scl <- as.numeric(scale(oakplots.sub$H.2005))
+oakplots.sub$shannons2014.scl <- as.numeric(scale(oakplots.sub$H.2014))
+oakplots.sub$twi15m.scl <- as.numeric(scale(oakplots.sub$twi15m))
+oakplots.sub$twi15m.log.scl <- as.numeric(scale(oakplots.sub$twi15m.log))
 ```
@@
 
-      slc <- lmer(tot_lfct.log.scl ~ tot_bay.log.scl + hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      slc <- lmer(tot_lfct.log.scl ~ umca_density_ha.log.scl + hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
-      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ tot_bay.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ umca_density_ha.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
@@
 
-      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      umca_density <- lmer(tot_bay.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
@@
 plot(slc, main = "LMM of symptomatic leaf count residuals")
+plot(slc@frame$umca_density_ha.log, resid(slc),
+     xlab = "bay laurel density",
+     ylab = "model residuals",
+     main = "SLC model residuals vs. bay laurel density predictor")
+plot(slc@frame$hrs_abv25ds, resid(slc),
+     xlab = "# dry-season hours >25 C",
+     ylab = "model residuals",
+     main = "SLC model residuals vs. dry-season hours >25 C predictor")
+plot(slc@frame$avg_hrs1422_wet_t_t1.log,
+     xlab = "avg. wet-hours 14-22C",
+     ylab = "model residuals",
+     main = "SLC model residuals vs. wet-hours predictor")
+plot(slc@frame$H.2005, resid(slc),
+     xlab = "Shannon's diversity index",
+     ylab = "model residuals",
+     main = "SLC model residuals vs. Shannon's diversity predictor")
+plot(slc@frame$twi15m.log, resid(slc),
+     xlab = "wetness index",
+     ylab = "model residuals",
+     main = "SLC model residuals vs. wetness index predictor")
 ```
@@
 plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
+plot(wet_hours_1422@frame$umca_density_ha.log, resid(wet_hours_1422),
+     xlab = "bay laurel density",
+     ylab = "model residuals",
+     main = "Wet-hours model residuals vs. bay laurel density predictor")
 ```
@@
 plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
+plot(hrs_abv25@frame$twi15m.log, resid(hrs_abv25),
+     xlab = "wetness index",
+     ylab = "model residuals",
+     main = "Dry-season hours model residuals vs. wetness index predictor")
 ```
@@
 plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
+plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
+     xlab = "symptomatic leaf count",
+     ylab = "model residuals",
+     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
+plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
+     xlab = "avg. wet-hours 14-22C",
+     ylab = "model residuals",
+     main = "Oak infectin model residuals vs. wet-hours predictor")
+plot(oak_infection@frame$H.2005, resid(oak_infection),
+     xlab = "Shannon's diversity index",
+     ylab = "model residuals",
+     main = "Oak infection model residuals vs. diversity predictor")
 ```
@@
 
-sem.fit(oak_and_bay_list1, oak_and_bay_plots, 
-        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
-                        "tot_bay.log ~~ H.2005"))
+# sem.fit(oak_and_bay_list1, oak_and_bay_plots, 
+#         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
+#                         "tot_bay.log ~~ H.2005"))
 ```
@@
 
-sem.coefs(oak_and_bay_list1, oak_and_bay_plots, standardize = "none",
-          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
-                          "tot_bay.log ~~ H.2005"))
+# sem.coefs(oak_and_bay_list1, oak_and_bay_plots, standardize = "none",
+#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log",
+#                           "tot_bay.log ~~ H.2005"))
 ```

### 2016-07-19 pathogen-spillover-diss-ch.Rmd
Add write up of pathogen spillover path model results.
These results are for the path model using the number of hours >25C during the dry-season and number of wet-hours 14-22C

---
title: "Pathogen Spillover - dissertation chapter"
author: "Whalen Dillon"
date: "July 18, 2016"
output: html_document
---

# CHAPTER 3: EFFECTS OF DIVERSITY, TOPOGRAPHY, AND INTERANNUAL CLIMATE VARIABILITY ON PATHOGEN SPILLOVER

## Results

```{r load data, echo=F, message=F}
library(knitr)
load("climate_path_model_data.RData")
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
library(dplyr, verbose = F)
library(lme4, verbose = F)
library(lmerTest, verbose = F)

# Calculate oak density variable as stems per hectare 
oakplots.sub$umca_density_ha <- (oakplots.sub$tot_bay / (15*15))*10000
oakplots.sub$umca_density_ha.log <- log(oakplots.sub$umca_density_ha+1)
```

```{r rescale variables, echo=F}
## rescale variables
# oakplots.sub2$inf_bay.scl <- scale((oakplots.sub2$infected_bay_ct))
oakplots.sub$tot_bay.scl <- as.numeric(scale(oakplots.sub$tot_bay))
oakplots.sub$tot_bay.log.scl <- as.numeric(scale(oakplots.sub$tot_bay.log))
oakplots.sub$umca_density_ha.scl <- as.numeric(scale(oakplots.sub$umca_density_ha))
oakplots.sub$umca_density_ha.log.scl <- as.numeric(scale(oakplots.sub$umca_density_ha.log))
oakplots.sub$umca_basal_area.scl <- as.numeric(scale(oakplots.sub$umca_basal_area))
oakplots.sub$umca_ba.log.scl <- as.numeric(scale(oakplots.sub$umca_ba.log))
oakplots.sub$tot_lfct.scl <- as.numeric(scale(oakplots.sub$tot_lfct))
oakplots.sub$tot_lfct.log.scl <- as.numeric(scale(oakplots.sub$tot_lfct.log))
# oakplots.sub$avg_lfct.scl <- scale(oakplots.sub$avg_lfct)
oakplots.sub$hrs_abv25ds.scl <- as.numeric(scale(oakplots.sub$hrs_abv25ds))
oakplots.sub$dys_abv25ds.scl <- as.numeric(scale(oakplots.sub$dys_abv25ds))
oakplots.sub$avg_tmax_ds.scl <- as.numeric(scale(oakplots.sub$avg_tmax_ds))
oakplots.sub$avg_hrsblw14_wet_tminus1.scl <- as.numeric(scale(oakplots.sub$avg_hrsblw14_wet_tminus1))
oakplots.sub$hrs1422_wet_tminus1.log.scl <- as.numeric(scale(oakplots.sub$hrs1422_wet_tminus1.log))
# oakplots.sub$avg_hrs1422_wet_t_t1.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1)
oakplots.sub$avg_hrs1422_wet_t_t1.log.scl <- as.numeric(scale(oakplots.sub$avg_hrs1422_wet_t_t1.log))

## static variables
oakplots.sub$candens15m.scl <- as.numeric(scale(oakplots.sub$candens15m))
oakplots.sub$shannons2005.scl <- as.numeric(scale(oakplots.sub$H.2005))
oakplots.sub$shannons2014.scl <- as.numeric(scale(oakplots.sub$H.2014))
oakplots.sub$twi15m.scl <- as.numeric(scale(oakplots.sub$twi15m))
oakplots.sub$twi15m.log.scl <- as.numeric(scale(oakplots.sub$twi15m.log))
```

### Path model 
Path model with the continuous variables standardized using the default settings of the `scale` function, resulting in each having a mean of zero and standard deviation of one. Each variable is first  centered by subtracting its mean and then scaled by dividing by its standard deviation.
```{r hours >25 path model all oak plots unscaled, echo=F}
model_list1 <- list(
      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),

      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ umca_density_ha.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)
```

```{r hours >25 path model all oak plots scaled}
model_list1.scl <- list(
      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),

      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      slc <- lmer(tot_lfct.log.scl ~ umca_density_ha.log.scl + hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ umca_density_ha.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),

      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
)
```

### Examining the residual plots of each piece of the path model
**Residual plots of Shannon's Diversity Index model**
```{r shannons residuals plot, echo=F}
plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
plot(shannons$model$twi15m.log.scl, resid(shannons),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Diversity model residuals vs. wetness index predictor")
```

**Residual plots of bay laurel density model**
```{r bay laurel density residuals plot, echo=F}
plot(umca_density, main = "LMM of bay laurel density residuals")
plot(umca_density@frame$twi15m.log.scl, resid(umca_density),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Bay laurel density model residuals vs. wetness index predictor")
```

**Residual plots of symptomatic leaf count model**
```{r symptomatic leaf count residuals plot, echo=F}
plot(slc, main = "LMM of symptomatic leaf count residuals")
plot(slc@frame$umca_density_ha.log.scl, resid(slc),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "SLC model residuals vs. bay laurel density predictor")
plot(slc@frame$hrs_abv25ds.scl, resid(slc),
     xlab = "# dry-season hours >25 C",
     ylab = "model residuals",
     main = "SLC model residuals vs. dry-season hours >25 C predictor")
plot(slc@frame$avg_hrs1422_wet_t_t1.log.scl,
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "SLC model residuals vs. wet-hours predictor")
plot(slc@frame$shannons2005.scl, resid(slc),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "SLC model residuals vs. Shannon's diversity predictor")
plot(slc@frame$twi15m.log.scl, resid(slc),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "SLC model residuals vs. wetness index predictor")
```

**Residual plots of average number of wet-hours model**
```{r wet-hours residuals plot, echo=F}
plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
     xlab = "bay laurel density",
     ylab = "model residuals",
     main = "Wet-hours model residuals vs. bay laurel density predictor")
```

**Residual plots of dry-season temperature >25C model**
```{r dry-season hours residuals plot, echo=F}
plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
plot(hrs_abv25@frame$twi15m.log.scl, resid(hrs_abv25),
     xlab = "wetness index",
     ylab = "model residuals",
     main = "Dry-season hours model residuals vs. wetness index predictor")
```

**Residual plots of oak infection model**
```{r oak infection residuals plot, echo=F}
plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
plot(oak_infection@frame$tot_lfct.log.scl, resid(oak_infection),
     xlab = "symptomatic leaf count",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log.scl, resid(oak_infection),
     xlab = "avg. wet-hours 14-22C",
     ylab = "model residuals",
     main = "Oak infectin model residuals vs. wet-hours predictor")
plot(oak_infection@frame$shannons2005.scl, resid(oak_infection),
     xlab = "Shannon's diversity index",
     ylab = "model residuals",
     main = "Oak infection model residuals vs. diversity predictor")
```

### Fitting the path model
Fitting the path model by implementing the d-sep test using standardized and unstandardized variables with the assumption of correlated errors between the two climate variables. All continuous variables were standardized to a mean of zero and standard deviation of one. The results for the conditional claims of independence for the standardized and unstandardized variables are reported in the tables. Statistically significant p-values indicate possible missing paths, however, the overall model may still be accepted if the p-value of the C-statistic ("fisher.c") is >0.05. In this case, the significant missing path between diversity and bay laurel density makes ecological sense, but neither variable can be determined as causal to the other. I chose to maintain the initial hypothesized structure of the path model since the overall fit of the model was still acceptable, though this relationship could also be included as a correlated error.
```{r fit and compare all oak plots, echo=F, }
raw_variables_model <- sem.fit(model_list1, oakplots.sub, 
        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_variables_model <- sem.fit(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
```

The overall model fit is reported to the right of of independence claims in these tables where 'fisher.c' and additional model results are reported.
```{r tables of d-sep test results from sem.fit, echo=F}
kable(raw_variables_model, 
      caption = "Values for the d-sep test using raw variables.")
kable(scaled_variables_model,
      caption = "Values for the d-sep test using standardized variables.")
```

### Path coefficients
Tables of the path coeffcients for the fitted path models with raw and standardized variables. The standardized coefficients of the predictors that have the same response variable can be interpreted relative to each other in terms of influence on the response variable.
```{r get path coefficients, echo=F}
raw_coefficients <- sem.coefs(model_list1, oakplots.sub,
          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
scaled_coefficients <- sem.coefs(model_list1.scl, oakplots.sub,
          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
```
```{r tables of path coefficients, echo=F}
kable(raw_coefficients,
      caption = "Raw path coefficients")
kable(scaled_coefficients,
      caption = "Standardized path coefficients")
```
There were five non-significant path coefficients at the p=0.05 level. Two of these (symptomatic leaf count ~ wet-hours and wet-hours ~ bay laurel density) were marginal, with p-values <0.10, which may warrant further investigation and I will address the ecological relevance in the Discussion section. The topographic wetness index was not a statistically significant predictor of bay laurel density nor symptomatic leaf count after accounting for other predictor variables and additional variability due to unmeasured plot-level factors and repeated measures. The diversity index was not a statistically significant predictor of symptomatic leaf count

```{r print session info}
print(sessionInfo())
```

### 2016-07-19 climate-path-model.R
Rename RData object export.

@@ -406,7 +406,7 @@ anyDuplicated(oakplots.2014)
 #       lower.panel = panel.cor, diag.panel = panel.hist,
 #       main = "Oak Plots Data 2014 Sample Season")
 
-save.image("climate-path-model-data.RData")
+save.image("climate_path_model_data.RData")
 #'
 
 

### 2016-07-19 pathogen-spillover-diss-ch.Rmd
Add how to interpret path coefficients.
Also, began reporting of results with non-significant path coefficients.

@@
 Tables of the path coeffcients for the fitted path models with raw and standardized variables.
-The standardized coefficients of the predictors that have the same response variable can be interpreted relative to each other in terms of influence on the response variable.
+The path model provides direct and indirect effects of predictor variables on multiple responses, which can be combined to calculate the total effect of a predictor variable on a response.
+Direct effects are from the variables that have a path pointing directly to a response variable, so effect on the response is the value of the path coefficient.
+Variables with indirect effects are those that are at least once removed from another response variable, that is the response variable of a predictor with an indirect effect is the predictor variable along another path in the model.
+Indirect effects are calculated by multiplying the coefficients along the path to the response variable of interest.
+The total effect of variables that have a direct and indirect effect(s) on a response is the sum of the direct and indirect effects.
+
+The standardized coefficients of the predictors that have the same response variable (i.e.
+direct effects) can be interpreted relative to each other in terms of influence on the response variable.
+Indirect effects
 ```{r get path coefficients, echo=F}
@@
 The topographic wetness index was not a statistically significant predictor of bay laurel density nor symptomatic leaf count after accounting for other predictor variables and additional variability due to unmeasured plot-level factors and repeated measures.
-The diversity index was not a statistically significant predictor of symptomatic leaf count
+The diversity index was not a statistically significant predictor of symptomatic leaf count, but was the strongest of the direct effects on oak infection.
 

### 2016-07-19 pathogen-spillover-diss-ch.Rmd
Add reporting of diversity and wet-hours results.

@@
 The topographic wetness index was not a statistically significant predictor of bay laurel density nor symptomatic leaf count after accounting for other predictor variables and additional variability due to unmeasured plot-level factors and repeated measures.
-The diversity index was not a statistically significant predictor of symptomatic leaf count, but was the strongest of the direct effects on oak infection.
+Locations with higher potential soil moisture (larger values of the wetness index) were significantly less diverse and had significanly fewer hours when the temperature was above 25 C during the dry-season.
+Based on the path model structure, the wetness index has only indirect effects on oak infection (pathogen spillover).
+
+* Total raw effect of the wetness index on pathogen spillover (significant paths only): `r (-.37 * -1.71) + (-124.21 * -.002 * .21)`
+
+* Total standardized effect of the wetness index on pathogen spillover (significant paths only): `r (-.24 * -.62) + (-21 * -.12 * .49)`
+
+The inference for the total effect of the wetness index on pathogen spillover was positive, indicating that oak trees located in areas with higher potential soil moisture are more likely to become infected.
+
+The diversity index was not a statistically significant predictor of symptomatic leaf count, but it was the strongest of the direct effects on oak infection.
+The negative effect of diversity on oak infection is evidence for a dilution effect.
+With the other hypothesized relationship of diversity in this system non-significant I only interpret the direct effect.
+
+The average number of hours between 14 C and 22 C on wet days during the current and sample year had a positive, but statistically weak direct effect (p = 0.068) on symptomatic leaf count.
+The direct effect of this variable on oak infection was positive and statistically significant, resulting in an overall positive effect.
 

### 2016-07-19 pathogen-spillover-diss-ch.Rmd
Add report of climate variable results.

@@
 The diversity index was not a statistically significant predictor of symptomatic leaf count, but it was the strongest of the direct effects on oak infection.
-The negative effect of diversity on oak infection is evidence for a dilution effect.
+The negative effect of diversity on oak infection is evidence for a dilution effect, where oak trees in plots with greater plant diversity are less likely to become infected.
 With the other hypothesized relationship of diversity in this system non-significant I only interpret the direct effect.
@@
 The direct effect of this variable on oak infection was positive and statistically significant, resulting in an overall positive effect.
+This indicates that oak trees in locations that have warm temperatures when it is wet are more likely to become infected.
+
+The number of hours greater than 25 C during the dry-season prior to the sample year was hypothesized to have a direct effect on bay laurel symptomatic leaf count (potential inoculum), and through this an indirect effect on pathogen spillover leading to oak infection.
+The direct effect on potential inoculum was negative, indicating that bay laurel trees in locations experiencing more time at desiccating temperatures during the dry-season had fewer infected leaves the during the following sample season.
+The total effect of the dry-season hours on pathogen spillover (oak infection) is negative, so oak trees in locations with more time at desiccating temperatures are less likely to become infected.
 

### 2016-07-19 pathogen-spillover-diss-ch.Rmd
Report results of bay laurel density and symptomatic leaf count.

@@
      ylab = "model residuals",
-     main = "Oak infectin model residuals vs. wet-hours predictor")
+     main = "Oak infection model residuals vs. wet-hours predictor")
 plot(oak_infection@frame$shannons2005.scl, resid(oak_infection),
@@
 I chose to maintain the initial hypothesized structure of the path model since the overall fit of the model was still acceptable, though this relationship could also be included as a correlated error.
-```{r fit and compare all oak plots, echo=F, }
+```{r fit and compare all oak plots, echo=F, message=F}
 raw_variables_model <- sem.fit(model_list1, oakplots.sub, 
@@
 The total effect of variables that have a direct and indirect effect(s) on a response is the sum of the direct and indirect effects.
-
 The standardized coefficients of the predictors that have the same response variable (i.e.
 direct effects) can be interpreted relative to each other in terms of influence on the response variable.
-Indirect effects
 ```{r get path coefficients, echo=F}
@@
 ```
+
 There were five non-significant path coefficients at the p=0.05 level.
@@
 The direct effect on potential inoculum was negative, indicating that bay laurel trees in locations experiencing more time at desiccating temperatures during the dry-season had fewer infected leaves the during the following sample season.
-The total effect of the dry-season hours on pathogen spillover (oak infection) is negative, so oak trees in locations with more time at desiccating temperatures are less likely to become infected.
+The total effect of the dry-season hours on pathogen spillover (oak infection) is negative, so oak trees in locations that experienced more time at desiccating temperatures are less likely to become infected.
+
+Expectedly, bay laurel density and potential inoculum load are tightly related through the direct effect of density on symptomatic leaf count.
+I structured the path model hypothesizing that bay laurel density indirectly influences pathogen spillover through its direct effect on inoculum load.
+The total effect of bay laurel density effect on pathogen spillover is positive through its direct influences on inoculum load and average number of wet-hours.
+
+The symptomatic leaf count represents potential inoculum load, and ecologically is the primary driver that I measured of the pathogen spillover process leading to oak infection.
+Oaks in locations with greater inoculum load are more likely to be infected.
+Additional factors have direct and indirect effects on the success of the pathogen spillover process as indicated by infection of oak trees.
 

### 2016-07-29 pathogen-spillover-diss-ch.Rmd
Add pairs plots of predictor variabls from multivariate models.
I also added some figures summarizing temperature and leaf symptom variables for each season.

@@
 date: "July 18, 2016"
-output: html_document
+output:
+  html_document:
+    self_contained: no
+  pdf_document: default
+  word_document: default
 ---
@@
 ### Path model 
-Path model with the continuous variables standardized using the default settings of the `scale` function, resulting in each having a mean of zero and standard deviation of one.
+This is the list of models that make up the path model to be fit using the methods implemented in the 'piecewiseSEM' packag.
+The continuous variables in the models were standardized using the default settings of the 'scale' function in R, resulting in each having a mean of zero and standard deviation of one.
 Each variable is first centered by subtracting its mean and then scaled by dividing by its standard deviation.
+
 ```{r hours >25 path model all oak plots unscaled, echo=F}
@@
 
+In my assessment of model residuals I found no substantial violations of the assumption of homoscedacity for individual models, and no relationships between the model residuals and the predictor variable(s).
+The patterning in the residual plots for some of the models is likely in part due to observations of count variables where the value is zero.
+Since the dataset in this analysis is all plots with one or more susceptible oak species there are some plots that have zero bay laurel and zero symptomatic bay laurel leaves.
+
 ### Examining the residual plots of each piece of the path model
+
 **Residual plots of Shannon's Diversity Index model**
@@
 ```
+```{r pairs plot of slc predictor variables}
+pairs(select(oakplots.sub, umca_density_ha.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl, twi15m.log.scl),
+      lower.panel = panel.cor, diag.panel = panel.hist,
+      main = "Predictor variables for symptomatic leaf count models")
+```
 
@@
 ```
+```{r pairs plot of oak infection predictor variables, echo=F}
+pairs(select(oakplots.sub, tot_lfct.log.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl),
+      lower.panel = panel.cor, diag.panel = panel.hist,
+      main = "Predictor variables for oak infection model")
+```
 
 ### Fitting the path model
-Fitting the path model by implementing the d-sep test using standardized and unstandardized variables with the assumption of correlated errors between the two climate variables.
-All continuous variables were standardized to a mean of zero and standard deviation of one.
-The results for the conditional claims of independence for the standardized and unstandardized variables are reported in the tables.
-Statistically significant p-values indicate possible missing paths, however, the overall model may still be accepted if the p-value of the C-statistic ("fisher.c") is >0.05.
-In this case, the significant missing path between diversity and bay laurel density makes ecological sense, but neither variable can be determined as causal to the other.
+
+I fit the path model by implementing the d-sep test using the 'sem.fit' function separately for standardized and unstandardized variables, with the assumption of correlated errors between the two climate variables.
+In the model with standardized variables all continuous variables were scaled to a mean of zero and standard deviation of one.
+The results for the conditional claims of independence (the d-sep test) for the standardized and unstandardized variables are reported in the tables.
+Statistically significant p-values indicate possible missing paths, however, the overall model may still be accepted if the p-value of Fisher's C-statistic is >0.05.
+The results of the d-sep test assessing overall model fit is reported with Fisher's C-statistic (fisher.c), degrees of freedom (df), the p-value of Fisher's C based on comparison to a Chi-squared distribution, the general Akaike information criterion (AIC) and corrected for sample size (AICc), the number of parameters (K) estimated, the sample size (n).
+The sample size and number of parameters are used in calculating AIC.
+
+In this case, the significant missing path between diversity and bay laurel density may be explained ecologically, but neither variable can be determined as causal to the other.
 I chose to maintain the initial hypothesized structure of the path model since the overall fit of the model was still acceptable, though this relationship could also be included as a correlated error.
+
 ```{r fit and compare all oak plots, echo=F, message=F}
@@
 
-The overall model fit is reported to the right of of independence claims in these tables where 'fisher.c' and additional model results are reported.
 ```{r tables of d-sep test results from sem.fit, echo=F}
 kable(raw_variables_model, 
-      caption = "Values for the d-sep test using raw variables.")
+      caption = "Results of the d-sep test using raw variables.")
 kable(scaled_variables_model,
-      caption = "Values for the d-sep test using standardized variables.")
+      caption = "Results of the d-sep test using standardized variables.")
 ```
@@
 ### Path coefficients
-Tables of the path coeffcients for the fitted path models with raw and standardized variables.
 The path model provides direct and indirect effects of predictor variables on multiple responses, which can be combined to calculate the total effect of a predictor variable on a response.
@@
 direct effects) can be interpreted relative to each other in terms of influence on the response variable.
+
 ```{r get path coefficients, echo=F}
@@
 ```
+
+**Tables of the path coeffcients for the fitted path models with raw and standardized variables.**
+
 ```{r tables of path coefficients, echo=F}
@@
 Based on the path model structure, the wetness index has only indirect effects on oak infection (pathogen spillover).
-
-* Total raw effect of the wetness index on pathogen spillover (significant paths only): `r (-.37 * -1.71) + (-124.21 * -.002 * .21)`
-
-* Total standardized effect of the wetness index on pathogen spillover (significant paths only): `r (-.24 * -.62) + (-21 * -.12 * .49)`
-
 The inference for the total effect of the wetness index on pathogen spillover was positive, indicating that oak trees located in areas with higher potential soil moisture are more likely to become infected.
@@
 
+## Temperature Figures
+```{r temperature figures}
+library(ggplot2)
+library(scales)
+temperatures <- left_join(
+      select(summer_temps, plotid, sample_year, hrs_abv25ds),
+      select(wet_hrs, plotid, sample_year, hrs1422_wet),
+      by = c("plotid","sample_year"))
+oak_sod$year <- as.Date(oak_sod$sample_year)
+
+temps_figure <- ggplot(oak_sod, aes(group = sample_year))
+
+sstemps_figure <- temps_figure + 
+      geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds)) +
+      theme_bw() +
+      xlab("Sample Year") +
+      ylab("Number of hours") +
+      ggtitle("Dry-season Hours >25 C")
+
+slc_avg_figure <- temps_figure + 
+      geom_boxplot(aes(as.factor(sample_year), avg_lfct)) +
+      theme_bw() +
+      xlab("Sample Year") +
+      ylab("Symptomatic leaf count") +
+      ggtitle("California bay laurel leaf symptoms - plot averages")
+
+slc_total_figure <- temps_figure + 
+      geom_boxplot(aes(as.factor(sample_year), tot_lfct)) +
+      theme_bw() +
+      xlab("Sample Year") +
+      ylab("Symptomatic leaf count") +
+      ggtitle("California bay laurel leaf symptoms - plot totals")
+
+rstemps_figure <- temps_figure + 
+      geom_boxplot(aes(as.factor(sample_year), hrs1422_wet)) +
+      theme_bw() +
+      xlab("Sample Year") +
+      ylab("Number of hours") +
+      ggtitle("Wet-ours 14-22 C")
+
+```
+
+## Rainfall Figures
+```{r rainfall figures}
+summary(rainy_season_ppt)
+```
+
 ```{r print session info}

### 2016-07-29 pathogen-spillover-diss-ch.Rmd
Add figure of oak infection.

@@
       ylab("Number of hours") +
-      ggtitle("Wet-ours 14-22 C")
+      ggtitle("Wet-hours 14-22 C")
 
+oak_infection_figure <- temps_figure +
+      geom_bar(aes(as.factor(sample_year), inf_oak_ct), stat = "identity") +
+      # geom_boxplot(aes(as.factor(sample_year), inf_oak_ct)) +
+      theme_bw() +
+      xlab("Sample Year") +
+      ylab("Number of infected oaks") +
+      ggtitle("Incident oak infection")
 ```

### 2016-08-15 pathogen-spillover-diss-ch.Rmd
Revise output.

@@
 ```{r tables of path coefficients, echo=F}
+library(knitr)
 kable(raw_coefficients,
@@
 ## Temperature Figures
-```{r temperature figures}
+```{r temperature figures, echo=F}
 library(ggplot2)
@@
       by = c("plotid","sample_year"))
-oak_sod$year <- as.Date(oak_sod$sample_year)
+# oak_sod$year <- as.Date(oak_sod$sample_year)
 
@@
 ```{r rainfall figures}
-summary(rainy_season_ppt)
+# summary(rainy_season_ppt)
 ```

### 2016-08-15 path-temp-compare.Rmd
Revise models and examine for spatial autocorrelation.

@@
 ```{r load data, echo=F}
-load("climate-path-model-data.RData")
+load("climate_path_model_data.RData")
 library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
@@
 oakplots.sub$twi15m.log.scl <- as.numeric(scale(oakplots.sub$twi15m.log))
+
+plot_coords <- select(plots_shp@data, PLOT_ID, X, Y)
+plot_coords$PLOT_ID <- tolower(plot_coords$PLOT_ID)
+oakplots.sub <- left_join(oakplots.sub, plot_coords, by = c("plotid"="PLOT_ID"))
+oakplots.2006 <- left_join(oakplots.2006, plot_coords, by = c("plotid"="PLOT_ID"))
+
+oakplots.2006$twi15m.log.scl <- as.numeric(scale(oakplots.2006$twi15m.log))
+oakplots.2006$shannons2005.scl <- as.numeric(scale(oakplots.2006$H.2005))
+
+## convert grouping variables to factors
+oakplots.sub$plotid <- as.factor(oakplots.sub$plotid)
+oakplots.sub$sample_year <- as.factor(oakplots.sub$sample_year)
 ```
@@
 model_list1.scl <- list(
-      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),
+      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.2006),
 
@@
 model_list1 <- list(
-      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
+      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.2006),
 
-      umca_density <- lmer(tot_bay.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
-      slc <- lmer(tot_lfct.log ~ tot_bay.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      slc <- lmer(tot_lfct.log ~ umca_density_ha.log.scl + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
-      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ tot_bay.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ umca_density_ha.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
@@
 
-Examining the residual plots of each piece of the path model
-```{r shannons residuals plot}
+Examining the residual plots of each piece of the path model and checking for spatial autocorrelation in the residuals.
+```{r general code for testing SAC, echo=F, eval=F}
+### Check for spatial autocorrelation in residuals
+
+#install.packages('ape')
+library(ape)
+
+### Create inverse distance matrix
+points <- as.matrix(dist(cbind(data$x, data$y)))
+points.inv <- 1/points
+diag(points.inv) <- 0
+# points.inv [is.infinite(points.inv)]<-0 
+# WAS GETTING ERROR - if (obs <= ei) 2 * pv else 2 * (1 - pv) : 
+# missing value where TRUE/FALSE needed, this fixed the infinite values in the data
+
+### Calculate Moran's I
+Moran.I(model.residuals, points.inv) 
+### Additional description calculating Moran's I in R: http://www.ats.ucla.edu/stat/r/faq/morans_i.htm
+### Where model.residuals = residuals(model)
+```
+
+```{r shannons residuals plot, echo=F}
 par(mfrow=c(2,2))
@@
 
-```{r bay laurel density residuals plot}
+```{r test SAC diversity model residuals, echo=F}
+library(spBayes); library(classInt); library(RColorBrewer); library(geoR)
+summary(na.omit(oakplots.2006))
+points <- select(na.omit(oakplots.2006), X, Y)
+points <- as.matrix(dist(cbind(points$X, points$Y)))
+points.inv <- 1/points
+diag(points.inv) <- 0
+shannons.resids <- resid(shannons)
+max.dist <- 0.33 * max(iDist(points))
+bins <- 100
+
+vario.shannons <- variog(coords = points, 
+                     data = na.omit(oakplots.2006$shannons2005.scl),
+                     uvec = (seq(0, max.dist, length = bins)))
+fit.vario.shannons <- variofit(
+      vario.shannons,
+      ini.cov.pars = c(300, 200/-log(0.5)),
+      cov.model = "exponential",
+      minimisation.function = "optim",
+      weights = "equal"
+)
+par(mfrow = c(1,2))
+plot(vario.shannons)
+lines(fit.vario.shannons)
+abline(h = fit.vario.shannons$nugget, col = "blue")# nugget
+abline(h = fit.vario.shannons$cov.pars[1] + fit.vario.shannons$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.shannons$cov.pars[2], col = "red3")# range
+
+vario.shannons.resid <- variog(coords = points,
+                               data = shannons.resids,
+                               uvec = (seq(0, max.dist, length = bins)))
+fit.vario.shannons.resid <- variofit(
+      vario.shannons.resid,
+      cov.model = "exponential",
+      minimisation.function = "optim",
+      weights = "equal"
+)
+par(mfrow = c(1,2))
+plot(vario.shannons.resid)
+lines(fit.vario.shannons.resid)
+abline(h = fit.vario.shannons.resid$nugget, col = "blue")# nugget
+abline(h = fit.vario.shannons.resid$cov.pars[1] + fit.vario.shannons.resid$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.shannons.resid$cov.pars[2], col = "red3")# range
+
+Moran.I(shannons.resids, points.inv)
+# Significant p-value, but no pattern in the model residuals 
+```
+
+```{r bay laurel density residuals plot, echo=F}
 plot(umca_density, main = "LMM of bay laurel density residuals")
@@
 
-```{r symptomatic leaf count residuals plot}
+```{r test SAC bay density model, echo=F}
+# Test SAC of residuals
+umca_density.resids <- resid(umca_density)
+plots1 <- select(umca_density@frame, plotid)
+plots1 <- as.vector(plots1$plotid)
+plots1 <- as.data.frame(plots1)
+names(plots1) <- "plotid"
+# plots1$plotid <- as.character(plots1$plotid)
+str(plots1)
+
+plots2 <- select(plots_shp@data, PLOT_ID, X, Y)
+# plots2$plotid <- as.character(plots2$plotid)
+names(plots2) <- tolower(names(plots2))
+plots2$plot_id <- tolower(plots2$plot_id)
+str(plots2)
+
+# all.equal(plots1$plotid,plots2$plotid)
+points_rep <- left_join(plots1, plots2, by = c("plotid"="plot_id"))
+summary(points_rep)
+points_rep <- as.matrix(dist(cbind(points_rep$x, points_rep$y)))
+points_rep.inv <- 1/points_rep
+points_rep.inv[points_rep.inv == "Inf"] <- 0
+summary(points_rep.inv)
+
+# Plot variograms
+vario.umca_density <- variog(coords = points_rep, 
+                     data = na.omit(umca_density@frame$tot_bay.log),
+                     uvec = (seq(0, max.dist, length = bins)))
+fit.vario.umca_density <- variofit(
+      vario.umca_density,
+      # ini.cov.pars = c(300, 200/-log(0.5)),
+      cov.model = "exponential",
+      minimisation.function = "optim",
+      weights = "equal"
+)
+par(mfrow = c(1,2))
+plot(vario.umca_density)
+lines(fit.vario.umca_density)
+abline(h = fit.vario.umca_density$nugget, col = "blue")# nugget
+abline(h = fit.vario.umca_density$cov.pars[1] + fit.vario.umca_density$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.umca_density$cov.pars[2], col = "red3")# range
+
+vario.umca_density.resids <- variog(coords = points_rep,
+                               data = umca_density.resids,
+                               uvec = (seq(0, max.dist, length = bins)))
+fit.vario.umca_density.resids <- variofit(
+      vario.umca_density.resids,
+      cov.model = "exponential",
+      minimisation.function = "optim",
+      weights = "equal"
+)
+# par(mfrow = c(1,2))
+plot(vario.umca_density.resids)
+lines(fit.vario.umca_density.resids)
+abline(h = fit.vario.umca_density.resids$nugget, col = "blue")# nugget
+abline(h = fit.vario.umca_density.resids$cov.pars[1] + fit.vario.umca_density.resids$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.umca_density.resids$cov.pars[2], col = "red3")# range
+
+Moran.I(umca_density.resids, points_rep.inv)
+
+# length(umca_density.resids)
+# dim(points_rep)
+```
+
+```{r symptomatic leaf count residuals plot, echo=F}
 plot(slc, main = "LMM of symptomatic leaf count residuals")
@@
 
-```{r wet-hours residuals plot}
+```{r test SAC of slc model residuals, echo=F}
+# Test SAC of residuals
+slc.resids <- resid(slc)
+plots1 <- select(slc@frame, plotid)
+plots1 <- as.vector(plots1$plotid)
+plots1 <- as.data.frame(plots1)
+names(plots1) <- "plotid"
+# plots1$plotid <- as.character(plots1$plotid)
+str(plots1)
+
+plots2 <- select(plots_shp@data, PLOT_ID, X, Y)
+# plots2$plotid <- as.character(plots2$plotid)
+names(plots2) <- tolower(names(plots2))
+plots2$plot_id <- tolower(plots2$plot_id)
+str(plots2)
+
+# all.equal(plots1$plotid,plots2$plotid)
+points_rep <- left_join(plots1, plots2, by = c("plotid"="plot_id"))
+summary(points_rep)
+points_rep <- as.matrix(dist(cbind(points_rep$x, points_rep$y)))
+points_rep.inv <- 1/points_rep
+points_rep.inv[points_rep.inv == "Inf"] <- 0
+summary(points_rep.inv)
+
+# Plot variograms
+vario.slc <- variog(coords = points_rep, 
+                     data = na.omit(slc@frame$tot_lfct.log),
+                     uvec = (seq(0, max.dist, length = bins)))
+fit.vario.slc <- variofit(
+      vario.slc,
+      # ini.cov.pars = c(300, 200/-log(0.5)),
+      cov.model = "exponential",
+      minimisation.function = "optim",
+      weights = "equal"
+)
+par(mfrow = c(1,2))
+plot(vario.slc)
+lines(fit.vario.slc)
+abline(h = fit.vario.slc$nugget, col = "blue")# nugget
+abline(h = fit.vario.slc$cov.pars[1] + fit.vario.slc$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.slc$cov.pars[2], col = "red3")# range
+
+vario.slc.resids <- variog(coords = points_rep,
+                               data = slc.resids,
+                               uvec = (seq(0, max.dist, length = bins)))
+fit.vario.slc.resids <- variofit(
+      vario.slc.resids,
+      cov.model = "exponential",
+      minimisation.function = "optim",
+      weights = "equal"
+)
+# par(mfrow = c(1,2))
+plot(vario.slc.resids)
+lines(fit.vario.slc.resids)
+abline(h = fit.vario.slc.resids$nugget, col = "blue")# nugget
+abline(h = fit.vario.slc.resids$cov.pars[1] + fit.vario.slc.resids$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.slc.resids$cov.pars[2], col = "red3")# range
+
+Moran.I(slc.resids, points_rep.inv)
+```
+
+```{r wet-hours residuals plot, echo=F}
 plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
@@
 
-```{r dry-season hours residuals plot}
+```{r test SAC wet-hours residuals, echo=F}
+# Test SAC of residuals
+wet_hours.resids <- resid(wet_hours_1422)
+plots1 <- select(wet_hours_1422@frame, plotid)
+plots1 <- as.vector(plots1$plotid)
+plots1 <- as.data.frame(plots1)
+names(plots1) <- "plotid"
+# plots1$plotid <- as.character(plots1$plotid)
+str(plots1)
+
+plots2 <- select(plots_shp@data, PLOT_ID, X, Y)
+# plots2$plotid <- as.character(plots2$plotid)
+names(plots2) <- tolower(names(plots2))
+plots2$plot_id <- tolower(plots2$plot_id)
+str(plots2)
+
+# all.equal(plots1$plotid,plots2$plotid)
+points_rep <- left_join(plots1, plots2, by = c("plotid"="plot_id"))
+summary(points_rep)
+points_rep <- as.matrix(dist(cbind(points_rep$x, points_rep$y)))
+points_rep.inv <- 1/points_rep
+points_rep.inv[points_rep.inv == "Inf"] <- 0
+summary(points_rep.inv)
+
+# Plot variograms
+vario. <- variog(coords = points_rep, 
+                     data = na.omit(umca_density@frame$tot_bay.log),
+                     uvec = (seq(0, max.dist, length = bins)))
+fit.vario.umca_density <- variofit(
+      vario.umca_density,
+      # ini.cov.pars = c(300, 200/-log(0.5)),
+      cov.model = "exponential",
+      minimisation.function = "optim",
+      weights = "equal"
+)
+par(mfrow = c(1,2))
+plot(vario.umca_density)
+lines(fit.vario.umca_density)
+abline(h = fit.vario.umca_density$nugget, col = "blue")# nugget
+abline(h = fit.vario.umca_density$cov.pars[1] + fit.vario.umca_density$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.umca_density$cov.pars[2], col = "red3")# range
+
+vario.umca_density.resids <- variog(coords = points_rep,
+                               data = umca_density.resids,
+                               uvec = (seq(0, max.dist, length = bins)))
+fit.vario.umca_density.resids <- variofit(
+      vario.umca_density.resids,
+      cov.model = "exponential",
+      minimisation.function = "optim",
+      weights = "equal"
+)
+# par(mfrow = c(1,2))
+plot(vario.umca_density.resids, 
+     main = "Variogram - UMCA deensity model residuals",
+     xlab = "distance (m)")
+lines(fit.vario.umca_density.resids)
+abline(h = fit.vario.umca_density.resids$nugget, col = "blue")# nugget
+abline(h = fit.vario.umca_density.resids$cov.pars[1] + fit.vario.umca_density.resids$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.umca_density.resids$cov.pars[2], col = "red3")# range
+
+Moran.I(wet_hours.resids, points_rep.inv)
+```
+
+```{r dry-season hours residuals plot, echo=F}
 plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
@@
 
-```{r oak infection residuals plot}
+```{r test SAC dry-season model residuals, echo=F}
+# Test SAC of residuals
+hrs_abv25.resids <- resid(hrs_abv25)
+plots1 <- select(hrs_abv25@frame, plotid)
+plots1 <- as.vector(plots1$plotid)
+plots1 <- as.data.frame(plots1)
+names(plots1) <- "plotid"
+# plots1$plotid <- as.character(plots1$plotid)
+str(plots1)
+
+plots2 <- select(plots_shp@data, PLOT_ID, X, Y)
+# plots2$plotid <- as.character(plots2$plotid)
+names(plots2) <- tolower(names(plots2))
+plots2$plot_id <- tolower(plots2$plot_id)
+str(plots2)
+
+# all.equal(plots1$plotid,plots2$plotid)
+points_rep <- left_join(plots1, plots2, by = c("plotid"="plot_id"))
+summary(points_rep)
+points_rep <- as.matrix(dist(cbind(points_rep$x, points_rep$y)))
+points_rep.inv <- 1/points_rep
+points_rep.inv[points_rep.inv == "Inf"] <- 0
+summary(points_rep.inv)
+
+Moran.I(hrs_abv25.resids, points_rep.inv)
+```
+
+```{r oak infection residuals plot, echo=F}
 plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
@@
      main = "Oak infection model residuals vs. diversity predictor")
+```
+
+```{r test SAC of oak infection residuals, echo=F}
+# Test SAC of residuals
+oak_infection.resids <- resid(oak_infection)
+plots1 <- select(oak_infection@frame, plotid)
+plots1 <- as.vector(plots1$plotid)
+plots1 <- as.data.frame(plots1)
+names(plots1) <- "plotid"
+# plots1$plotid <- as.character(plots1$plotid)
+str(plots1)
+
+plots2 <- select(plots_shp@data, PLOT_ID, X, Y)
+# plots2$plotid <- as.character(plots2$plotid)
+names(plots2) <- tolower(names(plots2))
+plots2$plot_id <- tolower(plots2$plot_id)
+str(plots2)
+
+# all.equal(plots1$plotid,plots2$plotid)
+points_rep <- left_join(plots1, plots2, by = c("plotid"="plot_id"))
+summary(points_rep)
+points_rep <- as.matrix(dist(cbind(points_rep$x, points_rep$y)))
+points_rep.inv <- 1/points_rep
+points_rep.inv[points_rep.inv == "Inf"] <- 0
+summary(points_rep.inv)
+
+Moran.I(oak_infection.resids, points_rep.inv)
 ```
@@
 # Comparing fits of the piecewise models using all oak plots vs. plots with oaks and at least one bay laurel
-sem.fit(model_list1, oakplots.sub, 
-        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
-sem.fit(model_list1.scl, oakplots.sub,
-          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+(sem.fit(model_list1, oakplots.sub, 
+        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log")))
+(sem.fit(model_list1.scl, oakplots.sub,
+          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl")))
 

### 2016-08-24 microclimate-correlations.Rmd
Report microclimate correlations.

---
title: "Microclimate Variable Correlations"
author: "Whalen Dillon"
date: "August 24, 2016"
output: html_document
---

```{r load data, echo=F, message=F}
library(knitr)
load("climate_path_model_data.RData")
library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
library(plyr); library(dplyr, verbose = F)
library(lme4, verbose = F)
library(lmerTest, verbose = F)

# Calculate oak density variable as stems per hectare 
oakplots.sub$umca_density_ha <- (oakplots.sub$tot_bay / (15*15))*10000
oakplots.sub$umca_density_ha.log <- log(oakplots.sub$umca_density_ha+1)
```

```{r rescale variables, echo=F}
## rescale variables
# oakplots.sub2$inf_bay.scl <- scale((oakplots.sub2$infected_bay_ct))
oakplots.sub$tot_bay.scl <- as.numeric(scale(oakplots.sub$tot_bay))
oakplots.sub$tot_bay.log.scl <- as.numeric(scale(oakplots.sub$tot_bay.log))
oakplots.sub$umca_density_ha.scl <- as.numeric(scale(oakplots.sub$umca_density_ha))
oakplots.sub$umca_density_ha.log.scl <- as.numeric(scale(oakplots.sub$umca_density_ha.log))
oakplots.sub$umca_basal_area.scl <- as.numeric(scale(oakplots.sub$umca_basal_area))
oakplots.sub$umca_ba.log.scl <- as.numeric(scale(oakplots.sub$umca_ba.log))
oakplots.sub$tot_lfct.scl <- as.numeric(scale(oakplots.sub$tot_lfct))
oakplots.sub$tot_lfct.log.scl <- as.numeric(scale(oakplots.sub$tot_lfct.log))
# oakplots.sub$avg_lfct.scl <- scale(oakplots.sub$avg_lfct)
oakplots.sub$hrs_abv25ds.scl <- as.numeric(scale(oakplots.sub$hrs_abv25ds))
oakplots.sub$dys_abv25ds.scl <- as.numeric(scale(oakplots.sub$dys_abv25ds))
oakplots.sub$avg_tmax_ds.scl <- as.numeric(scale(oakplots.sub$avg_tmax_ds))
oakplots.sub$avg_hrsblw14_wet_tminus1.scl <- as.numeric(scale(oakplots.sub$avg_hrsblw14_wet_tminus1))
oakplots.sub$hrs1422_wet_tminus1.log.scl <- as.numeric(scale(oakplots.sub$hrs1422_wet_tminus1.log))
# oakplots.sub$avg_hrs1422_wet_t_t1.scl <- scale(oakplots.sub$avg_hrs1422_wet_t_t1)
oakplots.sub$avg_hrs1422_wet_t_t1.log.scl <- as.numeric(scale(oakplots.sub$avg_hrs1422_wet_t_t1.log))

## static variables
oakplots.sub$candens15m.scl <- as.numeric(scale(oakplots.sub$candens15m))
oakplots.sub$shannons2005.scl <- as.numeric(scale(oakplots.sub$H.2005))
oakplots.sub$shannons2014.scl <- as.numeric(scale(oakplots.sub$H.2014))
oakplots.sub$twi15m.scl <- as.numeric(scale(oakplots.sub$twi15m))
oakplots.sub$twi15m.log.scl <- as.numeric(scale(oakplots.sub$twi15m.log))
```

```{r microclimate correlations, echo=F}
library(ggplot2)
panel.cor <- function(x, y, digits = 2, cex.cor, ...){
      usr <- par("usr"); on.exit(par(usr))
      par(usr = c(0, 1, 0, 1))
      # correlation coefficient
      r <- cor(x, y)
      txt <- format(c(r, 0.123456789), digits = digits)[1]
      txt <- paste("r= ", txt, sep = "")
      text(0.5, 0.5, txt, cex = 1.2)

      # p-value calculation
      # p <- cor.test(x, y)$p.value
      # txt2 <- format(c(p, 0.123456789), digits = digits)[1]
      # txt2 <- paste("p= ", txt2, sep = "")
      # if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
      # text(0.5, 0.4, txt2)
}

microclimate_variables <- oakplots.sub %>% 
      select(sample_year, plotid, tot_lfct.log, avg_lfct.log, contains("hrs"), contains("tmax"), contains("tmin"))
summary(microclimate_variables)

# umca_plots <- left_join(umca_plots, select(rs_temps_lag, -avg_tmax_rs), by = c("plotid","sample_year"))
# umca_plots <- left_join(umca_plots, wet_hrs_lag, by = c("plotid","sample_year"))

pairs(select(filter(microclimate_variables, sample_year == 2005), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2005 Sample Year \n n=165 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2006), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2006 Sample Year \n n=167 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2007), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2007 Sample Year \n n=164 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2008), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2008 Sample Year \n n=163 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2009), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2009 Sample Year \n n=162 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2010), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2010 Sample Year \n n=159 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2011), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2011 Sample Year \n n=156 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2012), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2012 Sample Year \n n=156 oak plots")
pairs(select(filter(microclimate_variables, sample_year == 2014), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmin_rs, avg_tmax_ds, avg_tmin_ds), lower.panel = panel.cor,
      main = "Correlations Between Variables for 2014 Sample Year \n n=156 oak plots")
pairs(na.omit(select(filter(microclimate_variables, sample_year != 2004), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds, avgtmax_wet)), lower.panel = panel.cor,
      main = "Correlations Between Variables \n all oak plots 2005-2014")

# pairs(select(filter(umca_plots, sample_year == 2005), tot_lfct, avg_lfct, avg_hrs1422_wet_t_t1, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
#       main = "Correlations Between Variables for 2005 Sample Year \n n=179 UMCA plots")

reg_eqn <- function(df,y,x) {
      y = df$avg_hrs1422_wet_t_t1.log
      x = df$avgtmax_wet
      # y = y
      # x = x
      m <- lm(y ~ x, data = df)

      a <- coef(m)[1]
      b <- coef(m)[2]
      r2 <- summary(m)$r.squared

      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                         list(a = format(a, digits = 2),
                              b = format(b, digits = 2),
                              r2 = format(r2, digits = 3)))
      c(lab = as.character(as.expression(expr)))
}
# y <- "avg_hrs1422_wet_t_t1.log"
# x <- "avgtmax_wet"
# reg_eqn(df = na.omit(microclimate_variables), y = y, x = x)

# eqns <- ddply(na.omit(microclimate_variables), "sample_year", reg_eqn)

ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
             group_by(sample_year), 
       aes(avgtmax_wet, avg_hrs1422_wet_t_t1.log)) +
      geom_point() +
      facet_grid(sample_year ~ .) +
      geom_smooth(method = "lm", size = 0.5) +
      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T, 
                hjust = 0, vjust = -1.5) + 
      facet_grid(sample_year ~ .)

ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
             group_by(sample_year),
       aes(y = tot_lfct)) +
      geom_point(aes(x = avgtmax_wet)) +
      # geom_point(aes(x = avgtmin_wet)) +
      facet_grid(sample_year ~ .)

ggplot(filter(microclimate_variables, sample_year != 2004), 
       aes(avgtmin_wet, avg_hrs1422_wet_t_t1.log)) +
      geom_point() +
      facet_grid(sample_year ~ .)

ggplot(filter(microclimate_variables, sample_year != 2004), 
       aes(avg_tmax_rs, avg_hrs1422_wet_t_t1.log)) +
      geom_point() +
      facet_grid(sample_year ~ .)

ggplot(filter(microclimate_variables, sample_year != 2004), 
       aes(avg_tmax_ds, avg_hrs1422_wet_t_t1.log)) +
      geom_point() +
      facet_grid(sample_year ~ .)

```

### 2016-08-24 microclimate-correlations.Rmd
Add bivariate linear models/plots.

@@
 date: "August 24, 2016"
-output: html_document
+output: 
+  html_document: 
+    theme: journal
+    toc: yes
 ---
 
+```{r include=F}
+knitr::opts_chunk$set(warning = F)
+```
+
 ```{r load data, echo=F, message=F}
+load("climate_path_model_data.RData")
+library(plyr); library(dplyr, verbose = F); library(ggplot2, verbose = F)
 library(knitr)
-load("climate_path_model_data.RData")
-library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
-library(plyr); library(dplyr, verbose = F)
-library(lme4, verbose = F)
-library(lmerTest, verbose = F)
+panel.cor <- function(x, y, digits = 2, cex.cor, ...){
+      usr <- par("usr"); on.exit(par(usr))
+      par(usr = c(0, 1, 0, 1))
+      # correlation coefficient
+      r <- cor(x, y)
+      txt <- format(c(r, 0.123456789), digits = digits)[1]
+      txt <- paste("r= ", txt, sep = "")
+      text(0.5, 0.5, txt, cex = 1.2)
+
+      # p-value calculation
+      # p <- cor.test(x, y)$p.value
+      # txt2 <- format(c(p, 0.123456789), digits = digits)[1]
+      # txt2 <- paste("p= ", txt2, sep = "")
+      # if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
+      # text(0.5, 0.4, txt2)
+}
 
@@
 oakplots.sub$twi15m.log.scl <- as.numeric(scale(oakplots.sub$twi15m.log))
+
+microclimate_variables <- oakplots.sub %>% 
+      select(sample_year, plotid, tot_lfct.log, avg_lfct.log, contains("hrs"), contains("tmax"), contains("tmin"))
+# summary(microclimate_variables)
+```
+
+## Correlation Plots
+These plots show the correlations and scatter plots among a selection of the microclimate variables that I calculated, and the relationship with symptomatic leaf count for a plot (total and average).
+
+These data are for the plots that I used in the path model, which means they had to have at least one susceptible oak tree, but could (and do in some cases) have zero California bay laurel.
+
+I broke down the plots into each sample year, and then provide a plot with all the data aggregated across sample years at the end.
+The earliest sampling that enables applying a one year lag of the microclimate data is 2005, so I start with that season.
+It is also the first full sampling year after plot establishment in 2003-2004.
+
+#### Variable Name Descriptions
+ * The dry-season corresponds to the months June-September and the rainy season corresponds to the months November-May.
+
+ * Wet-days are days when the nearest rain gauge recorded precipitation.
+
+```{r var names, echo=F}
+var_names <- as.data.frame(names(microclimate_variables %>% select(tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds, avgtmax_wet)))
+var_names <- var_names %>% select(variable_name = starts_with("names"))
+var_names$description <- c("Natural log of the total leaf count", "Natural log of the average leaf count", "Averaged number of hours the temperature was 14-22 C on wet days during the current & previous year", "Number of hours the temperature was >25 C during the dry season", "Average daily maximum temperature during the rainy season", "Average daily maximum temperature during the dry season", "Average daily minimum temperature during the rainy season", "Average daily minimum temperature during the dry season", "Average daily maximum temperature during wet days")
+
+kable(var_names)
 ```
@@
 ```{r microclimate correlations, echo=F}
-library(ggplot2)
-panel.cor <- function(x, y, digits = 2, cex.cor, ...){
-      usr <- par("usr"); on.exit(par(usr))
-      par(usr = c(0, 1, 0, 1))
-      # correlation coefficient
-      r <- cor(x, y)
-      txt <- format(c(r, 0.123456789), digits = digits)[1]
-      txt <- paste("r= ", txt, sep = "")
-      text(0.5, 0.5, txt, cex = 1.2)
-
-      # p-value calculation
-      # p <- cor.test(x, y)$p.value
-      # txt2 <- format(c(p, 0.123456789), digits = digits)[1]
-      # txt2 <- paste("p= ", txt2, sep = "")
-      # if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
-      # text(0.5, 0.4, txt2)
-}
-
-microclimate_variables <- oakplots.sub %>% 
-      select(sample_year, plotid, tot_lfct.log, avg_lfct.log, contains("hrs"), contains("tmax"), contains("tmin"))
-summary(microclimate_variables)
-
-# umca_plots <- left_join(umca_plots, select(rs_temps_lag, -avg_tmax_rs), by = c("plotid","sample_year"))
-# umca_plots <- left_join(umca_plots, wet_hrs_lag, by = c("plotid","sample_year"))
-
 pairs(select(filter(microclimate_variables, sample_year == 2005), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
@@
       main = "Correlations Between Variables for 2014 Sample Year \n n=156 oak plots")
-pairs(na.omit(select(filter(microclimate_variables, sample_year != 2004), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds, avgtmax_wet)), lower.panel = panel.cor,
+pairs(na.omit(select(filter(microclimate_variables, sample_year != 2004), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds, avgtmax_wet)),
+      lower.panel = panel.cor,
       main = "Correlations Between Variables \n all oak plots 2005-2014")
@@
 #       main = "Correlations Between Variables for 2005 Sample Year \n n=179 UMCA plots")
+```
 
+## Bivariate Linear Models
+This is another way to assess the relationships between microclimate and pathogen load.
+It will take more effort to make one of these for each variable, so I did it just for two that generally had the strongest correlations with pathogen load.
+
+```{r bivariate linear model avgtmaxds, echo=F, fig.height=8, message=F}
 reg_eqn <- function(df,y,x) {
-      y = df$avg_hrs1422_wet_t_t1.log
-      x = df$avgtmax_wet
+      y = df$tot_lfct.log
+      x = df$avg_tmax_ds
       # y = y
@@
 
-# eqns <- ddply(na.omit(microclimate_variables), "sample_year", reg_eqn)
+eqns <- ddply(na.omit(microclimate_variables), "sample_year", reg_eqn)
+
+# ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
+#              group_by(sample_year), 
+#        aes(avgtmax_wet, avg_hrs1422_wet_t_t1.log)) +
+#       geom_point() +
+#       facet_grid(sample_year ~ .) +
+#       geom_smooth(method = "lm", size = 0.5) +
+#       geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T, 
+#                 hjust = 0, vjust = -1.5) + 
+#       facet_grid(sample_year ~ .)
 
@@
              group_by(sample_year),
-       aes(avgtmax_wet, avg_hrs1422_wet_t_t1.log)) +
+       aes(x = avg_tmax_ds, y = tot_lfct.log)) +
       geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = -1) +
       facet_grid(sample_year ~ .) +
-      geom_smooth(method = "lm", size = 0.5) +
-      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T, 
-                hjust = 0, vjust = -1.5) + 
-      facet_grid(sample_year ~ .)
+      # scale_x_continuous(limits = c(20,30), breaks = c(seq(15,35,5))) +
+      xlab("Average Daily Maximum Temperature - Dry Season") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("Oak Plots") +
+      theme_bw()
+
+# ggplot(filter(microclimate_variables, sample_year != 2004), 
+#        aes(avgtmin_wet, avg_hrs1422_wet_t_t1.log)) +
+#       geom_point() +
+#       facet_grid(sample_year ~ .)
+
+# ggplot(filter(microclimate_variables, sample_year != 2004), 
+#        aes(avg_tmax_rs, avg_hrs1422_wet_t_t1.log)) +
+#       geom_point() +
+#       facet_grid(sample_year ~ .)
+
+# ggplot(filter(microclimate_variables, sample_year != 2004), 
+#        aes(avg_tmax_ds, avg_hrs1422_wet_t_t1.log)) +
+#       geom_point() +
+#       facet_grid(sample_year ~ .)
+```
+
+```{r bivariate linear model hrsabv25ds, echo=F, fig.height=8, message=F}
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrs_abv25ds
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+# y <- "avg_hrs1422_wet_t_t1.log"
+# x <- "avgtmax_wet"
+# reg_eqn(df = na.omit(microclimate_variables), y = y, x = x)
+
+eqns <- ddply(na.omit(microclimate_variables), "sample_year", reg_eqn)
 
@@
              group_by(sample_year),
-       aes(y = tot_lfct)) +
-      geom_point(aes(x = avgtmax_wet)) +
-      # geom_point(aes(x = avgtmin_wet)) +
-      facet_grid(sample_year ~ .)
+       aes(x = hrs_abv25ds, y = tot_lfct.log)) +
+      geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Dry-season Hours >25 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("Oak Plots") +
+      theme_bw()
+```
 
-ggplot(filter(microclimate_variables, sample_year != 2004), 
-       aes(avgtmin_wet, avg_hrs1422_wet_t_t1.log)) +
+Neither do a very good job of explaining the variation in leaf count. I wonder if this is at least partly due to the plots with zeroes.
+
+#### Below is a set using plots that have one or more bay laurel trees. This also boosts the number of plots in the sample to ~170.
+
+```{r bivariate models using umca plots, echo=F, fig.height=8, message=F}
+umca_plots <- left_join(umca_plots, select(rs_temps_lag, -avg_tmax_rs), by = c("plotid","sample_year"))
+umca_plots <- left_join(umca_plots, wet_hrs_lag, by = c("plotid","sample_year"))
+umca_plots$tot_lfct.log <- log(umca_plots$tot_lfct+1)
+umca_plots$avg_lfct.log <- log(umca_plots$avg_lfct+1)
+
+umca_plots <- filter(umca_plots, sample_year >= 2005)
+
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrs_abv25ds
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+eqns <- ddply(umca_plots, "sample_year", reg_eqn)
+ggplot(umca_plots %>% filter(sample_year > 2004) %>% 
+             group_by(sample_year),
+       aes(x = hrs_abv25ds, y = tot_lfct.log)) +
       geom_point() +
-      facet_grid(sample_year ~ .)
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Dry-season Hours >25 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("UMCA Plots") +
+      theme_bw()
 
-ggplot(filter(microclimate_variables, sample_year != 2004), 
-       aes(avg_tmax_rs, avg_hrs1422_wet_t_t1.log)) +
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$avg_tmax_ds
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+eqns <- ddply(umca_plots, "sample_year", reg_eqn)
+ggplot(filter(umca_plots, sample_year > 2004) %>% 
+             group_by(sample_year),
+       aes(x = avg_tmax_ds, y = tot_lfct.log)) +
       geom_point() +
-      facet_grid(sample_year ~ .)
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = -1) +
+      facet_grid(sample_year ~ .) +
+      # scale_x_continuous(limits = c(20,30), breaks = c(seq(15,35,5))) +
+      xlab("Average Daily Maximum Temperature - Dry Season") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("UMCA Plots") +
+      theme_bw()
+```
 
-ggplot(filter(microclimate_variables, sample_year != 2004), 
-       aes(avg_tmax_ds, avg_hrs1422_wet_t_t1.log)) +
-      geom_point() +
-      facet_grid(sample_year ~ .)
-
-```
+These appear to do a little better in explaining the variation, but not for all years.

### 2016-08-26 sem-data-development.Rmd
Update lagged temperature variables.

@@
 rs_temps <- rename(rs_temps, 
-                   hrs_14_20_rs = hrs_14_20, hrs_abv25rs = hrs_abv25,
-                   hrs_abv20_rs = hrs_abv20, hrs_blw10rs = hrs_blw10,
-                   avg_tmax_rs = avg_tmax, avg_tmin_rs = avg_tmin)
+                   hrs_14_20_rs = hrs_14_20, hrs_10_20_rs = hrs_10_20,
+                   hrs_abv25rs = hrs_abv25, hrs_abv20_rs = hrs_abv20,
+                   hrs_blw20rs = hrs_blw20, hrs_blw10rs = hrs_blw10,
+                   hrs_blw14rs = hrs_blw14, avg_tmax_rs = avg_tmax,
+                   avg_tmin_rs = avg_tmin)
 
@@
                        hrs_abv25ds = hrs_abv25, hrs_abv20ds = hrs_abv20,
-                       hrs_blw18ds = hrs_blw18, avg_tmax_ds = avg_tmax,
+                       hrs_blw20ds = hrs_blw20, avg_tmax_ds = avg_tmax,
                        avg_tmin_ds = avg_tmin)
@@
 hrsblw14_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrsblw14_wet))
+hrs1422_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrs1422_wet))
+hrsabv22_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrsabv22_wet))
+hrsblw22_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrsblw22_wet))
+hrsblw10_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrsblw10_wet))
+hrs1020_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrs1020_wet))
+hrs1022_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrs1022_wet))
+hrsabv20_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrsabv20_wet))
+hrsblw20_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrsblw20_wet))
+
 avgtmax_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, avgtmax_wet))
-hrs1422_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, hrs1422_wet))
+avgtmin_wet_lag <- as.tbl(select(wet_hrs, plotid, sample_year, avgtmin_wet))
+
 wet_hrs_lag <- select(wet_hrs, plotid, sample_year)
@@
 
+## Number of wet-hours below 20 C lagged ####
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw20_wet_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrsblw20_wet_tminus1 = hrsblw20_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw20_wet_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrsblw20_wet_tminus2 = hrsblw20_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw20_wet_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrsblw20_wet_tminus3 = hrsblw20_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw20_wet_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## Number of wet-hours below 10 C lagged ####
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw10_wet_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrsblw10_wet_tminus1 = hrsblw10_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw10_wet_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrsblw10_wet_tminus2 = hrsblw10_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw10_wet_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrsblw10_wet_tminus3 = hrsblw10_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrsblw10_wet_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
 ## Average daily maximum temperature on wet days lagged ####
@@
 
+## Average daily minimum temperature on wet days lagged ####
+wet_hrs_lag <- left_join(wet_hrs_lag, avgtmin_wet_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(avgtmin_wet_tminus1 = avgtmin_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, avgtmin_wet_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(avgtmin_wet_tminus2 = avgtmin_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, avgtmin_wet_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(avgtmin_wet_tminus3 = avgtmin_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, avgtmin_wet_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
 ## Number of hours 14-22 C on wet days ####
@@
 wet_hrs_lag <- left_join(wet_hrs_lag, hrs1422_wet_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## Number of hours 10-20 C on wet days ####
+wet_hrs_lag <- left_join(wet_hrs_lag, hrs1020_wet_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrs1020_wet_tminus1 = hrs1020_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrs1020_wet_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrs1020_wet_tminus2 = hrs1020_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrs1020_wet_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrs1020_wet_tminus3 = hrs1020_wet)
+wet_hrs_lag <- left_join(wet_hrs_lag, hrs1020_wet_lag,
                 by = c("plotid", "sample_year"="sample_year"))
@@
 
+            avg_hrsblw10_wet_tminus1 = rowMeans(cbind(hrsblw10_wet, hrsblw10_wet_tminus1)),
+             avg_hrsblw10_wet_tminus2 = rowMeans(cbind(hrsblw10_wet, hrsblw10_wet_tminus1, hrsblw10_wet_tminus2)),
+             avg_hrsblw10_wet_tminus3 = rowMeans(cbind(hrsblw10_wet, hrsblw10_wet_tminus1, hrsblw10_wet_tminus2, hrsblw10_wet_tminus3)),
+
+            avg_hrsblw20_wet_tminus1 = rowMeans(cbind(hrsblw20_wet, hrsblw20_wet_tminus1)),
+             avg_hrsblw20_wet_tminus2 = rowMeans(cbind(hrsblw20_wet, hrsblw20_wet_tminus1, hrsblw20_wet_tminus2)),
+             avg_hrsblw20_wet_tminus3 = rowMeans(cbind(hrsblw20_wet, hrsblw20_wet_tminus1, hrsblw20_wet_tminus2, hrsblw20_wet_tminus3)),
+
             avgtmax_wet_t_t1 = rowMeans(cbind(avgtmax_wet, avgtmax_wet_tminus1)),
@@
              avgtmax_wet_t1t2 = rowMeans(cbind(avgtmax_wet_tminus1, avgtmax_wet_tminus2)),
+
+            avg_hrs1020_wet_t_t1 = rowMeans(cbind(hrs1020_wet, hrs1020_wet_tminus1)),
+             avg_hrs1020_wet_t_t1t2 = rowMeans(cbind(hrs1020_wet, hrs1020_wet_tminus1, hrs1020_wet_tminus2)),
+             avg_hrs1020_wet_t1t2 = rowMeans(cbind(hrs1020_wet_tminus1, hrs1020_wet_tminus2)),
 
@@
 hrs1420_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, hrs_14_20_rs))
+hrs1020_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, hrs_10_20_rs))
+avgtmin_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, avg_tmin_rs))
 avgtmax_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, avg_tmax_rs))
@@
 hrsblw10_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, hrs_blw10rs))
+hrsblw14_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, hrs_blw14rs))
+hrsblw20_rs_lag <- as.tbl(select(rs_temps, plotid, sample_year, hrs_blw20rs))
 
@@
 
+## Number of rainy season hours 10-20 C ####
+rs_temps_lag <- left_join(rs_temps_lag, hrs1020_rs_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrs1020_rs_tminus1 = hrs_10_20_rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrs1020_rs_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrs1020_rs_tminus2 = hrs_10_20_rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrs1020_rs_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrs1020_rs_tminus3 = hrs_10_20_rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrs1020_rs_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
 ## Average daily maximum temperature during rainy season lagged ####
@@
 
+## Average daily minimum temperature during rainy season lagged ####
+rs_temps_lag <- left_join(rs_temps_lag, avgtmin_rs_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(avgtmin_rs_tminus1 = avg_tmin_rs)
+rs_temps_lag <- left_join(rs_temps_lag, avgtmin_rs_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(avgtmin_rs_tminus2 = avg_tmin_rs)
+rs_temps_lag <- left_join(rs_temps_lag, avgtmin_rs_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(avgtmin_rs_tminus3 = avg_tmin_rs)
+rs_temps_lag <- left_join(rs_temps_lag, avgtmin_rs_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
 ## Number of hours above 20 C during rainy season lagged ####
@@
 rs_temps_lag <- left_join(rs_temps_lag, hrsabv20_rs_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## Number of hours below 20 C during rainy season lagged ####
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw20_rs_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrsblw20_rs_tminus1 = hrs_blw20rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw20_rs_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrsblw20_rs_tminus2 = hrs_blw20rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw20_rs_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrsblw20_rs_tminus3 = hrs_blw20rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw20_rs_lag,
+                by = c("plotid", "sample_year"="sample_year"))
+
+## Number of hours below 14 C during rainy season lagged ####
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw14_rs_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrsblw14_rs_tminus1 = hrs_blw14rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw14_rs_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrsblw14_rs_tminus2 = hrs_blw14rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw14_rs_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrsblw14_rs_tminus3 = hrs_blw14rs)
+rs_temps_lag <- left_join(rs_temps_lag, hrsblw14_rs_lag,
                 by = c("plotid", "sample_year"="sample_year"))
@@
 
+            avgtmin_rs_t_t1 = rowMeans(cbind(avg_tmin_rs, avgtmin_rs_tminus1)),
+             avgtmin_rs_t_t1t2 = rowMeans(cbind(avg_tmin_rs, avgtmin_rs_tminus1, avgtmin_rs_tminus2)),
+             avgtmin_rs_t_t1t2t3 = rowMeans(cbind(avg_tmin_rs, avgtmin_rs_tminus1, avgtmin_rs_tminus2, avgtmin_rs_tminus3)),
+             avgtmin_rs_t1t2 = rowMeans(cbind(avgtmin_rs_tminus1, avgtmin_rs_tminus2)),
+
             avg_hrs1420_rs_tminus1 = rowMeans(cbind(hrs_14_20_rs, hrs1420_rs_tminus1)),
@@
 
+            avg_hrs1020_rs_tminus1 = rowMeans(cbind(hrs_10_20_rs, hrs1020_rs_tminus1)),
+             avg_hrs1020_rs_tminus2 = rowMeans(cbind(hrs_10_20_rs, hrs1020_rs_tminus1, hrs1020_rs_tminus2)),
+             avg_hrs1020_rs_tminus3 = rowMeans(cbind(hrs_10_20_rs, hrs1020_rs_tminus1, hrs1020_rs_tminus2, hrs1020_rs_tminus3)),
+             avg_hrs1020_rs_t1t2 = rowMeans(cbind(hrs1020_rs_tminus1, hrs1020_rs_tminus2)),
+
             avg_hrsblw10_rs_tminus1 = rowMeans(cbind(hrs_blw10rs, hrsblw10_rs_tminus1)),
@@
              avg_hrsblw10_rs_t1t2 = rowMeans(cbind(hrsblw10_rs_tminus1, hrsblw10_rs_tminus2)),
+
+            avg_hrsblw14_rs_tminus1 = rowMeans(cbind(hrs_blw14rs, hrsblw14_rs_tminus1)),
+             avg_hrsblw14_rs_tminus2 = rowMeans(cbind(hrs_blw14rs, hrsblw14_rs_tminus1, hrsblw14_rs_tminus2)),
+             avg_hrsblw14_rs_tminus3 = rowMeans(cbind(hrs_blw14rs, hrsblw14_rs_tminus1, hrsblw14_rs_tminus2, hrsblw14_rs_tminus3)),
+             avg_hrsblw14_rs_t1t2 = rowMeans(cbind(hrsblw14_rs_tminus1, hrsblw14_rs_tminus2)),
+
+            avg_hrsblw20_rs_tminus1 = rowMeans(cbind(hrs_blw20rs, hrsblw20_rs_tminus1)),
+             avg_hrsblw20_rs_tminus2 = rowMeans(cbind(hrs_blw20rs, hrsblw20_rs_tminus1, hrsblw20_rs_tminus2)),
+             avg_hrsblw20_rs_tminus3 = rowMeans(cbind(hrs_blw20rs, hrsblw20_rs_tminus1, hrsblw20_rs_tminus2, hrsblw20_rs_tminus3)),
+             avg_hrsblw20_rs_t1t2 = rowMeans(cbind(hrsblw20_rs_tminus1, hrsblw20_rs_tminus2)),
 
@@
 hrs_abv25ds_lag <- as.tbl(select(summer_temps, plotid, sample_year, hrs_abv25ds))
+hrs_abv20ds_lag <- as.tbl(select(summer_temps, plotid, sample_year, hrs_abv20ds))
+hrs_blw20ds_lag <- as.tbl(select(summer_temps, plotid, sample_year, hrs_blw20ds))
 avg_tmax_ds_lag <- as.tbl(select(summer_temps, plotid, sample_year, avg_tmax_ds))
-hrs_blw18ds_lag <- as.tbl(select(summer_temps, plotid, sample_year, hrs_blw18ds))
+avg_tmin_ds_lag <- as.tbl(select(summer_temps, plotid, sample_year, avg_tmin_ds))
 
@@
 
+## Number of dry-season hours above 20 C lagged ####
+summer_temps_lag <- left_join(summer_temps_lag, hrs_abv20ds_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrs_abv20ds_tminus1 = hrs_abv20ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_abv20ds_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrs_abv20ds_tminus2 = hrs_abv20ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_abv20ds_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrs_abv20ds_tminus3 = hrs_abv20ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_abv20ds_lag,
+                              by = c("plotid", "sample_year"="sample_year"))
+
 ## Average daily maximum temperature during dry-season lagged ####
@@
 
-## Number of dry-season hours below 18 C lagged ####
-summer_temps_lag <- left_join(summer_temps_lag, hrs_blw18ds_lag,
-                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
-      rename(hrs_blw18ds_tminus1 = hrs_blw18ds)
-summer_temps_lag <- left_join(summer_temps_lag, hrs_blw18ds_lag,
-                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
-      rename(hrs_blw18ds_tminus2 = hrs_blw18ds)
-summer_temps_lag <- left_join(summer_temps_lag, hrs_blw18ds_lag,
-                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
-      rename(hrs_blw18ds_tminus3 = hrs_blw18ds)
-summer_temps_lag <- left_join(summer_temps_lag, hrs_blw18ds_lag,
+## Average daily minimum temperature during dry-season lagged ####
+summer_temps_lag <- left_join(summer_temps_lag, avg_tmin_ds_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(avg_tmin_ds_tminus1 = avg_tmin_ds)
+summer_temps_lag <- left_join(summer_temps_lag, avg_tmin_ds_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(avg_tmin_ds_tminus2 = avg_tmin_ds)
+summer_temps_lag <- left_join(summer_temps_lag, avg_tmin_ds_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(avg_tmin_ds_tminus3 = avg_tmin_ds)
+summer_temps_lag <- left_join(summer_temps_lag, avg_tmin_ds_lag,
+                              by = c("plotid","sample_year"="sample_year"))
+
+## Number of dry-season hours below 20 C lagged ####
+summer_temps_lag <- left_join(summer_temps_lag, hrs_blw20ds_lag,
+                by = c("plotid", "sample_year_minus1"="sample_year")) %>%
+      rename(hrs_blw20ds_tminus1 = hrs_blw20ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_blw20ds_lag,
+                by = c("plotid", "sample_year_minus2"="sample_year")) %>%
+      rename(hrs_blw20ds_tminus2 = hrs_blw20ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_blw20ds_lag,
+                by = c("plotid", "sample_year_minus3"="sample_year")) %>%
+      rename(hrs_blw20ds_tminus3 = hrs_blw20ds)
+summer_temps_lag <- left_join(summer_temps_lag, hrs_blw20ds_lag,
                               by = c("plotid", "sample_year"="sample_year"))
@@
       mutate(
-            avg_hrsblw18_ds_tminus1 = rowMeans(cbind(hrs_blw18ds, hrs_blw18ds_tminus1)),
-             avg_hrsblw18_ds_tminus2 = rowMeans(cbind(hrs_blw18ds, hrs_blw18ds_tminus1, hrs_blw18ds_tminus2)),
-             avg_hrsblw18_ds_tminus3 = rowMeans(cbind(hrs_blw18ds, hrs_blw18ds_tminus1, hrs_blw18ds_tminus2, hrs_blw18ds_tminus3)),
-             avg_hrsblw18_ds_t1t2 = rowMeans(cbind(hrs_blw18ds_tminus1, hrs_blw18ds_tminus2)),
+            avg_hrsblw20_ds_tminus1 = rowMeans(cbind(hrs_blw20ds, hrs_blw20ds_tminus1)),
+             avg_hrsblw20_ds_tminus2 = rowMeans(cbind(hrs_blw20ds, hrs_blw20ds_tminus1, hrs_blw20ds_tminus2)),
+             avg_hrsblw20_ds_tminus3 = rowMeans(cbind(hrs_blw20ds, hrs_blw20ds_tminus1, hrs_blw20ds_tminus2, hrs_blw20ds_tminus3)),
+             avg_hrsblw20_ds_t1t2 = rowMeans(cbind(hrs_blw20ds_tminus1, hrs_blw20ds_tminus2)),
 
@@
              avgtmax_ds_t1t2 = rowMeans(cbind(avg_tmax_ds_tminus1, avg_tmax_ds_tminus2)),
+
+            avgtmin_ds_t_t1 = rowMeans(cbind(avg_tmin_ds, avg_tmin_ds_tminus1)),
+             avgtmin_ds_t_t1t2 = rowMeans(cbind(avg_tmin_ds, avg_tmin_ds_tminus1, avg_tmin_ds_tminus2)),
+             avgtmin_ds_t_t1t2t3 = rowMeans(cbind(avg_tmin_ds, avg_tmin_ds_tminus1, avg_tmin_ds_tminus2, avg_tmin_ds_tminus3)),
+             avgtmin_ds_t1t2 = rowMeans(cbind(avg_tmin_ds_tminus1, avg_tmin_ds_tminus2)),
+
+            avg_hrsabv20_ds_tminus1 = rowMeans(cbind(hrs_abv20ds, hrs_abv20ds_tminus1)),
+             avg_hrsabv20_ds_tminus2 = rowMeans(cbind(hrs_abv20ds, hrs_abv20ds_tminus1, hrs_abv20ds_tminus2)),
+             avg_hrsabv20_ds_tminus3 = rowMeans(cbind(hrs_abv20ds, hrs_abv20ds_tminus1, hrs_abv20ds_tminus2, hrs_abv20ds_tminus3)),
+             avg_hrsabv20_ds_t1t2 = rowMeans(cbind(hrs_abv20ds_tminus1, hrs_abv20ds_tminus2)),
 

### 2016-08-26 climate-path-model.R
Update source data.

@@ -46,7 +46,7 @@ summary(wet_hrs)# since these are counts the NAs are in fact zeroes
 wet_hrs[is.na(wet_hrs)] <- 0# introduce a lot of zeroes for some columns
 anyDuplicated(wet_hrs)
 oak_sod <- left_join(
-      oak_sod, select(wet_hrs, -X), by = c("plotid","sample_year"))
+      oak_sod, wet_hrs, by = c("plotid","sample_year"))
 anyDuplicated(oak_sod)
 # oak_sod$plotid <- as.factor(oak_sod$plotid)
 

### 2016-09-01 pathogen-spillover-diss-ch.Rmd
Add back bay laurel density variable.

@@
 library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
-library(dplyr, verbose = F)
+library(plyr); library(dplyr, verbose = F)
 library(lme4, verbose = F)
@@
 
-      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ umca_density_ha.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ umca_density_ha.log + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
-      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      hrs_abv25 <- lmer(hrs_abv25ds ~ umca_density_ha.log + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
@@
 
-      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ umca_density_ha.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ umca_density_ha.log.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
 
-      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ umca_density_ha.log.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
 

### 2016-09-01 pathogen-spillover-diss-ch.Rmd
Add separate models without bay laurel density.

@@
 Each variable is first centered by subtracting its mean and then scaled by dividing by its standard deviation.
+
+```{r hours >25 path model no umca density unscaled, echo=F}
+model_list2 <- list(
+      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
+
+      # umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      slc <- lmer(tot_lfct.log ~ hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+)
+```
+
+```{r hours >25 path model no umca density scaled}
+model_list2.scl <- list(
+      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),
+
+      # umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      slc <- lmer(tot_lfct.log.scl ~ hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+)
+```
 
@@
           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+
+raw_variables_model_noUMCA <- sem.fit(model_list2, oakplots.sub, 
+        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+scaled_variables_model_noUMCA <- sem.fit(model_list2.scl, oakplots.sub,
+          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
 ```

### 2016-09-01 microclimate-correlations.Rmd
Revise variables in correlations and add bivariate plots.

@@
 microclimate_variables <- oakplots.sub %>%
-      select(sample_year, plotid, tot_lfct.log, avg_lfct.log, contains("hrs"), contains("tmax"), contains("tmin"))
+      select(sample_year, plotid, tot_lfct.log, avg_lfct.log, inf_oak_ct, uninf_oak_ct, contains("hrs"), contains("tmax"), contains("tmin")) %>% 
+      mutate(inf_oak.pct = inf_oak_ct/(inf_oak_ct + uninf_oak_ct))
+microclimate_variables <- na.omit(microclimate_variables)
 # summary(microclimate_variables)
@@
 
-```{r microclimate correlations, echo=F}
-pairs(select(filter(microclimate_variables, sample_year == 2005), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
+```{r microclimate correlations, echo=F, fig.width=12, fig.height=8, eval=T}
+pairs(select(filter(microclimate_variables, sample_year == 2005), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
       main = "Correlations Between Variables for 2005 Sample Year \n n=165 oak plots")
-pairs(select(filter(microclimate_variables, sample_year == 2006), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
+pairs(select(filter(microclimate_variables, sample_year == 2006), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
       main = "Correlations Between Variables for 2006 Sample Year \n n=167 oak plots")
-pairs(select(filter(microclimate_variables, sample_year == 2007), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
+pairs(select(filter(microclimate_variables, sample_year == 2007), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
       main = "Correlations Between Variables for 2007 Sample Year \n n=164 oak plots")
-pairs(select(filter(microclimate_variables, sample_year == 2008), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
+pairs(select(filter(microclimate_variables, sample_year == 2008), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
       main = "Correlations Between Variables for 2008 Sample Year \n n=163 oak plots")
-pairs(select(filter(microclimate_variables, sample_year == 2009), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avgtmax_rs_t_t1, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
+pairs(select(filter(microclimate_variables, sample_year == 2009), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
       main = "Correlations Between Variables for 2009 Sample Year \n n=162 oak plots")
-pairs(select(filter(microclimate_variables, sample_year == 2010), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
+pairs(select(filter(microclimate_variables, sample_year == 2010), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
       main = "Correlations Between Variables for 2010 Sample Year \n n=159 oak plots")
-pairs(select(filter(microclimate_variables, sample_year == 2011), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
+pairs(select(filter(microclimate_variables, sample_year == 2011), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
       main = "Correlations Between Variables for 2011 Sample Year \n n=156 oak plots")
-pairs(select(filter(microclimate_variables, sample_year == 2012), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds), lower.panel = panel.cor,
+pairs(select(filter(microclimate_variables, sample_year == 2012), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
       main = "Correlations Between Variables for 2012 Sample Year \n n=156 oak plots")
-pairs(select(filter(microclimate_variables, sample_year == 2014), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmin_rs, avg_tmax_ds, avg_tmin_ds), lower.panel = panel.cor,
+pairs(select(filter(microclimate_variables, sample_year == 2014), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1), lower.panel = panel.cor,
       main = "Correlations Between Variables for 2014 Sample Year \n n=156 oak plots")
-pairs(na.omit(select(filter(microclimate_variables, sample_year != 2004), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds, avgtmax_wet)),
+pairs(na.omit(select(filter(microclimate_variables, sample_year != 2004), tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw10_wet, avg_tmax_rs, avg_tmax_ds, hrsblw10_wet_tminus1, hrsblw14_wet_tminus1)),
       lower.panel = panel.cor,
       main = "Correlations Between Variables \n all oak plots 2005-2014")
+
+# pairs(na.omit(select(filter(microclimate_variables, sample_year != 2004), inf_oak_ct, tot_lfct.log, avg_lfct.log, avg_hrs1422_wet_t_t1.log, hrs_abv25ds, hrs_blw10rs, hrs1020_wet, hrs_blw14rs, hrsblw14_wet, hrsblw14_wet_tminus1, hrsblw10_rs_tminus1, avg_tmax_rs, avg_tmax_ds, avg_tmin_rs, avg_tmin_ds, avgtmax_wet, avg_hrsblw14_wet_tminus1, avg_hrsblw10_rs_tminus1)),
+#       lower.panel = panel.cor,
+#       main = "Correlations Between Variables \n all oak plots 2005-2014")
 
@@
 
+```{r bivariate linear model hrsblw10rs, echo=F, fig.height=8, message=F}
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrs_blw10rs
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+# y <- "avg_hrs1422_wet_t_t1.log"
+# x <- "avgtmax_wet"
+# reg_eqn(df = na.omit(microclimate_variables), y = y, x = x)
+
+eqns <- ddply(na.omit(microclimate_variables), "sample_year", reg_eqn)
+
+ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
+             group_by(sample_year),
+       aes(x = hrs_blw10rs, y = tot_lfct.log)) +
+      geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Rainy-season Hours <10 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("Oak Plots") +
+      theme_bw()
+```
+
+```{r bivariate linear model hrsblw14rs, echo=F, fig.height=8, message=F}
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrs_blw14rs
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+# y <- "avg_hrs1422_wet_t_t1.log"
+# x <- "avgtmax_wet"
+# reg_eqn(df = na.omit(microclimate_variables), y = y, x = x)
+
+eqns <- ddply(na.omit(microclimate_variables), "sample_year", reg_eqn)
+
+ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
+             group_by(sample_year),
+       aes(x = hrs_blw14rs, y = tot_lfct.log)) +
+      geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Rainy-season Hours <14 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("Oak Plots") +
+      theme_bw()
+```
+
+```{r bivariate linear model hrsblw14_wet, echo=F, fig.height=8, message=F}
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrsblw14_wet
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+# y <- "avg_hrs1422_wet_t_t1.log"
+# x <- "avgtmax_wet"
+# reg_eqn(df = na.omit(microclimate_variables), y = y, x = x)
+
+eqns <- ddply(na.omit(microclimate_variables), "sample_year", reg_eqn)
+
+ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
+             group_by(sample_year),
+       aes(x = hrsblw14_wet, y = tot_lfct.log)) +
+      geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Wet-hours <14 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("Oak Plots") +
+      theme_bw()
+```
+
+```{r bivariate linear model hrs1020_wet, echo=F, fig.height=8, message=F}
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrs1020_wet
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+# y <- "avg_hrs1422_wet_t_t1.log"
+# x <- "avgtmax_wet"
+# reg_eqn(df = na.omit(microclimate_variables), y = y, x = x)
+
+eqns <- ddply(na.omit(microclimate_variables), "sample_year", reg_eqn)
+
+ggplot(filter(microclimate_variables, sample_year != 2004) %>% 
+             group_by(sample_year),
+       aes(x = hrs1020_wet, y = tot_lfct.log)) +
+      geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Wet-hours 10-20 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("Oak Plots") +
+      theme_bw()
+```
+
 Neither do a very good job of explaining the variation in leaf count. I wonder if this is at least partly due to the plots with zeroes.
@@
 ```{r bivariate models using umca plots, echo=F, fig.height=8, message=F}
-umca_plots <- left_join(umca_plots, select(rs_temps_lag, -avg_tmax_rs), by = c("plotid","sample_year"))
-umca_plots <- left_join(umca_plots, wet_hrs_lag, by = c("plotid","sample_year"))
+# umca_plots <- left_join(umca_plots, select(rs_temps_lag, -avg_tmax_rs), by = c("plotid","sample_year"))
+umca_plots <- left_join(umca_plots, wet_hrs, by = c("plotid","sample_year"))
 umca_plots$tot_lfct.log <- log(umca_plots$tot_lfct+1)
@@
 
+```{r bivariate model umca plots hrsblw10rs, echo=F, fig.height=8, message=F}
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrs_blw10rs
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+# y <- "avg_hrs1422_wet_t_t1.log"
+# x <- "avgtmax_wet"
+# reg_eqn(df = na.omit(umca_plots), y = y, x = x)
+
+eqns <- ddply(na.omit(umca_plots), "sample_year", reg_eqn)
+
+ggplot(filter(umca_plots, sample_year != 2004) %>% 
+             group_by(sample_year),
+       aes(x = hrs_blw10rs, y = tot_lfct.log)) +
+      geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Rainy-season Hours <10 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("UMCA Plots") +
+      theme_bw()
+```
+
+```{r bivariate model umca plots hrsblw14rs, echo=F, fig.height=8, message=F}
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrs_blw14rs
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+# y <- "avg_hrs1422_wet_t_t1.log"
+# x <- "avgtmax_wet"
+# reg_eqn(df = na.omit(umca_plots), y = y, x = x)
+
+eqns <- ddply(na.omit(umca_plots), "sample_year", reg_eqn)
+
+ggplot(filter(umca_plots, sample_year != 2004) %>% 
+             group_by(sample_year),
+       aes(x = hrs_blw14rs, y = tot_lfct.log)) +
+      geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Rainy-season Hours <14 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("UMCA Plots") +
+      theme_bw()
+```
+
+```{r bivariate model umca plots hrsblw14_wet, echo=F, fig.height=8, message=F}
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrsblw14_wet
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+# y <- "avg_hrs1422_wet_t_t1.log"
+# x <- "avgtmax_wet"
+# reg_eqn(df = na.omit(umca_plots), y = y, x = x)
+
+eqns <- ddply(na.omit(umca_plots), "sample_year", reg_eqn)
+
+ggplot(filter(umca_plots, sample_year != 2004) %>% 
+             group_by(sample_year),
+       aes(x = hrsblw14_wet, y = tot_lfct.log)) +
+      geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Wet-hours <14 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("UMCA Plots") +
+      theme_bw()
+```
+
+```{r bivariate model umca hrs1020_wet, echo=F, fig.height=8, message=F}
+reg_eqn <- function(df,y,x) {
+      y = df$tot_lfct.log
+      x = df$hrs1020_wet
+      # y = y
+      # x = x
+      m <- lm(y ~ x, data = df)
+
+      a <- coef(m)[1]
+      b <- coef(m)[2]
+      r2 <- summary(m)$r.squared
+
+      expr <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
+                         list(a = format(a, digits = 2),
+                              b = format(b, digits = 2),
+                              r2 = format(r2, digits = 3)))
+      c(lab = as.character(as.expression(expr)))
+}
+# y <- "avg_hrs1422_wet_t_t1.log"
+# x <- "avgtmax_wet"
+# reg_eqn(df = na.omit(umca_plots), y = y, x = x)
+
+eqns <- ddply(na.omit(umca_plots), "sample_year", reg_eqn)
+
+ggplot(filter(umca_plots, sample_year != 2004) %>% 
+             group_by(sample_year),
+       aes(x = hrs1020_wet, y = tot_lfct.log)) +
+      geom_point() +
+      geom_smooth(method = "lm", size = 0.7) +
+      geom_text(aes(x = 0, y = 3, label = lab), data = eqns, parse = T,
+                hjust = 0, vjust = 1) +
+      facet_grid(sample_year ~ .) +
+      xlab("Wet-hours 10-20 C") +
+      ylab("Pathogen Inoculum Load (ln(Total Symptomatic Leaf Count))") +
+      ggtitle("UMCA Plots") +
+      theme_bw()
+```
+
 These appear to do a little better in explaining the variation, but not for all years.

### 2016-09-09 climate-path-model.R
Comment out p-value reporting for pairs plots.

@@ -18,11 +18,11 @@ panel.cor <- function(x, y, digits = 2, cex.cor, ...)
       text(0.5, 0.6, txt)
       
       # p-value calculation
-      p <- cor.test(x, y)$p.value
-      txt2 <- format(c(p, 0.123456789), digits = digits)[1]
-      txt2 <- paste("p= ", txt2, sep = "")
-      if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
-      text(0.5, 0.4, txt2)
+      # p <- cor.test(x, y)$p.value
+      # txt2 <- format(c(p, 0.123456789), digits = digits)[1]
+      # txt2 <- paste("p= ", txt2, sep = "")
+      # if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
+      # text(0.5, 0.4, txt2)
 }
 
 panel.hist <- function(x, ...)

### 2016-09-09 pathogen-spillover-diss-ch.Rmd
Add figure printing for temperature and rainfall.

@@
 library(lmerTest, verbose = F)
+
+panel.cor <- function(x, y, digits = 2, cex.cor, ...)
+{
+      usr <- par("usr"); on.exit(par(usr))
+      par(usr = c(0, 1, 0, 1))
+      # correlation coefficient
+      r <- cor(x, y)
+      txt <- format(c(r, 0.123456789), digits = digits)[1]
+      txt <- paste("r= ", txt, sep = "")
+      text(0.5, 0.6, txt)
+
+      # p-value calculation
+      # p <- cor.test(x, y)$p.value
+      # txt2 <- format(c(p, 0.123456789), digits = digits)[1]
+      # txt2 <- paste("p= ", txt2, sep = "")
+      # if(p<0.01) txt2 <- paste("p= ", "<0.01", sep = "")
+      # text(0.5, 0.4, txt2)
+}
 
@@
 
-```{r hours >25 path model no umca density scaled}
+```{r hours >25 path model no umca density scaled, echo=F}
 model_list2.scl <- list(
@@
 ```{r symptomatic leaf count residuals plot, echo=F}
-plot(slc, main = "LMM of symptomatic leaf count residuals")
+plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
+# qqnorm(resid(slc), main="Q-Q plot for conditional residuals - SLC")
+# qqnorm(ranef(slc)$sample_year$`(Intercept)`, 
+#        main="Q-Q plot for Sample Year")
+# qqnorm(ranef(slc)$plotid$`(Intercept)`, 
+#        main="Q-Q plot for plot ID")
+
 plot(slc@frame$umca_density_ha.log.scl, resid(slc),
@@
 ```
-```{r pairs plot of slc predictor variables}
-pairs(select(oakplots.sub, umca_density_ha.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl, twi15m.log.scl),
+```{r pairs plot of slc predictor variables, echo=F}
+pairs(na.omit(select(oakplots.sub, umca_density_ha.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl, twi15m.log.scl)),
       lower.panel = panel.cor, diag.panel = panel.hist,
@@
 ```{r pairs plot of oak infection predictor variables, echo=F}
-pairs(select(oakplots.sub, tot_lfct.log.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl),
+pairs(na.omit(select(oakplots.sub, tot_lfct.log.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl)),
       lower.panel = panel.cor, diag.panel = panel.hist,
@@
 
-```{r fit and compare all oak plots, echo=F, message=F}
+```{r fit and compare all oak plots, echo=F, message=F, warning=F}
 raw_variables_model <- sem.fit(model_list1, oakplots.sub, 
-        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+        .progressBar = F)
 scaled_variables_model <- sem.fit(model_list1.scl, oakplots.sub,
-          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
+          .progressBar = F)
 
 raw_variables_model_noUMCA <- sem.fit(model_list2, oakplots.sub, 
-        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+        .progressBar = F)
 scaled_variables_model_noUMCA <- sem.fit(model_list2.scl, oakplots.sub,
-          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
+          .progressBar = F)
 ```
@@
       caption = "Results of the d-sep test using standardized variables.")
+
+kable(raw_variables_model_noUMCA, 
+      caption = "Results of the d-sep test using raw variables - no UMCA density.")
+kable(scaled_variables_model_noUMCA,
+      caption = "Results of the d-sep test using standardized variables - no UMCA density.")
 ```
@@
           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+
+raw_coefficients_noUMCA <- sem.coefs(model_list2, oakplots.sub,
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+scaled_coefficients_noUMCA <- sem.coefs(model_list2.scl, oakplots.sub,
+          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
 ```
@@
 ```{r tables of path coefficients, echo=F}
-library(knitr)
 kable(raw_coefficients,
       caption = "Raw path coefficients")
+kable(raw_coefficients_noUMCA,
+      caption = "Raw path coefficients - no UMCA density")
+
 kable(scaled_coefficients,
       caption = "Standardized path coefficients")
+kable(scaled_coefficients_noUMCA,
+      caption = "Standardized path coefficients - no UMCA density")
 ```
@@
 ## Temperature Figures
-```{r temperature figures, echo=F}
+```{r temperature figures, echo=F, warning=F, eval=F}
 library(ggplot2)
@@
 
-sstemps_figure <- temps_figure + 
+(sstemps_figure <- temps_figure + 
       geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds)) +
@@
       ggtitle("Dry-season Hours >25 C")
-
-slc_avg_figure <- temps_figure + 
+)     
+(slc_avg_figure <- temps_figure + 
       geom_boxplot(aes(as.factor(sample_year), avg_lfct)) +
@@
       ggtitle("California bay laurel leaf symptoms - plot averages")
-
-slc_total_figure <- temps_figure + 
+)
+(slc_total_figure <- temps_figure + 
       geom_boxplot(aes(as.factor(sample_year), tot_lfct)) +
@@
       ggtitle("California bay laurel leaf symptoms - plot totals")
-
-rstemps_figure <- temps_figure + 
+)
+(rstemps_figure <- temps_figure + 
       geom_boxplot(aes(as.factor(sample_year), hrs1422_wet)) +
@@
       ggtitle("Wet-hours 14-22 C")
-
-oak_infection_figure <- temps_figure +
+)
+(oak_infection_figure <- temps_figure +
       geom_bar(aes(as.factor(sample_year), inf_oak_ct), stat = "identity") +
@@
       ggtitle("Incident oak infection")
+)
 ```
 
-## Rainfall Figures
-```{r rainfall figures}
+## Rainfall Figure
+Total annual rainfall estimated at plot locations using two-dimensional interpolation methods implemented in GRASS-GIS. The red line indicates the average rainfall from 2005-2014
+```{r plot rainfall figure, echo=F}
+library(ggplot2)
 # summary(rainy_season_ppt)
+# rainy_season_ppt$year <- as.factor(rainy_season_ppt$year)
+rainfall_figure <- ggplot(filter(rainy_season_ppt, year > 2004), 
+                          aes(year, rain2d_total))
+rainfall_figure +
+      geom_boxplot(aes(group = year)) +
+      geom_hline(aes(yintercept = mean(rain2d_total)),
+                 colour = "red") +
+      xlab("Sample Year") +
+      ylab("November-May Total Rainfall (cm)") +
+      ggtitle("Rainfall Estimated at Plots from Rain Gauges")
 ```
 
-```{r print session info}
+### Session Information
+```{r print session info, echo=FALSE}
 print(sessionInfo())

### 2016-09-14 analysis/species_diversity_calc.R
Minor change to species richness subsetting.

@@ -113,11 +113,12 @@ summary(us_richness)
 filter(us_richness, is.na(us_rich05)); filter(us_richness, is.na(us_rich11))
 us_richness <- filter(us_richness, plotid != "bush01", plotid != "halow01",
                       plotid != "sweet01")
+
 us_richness$us_rich05[us_richness$plotid=="jlsp05"] <- 0
-us_richness$us_rich05[us_richness$plotid=="ann06"] <- 0
 us_richness$us_rich11[us_richness$plotid=="ann06"] <- 0
-us_richness$us_rich05[us_richness$plotid=="mroth03"] <- 0
 us_richness$us_rich11[us_richness$plotid=="mroth03"] <- 0
+# us_richness$us_rich11[us_richness$plotid=="sumtv02"] <- 0
+# us_richness$us_rich11[us_richness$plotid=="mcneil01"] <- 0
 
 
 us_sub_richness <- full_join(
@@ -129,7 +130,7 @@ us_sub_richness <- full_join(
 summary(us_sub_richness)
 filter(us_sub_richness, is.na(us_rich11)); filter(us_sub_richness, is.na(us_rich05))
 # remove plots abandoned early in study
-us_sub_richness <- filter(us_sub_richness, 
+us_sub_richness <- filter(us_sub_richness,
                       plotid != "bush01", plotid != "halow01",
                       plotid != "sweet01")
 # mcneil01 was lost in 2009 (road built through plot), so NA for later years
@@ -138,6 +139,7 @@ us_sub_richness <- filter(us_sub_richness,
 us_sub_richness$us_rich11[us_sub_richness$plotid=="ann06"] <- 0
 us_sub_richness$us_rich11[us_sub_richness$plotid=="mroth03"] <- 0
 us_sub_richness$us_rich05[us_sub_richness$plotid=="jlsp05"] <- 0
+# us_sub_richness$us_rich11[us_sub_richness$plotid=="sumtv02"] <- 0
 # This leaves two legitimate NAs for understory species in 2011
 
 

### 2016-09-14 pathogen-spillover-diss-ch.Rmd
Add temperature and infection figures.
I also changed the title of the document.

@@
   html_document:
-    self_contained: no
+    toc: yes
   pdf_document: default
@@
 
-# CHAPTER 3: EFFECTS OF DIVERSITY, TOPOGRAPHY, AND INTERANNUAL CLIMATE VARIABILITY ON PATHOGEN SPILLOVER
+# Appendix 2: Analysis of Environmental Controls on Pathogen Spillover
+Effects of Diversity, Topography, and Interannual Climate Variability
 
@@
 
+## Disease Figures
+```{r disease figures, echo=F, warning=F}
+library(ggplot2)
+temperatures <- left_join(
+      select(summer_temps, plotid, sample_year, hrs_abv25ds),
+      select(wet_hrs, plotid, sample_year, hrs1422_wet),
+      by = c("plotid","sample_year"))
+# oak_sod$year <- as.Date(oak_sod$sample_year)
+temps_figure <- ggplot(oak_sod, aes(group = sample_year))
+
+(oak_infection_figure <- temps_figure +
+      geom_bar(aes(as.factor(sample_year), inf_oak_ct), stat = "identity") +
+      # geom_boxplot(aes(as.factor(sample_year), inf_oak_ct)) +
+      theme_bw() +
+      xlab("Sample Year") +
+      ylab("Number of infected oaks") +
+      ggtitle("Oak infection")
+)
+
+(slc_avg_figure <- temps_figure + 
+      geom_boxplot(aes(as.factor(sample_year), avg_lfct)) +
+      theme_bw() +
+      xlab("Sample Year") +
+      ylab("Symptomatic leaf count") +
+      ggtitle("California bay laurel leaf symptoms - plot averages")
+)
+
+(slc_total_figure <- temps_figure + 
+      geom_boxplot(aes(as.factor(sample_year), tot_lfct)) +
+      theme_bw() +
+      xlab("Sample Year") +
+      ylab("Symptomatic leaf count") +
+      ggtitle("California bay laurel leaf symptoms - plot totals")
+)
+```
+
 ## Temperature Figures
-```{r temperature figures, echo=F, warning=F, eval=F}
+```{r temperature figures, echo=F, warning=F, eval=T}
 library(ggplot2)
@@
 )
-(slc_avg_figure <- temps_figure + 
-      geom_boxplot(aes(as.factor(sample_year), avg_lfct)) +
-      theme_bw() +
-      xlab("Sample Year") +
-      ylab("Symptomatic leaf count") +
-      ggtitle("California bay laurel leaf symptoms - plot averages")
-)
-(slc_total_figure <- temps_figure + 
-      geom_boxplot(aes(as.factor(sample_year), tot_lfct)) +
-      theme_bw() +
-      xlab("Sample Year") +
-      ylab("Symptomatic leaf count") +
-      ggtitle("California bay laurel leaf symptoms - plot totals")
-)
+
 (rstemps_figure <- temps_figure + 
@@
 )
-(oak_infection_figure <- temps_figure +
-      geom_bar(aes(as.factor(sample_year), inf_oak_ct), stat = "identity") +
-      # geom_boxplot(aes(as.factor(sample_year), inf_oak_ct)) +
-      theme_bw() +
-      xlab("Sample Year") +
-      ylab("Number of infected oaks") +
-      ggtitle("Incident oak infection")
-)
+
 ```
@@
 ## Rainfall Figure
-Total annual rainfall estimated at plot locations using two-dimensional interpolation methods implemented in GRASS-GIS. The red line indicates the average rainfall from 2005-2014
+Total annual rainfall estimated at plot locations using two-dimensional interpolation methods implemented in GRASS-GIS. The red line indicates the average rainfall from 2005-2014.
+
 ```{r plot rainfall figure, echo=F}
@@
                  colour = "red") +
+      theme_bw() +
       xlab("Sample Year") +

### 2016-10-31 path-temp-compare.Rmd
Add plot of total rainfall by year.

@@
 oakplots.sub$umca_density_ha.log <- log(oakplots.sub$umca_density_ha+1)
+```
+
+```{r plot rainy season totals, echo=F}
+library(ggplot2)
+ggplot(rainy_season_ppt, aes(year, rain_v_total, group = year)) +
+      geom_boxplot() +
+      ylab("Rainfall (mm)") +
+      xlab("Rainy Season Year") +
+      ggtitle("Rainy Season Rainfall at Plots (nearest-neighbor assignment)") +
+      theme_bw()
+# ggsave("figures/rainy-season-ppt.png", dpi = 600)
 ```

### 2016-10-31 d-sep_test.Rmd
Add newline.

@@
 Choosing the correct modeling procedure requires identifying the appropriate distribution of the dependent variable, in this case `Date`.
-Using some helper fucntions and the `pairs()` function in the base `graphics` package I can look at the correlations and distributions simultaneously.
+Using some helper functions and the `pairs()` function in the base `graphics` package I can look at the correlations and distributions simultaneously.
+
 ```{r helper functions for scatterplot matrix, echo=FALSE}

### 2016-10-31 pathogen-spillover-diss-ch.Rmd
Add model data frame & revise output figures.

@@
   html_document:
+    self_contained: no
     toc: yes
@@
 # Appendix 2: Analysis of Environmental Controls on Pathogen Spillover
-Effects of Diversity, Topography, and Interannual Climate Variability
+**Effects of Diversity, Topography, and Interannual Climate Variability**
+
+[Conceptual Path Model](figures/conceptual-path-model-updated.png)
 
@@
 oakplots.sub$twi15m.log.scl <- as.numeric(scale(oakplots.sub$twi15m.log))
+
+## make consolidated data frame only of variables in the path model
+oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct.log, tot_lfct.log.scl, hrs_abv25ds, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log, umca_density_ha.log.scl, H.2005, shannons2005.scl, twi15m.log, twi15m.log.scl)
+
+oakplots.model.data <- filter(oakplots.model.data, is.na(H.2005)==F)
+write.csv(oakplots.model.data, "analysis/data/oakplots-pathmodel-data.csv",
+          row.names = F)
 ```
@@
 model_list2 <- list(
-      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
+      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.model.data, sample_year == 2006)),
 
@@
 
-      slc <- lmer(tot_lfct.log ~ hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      slc <- lmer(tot_lfct.log ~ hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      hrs_abv25 <- lmer(hrs_abv25ds ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
 )
@@
 model_list2.scl <- list(
-      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),
+      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),
 
-      # umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      # umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      slc <- lmer(tot_lfct.log.scl ~ hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      slc <- lmer(tot_lfct.log.scl ~ hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
 )
@@
 ```{r hours >25 path model all oak plots unscaled, echo=F}
+
 model_list1 <- list(
-      shannons <- lm(H.2005 ~ twi15m.log, data = oakplots.sub),
+      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.model.data, sample_year == 2006)),
 
-      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ umca_density_ha.log + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log ~ 
+                                   # umca_density_ha.log + twi15m.log +
+                                   twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      hrs_abv25 <- lmer(hrs_abv25ds ~ umca_density_ha.log + twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      hrs_abv25 <- lmer(hrs_abv25ds ~ 
+                              # umca_density_ha.log + 
+                              twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
 )
@@
 model_list1.scl <- list(
-      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = oakplots.sub),
+      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),
 
-      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      slc <- lmer(tot_lfct.log.scl ~ umca_density_ha.log.scl + hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      slc <- lmer(tot_lfct.log.scl ~ umca_density_ha.log.scl + hrs_abv25ds.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~ umca_density_ha.log.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~
+                                   # umca_density_ha.log.scl + 
+                                   twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ umca_density_ha.log.scl + twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.sub),
+      hrs_abv25 <- lmer(hrs_abv25ds.scl ~ 
+                              # umca_density_ha.log.scl + 
+                              twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.sub)
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + avg_hrs1422_wet_t_t1.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
+)
+
+## remove microclimate variables
+model_list1.scla <- list(
+      shannons <- lm(shannons2005.scl ~ twi15m.log.scl, data = filter(oakplots.model.data, sample_year == 2006)),
+
+      umca_density <- lmer(umca_density_ha.log.scl ~ twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+
+      slc <- lmer(tot_lfct.log.scl ~ umca_density_ha.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+
+      # wet_hours_1422 <- lmer(avg_hrs1422_wet_t_t1.log.scl ~
+      #                              # umca_density_ha.log.scl + 
+      #                              twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+      # 
+      # hrs_abv25 <- lmer(hrs_abv25ds.scl ~ 
+      #                         # umca_density_ha.log.scl + 
+      #                         twi15m.log.scl + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log.scl + shannons2005.scl + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
 )
@@
      main = "SLC model residuals vs. Shannon's diversity predictor")
-plot(slc@frame$twi15m.log.scl, resid(slc),
-     xlab = "wetness index",
-     ylab = "model residuals",
-     main = "SLC model residuals vs. wetness index predictor")
+# plot(slc@frame$twi15m.log.scl, resid(slc),
+#      xlab = "wetness index",
+#      ylab = "model residuals",
+#      main = "SLC model residuals vs. wetness index predictor")
 ```
@@
 plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
-plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
+# plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
+#      xlab = "bay laurel density",
+#      ylab = "model residuals",
+#      main = "Wet-hours model residuals vs. bay laurel density predictor")
+plot(wet_hours_1422@frame$twi15m.log.scl, resid(wet_hours_1422),
      xlab = "bay laurel density",
@@
      main = "Dry-season hours model residuals vs. wetness index predictor")
+# plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
+#      xlab = "wetness index",
+#      ylab = "model residuals",
+#      main = "Dry-season hours model residuals vs. wetness index predictor")
 ```
@@
 ```{r fit and compare all oak plots, echo=F, message=F, warning=F}
-raw_variables_model <- sem.fit(model_list1, oakplots.sub, 
+raw_variables_model <- sem.fit(model_list1, oakplots.model.data, 
         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-        .progressBar = F)
-scaled_variables_model <- sem.fit(model_list1.scl, oakplots.sub,
+        conditional = T, .progressBar = F)
+scaled_variables_model <- sem.fit(model_list1.scl, oakplots.model.data,
           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
-          .progressBar = F)
+          conditional = T, .progressBar = F)
 
-raw_variables_model_noUMCA <- sem.fit(model_list2, oakplots.sub, 
-        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-        .progressBar = F)
-scaled_variables_model_noUMCA <- sem.fit(model_list2.scl, oakplots.sub,
-          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
-          .progressBar = F)
+scaled_variables_modela <- sem.fit(model_list1.scla, oakplots.model.data,
+          # corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
+          conditional = F, .progressBar = F)
+
+# raw_variables_model_noUMCA <- sem.fit(model_list2, oakplots.sub, 
+#         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+#         .progressBar = F)
+# scaled_variables_model_noUMCA <- sem.fit(model_list2.scl, oakplots.sub,
+#           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
+#           .progressBar = F)
 ```
@@
 ```{r tables of d-sep test results from sem.fit, echo=F}
-kable(raw_variables_model, 
+kable(raw_variables_model, digits = 4,
       caption = "Results of the d-sep test using raw variables.")
@@
 
-kable(raw_variables_model_noUMCA, 
-      caption = "Results of the d-sep test using raw variables - no UMCA density.")
-kable(scaled_variables_model_noUMCA,
-      caption = "Results of the d-sep test using standardized variables - no UMCA density.")
+# kable(raw_variables_model_noUMCA, 
+#       caption = "Results of the d-sep test using raw variables - no UMCA density.")
+# kable(scaled_variables_model_noUMCA,
+#       caption = "Results of the d-sep test using standardized variables - no UMCA density.")
 ```
@@
 ```{r get path coefficients, echo=F}
-raw_coefficients <- sem.coefs(model_list1, oakplots.sub,
+raw_coefficients <- sem.coefs(model_list1, oakplots.model.data,
           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
-scaled_coefficients <- sem.coefs(model_list1.scl, oakplots.sub,
+scaled_coefficients <- sem.coefs(model_list1.scl, oakplots.model.data,
           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
 
-raw_coefficients_noUMCA <- sem.coefs(model_list2, oakplots.sub,
-          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
-scaled_coefficients_noUMCA <- sem.coefs(model_list2.scl, oakplots.sub,
-          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+# raw_coefficients_noUMCA <- sem.coefs(model_list2, oakplots.sub,
+#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+# scaled_coefficients_noUMCA <- sem.coefs(model_list2.scl, oakplots.sub,
+#           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
 ```
@@
 ```{r tables of path coefficients, echo=F}
-kable(raw_coefficients,
+kable(raw_coefficients, digits = 4,
       caption = "Raw path coefficients")
-kable(raw_coefficients_noUMCA,
-      caption = "Raw path coefficients - no UMCA density")
+# kable(raw_coefficients_noUMCA,
+#       caption = "Raw path coefficients - no UMCA density")
 
-kable(scaled_coefficients,
+kable(scaled_coefficients, digits = 4,
       caption = "Standardized path coefficients")
-kable(scaled_coefficients_noUMCA,
-      caption = "Standardized path coefficients - no UMCA density")
+# kable(scaled_coefficients_noUMCA,
+#       caption = "Standardized path coefficients - no UMCA density")
 ```
@@
 (slc_avg_figure <- temps_figure + 
-      geom_boxplot(aes(as.factor(sample_year), avg_lfct)) +
+      geom_boxplot(aes(as.factor(sample_year), avg_lfct), notch = F) +
       theme_bw() +
       xlab("Sample Year") +
-      ylab("Symptomatic leaf count") +
+      ylab("Pathogen Load \n(average symptomatic leaf count)") +
       ggtitle("California bay laurel leaf symptoms - plot averages")
@@
 (slc_total_figure <- temps_figure + 
-      geom_boxplot(aes(as.factor(sample_year), tot_lfct)) +
+      geom_boxplot(aes(as.factor(sample_year), tot_lfct), notch = F) +
       theme_bw() +
       xlab("Sample Year") +
-      ylab("Symptomatic leaf count") +
+      ylab("Pathogen Load \n(total symptomatic leaf count)") +
       ggtitle("California bay laurel leaf symptoms - plot totals")
@@
 (sstemps_figure <- temps_figure + 
-      geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds)) +
+      geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds), notch = T) +
       theme_bw() +
@@
       ylab("Number of hours") +
-      ggtitle("Dry-season Hours >25 C")
+      ggtitle("Heat Exposure (dry-season Hours >25 C)")
 )
@@
 (rstemps_figure <- temps_figure + 
-      geom_boxplot(aes(as.factor(sample_year), hrs1422_wet)) +
+      geom_boxplot(aes(as.factor(sample_year), hrs1422_wet), notch = T) +
       theme_bw() +
@@
       ylab("Number of hours") +
-      ggtitle("Wet-hours 14-22 C")
+      ggtitle("Warm & Wet Conditions (wet-hours 14-22 C)")
 )
@@
 rainfall_figure +
-      geom_boxplot(aes(group = year)) +
+      geom_boxplot(aes(group = year), notch = T) +
       geom_hline(aes(yintercept = mean(rain2d_total)),
@@
       xlab("Sample Year") +
-      ylab("November-May Total Rainfall (cm)") +
-      ggtitle("Rainfall Estimated at Plots from Rain Gauges")
+      ylab("Total Rainfall (cm)") +
+      ggtitle("Rainfall Estimated at Plots from Rain Gauges (November - May)")
 ```

### 2016-10-31 sem-data-development.Rmd
Add newline.

@@
 LiDAR products, variables derived from DEMs
+
 ```{r join remotely sensed data}
@@
 This means that I will need to exclude disease-related sampling from 2003 of the ecological data (NAs should fill-in if missing data when joining).
+
 ```{r examine temperature & rain data}

### 2018-01-11 climate-path-model.R
Line changes.

@@ -182,6 +182,7 @@ oak_sod$avg_lfct[oak_sod$plotid == "yahng02" & oak_sod$sample_year == 2010] <- N
 #' 
 #' I am going to model each year independently, and use a "repeated measures" approach that accounts for correlation among plots sampled during the same season by applying the sample year as a random effect. This is effectively a single model implemented to account for the same plots being sampled repeatedly, i.e. once each year.
 #' 
+#' 
 #+ repeated measures plot-level path model ####
 # pairs(na.omit(select(oak_sod, inf_oak_ct, tot_bay, tot_lfct, H.2005, H.2014, us.H.2005, us.H.2011, hrsblw14_wet, hrs1020_wet, avg_tmax_ds, avg_tmin_ds, hrs_abv25ds, rs2d_total, rs_v_total, twi15m)),
 #       lower.panel = panel.cor, diag.panel = panel.hist)

### 2018-01-11 pathogen-spillover-diss-ch.Rmd
Don't evaluate some models for document output.

@@
 
-[Conceptual Path Model](figures/conceptual-path-model-updated.png)
+![Conceptual Path Model](figures/temperature-path-model-updated.png)
 
 ## Results
+
+```{r set options, echo=F}
+# knitr::opts_chunk$set(fig.width = 7.5, fig.height = 7)
+knitr::opts_chunk$set(dpi = 300)
+knitr::opts_chunk$set(echo = F)
+knitr::opts_chunk$set(warning = F)
+knitr::opts_chunk$set(message = F)
+```
 
@@
 library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
+## Avoided this error in 1.1.2 by removing observations with NA values, which gives the same results as before, so maybe it wasn't properly dropping missing values when extracting info from the models.
 library(plyr); library(dplyr, verbose = F)
-library(lme4, verbose = F)
-library(lmerTest, verbose = F)
+library(lme4, verbose = F); library(nlme, verbose = F)
+# library(lmerTest, verbose = F)
 
@@
 ## make consolidated data frame only of variables in the path model
-oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct.log, tot_lfct.log.scl, hrs_abv25ds, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log, umca_density_ha.log.scl, H.2005, shannons2005.scl, twi15m.log, twi15m.log.scl)
+oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
+                              )
 
-oakplots.model.data <- filter(oakplots.model.data, is.na(H.2005)==F)
+oakplots.model.data <- filter(oakplots.model.data, avg_hrs1422_wet_t_t1 !="NA", H.2005 != "NA", tot_lfct != "NA")
 write.csv(oakplots.model.data, "analysis/data/oakplots-pathmodel-data.csv",
           row.names = F)
+
+oakplots.model.data.sub <- oakplots.model.data %>% 
+      filter(plotid!="hood03", plotid!="hood05", plotid!="jlsp04", plotid!="johns02", plotid!="sugar17")
+
+write.csv(oakplots.model.data.sub, "pathogen_spillover_data.csv", row.names = F)
 ```
@@
 
-```{r hours >25 path model no umca density unscaled, echo=F}
+```{r hours >25 path model no umca density unscaled, echo=F, eval=F}
 model_list2 <- list(
@@
 
-```{r hours >25 path model no umca density scaled, echo=F}
+```{r hours >25 path model no umca density scaled, echo=F, eval=F}
 model_list2.scl <- list(
@@
 ```{r hours >25 path model all oak plots unscaled, echo=F}
+# slc_nlme <- lme(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005, random = ~1|plotid, data = oakplots.model.data)
 
 model_list1 <- list(
-      shannons <- lm(H.2005 ~ twi15m.log, data = filter(oakplots.model.data, sample_year == 2006)),
+      shannons <- lm(
+            H.2005 ~ twi15m, 
+            data = filter(oakplots.model.data, sample_year == 2006)),
 
-      umca_density <- lmer(umca_density_ha.log ~ twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+      umca_density <- lmer(
+            umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), 
+            data = oakplots.model.data),
 
-      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
+                  data = oakplots.model.data),
+
+      wet_hours_1422 <- lmer(
+            avg_hrs1422_wet_t_t1.log ~ twi15m + (1|plotid) + (1|sample_year), 
+            data = oakplots.model.data),
+
+      hrs_abv25 <- lmer(
+            hrs_abv25ds ~ twi15m + (1|plotid) + (1|sample_year), 
+            data = oakplots.model.data),
+
+      oak_infection <- glmer(
+            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), 
+            data = oakplots.model.data)
+)
+
+# refit model with missing path from diversity to umca density
+model_list1_refit <- list(
+      shannons_refit <- lm(
+            H.2005 ~ twi15m, 
+            data = filter(oakplots.model.data, sample_year == 2006)),
+
+      umca_density_refit <- lmer(
+            umca_density_ha.log ~ H.2005 + twi15m + (1|plotid) + (1|sample_year), 
+            data = oakplots.model.data),
+
+      slc_refit <- lmer(
+            tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
+                  data = oakplots.model.data),
+
+      wet_hours_1422_refit <- lmer(
+            avg_hrs1422_wet_t_t1.log ~ twi15m + (1|plotid) + (1|sample_year), 
+                             data = oakplots.model.data),
+
+      hrs_abv25_refit <- lmer(
+            hrs_abv25ds ~ twi15m + (1|plotid) + (1|sample_year), 
+            data = oakplots.model.data),
+
+      oak_infection_refit <- glmer(
+            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
+            family = binomial(link = "logit"), 
+            data = oakplots.model.data)
+)
+```
+
+```{r path-model-no-lide-plots, echo=F}
+model_list1.noLIDE <- list(
+      shannons <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data.sub, sample_year == 2006)),
+
+      umca_density <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),
+
+      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),
 
@@
                                    # umca_density_ha.log + twi15m.log +
-                                   twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+                                   twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),
 
@@
                               # umca_density_ha.log + 
-                              twi15m.log + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),
 
-      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data.sub)
 )
+
 ```
 
-```{r hours >25 path model all oak plots scaled}
+```{r hours >25 path model all oak plots scaled, eval = F}
 model_list1.scl <- list(
@@
 
-In my assessment of model residuals I found no substantial violations of the assumption of homoscedacity for individual models, and no relationships between the model residuals and the predictor variable(s).
+In my assessment of model residuals I found no substantial violations of the assumption of homoscedacity for individual models (although there was some), and no relationships between the model residuals and the predictor variable(s).
 The patterning in the residual plots for some of the models is likely in part due to observations of count variables where the value is zero.
@@
 plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
-plot(shannons$model$twi15m.log.scl, resid(shannons),
+plot(shannons$model$twi15m, resid(shannons),
      xlab = "wetness index",
@@
 plot(umca_density, main = "LMM of bay laurel density residuals")
-plot(umca_density@frame$twi15m.log.scl, resid(umca_density),
+plot(umca_density@frame$twi15m, resid(umca_density),
      xlab = "wetness index",
@@
 
-plot(slc@frame$umca_density_ha.log.scl, resid(slc),
+plot(slc@frame$umca_density_ha.log, resid(slc),
      xlab = "bay laurel density",
@@
      main = "SLC model residuals vs. bay laurel density predictor")
-plot(slc@frame$hrs_abv25ds.scl, resid(slc),
+plot(slc@frame$hrs_abv25ds, resid(slc),
      xlab = "# dry-season hours >25 C",
@@
      main = "SLC model residuals vs. dry-season hours >25 C predictor")
-plot(slc@frame$avg_hrs1422_wet_t_t1.log.scl,
+plot(slc@frame$avg_hrs1422_wet_t_t1.log,
      xlab = "avg. wet-hours 14-22C",
@@
      main = "SLC model residuals vs. wet-hours predictor")
-plot(slc@frame$shannons2005.scl, resid(slc),
+plot(slc@frame$H.2005, resid(slc),
      xlab = "Shannon's diversity index",
@@
 ```{r pairs plot of slc predictor variables, echo=F}
-pairs(na.omit(select(oakplots.sub, umca_density_ha.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl, twi15m.log.scl)),
+pairs(na.omit(select(oakplots.sub, umca_density_ha.log, hrs_abv25ds, avg_hrs1422_wet_t_t1.log, H.2005, twi15m)),
       lower.panel = panel.cor, diag.panel = panel.hist,
@@
 #      main = "Wet-hours model residuals vs. bay laurel density predictor")
-plot(wet_hours_1422@frame$twi15m.log.scl, resid(wet_hours_1422),
+plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
      xlab = "bay laurel density",
@@
 plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
-plot(hrs_abv25@frame$twi15m.log.scl, resid(hrs_abv25),
+plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
      xlab = "wetness index",
@@
 plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
-plot(oak_infection@frame$tot_lfct.log.scl, resid(oak_infection),
+plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
      xlab = "symptomatic leaf count",
@@
      main = "Oak infection model residuals vs. symptomatic leaf count predictor")
-plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log.scl, resid(oak_infection),
+plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
      xlab = "avg. wet-hours 14-22C",
@@
      main = "Oak infection model residuals vs. wet-hours predictor")
-plot(oak_infection@frame$shannons2005.scl, resid(oak_infection),
+plot(oak_infection@frame$H.2005, resid(oak_infection),
      xlab = "Shannon's diversity index",
@@
 ```{r pairs plot of oak infection predictor variables, echo=F}
-pairs(na.omit(select(oakplots.sub, tot_lfct.log.scl, avg_hrs1422_wet_t_t1.log.scl, shannons2005.scl)),
+pairs(na.omit(select(oakplots.sub, tot_lfct.log, avg_hrs1422_wet_t_t1.log, H.2005)),
       lower.panel = panel.cor, diag.panel = panel.hist,
@@
         conditional = T, .progressBar = F)
-scaled_variables_model <- sem.fit(model_list1.scl, oakplots.model.data,
-          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
+
+raw_vars_refit <- sem.fit(model_list1_refit, oakplots.model.data, 
+        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
         conditional = T, .progressBar = F)
 
-scaled_variables_modela <- sem.fit(model_list1.scla, oakplots.model.data,
+## Model with all LIDE plots excluded
+# sem.fit(model_list1.noLIDE, oakplots.model.data.sub, 
+#         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+#         conditional = T, .progressBar = F)
+
+# scaled_variables_model <- sem.fit(model_list1.scl, oakplots.model.data,
 #           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
-          conditional = F, .progressBar = F)
+#           conditional = T, .progressBar = F)
+# scaled_variables_modela <- sem.fit(model_list1.scla, oakplots.model.data,
+#           # corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"),
+#           conditional = F, .progressBar = F)
 
@@
       caption = "Results of the d-sep test using raw variables.")
-kable(scaled_variables_model,
-      caption = "Results of the d-sep test using standardized variables.")
+# kable(scaled_variables_model,
+#       caption = "Results of the d-sep test using standardized variables.")
 
@@
           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
-scaled_coefficients <- sem.coefs(model_list1.scl, oakplots.model.data,
-          corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+scaled_coefficients <- sem.coefs(model_list1, oakplots.model.data,
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+          standardize = "scale")
+
+raw_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+scaled_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+          standardize = "scale")
+
+## Model with LIDE plots excluded
+# sem.coefs(model_list1.noLIDE, oakplots.model.data.sub,
+#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+# sem.coefs(model_list1.noLIDE, oakplots.model.data.sub,
+#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+#           standardize = "scale")
 
@@
 Additional factors have direct and indirect effects on the success of the pathogen spillover process as indicated by infection of oak trees.
+
+By calculating oak infection at the plot-level we aggregated any variation in susecptibility that may be present in different species or individuals.
 
@@
       geom_boxplot(aes(as.factor(sample_year), hrs_abv25ds), notch = T) + 
+      geom_hline(aes(yintercept = mean(hrs_abv25ds)), colour = "red") +
       theme_bw() +
@@
       ylab("Number of hours") +
-      ggtitle("Heat Exposure (dry-season Hours >25 C)")
+      ggtitle("Heat Exposure (dry-season Hours >25 C)") +
+      theme(panel.grid = element_blank())
 )
@@
       geom_boxplot(aes(as.factor(sample_year), hrs1422_wet), notch = T) +
+      geom_hline(aes(yintercept = mean(hrs1422_wet)), colour = "red") +
       theme_bw() +
@@
       ylab("Number of hours") +
-      ggtitle("Warm & Wet Conditions (wet-hours 14-22 C)")
+      ggtitle("Warm & Wet Conditions (wet-hours 14-22 C)") +
+      theme(panel.grid = element_blank())
 )
-
 ```
@@
 library(ggplot2)
+rainfall
 # summary(rainy_season_ppt)
@@
       geom_boxplot(aes(group = year), notch = T) +
-      geom_hline(aes(yintercept = mean(rain2d_total)),
+      geom_hline(aes(yintercept = mean(rain2d_total)), linetype = "dashed") +
+      theme_bw() +
+      xlab("Sample Year") +
+      ylab("Total Rainfall (mm)") +
+      # ggtitle("Rainfall Estimated at Plots from Rain Gauges (November - May)") +
+      theme(panel.grid = element_blank(),
+            axis.title = element_text(size = 12),
+            axis.text = element_text(size = 10))
+# ggsave("figures/rainy-season-ppt.eps", width = 6.25, height = 4.5)
+
+rainy_days_figure <- ggplot(filter(rainy_season_ppt, year > 2004), 
+                          aes(year, rain_days_v))
+rainy_days_figure +
+      geom_boxplot(aes(group = year), notch = F) +
+      geom_hline(aes(yintercept = mean(rain_days_v)),
                  colour = "red") +
@@
       xlab("Sample Year") +
-      ylab("Total Rainfall (cm)") +
-      ggtitle("Rainfall Estimated at Plots from Rain Gauges (November - May)")
+      ylab("Number of days w/precipitation") +
+      ggtitle("Rainy Days at Plots (November - May)")
 ```
 
-### Session Information
+## Session Information
 ```{r print session info, echo=FALSE}
@@
 ```
+
+## Model a subset of the data
+
+### Bay laurel density ~ Diveristy relationship
+
+```{r}
+ggplot(oakplots.model.data,
+       aes(umca_density_ha, H.2005)) +
+      geom_point() +
+      geom_smooth(method = "loess") +
+      facet_grid(sample_year~.) +
+      theme_bw()
+
+d1 <- select(oakplots.sub, plotid:inf_oak_ct, H.2005)
+d1 <- mutate(d1, tot_oak = rowSums(cbind(inf_oak_ct, uninf_oak_ct)))
+ggplot(filter(d1, sample_year == 2006), aes(log1p(tot_oak), H.2005)) +
+      geom_point() +
+      geom_smooth(method = "lm") +
+      theme_bw()
+
+```
+
+I want to examine the different matrices of the mixed effects models using a smaller datset that will make them easier to visualize.
+```{r model subset data, eval=F}
+small_data <- oakplots.model.data[1:103,]
+xtabs(~sample_year + plotid, droplevels(small_data))
+## shows a somewhat unbalanced data set
+
+small_data_model_list <- list(
+      shannons_small <- lm(H.2005 ~ twi15m, data = filter(small_data, sample_year == 2006)),# this model avoids pseudoreplication
+
+      umca_density_small <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = small_data),
+
+      slc_small <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = small_data),
+
+      wet_hours_1422_small <- lmer(avg_hrs1422_wet_t_t1.log ~ 
+                                   # umca_density_ha.log + twi15m.log +
+                                   twi15m + (1|plotid) + (1|sample_year), data = small_data),
+
+      hrs_abv25_small <- lmer(hrs_abv25ds ~ 
+                              # umca_density_ha.log + 
+                              twi15m + (1|plotid) + (1|sample_year), data = small_data),
+
+      oak_infection_small <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = small_data)
+)
+
+summary(umca_density_small)
+summary(slc_small)
+summary(hrs_abv25_small)
+anova(umca_density,update(umca_density, .~ 1+. -twi15m))
+
+getME(umca_density_small, "X")# fixed-effects model matrix
+getME(umca_density_small, "Z")# random-effects model matrix
+getME(umca_density_small, "Zt")# transposed random-effects model matrix, diagonal values = 1
+# image(getME(umca_density_small, "Z"))#
+getME(umca_density_small, "Ztlist")# list of components of Zt separated by individual variance component
+getME(umca_density_small, "mmList")# list of raw model matrices associated w/random effects terms
+getME(umca_density_small, "u")# conditional mode of "spherical" random effects variable
+getME(umca_density_small, "b")# conditional mode of random effects variable
+getME(umca_density_small, "Gp")
+
+image(getME(umca_density_small, "L"))# sparse Cholesky factor matrix of the penalized random-effects model
+image(getME(umca_density_small, "Lambda"))# relative covariance factor of the random effects
+image(getME(umca_density_small, "Lambdat"))# cross-product of random effects model matrix, diagonal values = random-effects parameter estimates
+
+getME(umca_density_small, "A")# scaled sparse model matrix
+getME(umca_density_small, "RX")# Cholesky factor for fixed-effects parameters
+getME(umca_density_small, "RZX")#cross-term in the full Cholesky factor
+
+getME(umca_density_small, "sigma")# residual standard error
+
+getME(umca_density_small, "theta")# random-effects parameter estimates
+
+getME(umca_density_small, "q")# number of coluns of RE model matrix Z
+getME(umca_density_small, "l_i")# number of levels of grouping factors
+getME(umca_density_small, "k")# number of RE terms
+getME(umca_density_small, "m_i")# number of covariance parameters in each RE term
+getME(umca_density_small, "devcomp")# list of items describing fitted model
+
+# image(getME(oak_infection, "L"))
+```

### 2018-01-11 analysis/explore_data.R
Change lines.

@@ -63,6 +63,7 @@ summary(stems$species) # 5 host species listed, so this appears legit
 summary(stems$status) # There are a lot of different `X` statuses, which I am not particularly concerned about. These will all just be treated as missing. However, there are 26 stems that appear to have a `blank` status type. 
 class(stems$status)
 View(filter(stems, status == ""))
+
 ## A good number of these stems needed an `X` status, but there were at least 8 that needed correcting in the database. The primary cause for the blank status appeared to be a duplication of the visit date for a single stem. Once one of these records was deleted the data appeared. So, I am going to export the query from the data base again and reassess the data with these corrections.
 stems <- read.csv("analysis/data/stem_summary_qry2_04-14.csv")
 summary(stems)
@@ -220,7 +221,7 @@ stems <- read.csv("analysis/data/all_stems_corrected.csv")
 library(dplyr)
 
 plots_umca <- stems %>%
-      select(plot, Date, species, slc, year, status) %>%
+      select(plot, date, species, slc, year, status) %>%
       group_by(plot, year, species) %>%
       filter(species == "UMCA", status == "Alive") %>%
       summarise(uninfected_bay_ct = length(which(slc==0)), infected_bay_ct = length(which(slc > 0)), tot_bay = length(species))
@@ -234,7 +235,7 @@ length(which(plots_umca$ct_bay_NA > 0))
 # filter(plots_umca, infected == 0)
 summary(filter(plots_umca, infected_bay_ct == 0))# 49 plots, max year = 2010
 
-write.csv(plots_umca, "analysis/data/plots_umca_infection.csv")
+# write.csv(plots_umca, "analysis/data/plots_umca_infection.csv")
 # I sent this file to Francesco for informing the spread model
 
 
@@ -253,7 +254,7 @@ spread(plot_qusp, key = species, value = infected_oak_ct, drop = TRUE)
 
 gather(plot_qusp, inf_status, count, ends_with("ct")) %>% 
       unite(species, inf_status) %>% 
-      spread(inf_status, )
+      spread(inf_status)
 
 library(reshape2)
 tmp <- plot_qusp %>%

### 2018-01-11 analysis/repeated_measures.R

@@ -103,10 +103,12 @@ plot(m2)
 plot(m2a)
 plot(m2b)
 plot(m2c)
+
 #' All of them show a lot of spread - expected because I know the distributions for leaf count and density aren't even close to normal
 par(mfrow = c(2,2))
 hist(plot_umca$tot_slc)
 hist(plot_umca$live_dens)
+
 #' Log transformation improves the distributions, but with some zeroes for symptomatic leaf count I need to add a constant. The `log1p` function computes log(1+x) accurately for |x| much less than one (automatically adds the constant 1) 
 log_tot_slc <- log1p(plot_umca$tot_slc); hist(log_tot_slc)
 log_live_dens <- log(plot_umca$live_dens); hist(log_live_dens)

### 2018-09-28 pathogen-spillover-diss-ch.Rmd
Revisit modeling and make model output figures.

@@
 author: "Whalen Dillon"
-date: "July 18, 2016"
+date: "`r Sys.Date()"
 output:
@@
     self_contained: no
-    toc: yes
+    toc: true
+    toc_float: true
   pdf_document: default
@@
 
-![Conceptual Path Model](figures/temperature-path-model-updated.png)
+![Conceptual Path Model](figures/Fig2-conceptual-path-model-updated.png)
 
@@
 load("climate_path_model_data.RData")
+
+# devtools::install_github("jslefche/piecewiseSEM@devel", build_vignette = TRUE)
 library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
@@
 library(lme4, verbose = F); library(nlme, verbose = F)
+library(ggplot2)
+library(viridisLite); library(viridisLite)
 # library(lmerTest, verbose = F)
@@
 ## make consolidated data frame only of variables in the path model
-oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
+oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, inf_oak_ct, uninf_oak_ct, tot_bay, umca_density_ha, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
                               )
@@
 
+# Exclude LIDE plots
 oakplots.model.data.sub <- oakplots.model.data %>% 
@@
 
-write.csv(oakplots.model.data.sub, "pathogen_spillover_data.csv", row.names = F)
+# write.csv(oakplots.model.data.sub, "pathogen_spillover_data.csv", row.names = F)
+
 ```
@@
 
-```{r hours >25 path model all oak plots unscaled, echo=F}
+```{r hours >25 path model all oak plots unscaled, echo=F, eval=F}
 # slc_nlme <- lme(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005, random = ~1|plotid, data = oakplots.model.data)
@@
             H.2005 ~ twi15m, 
-            data = filter(oakplots.model.data, sample_year == 2006)),
+            data = filter(oakplots.model.data, sample_year==2006)),
 
@@
 )
+```
+
+Add some re-calcualted variables and explore the bay density~diversity relationship.
+```{r re-examine data}
 
 # refit model with missing path from diversity to umca density
-model_list1_refit <- list(
-      shannons_refit <- lm(
-            H.2005 ~ twi15m, 
-            data = filter(oakplots.model.data, sample_year == 2006)),
+oaks.data <- oakplots.model.data %>% 
+      rename(hrs25 = hrs_abv25ds, 
+             hrs1422 = avg_hrs1422_wet_t_t1.log, 
+             lfct = tot_lfct.log, 
+             umcadens = umca_density_ha.log, 
+             h2005 = H.2005, 
+             oakinf = oak.inf, 
+             year = sample_year)
 
-      umca_density_refit <- lmer(
-            umca_density_ha.log ~ H.2005 + twi15m + (1|plotid) + (1|sample_year), 
-            data = oakplots.model.data),
+oaks.data <- oaks.data %>% 
+      mutate(tot_bay.ln = log(tot_bay),
+             tot_bay.ln1 = log1p(tot_bay),
+             tot_bay0.1 = ifelse(tot_bay==0, 0.1, tot_bay),
+             tot_bay.ln.1 = log(tot_bay0.1)
+             )
 
-      slc_refit <- lmer(
-            tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
-                  data = oakplots.model.data),
+GGally::ggpairs(select(oaks.data, h2005, twi15m, tot_bay, tot_bay.ln, tot_bay.ln1))
 
-      wet_hours_1422_refit <- lmer(
-            avg_hrs1422_wet_t_t1.log ~ twi15m + (1|plotid) + (1|sample_year), 
-                             data = oakplots.model.data),
+ggplot(oaks.data, 
+       aes(h2005, tot_bay)) +
+      geom_point() +
+      geom_smooth(method = "lm") +
+      scale_fill_viridis_d() +
+      scale_color_viridis_d() +
+      # geom_density(alpha = .8) +
+      theme_classic() +
+      NULL
 
-      hrs_abv25_refit <- lmer(
-            hrs_abv25ds ~ twi15m + (1|plotid) + (1|sample_year), 
-            data = oakplots.model.data),
+ggplot(oaks.data, aes(tot_bay)) +
+      geom_density() +
+      theme_classic()
+ggplot(oaks.data, aes(tot_bay.ln)) +
+      geom_density() +
+      geom_density(aes(tot_bay.ln1), color = "blue") +
+      theme_classic()
 
-      oak_infection_refit <- glmer(
-            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
-            family = binomial(link = "logit"), 
-            data = oakplots.model.data)
-)
+ggplot(oaks.data) +
+      geom_density(aes(h2005)) +
+      theme_classic()
+
+oaks.data %>% 
+      # filter(tot_bay>0) %>% 
+      filter(year==2006) %>%
+ggplot(aes(h2005, tot_bay.ln)) +
+      geom_point() +
+      geom_smooth(method = "lm") +
+      geom_smooth(aes(y=tot_bay.ln1), method = "lm", color = "red") +
+      # geom_smooth(aes(y=tot_bay.ln.001), method = "lm", color = "tan4") +
+      # geom_density(alpha = .8) +
+      theme_classic() +
+      NULL
+## DON'T ADD 1 & LOG-TRANSFORM
+summary(oaks.data)
+filter(oaks.data, is.infinite(tot_bay.ln))$tot_bay
+oaks.data$tot_bay.ln[is.infinite(oaks.data$tot_bay.ln)] <- 0
+
+m1 <- lmer(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
+           data = oaks.data, na.action = na.omit)
+m1a <- lmer(tot_bay.ln ~ h2005 + twi15m + (1|plotid) + (1|year),
+           data = oaks.data, na.action = na.omit)
+# anova(m1,m1a)
+summary(m1)
+plot(m1)
+# plot(m1a)
+# summary(m1a)
+
+m2 <- mgcv::gamm(tot_bay ~ s(h2005) + twi15m, random=list(year=~1, plotid=~1), data = oaks.data, na.action = na.omit)
+summary(m1)
+summary(m2[[1]])
+summary(m2[[2]])
+plot(m2[[1]])
+plot(m2[[2]])
+mgcv::vis.gam(m2)
+m3 <- gamm4::gamm4(tot_bay ~ s(h2005) + twi15m, random = ~(1|plotid)+(1|year), data = oaks.data, na.action = na.omit)
+summary(m3[[1]])
+plot(m3[[1]])
+summary(m3[[2]])
+plot(m3[[2]])
+
+m4 <- glmer(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
+           data = oaks.data, na.action = na.omit, family = poisson)
+# par(mfrow=c(2,1))
+plot(m1); plot(m4)
+summary(m1)
+summary(m4)
+anova(m1, m4)
+
+m5 <- glmer.nb(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
+           data = oaks.data, na.action = na.omit)
+summary(m5)
+plot(m5)
+
+dat <- filter(oaks.data, year==2006)
+m6 <- glm(tot_bay ~ h2005 + twi15m, data = dat, family = quasipoisson, na.action = na.fail)
+summary(m6)
+# plot(m6)
+(c_hat <- deviance(m6) / df.residual(m6))
+library(MuMIn)
+dredge(m6, rank = "QAIC", chat = c_hat)
+dredge(m6, rank = "AIC")
+hacked.quasipoisson <- function(...) {
+    res <- quasipoisson(...)
+    res$aic <- poisson(...)$aic
+    res
+}
+QAIC(update(m6, family = hacked.quasipoisson, chat = c_hat)
+
+# library(gamlss)
+# m7 <- gamlss(tot_bay ~ h2005 + twi15m, data = oaks.data)
+# summary(m7)
+# plot(m7)
+
+# GGally::ggpairs(select(oaks.data, tot_lfct,tot_bay,h2005))
+m1 <- glmer(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), family = poisson, data = oaks.data, na.action = na.omit)
+summary(m1)
+m2 <- glmer.nb(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit)
+summary(m2)
+anova(m2,lfctmod)
+plot(lfctmod)
 ```
 
-```{r path-model-no-lide-plots, echo=F}
+```{r refit model with H bay density path, eval=F}
+# The refit model has path between H and UMCA density ####
+model_list1_refit <- list(
+      shannonmod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit),
+
+      densmod <- lmer(umcadens ~ h2005 + twi15m + (1|plotid) + (1|year),
+           data = oaks.data, na.action = na.omit),
+
+      lfctmod <- lmer(lfct ~ umcadens + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+      wetmod <- lmer(hrs1422 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+      hotmod <- lmer(hrs25 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+      # hrs25 %~~% hrs1422,
+
+      infmod <- glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)
+)
+summary(lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit))
+summary(densmod)
+summary(lfctmod)
+summary(wetmod)
+summary(hotmod)
+summary(infmod)
+
+# summary(oaks.data)
+
+# ?psem I get errors with row matching when trying to use the new fucntion.
+I think it may have more to do with the new version of R and how that affects the internal sub-modeling when there are complex mixed effects models.
+
+```
+
+```{r Update model list}
+
+# Updated SEM uses untransformed UMCA variables ####
+# lfct is the ln(tot_lfct+1)
+      # lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+m_list_update <- list(
+
+            shannon.mod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit),
+      # changed the bay response to be log-transformed 
+      umca.mod <- lmer(tot_bay.ln1 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+      lfct.mod <- lmer(lfct ~ tot_bay.ln1 + hrs25 + hrs1422 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+      wet.mod <- lmer(hrs1422 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+      hot.mod <- lmer(hrs25 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+      # hrs25 %~~% hrs1422,
+
+      inf.mod <- glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)
+
+)
+
+# ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
+#       geom_point() +
+#       geom_smooth(method = "lm", color = "red") +
+#       geom_smooth() +
+#       theme_classic()
+
+```
+
+```{r path-model-no-lide-plots, echo=F, eval=F}
 model_list1.noLIDE <- list(
@@
 **Residual plots of Shannon's Diversity Index model**
-```{r shannons residuals plot, echo=F}
-plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
-plot(shannons$model$twi15m, resid(shannons),
-     xlab = "wetness index",
-     ylab = "model residuals",
-     main = "Diversity model residuals vs. wetness index predictor")
+```{r shannon.mod plot, echo=F}
+
+library(sjPlot)
+plot_model(shannon.mod, type = "diag", show.values = T, grid = TRUE)
+plot_model(shannon.mod, type = "slope", show.values = T)
+plot_model(shannon.mod, type = "resid", show.values = T)
+plot_model(shannon.mod, type = "std", show.values = T, 
+           title = "Standardized coefficient estimate")
+S.plot <- plot_model(shannon.mod, type = "slope")
+
+# line_color <- "black"
+# S.plot$data
+# str(S.plot)
+
+(shannon.plot <- ggplot(oaks.data, aes(twi15m, h2005)) +
+      geom_point(alpha = .7) +
+      # geom_smooth(method = "loess", se = F, color = "gray50", linetype = "dashed") +
+      geom_smooth(method = "lm", color = "black") +
+      theme_classic() +
+      xlab("Topographic index") +
+      ylab("Tree diversity (H')") +
+      theme(strip.background = element_blank(),
+            strip.text = element_blank()) +
+      NULL
+      )
+
+# plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
+# plot(shannons$model$twi15m, resid(shannons),
+#      xlab = "wetness index",
+#      ylab = "model residuals",
+#      main = "Diversity model residuals vs. wetness index predictor")
+
 ```
@@
 **Residual plots of bay laurel density model**
-```{r bay laurel density residuals plot, echo=F}
-plot(umca_density, main = "LMM of bay laurel density residuals")
-plot(umca_density@frame$twi15m, resid(umca_density),
-     xlab = "wetness index",
-     ylab = "model residuals",
-     main = "Bay laurel density model residuals vs. wetness index predictor")
+```{r umca.mod plot, echo=F}
+
+plot_model(umca.mod, type = "diag")
+plot_model(umca.mod, show.values = T, type = "est", 
+           title = "Coefficient estimates - UMCA density")
+plot_model(umca.mod, show.values = T, type = "std", 
+           title = "Standardized coefficient estimates - UMCA density")
+plot_model(umca.mod, show.values = T, type = "re")
+
+# umca_vs_h2005 <- plot_model(umca.mod, show.values = T, type = "pred", terms = c("twi15m"), pred.type = "fe")
+# umca_vs_h2005 +
+#       theme_classic()
+
+# oaks.data %>% 
+#       filter() %>% 
+#       ggplot(aes(h2005, log1p(tot_bay))) +
+#       geom_point() +
+#       geom_smooth(method = "lm") +
+      # facet_grid(.~year)
+# qplot(x=h2005, y=tot_bay, data = oaks.data)
+
+# library(ggeffects)
+# umca.mod_predict <- ggpredict(model = umca.mod.update, terms = "h2005", type = "re", full.data = T)
+# ggplot(umca.mod_predict, aes(x, predicted)) +
+#       # data = umca.mod.update@frame,
+#       geom_point(aes(x, predicted), show.legend = F) +
+#       # geom_smooth(se = F) +
+#       geom_smooth(method = "lm") +
+#       # facet_grid(.~year) +
+#       xlab("Shannon's diversity index (H')") +
+#       ylab("Bay laurel density (stems/ha)") +
+#       theme_classic() +
+#       NULL
+
+# oaks.data$h2005.sq <- oaks.data$h2005^2
+# oaks.data$tot_bay.log <- log1p(oaks.data$tot_bay)
+# umca.mod.update <- lmer(tot_bay.log ~ h2005 + twi15m + (1 | plotid) + (1 | year), data = oaks.data)
+# plot(umca.mod)
+# plot_model(umca.mod.update, show.values = T, type = "resid")
+# plot_model(umca.mod.update, show.values = T, type = "diag")
+
+# umca_vs_twi15m <- plot_model(densmod, show.values = T, type = "eff", terms = c("twi15m"), pred.type = "re")
+
+# plot(umca_density, main = "LMM of bay laurel density residuals")
+# plot(umca_density@frame$twi15m, resid(umca_density),
+#      xlab = "wetness index",
+#      ylab = "model residuals",
+#      main = "Bay laurel density model residuals vs. wetness index predictor")
+
 ```
@@
 **Residual plots of symptomatic leaf count model**
-```{r symptomatic leaf count residuals plot, echo=F}
-plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
+```{r lfct.mod plot, echo=F}
+
+# diagnostic plots
+plot_model(lfct.mod, type = "diag")
+plot_model(lfct.mod, type = "resid")
+
+# estimate plots
+plot_model(lfct.mod, show.values = T, type = "est", 
+           title = "Coefficient estimates - leaf count")
+plot_model(lfct.mod, show.values = T, type = "std", 
+           title = "Standardized coefficient estimates - leaf count")
+plot_model(lfct.mod, show.values = T, type = "re")
+
+# effect plots
+plot_model(lfct.mod, type = "slope") +
+      theme_bw() +
+      ylab("log (leaf count + 1)") +
+      theme(strip.background = element_blank()) +
+      NULL
+
+plot_model(lfct.mod, type = "pred", terms = c("hrs25", "tot_bay.ln1")) +
+      theme_classic() +
+      ylab("log( leaf count + 1 )") +
+      xlab("Hours >25 ºC") +
+      ggtitle("Predicted values") +
+      NULL
+
+slc_vs_bay_plot <- ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
+      geom_point(alpha = .7) +
+      geom_smooth(method = "lm", color = "black") +
+      xlab("log( Bay laurel count )") +
+      ylab("log( Symptomatic leaf count )") +
+      theme_classic()
+slc_vs_hrs25_plot <- ggplot(oaks.data, aes(hrs25, lfct)) +
+      geom_point(alpha = .7) +
+      geom_smooth(method = "lm", color = "black") +
+      xlab("Number of hours >25 ºC") +
+      ylab("log( Symptomatic leaf count )") +
+      theme_classic()
+
+cowplot::plot_grid(slc_vs_bay_plot, slc_vs_hrs25_plot, nrow = 2)
+
+# plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
 # qqnorm(resid(slc), main="Q-Q plot for conditional residuals - SLC")
@@
 
-plot(slc@frame$umca_density_ha.log, resid(slc),
-     xlab = "bay laurel density",
-     ylab = "model residuals",
-     main = "SLC model residuals vs. bay laurel density predictor")
-plot(slc@frame$hrs_abv25ds, resid(slc),
-     xlab = "# dry-season hours >25 C",
-     ylab = "model residuals",
-     main = "SLC model residuals vs. dry-season hours >25 C predictor")
-plot(slc@frame$avg_hrs1422_wet_t_t1.log,
-     xlab = "avg. wet-hours 14-22C",
-     ylab = "model residuals",
-     main = "SLC model residuals vs. wet-hours predictor")
-plot(slc@frame$H.2005, resid(slc),
-     xlab = "Shannon's diversity index",
-     ylab = "model residuals",
-     main = "SLC model residuals vs. Shannon's diversity predictor")
+# plot(slc@frame$umca_density_ha.log, resid(slc),
+#      xlab = "bay laurel density",
+#      ylab = "model residuals",
+#      main = "SLC model residuals vs. bay laurel density predictor")
+# plot(slc@frame$hrs_abv25ds, resid(slc),
+#      xlab = "# dry-season hours >25 C",
+#      ylab = "model residuals",
+#      main = "SLC model residuals vs. dry-season hours >25 C predictor")
+# plot(slc@frame$avg_hrs1422_wet_t_t1.log,
+#      xlab = "avg. wet-hours 14-22C",
+#      ylab = "model residuals",
+#      main = "SLC model residuals vs. wet-hours predictor")
+# plot(slc@frame$H.2005, resid(slc),
+#      xlab = "Shannon's diversity index",
+#      ylab = "model residuals",
+#      main = "SLC model residuals vs. Shannon's diversity predictor")
 # plot(slc@frame$twi15m.log.scl, resid(slc),
@@
 #      main = "SLC model residuals vs. wetness index predictor")
+
 ```
-```{r pairs plot of slc predictor variables, echo=F}
+
+```{r pairs plot of slc predictor variables, echo=F, eval=F}
+
 pairs(na.omit(select(oakplots.sub, umca_density_ha.log, hrs_abv25ds, avg_hrs1422_wet_t_t1.log, H.2005, twi15m)),
@@
       main = "Predictor variables for symptomatic leaf count models")
+
 ```
@@
 **Residual plots of average number of wet-hours model**
-```{r wet-hours residuals plot, echo=F}
-plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
+```{r wet.mod plot, echo=F}
+# diagnostic plots
+plot_model(wet.mod, type = "diag")
+plot_model(wet.mod, type = "resid")
+
+# estimate plots
+plot_model(wet.mod, show.values = T, type = "est", 
+           title = "Coefficient estimates - wet hours")
+plot_model(wet.mod, show.values = T, type = "std", 
+           title = "Standardized coefficient estimates - wet hours")
+plot_model(wet.mod, show.values = T, type = "re")
+
+# effect plots
+plot_model(wet.mod, type = "slope") +
+      theme_bw() +
+      # ylab("log (leaf count + 1)") +
+      theme(strip.background = element_blank()) +
+      NULL
+
+# plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
 # plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
@@
 #      main = "Wet-hours model residuals vs. bay laurel density predictor")
-plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
-     xlab = "bay laurel density",
-     ylab = "model residuals",
-     main = "Wet-hours model residuals vs. bay laurel density predictor")
+# plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
+#      xlab = "bay laurel density",
+#      ylab = "model residuals",
+#      main = "Wet-hours model residuals vs. bay laurel density predictor")
+
 ```
@@
 **Residual plots of dry-season temperature >25C model**
-```{r dry-season hours residuals plot, echo=F}
-plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
-plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
-     xlab = "wetness index",
-     ylab = "model residuals",
-     main = "Dry-season hours model residuals vs. wetness index predictor")
+```{r hot.mod plot, echo=F}
+
+# diagnostic plots
+plot_model(hot.mod, type = "diag")
+plot_model(hot.mod, type = "resid")
+
+# estimate plots
+plot_model(hot.mod, show.values = T, type = "est", 
+           title = "Coefficient estimates - hours >25 ºC")
+plot_model(hot.mod, show.values = T, type = "std", 
+           title = "Standardized coefficient estimates - hours >25 ºC")
+plot_model(hot.mod, show.values = T, type = "re")
+
+# effect plots
+plot_model(hot.mod, type = "slope") +
+      # geom_point() +
+      theme_classic() +
+      ylab("Number of hours >25 ºC") +
+      xlab("Topographic index") +
+      theme(strip.background = element_blank(),
+            strip.text = element_blank()) +
+      NULL
+
+ggplot(oaks.data, aes(twi15m, hrs25)) +
+      geom_point(alpha = .3) +
+      geom_smooth(method = "lm", color = "black") +
+      xlab("Topographic index") +
+      ylab("Number of hours >25 ºC") +
+      theme_classic()
+
+# plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
+# plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
+#      xlab = "wetness index",
+#      ylab = "model residuals",
+#      main = "Dry-season hours model residuals vs. wetness index predictor")
 # plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
@@
 #      main = "Dry-season hours model residuals vs. wetness index predictor")
+
 ```
 
-**Residual plots of oak infection model**
-```{r oak infection residuals plot, echo=F}
-plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
-plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
-     xlab = "symptomatic leaf count",
-     ylab = "model residuals",
-     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
-plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
-     xlab = "avg. wet-hours 14-22C",
-     ylab = "model residuals",
-     main = "Oak infection model residuals vs. wet-hours predictor")
-plot(oak_infection@frame$H.2005, resid(oak_infection),
-     xlab = "Shannon's diversity index",
-     ylab = "model residuals",
-     main = "Oak infection model residuals vs. diversity predictor")
+**Plots of oak infection model**
+```{r inf.mod plot, echo=F}
+
+plot_model(inf.mod, type = "diag")
+plot_model(inf.mod, type = "resid")
+
+cowplot::plot_grid(
+      plot_model(inf.mod, type = "est", show.values = T, 
+                 title = "Oak infection - est."),
+      plot_model(inf.mod, type = "std", show.values = T, 
+                 title = "Oak infection std. est."),
+      nrow = 2
+      )
+
+plot_model(inf.mod, type = "pred", terms = c("lfct")) +
+      # geom_point() +
+      theme_classic() +
+      xlab("log( symptomatic leaf count + 1 )") +
+      ylab("Predicted probability for oak infection") +
+      ggtitle("") +
+      NULL
+
+ggplot(oaks.data, aes(lfct, inf_oak_ct/(inf_oak_ct+uninf_oak_ct))) +
+      geom_point() +
+      geom_smooth(method = "loess", color = "black") +
+      theme_classic() +
+      NULL
+
+# plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
+# plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
+#      xlab = "symptomatic leaf count",
+#      ylab = "model residuals",
+#      main = "Oak infection model residuals vs. symptomatic leaf count predictor")
+# plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
+#      xlab = "avg. wet-hours 14-22C",
+#      ylab = "model residuals",
+#      main = "Oak infection model residuals vs. wet-hours predictor")
+# plot(oak_infection@frame$H.2005, resid(oak_infection),
+#      xlab = "Shannon's diversity index",
+#      ylab = "model residuals",
+#      main = "Oak infection model residuals vs. diversity predictor")
+
 ```
-```{r pairs plot of oak infection predictor variables, echo=F}
+
+```{r pairs plot of oak infection predictor variables, echo=F, eval=F}
+
 pairs(na.omit(select(oakplots.sub, tot_lfct.log, avg_hrs1422_wet_t_t1.log, H.2005)),
@@
       main = "Predictor variables for oak infection model")
+
 ```
+
+## Check for spatial autocorrelation
+Check oak infection and symptomatic leaf count response variables and model residuals.
+
+**Oak infection autocorrelation**
+```{r spatial autocorrelation inf.mod}
+
+plots_coords <- select(plots_shp@data, PLOT_ID, X, Y)
+names(plots_coords) <- tolower(names(plots_coords))
+plots_coords <- plots_coords %>% 
+      rename(plotid = plot_id) %>% 
+      mutate(plotid = tolower(plotid))
+
+oaks.data <- left_join(
+      oaks.data, plots_coords
+)
+summary(oaks.data)
+
+library(spBayes); library(classInt); library(RColorBrewer); library(geoR)
+library(ape); library(ncf)
+
+oakplots.coords <- as.matrix(oaks.data[,c("x","y")])
+pl.dists <- as.matrix(dist(cbind(oaks.data$x, oaks.data$y)))
+pl.dists[1:20,1:20]
+dim(pl.dists)
+# image(pl.dists)
+pl.dists.inv <- 1/pl.dists
+# image(pl.dists.inv)
+diag(pl.dists.inv) <- 0
+pl.dists.inv[is.infinite(pl.dists.inv)] <- 0
+pl.dists.inv[1:20,1:20]
+# summary(pl.dists.inv)
+
+plot.idist <- iDist(cbind(plots_coords$x, plots_coords$y))
+plot.idist[plot.idist==0] <- NA
+min(plot.idist, na.rm = T)
+max.dist <- .33*max(iDist(cbind(plots_coords$x, plots_coords$y)))
+
+oaks.data <- oaks.data %>% 
+      mutate(p.oak.inf = inf_oak_ct/(inf_oak_ct+uninf_oak_ct))
+summary(oaks.data)
+# ggplot(oakplots.model.data %>% group_by(sample_year), 
+#        aes(sample_year, p.oak.inf)) +
+#       geom_point(position = "jitter")
+
+Moran.I(oaks.data$p.oak.inf, pl.dists.inv)
+cgram.inf <- correlog(oaks.data$x, oaks.data$y, oaks.data$p.oak.inf, increment = 300, resamp = 99)
+plot(cgram.inf, main = "Oak infection response correlogram")
+# abline(h = mean(cgram.inf$correlation))
+max(cgram.inf$correlation)
+
+vario.inf <- variog(coords = oakplots.coords,
+                            data = oaks.data$p.oak.inf)
+plot(vario.inf)
+fit.vario.inf <- variofit(vario.inf)
+lines(fit.vario.inf)
+abline(h = fit.vario.inf$nugget, col = "purple")
+abline(h = fit.vario.inf$cov.pars[1], col = "darkblue")
+abline(v = -log(.05)*fit.vario.inf$cov.pars[2], col = "tan4")
+
+infmod.resids <- resid(infmod)
+Moran.I(infmod.resids, pl.dists.inv)
+cgram.inf.resids <- correlog(oaks.data$x, oaks.data$y, infmod.resids, increment = 300, resamp = 99)
+plot(cgram.inf.resids, main = "Oak infection residuals correlogram")
+
+vario.inf.resids <- variog(coords = oakplots.coords,
+                            data = infmod.resids)
+plot(vario.inf.resids)
+fit.vario.inf.resids <- variofit(vario.inf.resids)
+lines(fit.vario.inf.resids)
+abline(h = fit.vario.inf.resids$nugget, col = "purple")
+abline(h = fit.vario.inf.resids$cov.pars[1], col = "darkblue")
+abline( v = -log(.05)*fit.vario.inf.resids$cov.pars[2], col = "tan4")
+```
+
+There was minimal spatial autocorrelation in the response variable. The spatial autocorrelation that was present was dramitically reduced in the model residuals.
+
+**Symptomatic leaf count**
+```{r spatial autocorrelation lfct.mod}
+# lfct.data <- oakplots.model.data %>% 
+#       select(sample_year, tot_lfct.log) %>% 
+#       arrange(sample_year)
+# acf(lfct.data$tot_lfct.log)
+
+# qplot(x = sample_year, y = tot_lfct, data = oakplots.model.data)
+
+Moran.I(oaks.data$lfct, pl.dists.inv)
+cgram.lfct <- correlog(oaks.data$x, oaks.data$y, oaks.data$tot_lfct.log, increment = 300, resamp = 99)
+plot(cgram.lfct, main = "Leaf count response correlogram")
+
+r.lfctmod <- resid(lfctmod)
+Moran.I(r.lfctmod, pl.dists.inv)
+cgram.lfct.resids <- correlog(oaks.data$x, oaks.data$y, r.lfctmod, increment = 300, resamp = 99)
+plot(cgram.lfct.resids, main = "Leaf count residuals correlogram")
+
+```
+
+There was minimal spatial autocorrelation in the response variable. The spatial autocorrelation that was present was dramitically reduced in the model residuals.
 
@@
 ```{r fit and compare all oak plots, echo=F, message=F, warning=F}
-raw_variables_model <- sem.fit(model_list1, oakplots.model.data, 
-        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-        conditional = T, .progressBar = F)
+# raw_variables_model <- sem.fit(model_list1, oakplots.model.data, 
+#         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+#         conditional = F, .progressBar = F)
+# 
+# raw_vars_refit <- sem.fit(
+#       model_list1_refit, oaks.data, corr.errors = c("hrs25 ~~ hrs1422"),
+#       conditional = F, .progressBar = F)
 
-raw_vars_refit <- sem.fit(model_list1_refit, oakplots.model.data, 
-        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-        conditional = T, .progressBar = F)
+raw_vars_update <- sem.fit(
+      m_list_update, oaks.data,
+      corr.errors = c("hrs25 ~~ hrs1422"),
+      conditional = F, .progressBar = F
+)
 
-## Model with all LIDE plots excluded
+## Fit SEM with all LIDE plots excluded ####
 # sem.fit(model_list1.noLIDE, oakplots.model.data.sub, 
@@
 #           .progressBar = F)
+
 ```
@@
 ```{r tables of d-sep test results from sem.fit, echo=F}
-kable(raw_variables_model, digits = 4,
+
+# kable(raw_variables_model, digits = 4,
+      # caption = "Results of the d-sep test using raw variables.")
+
+# kable(raw_vars_refit, digits = 4,
+      # caption = "Results of the d-sep test using raw variables.")
+
+kable(raw_vars_update, digits = 3,
       caption = "Results of the d-sep test using raw variables.")
+
 # kable(scaled_variables_model,
@@
 #       caption = "Results of the d-sep test using standardized variables - no UMCA density.")
+
 ```
@@
 ```{r get path coefficients, echo=F}
-raw_coefficients <- sem.coefs(model_list1, oakplots.model.data,
-          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
-scaled_coefficients <- sem.coefs(model_list1, oakplots.model.data,
-          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-          standardize = "scale")
 
-raw_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
-          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
-scaled_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
-          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-          standardize = "scale")
+# raw_coefficients <- sem.coefs(model_list1, oakplots.model.data,
+#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+# scaled_coefficients <- sem.coefs(model_list1, oakplots.model.data,
+#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+#           standardize = "scale")
+
+# raw_coeff_refit <- sem.coefs(
+#       model_list1_refit, oaks.data
+#       )
+# scaled_coeff_refit <- sem.coefs(
+#       model_list1_refit, oaks.data,
+#       standardize = "scale"
+#       )
+
+raw_coeff_update <- sem.coefs(
+      m_list_update, oaks.data
+      )
+scaled_coeff_update <- sem.coefs(
+      m_list_update, oaks.data,
+      standardize = "scale"
+      )
 
@@
 #           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
+
 ```
@@
 ```{r tables of path coefficients, echo=F}
-kable(raw_coefficients, digits = 4,
+
+kable(raw_coeff_update, digits = 3,
       caption = "Raw path coefficients")
+# kable(raw_coeff_refit, digits = 3,
+#       caption = "Raw path coefficients")
 # kable(raw_coefficients_noUMCA,
@@
 
-kable(scaled_coefficients, digits = 4,
+kable(scaled_coeff_update, digits = 3,
       caption = "Standardized path coefficients")
+# kable(scaled_coeff_refit, digits = 2,
+#       caption = "Standardized path coefficients")
 # kable(scaled_coefficients_noUMCA,
@@
 ```{r disease figures, echo=F, warning=F}
-library(ggplot2)
+
 temperatures <- left_join(

### 2018-10-08 pathogen-spillover-diss-ch.Rmd
Revise formatting and output model plots.

@@
 ![Conceptual Path Model](figures/Fig2-conceptual-path-model-updated.png)
-
-## Results
 
@@
 
-### Path model 
+# Path modeling 
 This is the list of models that make up the path model to be fit using the methods implemented in the 'piecewiseSEM' packag.
@@
 Add some re-calcualted variables and explore the bay density~diversity relationship.
-```{r re-examine data}
-
-# refit model with missing path from diversity to umca density
+
+```{r revise data names}
 oaks.data <- oakplots.model.data %>% 
@@
              )
-
-GGally::ggpairs(select(oaks.data, h2005, twi15m, tot_bay, tot_bay.ln, tot_bay.ln1))
+```
+
+```{r explore diveristy density}
+
+## refit model with missing path from diversity to umca density
+
+library(GGally)
+ggpairs(select(oaks.data, h2005, twi15m, tot_bay, tot_bay.ln, tot_bay.ln1))
 
@@
       theme_classic()
+```
+
+```{r explore models, eval=F}
 
@@
 (c_hat <- deviance(m6) / df.residual(m6))
+
 library(MuMIn)
@@
 dredge(m6, rank = "AIC")
+
 hacked.quasipoisson <- function(...) {
@@
 }
+
 QAIC(update(m6, family = hacked.quasipoisson, chat = c_hat)
@@
 # GGally::ggpairs(select(oaks.data, tot_lfct,tot_bay,h2005))
+
 m1 <- glmer(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), family = poisson, data = oaks.data, na.action = na.omit)
 summary(m1)
+
 m2 <- glmer.nb(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit)
+
 summary(m2)
@@
 plot(lfctmod)
+
 ```
@@
 
-### Examining the residual plots of each piece of the path model
-
-**Residual plots of Shannon's Diversity Index model**
-```{r shannon.mod plot, echo=F}
+# Examining the residual plots of each piece of the path model
+
+## **Residual plots of Shannon's Diversity Index model**
+
+### Diagnostic plots 
+```{r shannon.mod diagnostic plots, echo=F}
 
@@
 plot_model(shannon.mod, type = "diag", show.values = T, grid = TRUE)
+plot_model(shannon.mod, type = "resid", show.values = T)
+
+```
+
+### Fitted model plots
+```{r shannon.mod fit plots}
+
 plot_model(shannon.mod, type = "slope", show.values = T)
-plot_model(shannon.mod, type = "resid", show.values = T)
 plot_model(shannon.mod, type = "std", show.values = T, 
@@
 
-**Residual plots of bay laurel density model**
-```{r umca.mod plot, echo=F}
+## **Residual plots of bay laurel density model**
+
+### Diagnostic plots
+```{r umca.mod diagnostic plots, echo=F}
 
 plot_model(umca.mod, type = "diag")
+plot_model(umca.mod, type = "resid")
+
+```
+
+### Fitted model plots
+```{r umca.mod fit plots}
+
 plot_model(umca.mod, show.values = T, type = "est", 
@@
 
-**Residual plots of symptomatic leaf count model**
+## **Residual plots of symptomatic leaf count model**
+
+### Diagnostic plots
 ```{r lfct.mod plot, echo=F}
@@
 
+```
+
+### Fitted model plots
+```{r lfct.mod fit plots}
 # estimate plots
@@
 
-**Residual plots of average number of wet-hours model**
-```{r wet.mod plot, echo=F}
+## **Residual plots of average number of wet-hours model**
+
+### Diagnostic plots
+```{r wet.mod diganostic plots, echo=F}
 # diagnostic plots
@@
 
+```
+
+### Fitted model plots
+```{r wet.mod fit plots}
 # estimate plots
@@
 
-**Residual plots of dry-season temperature >25C model**
-```{r hot.mod plot, echo=F}
+## **Residual plots of dry-season temperature >25C model**
+
+### Diagnostic plots
+```{r hot.mod diagnostic plots, echo=F}
 
@@
 
+```
+
+### Fitted model plots
+```{r hot.mod fit plots}
 # estimate plots
@@
 # effect plots
-plot_model(hot.mod, type = "slope") +
+plot_model(hot.mod, type = "slope", show.data = T) +
       # geom_point() +
@@
 
-**Plots of oak infection model**
-```{r inf.mod plot, echo=F}
+## **Plots of oak infection model**
+
+### Diagnostic plots
+```{r inf.mod diagnostic plots, echo=F}
 
@@
 
+```
+
+### Fitted model plots
+```{r inf.mod fit plots, echo=F}
+
 cowplot::plot_grid(
       plot_model(inf.mod, type = "est", show.values = T, 
-                 title = "Oak infection - est."),
+                 title = "Oak infection coefficient estimates") +
+            # geom_hline(yintercept = 1, color = "gray80",linetype = "dashed") +
+            theme_classic(),
       plot_model(inf.mod, type = "std", show.values = T, 
-                 title = "Oak infection std. est."),
+                 title = "Oak infection std. coefficient estimates",
+                 # vline.color = "gray80"
+                 ) +
+            theme_classic(),
       nrow = 2
@@
 
-plot_model(inf.mod, type = "pred", terms = c("lfct")) +
+plot_model(inf.mod, type = "pred", terms = c("lfct"), show.data = T) +
       # geom_point() +
@@
 
-ggplot(oaks.data, aes(lfct, inf_oak_ct/(inf_oak_ct+uninf_oak_ct))) +
-      geom_point() +
+ggplot(oaks.data, aes(exp(lfct), inf_oak_ct/(inf_oak_ct+uninf_oak_ct))) +
+      # geom_point(alpha = .5) +
+      stat_sum() +
+      facet_grid(year~.) +
       geom_smooth(method = "loess", color = "black") +
@@
 
-## Check for spatial autocorrelation
+# Check for spatial autocorrelation
 Check oak infection and symptomatic leaf count response variables and model residuals.
 
-**Oak infection autocorrelation**
+## **Oak infection autocorrelation**
 ```{r spatial autocorrelation inf.mod}
@@
 
-infmod.resids <- resid(infmod)
-Moran.I(infmod.resids, pl.dists.inv)
-cgram.inf.resids <- correlog(oaks.data$x, oaks.data$y, infmod.resids, increment = 300, resamp = 99)
+inf.mod.resids <- resid(inf.mod)
+Moran.I(inf.mod.resids, pl.dists.inv)
+cgram.inf.resids <- correlog(oaks.data$x, oaks.data$y, inf.mod.resids, increment = 300, resamp = 99)
 plot(cgram.inf.resids, main = "Oak infection residuals correlogram")
@@
 vario.inf.resids <- variog(coords = oakplots.coords,
-                            data = infmod.resids)
+                            data = inf.mod.resids)
 plot(vario.inf.resids)
@@
 
-**Symptomatic leaf count**
+## **Symptomatic leaf count**
 ```{r spatial autocorrelation lfct.mod}
+
 # lfct.data <- oakplots.model.data %>% 
@@
 Moran.I(oaks.data$lfct, pl.dists.inv)
-cgram.lfct <- correlog(oaks.data$x, oaks.data$y, oaks.data$tot_lfct.log, increment = 300, resamp = 99)
+cgram.lfct <- correlog(oaks.data$x, oaks.data$y, oaks.data$lfct, increment = 300, resamp = 99)
 plot(cgram.lfct, main = "Leaf count response correlogram")
 
-r.lfctmod <- resid(lfctmod)
+r.lfctmod <- resid(lfct.mod)
 Moran.I(r.lfctmod, pl.dists.inv)
@@
 
-### Fitting the path model
+# Fitting the path model
 
@@
 
-### Path coefficients
+## Path coefficients
 The path model provides direct and indirect effects of predictor variables on multiple responses, which can be combined to calculate the total effect of a predictor variable on a response.
@@
 library(ggplot2)
-rainfall
+# rainfall
 # summary(rainy_season_ppt)

### 2018-10-10 pathogen-spillover-diss-ch.Rmd
Add figure of significant relationships in path model.

@@
 author: "Whalen Dillon"
-date: "`r Sys.Date()"
+date: "`r Sys.Date()`"
 output:
@@
            title = "Standardized coefficient estimate")
-S.plot <- plot_model(shannon.mod, type = "slope")
+S.plot <- plot_model(shannon.mod, type = "pred", terms = "twi15m")
 
@@
 
-(shannon.plot <- ggplot(oaks.data, aes(twi15m, h2005)) +
-      geom_point(alpha = .7) +
+(shannon.plot <- plot_model(shannon.mod, type = "pred", terms = "twi15m") +
+      geom_point(data = oaks.data, 
+                 aes(twi15m, h2005), 
+                 alpha = .1, size = 1, position = "jitter") +
       # geom_smooth(method = "loess", se = F, color = "gray50", linetype = "dashed") +
-      geom_smooth(method = "lm", color = "black") +
+      # geom_smooth(method = "lm", color = "black") +
       theme_classic() +
@@
 # effect plots
-plot_model(lfct.mod, type = "slope") +
+plot_model(lfct.mod, type = "slope", show.data = T, colors = "blambus") +
+      # geom_point(color = "black", alpha = .1) +
+      # geom_smooth(method = "lm", color = "black", fill = "black") +
+      # geom_smooth(method = "loess", color = NA) +
       theme_bw() +
       ylab("log (leaf count + 1)") +
-      theme(strip.background = element_blank()) +
+      # theme(strip.text = element_blank()) +
       NULL
 
-plot_model(lfct.mod, type = "pred", terms = c("hrs25", "tot_bay.ln1")) +
+slc_vs_bay_plot <- plot_model(lfct.mod, type = "pred", terms = c("tot_bay.ln1"), show.data = F) +
+      geom_point(data = oaks.data, aes(x = tot_bay.ln1, y = lfct),alpha = .1, size = 1) +
       theme_classic() +
-      ylab("log( leaf count + 1 )") +
+      ylab("Symptomatic leaf count") +
+      xlab("Bay laurel density") +
+      ggtitle("Model fit") +
+      NULL
+
+slc_vs_hrs25_plot <- plot_model(lfct.mod, type = "pred", terms = c("hrs25"), show.data = F) +
+      geom_point(data = oaks.data, 
+                 aes(x = hrs25, y = lfct),
+                 alpha = .1, size = 1) +
+      theme_classic() +
+      ylab("Symptomatic leaf count") +
       xlab("Hours >25 ºC") +
-      ggtitle("Predicted values") +
+      ggtitle("Model fit") +
       NULL
 
-slc_vs_bay_plot <- ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
-      geom_point(alpha = .7) +
-      geom_smooth(method = "lm", color = "black") +
-      xlab("log( Bay laurel count )") +
-      ylab("log( Symptomatic leaf count )") +
-      theme_classic()
-slc_vs_hrs25_plot <- ggplot(oaks.data, aes(hrs25, lfct)) +
-      geom_point(alpha = .7) +
-      geom_smooth(method = "lm", color = "black") +
-      xlab("Number of hours >25 ºC") +
-      ylab("log( Symptomatic leaf count )") +
-      theme_classic()
-
-cowplot::plot_grid(slc_vs_bay_plot, slc_vs_hrs25_plot, nrow = 2)
+# ggeffects::ggpredict(lfct.mod)
+
+# slc_vs_bay_plot <- ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
+#       geom_point(alpha = .7) +
+#       geom_smooth(method = "lm", color = "black") +
+#       xlab("log( Bay laurel count )") +
+#       ylab("log( Symptomatic leaf count )") +
+#       theme_classic()
+# 
+# slc_vs_hrs25_plot <- ggplot(oaks.data, aes(hrs25, lfct)) +
+#       geom_point(alpha = .7) +
+#       geom_smooth(method = "lm", color = "black") +
+#       xlab("Number of hours >25 ºC") +
+#       ylab("log( Symptomatic leaf count )") +
+#       theme_classic()
+
+cowplot::plot_grid(slc_vs_bay_plot, 
+                   slc_vs_hrs25_plot,
+                   labels = "auto",
+                   nrow = 2)
 
@@
 
+# ggplot(oaks.data, aes(tot_bay, tot_lfct)) +
+#       geom_point() +
+#       scale_x_log10() +
+#       scale_y_log10()
+
 ```
@@
 
-ggplot(oaks.data, aes(twi15m, hrs25)) +
-      geom_point(alpha = .3) +
-      geom_smooth(method = "lm", color = "black") +
+# Fitted line with the observed data.
+hot.mod.plot <- plot_model(hot.mod, type = "pred", terms = "twi15m", show.data = F) +
+      geom_point(data = oaks.data, 
+                         aes(x = twi15m, y = hrs25),
+                         alpha = .1) +
+      theme_classic() +
+      ylab("Number of hours >25 ºC") +
+      xlab("Topographic index") +
+      theme(strip.background = element_blank(),
+            strip.text = element_blank()) +
+      NULL
+
+(hot.mod.plot2 <- ggplot(oaks.data, 
+                         aes(twi15m, hrs25,
+                             # color = as.factor(year)
+                             )
+                         ) +
+      geom_point(alpha = .1) +
+      # geom_smooth(method = "lm", se = F) +
+      geom_smooth(method = "lm", se = T, color = "black") +
+      scale_color_viridis_d() +
       xlab("Topographic index") +
       ylab("Number of hours >25 ºC") +
-      theme_classic()
+      theme_classic() +
+      NULL
+      )
 
@@
       plot_model(inf.mod, type = "est", show.values = T, 
-                 title = "Oak infection coefficient estimates") +
+                 title = "Oak infection coefficient estimates",
+                 vline.color = "gray80") +
             # geom_hline(yintercept = 1, color = "gray80",linetype = "dashed") +
@@
                  title = "Oak infection std. coefficient estimates",
-                 # vline.color = "gray80"
+                 vline.color = "gray80"
                  ) +
@@
 
-plot_model(inf.mod, type = "pred", terms = c("lfct"), show.data = T) +
+inf.lfct.plot <- plot_model(inf.mod, type = "pred", terms = c("lfct"), show.data = T) +
       # geom_point() +
       theme_classic() +
-      xlab("log( symptomatic leaf count + 1 )") +
-      ylab("Predicted probability for oak infection") +
+      xlab("Symptomatic leaf count") +
+      ylab("Infected oaks (%)") +
       ggtitle("") +
@@
 
-ggplot(oaks.data, aes(exp(lfct), inf_oak_ct/(inf_oak_ct+uninf_oak_ct))) +
-      # geom_point(alpha = .5) +
-      stat_sum() +
-      facet_grid(year~.) +
-      geom_smooth(method = "loess", color = "black") +
+# marginal effects of each significant predictor variable
+cowplot::plot_grid(
+inf.lfct.plot <- inf.lfct.plot +
+      geom_point(data = oaks.data, 
+                 aes(x = lfct, 
+                     y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)
+                     ),
+                 alpha = .1, size = 1),
+# inf.lfct.plot +
+#       geom_point(data = filter(oaks.data, inf_oak_ct/(inf_oak_ct+uninf_oak_ct) <.5), 
+#                  aes(x = lfct, 
+#                      y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)
+#                      ),
+#                  alpha = .4, size = 1),
+
+inf.h2005.plot <- plot_model(inf.mod, type = "pred", terms = c("h2005"), show.data = T) +
+      geom_point(data = oaks.data, 
+                 aes(x = h2005, 
+                     y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)
+                     ),
+                 alpha = .1, size = 1) +
+      theme_classic() +
+      xlab("Shannon's diversity index") +
+      ylab("Infected oaks (%)") +
+      ggtitle("") +
+      NULL,
+
+inf.hrs1422.plot <- plot_model(inf.mod, type = "pred", terms = c("hrs1422")) +
+      geom_point(data = oaks.data, aes(x = hrs1422, y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)), alpha = .1, size = 1) +
+      theme_classic() +
+      xlab("Hours 14-22 ºC") +
+      ylab("Infected oaks (%)") +
+      ggtitle("") +
+      NULL,
+labels = "auto"
+)
+
+ggplot(oaks.data, aes((lfct), inf_oak_ct/(inf_oak_ct+uninf_oak_ct), color = as.factor(year))) +
+      geom_point(alpha = .5) +
+      # geom_line(aes(x=lfct, y=predict(inf.mod))) +
+      geom_smooth(aes(y=predict(inf.mod, type = "response")), method = "gam", formula = y~s(x, bs="cs"), se=F) +
+      scale_color_viridis_d() +
+      # stat_sum() +
+      # facet_grid(year~.) +
+      # geom_smooth(method = "loess", color = "black") +
       theme_classic() +
       NULL
+# 
+# 
+# inf.pred <- ggeffects::ggpredict(inf.mod, full.data = T)
+# 
+# ggplot(inf.pred$lfct, aes(x, predicted)) +
+#       # geom_point(data = oaks.data, aes(x = lfct, y = )) +
+#       geom_path()
+#       
+# inf.pred$h2005
+# 
+# inf.mod.predict <- predict(inf.mod)
 
@@
       main = "Predictor variables for oak infection model")
+
+```
+
+# Combine significant relationships from fitted models
+
+```{r significant relationships figure}
+
+cowplot::plot_grid(
+      inf.lfct.plot,
+      inf.h2005.plot +
+            xlab("Tree diversity (H')"),
+      inf.hrs1422.plot,
+      slc_vs_bay_plot +
+            ggtitle(""),
+      slc_vs_hrs25_plot +
+            ggtitle(""),
+      shannon.plot +
+            ggtitle(""),
+      hot.mod.plot +
+            ggtitle(""),
+      labels = "auto", label_x = .01
+      )
 

### 2018-10-10 analysis/plotting.R
Change to 'readr' function.

@@ -2,22 +2,20 @@
 
 library(plyr)
 library(dplyr)
+library(readr)
 library(tidyr)
 library(ggplot2)
 
-umca_stem <- read.csv("SH_data/UMCA_stems_live_2004-11.csv")
+umca_stem <- read_csv("SH_data/UMCA_stems_live_2004-11.csv")
 
 head(umca_stem)
 summary(umca_stem %>% group_by(year, tag))
 summary(umca_stem %>% group_by(year))
 
-#' Some observations repeated in this file, so subset to unique
-umca_stem <- unique(umca_stem[ , 1:8] )
-
 #' Get specific quantile values for SLC for each year ####
 #' If the same stems are consistently in the upper quantile then they may be superspreaders because they support more infections compared to the population. 
 #' The `plyr` method with sample code from <http://stackoverflow.com/questions/5473537/how-to-calculate-95th-percentile-of-values-with-grouping-variable-in-r-or-excel>
-library(plyr)
+
 #Random seed
 set.seed(42)
 #Sample data
@@ -25,8 +23,9 @@ dat <- data.frame(Watershed = sample(letters[1:2], 100, TRUE), WQ = rnorm(100))
 #plyr call
 ddply(dat, "Watershed", summarise, WQ95 = quantile(WQ, .95))
 
-library(dplyr)
-dat %>% group_by(Watershed) %>% summarise(WQ95 = quantile(slc, 0.95))
+dat %>% 
+      group_by(Watershed) %>% 
+      summarise(WQ95 = quantile(WQ, 0.95))
 
 ddply(umca_stem, "year", summarise, slc95 = quantile(slc, .95))
 
@@ -71,8 +70,7 @@ p2 + geom_text(aes(label = tag, color = ))
 
 #' "Seasonality" and symptoms ####
 #' Note that this has to be done within a sample season, so it's a bad assessment of true seasonality of symptoms. Jennifer Davidson's work addressed this pretty well, so it would be a proper reference.
-library(dplyr)
-library(ggplot2)
+
 ggplot(data = stems %>% select(Date, slc, species, status, year) %>% filter(species == "UMCA" & status == "Alive") %>% group_by(Date)) + 
       geom_point(aes(x=Date,y=slc)) +
       facet_grid(year ~ .)
@@ -109,7 +107,8 @@ library(ggplot2)
[…]

### 2018-10-10 analysis/repeated_measures.R
Edit comment for clarity.

@@ -60,7 +60,7 @@ library(lme4)
 #' The random intercept model implies that there is one average curve (the population estimate) that may be shifted up or down by something normally distributed with variance d^2
 
 #' Yi = Xi * B + Zi * bi + Ei
-#'   bi ~ N(0, d^2), 
+#'   bi ~ N(0, d^2^), 
 #'   Ei ~ N(0, covar)
 
 #' The fitted lines for each group are parallel to the average curve
@@ -89,7 +89,7 @@ m2c <- lmer(tot_slc ~ live_dens + (1 | year) + (1 + live_dens | plot), data = pl
 #' Compare all models using Chi-square test of ANOVA ####
 #' Refits model(s) with maximum likelihood instead of REML
 anova(m0, m1, m2, m0a, m0b, m1a, m1b, m2a, m2b, m2c)
-#' The random slope, random intercept model with the density:year interaction and the additional random effect is the best (lowest AIC), but the high values for the metrics may indicate some misspecification.
+#' The random slope, random intercept model with the density:year interaction and the additional random effect (m2b) is the best (lowest AIC), but the high values for the metrics may indicate some misspecification.
 
 #' Plot the fitted models
 par(mfrow = c(2,2))

### 2018-10-10 pathogen-spillover-diss-ch.Rmd
Edit exported figure dimensions.

@@
 
-cowplot::plot_grid(
+(sig.rel.fig <- cowplot::plot_grid(
       inf.lfct.plot,
@@
             ggtitle(""),
+      hot.mod.plot +
+            ggtitle(""),
       shannon.plot +
             ggtitle(""),
-      hot.mod.plot +
-            ggtitle(""),
       labels = "auto", label_x = .01
       )
+)
+ggsave("figures/path-significant-relationships.png", plot = sig.rel.fig,
+       width = 7, height = 7, units = "in")
 

### 2018-11-12 Responses-to-Ecology-reviews.Rmd
Add to repo.

---
title: "Response to reviewer comments Ecology manuscript ECY18-0093"
author: "Whalen Dillon"
date: "`r Sys.Date()`"
output: word_document
#bibliography: references.bib
---

Dear Editor,

Enclosed is our substantially revised manuscript, **Direct and indirect effects of local microclimate on pathogen spillover** (ECY18-0093) for further consideration as an **Article** for publication in *Ecology*. In their major comments, Reviewer 1 noted that "*it is difficult to discern which results are novel and what is exciting or generalizable about the results from the paper as it is written.*" In their comments, Reviewer 2 asked about spatial independence of our sampling, which we did neglect to address in the original text and have remedied in this revised manuscript; and stated that "*abiotic/biotic interactions at the local scale and would be potentially very valuable information to have.*" These broad comments along with the more detailed comments provided inspired major revisions to the Introduction and Results sections to emphasize our analysis of the effects of local (plot-level) microclimate on local (plot-level) pathogen spillover.

Below is a detailed account of how we implemented the requested revisions and responses for each reviewer comment. We organized the letter below to show the reviewer's comment in bold, with emphasis on the part(s) we are addressing, followed by our Response. To assist you with assessing our revisions, we refer to paragraph order within sections and line numbering and include added changed text (italics).

Thank you again for your consideration of this revised manuscript for publication.

Sincerely,

Whalen W. Dillon and Ross K. Meentemeyer

# Revisions following reviewer comments

## Reviewer 1 Comments:

**In this paper, the authors use a path analysis on observational data collected from 2003 - 2014 to explore the abiotic and biotic divers of pathogen spillover of the Sudden Oak Death pathogen onto susceptible oak trees in California. It is an ecologically important disease, pathogen spillover is an understudied problem, and the observational study is impressively long. Despite these positive features, *it is difficult to discern which results are novel and what is exciting or generalizable about the results from the paper as it is written.* The major challenge with this paper is distinguishing the current work from previously published work in this system from the same research group. It wasn't clear to me which were novel results and which were supporting previous findings, but with a longer dataset. *The Results section is very brief, and most of the results were expected a priori from previous work in the same system.* It is very important to establish up front what is known from previous studies, what the major research gaps are, and how this study fills those gaps (i.e., how it differs from previous studies).**

Response: Few studies have presented an analysis of measurements made at epidemiologically relevant scales for each aspect of the disease triangle over a time period that captures multiple fluctuations in these factors. For example, exposure measurements (e.g. time above a threshold temperature) are rarely available at the same location as disease measurements for a decade. We lay out how this study provides new insights in our revisions to the Introduction (paragraph 2, lines 60-82, text below in response to a different comment).

While many of the relationships in our results were expected a priori from previous work we would like to point out that having prior knowledge is a strength of path analysis. A key addition of this work is testing the relationships hypothesized by previous work that in most cases depended on just a few years of data or a few experiments. We also expanded the results section to add greater detail on the relationships and added/replaced a figure (Fig. 4).

**As written, the paper does not appear to be very hypothesis-driven or focused on understanding mechanisms driving disease transmission (e.g., between lines 44 and 45 would be a good place to state them). Instead, it builds from the assumption that environmental heterogeneity is an interesting and important driver (lines 34-35) without really discussing what aspects of the environment are most likely to be important, or what is generalizable about this study. Phrases like "some key biotic and abiotic factors" (line 74) are too vague and generic to inspire much interest.**

Response: We made a major revision/addition to the Introduction where we discuss in greater what environmental aspects are important and why in the second paragraph (text below, lines 60-82) 
*Disease transmission and spillover fundamentally depends on the overlap of a pathogen, susceptible hosts, and abiotic environmental conditions that are favorable for disease. Abiotic factors such as temperature, rainfall, and physical landscape structure can all directly and/or indirectly affect the host-pathogen interactions that lead to disease. Abiotic conditions can directly affect pathogen survival especially when the pathogen is transmitted indirectly and must survive outside of the host. Landscape characteristics and changes in land use affect host density by influencing where hosts aggregate. Animal hosts may have preferred forage areas where they aggregate, while the aggregation of plant hosts is more strongly determined by landscape structure because individuals don’t move. Plant host density only increases in areas on the landscape where conditions are favorable or where conditions are made favorable by human action. The hosts and the pathogen experience the same climatic conditions but respond at different temporal scales to exposure to hot/cold or wet/dry extremes. Extreme conditions for the pathogen can reduce pathogen survival thereby reducing transmission. Extreme conditions for the host may increase its susceptibility to infection at a later time by stressing the host and compromising its immune response. Pathogen reproduction is also dependent on the host community because of differences in species’ competency for pathogen reproduction and tolerance of pathogen infection. Greater diversity may present more low-competency or non-hosts, which would reduce disease transmission through processes such as spillover. Although abiotic environmental conditions are a key component for many diseases, plant disease systems may be more regularly affected by abiotic environmental heterogeneity because the plants and their pathogens generally can’t move to change the conditions they are experiencing.*

**In lines 79-83, the authors state what they will measure and assess, but again without giving rationale for why things like topography, microclimate, and community composition would be important for pathogen spillover or hypotheses for their directional effects and direct or indirect effects. There are lots of interesting potential mechanisms here, and they could use greater explanation up front for a broad ecological audience. It's not clear why host diversity per se should be an important driver, and not just the abundance of bay laurel (the reservoir host) and oak (the susceptible host). This rationale should be explained better.**

Response we revised the last two paragraphs of the Introduction to build the rationale for why these factors are important components of pathogen spillover and explicitly lay out the hypothesized relationships and the directions of effects:

*Landscape characteristics, such as topography and geology, influence local moisture persistence, the vegetation community and, indirectly, disease dynamics. Local microclimate conditions, i.e. temperature and moisture, directly influence pathogen reproduction and may also affect host susceptibility. While the hosts and the pathogen experience the same abiotic conditions, they respond differently to extremes (too hot or too cold, too wet or too warm) because what is “extreme” for a microscopic pathogen is not the same as an “extreme” for a tree. Pathogen reproduction is also dependent on the host community because of differences in competency for pathogen reproduction and tolerance of pathogen infection.*

*We hypothesize that biotic factors, in this case the density of bay laurel and diversity of  the tree community, affect spillover more strongly than abiotic factors of microclimate and landscape topography, but that abiotic factors still significantly affect the spillover process. Specifically, as a very competent host greater bay laurel density will generate greater pathogen loads and directly drive spillover to susceptible oak species. At the same time, infection of oaks is expected to be reduced by greater tree diversity (i.e. a dilution effect) because all other tree species in the community have low competency compared to bay laurel. Greater exposure to warm and wet conditions is hypothesized to directly increase pathogen loads and disease prevalence of oaks by simultaneously creating favorable conditions for the pathogen and increasing the susceptibility of oaks. At the same time, we hypothesize that heat exposure will directly reduce pathogen load in the bay laurel canopy due to poor conditions for pathogen reproduction and increased abscission of leaves damaged by infection. Finally, we hypothesize that the landscape context (measured as the topographic index) will indirectly affect pathogen spillover by influencing bay laurel density, heat exposure, exposure to warm and wet conditions, and tree diversity. Higher values of the topographic index will correlate with greater bay laurel density, less heat exposure, more exposure to warm and wet conditions, and lower tree diversity. Our results provide insights into the relative effects of biotic and abiotic environmental heterogeneity on disease dynamics, particularly the process of pathogen spillover, which can improve predictions of disease risk and strategies for disease control.*

**In Fig. 2, it would be useful to walk through all the hypothesized mechanisms and explain the rationale for each (lines 190-191).**

Response: The hypothesized mechanisms were previously laid out in the Methods section with the calculation of each variable. During revision these were moved to the last paragraph of the Introduction (lines 148-161; text above) and we reference the figure in that paragraph.

**It is unsatisfying to see just a path diagram with estimated coefficients as the only results presented visually (Fig. 3). It would be much easier to follow the results with a direct visualization of the data and model fits (e.g., scatterplots of driver - response relationships). I realize this could be challenging with this type of path model, but seeing the data and how they align with the model is critical for gaining intuition into the results. Why are the rainfall totals shown in Fig. 4? Is this figure cited in the main text? It doesn't appear to contribute much.**

Response: We agree with the reviewer comment on Fig. 4 and replaced it with the suggested figure of driver - response curves from the model fits plotted over the data points. In the new Fig. 4 we included only significant relationships from the path model.

**The title is jargony and hard to engage with, as is the first sentence of the Abstract and lines 30-31. Why should we care about abiotic environmental heterogeneity? Why is this an interesting question to ask?**

Response: revised title, *Microclimate has a significant but relatively weak effect on pathogen spillover*

**Line 135: unnecessary comma.**

Response: fixed

**Line 172: seems kind of late to be first defining amplification and spillover.**

Response: Amplification and spillover were defined in the first paragraph of the Introduction. We made the role of amplification more clear *Reservoir hosts support high levels of pathogen reproduction (amplification) and transmission*

**Line 193: define 'confirmatory approach'.**

Response: This text was removed during revision of the Methods as it caused some confoundment with the general principle of structural equation modeling. Structural equation modeling is generally a "confirmatory approach" because it lays out a set of hypotheses in the structural model that are then tested using data. This SEM "philosophy" has been well-covered in the literature, including references in this manuscript, and a full description is beyond the scope of our research paper.

**Lines 207-211: explain why independence is important and what happens when it fails.**

Response: this text was removed during revision in the interest of length. The independence test is related to some of the assumptions of statitistical regression models where inference is the goal. The independence test of the SEM is a check on the assumption that each model has all the important explantory variables from the pool of variables that are being examined in the system. A failure of the independence test indicates one or more missing paths in the structural model, i.e. the initial set of hypotheses of the relationships in the systems isn't entirely accurate.

**Line 212: what are the response variables?**

Response: We refer to what the response variables are in Fig. 1 in the first paragraph of the revised *Statistical analyses* section: *Each box in Fig. 1 that has at least one arrow pointing to it is a response variable in the model and each of the boxes where that arrow is rooted is a predictor variable.*

**Lines 216-222: the model section is hard to follow.**

Response: We made substantial revisions to this section of the Methods, which we hope improved the flow of the text and made it easier to follow.

**Line 225: what are the random effects?**

Response: The random effects (sample year and plot) are identified and described in the second paragraph of the *Statistical analyses* section: *For variables repeatedly measured at each plot during each year we used mixed-effects models with scalar random effects for the sample year and plot, which account for this cross-replication by allowing the intercepts for these grouping factors to vary. The random effect for year accounts for correlation between measurements made during the same year, while the random effect for plot accounts for correlation between the measurements from the same plot. Accounting for this induced correlation provides more accurate estimates of confidence intervals and p-values by preventing variation from being attributed to measured variables.*

**Line 250-251: it would have been helpful for this explanation to come earlier.**

Response: We moved this sentence to the first paragraph of the *Statistical analyses* section and added to the description in the caption of Fig.1. *The curved, double-headed arrow indicates a correlative but not causal relationship between variables, and this relationship is excluded from the model fitting.*

**Lines 307-309: hard to understand this sentence.**

Response: We revised this paragraph of the Discussion and this sentence now reads, *Previous studies indicated  individual variation in bay laurel susceptibility and competency (Anacker et al. 2008, Hüberli et al. 2011) that would be obscured by aggregation to the plot level in our study.*

**Line 315: subject-verb agreement problem. Also, this point is speculative based on the data presented here.**

Response: We revised the sentence, *Our analysis using a longer data set from a geographically distinct plot network with different forest types reinforces this previous finding in this system.*

**Line 318: this dilution effect result was shown in previous work. A broader discussion of the dilution effect in this system (and why it goes beyond just the abundance of bay laurel vs. oak) would be interesting, and would frame the paper within the broader community ecology of disease context.**

Response: We added the following text to this paragraph *An interesting contrast in our results is that the dilution effect did not extend to pathogen load (symptomatic leaf count) on bay laurel. One possible reason for observing a dilution effect of diversity on oak disease prevalence but no effect on pathogen load is because the P. ramorum infection on bay laurel occurs at the scale of the leaf, while on oaks it occurs at the scale of the tree (i.e., bole-cankers). At the leaf-scale, a single bay laurel tree with thousands of leaves could support a very large pathogen load even as the only bay laurel amongst a group of susceptible oaks. In this low-diversity scenario all of the oaks nearby the bay laurel tree would be at high risk of becoming infected. Therefore, increasing tree diversity to include more species with low or no competency (essentially any species that is not bay laurel) for P. ramorum infection would likely reduce the encounter rate between dispersed spores and susceptible oak species.*

**Lines 335-337: What about rainfall changes? Seems like they would have a greater impact.**

Response: We added two sentences discussing rainfall changes to this paragraph *Additionally, changes in rainfall patterns, such as shifts in the timing and length of the wet and dry seasons and the amount of rainfall, would dramatically affect sudden oak death disease dynamics. For example, extended drought would reduce pathogen loads and stress trees, potentially increasing oak tree susceptibility to infection during epidemic outbreaks by compromising their immune response.*

**Lines 338-340: What is the evidence for this claim?**

Response: We revised this sentence to reiterate our results *While these conditions could offset each other, the relative strength of the effects of heat exposure on bay laurel pathogen load and of optimal warm and wet temperatures on oak infection in our results suggests that increasing exposure to optimal warm and wet temperatures could override the limiting influence of summer heat exposure on pathogen load.*

**Lines 343-344: This is a coin toss, indicating that the effect of drought is too variable to predict at this point.**

Response: We agree that the effect of drought is variable and difficult to predict, in large part because it is a large-scale event that has direct and indirect effects on the pathogen and the host. We expanded the discussion of this context in this paragraph: *However, only 50% of the studies on Phytophthora related diseases in this review reported positive drought-disease interactions. The findings of positive interactions were frequently attributed to indirect effects of drought on the pathogen where host susceptibility to disease was increased by drought. A possible reason for this 50/50 finding for drought interactions with Phytophthora diseases is that in some situations the cost of drought to the host is greater than the cost of drought to the pathogen and in some cases the cost is greater for the pathogen than the host. The effect of drought also depends on the size and life-history of the organism – an effective drought for a pathogen would generally be at a smaller spatial and temporal scale than a drought for a tree. Our results suggest an antagonistic interaction between drought and disease for a Phytophthora-caused disease, where greater exposure to temperatures >25 ºC during the summer dry season (June – September) reduced pathogen load in bay laurel canopy foliage and therefore spillover that causes disease on susceptible oaks; i.e. the cost of “drought” was greater for the pathogen than for the host.*

**Lines 387-389: This study doesn't make this clear. This comes off as a vacuous conclusion without much substance.**

Response: Text removed/revised.

**Table 1: What is the d-sep test? This caption does not give enough information to interpret the results presented.**

Response: We added greater detail to the caption *Results for the fitted path model indicating relationships between pairs of variables with missing path(s) and the metrics to assess the fit of the entire path model. All missing paths are statistically insignificant and the overall model fit is acceptable (p=0.409)*

## Reviewer 2 Comments

**This paper examines the relationship between biotic and abiotic variables in explaining prevalence of an economic and ecologically important pathogen. The paper is very well written, with excellent descriptions of the importance of the study, the measurements that went into the fit SEM, how the model was fit and what the implications are for the study.**

**SEM is an excellent technique for integrating a deep research program on a topic and overall, I think I agree with the analysis and the interpretation.**

**Two related points which are conceptual and technial stand out, however, and I would like the authors to address them. On the technical side, the authors have assumed independence of each spatial location.  Without knowing how far the pathogen disperses (or indeed, other spatial relationships expressed in the SEM), I cannot assess whether this assumption is valid.  Have the authors examined their residuals for spatial autocorrelation?  Is there other evidence that these sample locations should be independent?**

Response: We made a few revisions to address this point. In the Introduction we add the detail that it is a primarily wind-dispersed pathogen, *The pathogen is wind-dispersed and thought to spread primarily between individuals of the foliar host. The majority of disease spread occurs locally (<60m; Swiecki and Bernhardt 2007, Dillon et al. 2014), though long-range dispersal events are an important contribution to regional spread (Meentemeyer et al. 2011).* In the Methods we added the assessment of spatial autocorrelation using correlograms (lines 313-317): *Models were checked for spatial autocorrelation by plotting and comparing correlograms of the response variable and the residuals from the fitted models. Models were checked for spatial autocorrelation by plotting and comparing correlograms of the response variable and the residuals from the fitted models. There was some minor spatial correlation in the modeled response variables, but the model residuals showed no significant spatial autocorrelation, indicating that the models accounted for any significant spatial correlation in the response.*

**A second related issue, is that the SEM has separated the direct links between biotic and abiotic variables and prevalence. The presented information strongly suggests that all (or nearly all) oak infection is caused by spillover, so the current SEM can be interpreted as follows, I think: abiotic factors are responsible for determining long distance spillover into the patches, while biotic factors determine local spillover.**

Response: The known epidemiology of the disease indicates that infection of oaks (*Quercus*) must happen via spillover. We think that a major advantage of the SEM approach is the separation of direct linkages between variables (i.e. examining indirect effects) because it can provide insights into, and better estimates for, biological mechanisms. For example, in our model the heat exposure affects pathogen load but not disease prevalene, while exposure to warm and wet ("favorable") pathogen conditions affected disease prevalence on oaks but not pathogen load. Addression long-distance spillover of the pathogen was beyond the scope of this study, though estimates from this study could potentially be used to inform long-distance spread modeling under different future climate scenarios.

**The point I'm getting to is that it seems that there is the possibility for abiotic/biotic interactions at the local scale and this would be potentially very valuable information to have. Can the authors exclude such interactions or are they simply to difficult to fit due to undersampling?  Please discuss!**

Response: We didn't effectively communicate a key feature of this study, which is that we did measure and assess the interaction between biotic and abiotic factors at the local scale. That is, the temperature measurements were made in at each plot and then aggregated over a time period to create the threshold variables. While the temperature recorded would be responding to the broader climate pattern the observed pathogen load in the plot would be more closely responding to the local temperature variation and extremes. We made revisions in key areas throughout the manuscript to highlight that we are focused on local-scale pathogen spillover:

1. *Parsing these relationships necessitates long-term monitoring of disease dynamics under heterogeneous environmental conditions and taking measurements at epidemiologically relevant spatial and temporal scales, including interacting biotic and abiotic factors at **local scales**.*

2. *We analyzed the relative effects of biotic and abiotic factors on **local-scale** pathogen spillover using sudden oak death (SOD) as a case study of an emerging infectious disease with asymmetric host competency and transmission patterns correlated with environmental heterogeneity*

3. *By observing fluctuations of environmental factors influencing pathogen reproduction and host susceptibility over time at multiple locations across a landscape, we aimed to disentangle the relative contributions of key biotic and abiotic factors on **local-scale** pathogen spillover in this disease system.*

### 2018-11-12 pathogen-spillover-diss-ch.Rmd
Correct lfct.mod predictor and add spatial autocorrelation checks for other models.

@@
 output:
+  word_document: default
+  pdf_document: default
   html_document:
@@
     toc_float: true
-  pdf_document: default
-  word_document: default
 ---
@@
 
-      lfct.mod <- lmer(lfct ~ tot_bay.ln1 + hrs25 + hrs1422 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+      lfct.mod <- lmer(lfct ~ tot_bay.ln1 + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
 
@@
 min(plot.idist, na.rm = T)
+length(plot.idist[plot.idist<=250 & is.na(plot.idist)=="FALSE"])/2
 max.dist <- .33*max(iDist(cbind(plots_coords$x, plots_coords$y)))
@@
 
-There was minimal spatial autocorrelation in the response variable. The spatial autocorrelation that was present was dramitically reduced in the model residuals.
+There was minimal spatial autocorrelation in the response variable. The spatial autocorrelation that was present was dramatically reduced in the model residuals.
+
+## **Optimal warm-wet conditions**
+```{r spatial autocorrelation wet.mod}
+Moran.I(oaks.data$hrs1422, pl.dists.inv)
+Moran.I(resid(wet.mod), pl.dists.inv)
+
+cgram.wet.mod <- correlog(oaks.data$x, oaks.data$y, oaks.data$hrs1422, increment = 300, resamp = 99)
+plot(cgram.wet.mod)
+
+cgram.wet.mod.resids <- correlog(oaks.data$x, oaks.data$y, resid(wet.mod), increment = 300, resamp = 99)
+plot(cgram.wet.mod.resids)
+
+```
+
+## **Hot summer temperatures**
+```{r spatial autocorrelation hot.mod}
+Moran.I(oaks.data$hrs25, pl.dists.inv)
+Moran.I(resid(hot.mod), pl.dists.inv)
+
+cgram.hot.mod <- correlog(oaks.data$x, oaks.data$y, oaks.data$hrs25, increment = 300, resamp = 99)
+plot(cgram.hot.mod)
+
+cgram.hot.mod.resids <- correlog(oaks.data$x, oaks.data$y, resid(hot.mod), increment = 300, resamp = 99)
+plot(cgram.hot.mod.resids)
+
+```
+
+## UMCA density
+```{r spatial autocorrelation umca.mod}
+
+Moran.I(oaks.data$tot_bay.ln1, pl.dists.inv)
+Moran.I(resid(umca.mod), pl.dists.inv)
+
+cgram.umca.mod <- correlog(oaks.data$x, oaks.data$y, oaks.data$hrs25, increment = 300, resamp = 99)
+plot(cgram.umca.mod)
+
+cgram.umca.mod.resids <- correlog(oaks.data$x, oaks.data$y, resid(umca.mod), increment = 300, resamp = 99)
+plot(cgram.umca.mod.resids)
+
+```
+
+```{r spatial autocorrelation shannon.mod}
+
+Moran.I(oaks.data$h2005, pl.dists.inv)
+Moran.I(resid(shannon.mod), pl.dists.inv)
+
+cgram.shannon.mod <- correlog(oaks.data$x, oaks.data$y, oaks.data$h2005, increment = 300, resamp = 99)
+plot(cgram.shannon.mod)
+
+cgram.shannon.mod.resids <- correlog(oaks.data$x, oaks.data$y, resid(shannon.mod), increment = 300, resamp = 99)
+plot(cgram.shannon.mod.resids)
+
+```
 

### 2019-04-07 d-sep_test.Rmd
??

@@
 ---
-output: html_document
+output:
+  word_document: default
+  html_document: default
 ---
@@
 3.
-For each of the *k* pairs of variables (*Xi,Xj*), list the set of other variables, {*Z*} that are direct causes of either *Xi* or *Xj*.
-The pair of variables, (*Xi,Xj*), along with its conditioning set, {*Z*}, define an independence claim, (*Xi,Xj*) | {*Z*}, and the full set of the *k* independence claims defines the basis set, *Bu*.
+For each of the *k* pairs of variables (*X$_i$,X$_j$*), list the set of other variables, {*Z*} that are direct causes of either *X$_i$* or *X$_j$*.
+The pair of variables, (*X$_i$,X$_j$*), along with its conditioning set, {*Z*}, define an independence claim, (*X$_i$,X$_j$*) | {*Z*}, and the full set of the *k* independence claims defines the basis set, *B$_u$*.
 
 4.
-For each element in this basis set, obtain the _**exact**_ probability, *Pk* that the pair (*Xi,Xj*) is statistically independent conditional on the variables *Z*.
+For each element in this basis set, obtain the _**exact**_ probability, *P$_k$* that the pair (*X$_i$,X$_j$*) is statistically independent conditional on the variables *Z*.
 In other words, perform a regression model using an appropriate method.
@@
 $$ 
+6.
+Compare the value for *C* to a $\chi^2$ (chi-square) distribution with 2*k* degrees of freedom.
+If the P-value is greater than the *a priori* selected value, e.g.
+0.05, then the data provides sufficient support for the model structure.
+
+7. If the data provide sufficent support for the model structure, then each component may be assessed and coefficients interpreted.
 
@@
 ### Establish Basis Set
+
 ```{r get basis set, eval=TRUE}
@@
 Following the directions of the Shipley (2009) supplement I will fit separate mixed models for each independence claim in the basis set.
+
 ```{r print basis set, echo=TRUE}
@@
 ```
+
 Six claims equals six models. The first two variables listed in each item are claimed to be independent based on the following set of zero or more variables.
@@
 ```
+
 ```{r correlations & distributions}
@@
 I am going to check to see if that is still true.
+
 ```{r independence claim 1 test}
@@
 ```
+
 It is correct that `lmer()` does not provide degrees of freedom, and thus no probability estimate (p-value) for the coefficients.
@@
 Now to do the same for the other normally distributed dependent variables
+
 ```{r independence claim test 2 & 3}

### 2019-04-07 path-temp-compare.Rmd
Analyse SAC

@@
 library(spBayes); library(classInt); library(RColorBrewer); library(geoR)
-summary(na.omit(oakplots.2006))
-points <- select(na.omit(oakplots.2006), X, Y)
-points <- as.matrix(dist(cbind(points$X, points$Y)))
-points.inv <- 1/points
+summary(na.omit(as.data.frame(oakplots.2006)))
+points <- select(na.omit(as.data.frame(oakplots.2006)), X, Y)
+points <- as.matrix((cbind(points$X, points$Y)))
+points.inv <- 1/(iDist(points))
 diag(points.inv) <- 0
@@
       vario.shannons,
-      ini.cov.pars = c(300, 200/-log(0.5)),
+      # ini.cov.pars = c(300, 200/-log(0.5)),
       cov.model = "exponential",
@@
 summary(points_rep)
-points_rep <- as.matrix(dist(cbind(points_rep$x, points_rep$y)))
-points_rep.inv <- 1/points_rep
-points_rep.inv[points_rep.inv == "Inf"] <- 0
+points_rep <- as.matrix(cbind(points_rep$x, points_rep$y))
+points_rep.inv <- 1/(iDist(points_rep))
+points_rep.inv[points_rep.inv=="Inf"] <- 0
 summary(points_rep.inv)
@@
 vario.umca_density <- variog(coords = points_rep, 
-                     data = na.omit(umca_density@frame$tot_bay.log),
+                     data = umca_density@frame$umca_density_ha.log.scl,
                      uvec = (seq(0, max.dist, length = bins)))
@@
 summary(points_rep)
-points_rep <- as.matrix(dist(cbind(points_rep$x, points_rep$y)))
-points_rep.inv <- 1/points_rep
+points_rep <- cbind(points_rep$x, points_rep$y)
+points_rep.inv <- 1/iDist(points_rep)
 points_rep.inv[points_rep.inv == "Inf"] <- 0
-summary(points_rep.inv)
+str(points_rep.inv)
 
@@
 vario.slc <- variog(coords = points_rep, 
-                     data = na.omit(slc@frame$tot_lfct.log),
+                     data = slc@frame$tot_lfct.log,
                      uvec = (seq(0, max.dist, length = bins)))
@@
 summary(points_rep)
-points_rep <- as.matrix(dist(cbind(points_rep$x, points_rep$y)))
-points_rep.inv <- 1/points_rep
+points_rep <- cbind(points_rep$x, points_rep$y)
+points_rep.inv <- 1/iDist(points_rep)
 points_rep.inv[points_rep.inv == "Inf"] <- 0
@@
 # Plot variograms
-vario. <- variog(coords = points_rep, 
-                     data = na.omit(umca_density@frame$tot_bay.log),
+vario <- variog(coords = points_rep, 
+                     data = wet_hours_1422@frame$avg_hrs1422_wet_t_t1.log,
                      uvec = (seq(0, max.dist, length = bins)))
-fit.vario.umca_density <- variofit(
-      vario.umca_density,
+fit.vario.wet_hours <- variofit(
+      vario,
       # ini.cov.pars = c(300, 200/-log(0.5)),
@@
 par(mfrow = c(1,2))
-plot(vario.umca_density)
-lines(fit.vario.umca_density)
-abline(h = fit.vario.umca_density$nugget, col = "blue")# nugget
-abline(h = fit.vario.umca_density$cov.pars[1] + fit.vario.umca_density$nugget, col = "green")# sill
-abline(v = -log(.05) * fit.vario.umca_density$cov.pars[2], col = "red3")# range
+plot(vario)
+lines(fit.vario.wet_hours)
+abline(h = fit.vario.wet_hours$nugget, col = "blue")# nugget
+abline(h = fit.vario.wet_hours$cov.pars[1] + fit.vario.wet_hours$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.wet_hours$cov.pars[2], col = "red3")# range
 
-vario.umca_density.resids <- variog(coords = points_rep,
-                               data = umca_density.resids,
+vario.wet_hours.resids <- variog(coords = points_rep,
+                               data = wet_hours.resids,
                                uvec = (seq(0, max.dist, length = bins)))
-fit.vario.umca_density.resids <- variofit(
-      vario.umca_density.resids,
+fit.vario.wet_hours.resids <- variofit(
+      vario.wet_hours.resids,
       cov.model = "exponential",
@@
 # par(mfrow = c(1,2))
-plot(vario.umca_density.resids, 
-     main = "Variogram - UMCA deensity model residuals",
+plot(vario.wet_hours.resids, 
+     main = "Variogram - Wet-hours model residuals",
      xlab = "distance (m)")
-lines(fit.vario.umca_density.resids)
-abline(h = fit.vario.umca_density.resids$nugget, col = "blue")# nugget
-abline(h = fit.vario.umca_density.resids$cov.pars[1] + fit.vario.umca_density.resids$nugget, col = "green")# sill
-abline(v = -log(.05) * fit.vario.umca_density.resids$cov.pars[2], col = "red3")# range
+lines(fit.vario.wet_hours.resids)
+abline(h = fit.vario.wet_hours.resids$nugget, col = "blue")# nugget
+abline(h = fit.vario.wet_hours.resids$cov.pars[1] + fit.vario.wet_hours.resids$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.wet_hours.resids$cov.pars[2], col = "red3")# range
 
@@
 Moran.I(oak_infection.resids, points_rep.inv)
+
+vario.oak_infection.resids <- variog(coords = points_rep,
+                               data = oak_infection.resids,
+                               uvec = (seq(0, max.dist, length = bins)))
+fit.vario.oak_infection.resids <- variofit(
+      vario.oak_infection.resids,
+      cov.model = "exponential",
+      minimisation.function = "optim",
+      weights = "equal"
+)
+# par(mfrow = c(1,2))
+plot(vario.oak_infection.resids, 
+     main = "Variogram - Wet-hours model residuals",
+     xlab = "distance (m)")
+lines(fit.vario.oak_infection.resids)
+abline(h = fit.vario.oak_infection.resids$nugget, col = "blue")# nugget
+abline(h = fit.vario.oak_infection.resids$cov.pars[1] + fit.vario.oak_infection.resids$nugget, col = "green")# sill
+abline(v = -log(.05) * fit.vario.oak_infection.resids$cov.pars[2], col = "red3")# range
 ```

### 2019-04-07 pathogen-spillover-diss-ch.Rmd
Old update?

--- 2018-10-10 pathogen-spillover-diss-ch.Rmd
@@
 author: "Whalen Dillon"
-date: "`r Sys.Date()`"
+date: "July 18, 2016"
 output:
@@
     self_contained: no
-    toc: true
-    toc_float: true
+    toc: yes
   pdf_document: default
@@
 
-![Conceptual Path Model](figures/Fig2-conceptual-path-model-updated.png)
+[Conceptual Path Model](figures/conceptual-path-model-updated.png)
+
+## Results
 
@@
 load("climate_path_model_data.RData")
-
-# devtools::install_github("jslefche/piecewiseSEM@devel", build_vignette = TRUE)
 library(piecewiseSEM)# version 1.1.1, tried most recent version & kept getting errror 7/18/2016 b/c matrices rows were unequal in sem.missing.paths function
-## Avoided this error in 1.1.2 by removing observations with NA values, which gives the same results as before, so maybe it wasn't properly dropping missing values when extracting info from the models.
+## Avoided this error in 1.1.2 by removing observations with NA values, which gives the same results as before, so maybe it wasn't properly dropping levels when extracting info from the models.
 library(plyr); library(dplyr, verbose = F)
 library(lme4, verbose = F); library(nlme, verbose = F)
-library(ggplot2)
-library(viridisLite); library(viridisLite)
 # library(lmerTest, verbose = F)
@@
 ## make consolidated data frame only of variables in the path model
-oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, inf_oak_ct, uninf_oak_ct, tot_bay, umca_density_ha, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
+oakplots.model.data <- select(oakplots.sub, plotid, sample_year, oak.inf, tot_lfct, tot_lfct.log, hrs_abv25ds, avg_hrs1422_wet_t_t1, avg_hrs1422_wet_t_t1.log, umca_density_ha, umca_density_ha.log, H.2005, twi15m, twi15m.log#, twi15m.log.scl, tot_lfct.log.scl, hrs_abv25ds.scl, avg_hrs1422_wet_t_t1.log.scl, umca_density_ha.log.scl, shannons2005.scl
                               )
@@
           row.names = F)
-
-# Exclude LIDE plots
-oakplots.model.data.sub <- oakplots.model.data %>% 
-      filter(plotid!="hood03", plotid!="hood05", plotid!="jlsp04", plotid!="johns02", plotid!="sugar17")
-
-# write.csv(oakplots.model.data.sub, "pathogen_spillover_data.csv", row.names = F)
-
 ```
 
-# Path modeling 
+### Path model 
 This is the list of models that make up the path model to be fit using the methods implemented in the 'piecewiseSEM' packag.
@@
 
-```{r hours >25 path model all oak plots unscaled, echo=F, eval=F}
+```{r hours >25 path model all oak plots unscaled, echo=F}
 # slc_nlme <- lme(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005, random = ~1|plotid, data = oakplots.model.data)
@@
 model_list1 <- list(
-      shannons <- lm(
-            H.2005 ~ twi15m, 
-            data = filter(oakplots.model.data, sample_year==2006)),
+      shannons <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data, sample_year == 2006)),
 
-      umca_density <- lmer(
-            umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), 
-            data = oakplots.model.data),
+      umca_density <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), 
-                  data = oakplots.model.data),
-
-      wet_hours_1422 <- lmer(
-            avg_hrs1422_wet_t_t1.log ~ twi15m + (1|plotid) + (1|sample_year), 
-            data = oakplots.model.data),
-
-      hrs_abv25 <- lmer(
-            hrs_abv25ds ~ twi15m + (1|plotid) + (1|sample_year), 
-            data = oakplots.model.data),
-
-      oak_infection <- glmer(
-            oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), 
-            data = oakplots.model.data)
-)
-```
-
-Add some re-calcualted variables and explore the bay density~diversity relationship.
-
-```{r revise data names}
-oaks.data <- oakplots.model.data %>% 
-      rename(hrs25 = hrs_abv25ds, 
-             hrs1422 = avg_hrs1422_wet_t_t1.log, 
-             lfct = tot_lfct.log, 
-             umcadens = umca_density_ha.log, 
-             h2005 = H.2005, 
-             oakinf = oak.inf, 
-             year = sample_year)
-
-oaks.data <- oaks.data %>% 
-      mutate(tot_bay.ln = log(tot_bay),
-             tot_bay.ln1 = log1p(tot_bay),
-             tot_bay0.1 = ifelse(tot_bay==0, 0.1, tot_bay),
-             tot_bay.ln.1 = log(tot_bay0.1)
-             )
-```
-
-```{r explore diveristy density}
-
-## refit model with missing path from diversity to umca density
-
-library(GGally)
-ggpairs(select(oaks.data, h2005, twi15m, tot_bay, tot_bay.ln, tot_bay.ln1))
-
-ggplot(oaks.data, 
-       aes(h2005, tot_bay)) +
-      geom_point() +
-      geom_smooth(method = "lm") +
-      scale_fill_viridis_d() +
-      scale_color_viridis_d() +
-      # geom_density(alpha = .8) +
-      theme_classic() +
-      NULL
-
-ggplot(oaks.data, aes(tot_bay)) +
-      geom_density() +
-      theme_classic()
-ggplot(oaks.data, aes(tot_bay.ln)) +
-      geom_density() +
-      geom_density(aes(tot_bay.ln1), color = "blue") +
-      theme_classic()
-
-ggplot(oaks.data) +
-      geom_density(aes(h2005)) +
-      theme_classic()
-```
-
-```{r explore models, eval=F}
-
-oaks.data %>% 
-      # filter(tot_bay>0) %>% 
-      filter(year==2006) %>%
-ggplot(aes(h2005, tot_bay.ln)) +
-      geom_point() +
-      geom_smooth(method = "lm") +
-      geom_smooth(aes(y=tot_bay.ln1), method = "lm", color = "red") +
-      # geom_smooth(aes(y=tot_bay.ln.001), method = "lm", color = "tan4") +
-      # geom_density(alpha = .8) +
-      theme_classic() +
-      NULL
-## DON'T ADD 1 & LOG-TRANSFORM
-summary(oaks.data)
-filter(oaks.data, is.infinite(tot_bay.ln))$tot_bay
-oaks.data$tot_bay.ln[is.infinite(oaks.data$tot_bay.ln)] <- 0
-
-m1 <- lmer(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
-           data = oaks.data, na.action = na.omit)
-m1a <- lmer(tot_bay.ln ~ h2005 + twi15m + (1|plotid) + (1|year),
-           data = oaks.data, na.action = na.omit)
-# anova(m1,m1a)
-summary(m1)
-plot(m1)
-# plot(m1a)
-# summary(m1a)
-
-m2 <- mgcv::gamm(tot_bay ~ s(h2005) + twi15m, random=list(year=~1, plotid=~1), data = oaks.data, na.action = na.omit)
-summary(m1)
-summary(m2[[1]])
-summary(m2[[2]])
-plot(m2[[1]])
-plot(m2[[2]])
-mgcv::vis.gam(m2)
-m3 <- gamm4::gamm4(tot_bay ~ s(h2005) + twi15m, random = ~(1|plotid)+(1|year), data = oaks.data, na.action = na.omit)
-summary(m3[[1]])
-plot(m3[[1]])
-summary(m3[[2]])
-plot(m3[[2]])
-
-m4 <- glmer(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
-           data = oaks.data, na.action = na.omit, family = poisson)
-# par(mfrow=c(2,1))
-plot(m1); plot(m4)
-summary(m1)
-summary(m4)
-anova(m1, m4)
-
-m5 <- glmer.nb(tot_bay ~ h2005 + twi15m + (1|plotid) + (1|year),
-           data = oaks.data, na.action = na.omit)
-summary(m5)
-plot(m5)
-
-dat <- filter(oaks.data, year==2006)
-m6 <- glm(tot_bay ~ h2005 + twi15m, data = dat, family = quasipoisson, na.action = na.fail)
-summary(m6)
-# plot(m6)
-(c_hat <- deviance(m6) / df.residual(m6))
-
-library(MuMIn)
-dredge(m6, rank = "QAIC", chat = c_hat)
-dredge(m6, rank = "AIC")
-
-hacked.quasipoisson <- function(...) {
-    res <- quasipoisson(...)
-    res$aic <- poisson(...)$aic
-    res
-}
-
-QAIC(update(m6, family = hacked.quasipoisson, chat = c_hat)
-
-# library(gamlss)
-# m7 <- gamlss(tot_bay ~ h2005 + twi15m, data = oaks.data)
-# summary(m7)
-# plot(m7)
-
-# GGally::ggpairs(select(oaks.data, tot_lfct,tot_bay,h2005))
-
-m1 <- glmer(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), family = poisson, data = oaks.data, na.action = na.omit)
-summary(m1)
-
-m2 <- glmer.nb(tot_lfct ~ tot_bay + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit)
-
-summary(m2)
-anova(m2,lfctmod)
-plot(lfctmod)
-
-```
-
-```{r refit model with H bay density path, eval=F}
-# The refit model has path between H and UMCA density ####
-model_list1_refit <- list(
-      shannonmod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit),
-
-      densmod <- lmer(umcadens ~ h2005 + twi15m + (1|plotid) + (1|year),
-           data = oaks.data, na.action = na.omit),
-
-      lfctmod <- lmer(lfct ~ umcadens + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
-
-      wetmod <- lmer(hrs1422 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
-
-      hotmod <- lmer(hrs25 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
-
-      # hrs25 %~~% hrs1422,
-
-      infmod <- glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)
-)
-summary(lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit))
-summary(densmod)
-summary(lfctmod)
-summary(wetmod)
-summary(hotmod)
-summary(infmod)
-
-# summary(oaks.data)
-
-# ?psem I get errors with row matching when trying to use the new fucntion.
-I think it may have more to do with the new version of R and how that affects the internal sub-modeling when there are complex mixed effects models.
-
-```
-
-```{r Update model list}
-
-# Updated SEM uses untransformed UMCA variables ####
-# lfct is the ln(tot_lfct+1)
-      # lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
-m_list_update <- list(
-
-            shannon.mod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit),
-      # changed the bay response to be log-transformed 
-      umca.mod <- lmer(tot_bay.ln1 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
-
-      lfct.mod <- lmer(lfct ~ tot_bay.ln1 + hrs25 + hrs1422 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
-
-      wet.mod <- lmer(hrs1422 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
-
-      hot.mod <- lmer(hrs25 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
-
-      # hrs25 %~~% hrs1422,
-
-      inf.mod <- glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)
-
-)
-
-# ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
-#       geom_point() +
-#       geom_smooth(method = "lm", color = "red") +
-#       geom_smooth() +
-#       theme_classic()
-
-```
-
-```{r path-model-no-lide-plots, echo=F, eval=F}
-model_list1.noLIDE <- list(
-      shannons <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data.sub, sample_year == 2006)),
-
-      umca_density <- lmer(umca_density_ha.log ~ twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),
-
-      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),
+      slc <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
@@
                                    # umca_density_ha.log + twi15m.log +
-                                   twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),
+                                   twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
@@
                               # umca_density_ha.log + 
-                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data.sub),
+                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),
 
-      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data.sub)
+      oak_infection <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
 )
 
+model_list1_refit <- list(
+      shannons_refit <- lm(H.2005 ~ twi15m, data = filter(oakplots.model.data, sample_year == 2006)),
+
+      umca_density_refit <- lmer(umca_density_ha.log ~ H.2005 + twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+
+      slc_refit <- lmer(tot_lfct.log ~ umca_density_ha.log + hrs_abv25ds +
+                        #avg_hrs1422_wet_t_t1.log + H.2005 + 
+                        (1|plotid) + (1|sample_year), 
+                  data = oakplots.model.data),
+
+      wet_hours_1422_refit <- lmer(avg_hrs1422_wet_t_t1.log ~ 
+                                   # umca_density_ha.log + twi15m.log +
+                                   twi15m +
+                                   (1|plotid) + (1|sample_year), 
+                             data = oakplots.model.data),
+
+      hrs_abv25_refit <- lmer(hrs_abv25ds ~ 
+                              # umca_density_ha.log + 
+                              twi15m + (1|plotid) + (1|sample_year), data = oakplots.model.data),
+
+      oak_infection_refit <- glmer(oak.inf ~ tot_lfct.log + avg_hrs1422_wet_t_t1.log + H.2005 + (1|plotid) + (1|sample_year), family = binomial(link = "logit"), data = oakplots.model.data)
+)
 ```
@@
 
-# Examining the residual plots of each piece of the path model
+### Examining the residual plots of each piece of the path model
 
-## **Residual plots of Shannon's Diversity Index model**
-
-### Diagnostic plots 
-```{r shannon.mod diagnostic plots, echo=F}
-
-library(sjPlot)
-plot_model(shannon.mod, type = "diag", show.values = T, grid = TRUE)
-plot_model(shannon.mod, type = "resid", show.values = T)
-
+**Residual plots of Shannon's Diversity Index model**
+```{r shannons residuals plot, echo=F}
+plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
+plot(shannons$model$twi15m, resid(shannons),
+     xlab = "wetness index",
+     ylab = "model residuals",
+     main = "Diversity model residuals vs. wetness index predictor")
 ```
 
-### Fitted model plots
-```{r shannon.mod fit plots}
-
-plot_model(shannon.mod, type = "slope", show.values = T)
-plot_model(shannon.mod, type = "std", show.values = T, 
-           title = "Standardized coefficient estimate")
-S.plot <- plot_model(shannon.mod, type = "pred", terms = "twi15m")
-
-# line_color <- "black"
-# S.plot$data
-# str(S.plot)
-
-(shannon.plot <- plot_model(shannon.mod, type = "pred", terms = "twi15m") +
-      geom_point(data = oaks.data, 
-                 aes(twi15m, h2005), 
-                 alpha = .1, size = 1, position = "jitter") +
-      # geom_smooth(method = "loess", se = F, color = "gray50", linetype = "dashed") +
-      # geom_smooth(method = "lm", color = "black") +
-      theme_classic() +
-      xlab("Topographic index") +
-      ylab("Tree diversity (H')") +
-      theme(strip.background = element_blank(),
-            strip.text = element_blank()) +
-      NULL
-      )
-
-# plot(shannons, main = "OLS regression of diversity ~ wetness index residuals")
-# plot(shannons$model$twi15m, resid(shannons),
-#      xlab = "wetness index",
-#      ylab = "model residuals",
-#      main = "Diversity model residuals vs. wetness index predictor")
-
+**Residual plots of bay laurel density model**
+```{r bay laurel density residuals plot, echo=F}
+plot(umca_density, main = "LMM of bay laurel density residuals")
+plot(umca_density@frame$twi15m, resid(umca_density),
+     xlab = "wetness index",
+     ylab = "model residuals",
+     main = "Bay laurel density model residuals vs. wetness index predictor")
 ```
 
-## **Residual plots of bay laurel density model**
-
-### Diagnostic plots
-```{r umca.mod diagnostic plots, echo=F}
-
-plot_model(umca.mod, type = "diag")
-plot_model(umca.mod, type = "resid")
-
-```
-
-### Fitted model plots
-```{r umca.mod fit plots}
-
-plot_model(umca.mod, show.values = T, type = "est", 
-           title = "Coefficient estimates - UMCA density")
-plot_model(umca.mod, show.values = T, type = "std", 
-           title = "Standardized coefficient estimates - UMCA density")
-plot_model(umca.mod, show.values = T, type = "re")
-
-# umca_vs_h2005 <- plot_model(umca.mod, show.values = T, type = "pred", terms = c("twi15m"), pred.type = "fe")
-# umca_vs_h2005 +
-#       theme_classic()
-
-# oaks.data %>% 
-#       filter() %>% 
-#       ggplot(aes(h2005, log1p(tot_bay))) +
-#       geom_point() +
-#       geom_smooth(method = "lm") +
-      # facet_grid(.~year)
-# qplot(x=h2005, y=tot_bay, data = oaks.data)
-
-# library(ggeffects)
-# umca.mod_predict <- ggpredict(model = umca.mod.update, terms = "h2005", type = "re", full.data = T)
-# ggplot(umca.mod_predict, aes(x, predicted)) +
-#       # data = umca.mod.update@frame,
-#       geom_point(aes(x, predicted), show.legend = F) +
-#       # geom_smooth(se = F) +
-#       geom_smooth(method = "lm") +
-#       # facet_grid(.~year) +
-#       xlab("Shannon's diversity index (H')") +
-#       ylab("Bay laurel density (stems/ha)") +
-#       theme_classic() +
-#       NULL
-
-# oaks.data$h2005.sq <- oaks.data$h2005^2
-# oaks.data$tot_bay.log <- log1p(oaks.data$tot_bay)
-# umca.mod.update <- lmer(tot_bay.log ~ h2005 + twi15m + (1 | plotid) + (1 | year), data = oaks.data)
-# plot(umca.mod)
-# plot_model(umca.mod.update, show.values = T, type = "resid")
-# plot_model(umca.mod.update, show.values = T, type = "diag")
-
-# umca_vs_twi15m <- plot_model(densmod, show.values = T, type = "eff", terms = c("twi15m"), pred.type = "re")
-
-# plot(umca_density, main = "LMM of bay laurel density residuals")
-# plot(umca_density@frame$twi15m, resid(umca_density),
-#      xlab = "wetness index",
-#      ylab = "model residuals",
-#      main = "Bay laurel density model residuals vs. wetness index predictor")
-
-```
-
-## **Residual plots of symptomatic leaf count model**
-
-### Diagnostic plots
-```{r lfct.mod plot, echo=F}
-
-# diagnostic plots
-plot_model(lfct.mod, type = "diag")
-plot_model(lfct.mod, type = "resid")
-
-```
-
-### Fitted model plots
-```{r lfct.mod fit plots}
-# estimate plots
-plot_model(lfct.mod, show.values = T, type = "est", 
-           title = "Coefficient estimates - leaf count")
-plot_model(lfct.mod, show.values = T, type = "std", 
-           title = "Standardized coefficient estimates - leaf count")
-plot_model(lfct.mod, show.values = T, type = "re")
-
-# effect plots
-plot_model(lfct.mod, type = "slope", show.data = T, colors = "blambus") +
-      # geom_point(color = "black", alpha = .1) +
-      # geom_smooth(method = "lm", color = "black", fill = "black") +
-      # geom_smooth(method = "loess", color = NA) +
-      theme_bw() +
-      ylab("log (leaf count + 1)") +
-      # theme(strip.text = element_blank()) +
-      NULL
-
-slc_vs_bay_plot <- plot_model(lfct.mod, type = "pred", terms = c("tot_bay.ln1"), show.data = F) +
-      geom_point(data = oaks.data, aes(x = tot_bay.ln1, y = lfct),alpha = .1, size = 1) +
-      theme_classic() +
-      ylab("Symptomatic leaf count") +
-      xlab("Bay laurel density") +
-      ggtitle("Model fit") +
-      NULL
-
-slc_vs_hrs25_plot <- plot_model(lfct.mod, type = "pred", terms = c("hrs25"), show.data = F) +
-      geom_point(data = oaks.data, 
-                 aes(x = hrs25, y = lfct),
-                 alpha = .1, size = 1) +
-      theme_classic() +
-      ylab("Symptomatic leaf count") +
-      xlab("Hours >25 ºC") +
-      ggtitle("Model fit") +
-      NULL
-
-# ggeffects::ggpredict(lfct.mod)
-
-# slc_vs_bay_plot <- ggplot(oaks.data, aes(tot_bay.ln1, lfct)) +
-#       geom_point(alpha = .7) +
-#       geom_smooth(method = "lm", color = "black") +
-#       xlab("log( Bay laurel count )") +
-#       ylab("log( Symptomatic leaf count )") +
-#       theme_classic()
-# 
-# slc_vs_hrs25_plot <- ggplot(oaks.data, aes(hrs25, lfct)) +
-#       geom_point(alpha = .7) +
-#       geom_smooth(method = "lm", color = "black") +
-#       xlab("Number of hours >25 ºC") +
-#       ylab("log( Symptomatic leaf count )") +
-#       theme_classic()
-
-cowplot::plot_grid(slc_vs_bay_plot, 
-                   slc_vs_hrs25_plot,
-                   labels = "auto",
-                   nrow = 2)
-
-# plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
+**Residual plots of symptomatic leaf count model**
+```{r symptomatic leaf count residuals plot, echo=F}
+plot(resid(slc), main = "LMM of symptomatic leaf count residuals")
 # qqnorm(resid(slc), main="Q-Q plot for conditional residuals - SLC")
@@
 
-# plot(slc@frame$umca_density_ha.log, resid(slc),
-#      xlab = "bay laurel density",
-#      ylab = "model residuals",
-#      main = "SLC model residuals vs. bay laurel density predictor")
-# plot(slc@frame$hrs_abv25ds, resid(slc),
-#      xlab = "# dry-season hours >25 C",
-#      ylab = "model residuals",
-#      main = "SLC model residuals vs. dry-season hours >25 C predictor")
-# plot(slc@frame$avg_hrs1422_wet_t_t1.log,
-#      xlab = "avg. wet-hours 14-22C",
-#      ylab = "model residuals",
-#      main = "SLC model residuals vs. wet-hours predictor")
-# plot(slc@frame$H.2005, resid(slc),
-#      xlab = "Shannon's diversity index",
-#      ylab = "model residuals",
-#      main = "SLC model residuals vs. Shannon's diversity predictor")
+plot(slc@frame$umca_density_ha.log, resid(slc),
+     xlab = "bay laurel density",
+     ylab = "model residuals",
+     main = "SLC model residuals vs. bay laurel density predictor")
+plot(slc@frame$hrs_abv25ds, resid(slc),
+     xlab = "# dry-season hours >25 C",
+     ylab = "model residuals",
+     main = "SLC model residuals vs. dry-season hours >25 C predictor")
+plot(slc@frame$avg_hrs1422_wet_t_t1.log,
+     xlab = "avg. wet-hours 14-22C",
+     ylab = "model residuals",
+     main = "SLC model residuals vs. wet-hours predictor")
+plot(slc@frame$H.2005, resid(slc),
+     xlab = "Shannon's diversity index",
+     ylab = "model residuals",
+     main = "SLC model residuals vs. Shannon's diversity predictor")
 # plot(slc@frame$twi15m.log.scl, resid(slc),
@@
 #      main = "SLC model residuals vs. wetness index predictor")
-
-# ggplot(oaks.data, aes(tot_bay, tot_lfct)) +
-#       geom_point() +
-#       scale_x_log10() +
-#       scale_y_log10()
-
 ```
-
-```{r pairs plot of slc predictor variables, echo=F, eval=F}
-
+```{r pairs plot of slc predictor variables, echo=F}
 pairs(na.omit(select(oakplots.sub, umca_density_ha.log, hrs_abv25ds, avg_hrs1422_wet_t_t1.log, H.2005, twi15m)),
@@
       main = "Predictor variables for symptomatic leaf count models")
-
 ```
 
-## **Residual plots of average number of wet-hours model**
-
-### Diagnostic plots
-```{r wet.mod diganostic plots, echo=F}
-# diagnostic plots
-plot_model(wet.mod, type = "diag")
-plot_model(wet.mod, type = "resid")
-
-```
-
-### Fitted model plots
-```{r wet.mod fit plots}
-# estimate plots
-plot_model(wet.mod, show.values = T, type = "est", 
-           title = "Coefficient estimates - wet hours")
-plot_model(wet.mod, show.values = T, type = "std", 
-           title = "Standardized coefficient estimates - wet hours")
-plot_model(wet.mod, show.values = T, type = "re")
-
-# effect plots
-plot_model(wet.mod, type = "slope") +
-      theme_bw() +
-      # ylab("log (leaf count + 1)") +
-      theme(strip.background = element_blank()) +
-      NULL
-
-# plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
+**Residual plots of average number of wet-hours model**
+```{r wet-hours residuals plot, echo=F}
+plot(wet_hours_1422, main = "LMM of avg. # of hours 14-22 C current & prev. year residuals")
 # plot(wet_hours_1422@frame$umca_density_ha.log.scl, resid(wet_hours_1422),
@@
 #      main = "Wet-hours model residuals vs. bay laurel density predictor")
-# plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
-#      xlab = "bay laurel density",
-#      ylab = "model residuals",
-#      main = "Wet-hours model residuals vs. bay laurel density predictor")
-
+plot(wet_hours_1422@frame$twi15m, resid(wet_hours_1422),
+     xlab = "bay laurel density",
+     ylab = "model residuals",
+     main = "Wet-hours model residuals vs. bay laurel density predictor")
 ```
 
-## **Residual plots of dry-season temperature >25C model**
-
-### Diagnostic plots
-```{r hot.mod diagnostic plots, echo=F}
-
-# diagnostic plots
-plot_model(hot.mod, type = "diag")
-plot_model(hot.mod, type = "resid")
-
-```
-
-### Fitted model plots
-```{r hot.mod fit plots}
-# estimate plots
-plot_model(hot.mod, show.values = T, type = "est", 
-           title = "Coefficient estimates - hours >25 ºC")
-plot_model(hot.mod, show.values = T, type = "std", 
-           title = "Standardized coefficient estimates - hours >25 ºC")
-plot_model(hot.mod, show.values = T, type = "re")
-
-# effect plots
-plot_model(hot.mod, type = "slope", show.data = T) +
-      # geom_point() +
-      theme_classic() +
-      ylab("Number of hours >25 ºC") +
-      xlab("Topographic index") +
-      theme(strip.background = element_blank(),
-            strip.text = element_blank()) +
-      NULL
-
-# Fitted line with the observed data.
-hot.mod.plot <- plot_model(hot.mod, type = "pred", terms = "twi15m", show.data = F) +
-      geom_point(data = oaks.data, 
-                         aes(x = twi15m, y = hrs25),
-                         alpha = .1) +
-      theme_classic() +
-      ylab("Number of hours >25 ºC") +
-      xlab("Topographic index") +
-      theme(strip.background = element_blank(),
-            strip.text = element_blank()) +
-      NULL
-
-(hot.mod.plot2 <- ggplot(oaks.data, 
-                         aes(twi15m, hrs25,
-                             # color = as.factor(year)
-                             )
-                         ) +
-      geom_point(alpha = .1) +
-      # geom_smooth(method = "lm", se = F) +
-      geom_smooth(method = "lm", se = T, color = "black") +
-      scale_color_viridis_d() +
-      xlab("Topographic index") +
-      ylab("Number of hours >25 ºC") +
-      theme_classic() +
-      NULL
-      )
-
-# plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
-# plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
-#      xlab = "wetness index",
-#      ylab = "model residuals",
-#      main = "Dry-season hours model residuals vs. wetness index predictor")
+**Residual plots of dry-season temperature >25C model**
+```{r dry-season hours residuals plot, echo=F}
+plot(hrs_abv25, main = "LMM of # of hours >25 C during prev. dry season residuals")
+plot(hrs_abv25@frame$twi15m, resid(hrs_abv25),
+     xlab = "wetness index",
+     ylab = "model residuals",
+     main = "Dry-season hours model residuals vs. wetness index predictor")
 # plot(hrs_abv25@frame$umca_density_ha.log.scl, resid(hrs_abv25),
@@
 #      main = "Dry-season hours model residuals vs. wetness index predictor")
-
 ```
 
-## **Plots of oak infection model**
-
-### Diagnostic plots
-```{r inf.mod diagnostic plots, echo=F}
-
-plot_model(inf.mod, type = "diag")
-plot_model(inf.mod, type = "resid")
-
+**Residual plots of oak infection model**
+```{r oak infection residuals plot, echo=F}
+plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
+plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
+     xlab = "symptomatic leaf count",
+     ylab = "model residuals",
+     main = "Oak infection model residuals vs. symptomatic leaf count predictor")
+plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
+     xlab = "avg. wet-hours 14-22C",
+     ylab = "model residuals",
+     main = "Oak infection model residuals vs. wet-hours predictor")
+plot(oak_infection@frame$H.2005, resid(oak_infection),
+     xlab = "Shannon's diversity index",
+     ylab = "model residuals",
+     main = "Oak infection model residuals vs. diversity predictor")
 ```
-
-### Fitted model plots
-```{r inf.mod fit plots, echo=F}
-
-cowplot::plot_grid(
-      plot_model(inf.mod, type = "est", show.values = T, 
-                 title = "Oak infection coefficient estimates",
-                 vline.color = "gray80") +
-            # geom_hline(yintercept = 1, color = "gray80",linetype = "dashed") +
-            theme_classic(),
-      plot_model(inf.mod, type = "std", show.values = T, 
-                 title = "Oak infection std. coefficient estimates",
-                 vline.color = "gray80"
-                 ) +
-            theme_classic(),
-      nrow = 2
-      )
-
-inf.lfct.plot <- plot_model(inf.mod, type = "pred", terms = c("lfct"), show.data = T) +
-      # geom_point() +
-      theme_classic() +
-      xlab("Symptomatic leaf count") +
-      ylab("Infected oaks (%)") +
-      ggtitle("") +
-      NULL
-
-# marginal effects of each significant predictor variable
-cowplot::plot_grid(
-inf.lfct.plot <- inf.lfct.plot +
-      geom_point(data = oaks.data, 
-                 aes(x = lfct, 
-                     y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)
-                     ),
-                 alpha = .1, size = 1),
-# inf.lfct.plot +
-#       geom_point(data = filter(oaks.data, inf_oak_ct/(inf_oak_ct+uninf_oak_ct) <.5), 
-#                  aes(x = lfct, 
-#                      y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)
-#                      ),
-#                  alpha = .4, size = 1),
-
-inf.h2005.plot <- plot_model(inf.mod, type = "pred", terms = c("h2005"), show.data = T) +
-      geom_point(data = oaks.data, 
-                 aes(x = h2005, 
-                     y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)
-                     ),
-                 alpha = .1, size = 1) +
-      theme_classic() +
-      xlab("Shannon's diversity index") +
-      ylab("Infected oaks (%)") +
-      ggtitle("") +
-      NULL,
-
-inf.hrs1422.plot <- plot_model(inf.mod, type = "pred", terms = c("hrs1422")) +
-      geom_point(data = oaks.data, aes(x = hrs1422, y = inf_oak_ct/(inf_oak_ct+uninf_oak_ct)), alpha = .1, size = 1) +
-      theme_classic() +
-      xlab("Hours 14-22 ºC") +
-      ylab("Infected oaks (%)") +
-      ggtitle("") +
-      NULL,
-labels = "auto"
-)
-
-ggplot(oaks.data, aes((lfct), inf_oak_ct/(inf_oak_ct+uninf_oak_ct), color = as.factor(year))) +
-      geom_point(alpha = .5) +
-      # geom_line(aes(x=lfct, y=predict(inf.mod))) +
-      geom_smooth(aes(y=predict(inf.mod, type = "response")), method = "gam", formula = y~s(x, bs="cs"), se=F) +
-      scale_color_viridis_d() +
-      # stat_sum() +
-      # facet_grid(year~.) +
-      # geom_smooth(method = "loess", color = "black") +
-      theme_classic() +
-      NULL
-# 
-# 
-# inf.pred <- ggeffects::ggpredict(inf.mod, full.data = T)
-# 
-# ggplot(inf.pred$lfct, aes(x, predicted)) +
-#       # geom_point(data = oaks.data, aes(x = lfct, y = )) +
-#       geom_path()
-#       
-# inf.pred$h2005
-# 
-# inf.mod.predict <- predict(inf.mod)
-
-# plot(oak_infection, main = "Binomial GLMM of oak infection residuals")
-# plot(oak_infection@frame$tot_lfct.log, resid(oak_infection),
-#      xlab = "symptomatic leaf count",
-#      ylab = "model residuals",
-#      main = "Oak infection model residuals vs. symptomatic leaf count predictor")
-# plot(oak_infection@frame$avg_hrs1422_wet_t_t1.log, resid(oak_infection),
-#      xlab = "avg. wet-hours 14-22C",
-#      ylab = "model residuals",
-#      main = "Oak infection model residuals vs. wet-hours predictor")
-# plot(oak_infection@frame$H.2005, resid(oak_infection),
-#      xlab = "Shannon's diversity index",
-#      ylab = "model residuals",
-#      main = "Oak infection model residuals vs. diversity predictor")
-
-```
-
-```{r pairs plot of oak infection predictor variables, echo=F, eval=F}
-
+```{r pairs plot of oak infection predictor variables, echo=F}
 pairs(na.omit(select(oakplots.sub, tot_lfct.log, avg_hrs1422_wet_t_t1.log, H.2005)),
@@
       main = "Predictor variables for oak infection model")
-
 ```
 
-# Combine significant relationships from fitted models
-
-```{r significant relationships figure}
-
-cowplot::plot_grid(
-      inf.lfct.plot,
-      inf.h2005.plot +
-            xlab("Tree diversity (H')"),
-      inf.hrs1422.plot,
-      slc_vs_bay_plot +
-            ggtitle(""),
-      slc_vs_hrs25_plot +
-            ggtitle(""),
-      shannon.plot +
-            ggtitle(""),
-      hot.mod.plot +
-            ggtitle(""),
-      labels = "auto", label_x = .01
-      )
-
-```
-
-# Check for spatial autocorrelation
-Check oak infection and symptomatic leaf count response variables and model residuals.
-
-## **Oak infection autocorrelation**
-```{r spatial autocorrelation inf.mod}
-
-plots_coords <- select(plots_shp@data, PLOT_ID, X, Y)
-names(plots_coords) <- tolower(names(plots_coords))
-plots_coords <- plots_coords %>% 
-      rename(plotid = plot_id) %>% 
-      mutate(plotid = tolower(plotid))
-
-oaks.data <- left_join(
-      oaks.data, plots_coords
-)
-summary(oaks.data)
-
-library(spBayes); library(classInt); library(RColorBrewer); library(geoR)
-library(ape); library(ncf)
-
-oakplots.coords <- as.matrix(oaks.data[,c("x","y")])
-pl.dists <- as.matrix(dist(cbind(oaks.data$x, oaks.data$y)))
-pl.dists[1:20,1:20]
-dim(pl.dists)
-# image(pl.dists)
-pl.dists.inv <- 1/pl.dists
-# image(pl.dists.inv)
-diag(pl.dists.inv) <- 0
-pl.dists.inv[is.infinite(pl.dists.inv)] <- 0
-pl.dists.inv[1:20,1:20]
-# summary(pl.dists.inv)
-
-plot.idist <- iDist(cbind(plots_coords$x, plots_coords$y))
-plot.idist[plot.idist==0] <- NA
-min(plot.idist, na.rm = T)
-max.dist <- .33*max(iDist(cbind(plots_coords$x, plots_coords$y)))
-
-oaks.data <- oaks.data %>% 
-      mutate(p.oak.inf = inf_oak_ct/(inf_oak_ct+uninf_oak_ct))
-summary(oaks.data)
-# ggplot(oakplots.model.data %>% group_by(sample_year), 
-#        aes(sample_year, p.oak.inf)) +
-#       geom_point(position = "jitter")
-
-Moran.I(oaks.data$p.oak.inf, pl.dists.inv)
-cgram.inf <- correlog(oaks.data$x, oaks.data$y, oaks.data$p.oak.inf, increment = 300, resamp = 99)
-plot(cgram.inf, main = "Oak infection response correlogram")
-# abline(h = mean(cgram.inf$correlation))
-max(cgram.inf$correlation)
-
-vario.inf <- variog(coords = oakplots.coords,
-                            data = oaks.data$p.oak.inf)
-plot(vario.inf)
-fit.vario.inf <- variofit(vario.inf)
-lines(fit.vario.inf)
-abline(h = fit.vario.inf$nugget, col = "purple")
-abline(h = fit.vario.inf$cov.pars[1], col = "darkblue")
-abline(v = -log(.05)*fit.vario.inf$cov.pars[2], col = "tan4")
-
-inf.mod.resids <- resid(inf.mod)
-Moran.I(inf.mod.resids, pl.dists.inv)
-cgram.inf.resids <- correlog(oaks.data$x, oaks.data$y, inf.mod.resids, increment = 300, resamp = 99)
-plot(cgram.inf.resids, main = "Oak infection residuals correlogram")
-
-vario.inf.resids <- variog(coords = oakplots.coords,
-                            data = inf.mod.resids)
-plot(vario.inf.resids)
-fit.vario.inf.resids <- variofit(vario.inf.resids)
-lines(fit.vario.inf.resids)
-abline(h = fit.vario.inf.resids$nugget, col = "purple")
-abline(h = fit.vario.inf.resids$cov.pars[1], col = "darkblue")
-abline( v = -log(.05)*fit.vario.inf.resids$cov.pars[2], col = "tan4")
-```
-
-There was minimal spatial autocorrelation in the response variable. The spatial autocorrelation that was present was dramitically reduced in the model residuals.
-
-## **Symptomatic leaf count**
-```{r spatial autocorrelation lfct.mod}
-
-# lfct.data <- oakplots.model.data %>% 
-#       select(sample_year, tot_lfct.log) %>% 
-#       arrange(sample_year)
-# acf(lfct.data$tot_lfct.log)
-
-# qplot(x = sample_year, y = tot_lfct, data = oakplots.model.data)
-
-Moran.I(oaks.data$lfct, pl.dists.inv)
-cgram.lfct <- correlog(oaks.data$x, oaks.data$y, oaks.data$lfct, increment = 300, resamp = 99)
-plot(cgram.lfct, main = "Leaf count response correlogram")
-
-r.lfctmod <- resid(lfct.mod)
-Moran.I(r.lfctmod, pl.dists.inv)
-cgram.lfct.resids <- correlog(oaks.data$x, oaks.data$y, r.lfctmod, increment = 300, resamp = 99)
-plot(cgram.lfct.resids, main = "Leaf count residuals correlogram")
-
-```
-
-There was minimal spatial autocorrelation in the response variable. The spatial autocorrelation that was present was dramitically reduced in the model residuals.
-
-# Fitting the path model
+### Fitting the path model
 
@@
 ```{r fit and compare all oak plots, echo=F, message=F, warning=F}
-# raw_variables_model <- sem.fit(model_list1, oakplots.model.data, 
-#         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-#         conditional = F, .progressBar = F)
-# 
-# raw_vars_refit <- sem.fit(
-#       model_list1_refit, oaks.data, corr.errors = c("hrs25 ~~ hrs1422"),
-#       conditional = F, .progressBar = F)
+raw_variables_model <- sem.fit(model_list1, oakplots.model.data, 
+        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+        conditional = T, .progressBar = F)
 
-raw_vars_update <- sem.fit(
-      m_list_update, oaks.data,
-      corr.errors = c("hrs25 ~~ hrs1422"),
-      conditional = F, .progressBar = F
-)
-
-## Fit SEM with all LIDE plots excluded ####
-# sem.fit(model_list1.noLIDE, oakplots.model.data.sub, 
-#         corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-#         conditional = T, .progressBar = F)
+raw_vars_refit <- sem.fit(model_list1_refit, oakplots.model.data, 
+        corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+        conditional = T, .progressBar = F)
 
@@
 #           .progressBar = F)
-
 ```
@@
 ```{r tables of d-sep test results from sem.fit, echo=F}
-
-# kable(raw_variables_model, digits = 4,
-      # caption = "Results of the d-sep test using raw variables.")
-
-# kable(raw_vars_refit, digits = 4,
-      # caption = "Results of the d-sep test using raw variables.")
-
-kable(raw_vars_update, digits = 3,
+kable(raw_variables_model, digits = 4,
       caption = "Results of the d-sep test using raw variables.")
-
 # kable(scaled_variables_model,
@@
 #       caption = "Results of the d-sep test using standardized variables - no UMCA density.")
-
 ```
 
-## Path coefficients
+### Path coefficients
 The path model provides direct and indirect effects of predictor variables on multiple responses, which can be combined to calculate the total effect of a predictor variable on a response.
@@
 ```{r get path coefficients, echo=F}
+raw_coefficients <- sem.coefs(model_list1, oakplots.model.data,
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+scaled_coefficients <- sem.coefs(model_list1, oakplots.model.data,
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+          standardize = "scale")
 
-# raw_coefficients <- sem.coefs(model_list1, oakplots.model.data,
-#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
-# scaled_coefficients <- sem.coefs(model_list1, oakplots.model.data,
-#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-#           standardize = "scale")
-
-# raw_coeff_refit <- sem.coefs(
-#       model_list1_refit, oaks.data
-#       )
-# scaled_coeff_refit <- sem.coefs(
-#       model_list1_refit, oaks.data,
-#       standardize = "scale"
-#       )
-
-raw_coeff_update <- sem.coefs(
-      m_list_update, oaks.data
-      )
-scaled_coeff_update <- sem.coefs(
-      m_list_update, oaks.data,
-      standardize = "scale"
-      )
-
-## Model with LIDE plots excluded
-# sem.coefs(model_list1.noLIDE, oakplots.model.data.sub,
-#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
-# sem.coefs(model_list1.noLIDE, oakplots.model.data.sub,
-#           corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
-#           standardize = "scale")
+raw_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"))
+scaled_coeff_refit <- sem.coefs(model_list1_refit, oakplots.model.data,
+          corr.errors = c("hrs_abv25ds ~~ avg_hrs1422_wet_t_t1.log"),
+          standardize = "scale")
 
@@
 #           corr.errors = c("hrs_abv25ds.scl ~~ avg_hrs1422_wet_t_t1.log.scl"))
-
 ```
@@
 ```{r tables of path coefficients, echo=F}
-
-kable(raw_coeff_update, digits = 3,
+kable(raw_coefficients, digits = 4,
       caption = "Raw path coefficients")
-# kable(raw_coeff_refit, digits = 3,
-#       caption = "Raw path coefficients")
 # kable(raw_coefficients_noUMCA,
@@
 
-kable(scaled_coeff_update, digits = 3,
+kable(scaled_coefficients, digits = 4,
       caption = "Standardized path coefficients")
-# kable(scaled_coeff_refit, digits = 2,
-#       caption = "Standardized path coefficients")
 # kable(scaled_coefficients_noUMCA,
@@
 ```{r disease figures, echo=F, warning=F}
-
+library(ggplot2)
 temperatures <- left_join(
@@
       ylab("Number of hours") +
-      ggtitle("Heat Exposure (dry-season Hours >25 C)") +
-      theme(panel.grid = element_blank())
+      ggtitle("Heat Exposure (dry-season Hours >25 C)")
 )
@@
       ylab("Number of hours") +
-      ggtitle("Warm & Wet Conditions (wet-hours 14-22 C)") +
-      theme(panel.grid = element_blank())
+      ggtitle("Warm & Wet Conditions (wet-hours 14-22 C)")
 )
@@
 library(ggplot2)
-# rainfall
 # summary(rainy_season_ppt)
@@
       geom_boxplot(aes(group = year), notch = T) +
-      geom_hline(aes(yintercept = mean(rain2d_total)), linetype = "dashed") +
+      geom_hline(aes(yintercept = mean(rain2d_total)),
+                 colour = "red") +
       theme_bw() +
@@
       ylab("Total Rainfall (mm)") +
-      # ggtitle("Rainfall Estimated at Plots from Rain Gauges (November - May)") +
-      theme(panel.grid = element_blank(),
-            axis.title = element_text(size = 12),
-            axis.text = element_text(size = 10))
-# ggsave("figures/rainy-season-ppt.eps", width = 6.25, height = 4.5)
+      ggtitle("Rainfall Estimated at Plots from Rain Gauges (November - May)")
 
@@
 ## Model a subset of the data
-
-### Bay laurel density ~ Diveristy relationship
-
-```{r}
-ggplot(oakplots.model.data,
-       aes(umca_density_ha, H.2005)) +
-      geom_point() +
-      geom_smooth(method = "loess") +
-      facet_grid(sample_year~.) +
-      theme_bw()
-
-d1 <- select(oakplots.sub, plotid:inf_oak_ct, H.2005)
-d1 <- mutate(d1, tot_oak = rowSums(cbind(inf_oak_ct, uninf_oak_ct)))
-ggplot(filter(d1, sample_year == 2006), aes(log1p(tot_oak), H.2005)) +
-      geom_point() +
-      geom_smooth(method = "lm") +
-      theme_bw()
-
-```
-
 I want to examine the different matrices of the mixed effects models using a smaller datset that will make them easier to visualize.

### 2019-04-07 analysis/plotting.R
Line deletion.

@@ -125,7 +125,6 @@ p2 <- p2 + geom_point(aes(color = plot.id), position = "jitter") +
 p2 + geom_text(aes(label = tag, color = ))
 
 
-
 p1 + geom_boxplot(na.rm=TRUE) +
   stat_summary(fun.y=mean, na.rm=TRUE, geom="line", aes(group=1))  + 
   stat_summary(fun.y=mean, na.rm=TRUE,geom="point",shape=8) +

### 2019-04-07 pathogen-spillover-diss-ch.Rmd
Revert to pre-conflict file.

--- 2018-11-12 pathogen-spillover-diss-ch.Rmd
@@
       # lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+test.mod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit)
+
 m_list_update <- list(
 
-            shannon.mod <- lm(h2005 ~ twi15m, data = filter(oaks.data, year==2006), na.action = na.omit),
+      shannon.mod <- lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
       # changed the bay response to be log-transformed 
@@
       inf.mod <- glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)
+
+)
+
+test.psem <- psem(
+
+      lmer(h2005 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+      # lmer(tot_bay.ln1 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+      # 
+      # lmer(lfct ~ tot_bay.ln1 + hrs25 + hrs1422 + h2005 + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+      # 
+      # lmer(hrs1422 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+      # 
+      # lmer(hrs25 ~ twi15m + (1|plotid) + (1|year), data = oaks.data, na.action = na.omit),
+
+      # hrs25 %~~% hrs1422,
+
+      glmer(oakinf ~ lfct + hrs1422 + h2005 + (1|plotid) + (1|year), family = binomial(link = "logit"), data = oaks.data, na.action = na.omit)
 
@@
 # marginal effects of each significant predictor variable
-cowplot::plot_grid(
+
 inf.lfct.plot <- inf.lfct.plot +
@@
                      ),
-                 alpha = .1, size = 1),
+                 alpha = .1, size = 1)
 # inf.lfct.plot +
@@
       ggtitle("") +
-      NULL,
+      NULL
 
@@
       NULL,
-labels = "auto"
-)
 
@@
 )
-ggsave("figures/path-significant-relationships.png", plot = sig.rel.fig,
-       width = 7, height = 7, units = "in")
+ggsave("figures/Fig4-path_significant_relationships.png", plot = sig.rel.fig,
+       width = 7, height = 7, units = "in", dpi=600)
 
